### 2016-02-03 PaperThoughts.Rnw
add some thoughts on paper

\documentclass[a4paper]{report}
\usepackage[margin=.75in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,array}
\usepackage{booktabs}

%% load any required packages here

\usepackage{adjustbox}            %% to align tops of minipages
\usepackage[font=small,skip=5pt]{caption}
\usepackage{subcaption}
\usepackage{afterpage}

\usepackage{color}
\definecolor{purple}{rgb}{.4,0,.8}
\newcommand{\hh}[1]{{\color{magenta} #1}}
\newcommand{\st}[1]{{\color{purple} #1}}

\begin{document}

\LARGE{\bf Thoughts about Network Visual Inference Paper}

\begin{enumerate}
\item Need to complete lineups in 2-3 weeks
\item Lineups: 
  \begin{enumerate}
  \item What sizes? 3, 6, 9, 12
  \item What data? Football, olga's data, teenage friends data?
  \item Other things to consider: coloring of nodes? 
  \end{enumerate}
\item Get simulated networds from models from RSiena
\item Figure out another way to fit a model and then simulate from it? 
\end{enumerate}

Possible section outline for paper:
\begin{enumerate}
\item Introduction - visual inference, networks
\item Data and Model fitting & simulation
\item Lineup methods
\item Results
\item Conclusions and Discussion
\end{enumerate}

\end{document}

### 2016-02-03 paper/network-inference.Rnw
generic Rnw for vcgt

\documentclass[journal]{vgtc}                % final (journal style)
%\documentclass[review,journal]{vgtc}         % review (journal style)
%\documentclass[widereview]{vgtc}             % wide-spaced review
%\documentclass[preprint,journal]{vgtc}       % preprint (journal style)
%\documentclass[electronic,journal]{vgtc}     % electronic version, journal

%% Uncomment one of the lines above depending on where your paper is
%% in the conference process. ``review'' and ``widereview'' are for review
%% submission, ``preprint'' is for pre-publication, and the final version
%% doesn't use a specific qualifier. Further, ``electronic'' includes
%% hyperreferences for more convenient online viewing.

%% Please use one of the ``review'' options in combination with the
%% assigned online id (see below) ONLY if your paper uses a double blind
%% review process. Some conferences, like IEEE Vis and InfoVis, have NOT
%% in the past.

%% Please note that the use of figures other than the optional teaser is not permitted on the first page
%% of the journal version.  Figures should begin on the second page and be
%% in CMYK or Grey scale format, otherwise, colour shifting may occur
%% during the printing process.  Papers submitted with figures other than the optional teaser on the
%% first page will be refused.

%% These three lines bring in essential packages: ``mathptmx'' for Type 1
%% typefaces, ``graphicx'' for inclusion of EPS figures. and ``times''
%% for proper handling of the times font family.

\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{times}

%% We encourage the use of mathptmx for consistent usage of times font
%% throughout the proceedings. However, if you encounter conflicts
%% with other math-related packages, you may want to disable it.

%% This turns references into clickable hyperlinks.
\usepackage[bookmarks,backref=true,linkcolor=black]{hyperref} %,colorlinks
\hypersetup{
  pdfauthor = {},
  pdftitle = {},
  pdfsubject = {},
  pdfkeywords = {},
  colorlinks=true,
  linkcolor= black,
  citecolor= black,
  pageanchor=true,
  urlcolor = black,
  plainpages = false,
  linktocpage
}

%% If you are submitting a paper to a conference for review with a double
%% blind reviewing process, please replace the value ``0'' below with your
%% OnlineID. Otherwise, you may safely leave it at ``0''.
\onlineid{0}

%% declare the category of your paper, only shown in review mode
\vgtccategory{Research}

%% allow for this line if you want the electronic option to work properly
\vgtcinsertpkg

%% In preprint mode you may define your own headline.
%\preprinttext{To appear in an IEEE VGTC sponsored conference.}

%% Paper title.

\title{Variations of boxplots}

%% This is how authors are specified in the journal style

%% indicate IEEE Member or Student Member in form indicated below
\author{Susan VanderPlas, Eric Hare, Heike Hofmann \textit{Member, IEEE}, and Marie Vendettuoli}
\authorfooter{
%% insert punctuation at end of each item
\item
 Roy G. Biv is with Starbucks Research. E-mail: roy.g.biv@aol.com.
\item
 Ed Grimley is with Grimley Widgets, Inc.. E-mail: ed.grimley@aol.com.
\item
 Martha Stewart is with Martha Stewart Enterprises at Microsoft
 Research. E-mail: martha.stewart@marthastewart.com.
}

%other entries to be set up for journal
\shortauthortitle{VanderPlas \MakeLowercase{\textit{et al.}}: Boxplot variations}
%\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}

%% Abstract section.
\abstract{Boxplots have been around for quite a while now. There are a ton of variations but are those improvements? We find out using the visual inference protocol.} % end of abstract

%% Keywords that describe your work. Will show as 'Index Terms' in journal
%% please capitalize first letter and insert punctuation after last keyword
\keywords{Human subject testing, visual inference, comparison of designs}

%% ACM Computing Classification System (CCS). 
%% See <http://www.acm.org/class/1998/> for details.
%% The ``\CCScat'' command takes four arguments.

% HH: do we need those classifications?
% \CCScatlist{ % not used in journal version
%  \CCScat{K.6.1}{Management of Computing and Information Systems}%
% {Project and People Management}{Life Cycle};
%  \CCScat{K.7.m}{The Computing Profession}{Miscellaneous}{Ethics}
% }

%% Uncomment below to include a teaser figure.
  \teaser{
 \centering
 \includegraphics[width=16cm]{CypressView}
  \caption{In the Clouds: Vancouver from Cypress Mountain.}
  }

%% Uncomment below to disable the manuscript note
%\renewcommand{\manuscriptnotetxt}{}

%% Copyright space is enabled by default as required by guidelines.
%% It is disabled by the 'review' option or via the following command:
% \nocopyrightspace

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% START OF THE PAPER %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% The ``\maketitle'' command must be the first command after the
%% ``\begin{document}'' command. It prepares and prints the title block.

%% the only exception to this rule is the \firstsection command
\firstsection{Introduction}

\maketitle

%% \section{Introduction} %for journal use above \firstsection{..} instead

\section{Exposition}

Equations are cool.

\begin{equation}
\sum_{j=1}^{z} j = \frac{z(z+1)}{2}
\end{equation}

Notice the caption at the top of Table~\ref{vis_accept}.

\begin{table}
%% Table captions on top in journal version
 \caption{SciVis Paper Acceptance Rate: 1994-2006}
 \label{vis_accept}
 \scriptsize
 \begin{center}
   \begin{tabular}{cccc}
     Year & Submitted & Accepted & Accepted (\%)\\
   \hline
     2004 & 167 & 46 & 27.6\\
     2005 & 268 & 88 & 32.8\\
     2006 & 228 & 63 & 27.6
   \end{tabular}
 \end{center}
\end{table}

\subsection{Mezcal Head}

Citing happens only with cite - like this~\cite{Lorensen:1987:MCA}. Footnotes are ok%
\footnote{Footnotes appear at the bottom of the column}.

\section{Conclusion}

%% if specified like this the section will be committed in review mode
\acknowledgments{
The authors wish to thank A, B, C. This work was supported in part by
a grant from XYZ.}

\bibliographystyle{abbrv}
%%use following if all content of bibtex file should be shown
%\nocite{*}
\bibliography{template}
\end{document}

### 2016-02-03 paper/network-inference.Rnw
basics in

@@
 
-\title{Variations of boxplots}
+\title{Visual inference for networks}
 
@@
 %% indicate IEEE Member or Student Member in form indicated below
-\author{Susan VanderPlas, Eric Hare, Heike Hofmann \textit{Member, IEEE}, and Marie Vendettuoli}
+\author{Samantha Tyner, Heike Hofmann \textit{Member, IEEE}}
 \authorfooter{
@@
 %other entries to be set up for journal
-\shortauthortitle{VanderPlas \MakeLowercase{\textit{et al.}}: Boxplot variations}
+\shortauthortitle{Tyner \MakeLowercase{\textit{et al.}}: Visual inference for networks}
 %\shortauthortitle{Firstauthor \MakeLowercase{\textit{et al.}}: Paper Title}
@@
 %% Abstract section.
-\abstract{Boxplots have been around for quite a while now. There are a ton of variations but are those improvements? We find out using the visual inference protocol.} % end of abstract
+\abstract{abstract is here} % end of abstract
 
@@
 %% please capitalize first letter and insert punctuation after last keyword
-\keywords{Human subject testing, visual inference, comparison of designs}
+\keywords{Human subject testing, visual inference, Siena models, network model diagnostics}
 
@@
  \centering
- \includegraphics[width=16cm]{CypressView}
+ Teaser picture goes here
   \caption{In the Clouds: Vancouver from Cypress Mountain.}
@@
 
-Citing happens only with cite - like this~\cite{Lorensen:1987:MCA}. Footnotes are ok%
+Citing happens only with cite - like this~\cite{buja:2009}. Footnotes are ok%
 \footnote{Footnotes appear at the bottom of the column}.
@@
 %\nocite{*}
-\bibliography{template}
+\bibliography{references}
 \end{document}

### 2016-02-04 paper/network-inference.Rnw
reference added
copied the list into the paper

@@
 
+\begin{enumerate}
+\item Need to complete lineups in 2-3 weeks
+\item Lineups: 
+  \begin{enumerate}
+  \item What sizes? 3, 6, 9, 12
+  \item What data? Football, olga's data, teenage friends data?
+  \item Other things to consider: coloring of nodes? 
+  \end{enumerate}
+\item Get simulated networds from models from RSiena
+\item Figure out another way to fit a model and then simulate from it? 
+\end{enumerate}
+
+Possible section outline for paper:
+\begin{enumerate}
+\item Introduction - visual inference, networks
+\item Data and Model fitting \& simulation
+\item Lineup methods
+\item Results
+\item Conclusions and Discussion
+\end{enumerate}
+
 \section{Exposition}
@@
 
-Citing happens only with cite - like this~\cite{buja:2009}. Footnotes are ok%
+Citing happens only with cite - like this for the lineup protocol~\cite{buja:2009}, and for RSiena~\cite{RSiena}. Footnotes are ok%
 \footnote{Footnotes appear at the bottom of the column}.

### 2016-02-09 Code/RSiena_mySimFn.R
first go at simulation in RSiena

@@ -1,0 +1,66 @@
+# simulate the networks 
+
+# M = number of waves to simulate
+# init.waves = initial array of network observations
+# V0 = categorical actor variable
+# 
+
+SimulateNextWave <- function(init.waves, V0, 
+                                     rate.params.names, rate.params.vals,
+                                     eval.params.names, eval.params.vals,
+                                     eval.int.params.names, eval.int.params.vals){
+  if (ncol(init.waves[,,1]) != nrow(init.waves[,,1])){
+    return("Error: Non-square adjacency matrix provided")
+  } 
+  if (length(eval.params.vals) != length(eval.params.names)){
+    return("Error: Number of eval parameters to include in model does not
+           match number of eval parameter values provided.")
+  }
+  if (length(rate.params.vals) != length(rate.params.names)){
+    return("Error: Number of rate parameters to include in model does not
+           match number of rate parameter values provided.")
+  }
+  if (length(eval.int.params.vals) != length(eval.int.params.names)){
+    return("Error: Number of interaction eval parameters to include in model does not
+           match number of interaction eval parameter values provided.")
+  }
+  init.waves <- friendshipData
+  n <- ncol(init.waves[,,1])
+  V0 <- drink
+  c <- length(unique(V0[,1]))
+  p.e <- length(eval.params.names)
+  p.r <- length(rate.params.names)
+  p.e.i <- length(eval.int.params.names)
+  # Simulates M consecutive network and behavior waves, with n actors,
+  # with c categories of the behavior variable, from initial network init.net.adj
+  # according to a stochastic actor-oriented model with actor covariate V0
+  # with parameter values params.values for the parameters of interest params.names
+  Va <- varCovar(V0)
+  X   <- sienaDependent(init.waves)
+  InitData <- sienaDataCreate(X, Va)
+  InitEff0 <- getEffects(InitData)
+  # sink to avoid printing to the screen
+  #sink("eff.txt")
+  # Specify the parameters.
+  # The rate parameters are first multiplied by 10,
+  # which will be used only to get from the totally random network XX[,,1] = X0
+  # to the network that will be the simulated first wave.
+  
+  for (i in 1:p.r){
+    InitEff0 <- setEffect(InitEff0, rate.params.names[i], type = 'rate', initialValue = rate.params.vals[i], period = i, character = T)
[…]

### 2016-02-09 Code/friendship.R

@@ -1,0 +1,50 @@
+# Using the friends example in RSiena 
+
+# Use the three effects presented as the alternative model, then used 
+# the similarity effects presented in the "assignment" section of the code
+# as the null model
+
+setwd("s50_data")
+library(RSiena) # or RSienaTest
+# read in network data
+friend.data.w1 <- as.matrix(read.table("s50-network1.dat"))
+friend.data.w2 <- as.matrix(read.table("s50-network2.dat"))
+friend.data.w3 <- as.matrix(read.table("s50-network3.dat"))
+# read in covariate data
+drink <- as.matrix(read.table("s50-alcohol.dat"))
+smoke <- as.matrix(read.table("s50-smoke.dat"))
+# First create a 50 * 50 * 3 array composed of the 3 adjacency matrices
+friendshipData <- array( c( friend.data.w1, friend.data.w2, friend.data.w3 ),
+                         dim = c( 50, 50, 3 ) )
+# and next give this the role of the dependent variable:
+friendship <- sienaDependent(friendshipData)
+alcohol <- varCovar( as.matrix(drink[,1:2]) )
+# create siena data object
+mydata <- sienaDataCreate( friendship, alcohol)
+
+myeff <- getEffects( mydata )
+myeff1 <- includeEffects( myeff, egoX, altX, egoXaltX, interaction1 = "alcohol" )
+myeff2 <- includeEffects( myeff, simX, interaction1 = "alcohol")
+# Create object with algorithm settings
+myalgorithm <- sienaAlgorithmCreate( projname = 's50' )
+#Estimate parameters
+ans1 <- siena07( myalgorithm, data = mydata, effects = myeff1)
+ans2 <- siena07( myalgorithm, data = mydata, effects = myeff2)
+# view estimates
+ans1$rate
+ans1$theta
+
+rate.params <- data.frame(rate.param = c("Rate", "Rate"),
+           rate.param.val = ans1$rate)
+eval.params <- data.frame(eval.param = c("density", "recip"), eval.param.val = ans1$theta[1:2])
+eval.int.params <- data.frame(eval.int.param = c("altX", "egoX", "egoXaltX"), 
+                              eval.int.param.val = ans1$theta[3:5])
+
+sim1 <- SimulateNextWave(init.waves = friendshipData, V0 = drink, 
+                 rate.params.names = rate.params$rate.param, 
+                 rate.params.vals = rate.params$rate.param.val,
+                 eval.params.names = eval.params$eval.param,
[…]

### 2016-02-09 paper/network-inference.Rnw

@@
 \item
- Roy G. Biv is with Starbucks Research. E-mail: roy.g.biv@aol.com.
+ Samantha Tyner is with the Department of Statistics, Iowa State University. E-mail: sctyner@iastate.edu.
 \item
- Ed Grimley is with Grimley Widgets, Inc.. E-mail: ed.grimley@aol.com.
-\item
- Martha Stewart is with Martha Stewart Enterprises at Microsoft
- Research. E-mail: martha.stewart@marthastewart.com.
+ Heike Hofmann is with the Department of Statistics, Iowa State University. E-mail: hofmann@iastate.edu.
 }
@@
   \item Other things to consider: coloring of nodes? 
+  \item use replicates: different seeds for the same data produces different layouts -
+  different draws from the model
   \end{enumerate}

### 2016-02-09 Code/RSiena_mySimFn.R
created simulated network from fitted friend params

@@ -62,5 +62,5 @@ SimulateNextWave <- function(init.waves, V0,
   # Simulate the first wave.
   InitSim   <- siena07(InitAlg, data=InitData, eff=InitEff0,
                        returnDeps=TRUE, batch=TRUE, silent=TRUE)
-  return(InitSim) 
-}  
+  return(data.frame(InitSim$sims[[1000]][[1]][[1]][[1]]))
+}   

### 2016-02-09 Code/friendship.R

@@ -47,4 +47,8 @@ sim1 <- SimulateNextWave(init.waves = friendshipData, V0 = drink,
                  eval.params.val = eval.params$eval.param.val,
                  eval.int.params.names = eval.int.params$eval.int.param,
                  eval.int.params.vals = eval.int.params$eval.int.param.val)
+sim1_net <- merge(sim1, data.frame(id =1:50), by.x = "X1", by.y = "id", all = T)
+
+ggplot(sim1_net, aes(from_id = X1, to_id = X2)) + 
+  geom_net(label = T)
   

### 2016-02-09 Code/friendship.R
create 1 small lineup

@@ -50,5 +50,57 @@ sim1 <- SimulateNextWave(init.waves = friendshipData, V0 = drink,
 sim1_net <- merge(sim1, data.frame(id =1:50), by.x = "X1", by.y = "id", all = T)
 
 ggplot(sim1_net, aes(from_id = X1, to_id = X2)) + 
-  geom_net(label = T)
-  
+  geom_net(label = T, size = .1)
+
+sim2 <- SimulateNextWave(init.waves = friendshipData, V0 = drink, 
+                         rate.params.names = rate.params$rate.param, 
+                         rate.params.vals = rate.params$rate.param.val,
+                         eval.params.names = eval.params$eval.param,
+                         eval.params.val = eval.params$eval.param.val,
+                         eval.int.params.names = eval.int.params$eval.int.param,
+                         eval.int.params.vals = eval.int.params$eval.int.param.val)
+sim2_net <- merge(sim2, data.frame(id =1:50), by.x = "X1", by.y = "id", all = T)
+ggplot(sim2_net, aes(from_id = X1, to_id = X2)) + 
+  geom_net(label = T, size = .1)
+
+#### Using the different model (ans2) #######
+rate.params2 <- data.frame(rate.param = c("Rate", "Rate"),
+                          rate.param.val = ans2$rate)
+eval.params2 <- data.frame(eval.param = c("density", "recip"), eval.param.val = ans2$theta[1:2])
+eval.int.params2 <- data.frame(eval.int.param = "simX", 
+                              eval.int.param.val = ans2$theta[3])
+
+sim1.2 <- SimulateNextWave(init.waves = friendshipData, V0 = drink, 
+                         rate.params.names = rate.params2$rate.param, 
+                         rate.params.vals = rate.params2$rate.param.val,
+                         eval.params.names = eval.params2$eval.param,
+                         eval.params.val = eval.params2$eval.param.val,
+                         eval.int.params.names = eval.int.params2$eval.int.param,
+                         eval.int.params.vals = eval.int.params2$eval.int.param.val)
+sim1.2_net <- merge(sim1.2, data.frame(id =1:50), by.x = "X1", by.y = "id", all = T)
+
+ggplot(sim1.2_net, aes(from_id = X1, to_id = X2)) + 
+  geom_net(label = T, size = .1)
+
+sim2.2 <- SimulateNextWave(init.waves = friendshipData, V0 = drink, 
[…]

### 2016-02-10 paper/network-inference.Rnw
rearrange folders & put code in paper

@@
 \usepackage{times}
+\usepackage{color}
+\definecolor{purple}{rgb}{.4,0,.8}
+\newcommand{\hh}[1]{{\color{magenta} #1}}
+\newcommand{\st}[1]{{\color{purple} #1}}
 
@@
 
+<<packages, echo=FALSE, warning=FALSE, message=FALSE>>=
+library(RSiena)
+library(network)
+library(ggplot2)
+library(geomnet)
+@
 %% \section{Introduction} %for journal use above \firstsection{..} instead
@@
 
-\section{Exposition}
+\section{Examples \st{??}}
 
-Equations are cool.
+\st{
+One of the data sets provided with RSiena is a subset of 50 female respondants from the \emph{Teenage Friends and Lifestyle Study} XXX citation???
+XXX.
+There are three friendship networks provided at three timepoints representing the relationships between the 50 teenage girls in the data.
+There are also actor-level variables observed at each time period, such as a girl's drinking and smoking behavior.
+These networks are shown in \autoref{fig:friendship_waves}
+XX autoref not working?? XX
+<<friendship_waves, echo=FALSE, message=FALSE, warning=FALSE>>=
+friend.data.w1 <- as.matrix(read.table("../Data/s50_data/s50-network1.dat"))
+friend.data.w2 <- as.matrix(read.table("../Data/s50_data/s50-network2.dat"))
+friend.data.w3 <- as.matrix(read.table("../Data/s50_data/s50-network3.dat"))
+drink <- as.matrix(read.table("../Data/s50_data/s50-alcohol.dat"))
 
-\begin{equation}
-\sum_{j=1}^{z} j = \frac{z(z+1)}{2}
-\end{equation}
+friend.data.edges <- data.frame(as.edgelist(as.network(friend.data.w1)))
+friend.data.nodes <- data.frame(ids = 1:50, drink = drink[,1])
+friend.data.net1 <- merge(friend.data.edges, friend.data.nodes, by.x="X1", by.y = "ids", all = T)
+friend.data.edges <- data.frame(as.edgelist(as.network(friend.data.w2)))
+friend.data.nodes <- data.frame(ids = 1:50, drink = drink[,2])
+friend.data.net2 <- merge(friend.data.edges, friend.data.nodes, by.x="X1", by.y = "ids", all = T)
+friend.data.edges <- data.frame(as.edgelist(as.network(friend.data.w3)))
+friend.data.nodes <- data.frame(ids = 1:50, drink = drink[,3])
+friend.data.net3 <- merge(friend.data.edges, friend.data.nodes, by.x="X1", by.y = "ids", all = T)
 
+friend.data.net.all <- rbind(friend.data.net1, friend.data.net2, friend.data.net3)
+names(friend.data.net.all)[1:2] <- c('from', 'to')
+friend.data.net.all$period <- rep(1:3, c(nrow(friend.data.net1),
+                                  nrow(friend.data.net2),
+                                  nrow(friend.data.net3)))
+
+ggplot(data=friend.data.net.all, aes(from_id = from, to_id = to)) +
+  geom_net(aes(colour = as.factor(drink)), size = .75, vjust = .5,
+           fontsize = 6, curvature = .25, ealpha = .6, directed = T,
+           arrowsize = .5, arrowgap = .01) + theme_net() + 
+  scale_color_brewer(palette = "YlOrRd", name = "Drinking Frequency", labels = c("never", "about 2x/year", "about 1x/month", "about 1x/week", "More than 1x/week")) + 
+  facet_wrap(~period, nrow = 1)
+@
+
+Several actor-oriented longitudinal network models can be fit to this data in RSiena.
+In order to determine the importance of certain actor-level variable effects included in the models, we simulated networks from different fitted models in Siena, and placed them in lineups to determine if there is any difference between the models.
+}
+
+<<fit_models>>=
+# First create a 50 * 50 * 3 array composed of the 3 adjacency matrices
+friendshipData <- array( c( friend.data.w1, friend.data.w2, friend.data.w3 ),
+                         dim = c( 50, 50, 3 ) )
+# and next give this the role of the dependent variable:
+friendship <- sienaDependent(friendshipData)
+alcohol <- varCovar( as.matrix(drink[,1:2]) )
+# create siena data object
+mydata <- sienaDataCreate( friendship, alcohol)
+
+myeff <- getEffects( mydata )
+# 3 actor effects
+myeff1 <- includeEffects( myeff, egoX, altX, egoXaltX, interaction1 = "alcohol" )
+# 1 actor effect 
+myeff2 <- includeEffects( myeff, egoX, interaction1 = "alcohol")
+# Create object with algorithm settings
+myalgorithm <- sienaAlgorithmCreate( projname = 's50' )
+#Estimate parameters
+ans1 <- siena07( myalgorithm, data = mydata, effects = myeff1)
+ans2 <- siena07( myalgorithm, data = mydata, effects = myeff2)
+
+# create data frames with fitted parameters to use later (need to find a more automated way to do this.)
+rate.params <- data.frame(rate.param = c("Rate", "Rate"),
+           rate.param.val = ans1$rate)
+eval.params <- data.frame(eval.param = c("density", "recip"), eval.param.val = ans1$theta[1:2])
+eval.int.params <- data.frame(eval.int.param = c("altX", "egoX", "egoXaltX"), 
+                              eval.int.param.val = ans1$theta[3:5])
+rate.params2 <- data.frame(rate.param = c("Rate", "Rate"),
+                          rate.param.val = ans2$rate)
+eval.params2 <- data.frame(eval.param = c("density", "recip"), eval.param.val = ans2$theta[1:2])
+eval.int.params2 <- data.frame(eval.int.param = "egoX", 
+                              eval.int.param.val = ans2$theta[3])
+@
+
+<<simulation_fn, echo=FALSE>>=
+SimulateNextWave <- function(init.waves, V0, 
+                                     rate.params.names, rate.params.vals,
+                                     eval.params.names, eval.params.vals,
+                                     eval.int.params.names, eval.int.params.vals){
+  if (ncol(init.waves[,,1]) != nrow(init.waves[,,1])){
+    return("Error: Non-square adjacency matrix provided")
+  } 
+  if (length(eval.params.vals) != length(eval.params.names)){
+    return("Error: Number of eval parameters to include in model does not
+           match number of eval parameter values provided.")
+  }
+  if (length(rate.params.vals) != length(rate.params.names)){
+    return("Error: Number of rate parameters to include in model does not
+           match number of rate parameter values provided.")
+  }
+  if (length(eval.int.params.vals) != length(eval.int.params.names)){
+    return("Error: Number of interaction eval parameters to include in model does not
+           match number of interaction eval parameter values provided.")
+  }
+  init.waves <- friendshipData
+  n <- ncol(init.waves[,,1])
+  V0 <- drink
+  c <- length(unique(V0[,1]))
+  p.e <- length(eval.params.names)
+  p.r <- length(rate.params.names)
+  p.e.i <- length(eval.int.params.names)
+  # Simulates M consecutive network and behavior waves, with n actors,
+  # with c categories of the behavior variable, from initial network init.net.adj
+  # according to a stochastic actor-oriented model with actor covariate V0
+  # with parameter values params.values for the parameters of interest params.names
+  Va <- varCovar(V0)
+  X   <- sienaDependent(init.waves)
+  InitData <- sienaDataCreate(X, Va)
+  InitEff0 <- getEffects(InitData)
+  # sink to avoid printing to the screen
+  #sink("eff.txt")
+  # Specify the parameters.
+  # The rate parameters are first multiplied by 10,
+  # which will be used only to get from the totally random network XX[,,1] = X0
+  # to the network that will be the simulated first wave.
+
+  for (i in 1:p.r){
+    InitEff0 <- setEffect(InitEff0, rate.params.names[i], type = 'rate', initialValue = rate.params.vals[i], period = i, character = T)
+  }
+  for (i in 1:p.e){
+    InitEff0 <- setEffect(InitEff0, eval.params.names[i], type = 'eval', initialValue = eval.params.vals[i], character = T)
+  }
+  for (i in 1:p.e.i){
+    InitEff0 <- setEffect(InitEff0, eval.int.params.names[i], type = 'eval', interaction1 = "Va", initialValue = eval.int.params.vals[i], character = T)
+  }
+
+  ## The parameter given for n3 should be larger than sum(InitEff0$include)
+  InitAlg <- sienaAlgorithmCreate(projname="Init", nsub=0,
+                                  cond=FALSE, simOnly=TRUE)
+  # Simulate the first wave.
+  InitSim   <- siena07(InitAlg, data=InitData, eff=InitEff0,
+                       returnDeps=TRUE, batch=TRUE, silent=TRUE)
+  return(data.frame(InitSim$sims[[1000]][[1]][[1]][[1]]))
+}  
+@
+
+<<simulate_network>>=
+# one from 3-effect model (the "data" model)
+sim_dat <- SimulateNextWave(init.waves = friendshipData, V0 = drink, 
+                 rate.params.names = rate.params$rate.param, 
+                 rate.params.vals = rate.params$rate.param.val,
+                 eval.params.names = eval.params$eval.param,
+                 eval.params.val = eval.params$eval.param.val,
+                 eval.int.params.names = eval.int.params$eval.int.param,
+                 eval.int.params.vals = eval.int.params$eval.int.param.val)
+sim_dat_net <- merge(sim_dat, data.frame(id =1:50), by.x = "X1", by.y = "id", all = T)
+
+# two from the 1-effect model (the "null" model)
+sim_null1 <- SimulateNextWave(init.waves = friendshipData, V0 = drink,                         rate.params.names = rate.params2$rate.param,
+                    rate.params.vals = rate.params2$rate.param.val,
+                      eval.params.names = eval.params2$eval.param,
+                     eval.params.val = eval.params2$eval.param.val,
+           eval.int.params.names = eval.int.params2$eval.int.param,
+        eval.int.params.vals = eval.int.params2$eval.int.param.val)
+sim_null1_net <- merge(sim_null1, data.frame(id =1:50), by.x = "X1", by.y = "id", all = T)
+sim_null2 <- SimulateNextWave(init.waves = friendshipData, V0 = drink, 
+                       rate.params.names = rate.params2$rate.param, 
+                     rate.params.vals = rate.params2$rate.param.val,
+                       eval.params.names = eval.params2$eval.param,
+                     eval.params.val = eval.params2$eval.param.val,             eval.int.params.names = eval.int.params2$eval.int.param,         eval.int.params.vals = eval.int.params2$eval.int.param.val)
+sim_null2_net <- merge(sim_null2, data.frame(id =1:50), by.x = "X1", by.y = "id", all = T)
+
+# null is smaller model
+small_lu <- rbind(sim_dat_net, sim_null1_net,sim_null2_net)
+small_lu$group <- rep(sample.int(3), c(nrow(sim_dat_net), nrow(sim_null1_net), nrow(sim_null2_net)))
+
+ggplot(small_lu, aes(from_id = X1, to_id = X2)) + 
+  geom_net(ealpha = .7, fiteach = T, directed = T, size = 1, arrowsize = .5) + 
+  theme_net() + 
+  facet_wrap(~group, nrow=1)
+@
+
+The ``data" model is facet \Sexpr{small_lu$group[1]}.
 Notice the caption at the top of Table~\ref{vis_accept}.

### 2016-02-10 paper/network-inference.Rnw
just worked through the code and started labelling the first example

@@
 \maketitle
+<<knitr-setup, fig.keep='all', cache=FALSE, echo=FALSE, eval=TRUE, include=FALSE>>=
+#rm(list=ls())
+options(replace.assign=TRUE,width=70)
+require(knitr)
+opts_chunk$set(fig.path='figure/', cache.path='cache/', fig.align='center', fig.width=5, fig.height=5, par=TRUE, cache=TRUE, concordance=TRUE, autodep=TRUE, message=F, warning=F)
+@
 
@@
 \st{
-One of the data sets provided with RSiena is a subset of 50 female respondants from the \emph{Teenage Friends and Lifestyle Study} XXX citation???
-XXX.
+One of the data sets provided with RSiena is a subset of 50 female respondents from the \emph{Teenage Friends and Lifestyle Study} XXX citation???
+from \url{https://www.stats.ox.ac.uk/~snijders/siena/s50_data.htm} XXX.
 There are three friendship networks provided at three timepoints representing the relationships between the 50 teenage girls in the data.
 There are also actor-level variables observed at each time period, such as a girl's drinking and smoking behavior.
-These networks are shown in \autoref{fig:friendship_waves}
-XX autoref not working?? XX
-<<friendship_waves, echo=FALSE, message=FALSE, warning=FALSE>>=
+These networks are shown in Figure~\ref{fig:friendship_waves}.
+
+\begin{figure}
+<<friendship_waves, echo=FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=
 friend.data.w1 <- as.matrix(read.table("../Data/s50_data/s50-network1.dat"))
@@
                                   nrow(friend.data.net3)))
+friend.data.net.all$year <- friend.data.net.all$period
 
@@
   scale_color_brewer(palette = "YlOrRd", name = "Drinking Frequency", labels = c("never", "about 2x/year", "about 1x/month", "about 1x/week", "More than 1x/week")) + 
-  facet_wrap(~period, nrow = 1)
+  facet_wrap(~year, nrow = 1, labeller="label_both") +
+  theme(legend.position="bottom") +
+  guides(col = guide_legend(nrow = 2, override.aes = list(size=3.5)))
 @
-
+\caption{\label{fig:friendship_waves}\hh{This figure needs some caption XXX}}
+\end{figure}
 Several actor-oriented longitudinal network models can be fit to this data in RSiena.
@@
 
-<<fit_models>>=
+<<fit_models, echo=FALSE, message=FALSE, results='hide'>>=
 # First create a 50 * 50 * 3 array composed of the 3 adjacency matrices
@@
 
-<<simulation_fn, echo=FALSE>>=
+<<simulation_fn, echo=FALSE, message=FALSE>>=
 SimulateNextWave <- function(init.waves, V0, 
@@
 
-<<simulate_network>>=
+<<simulate_network, echo=FALSE, message=FALSE, results='hide'>>=
 # one from 3-effect model (the "data" model)

### 2016-02-11 paper/network-inference.Rnw
simulated figure network up

@@
 \begin{figure}
-<<friendship_waves, echo=FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=3.5, out.width='\\linewidth'>>=
+<<friendship_waves, echo=FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=3.35, out.width='\\linewidth'>>=
 friend.data.w1 <- as.matrix(read.table("../Data/s50_data/s50-network1.dat"))
@@
 small_lu$group <- rep(sample.int(3), c(nrow(sim_dat_net), nrow(sim_null1_net), nrow(sim_null2_net)))
+@
 
+\begin{figure}
+<<friendship_simulated, dependson='simulate_network', echo=FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=3.35, out.width='\\linewidth'>>=
 ggplot(small_lu, aes(from_id = X1, to_id = X2)) + 
@@
 @
+\caption{\label{fig:simulated}}
+\end{figure}
 
 The ``data" model is facet \Sexpr{small_lu$group[1]}.
+
 Notice the caption at the top of Table~\ref{vis_accept}.

### 2016-02-11 paper/network-inference.Rnw
references added

@@
 
+References for the intro: lineup protocol~\cite{buja:2009}, RSiena~\cite{RSiena}.
+
 \begin{enumerate}
@@
 
-Possible section outline for paper:
+Possible section outline for paper: \hh{XXX the  outline looks good XXX }
 \begin{enumerate}
@@
 \item Data and Model fitting \& simulation
-\item Lineup methods
+\item Lineup methods \hh{XXX what do you mean by `methods' here? Do you mean the Experimental Design?}
 \item Results
@@
 
-\section{Examples \st{??}}
+\section{Examples \st{??} \hh{this might be able to go into the data and model fitting section.
+We need both a theoretical bit on the models (at the beginning of the section), and then a discussion of the examples used in the paper.
+}}
 
 \st{
-One of the data sets provided with RSiena is a subset of 50 female respondents from the \emph{Teenage Friends and Lifestyle Study} XXX citation???
-from \url{https://www.stats.ox.ac.uk/~snijders/siena/s50_data.htm} XXX.
-There are three friendship networks provided at three timepoints representing the relationships between the 50 teenage girls in the data.
-There are also actor-level variables observed at each time period, such as a girl's drinking and smoking behavior.
+One of the data sets provided with RSiena is a subset of 50 female respondents from the \emph{Teenage Friends and Lifestyle Study} \cite{michell:1997, pearson:2000, pearson:2003}.
+There are three friendship networks provided  \hh{are the networks really provided?
+I thought RSiena estimated the networks based on the girls' drinking and smoking behaviour?} at three timepoints representing the relationships between the 50 teenage girls in the data.
+There are also actor-level variables observed at each time period, such as a girls' drinking and smoking behavior.
 These networks are shown in Figure~\ref{fig:friendship_waves}.
@@
 
-The ``data" model is facet \Sexpr{small_lu$group[1]}.
-
-Notice the caption at the top of Table~\ref{vis_accept}.
-
-\begin{table}
-%% Table captions on top in journal version
- \caption{SciVis Paper Acceptance Rate: 1994-2006}
- \label{vis_accept}
- \scriptsize
- \begin{center}
-   \begin{tabular}{cccc}
-     Year & Submitted & Accepted & Accepted (\%)\\
-   \hline
-     2004 & 167 & 46 & 27.6\\
-     2005 & 268 & 88 & 32.8\\
-     2006 & 228 & 63 & 27.6
-   \end{tabular}
- \end{center}
-\end{table}
+The ``data" model is in panel \#\Sexpr{small_lu$group[1]}.
 
@@
 
-Citing happens only with cite - like this for the lineup protocol~\cite{buja:2009}, and for RSiena~\cite{RSiena}. Footnotes are ok%
+Footnotes are ok%
 \footnote{Footnotes appear at the bottom of the column}.

### 2016-02-15 paper/network-inference.Rnw
add some writing to paper

@@
 @
-\caption{\label{fig:friendship_waves}\hh{This figure needs some caption XXX}}
+\caption{\label{fig:friendship_waves}\st{The friendship network of 50 girls in the \emph{Teenage Friends and Lifestyle Study} over three time periods.
+The nodes, which represent the girls in the study, are colored by their level of alcohol use.}}
 \end{figure}
@@
 In order to determine the importance of certain actor-level variable effects included in the models, we simulated networks from different fitted models in Siena, and placed them in lineups to determine if there is any difference between the models.
+
+\par In the fitting of models to the Friends data, there are four effects that are automatically included in model estimation.
+Two of these are rate parameters for moving from time point one to two and from time point two to three.
+The other two are the outdegree and the reciprocity effects.
+The outdegree is just an edge count from each node, and the reciprocity is the count of reciprocated ties from each node.
+XXX insert more about the math!
+XXX The documentation and code examples from the RSiena webpage include models to fit to the Friends' data.
+One of these includes the effects covariate-alter, covariate-ego, and covariate-alter X ego, which are measures of the behavior of a node and its neighbors.
+XXX insert more math XXX.
+The RSiena and RSienaTest packages include functions to test the significance of fitted parameters via t-tests, wald tests, and score-type tests.
+I used the wald tests and the score-type test to assess the significance of the parameters in this friends model with three covariate effects.
+The tests found the two rate parameters and the outdegree and reciprocity parameters were significant, but of the covariate effects, only the covariate-alter X ego effect was significant.
+So, we try to determine if there was a difference visually between the networks with just the four network level variables included and the networks with the significant covariate parameter included.
 }

### 2016-02-15 Code/RSiena_mySimFn.R

@@ -62,5 +62,5 @@ SimulateNextWave <- function(init.waves, V0,
   # Simulate the first wave.
   InitSim   <- siena07(InitAlg, data=InitData, eff=InitEff0,
                        returnDeps=TRUE, batch=TRUE, silent=TRUE)
-  return(data.frame(InitSim$sims[[1000]][[1]][[1]][[1]]))
+  return(InitSim$sims)
 }   

### 2016-02-15 Code/friendship.R

@@ -7,32 +7,58 @@
 setwd("s50_data")
 library(RSiena) # or RSienaTest
 # read in network data
-friend.data.w1 <- as.matrix(read.table("s50-network1.dat"))
-friend.data.w2 <- as.matrix(read.table("s50-network2.dat"))
-friend.data.w3 <- as.matrix(read.table("s50-network3.dat"))
+friend.data.w1 <- as.matrix(read.table("Data/s50_data/s50-network1.dat"))
+friend.data.w2 <- as.matrix(read.table("Data/s50_data/s50-network2.dat"))
+friend.data.w3 <- as.matrix(read.table("Data/s50_data/s50-network3.dat"))
 # read in covariate data
-drink <- as.matrix(read.table("s50-alcohol.dat"))
-smoke <- as.matrix(read.table("s50-smoke.dat"))
+drink <- as.matrix(read.table("Data/s50_data/s50-alcohol.dat"))
+smoke <- as.matrix(read.table("Data/s50_data/s50-smoke.dat"))
 # First create a 50 * 50 * 3 array composed of the 3 adjacency matrices
 friendshipData <- array( c( friend.data.w1, friend.data.w2, friend.data.w3 ),
                          dim = c( 50, 50, 3 ) )
 # and next give this the role of the dependent variable:
 friendship <- sienaDependent(friendshipData)
-alcohol <- varCovar( as.matrix(drink[,1:2]) )
+alcohol <- varCovar( as.matrix(drink) )
 # create siena data object
 mydata <- sienaDataCreate( friendship, alcohol)
 
-myeff <- getEffects( mydata )
-myeff1 <- includeEffects( myeff, egoX, altX, egoXaltX, interaction1 = "alcohol" )
-myeff2 <- includeEffects( myeff, simX, interaction1 = "alcohol")
+null_model_eff <- getEffects(mydata)
+test_model_eff <- includeEffects( null_model_eff, egoXaltX, interaction1 = "alcohol")
+
 # Create object with algorithm settings
 myalgorithm <- sienaAlgorithmCreate( projname = 's50' )
 #Estimate parameters
-ans1 <- siena07( myalgorithm, data = mydata, effects = myeff1)
-ans2 <- siena07( myalgorithm, data = mydata, effects = myeff2)
[…]

### 2016-02-17 paper/network-inference.Rnw
couple of comments

@@
 @
-\caption{\label{fig:friendship_waves}\st{The friendship network of 50 girls in the \emph{Teenage Friends and Lifestyle Study} over three time periods.
-The nodes, which represent the girls in the study, are colored by their level of alcohol use.}}
+\caption{\label{fig:friendship_waves}Friendship networks of 50 girls in the \emph{Teenage Friends and Lifestyle Study} over three time periods.
+Each node represents one of the girls.
+Nodes are colored by the girl's level of alcohol use.}
 \end{figure}
@@
 The outdegree is just an edge count from each node, and the reciprocity is the count of reciprocated ties from each node.
-XXX insert more about the math!
+\hh{I'm not sure whether the math should go here or at the beginning - actor models still need to be introduced.} XXX insert more about the math!
 XXX The documentation and code examples from the RSiena webpage include models to fit to the Friends' data.
@@
 XXX insert more math XXX.
-The RSiena and RSienaTest packages include functions to test the significance of fitted parameters via t-tests, wald tests, and score-type tests.
-I used the wald tests and the score-type test to assess the significance of the parameters in this friends model with three covariate effects.
+The RSiena and RSienaTest packages include functions to test the significance of fitted parameters via t-tests, Wald tests, and score-type tests.
+We used the wald tests and the score-type test to assess the significance of the parameters in this friends model with three covariate effects.
 The tests found the two rate parameters and the outdegree and reciprocity parameters were significant, but of the covariate effects, only the covariate-alter X ego effect was significant.

### 2016-02-17 paper/network-inference.Rnw
working on sims

--- 2016-02-15 paper/network-inference.Rnw
@@
 \end{figure}
+
 Several actor-oriented longitudinal network models can be fit to this data in RSiena.
@@
 So, we try to determine if there was a difference visually between the networks with just the four network level variables included and the networks with the significant covariate parameter included.
+
+\par To simulate networks from these two models, we first fit RSiena models with the default effects and with the default effects plus the significant actor-level covariate variable.
+After obtaining the estimates of the effects, we simulated
 }

### 2016-02-17 Code/friendship.R

@@ -26,10 +26,60 @@ null_model_eff <- getEffects(mydata)
 test_model_eff <- includeEffects( null_model_eff, egoXaltX, interaction1 = "alcohol")
 
 # Create object with algorithm settings
-myalgorithm <- sienaAlgorithmCreate( projname = 's50' )
+myalgorithm <- sienaAlgorithmCreate( projname = 's50' , n3 = 1500)
 #Estimate parameters
-ests_null <- siena07( myalgorithm, data = mydata, effects = null_model_eff)
-ests_test <- siena07( myalgorithm, data = mydata, effects = test_model_eff)
+ests_null <- siena07( myalgorithm, data = mydata, returnDeps = TRUE, effects = null_model_eff, batch=TRUE, verbose = FALSE)
+ests_test <- siena07( myalgorithm, data = mydata, returnDeps = TRUE, effects = test_model_eff, batch=TRUE, verbose = FALSE)
+
+# How do the simulations look compared to each other? 
+# i.e. how much does the network evolve in simulation
+
+# pull out the last 10 networks for periods 1 (wave 1->2) and 2 (wave 2->3)
+period1 <- NULL
+for (i in 1191:1200){
+  getnet <- merge(data.frame(ests_null$sims[[i]][[1]][[1]][[1]])[,-3], 
+                  data.frame(id = 1:50), by.x = "X1", by.y = "id",
+                  all = T)
+  getnet$count <- paste(i, "wave2") 
+  period1 <- rbind(period1, getnet)
+}
+
+period2 <- NULL
+for (i in 1191:1200){
+  getnet <- merge(data.frame(ests_null$sims[[i]][[1]][[1]][[2]])[,-3], 
+                  data.frame(id = 1:50), by.x = "X1", by.y = "id",
+                  all = T)
+  getnet$count <- paste(i, "wave3") 
+  period2 <- rbind(period2, getnet)
+}
+
+library(geomnet)
+
+# scroll down for actual2, actual3
+period1 <- rbind(period1, actual2)
+ggplot(data = period1, aes(from_id = X1, to_id = X2)) +
+  geom_net(fiteach = TRUE) + theme_net() +
+  facet_wrap(~count)
+
+period2 <- rbind(period2, actual3)
+ggplot(data = period2, aes(from_id = X1, to_id = X2)) +
[…]

### 2016-02-17 paper/network-inference.Rnw

--- 2016-02-17 paper/network-inference.Rnw
@@
 So, we try to determine if there was a difference visually between the networks with just the four network level variables included and the networks with the significant covariate parameter included.
+
+\par To simulate networks from these two models, we first fit RSiena models with the default effects and with the default effects plus the significant actor-level covariate variable.
+After obtaining the estimates of the effects, we simulated
+
 }

### 2016-02-17 paper/network-inference.Rnw
add blurb to paper about models

@@
 \item Data and Model fitting \& simulation
-\item Lineup methods \hh{XXX what do you mean by `methods' here? Do you mean the Experimental Design?}
+\item Lineup methods \hh{XXX what do you mean by `methods' here? Do you mean the Experimental Design?} \st{XXX I think that's what I mean.}
 \item Results
@@
 
-\section{Examples \st{??} \hh{this might be able to go into the data and model fitting section.
-We need both a theoretical bit on the models (at the beginning of the section), and then a discussion of the examples used in the paper.
-}}
+\section{Data, Model Fit, and Simulation}
+
+\hh{We need both a theoretical bit on the models (at the beginning of the section), and then a discussion of the examples used in the paper. }
 
 \st{
+\subsection{Network Models}
+
+XXX should we just focus on the stochastic, actor oriented ones? or should we add additional ones? XXX
+
+\par One network model we examine is the stochastic actor-oriented models introduced by Snijders~\cite{Snijders:1996}.
+Social networks are generally not static: they evolve over time and this evolution is influenced by the actors and the network structure.
+Estimation and implementation of these models can be done in the RSIENA software~\cite{RSiena}.
+The software allows many possible effects to be included in this type of model, given below.
+XX insert math after I read the Snijders papers XX  Some of the network effects and actor effects we look at are XXX insert list of effects XXX.
+
+\subsection{Data}
 One of the data sets provided with RSiena is a subset of 50 female respondents from the \emph{Teenage Friends and Lifestyle Study} \cite{michell:1997, pearson:2000, pearson:2003}.

### 2016-02-22 paper/network-inference.Rnw
add model description

@@
 
-\par One network model we examine is the stochastic actor-oriented models introduced by Snijders~\cite{Snijders:1996}.
+\par One network model we examine is the stochastic actor-oriented models (SAOM) introduced by Snijders~\cite{Snijders:1996}.
 Social networks are generally not static: they evolve over time and this evolution is influenced by the actors and the network structure.
-Estimation and implementation of these models can be done in the RSIENA software~\cite{RSiena}.
+In SAOMs, each node, or actor, has its own objective function, $f_i(\beta, x)$ where $i$ is the node index, and $\beta$ is a vector of $p$ coefficients for the $p$ effects in a model, $s_{ki}(x)$ for a given network state $x$, effect $k = 1, ..., p$ and node $i$.
+The values of each $\beta_k$ are estimated with all nodes and ties in consideration and are applied to all nodes' objective functions.
+\par At random moments (occurring with rate $\rho$), a single actor $i$ is permitted to change one of its ties, $x_{ij}$.
+Or, the actor can do nothing.
+These single edge change moments are unobserved.
+They are only observed in aggregate when the network is observed at discrete time points.
+\par When given the opportunity to change or not, each actor tries to optimize its own objective function $f_i(\beta,x)$ under the current configuration of the network (denoted by $x$) in the observed time point.
+The probability that node $i$ will change a tie $x_ij$ when given the opportunity to act is $$ p_{ij} = \frac{\exp(f(\beta,x| x_{ij}=x'_{ij}) - f(\beta, x |x_{ij}=x_{ij}))}{\sum_{h=1}^n \exp(f(\beta, x |x_{ih}=x'_{ih}) - f(\beta, x|x_{ih}=x_{ih}))}$$
+
+where $x'_{ij}$ is the opposite of the current state of $x_{ij}$.
+
+\par Estimation and implementation of these models can be done in the RSIENA software~\cite{RSiena}.
 The software allows many possible effects to be included in this type of model, given below.

### 2016-03-10 Code/small_friends_lineup.R
change border color in svgs

@@ -26,13 +26,13 @@ create_smfriend_lu <- function(null_eff_struct, test_eff_struct, M, my_dat=mysma
   to_plot <- rbind(null_data, test_data)
   to_plot$plot_order <- rep(sample(M), as.vector(table(to_plot$count)))
   plot <- ggplot(data = to_plot, aes(from_id = X1, to_id = X2)) + 
-            geom_net(fiteach = TRUE, directed = T, size = 1, arrowsize = .5) +
-            facet_wrap(~plot_order) + theme_net()
+            geom_net(fiteach = TRUE, directed = F, size = 1, arrowsize = .5) +
+            facet_wrap(~plot_order) + theme_net() + theme(panel.background = element_rect(fill = "white", color = 'white'))
   data_plot_id <- unique(to_plot$plot_order[which(to_plot$count == M)])
   return(list(data = to_plot, lineup = plot, test_id = data_plot_id))
 }
 
-lu1 <- create_smfriend_lu(null_eff_struct = null_model_eff2, test_eff_struct = eff_models_smallFriends[[39]], M = 9)
+lu1 <- create_smfriend_lu(null_eff_struct = null_model_eff2, test_eff_struct = eff_models_smallFriends[[39]], M = 6)
 lu1$data
 lu1$lineup
 lu1$test_id

### 2016-03-10 paper/network-inference.Rnw
add some math

--- 2016-02-17 paper/network-inference.Rnw
@@
 \begin{enumerate}
-\item Need to complete lineups in 2-3 weeks
-\item Lineups: 
-  \begin{enumerate}
+\item Lineups to test
   \item What sizes? 3, 6, 9, 12
-  \item What data? Football, olga's data, teenage friends data?
-  \item Other things to consider: coloring of nodes? 
-  \item use replicates: different seeds for the same data produces different layouts -
-  different draws from the model
-  \end{enumerate}
-\item Get simulated networds from models from RSiena
-\item Figure out another way to fit a model and then simulate from it? 
+  \item Maybe one other really significant effect
+  \item facetting - same layout for all or different layouts in each? 
 \end{enumerate}
 
-Possible section outline for paper: \hh{XXX the  outline looks good XXX }
+Outline for paper:
 \begin{enumerate}
@@
 \item Data and Model fitting \& simulation
-\item Lineup methods \hh{XXX what do you mean by `methods' here? Do you mean the Experimental Design?} \st{XXX I think that's what I mean.}
+\item Experiment / Experimental Design
 \item Results
@@
 \st{
-\subsection{Network Models}
+\subsection{Stochastic Actor-Oriented Models}
 
-XXX should we just focus on the stochastic, actor oriented ones? or should we add additional ones? XXX
+\par Stochastic actor-oriented models (SAOMs) were introduced in Snijders~\cite{Snijders:1996}.
+Social networks are generally not static: they evolve over time.
+This evolution is influenced by both the actors' behaviors and the network structure.
+In SAOMs, each actor, or node, has its own objective function, $f_i(\beta, x)$ where $i$ is the node index, and $\beta$ is a vector of $p$ coefficients for the $p$ effects in a model, $s_{ki}(x)$ for a given network state $x$, effect $k = 1, ..., p$ and node $i$.
+The full form of the objective function is $$ f_i(\beta, x) = \sum_k \beta_k s_{ik}(x).
+$$ The values of each $\beta_k$ are estimated with all nodes and ties in consideration and are applied to all nodes' objective functions.
+\par At random moments occurring with rate $\rho_m$, $m = 1, \dots, M-1$ where $M$ is the number of observed waves of a network, a single actor $i$ is permitted to change one of its ties, $x_{ij}$.
+Or, the actor can do nothing.
+These single edge change moments are unobserved.
+They are only observed in aggregate when the network is observed at discrete time points.
+So, multiple actors may change their ties between wave $m-1$ and wave $m$.
+\par When given the opportunity to change or not, each actor tries to optimize its own objective function $f_i(\beta,x)$ under the current configuration of the network (denoted by $x$) in the observed time point.
+The probability that node $i$ will change a tie $x_ij$ when given the opportunity to act is $$ p_{ij} = \frac{\exp(f(\beta,x| x_{ij}=x'_{ij}) - f(\beta, x |x_{ij}=x_{ij}))}{\sum_{h=1}^n \exp(f(\beta, x |x_{ih}=x'_{ih}) - f(\beta, x|x_{ih}=x_{ih}))}$$
+where $x'_{ij}$ is the opposite of the current state of $x_{ij} ~\cite{Snijders:1996}$.
 
-\par One network model we examine is the stochastic actor-oriented models introduced by Snijders~\cite{Snijders:1996}.
-Social networks are generally not static: they evolve over time and this evolution is influenced by the actors and the network structure.
-Estimation and implementation of these models can be done in the RSIENA software~\cite{RSiena}.
-The software allows many possible effects to be included in this type of model, given below.
-XX insert math after I read the Snijders papers XX  Some of the network effects and actor effects we look at are XXX insert list of effects XXX.
+\par Estimation and implementation of these models can be done in the RSIENA software~\cite{RSiena}.
+The software allows many possible effects to be included in this type of model.
+Some of the network effects and actor effects we look at are:
+\begin{table}  
+  \begin{tabular}{llc}
+  Name & Type & Formula \\ 
+  \hline
+  period 1 rate & Rate & $\rho_1$ \\
+  outdegree & Structural & $s_{i1}(x) = \sum_j x_{ij}$ \\
+  reciprocity & Structural & $s_{i2}(x) = \sum_j x_{ij}x_{ji} $
+  \end{tabular}
+\end{table}
 

### 2016-03-10 paper/network-inference.Rnw
more examples of effects

@@
 The software allows many possible effects to be included in this type of model.
-Some of the network effects and actor effects we look at are:
+Some of the network effects and actor effects that are possible to include in SAOMs are given in Table ~\ref{tab:effects}.
 \begin{table}  
@@
   \hline
-  period 1 rate & Rate & $\rho_1$ \\
+  period $m$ rate & Rate & $\rho_{m-1}$ \\
   outdegree & Structural & $s_{i1}(x) = \sum_j x_{ij}$ \\
-  reciprocity & Structural & $s_{i2}(x) = \sum_j x_{ij}x_{ji} $
+  reciprocity & Structural & $s_{i2}(x) = \sum_j x_{ij}x_{ji} $ \\
+  transitive triplets & Structural & $s_{i3}(x) = \sum_{j,h} x_{ij}x_{jh}x_{ih}$
+
   \end{tabular}
+  \caption{\label{tab:effects}}
 \end{table}

### 2016-03-11 paper/network-inference.Rnw
more effects

@@
 
-\usepackage{mathptmx}
+\usepackage{mathptmx, amsmath, amssymb}
 \usepackage{graphicx}
@@
 $$ The values of each $\beta_k$ are estimated with all nodes and ties in consideration and are applied to all nodes' objective functions.
-\par At random moments occurring with rate $\rho_m$, $m = 1, \dots, M-1$ where $M$ is the number of observed waves of a network, a single actor $i$ is permitted to change one of its ties, $x_{ij}$.
+\par At random moments occurring with rate $\rho_{m-1}$, $m = 2, \dots, M$ where $M \geq 2$ is the number of observed waves of a network, a single actor $i$ is permitted to change one of its ties, $x_{ij}$.
 Or, the actor can do nothing.
@@
   reciprocity & Structural & $s_{i2}(x) = \sum_j x_{ij}x_{ji} $ \\
-  transitive triplets & Structural & $s_{i3}(x) = \sum_{j,h} x_{ij}x_{jh}x_{ih}$
-
+  transitive triplets & Structural & $s_{i3}(x) = \sum_{j,h} x_{ij}x_{jh}x_{ih}$ \\
+  covariate-alter & Covariate & $s_{i4}(x) = \sum_j x_{ij}v_j $ \\
+  covariate-ego & Covariate & $ s_{i5}(x) = v_i\sum_j x_{ij}$ \\
+  same covariate & Covariate & $ s_{i6}(x) = \sum_j x_{ij} \mathbb{I}(v_i = v_j)$ \\
+  jumping $V$ transitive triplets & Covariate & $ s_{i7}(x) = \sum_{j \neq h} x_{ij}x_{ih}x_{hj} \mathbb{I}(v_i = v_h \neq v_j)$
   \end{tabular} 
-  \caption{\label{tab:effects}}
+  \caption{\label{tab:effects} Some of the possible effects to be included in the stochastic actor-oriented models in RSiena.
+There are many more possible effects, but we only consider a select few here.
+For a complete list, see the RSiena manual ~\cite{RSiena}.}
 \end{table}

### 2016-03-11 paper/network-inference.Rnw
typo

@@
 This evolution is influenced by both the actors' behaviors and the network structure.
-In SAOMs, each actor, or node, has its own objective function, $f_i(\beta, x)$ where $i$ is the node index, and $\beta$ is a vector of $p$ coefficients for the $p$ effects in a model, $s_{ki}(x)$ for a given network state $x$, effect $k = 1, ..., p$ and node $i$.
+In SAOMs, each actor, or node, has its own objective function, $f_i(\beta, x)$ where $i$ is the node index, and $\beta$ is a vector of $p$ coefficients for the $p$ effects in a model, $s_{ik}(x)$ for a given network state $x$, effect $k = 1, ..., p$ and node $i$.
 The full form of the objective function is $$ f_i(\beta, x) = \sum_k \beta_k s_{ik}(x).
@@
 \end{table}
+
+\par To estimate the parameter values, $\beta_k$, $k = 1, \dots, p$, the relevant statistics are
 

### 2016-03-13 paper/network-inference.Rnw
couple of comments

@@
 
-\st{
 \subsection{Stochastic Actor-Oriented Models}
 
-\par Stochastic actor-oriented models (SAOMs) were introduced in Snijders~\cite{Snijders:1996}.
-Social networks are generally not static: they evolve over time.
-This evolution is influenced by both the actors' behaviors and the network structure.
-In SAOMs, each actor, or node, has its own objective function, $f_i(\beta, x)$ where $i$ is the node index, and $\beta$ is a vector of $p$ coefficients for the $p$ effects in a model, $s_{ik}(x)$ for a given network state $x$, effect $k = 1, ..., p$ and node $i$.
-The full form of the objective function is $$ f_i(\beta, x) = \sum_k \beta_k s_{ik}(x).
-$$ The values of each $\beta_k$ are estimated with all nodes and ties in consideration and are applied to all nodes' objective functions.
-\par At random moments occurring with rate $\rho_{m-1}$, $m = 2, \dots, M$ where $M \geq 2$ is the number of observed waves of a network, a single actor $i$ is permitted to change one of its ties, $x_{ij}$.
-Or, the actor can do nothing.
+\par Social networks are generally not static: they evolve over time. This evolution is influenced by both the actors' behaviors and the network structure. To deal with this continuous evolution, 
+stochastic actor-oriented models (SAOMs) were introduced by Snijders in~\cite{Snijders:1996}. 
+In a SAOM each actor, represented by a node, has its own objective function, $f_i(\beta, x)$ where $i$ is the node index, and $\beta$ is a vector of $p$ coefficients for the $p$ effects in a model.
+The full form of the objective function is
+\begin{equation}\label{eq:SAOM}
+f_i(\beta, x) = \sum_k \beta_k s_{ik}(x),
+\end{equation}
+where $s_{ik}(x)$ \hh{is a structural function ... ??? XXX do these functions have a specific name?} for network state $x$, effect $k = 1, ..., p$ and node $i$.
+
+The values of each $\beta_k$ are estimated with all nodes and ties in consideration and are applied to all nodes' objective functions.
+\par
+\hh{XXX describe what happens to each individual actor first, then talk about the aggregate. Refer to the table to make the $s_{ik}$ more concrete.}
+\st{At random moments occurring with rate $\rho_{m-1}$, $m = 2, \dots, M$ where $M \geq 2$ is the number of observed waves of a network, a single actor $i$ is permitted to change one of its ties, $x_{ij}$.
+Alternatively, the actor can do nothing.
 These single edge change moments are unobserved.
@@
 The probability that node $i$ will change a tie $x_ij$ when given the opportunity to act is $$ p_{ij} = \frac{\exp(f(\beta,x| x_{ij}=x'_{ij}) - f(\beta, x |x_{ij}=x_{ij}))}{\sum_{h=1}^n \exp(f(\beta, x |x_{ih}=x'_{ih}) - f(\beta, x|x_{ih}=x_{ih}))}$$
-where $x'_{ij}$ is the opposite of the current state of $x_{ij} ~\cite{Snijders:1996}$.
+where $x'_{ij}$ is the opposite of the current state of $x_{ij} ~\cite{Snijders:1996}$. }
 
@@
 After obtaining the estimates of the effects, we simulated
-
-}
 

### 2016-03-15 paper/network-inference.Rnw
table re-formatted

@@
 Some of the network effects and actor effects that are possible to include in SAOMs are given in Table ~\ref{tab:effects}.
+%
 \begin{table}  
-  \begin{tabular}{llc}
-  Name & Type & Formula \\ 
-  \hline
-  period $m$ rate & Rate & $\rho_{m-1}$ \\
-  outdegree & Structural & $s_{i1}(x) = \sum_j x_{ij}$ \\
-  reciprocity & Structural & $s_{i2}(x) = \sum_j x_{ij}x_{ji} $ \\
-  transitive triplets & Structural & $s_{i3}(x) = \sum_{j,h} x_{ij}x_{jh}x_{ih}$ \\
-  covariate-alter & Covariate & $s_{i4}(x) = \sum_j x_{ij}v_j $ \\
-  covariate-ego & Covariate & $ s_{i5}(x) = v_i\sum_j x_{ij}$ \\
-  same covariate & Covariate & $ s_{i6}(x) = \sum_j x_{ij} \mathbb{I}(v_i = v_j)$ \\
-  jumping $V$ transitive triplets & Covariate & $ s_{i7}(x) = \sum_{j \neq h} x_{ij}x_{ih}x_{hj} \mathbb{I}(v_i = v_h \neq v_j)$
+  \begin{tabular}{ll}
+%\bf  Name  & \bf Formula \\ 
+%  \hline
+%  \multicolumn{2}{l}{\bf Rate parameter}\\
+%  period $m$ rate & $\rho_{m-1}$ \\[3pt]
+  \multicolumn{2}{l}{\bf Structural Effects}\\
+  outdegree  & $s_{i1}(x) = \sum_j x_{ij}$ \\
+  reciprocity  & $s_{i2}(x) = \sum_j x_{ij}x_{ji} $ \\
+  transitive triplets  & $s_{i3}(x) = \sum_{j,h} x_{ij}x_{jh}x_{ih}$ \\[3pt]
+  \multicolumn{2}{l}{\bf Covariate Effects}\\
+  covariate-alter  & $s_{i4}(x) = \sum_j x_{ij}v_j $ \\
+  covariate-ego  & $ s_{i5}(x) = v_i\sum_j x_{ij}$ \\
+  same covariate & $ s_{i6}(x) = \sum_j x_{ij} \mathbb{I}(v_i = v_j)$ \\
+  jumping $V$ transitive triplets  & $ s_{i7}(x) = \sum_{j \neq h} x_{ij}x_{ih}x_{hj} \mathbb{I}(v_i = v_h \neq v_j)$
   \end{tabular} 
@@
 \end{table}
-
+%
 \par To estimate the parameter values, $\beta_k$, $k = 1, \dots, p$, the relevant statistics are

### 2016-03-15 paper/network-inference.Rnw
table

@@
 stochastic actor-oriented models (SAOMs) were introduced by Snijders in~\cite{Snijders:1996}. 
-In a SAOM each actor, represented by a node, has its own objective function, $f_i(\beta, x)$ where $i$ is the node index, and $\beta$ is a vector of $p$ coefficients for the $p$ effects in a model.
+In a SAOM each actor is represented by a node.
+Each node has its own objective function, $f_i(\beta, x)$ where $i$ is the node index, and $\beta$ is a vector of $p$ coefficients for the $p$ effects in a model.
 The full form of the objective function is
@@
 \par When given the opportunity to change or not, each actor tries to optimize its own objective function $f_i(\beta,x)$ under the current configuration of the network (denoted by $x$) in the observed time point.
-The probability that node $i$ will change a tie $x_ij$ when given the opportunity to act is $$ p_{ij} = \frac{\exp(f(\beta,x| x_{ij}=x'_{ij}) - f(\beta, x |x_{ij}=x_{ij}))}{\sum_{h=1}^n \exp(f(\beta, x |x_{ih}=x'_{ih}) - f(\beta, x|x_{ih}=x_{ih}))}$$
-where $x'_{ij}$ is the opposite of the current state of $x_{ij} ~\cite{Snijders:1996}$. }
+The probability that actor $i$  changes a tie $x_{ij}$ \hh{to its opposite $x'_{ij}$ }  when given the opportunity to act is~\cite{Snijders:1996} \hh{could you first describe this function in words and then give the formula?
+It looks like a comparison of odds similar to a conditional logit model or a discrete choice model~\cite{benatkiva:1985}}
+$$ 
+p_{ij} = \frac{\exp(f(\beta,x| x_{ij}=x'_{ij}) - f(\beta, x |x_{ij}=x_{ij}))}{\sum_{h=1}^n \exp(f(\beta, x |x_{ih}=x'_{ih}) - f(\beta, x|x_{ih}=x_{ih}))}.
+$$ 
+\hh{Sam, could you check in the Snijder paper again about the actual form of the denominator? could it be two sums rather than the one?} }
 
@@
   \multicolumn{2}{l}{\bf Structural Effects}\\
-  outdegree  & $s_{i1}(x) = \sum_j x_{ij}$ \\
-  reciprocity  & $s_{i2}(x) = \sum_j x_{ij}x_{ji} $ \\
-  transitive triplets  & $s_{i3}(x) = \sum_{j,h} x_{ij}x_{jh}x_{ih}$ \\[3pt]
+ \ outdegree  & $s_{i1}(x) = \sum_j x_{ij}$ \\
+\   reciprocity  & $s_{i2}(x) = \sum_j x_{ij}x_{ji} $ \\
+ \  transitive triplets  & $s_{i3}(x) = \sum_{j,h} x_{ij}x_{jh}x_{ih}$ \\[3pt]
   \multicolumn{2}{l}{\bf Covariate Effects}\\
-  covariate-alter  & $s_{i4}(x) = \sum_j x_{ij}v_j $ \\
-  covariate-ego  & $ s_{i5}(x) = v_i\sum_j x_{ij}$ \\
-  same covariate & $ s_{i6}(x) = \sum_j x_{ij} \mathbb{I}(v_i = v_j)$ \\
-  jumping $V$ transitive triplets  & $ s_{i7}(x) = \sum_{j \neq h} x_{ij}x_{ih}x_{hj} \mathbb{I}(v_i = v_h \neq v_j)$
+ \ covariate-alter  & $s_{i4}(x) = \sum_j x_{ij}v_j $ \\
+\  covariate-ego  & $ s_{i5}(x) = v_i\sum_j x_{ij}$ \\
+\  same covariate & $ s_{i6}(x) = \sum_j x_{ij} \mathbb{I}(v_i = v_j)$ \\
+ \ jumping $V$ transitive triplets  & $ s_{i7}(x) = \sum_{j \neq h} x_{ij}x_{ih}x_{hj} \mathbb{I}(v_i = v_h \neq v_j)$
   \end{tabular} 

### 2016-03-15 paper/network-inference.Rnw
couple of questions

@@
 \hh{XXX describe what happens to each individual actor first, then talk about the aggregate. Refer to the table to make the $s_{ik}$ more concrete.}
-\st{At random moments occurring with rate $\rho_{m-1}$, $m = 2, \dots, M$ where $M \geq 2$ is the number of observed waves of a network, a single actor $i$ is permitted to change one of its ties, $x_{ij}$.
-Alternatively, the actor can do nothing.
-These single edge change moments are unobserved.
-They are only observed in aggregate when the network is observed at discrete time points.
-So, multiple actors may change their ties between wave $m-1$ and wave $m$.
-\par When given the opportunity to change or not, each actor tries to optimize its own objective function $f_i(\beta,x)$ under the current configuration of the network (denoted by $x$) in the observed time point.
-The probability that actor $i$  changes a tie $x_{ij}$ \hh{to its opposite $x'_{ij}$ }  when given the opportunity to act is~\cite{Snijders:1996} \hh{could you first describe this function in words and then give the formula?
+\st{At random moments occurring with rate $\rho_{m-1}$, $m = 2, \dots, M$ where $M \geq 2$ is the number of observed waves of a network, a single actor $i$ is permitted to act:  
+when given the opportunity to act, an actor tries to optimize its own objective function $f_i(\beta,x)$ under the current configuration of the network (denoted by $x$) in the observed time point.
+The probability that actor $i$  changes a tie $x_{ij}$ \hh{to its opposite $x'_{ij}$} is~\cite{Snijders:1996} \hh{XXX could you first describe this function in words and then give the formula?
 It looks like a comparison of odds similar to a conditional logit model or a discrete choice model~\cite{benatkiva:1985}}
@@
 $$ 
-\hh{Sam, could you check in the Snijder paper again about the actual form of the denominator? could it be two sums rather than the one?} }
+}
+\hh{XXX Sam, could you check in the Snijder paper again about the actual form of the denominator? could it be two sums rather than the one?}
+
+\st{Alternatively, the actor can do nothing.
+\hh{XXX how is that decided?
+somehow the probabilities have to be evaluated and a decision in favor of acting or not is made.
+from the description it's not clear yet how that happens.} These single edge change moments are unobserved.
+They are only observed in aggregate when the network is observed at discrete time points.
+So, multiple actors may change their ties between wave $m-1$ and wave $m$.
+\par
 

### 2016-03-15 paper/network-inference.Rnw
couple more comments

@@
 
+\hh{XXX define: node, tie}
+
 \subsection{Stochastic Actor-Oriented Models}
@@
 when given the opportunity to act, an actor tries to optimize its own objective function $f_i(\beta,x)$ under the current configuration of the network (denoted by $x$) in the observed time point.
-The probability that actor $i$  changes a tie $x_{ij}$ \hh{to its opposite $x'_{ij}$} is~\cite{Snijders:1996} \hh{XXX could you first describe this function in words and then give the formula?
-It looks like a comparison of odds similar to a conditional logit model or a discrete choice model~\cite{benatkiva:1985}}
+The probability that actor $i$  changes a tie $x_{ij}$ \hh{to its opposite $x'_{ij}$} is given in~\cite{Snijders:1996} as \hh{XXX could you first describe this formula in words and then state it?
+It looks like a comparison of odds similar to a conditional logit model or a discrete choice model~\cite{benatkiva:1985}.
+How many ties does the actor have, i.e.
+what's the domain of $j$?}
 $$ 
@@
 \hh{XXX how is that decided?
-somehow the probabilities have to be evaluated and a decision in favor of acting or not is made.
+somehow the probabilities $p_ij$, $j=1, ..., ?$ have to be evaluated and a decision in favor of acting or not is made.
 from the description it's not clear yet how that happens.} These single edge change moments are unobserved.
 They are only observed in aggregate when the network is observed at discrete time points.
-So, multiple actors may change their ties between wave $m-1$ and wave $m$.
+So, multiple actors may change (multiple of) their ties between wave $m-1$ and wave $m$.
 \par
@@
 %
-\par To estimate the parameter values, $\beta_k$, $k = 1, \dots, p$, the relevant statistics are
+\par To estimate the parameter values, $\beta_k$, $k = 1, \dots, p$, the relevant statistics are ... XXX
+
+}
 

### 2016-03-23 PaperThoughts.Rnw
start describing the model in detail

@@
 \usepackage{booktabs}
-
+\usepackage{tikz}
 %% load any required packages here
@@
 
-\LARGE{\bf Thoughts about Network Visual Inference Paper}
+{\Large \textbf Stochastic Actor-Oriented Models for Social Networks}
 
-\begin{enumerate}
-\item Need to complete lineups in 2-3 weeks
-\item Lineups: 
-  \begin{enumerate}
-  \item What sizes? 3, 6, 9, 12
-  \item What data? Football, olga's data, teenage friends data?
-  \item Other things to consider: coloring of nodes? 
-  \end{enumerate}
-\item Get simulated networds from models from RSiena
-\item Figure out another way to fit a model and then simulate from it? 
-\end{enumerate}
+\section{Terminology \& Notation}
 
-Possible section outline for paper:
-\begin{enumerate}
-\item Introduction - visual inference, networks
-\item Data and Model fitting & simulation
-\item Lineup methods
-\item Results
-\item Conclusions and Discussion
-\end{enumerate}
+\begin{itemize}
+\item \textbf{Network} - A collection of actors and ties between them.
+A whole network (all of the nodes and the ties between them) is denoted $x$.
+In SAOMs, the networks are observed in waves for time points $m = 1, \dots, M$.
+Sometimes, the networks observed are denoted as $x(t_m)$.
+\item \textbf{Node} or \textbf{Actor} - The entities that form ties in a network.
+The number of nodes in a network is denoted $n$.
+The set of nodes, $\{i : i = 1, \dots n\}$ is denoted $\mathcal{N}$.
+\item \textbf{Edge} or \textbf{Tie} - The connections between actors in a network.
+A tie from actor $i$ to actor $j$ is denoted by $x_{ij}$.
+In SAOMs, ties are almost always directed, so $x_{ij}$ does not necessarily equal $x_{ji}$.
+$x_{ij} = 1$ if there is a tie from actor $i$ to actor $j$ and $x_{ij} = 0 $ if there is not a tie from actor $i$ to actor $j$.
+Ties from an actor to itself, $x_{ii}$, are structurally zero, meaning that self-ties are not allowed.
+In a network of $n$ nodes, there are $n(n-1)$ possible ties.
+\item \textbf{Adjacency matrix} - The adjacency matrix of a network is a numerical representation of it.
+The adjaceny matrix of $x$ is denoted $\mathcal{A}(x)$, is of dimension $n \times n$ and has $ij^{th}$ entry equal to the value of $x_{ij}$ for all $i \neq j \in \mathcal{N}$.
+The diagonal entries of $\mathcal{A}(x)$, $ii$ for $i = 1, \dots, n$, are all structurally 0.
+\item $\mathbf{X(t)}$ - A continuous time Markov process \emph{in which} the observed panel of $M$ networks of size $n$ is embedded.
+For each $t = 1, 2, 3, \dots$, there is one tie change.
+In other words, the network $X(t+1)$ only differs from $X(t)$ by one tie, $x_{ij}$.
+Call this tie $x_{ij}(t)$ in network $X(t)$ and $x_{ij}(t+1)$ in network $X(t+1)$.
+Then, $x_{ij}(t+1) = 1-x_{ij}(t)$.
+Note that $t$ is different from $t_m$.
+Here is an illustrated example: \\
+
+\begin{centering}
+\textbf{\emph{$X(t)$, the continuous time Markov process}} \\
+\begin{tikzpicture}
+\draw (0,0) -- (12.5,0);
+\foreach \x in {0,1,2,3,4,5,6,7,8,9,10,11,12}
+\draw[shift={(\x,0)},color=black] (0pt,3pt) -- (0pt,0pt);
+\foreach \x in {0,1,2,3,4,5,6,7,8,9,10,11,12}
+\draw[shift={(\x,.05)},color=black] (0pt,0pt) -- (0pt,-3pt) node[above]
+{$\x$};
+\foreach \x in {0,1,2,3,4,5,6,7,8,9,10,11,12}
+\draw[shift={(\x,.5)},color=black] (0pt,0pt) -- (0pt,-3pt) node[above]
+{$X(\x)$};
+\draw (0,0) node[below]{$t_1$};
+\draw (4,0) node[below]{$t_2$};
+\draw (7,0) node[below]{$t_3$};
+\draw (10,0) node[below]{$t_4$};
+\draw (12,0) node[below]{$t_5$};
+\draw (0,-.5) node[below]{$x(t_1)$};
+\draw (4,-.5) node[below]{$x(t_2)$};
+\draw (7,-.5) node[below]{$x(t_3)$};
+\draw (10,-.5) node[below]{$x(t_4)$};
+\draw (12,-.5) node[below]{$x(t_5)$};
+\end{tikzpicture} \\
+
+\textbf{\emph{$x(t_m)$, the observed networks}} \\
+
+\end{centering}
+
+Thus, each network observation, $x(t_m)$ represents the concatenation of (usually) multiple consecutive steps in the Markov chain in SAOMs.
+The state space for the the Markov chain is $\mathcal{S}$, the large but finite set of all possible adjacency matrices for $n$ nodes with zeroes on the diagonal.
+Note that $|\mathcal{S}| = 2^{n(n-1)}$.\footnote{See http://oeis.org/A053763.} Also note that the time points $t$ are not strictly 1 unit of time apart.
+The space between them is dictated by the distribution of the waiting times in the Markov process.
+\item \textbf{Rate of Change} - One of the defining elements of SAOMs is the function for the rate of change of a network and of the actors in that network.
+For actor $i$, the rate function is commonly denoted $\lambda_i(\alpha, x)$, which dictates how quickly actor $i$ gets opportunities to change one of its ties, $x_{ij}$.
+In this function, $\alpha$ is a parameter and $x$ is the current network state.
+The rate function for the whole network is $\lambda(\alpha, x) = \sum_i \lambda_i(\alpha, x)$.
+For any time point, $t$, where $X(t) \equiv x$, the waiting time to the next change opportunity by \emph{any} actor has distribution \emph{Exponential}$(\lambda(\alpha, x))$.
+This is because the chain of network evolutions is Markovian.
+There are many possibilities for the rate function.
+The simplest is that it is constant between observations $x(t_m)$.
+The rate function can also depend on covariate values of the actors or structural network elements such as outdegree.
+For instance, $\lambda_i(\alpha, x) = \exp\{\sum_k \alpha_k r_{ik}(x)\}$, where $r_{ik}(x)$ is a function of the current network state and the node $i$ (one of the aforementioned possible dependencies), and $k = 1, \dots K$ could be any number of functions as chosen by the researcher from previous subject matter knowledge or otherwise.
+\item \textbf{Probability of Change (nodes)} - Given that a change, call it $X(t) \equiv x \rightarrow X(t+1)$ occurs, each node has its own probability of changing.
+Recall that from time $t$ to time $t+1$, only one tie is changing (call it $x_ij$), so that only one node, $i$, is given the power to change one of its $n-1$ available ties from $x_ij$ to $1-x_{ij}$.
+The probability that node $i$ will be the one to change a tie, given $x \rightarrow X(t+1)$ is $\pi_i(\alpha, x) \equiv \pi_i(\alpha, x|x\rightarrow X(t+1)) = \frac{\lambda_i(\alpha, x)}{\lambda(\alpha, x)}$.
+\item \textbf{Permitted Changes} - First note a slight abuse of notation.
+The current state of the network, which was previously denoted $X(t)$ and $x$, will now be denoted $x^0$.
+Then, let $x$ instead represent the new, as yet undetermined network $X(t+1)$.
+So, given that a change, $x^0 \rightarrow x$ will occur, and that node $i$ has been selected to change a tie, there is a set of permitted changes, $\mathcal{A}_i(x^0) = \{\mathcal{A}(x^0) \cup \mathcal{A}_i^r(x^0)\}$, where $\mathcal{A}(x^0)$ is the current network's adjacency matrix, and $ \mathcal{A}_i^r(x^0) = \{x | x_{ij} = 1 - x_{ij}^0 \text{for only one } j \in \mathcal{N},  j \neq i, \}$.
+So, the set of permitted changes includes the set of all adjacency matrices with all entries equivalent to the current adjacency matrix with the exception of one of the non-fixed, $n-1$ entries of the $i^{th}$ row and the current adjacency matrix, meaning that $|\mathcal{A}_i(x^0)| = n$.
+\item \textbf{Objective function} - So, once a change in node $i$ is given, \emph{how} does that node determine which of the $n$ permitted changes it will make?
+Its choice is determined in large part by its objective function, $f_i(\beta, x^0, x)$, where $\beta$ is a parameter.
+This function can be interpreted as the ``relative attractiveness" for node $i$ associated with making a change from the current network state, $x^0$ to a new network state, $x$, where $x \in \mathcal{A}_i(x^0)$.
+It is assumed that actors want to maximize this function when given an opportunity for change.
+For example, if changing from $x^0$ to $x$ will decrease the value of $f_i$ for all $x \in \mathcal{A}_i(x^0)$, the node will be very likely to \emph{not} change any ties.
+Many models don't include the current state, $x^0$, in the objective function, but instead only consider $x$, the future state, and just attempt to get closer to their optimum state regardless of where they currently are.
+The objective function is usually defined linearly as $f_i(\beta, x^0, x) = \sum_k \beta_k s_{ik}(x^0, x)$, where the functions $s_{ik}$ are determined from subject matter knowledge.
+These, like the $r_{ik}$ in the rate function, can be actor covariates or structural variables like outdegree or reciprocated ties.
+\item \textbf{Probability of Change (tie given node)} - Given the change of a tie from node $i$, the probability of the change $x^0 \rightarrow x$, where $x \in \mathcal{A}_i(x^0)$ is $p_i(\beta, x^0, x) \equiv p_i(\beta, x^0, x | x^0\rightarrow x, x \in \mathcal{A}_i(x^0)) = \frac{\exp \{f_i(\beta, x^0, x)\}}{\sum_{\tilde{x}} \exp \{f_i(\beta, x^0, \tilde{x})}$, where $\tilde{x} \in \mathcal{A}_i(x^0)$.
+For a specific tie, $x_{ij}$, with $j \neq i$, this expression can be re-written as $p_{ij}(\beta, x^0) = \frac{\exp \{f_i(\beta, x^0, x^j)\}}{f_i(\beta, x^0, x^0) + \sum_{j \neq i} \exp \{f_i(\beta, x^0, x^j)} $, where $x^j$ has ties $x^j_{ih} = x^0_{ih}$ for all $h \neq j$ and $x^j_{ih} = 1- x^0_{ih}$ for $h = j$.
+Using this new expression, we can write the probability of no change for node $i$ ($p_i(\beta, x^0, x^0)$ in the original notation) as $p_{i0} \equiv p_i(\beta, x^0, x^0) = 1 - \sum_{j \neq i} p_{ij}(\beta, x^0)$.
+\item \textbf{Intensity Matrix} - The intensity matrix, $Q$, of a continuous time Markov process describes the rate of change between states of the chain.
+For networks, there are a very large number of possible states: $2^{n(n-1)}$.
+But, due to the properties of continuous-time Markov processes, there are only $n$ possible states given the current state, $n-1$ of which are uniquely determined by the node $i$ that is given the opportunity to change.
+Thus, the intensity matrix $Q$ is a very sparse $2^{n(n-1)} \times 2^{n(n-1)}$ matrix, with only $n(n-1) + 1$ non-zero entries in each row.
+Note that $n(n-1)$ of these represent the possible states that are one edge different from a given state, and the additional non-zero entry is for the state to remain the same.
+All other entries in a row are zero because those column states cannot be reached from the row state by just one change.
+The entries of $Q$ are defined as follows: let $y \neq z \in \{1, 2, \dots, 2^{n(n-1)} \}$ be indices of two different possible states of the network, $x^y, x^z \in \mathcal{S}$.
+For a timepoint $t$ define $x^y = X(t)$ and let $x^z \in \mathcal{A}_i(x^y)$ for each $i \in \mathcal{N}$.
+Then the $yz^{th}$ entry of $Q$ is:
+\[ q(x^y, x^z) = \begin{cases} 
+      \lambda_i(\alpha, x^y)p_i(\beta, x^y, x^z) & \text{if } x^z \in \mathcal{A}_i(x^y), \forall i \in \mathcal{N} \\
+      0 & \text{if } x^z \notin \mathcal{A}_i(x^y) \text{ for any i } \in \mathcal{N}
+   \end{cases}
+\]
+Properties of intensity matrices dictate that the diagonal entry, where the state doesn't change, is defined as $q(x^y,x^y) = -\sum_{z\neq y} q(x^y, x^z)$ so that the row sum is 0.
+Thus, the rate of change between any two states that differ by only one tie from $i \to j$ is the product of the rate at which actor $i$ gets to change a tie and the probability that the tie that will change is the tie to node $j$.\footnote{Just to be clear, the change is from $x^y_{ij}$ to $x^z_{ij} = 1 - x^y_{ij}$.}
+\end{itemize}
 

### 2016-03-23 PaperThoughts.Rnw
my brain hurts.

@@
 Properties of intensity matrices dictate that the diagonal entry, where the state doesn't change, is defined as $q(x^y,x^y) = -\sum_{z\neq y} q(x^y, x^z)$ so that the row sum is 0.
-Thus, the rate of change between any two states that differ by only one tie from $i \to j$ is the product of the rate at which actor $i$ gets to change a tie and the probability that the tie that will change is the tie to node $j$.\footnote{Just to be clear, the change is from $x^y_{ij}$ to $x^z_{ij} = 1 - x^y_{ij}$.}
+Thus, the rate of change between any two states that differ by only one tie from $i \to j$ is the product of the rate at which actor $i$ gets to change a tie and the probability that the tie that will change is the tie to node $j$.\footnote{Just to be clear, the change is from $x^y_{ij}$ to $x^z_{ij} = 1 - x^y_{ij}$.} Furthermore, the theory of continuous time Markov chains gives that the matrix of transition probabilities between observation times $t_{m-1}$ and $t_{m}$ is dependent only on the difference between timepoints, $t_m - t_{m-1}$ and is equal to $e^{(t_m - t_{m-1})Q}$, where $Q$ is the matrix defined above and $e^X$ for a real or complex square matrix $X$ is equal to $\sum_{k=0}^{\infty} \frac{1}{k!} X^k$.
+There is a package in \texttt{R} to do this, but I can't imagine doing this for any $n > 5$.
 \end{itemize}

### 2016-03-23 PaperThoughts.Rnw
more maths

@@
 So, the set of permitted changes includes the set of all adjacency matrices with all entries equivalent to the current adjacency matrix with the exception of one of the non-fixed, $n-1$ entries of the $i^{th}$ row and the current adjacency matrix, meaning that $|\mathcal{A}_i(x^0)| = n$.
+Additionally, define the complete set of possible changes from $x^0 \rightarrow x$ as $\mathcal{A}_{1:n}(x^0) = \{\mathcal{A}(x^0) \cup \mathcal{A}_1(x^0) \cup \mathcal{A}_2(x^0) \cup \cdots \cup \mathcal{A}_n(x^0)\}$.
+This is the set of all networks that differ from $x^0$ by only one tie and has size $|\mathcal{A}_{1:n}(x^0)| = n(n-1) + 1$.
 \item \textbf{Objective function} - So, once a change in node $i$ is given, \emph{how} does that node determine which of the $n$ permitted changes it will make?
@@
 Thus, the rate of change between any two states that differ by only one tie from $i \to j$ is the product of the rate at which actor $i$ gets to change a tie and the probability that the tie that will change is the tie to node $j$.\footnote{Just to be clear, the change is from $x^y_{ij}$ to $x^z_{ij} = 1 - x^y_{ij}$.} Furthermore, the theory of continuous time Markov chains gives that the matrix of transition probabilities between observation times $t_{m-1}$ and $t_{m}$ is dependent only on the difference between timepoints, $t_m - t_{m-1}$ and is equal to $e^{(t_m - t_{m-1})Q}$, where $Q$ is the matrix defined above and $e^X$ for a real or complex square matrix $X$ is equal to $\sum_{k=0}^{\infty} \frac{1}{k!} X^k$.
-There is a package in \texttt{R} to do this, but I can't imagine doing this for any $n > 5$.
+There is a package in \texttt{R} to do this matrix exponential operation, but I can't imagine doing this for networks for any $n > 5$.
+\item \textbf{Moving from $x(t_{m-1})$ to $x(t_m)$} - First define $x(t_1)$ as the first network observed and $x(t_M)$ as the last network observed.
+Networks in between are written $x(t_{m})$ for $m = 2, \dots, M-1$.
+The observations $x(t_1), x(t_2), \dots, x(t_M)$ are embedded in the continuous-time Markov chain $\{X(t)|t_1 \leq t \leq t_M\}$.
+Moving from the observation $x(t_{m-1})$ to the observation $x(t_m)$ requires $C$ changes or steps in the Markov process $X(t)$ where $C = \sum_{i,j \in \mathcal{N}} |x_{ij}(t_m) - x_{ij}(t_{m-1})|$.
+Denote the $C-1$ timepoints in the the Markov process that correspond to the unobserved changes in the network as $T_1, T_2, \dots, T_{C-1}$ and define $T_0 = t_{m-1}$ and $T_C = t_{m}$.
+For each timepoint $T_c$, for $c = 1, \dots C-1$, there is a single actor, denoted $I_c$ that gets an opportunity to change at this timepoint.
+This actor changes one of its ties to one node, $J_c$.
+If there is no change, $I_c = J_c$.
+Formally, the pair $(I_c,J_c)$ is the only $(i,j)$ pair for which $x_{ij}(T_{c-1}) \neq x_{ij}(T_c)$ if such a pair exists, and $(I_c, J_c) = (I_c, I_c)$ otherwise.
+The triplet $(T_c, I_c, J_c)$ forms a stochastic process for $c = 1, \dots, C-1$ that, given the prior observation, $x(t_{m-1})$ for $m = 2, \dots, M$, completely determines $x(t)$ for $t_{m-1} < t < t_m$.
+\item \textbf{Augmenting Data} (see section on ML estimation) - The augmenting data consist of $C$ and the pairs $(I_c, J_c)$ for $c = 1, \dots, C-1$.
+(Note the timepoints $T_c$ are left out of the augmenting data.) The possible outcomes of the augmenting data are determined by the condition that if for a given tie, $x_{ij}(t_{m-1})$, the number of timepoints $1 \leq c \leq C-1$ at which $(i_c, j_c) = (i,j)$ is even if $x_{ij}(t_{m}) = x_{ij}(t_{m-1})$ and is odd if $x_{ij}(t_{m}) \neq x_{ij}(t_{m-1})$.
+\item \textbf{Sample Path} - The sample path is the stochastic process $V = ((I_c, J_c):c = 1, \dots, C-1)$.
+The elements of $V$ where $I_c = J_c$ are redundant.
+However, they are kept in the process because they help with computation of the likelihood.
+For notational convenience, define $x^{(c)} = x(T_C)$.
+Then, the networks $x^{(c)}$ and $x^{(c-1)}$ only differ in the element $(I_c, J_c) \in V$.
 \end{itemize}
 
+\section{Important Facts about SAOMs}
+\begin{itemize}
+\item In SAOMs, the observed networks in a longitudinal study, call them $x_{t_1}, \dots, x_{t_M}$, are discrete observations embedded in a continuous time process, $X(t)$.
+This means that the probability model is defined \emph{independently} of the observational design.
+Thus, the time between observations can be irregular and still be analyzed under this model.
+\item The continuous-time Markov process $\{X(t) | t_1 \leq t \leq t_M\}$ has a distribution that consitutes an exponential family, which are of the form $f(y|\theta) = \exp\{\sum_{j=1}^s q_j(\theta) T_j(y) \} c(\theta) h(y)$ where $y$ is the data vector and $\theta$ is the scalar or vector parameter.
+This means the SAOMs can be viewed as ``incompletely observed exponential family" models.
+\item Inference is computer-intensive and time-consuming because of ``elaborate MCMC procedures" that must be implemented.
+\item The definition of SAOMs means that data can be simulated directly from the probability distribution given the initial network state $x_{t_1}$. 
+\item However, SAOMs are ``too complicated for the calculation of likelihoods or estimators in closed form." The distribution, however, can be easily simulated from, so that estimation can proceed by MCMC simulation.
+\end{itemize}
+
+\section{Maximum Likelihood Estimation}
+Sketch of Algorithm:
+\begin{enumerate}
+\item Use the first network observation, $x(t_1)$, as a starting value. 
+\item For $m = 2, 3, \dots, M$ where $M$ is the total number of observed networks, augment the observed data with random draws from the sequence of intermediate steps in the Markov process, $x(t)$, that could have led from $x(t_{m-1})$ to $x(t_m)$.
+i.e.
+draw from the set {\small $$\{(x(t^*_1), x(t^*_2), \dots x(t^*_{C-1}))' | x(t^*_1) \in \mathcal{A}_{1:n}(x(t_{m-1})) \& x(t^*_2) \in \mathcal{A}_{1:n}(x(t^*_1)) \& x(t^*_3) \in \mathcal{A}_{1:n}(x(t^*_2)) \& \cdots \& x(t_{m}) \in \mathcal{A}_{1:n}(x(t^*_{C-1})) \}$$ } where $C = \sum_{i,j \in \mathcal{N}} |x_{ij}(t_m) - x_{ij}(t_{m-1})|$.
+Note that $C$ in the number of single edge changes needed to get from $x(t_{m-1})$ to $x(t_m)$, so $C-1$ is the number of intermediate networks.
+These draws can be simulated using a Metropolis-Hastings algorithm.
+\item Use the draws from 2 as updates in a Robbins-Monro algorithm (?) to find the solution of the likelihood equation. 
+\end{enumerate}
+
 \end{document}

### 2016-03-29 PaperThoughts.Rnw
managed to wade my way through the entire process of ML estimation for SAOMs. I think I understand it fully now.

@@
 \usepackage[T1]{fontenc}
+\usepackage{url}
 \usepackage{amsmath,amssymb,array}
@@
 \item \textbf{Probability of Change (nodes)} - Given that a change, call it $X(t) \equiv x \rightarrow X(t+1)$ occurs, each node has its own probability of changing.
-Recall that from time $t$ to time $t+1$, only one tie is changing (call it $x_ij$), so that only one node, $i$, is given the power to change one of its $n-1$ available ties from $x_ij$ to $1-x_{ij}$.
+Recall that from time $t$ to time $t+1$, only one tie is changing (call it $x_{ij}$), so that only one node, $i$, is given the power to change one of its $n-1$ available ties from $x_ij$ to $1-x_{ij}$.
 The probability that node $i$ will be the one to change a tie, given $x \rightarrow X(t+1)$ is $\pi_i(\alpha, x) \equiv \pi_i(\alpha, x|x\rightarrow X(t+1)) = \frac{\lambda_i(\alpha, x)}{\lambda(\alpha, x)}$.
@@
 The observations $x(t_1), x(t_2), \dots, x(t_M)$ are embedded in the continuous-time Markov chain $\{X(t)|t_1 \leq t \leq t_M\}$.
-Moving from the observation $x(t_{m-1})$ to the observation $x(t_m)$ requires $C$ changes or steps in the Markov process $X(t)$ where $C = \sum_{i,j \in \mathcal{N}} |x_{ij}(t_m) - x_{ij}(t_{m-1})|$.
-Denote the $C-1$ timepoints in the the Markov process that correspond to the unobserved changes in the network as $T_1, T_2, \dots, T_{C-1}$ and define $T_0 = t_{m-1}$ and $T_C = t_{m}$.
-For each timepoint $T_c$, for $c = 1, \dots C-1$, there is a single actor, denoted $I_c$ that gets an opportunity to change at this timepoint.
-This actor changes one of its ties to one node, $J_c$.
-If there is no change, $I_c = J_c$.
-Formally, the pair $(I_c,J_c)$ is the only $(i,j)$ pair for which $x_{ij}(T_{c-1}) \neq x_{ij}(T_c)$ if such a pair exists, and $(I_c, J_c) = (I_c, I_c)$ otherwise.
-The triplet $(T_c, I_c, J_c)$ forms a stochastic process for $c = 1, \dots, C-1$ that, given the prior observation, $x(t_{m-1})$ for $m = 2, \dots, M$, completely determines $x(t)$ for $t_{m-1} < t < t_m$.
-\item \textbf{Augmenting Data} (see section on ML estimation) - The augmenting data consist of $C$ and the pairs $(I_c, J_c)$ for $c = 1, \dots, C-1$.
-(Note the timepoints $T_c$ are left out of the augmenting data.) The possible outcomes of the augmenting data are determined by the condition that if for a given tie, $x_{ij}(t_{m-1})$, the number of timepoints $1 \leq c \leq C-1$ at which $(i_c, j_c) = (i,j)$ is even if $x_{ij}(t_{m}) = x_{ij}(t_{m-1})$ and is odd if $x_{ij}(t_{m}) \neq x_{ij}(t_{m-1})$.
-\item \textbf{Sample Path} - The sample path is the stochastic process $V = ((I_c, J_c):c = 1, \dots, C-1)$.
-The elements of $V$ where $I_c = J_c$ are redundant.
+Moving from the observation $x(t_{m-1})$ to the observation $x(t_m)$ requires $R$ changes or steps in the Markov process $X(t)$ where $R = \sum_{i,j \in \mathcal{N}} |x_{ij}(t_m) - x_{ij}(t_{m-1})|$.
+Note that the observed network change from $x(t_{m-1})$ to $x_(t_m)$ is conditionally independent of all other network observations given $x(t_{m-1})$.
+Denote the $R$ timepoints in the the Markov process that correspond to the unobserved changes in the network as $T_1, T_2, \dots, T_{R-1}$ and define $T_0 = t_{m-1}$ and $T_R \leq t_{m} < T_{R+1}$.
+For each timepoint $T_r$, for $r = 1, \dots R$, there is a single actor, denoted $I_r$ that gets an opportunity to change at this timepoint.
+This actor changes one of its ties to one node, $J_r$.
+If there is no change, $I_r = J_r$.
+Formally, the pair $(I_r,J_r)$ is the only $(i,j)$ pair for which $x_{ij}(T_{r-1}) \neq x_{ij}(T_r)$ if such a pair exists, and $(I_r, J_r) = (I_r, I_r)$ otherwise.
+The triplet $(T_r, I_r, J_r)$ forms a stochastic process for $r = 1, \dots, R$ that, given the prior observation, $x(t_{m-1})$ for $m = 2, \dots, M$, completely determines $x(t)$ for $t_{m-1} < t < t_m$.
+\item \textbf{Augmenting Data} (see section on ML estimation) - The augmenting data consist of $R$ and the pairs $(I_r, J_r)$ for $r= 1, \dots, R$.
+(Note the timepoints $T_r$ are left out of the augmenting data.) The possible outcomes of the augmenting data are determined by the condition that if for a given tie, $x_{ij}(t_{m-1})$, the number of timepoints $1 \leq r\leq R$ at which $(i_r, j_r) = (i,j)$ is even if $x_{ij}(t_{m}) = x_{ij}(t_{m-1})$ and is odd if $x_{ij}(t_{m}) \neq x_{ij}(t_{m-1})$.
+\item \textbf{Sample Path} - The sample path is the stochastic process $V = ((I_r, J_r):r = 1, \dots, R)$.
+Note that the elements of $V$ where $I_r = J_r$ are redundant.
 However, they are kept in the process because they help with computation of the likelihood.
-For notational convenience, define $x^{(c)} = x(T_C)$.
-Then, the networks $x^{(c)}$ and $x^{(c-1)}$ only differ in the element $(I_c, J_c) \in V$.
+For notational convenience, define $x^{(r)} = x(t_r)$.
+Using this notation, write $x(t_{m-1}) = x^(0)$ for the independent cases where $m = 2, \dots, M$.
+Then, the networks $x^{(r)}$ and $x^{(r-1)}$ only differ in the element $(I_r, J_r) \in V$, provided that $I_r\neq J_r$.
+\item \textbf{PDF of Sample Path} - The sample path has probability distribution function 
+  $$ p_{sp}(V|\alpha, \beta, x^{(0)} ) = Pr(T_R \leq t_m < T_{R+1} | x^{(0)}, V, \alpha, \beta) \times \prod_{r=1}^R \pi_{i_r}(\alpha, x^{(r-1)})p_{i_r,j_r}(\beta,x^{(r-1)}). $$
+The first component, $Pr(T_R \leq t_m < T_{R+1} | x^{(0)}, V, \alpha, \beta)$ is the probability that your next network observation at time point $t_m$ comes \emph{before} the next change in the continuous time Markov chain, $X(T_{R+1})$.
+From a discussion that follows, $Pr(T_R \leq t_m < T_{R+1} | x^{(0)}, V, \alpha, \beta)$, the probability that your next network observation at time point $t_m$ comes \emph{before} the next change in the continuous time Markov chain, $X(T_{R+1})$, can be approximated by
+$$ \kappa(\alpha, x(t_{m-1}), V) \approx \int_{t_{m-1}}^{t_m} p_{T_R}(s) \text{Pr}(\{T_{R+1} - T_R > t_{m} - T_R \} | T_R = s) ds $$
+$T_{R+1}$ has distribution Exponential$(\lambda(\alpha, x^{(R)}))$.
+So, Pr$(\{T_{R+1} - T_R > t_m - T_R \}) = \text{Pr}(T_{R+1} > t_m) = 1 - \text{Pr}(T_{R+1} < t_m) = 1 - (1 - \exp\{-\lambda(\alpha, x^{(R)})t_m\}) = \exp\{-\lambda(\alpha, x^{(R)})t_m\}$.
+So, the above integral can be written as
+$$ \kappa(\alpha, x(t_{m-1}), V) \approx \int_{t_{m-1}}^{t_m} p_{T_R}(s) \exp\{-\lambda(\alpha, x^{(R)})(t_m-s)\}. ds $$
+This quantity is approximately equal to $\frac{p_{T_R}(t_m)}{\lambda(\alpha, x^{(R)})}$.
+The second expression, $\prod_{r=1}^R \pi_{i_r}(\alpha, x^{(r-1)})p_{i_r,j_r}(\beta,x^{(r-1)})$, is the PDF of the stochastic process $V$.
+So, the probability of the sample path can be approximated by:
+$$ p_{sp}(V|\alpha, \beta, x^{(0)}) \approx \frac{p_{T_R}(t_m)}{\lambda(\alpha, x^{(R)})} \times \prod_{r=1}^R \pi_{i_r}(\alpha, x^{(r-1)})p_{i_r,j_r}(\beta,x^{(r-1)}). $$
+So, for observed data, augmented by the sample path, ``the likelihood conditional on $x(t_{m-1})$ can be expressed directly, either exactly [(for constant $\lambda_i$)] or in good approximation."
+\item \textbf{Distribution of waiting times between $x(t_{m-1})$ and $x(t_m)$} - Let $T_{r+1} - T_r$ be the waiting time between one of the $r = 1, \dots, R$ changes between $x(t_{m-1})$ and $x(t_m)$.
+After conditioning on $x(t_{m-1}) \equiv x^{(0)}$ and $V$, these waiting times have distribution $T_{r+1} - T_{r} \sim \text{Exponential}(\lambda(\alpha, x^{(r)}))$ for $r = 0, 1, \dots, R-1$.
+Following from this, the distribution of $T_R - t_{m-1}$, the time it takes for the network to move from state $x(t_{m-1})$ to state $x(t_{m})$ is the convolutions of the Exponential$(\lambda(\alpha, x^{(r)}))$ distributions for $r = 0, \dots, R-1$.
+The pdf of $T_R - t_{m-1}$ is\footnote{Source: \url{http://www.ccms.or.kr/data/pdfpaper/jcms21_4/21_4_501.pdf}}:
+$$ h(T_R - t_1 | \alpha, x(t_{m-1}), V) = \sum_{r=0}^{R-1} \frac{\lambda(\alpha, x^{(0)})\cdot \lambda(\alpha, x^{(1)}) \cdot \lambda(\alpha, x^{(R-1)})}{\prod_{q\neq r, q = 0}^{R-1} (\lambda(\alpha, x^{(q)}) - \lambda(\alpha, x^{(r)}))} \cdot \exp\{-(T_R - t_{m-1})\cdot \lambda(\alpha, x^{(r)})\}$$
+In the special case that the $\lambda_{i}(\alpha, x)$ rates are constant, call it $\alpha_1$, then the number of steps in the Markov chain to move from $x(t_{m-1})$ to $x(t_m)$, $R$ has distribution $R \sim \text{Poisson}(n\alpha_1(t_{m} - t_{m-1}))$.
+In this case, $Pr(T_R \leq t_m < T_{R+1} | x^{(0)}, V, \alpha, \beta) = \kappa(\alpha_1, x^{(0)}, V) = \exp\{-n\alpha_1(t_{m} - t_{m-1})\}\frac{(n\alpha_1(t_m - t_{m-1}))^R}{R!}$.
+However, when the individual rates are different for each actor, this probability must be approximated.
+\item \textbf{Distribution of $T_R$} - $T_R$ is the timepoint in the Markov chain at which the state $x(t_m)$ is acheived.
+Since $T_R - t_{m-1}$ is a convolution of exponentials with rate parameters $\lambda(\alpha, x^{(r)})$ for $r = 0, R-1$, by the Lyapounov Central Limit Theorem, the distribution of $T_R$ is approximately normal with mean $\mu_{\alpha}$ and variance $\sigma^2_{\alpha}$ where
+$$ \mu_{\alpha} = \sum_{r=0}^{R-1} \frac{1}{\lambda(\alpha, x^{(r)})} \quad \text{and} \quad  \sigma^2_{\alpha} = \sum_{r=0}^{R-1} \frac{1}{\lambda(\alpha, x^{(r)})^2} $$
+The PDF of $T_R$ is thus $p_{T_R} \approx \frac{1}{\sqrt{2\pi \sigma^2_{\alpha}}} \exp\left\{\frac{-((t-t_{m-1})-\mu_{\alpha})^2}{2\sigma^2_{\alpha}}\right\}$. 
 \end{itemize}
@@
 \item However, SAOMs are ``too complicated for the calculation of likelihoods or estimators in closed form." The distribution, however, can be easily simulated from, so that estimation can proceed by MCMC simulation.
+\item ML estimation of SAOMs is based on the missing data principle.
+The ``missing data" here are the paths from $x(t_{m-1})$ to $x(t_{m})$ for all $m = 2, \dots, M$.
+This set can be denoted for each $m$ by $V_m = \{(I_{m1}, J_{m1}),\dots, (I_{mR_m}, J_{mR_m})\}$.
+So, the complete set of augmenting data is $V = (V_2, \dots V_M)$.
 \end{itemize}
@@
 \begin{enumerate}
-\item Use the first network observation, $x(t_1)$, as a starting value. 
+\item Use the first network observation, $x(t_1)$, as a starting value. All analysis proceeds conditioning on $x(t_1)$.  
 \item For $m = 2, 3, \dots, M$ where $M$ is the total number of observed networks, augment the observed data with random draws from the sequence of intermediate steps in the Markov process, $x(t)$, that could have led from $x(t_{m-1})$ to $x(t_m)$.
@@
 
+Full Iterative Algorithm: For each iteration $N = 1, 2, \dots$
+\begin{enumerate}
+\item For each $m = 2, \dots, M$, make a large number of Metropolis-Hastings steps of the following form.
+Let $\underline{v} = ((i_1, j_1), \dots, (i_R, i_R))$ be a given path from $x(t_{m-1})$ to $x(t_m)$ in the whole possible set of paths $\mathcal{V}$.
+Then propose a new path, $\tilde{\underline{v}}$ from the proposal distribution that consists of all of the possible small changes:
+  \begin{enumerate}
+  \item ``Paired Deletions" - Of all pairs of indices $r_1, r_2$ such that $(i_{r_1}, j_{r_1}) = (i_{r_2}, j_{r_2})$, $i_{r_1} \neq j_{r_1}$, randomly select a pair $(r_1, r_2)$ and delete both from the current path.
+  \item ``Paired Insertions" - Randomly select an edge $(i,j) \in \mathcal{N}^2$ with $i \neq j$. Randomly choose 2 indices, $r_1, r_2$. Insert $(i,j)$ immediately before $r_1$ and $r_2$.  
+  \item ``Single Insertions" - At a random place in the current path, insert $(i,i)$ for a random $i \in \mathcal{N}$.
+  \item ``Single Deletions" - Of all elements in the current path that satisfy $i_r = j_r$, randomly delete one of them. 
+  \item ``Permutations" - For randomly chosen $r_1 < r_2$, where $r_2 - r_1$ is bounded from above by some smallish number to avoid lengthy computation (tuning parameter?), the segment of consecutive elements $(i_r,j_r)$, $r = r_1, \dots, r_2$ is randomly permuted.
+  \end{enumerate}
+Denote the proposal probabilities $u(\tilde{\underline{v}} | \underline{v})$ and the target probabilities $p(\underline{v})$.
+The target distribution is $p(\underline{v}) \propto p_{SP}(\underline{v} | \alpha, \beta, x(t_{m-1}))$.
+Then the acceptance probability for a proposal path $\tilde{\underline{v}}$ and a current path $\underline{v}$ is: $\min \left\{ 1, \frac{p(\tilde{\underline{v}})u(\underline{v}|\tilde{\underline{v}})}{p(\underline{v})u(\tilde{\underline{v}}| \underline{v})}\right\}$.
+The series of M-H steps will result in a new set of augmenting data, $(v^{(N)}) = (v_2^{(N)}, \dots, v_M^{(N)})$.
+\item Compute 
+  $$ S_{XV}(\hat{\theta^{(N)}}; x, v^{(N)}) = \sum_{m=2}^M S_m(\hat{\theta^{(N)}}; x(t_{m-1}), v_m^{(N)}) $$
+$S_{XV}(\hat{\theta^{(N)}}; x, v^{(N)})$ is the complete data score function.
+$S_m(\hat{\theta^{(N)}}; x(t_{m-1}), v_m^{(N)}) = \frac{\partial \log p_m(v_m; \theta | x(t_{m-1}))}{d\theta}$ is the total data score function at step $m$.
+$p_m$ is $p_{SP}(v_m; \theta = (\alpha, \beta)' | x(t_{m-1}))$, the probability of the sample path $v_m$.
+\item Update $$\theta^{(N+1)} =(\alpha^{(N+1)}, \beta^{(N+1)}) = \theta^{(N)} + a_N D^{-1} S_{XV}(\hat{\theta^{(N)}}; x, v^{(N)})$$
+where $D$ is a MC estimate of the complete data observed Fisher information matrix, $D_{XV}(\theta) = -\frac{\partial S_{XV}(\theta: x, V)}{\partial \theta}$ estimated for $\theta = \theta^{(1)}$.
+(i.e.
+evaluated at the starting value of $\theta$.) The starting value is the method of moments estimator, $\hat{\theta}^{(1)}$.
+(See Snijders 2001 for how to get the MoM estimate).
+The sequence $a_N$ is just a sequence that approaches 0 as $N \to \infty$.
+(I'm not sure exactly what?
+Maybe just $\frac{1}{N}$ would work?
+There's no specifics in Snijders 2001, and I'm looking through the RSiena manual for clues...).
+\end{enumerate}
+
 \end{document}

### 2016-03-31 PaperThoughts.Rnw
fixed small friend lineup so that singletons appear.

@@
 Note that the observed network change from $x(t_{m-1})$ to $x_(t_m)$ is conditionally independent of all other network observations given $x(t_{m-1})$.
-Denote the $R$ timepoints in the the Markov process that correspond to the unobserved changes in the network as $T_1, T_2, \dots, T_{R-1}$ and define $T_0 = t_{m-1}$ and $T_R \leq t_{m} < T_{R+1}$.
+Denote the $R$ timepoints in the the Markov process that correspond to the unobserved changes in the network as $T_1, T_2, \dots, T_{R}$ and define $T_0 = t_{m-1}$ and $T_R \leq t_{m} < T_{R+1}$.
 For each timepoint $T_r$, for $r = 1, \dots R$, there is a single actor, denoted $I_r$ that gets an opportunity to change at this timepoint.

### 2016-03-31 paper/network-inference.Rnw
now the comments get comments
fixed small friend lineup so that singletons appear.

@@
 stochastic actor-oriented models (SAOMs) were introduced by Snijders in~\cite{Snijders:1996}.
+A SAOM is a discrete process embedded in a continuous-time Markov process of network evolution.
+In the Markov chain, each step consists of adding or deleting one tie $(i,j)$.
+So, the change from the first observation of a network, call it $x(t_1)$, to the next observation, $x(t_2)$ is modeled as a series of $R$ changes, where $R = \sum_{i,j} |x_{ij}(t_2) - x_{ij}(t_1)|$.
+These individual changes are unobserved, and during the estimation of parameters, they serve as augmenting data.
+Denote the $R$ timepoints in the the Markov process that correspond to the unobserved changes in the network as $T_1, T_2, \dots, T_{R}$ and define $T_0 = t_{1}$ and $T_R \leq t_{2} < T_{R+1}$.
+For each timepoint $T_r$, for $r = 1, \dots R$, there is a single actor, denoted $I_r$ that gets an opportunity to change at this timepoint.
+This actor changes one of its ties to one node, $J_r$.
+If there is no change, $I_r = J_r$.
+The waiting times, $T_r - T_{r-1}$ for $r = 1, \dots, R$ have independent exponential distributions with expected value $(\lambda(\alpha, x^{(r-1)}))^{-1}$.
+The function $\lambda(\alpha, x^{(r-1)})$ is the rate at which the network changes.
+Here, $x^{(r)} \equiv x(T_r)$ for $r = 2, \dots, R$.
+It is the sum of the rates at which each node makes a change in one of its ties, $\lambda_i(\alpha, x^{(r-1)})$.
+Note that x^{(r)}
 In a SAOM each actor is represented by a node.
@@
 \hh{XXX how is that decided?
-somehow the probabilities $p_ij$, $j=1, ..., ?$ have to be evaluated and a decision in favor of acting or not is made.
+somehow the probabilities $p_ij$, $j=1, ..., ?$ have to be aggregated and a decision in favor of acting or not must be made.
 from the description it's not clear yet how that happens.} These single edge change moments are unobserved.

### 2016-03-31 Code/small_friends_lineup.R

@@ -10,6 +10,7 @@ library(geomnet)
 # source("Code/find_sig_eff_small_friends.R")
 
 create_smfriend_lu <- function(null_eff_struct, test_eff_struct, M, my_dat=mysmalldata, ...){
+  #browser()
   null_data <- NULL
   n <- M-1
   for (i in 1:n){
@@ -17,6 +18,11 @@ create_smfriend_lu <- function(null_eff_struct, test_eff_struct, M, my_dat=mysma
     net_df <- merge(data.frame(null_net$sims[[1000]][[1]][[1]][[1]])[,-3], 
                      data.frame(id = 1:16), by.x = "X1", by.y = "id",
                      all = T)
+    for (i in 1:nrow(net_df)){
+      if (!(net_df$X1[i] %in% net_df$X2) & is.na(net_df$X2[i])){
+        net_df$X2[i] <- net_df$X1[i]
+      } else {net_df$X2[i] <- net_df$X2[i]}
+      }
     net_df$count <- i 
     null_data <- rbind(null_data, net_df)
   }
@@ -24,9 +30,16 @@ create_smfriend_lu <- function(null_eff_struct, test_eff_struct, M, my_dat=mysma
   test_data <- merge(data.frame(test_net$sims[[1000]][[1]][[1]][[1]])[,-3], 
                      data.frame(id = 1:16), by.x = "X1", by.y = "id",
                      all = T)
+  for (i in 1:nrow(test_data)){
+  if (!(test_data$X1[i] %in% test_data$X2) & is.na(test_data$X2[i])){
+    test_data$X2[i] <- test_data$X1[i]
+  } else {test_data$X2[i] <- test_data$X2[i]}
+}
   test_data$count <- M  # test data is always indexed by M
   to_plot <- rbind(null_data, test_data)
   to_plot$plot_order <- rep(sample(M), as.vector(table(to_plot$count)))
+  to_plot$X1 <- as.factor(to_plot$X1)
+  to_plot$X2 <- as.factor(to_plot$X2)
   plot <- ggplot(data = to_plot, aes(from_id = X1, to_id = X2)) + 
             geom_net(fiteach = TRUE, directed = F, size = 1, arrowsize = .5) +
             facet_wrap(~plot_order) + theme_net() + theme(panel.background = element_rect(fill = "white", color = 'white'))

### 2016-04-03 Code/complete_lineup_creation.R
create all lineups

@@ -1,0 +1,135 @@
+library(RSiena)
+library(geomnet)
+library(grid)
+library(gridSVG)
+##### LOAD IN DATA #####
+#looks like 20:35 are pretty well connected, go with those for now
+setwd("Desktop/NetworksResearch/NetworksVizInference/")
+friend.data.w1 <- as.matrix(read.table("Data/s50_data/s50-network1.dat"))
+friend.data.w2 <- as.matrix(read.table("Data/s50_data/s50-network2.dat"))
+friend.data.w3 <- as.matrix(read.table("Data/s50_data/s50-network3.dat"))
+fd2.w1 <- friend.data.w1[20:35,20:35]
+fd2.w2 <- friend.data.w2[20:35,20:35]
+fd2.w3 <- friend.data.w3[20:35,20:35]
+# read in covariate data
+drink <- as.matrix(read.table("Data/s50_data/s50-alcohol.dat"))
+drink2 <- drink[20:35,]
+# siena data
+friendData2 <- array( c( fd2.w1, fd2.w2, fd2.w3 ),
+                      dim = c( 16, 16, 3 ) )
+# and next give this the role of the dependent variable:
+friend2 <- sienaDependent(friendData2)
+alcohol2 <- varCovar( as.matrix(drink2) )
+# create siena data object
+mysmalldata <- sienaDataCreate( friend2, alcohol2)
+null_model_eff2 <- getEffects(mysmalldata)
+myalgorithm2 <- sienaAlgorithmCreate( projname = 's50' , n3 = 1000)
+sig_eff_names <- read.csv("Data/sig_eff_names_smfriends.csv")
+runs_models_smallFriends <- readRDS("Data/runs_models_smallFriends.RDS")
+eff_models_smallFriends <- readRDS("Data/eff_models_smallFriends.RDS")
+# functions to create lineups 
+create_smfriend_lu <- function(null_eff_struct, test_eff_struct, M, my_dat=mysmalldata, ...){
+ # browser()
+  null_data <- NULL
+  n <- M-1
+  null_net <- siena07( myalgorithm2, data = my_dat, returnDeps = TRUE, effects = null_eff_struct, batch=TRUE, verbose = FALSE)
+  null_eff_struct2 <- null_eff_struct
+  null_eff_struct2$initialValue[null_eff_struct2$include] <- c(null_net$rate,null_net$theta)
+  sim_algorithm <- sienaAlgorithmCreate( projname = "sim_model", cond = FALSE,
[…]

### 2016-04-03 Code/find_sig_eff_small_friends.R

@@ -16,12 +16,13 @@ alcohol2 <- varCovar( as.matrix(drink2) )
 # create siena data object
 mysmalldata <- sienaDataCreate( friend2, alcohol2)
 null_model_eff2 <- getEffects(mysmalldata)
-includeEffects( null_model_eff2, sameXTransTrip, interaction1 = "alcohol2", character = TRUE)
+includeEffects( null_model_eff2, "sameXTransTrip", interaction1 = "alcohol2", character = TRUE)
 effectsDocumentation(null_model_eff2)
+library(dplyr)
 library(rvest)
 RSeffects <- read_html("null_model_eff2.html") %>% html_nodes("table") %>% html_table()
 RSeffects <- RSeffects[[1]]
-RSeffects$shortName
+
 
 effStructures <- NULL
 for (i in 1:nrow(RSeffects)){
@@ -60,17 +61,19 @@ for (i in 1:n){
 test_results[,4:6] = apply(test_results[,4:6], 2, function(x) as.numeric(as.character(x)))
 write.csv(test_results, file = "effects_significance_smallFriends.csv") 
 
-test_results %>% filter(Waldpval<.05)
+test_results %>% filter(Waldpval<.05) -> sig_eff_names
+write.csv(sig_eff_names, "Data/sig_eff_names_smfriends.csv")
 
-library(ggplot2)
-ggplot(test_results %>% filter(Waldpval <.05)) +
-  geom_point(aes(x = estimate, y = -Waldpval, color = inter1)) +
-  geom_hline(yintercept = -0.05) + 
-  facet_wrap(~type)
+# library(ggplot2)
+# ggplot(test_results %>% filter(Waldpval <.05)) +
+#   geom_point(aes(x = estimate, y = -Waldpval, color = inter1)) +
+#   geom_hline(yintercept = -0.05) + 
+#   facet_wrap(~type)
 
 null_model_smallFriends <- getEffects(mysmalldata)
 
-as.vector(test_results %>% filter(Waldpval < .05) %>% select(shortName, type, inter1, Waldpval)) -> sig_eff_names
+test_results %>% filter(Waldpval < .05 & Waldpval > 0) -> sig_eff_names
+write.csv(sig_eff_names, "Data/sig_eff_names.csv")
 
 eff_models_smallFriends <- NULL
 for (i in 1:nrow(sig_eff_names)){
@@ -88,28 +91,28 @@ for (i in 1:nrow(sig_eff_names)){
[…]

### 2016-04-03 Code/small_friends_lineup.R

@@ -113,9 +113,9 @@ write.csv(lu1$data, "Data/lineupdata/smallfriends-m-9-rep-1.csv", row.names=FALS
 file <- 1
 #dir.create(path = "lineups")
 #dir.create(path = "lineups/pdfs")
-ggsave(lu1$lineup, file=sprintf("lineups/pdfs/smallfriends-m-9-rep-%s.pdf",file), width = 7.2, height = 4.5, units = "in")
+ggsave(lu1$lineup, file=sprintf("Lineup-Images/pdfs/smallfriends-m-9-rep-%s.pdf",file), width = 7.2, height = 4.5, units = "in")
 #dir.create(path = "lineups/svgs")
-tmpfile <- sprintf("%s.svg",tempfile(tmpdir="lineups/svgs"))
+tmpfile <- sprintf("%s.svg",tempfile(tmpdir="Lineup-Images/svgs"))
 print(lu1$lineup)
 # isn't working. keeps throwing the same 2 errors
 # to do: fix arrows! 

### 2016-04-03 lineups.Rnw
all lineups in one document. 20 per M value

\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,mathrsfs,fancyhdr,syntonly,lastpage,hyperref,enumitem,graphicx, lscape, forloop}

\hypersetup{colorlinks=true,urlcolor=black}

\topmargin      -1.5cm   % read Lamport p.163
\oddsidemargin  -0.04cm  % read Lamport p.163
\evensidemargin -0.04cm  % same as oddsidemargin but for left-hand pages
\textwidth      16.59cm
\textheight     23.94cm
\parskip         7.2pt   % sets spacing between paragraphs
\parindent         0pt   % sets leading space for paragraphs
\pagestyle{empty}        % Uncomment if don't want page numbers
\pagestyle{fancyplain}

\begin{document}
\lhead{Graph Lineups}
\chead{}
\rhead{Samantha Tyner}

\newcounter{k}
\forloop{k}{1}{\value{k} < 21}{\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-m-3-rep-\arabic{k}} \\ }

\newcounter{k1}
\forloop{k1}{1}{\value{k1} < 21}{\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-m-6-rep-\arabic{k1}} \\ }

\newcounter{k2}
\forloop{k2}{1}{\value{k2} < 21}{\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-m-9-rep-\arabic{k2}} \\ }

\newcounter{k3}
\forloop{k3}{1}{\value{k3} < 21}{\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-m-12-rep-\arabic{k3}} \\ }

\end{document}

### 2016-04-06 lineups.Rnw
couple of comments on lineups

@@
 
+\newcommand{\hh}[1]{{\color{magenta} #1}}
+
 \begin{document}
@@
 \rhead{Samantha Tyner}
+\hh{Cool, you got all the singletons to display! Nice work!}
+\hh{Could you add a bit more information on the lineups?
+Maybe we could have a list of models/hypothesis at the beginning and then sections that summarize things a bit more.
+Right now I am not quite sure about how to refer to each individual lineup.
+- OK, fixed that for the first batch by including the names of the files.}
+
+\hh{Could you crop the white space from the lineup before saving it? - maybe this blog post helps: \url{http://robjhyndman.com/hyndsight/crop-r-figures/}}
 
 \newcounter{k}
-\forloop{k}{1}{\value{k} < 21}{\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-m-3-rep-\arabic{k}} \\ }
+\forloop{k}{1}{\value{k} < 21}{smallfriends-m-3-rep-\arabic{k}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-m-3-rep-\arabic{k}} \\ }
 

### 2016-04-06 lineups.Rnw
additional structure, so we know which figure we talk about

@@
 Right now I am not quite sure about how to refer to each individual lineup.
-- OK, fixed that for the first batch by including the names of the files.}
+- OK, fixed that by including the names of the files.}
 
-\hh{Could you crop the white space from the lineup before saving it? - maybe this blog post helps: \url{http://robjhyndman.com/hyndsight/crop-r-figures/}}
+\hh{Could you crop the white space from the lineup before saving it? - maybe this blog post helps: {http://robjhyndman.com/hyndsight/crop-r-figures/}}
 
+\hh{Please add frames around the panels - it helps with separating the graphs from one another}
+
+\tableofcontents
+
+\section{Model X versus Y lineup size 3}
 \newcounter{k}
@@
 
+\section{Model X versus Y lineup size 6}
 \newcounter{k1}
-\forloop{k1}{1}{\value{k1} < 21}{\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-m-6-rep-\arabic{k1}} \\ }
+\forloop{k1}{1}{\value{k1} < 21}{
+Lineup-Images/pdfs/smallfriends-m-6-rep-\arabic{k1}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-m-6-rep-\arabic{k1}} \\ }
 
+\section{Model X versus Y lineup size 9}
 \newcounter{k2}
-\forloop{k2}{1}{\value{k2} < 21}{\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-m-9-rep-\arabic{k2}} \\ }
+\forloop{k2}{1}{\value{k2} < 21}{Lineup-Images/pdfs/smallfriends-m-9-rep-\arabic{k2}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-m-9-rep-\arabic{k2}} \\ }
 
+\section{Model X versus Y lineup size 12}
 \newcounter{k3}
-\forloop{k3}{1}{\value{k3} < 21}{\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-m-12-rep-\arabic{k3}} \\ }
+\forloop{k3}{1}{\value{k3} < 21}{Lineup-Images/pdfs/smallfriends-m-12-rep-\arabic{k3}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-m-12-rep-\arabic{k3}} \\ }
 

### 2016-04-06 lineups.Rnw
add new code and rnw files

@@
 \newcommand{\hh}[1]{{\color{magenta} #1}}
+\newcommand{\st}[1]{{\color{purple} #1}}
 
@@
 
-\hh{Could you crop the white space from the lineup before saving it? - maybe this blog post helps: {http://robjhyndman.com/hyndsight/crop-r-figures/}}
+\hh{Could you crop the white space from the lineup before saving it?
+- maybe this blog post helps: {http://robjhyndman.com/hyndsight/crop-r-figures/}} \st{I cropped them using pdfcrop for now.
+I'll add the rest to the function.}
 
-\hh{Please add frames around the panels - it helps with separating the graphs from one another}
+\hh{Please add frames around the panels - it helps with separating the graphs from one another} \st{Okay - easy fix!}
 

### 2016-04-06 Code/complete_lineup_creation.R

@@ -66,11 +66,14 @@ create_smfriend_lu <- function(null_eff_struct, test_eff_struct, M, my_dat=mysma
   to_plot$X2 <- as.factor(to_plot$X2)
   plot <- ggplot(data = to_plot, aes(from_id = X1, to_id = X2)) + 
     geom_net(fiteach = TRUE, directed = F, size = 1, arrowsize = .5) +
-    facet_wrap(~plot_order) + theme_net() + theme(panel.background = element_rect(fill = "white", color = 'white'))
+    facet_wrap(~plot_order) + theme_net() + theme(panel.background = element_rect(fill = "white", color = 'black'))
   data_plot_id <- unique(to_plot$plot_order[which(to_plot$count == M)])
   return(list(data = to_plot, lineup = plot, test_id = data_plot_id))
 }
 lu1 <- create_smfriend_lu(null_eff_struct = null_model_eff2, test_eff_struct = eff_models_smallFriends[[39]], M = 6)
+savepdf(file = "Lineup-Images/pdfs/testcropping", width = 7.2*2.54, height = 4.5*2.54)
+print(lu1$lineup)
+dev.off()
 
 scriptURL = "http://www.hofroe.net/examples/lineup/action-back.js"
 
@@ -111,6 +114,15 @@ make_interactive <- function(filename2, script, toggle="toggle", high = "#c5c5c5
   grid.export(filename2)
 }
 
+# trim white space function {http://robjhyndman.com/hyndsight/crop-r-figures/}
+savepdf <- function(file, width=16, height=10)
+{
+  fname <- paste(file,".pdf",sep="")
+  pdf(fname, width=width/2.54, height=height/2.54,
+      pointsize=10)
+  par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(3.3,3.5,1.1,1))
+}
+  
 # function to create the lineups and linup data
 
 make_interactive_lineups <- function(M, reps, model.effects, null.effects){
@@ -118,8 +130,10 @@ make_interactive_lineups <- function(M, reps, model.effects, null.effects){
     lineup <- create_smfriend_lu(null_eff_struct = null.effects, test_eff_struct = model.effects, M = M)
     filename <- paste0("Data/lineupdata/smallfriends-m-", M, '-rep-', r, ".csv")
[…]

### 2016-04-07 lineups.Rnw
add some model description

--- 2016-04-06 lineups.Rnw
@@
 \documentclass[12pt]{article}
-\usepackage{amsmath,amssymb,mathrsfs,fancyhdr,syntonly,lastpage,hyperref,enumitem,graphicx, lscape, forloop}
+\usepackage{amsmath,amssymb,amsfonts, mathrsfs,fancyhdr,syntonly,lastpage,hyperref,enumitem,graphicx, lscape, forloop, url}
 
@@
 \newcommand{\hh}[1]{{\color{magenta} #1}}
+\newcommand{\st}[1]{{\color{red} #1}}
 
@@
 \rhead{Samantha Tyner}
-\hh{Cool, you got all the singletons to display! Nice work!}
-\hh{Could you add a bit more information on the lineups?
-Maybe we could have a list of models/hypothesis at the beginning and then sections that summarize things a bit more.
-Right now I am not quite sure about how to refer to each individual lineup.
-- OK, fixed that by including the names of the files.}
-
-\hh{Could you crop the white space from the lineup before saving it? - maybe this blog post helps: {http://robjhyndman.com/hyndsight/crop-r-figures/}}
-
-\hh{Please add frames around the panels - it helps with separating the graphs from one another}
 
 \tableofcontents
+
+\section{Models}
+
+We fit two different stochastic actor-oriented models to a subset of the 50 actor dataset from the ``Teenage Friends and Lifestyle Study" that is provided on the RSiena webpage.\footnote{\url{http://www.stats.ox.ac.uk/~snijders/siena/s50_data.htm}} The subset contained actors 20 through 35 and the ties between them, as well as the drinking behavior of each actor at each of the three waves.
+This specific subset was chosen because it showed somewhat higher connectivity than other subsets, as we've emphasized in the visualizations of the three network adjacency matrices below.
+For model fitting, we condition on wave 1 and estimate the parameters of our models from the second and third waves.
+
+<<get_sm_friends, echo=FALSE, fig.width=8, fig.height=3, message = FALSE, warning = FALSE>>=
+source("Code/small_friends.R")
+ggplot(data= view_rs2, aes(x = x, y=y)) + 
+  geom_tile(aes(fill = as.factor(value))) + scale_fill_manual(values = c('white', 'black')) +
+  geom_rect(data= NULL, inherit.aes = FALSE, 
+            aes(xmin = 20, xmax = 35, ymin = 20, ymax = 35, fill =NA), color = 'red') + 
+  facet_wrap(~wave) + theme_bw() + 
+  theme(aspect.ratio=1)
+
+@
+
+The first wave of the network, which is conditioned on in estimation, is given below.
+
+<<wave1, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3>>=
+library(sna)
+library(network)
+actual1 <- merge(data.frame(as.edgelist(as.network(fd2.w1))), 
+                 data.frame(id = 1:16, drink = drink2[,1]), 
+                 by.x = "X1", by.y = "id", all = T)
+for (j in 1:nrow(actual1)){
+      if (!(actual1$X1[j] %in% actual1$X2) & is.na(actual1$X2[j])){
+        actual1$X2[j] <- actual1$X1[j]
+      } else {actual1$X2[j] <- actual1$X2[j]}
+}
+actual2 <- merge(data.frame(as.edgelist(as.network(fd2.w2))), 
+                 data.frame(id = 1:16, drink = drink2[,2]), 
+                 by.x = "X1", by.y = "id", all = T)
+for (j in 1:nrow(actual2)){
+      if (!(actual2$X1[j] %in% actual2$X2) & is.na(actual2$X2[j])){
+        actual2$X2[j] <- actual2$X1[j]
+      } else {actual2$X2[j] <- actual2$X2[j]}
+    }
+actual1$wave <- 1
+actual2$wave <- 2
+
+waves <- rbind(actual1, actual2)
+library(geomnet)
+ggplot(data = actual1, aes(from_id = X1, to_id = X2)) + 
+  geom_net(label = TRUE, hjust = -.5, labelcolour = 'black', fiteach = T,
+           aes(color = as.factor(drink))) + 
+  theme_net() 
+#ggplot(data = waves, aes(from_id = X1, to_id = X2)) + 
+#  geom_net(label = TRUE, hjust = -.5, labelcolour = 'black', fiteach = T,
+#           aes(color = as.factor(drink))) + 
+#  theme_net() + theme(panel.background = element_rect(fill = "white", color = 'black')) + facet_wrap(~wave)
+@
+
+The two models we fit are a ``null model" (model X) and a ``data model" (model Y).
+The null model only includes the default parameters that are included in the RSiena R package.
+The data model includes one additional parameter that was determined to be significant by the Wald test in the RSienaTest package.
+Details on these effects are given in the table below.
+
+\begin{table}[h]
+\centering
+\tiny
+\begin{tabular}{lcccc}
+Effect name & Symbol & Corresponding Statistic & Model X Estimate & Model Y Estimate\\
+\hline
+\hline
+Rate 1 (wave 1 $\rightarrow$ 2) & $\alpha_1$ & ? \\
+Rate 2 (wave 2 $\rightarrow$ 3) & $\alpha_2$ & ? \\
+Outdegree & $\beta_1$ & $s_{i1}(x) = \sum\limits_{j=1}^n x_{ij}$ \\
+Reciprocity & $\beta_2$ & $s_{i2}(x) = \sum\limits_{j=1}^n x_{ij}x_{ji}$ \\
+Jumping Transitive Triplets & $\beta_3$ & $s_{i3}(x) = \sum\limits_{\forall j\neq h} x_{ij}x_{ih}x_{hj} \mathbb{I}(v_i = v_h \neq v_j) $
+\end{tabular}
+\end{table}
+
+After estimating the parameters in each of these models, we simulated from them to obtain realizations of each of the models. The objective function for each actor in each model is given below. 
+  \begin{align*}
+  f^{null}_{i}(x) & = \hat{\beta}_1^{null}s_{i1}(x) +  \hat{\beta}_2^{null}s_{i2}(x) \\
+  f^{data}_{i}(x) & = \hat{\beta}_1^{data}s_{i1}(x) +  \hat{\beta}_2^{data}s_{i2}(x) + \hat{\beta}_3^{data}s_{i3}(x) 
+  \end{align*}
 

### 2016-04-07 Code/small_friends.R

@@ -22,9 +22,12 @@ names(view_rs)[3:5] <- c('w1', 'w2','w3')
 view_rs %>% gather(wave, value, w1:w3) -> view_rs2
 
 # look at adjacency matrices
-ggplot(data= view_rs2, aes(x = x, y=y)) + 
-  geom_tile(aes(fill = as.factor(value))) + facet_wrap(~wave) +
-  theme(aspect.ratio=1)
+# ggplot(data= view_rs2, aes(x = x, y=y)) + 
+#   geom_tile(aes(fill = as.factor(value))) + scale_fill_manual(values = c('white', 'black')) +
+#   geom_rect(data= NULL, inherit.aes = FALSE, 
+#             aes(xmin = 20, xmax = 35, ymin = 20, ymax = 35, fill =NA), color = 'red') + 
+#   facet_wrap(~wave) + theme_bw() + 
+#   theme(aspect.ratio=1)
 
 #looks like 20:35 are pretty well connected, go with those for now
 fd2.w1 <- friend.data.w1[20:35,20:35]
@@ -44,124 +47,124 @@ friend2 <- sienaDependent(friendData2)
 alcohol2 <- varCovar( as.matrix(drink2) )
 # create siena data object
 mysmalldata <- sienaDataCreate( friend2, alcohol2)
-null_model_eff2 <- getEffects(mysmalldata)
+#null_model_eff2 <- getEffects(mysmalldata)
 #test_model_eff2 <- includeEffects( null_model_eff2, egoX, altX, egoXaltX, interaction1 = "alcohol2")
-test_model_eff2 <- includeEffects( null_model_eff2, sameXTransTrip, interaction1 = "alcohol2")
-myalgorithm2 <- sienaAlgorithmCreate( projname = 's50' , n3 = 1000)
-ests_null <- siena07( myalgorithm2, data = mysmalldata, returnDeps = TRUE, effects = null_model_eff2, batch=TRUE, verbose = FALSE)
-ests_test <- siena07( myalgorithm2, data = mysmalldata, returnDeps = TRUE, effects = test_model_eff2, batch=TRUE, verbose = FALSE)
+#test_model_eff2 <- includeEffects( null_model_eff2, sameXTransTrip, interaction1 = "alcohol2")
+#myalgorithm2 <- sienaAlgorithmCreate( projname = 's50' , n3 = 1000)
[…]

### 2016-04-08 lineups.Rnw
tiny stuff

@@
 After estimating the parameters in each of these models, we simulated from them to obtain realizations of each of the models. The objective function for each actor in each model is given below. 
+\hh{It might be better to re-phrase the two functions below from `null' and `data' to model $M_1$ and model $M_2$, where $M_1$ is nested within model $M_2$.
+Both $M_1$ and $M_2$ are based on the data - so the distinction between null and data might add a bit of confusion, when we flip the models around in the next step.
+}
   \begin{align*}

### 2016-04-09 lineups.Rnw
start writing intro

@@
 \tableofcontents
+
+\section{Introduction}
+
+Stochastic actor-oriented models (SAOMs) are a set of network models that consider both network and actor effects when determining how and why a network changes over time.
+Network effects, like reciprocity and outdegree, only consider the ties between actors as important to network change, while actor effects allow for covariate values associated with each actor to influence ties between actors.
+In the SAOM literature, a network is denoted by $x(t_m)$, has $n$ actors, and for two actors $i \neq j \in \{1, 2, \dots, n\}$, the relationship between them is denoted by $x_ij$.
+This relationship $x_{ij}$ is equal to 1 if there is a tie between actors $i$ and $j$ and it is 0 if there is not a tie between them.
+The time component, $t_m$, represents the $m = 1, 2, \dots, M$ points of observation of the network.
+These networks $x(t_1), \dots, x(t_M)$ are modeled as observations of a continuous-time Markov chain, where most of the steps in the chain are unobserved.
 
@@
 
-The two models we fit are a ``null model" (model X) and a ``data model" (model Y).
-The null model only includes the default parameters that are included in the RSiena R package.
-The data model includes one additional parameter that was determined to be significant by the Wald test in the RSienaTest package.
+<<modelfit, echo=FALSE, results = 'hide', message=FALSE, warning = FALSE, cache = T>>=
+source("Code/complete_lineup_creation.R")
+lineup1 <- create_smfriend_lu(null_eff_struct = null_model_eff2, test_eff_struct = eff_models_smallFriends[[39]], M = 3)
+@
+
+The two models we fit are a ``null model" (model $M_1$) and an ``alternative model" (model $M_2$).
+The null model only includes the default parameters that are included in the RSiena estimation.
+The alternative model includes one additional covariate parameter that was determined to be significant by the Wald test in the RSienaTest package.
 Details on these effects are given in the table below.
@@
 \begin{tabular}{lcccc}
-Effect name & Symbol & Corresponding Statistic & Model X Estimate & Model Y Estimate\\
+Effect name & Symbol & Corresponding Statistic & $M_1$ Estimate & Model Y Estimate\\
 \hline
 \hline
-Rate 1 (wave 1 $\rightarrow$ 2) & $\alpha_1$ & ? \\
-Rate 2 (wave 2 $\rightarrow$ 3) & $\alpha_2$ & ? \\
-Outdegree & $\beta_1$ & $s_{i1}(x) = \sum\limits_{j=1}^n x_{ij}$ \\
-Reciprocity & $\beta_2$ & $s_{i2}(x) = \sum\limits_{j=1}^n x_{ij}x_{ji}$ \\
-Jumping Transitive Triplets & $\beta_3$ & $s_{i3}(x) = \sum\limits_{\forall j\neq h} x_{ij}x_{ih}x_{hj} \mathbb{I}(v_i = v_h \neq v_j) $
+Rate 1 (wave 1 $\rightarrow$ 2) & $\alpha_1$ & $\sum\limits_{i,j = 1 i\neq j}^n (x_{ij}(t_2) - x_{ij}(t_1))^2 $ & \Sexpr{round(lineup1$null_mod_ests[1], 3)} & \Sexpr{round(lineup1$alt_mod_estimates[1],3)}\\
+Rate 2 (wave 2 $\rightarrow$ 3) & $\alpha_2$ & $\sum\limits_{i,j = 1 i\neq j}^n (x_{ij}(t_3) - x_{ij}(t_2))^2 $ & \Sexpr{round(lineup1$null_mod_ests[2], 3)} & \Sexpr{round(lineup1$alt_mod_estimates[2],3)}\\
+Outdegree & $\beta_1$ & $s_{i1}(x) = \sum\limits_{j=1}^n x_{ij}$ & \Sexpr{round(lineup1$null_mod_ests[3], 3)} & \Sexpr{round(lineup1$alt_mod_estimates[3],3)}\\
+Reciprocity & $\beta_2$ & $s_{i2}(x) = \sum\limits_{j=1}^n x_{ij}x_{ji}$ & \Sexpr{round(lineup1$null_mod_ests[4], 3)} & \Sexpr{round(lineup1$alt_mod_estimates[4],3)}\\
+Jumping Transitive Triplets & $\beta_3$ & $s_{i3}(x) = \sum\limits_{\forall j\neq h} x_{ij}x_{ih}x_{hj} \mathbb{I}(v_i = v_h \neq v_j)$ & -- & \Sexpr{round(lineup1$alt_mod_estimates[5],3)}
 \end{tabular}
@@
   f^{null}_{i}(x) & = \hat{\beta}_1^{null}s_{i1}(x) +  \hat{\beta}_2^{null}s_{i2}(x) \\
-  f^{data}_{i}(x) & = \hat{\beta}_1^{data}s_{i1}(x) +  \hat{\beta}_2^{data}s_{i2}(x) + \hat{\beta}_3^{data}s_{i3}(x) 
+  f^{alt}_{i}(x) & = \hat{\beta}_1^{alt}s_{i1}(x) +  \hat{\beta}_2^{alt}s_{i2}(x) + \hat{\beta}_3^{data}s_{i3}(x) 
   \end{align*}
+The rate parameters, $\alpha_1$ and $\alpha_2$ represent how many opportunities for change each actor gets when moving from wave 1 to 2 and from wave 2 to 3, respectively.
 

### 2016-04-09 Code/complete_lineup_creation.R

@@ -4,7 +4,7 @@ library(grid)
 library(gridSVG)
 ##### LOAD IN DATA #####
 #looks like 20:35 are pretty well connected, go with those for now
-setwd("Desktop/NetworksResearch/NetworksVizInference/")
+#setwd("Desktop/NetworksResearch/NetworksVizInference/")
 friend.data.w1 <- as.matrix(read.table("Data/s50_data/s50-network1.dat"))
 friend.data.w2 <- as.matrix(read.table("Data/s50_data/s50-network2.dat"))
 friend.data.w3 <- as.matrix(read.table("Data/s50_data/s50-network3.dat"))
@@ -68,12 +68,12 @@ create_smfriend_lu <- function(null_eff_struct, test_eff_struct, M, my_dat=mysma
     geom_net(fiteach = TRUE, directed = F, size = 1, arrowsize = .5) +
     facet_wrap(~plot_order) + theme_net() + theme(panel.background = element_rect(fill = "white", color = 'black'))
   data_plot_id <- unique(to_plot$plot_order[which(to_plot$count == M)])
-  return(list(data = to_plot, lineup = plot, test_id = data_plot_id))
+  return(list(data = to_plot, lineup = plot, test_id = data_plot_id, null_mod_ests = c(null_net$rate,null_net$theta), alt_mod_estimates = c(test_net$rate,test_net$theta)))
 }
-lu1 <- create_smfriend_lu(null_eff_struct = null_model_eff2, test_eff_struct = eff_models_smallFriends[[39]], M = 6)
-savepdf(file = "Lineup-Images/pdfs/testcropping", width = 7.2*2.54, height = 4.5*2.54)
-print(lu1$lineup)
-dev.off()
+# lu1 <- create_smfriend_lu(null_eff_struct = null_model_eff2, test_eff_struct = eff_models_smallFriends[[39]], M = 6)
+# savepdf(file = "Lineup-Images/pdfs/testcropping", width = 7.2*2.54, height = 4.5*2.54)
+# print(lu1$lineup)
+# dev.off()
 
 scriptURL = "http://www.hofroe.net/examples/lineup/action-back.js"
 
@@ -138,12 +138,13 @@ make_interactive_lineups <- function(M, reps, model.effects, null.effects){
     make_interactive(filename2 = tmpfile, script=scriptURL,  
                      high="#d5d5d5",  background="#ffffff")
   }
+  return(list(lineup = lineup))
 }
 
[…]

### 2016-04-09 lineups.Rnw
moar intro

@@
 
+\par Between any two discrete time points, $t_{m}$ and $t_{m-1}$ for $m = 2, \dots, M$, each actor gets the opportunity to change one of their ties at a rate of $\alpha_{m-1}$.
+This is due to the properties of the Markov chain, and it also assumes that the rate of change is constant for each actor between $t_{m-1}$ and $t_{m}$.
+When an actor is given an opportunity to change, it tries to maximize its objective function, $f_i(\beta, x)$, where $x$ is the potential new network state that actor $i$ can reach by changing one of its ties, $x_{ij}$.
+In model estimation, the parameters to be estimated are $\alpha_1, \dots, \alpha_{M-1}$ and the (possibly vector-valued) $\beta$
+
 \section{Models}
 
-We fit two different stochastic actor-oriented models to a subset of the 50 actor dataset from the ``Teenage Friends and Lifestyle Study" that is provided on the RSiena webpage.\footnote{\url{http://www.stats.ox.ac.uk/~snijders/siena/s50_data.htm}} The subset contained actors 20 through 35 and the ties between them, as well as the drinking behavior of each actor at each of the three waves.
+We fit two different stochastic actor-oriented models to a subset of the 50 actor dataset from the ``Teenage Friends and Lifestyle Study" that is provided on the RSiena webpage.\footnote{\url{http://www.stats.ox.ac.uk/~snijders/siena/s50_data.htm}} We chose to subset the data to decrease the cognitive load on our experiment's subjects.
+The subset contained actors 20 through 35 and the ties between them, as well as the drinking behavior of each actor at each of the three waves.
 This specific subset was chosen because it showed somewhat higher connectivity than other subsets, as we've emphasized in the visualizations of the three network adjacency matrices below.

### 2016-04-09 Code/reverse_lineups.R
Reverse the lineup structure

@@ -1,0 +1,68 @@
+source("Code/complete_lineup_creation.R")
+# function to reverse the models (from M-1 X (null) & 1 Y (alt) to  1 X & M-1 alt)
+create_smfriend_lu_rev <- function(null_eff_struct, test_eff_struct, M, my_dat=mysmalldata, ...){
+  # browser()
+  alt_data <- NULL
+  n <- M-1
+  alt_net <- siena07( myalgorithm2, data = my_dat, returnDeps = TRUE, effects = test_eff_struct, batch=TRUE, verbose = FALSE)
+  test_eff_struct2 <- test_eff_struct
+  test_eff_struct2$initialValue[test_eff_struct2$include] <- c(alt_net$rate,alt_net$theta)
+  sim_algorithm <- sienaAlgorithmCreate( projname = "sim_model", cond = FALSE,
+                                         useStdInits = FALSE, nsub = 0 , simOnly = TRUE)
+  sim_ans    <-  siena07( sim_algorithm, data = my_dat, effects = test_eff_struct2, returnDeps = TRUE)
+  for (i in 1:n){
+    net_df <- merge(data.frame(sim_ans$sims[[1000-(n-i)]][[1]][[1]][[1]])[,-3], 
+                    data.frame(id = 1:16), by.x = "X1", by.y = "id",
+                    all = T)
+    for (j in 1:nrow(net_df)){
+      if (!(net_df$X1[j] %in% net_df$X2) & is.na(net_df$X2[j])){
+        net_df$X2[j] <- net_df$X1[j]
+      } else {net_df$X2[j] <- net_df$X2[j]}
+    }
+    net_df$count <- i 
+    alt_data <- rbind(alt_data, net_df)
+  }
+  null_net <- siena07( myalgorithm2, data = my_dat, returnDeps = TRUE, effects =null_eff_struct, batch=TRUE, verbose = FALSE)
+  null_data <- merge(data.frame(null_net$sims[[1000]][[1]][[1]][[1]])[,-3], 
+                     data.frame(id = 1:16), by.x = "X1", by.y = "id",
+                     all = T)
+  for (j in 1:nrow(null_data)){
+    if (!(null_data$X1[j] %in% null_data$X2) & is.na(null_data$X2[j])){
+      null_data$X2[j] <- null_data$X1[j]
+    } else {null_data$X2[j] <- null_data$X2[j]}
+  }
+  null_data$count <- M  # test data is always indexed by M
[…]

### 2016-04-13 lineups.Rnw
Add reverse lineups

@@
 
+\section{Model X versus Y lineup size 16}
+\newcounter{k4}
+\forloop{k4}{1}{\value{k4} < 21}{smallfriends-m-16-rep-\arabic{k4}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-m-16-rep-\arabic{k4}} \\ }
+
+\section{Model Y versus X lineup size 3}
+\newcounter{k5}
+\forloop{k5}{1}{\value{k5} < 21}{smallfriendsrev-m-3-rep-\arabic{k5}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriendsrev-m-3-rep-\arabic{k5}} \\ }
+
+\section{Model Y versus X lineup size 6}
+\newcounter{k6}
+\forloop{k6}{1}{\value{k6} < 21}{smallfriendsrev-m-6-rep-\arabic{k6}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriendsrev-m-6-rep-\arabic{k6}} \\ }
+
+\section{Model Y versus X lineup size 9}
+\newcounter{k7}
+\forloop{k7}{1}{\value{k7} < 21}{smallfriendsrev-m-9-rep-\arabic{k7}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriendsrev-m-9-rep-\arabic{k7}} \\ }
+
+\section{Model Y versus X lineup size 12}
+\newcounter{k8}
+\forloop{k8}{1}{\value{k8} < 21}{smallfriendsrev-m-12-rep-\arabic{k8}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriendsrev-m-12-rep-\arabic{k8}} \\ }
+
+\section{Model Y versus X lineup size 16}
+\newcounter{k9}
+\forloop{k9}{1}{\value{k9} < 21}{smallfriendsrev-m-16-rep-\arabic{k9}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriendsrev-m-16-rep-\arabic{k9}} \\ }
 \end{document}

### 2016-04-13 Code/complete_lineup_creation.R

@@ -148,3 +148,5 @@ make_interactive_lineups <- function(M, reps, model.effects, null.effects){
 #make_interactive_lineups(M = 9, reps=20, model.effects = null_model_eff2, null.effects = eff_models_smallFriends[[39]])
 
 #make_interactive_lineups(M = 12, reps=20, model.effects = null_model_eff2, null.effects = eff_models_smallFriends[[39]])
+
+# make_interactive_lineups(M = 16, reps=20, model.effects = null_model_eff2, null.effects = eff_models_smallFriends[[39]])

### 2016-04-13 Code/reverse_lineups.R

@@ -66,3 +66,5 @@ make_interactive_lineups_rev <- function(M, reps, model.effects, null.effects){
 #make_interactive_lineups_rev(M = 9, reps=20, model.effects = null_model_eff2, null.effects = eff_models_smallFriends[[39]])
 
 #make_interactive_lineups_rev(M = 12, reps=20, model.effects = null_model_eff2, null.effects = eff_models_smallFriends[[39]])
+
+make_interactive_lineups_rev(M = 16, reps=20, model.effects = null_model_eff2, null.effects = eff_models_smallFriends[[39]])

### 2016-04-13 lineups.Rnw
couple of comments/questions for the meeting

@@
 
+\hh{XXX Sam, when you are creating these lineups, could you split the functions into two, and create and save the data into a csv file in a first function, and then in a second function get the data and put it into a network?
+- That way we don't have to create the data again, if/when we change the design.}
+
+\hh{XXX question for the meeting: right now the difference between models X and Y are the Jumping Transitive Triplets, which we might (or not) see in the graphs.
+Is there another effect that is significant at this stage as well?
+It would be good to have the same kind of lineups for a different effect.
+}
+
+\hh{XXX another question for tomorrow: in the first four reps of the reverse model it looks like one graph is more complex than the other two models ...
+which is not what I'd expect from this scenario, because ehere, we are looking for the simplest model, right?
+}
+
 \section{Model X versus Y lineup size 3}

### 2016-04-14 Code/get_coeff_dists.R
after 1000 runs of each model, here are the coefs.

@@ -1,0 +1,81 @@
+# estimate model effects
+
+library(RSiena)
+library(geomnet)
+library(grid)
+library(gridSVG)
+##### LOAD IN DATA #####
+#looks like 20:35 are pretty well connected, go with those for now
+#setwd("Desktop/NetworksResearch/NetworksVizInference/")
+friend.data.w1 <- as.matrix(read.table("../Data/s50_data/s50-network1.dat"))
+friend.data.w2 <- as.matrix(read.table("../Data/s50_data/s50-network2.dat"))
+friend.data.w3 <- as.matrix(read.table("../Data/s50_data/s50-network3.dat"))
+fd2.w1 <- friend.data.w1[20:35,20:35]
+fd2.w2 <- friend.data.w2[20:35,20:35]
+fd2.w3 <- friend.data.w3[20:35,20:35]
+# read in covariate data
+drink <- as.matrix(read.table("../Data/s50_data/s50-alcohol.dat"))
+drink2 <- drink[20:35,]
+# siena data
+friendData2 <- array( c( fd2.w1, fd2.w2, fd2.w3 ),
+                      dim = c( 16, 16, 3 ) )
+# and next give this the role of the dependent variable:
+friend2 <- sienaDependent(friendData2)
+alcohol2 <- varCovar( as.matrix(drink2) )
+# create siena data object
+mysmalldata <- sienaDataCreate( friend2, alcohol2)
+null_model_eff2 <- getEffects(mysmalldata)
+myalgorithm2 <- sienaAlgorithmCreate( projname = 's50' , n3 = 1000)
+sig_eff_names <- read.csv("../Data/sig_eff_names_smfriends.csv")
+runs_models_smallFriends <- readRDS("../Data/runs_models_smallFriends.RDS")
+eff_models_smallFriends <- readRDS("../Data/eff_models_smallFriends.RDS")
+
+
+# write a function to get dist of the parameter ests
+get_eff_val_dist <- function(B, eff_struct, sd = 2017){
+  require(RSiena)
+  set.seed(sd)
+  df.ests <- NULL
+  for (i in 1:B){
+    null_net <- siena07(myalgorithm2, data = mysmalldata, returnDeps = TRUE, effects = eff_struct, batch=TRUE, verbose = FALSE, silent = TRUE)
+    ests <- c(null_net$rate,null_net$theta)
+    df.ests <- rbind(df.ests, ests)
+  }
[…]

### 2016-04-14 lineups.Rnw
more questions
histograms of simulated distributions of model parameters

@@
 The alternative model includes one additional covariate parameter that was determined to be significant by the Wald test in the RSienaTest package.
-Details on these effects are given in the table below.
+Details on these effects are given in Table~\ref{tab:models}.
 
 \begin{table}[h]
+\caption{\label{tab:models}Parameters and estimates of models $M_1$ and $M_2$.}
 \centering
-\tiny
-\begin{tabular}{lcccc}
-Effect name & Symbol & Corresponding Statistic & $M_1$ Estimate & Model Y Estimate\\
+\scalebox{0.8}{
+\begin{tabular}{lccrr}
+Effect name & Symbol & Corresponding Statistic & $M_1$ Estimate & $M_2$ Estimate\\
 \hline
@@
 Jumping Transitive Triplets & $\beta_3$ & $s_{i3}(x) = \sum\limits_{\forall j\neq h} x_{ij}x_{ih}x_{hj} \mathbb{I}(v_i = v_h \neq v_j)$ & -- & \Sexpr{round(lineup1$alt_mod_estimates[5],3)}
-\end{tabular}
+\end{tabular}}
 \end{table}
 
-After estimating the parameters in each of these models, we simulated from them to obtain realizations of each of the models. The objective function for each actor in each model is given below. 
+\begin{figure}
+<<hist-estimates, echo=FALSE>>=
+# nulls <- read.csv("Data/distribution_null_model.csv")
+# names(nulls)[-1] <- c("alpha1", "alpha2", "beta1", "beta2")
+# nulls$Sim <- 1:nrow(nulls)
+# nulls$Model <- "M1"
+# 
+# alt <- read.csv("Data/distribution_jumpTT_model.csv")
+# names(alt)[-1] <- c("alpha1", "alpha2", "beta1", "beta2", "beta3")
+# alt$Sim <- 1:nrow(alt)
+# alt$Model <- "M2"
+# 
+# alt2 <- gather(alt, parameter, estimate, 2:6)
+# null2 <- gather(nulls, parameter, estimate, 2:5)
+# 
+# simu <- rbind(alt2[,-1], null2[,-1])
+# write.csv(simu, "Data/simulation-1000-M1-M2.csv", row.names=FALSE)
+
+simu <- read.csv("Data/simulation-1000-M1-M2.csv")
+qplot(data=simu, estimate, fill=Model, geom="density", alpha=I(0.75)) + facet_wrap(~parameter, scales="free") + theme_bw()
+@
+\caption{\label{fig:hist-simulations}Histogram of the distribution of model parameters based on 1,000 simulation runs.
+Model parameter $\beta_3$ for jumping transitive triplets in model $M_2$ is significantly different from zero, but its inclusion also leads to significant changes in the  other model parameters of model $M_1$.
+}
+\end{figure}
+
+After estimating the parameters in each of these models, we simulate from them to obtain realizations of each of the models. The objective function for each actor in each model is given below. 
 \hh{It might be better to re-phrase the two functions below from `null' and `data' to model $M_1$ and model $M_2$, where $M_1$ is nested within model $M_2$.
@@
 }
+
+\hh{XXX would a size 15 (= 5x3) lineup be more `square' than a size 16 one? }
 

### 2016-04-15 lineups.Rnw
teensy bit of a comment

@@
 \begin{table}[h]
-\caption{\label{tab:models}Parameters and estimates of models $M_1$ and $M_2$.}
+\caption{\label{tab:models}Parameters and estimates of models $M_1$ and $M_2$.
+\hh{Estimates stem from one of the simulations.
+XXX We probably should change that to the average of the 1000 simulations, and then simulate the lineups from those, like you suggested.
+}}
 \centering
@@
 \begin{tabular}{lccrr}
-Effect name & Symbol & Corresponding Statistic & $M_1$ Estimate & $M_2$ Estimate\\
+Effect name & Parameter & Corresponding Statistic & $M_1$ Estimate & $M_2$ Estimate\\
 \hline
@@
 \begin{figure}
-<<hist-estimates, echo=FALSE>>=
+\centering
+<<hist-estimates, echo=FALSE, fig.width=8, fig.height=5, out.width='.8\\linewidth'>>=
 # nulls <- read.csv("Data/distribution_null_model.csv")
@@
 After estimating the parameters in each of these models, we simulate from them to obtain realizations of each of the models. The objective function for each actor in each model is given below. 
-\hh{It might be better to re-phrase the two functions below from `null' and `data' to model $M_1$ and model $M_2$, where $M_1$ is nested within model $M_2$.
-Both $M_1$ and $M_2$ are based on the data - so the distinction between null and data might add a bit of confusion, when we flip the models around in the next step.
-}
   \begin{align*}
-  f^{null}_{i}(x) & = \hat{\beta}_1^{null}s_{i1}(x) +  \hat{\beta}_2^{null}s_{i2}(x) \\
-  f^{alt}_{i}(x) & = \hat{\beta}_1^{alt}s_{i1}(x) +  \hat{\beta}_2^{alt}s_{i2}(x) + \hat{\beta}_3^{data}s_{i3}(x) 
+  f^{M_1}_{i}(x) & = \hat{\beta}_1^{M_1}s_{i1}(x) +  \hat{\beta}_2^{M_1}s_{i2}(x) \\
+  f^{M_2}_{i}(x) & = \hat{\beta}_1^{M_2}s_{i1}(x) +  \hat{\beta}_2^{M_2}s_{i2}(x) + \hat{\beta}_3^{M_2}s_{i3}(x) 
   \end{align*}

### 2016-04-17 lineups.Rnw
table shows means from simulation

@@
 
-\begin{table}[h]
-\caption{\label{tab:models}Parameters and estimates of models $M_1$ and $M_2$.
-\hh{Estimates stem from one of the simulations.
-XXX We probably should change that to the average of the 1000 simulations, and then simulate the lineups from those, like you suggested.
-}}
-\centering
-\scalebox{0.8}{
-\begin{tabular}{lccrr}
-Effect name & Parameter & Corresponding Statistic & $M_1$ Estimate & $M_2$ Estimate\\
-\hline
-\hline
-Rate 1 (wave 1 $\rightarrow$ 2) & $\alpha_1$ & $\sum\limits_{i,j = 1 i\neq j}^n (x_{ij}(t_2) - x_{ij}(t_1))^2 $ & \Sexpr{round(lineup1$null_mod_ests[1], 3)} & \Sexpr{round(lineup1$alt_mod_estimates[1],3)}\\
-Rate 2 (wave 2 $\rightarrow$ 3) & $\alpha_2$ & $\sum\limits_{i,j = 1 i\neq j}^n (x_{ij}(t_3) - x_{ij}(t_2))^2 $ & \Sexpr{round(lineup1$null_mod_ests[2], 3)} & \Sexpr{round(lineup1$alt_mod_estimates[2],3)}\\
-Outdegree & $\beta_1$ & $s_{i1}(x) = \sum\limits_{j=1}^n x_{ij}$ & \Sexpr{round(lineup1$null_mod_ests[3], 3)} & \Sexpr{round(lineup1$alt_mod_estimates[3],3)}\\
-Reciprocity & $\beta_2$ & $s_{i2}(x) = \sum\limits_{j=1}^n x_{ij}x_{ji}$ & \Sexpr{round(lineup1$null_mod_ests[4], 3)} & \Sexpr{round(lineup1$alt_mod_estimates[4],3)}\\
-Jumping Transitive Triplets & $\beta_3$ & $s_{i3}(x) = \sum\limits_{\forall j\neq h} x_{ij}x_{ih}x_{hj} \mathbb{I}(v_i = v_h \neq v_j)$ & -- & \Sexpr{round(lineup1$alt_mod_estimates[5],3)}
-\end{tabular}}
-\end{table}
-
-\begin{figure}
-\centering
-<<hist-estimates, echo=FALSE, fig.width=8, fig.height=5, out.width='.8\\linewidth'>>=
+<<readsimu, echo=FALSE>>=
 # nulls <- read.csv("Data/distribution_null_model.csv")
@@
 simu <- read.csv("Data/simulation-1000-M1-M2.csv")
+means <- simu %>% group_by(Model, parameter) %>% summarize(
+  mean = mean(estimate)
+)
+@
+
+\begin{table}[h]
+\caption{\label{tab:models}Parameters and estimates of models $M_1$ and $M_2$.
+\hh{Estimates stem from one of the simulations.
+XXX We probably should change that to the average of the 1000 simulations, and then simulate the lineups from those, like you suggested.
+}}
+\centering
+\scalebox{0.8}{
+\begin{tabular}{lccrr}
+Effect name & Parameter & Corresponding Statistic & $M_1$ Estimate & $M_2$ Estimate\\
+\hline
+\hline
+Rate 1 (wave 1 $\rightarrow$ 2) & $\alpha_1$ & $\sum\limits_{i,j = 1 i\neq j}^n (x_{ij}(t_2) - x_{ij}(t_1))^2 $ & 
+\Sexpr{round(subset(means, Model=="M1" & parameter=="alpha1")$mean, 2)} &
+\Sexpr{round(subset(means, Model=="M2" & parameter=="alpha1")$mean, 2)}
+\\
+Rate 2 (wave 2 $\rightarrow$ 3) & $\alpha_2$ & $\sum\limits_{i,j = 1 i\neq j}^n (x_{ij}(t_3) - x_{ij}(t_2))^2 $ & 
+\Sexpr{round(subset(means, Model=="M1" & parameter=="alpha2")$mean, 2)} &
+\Sexpr{round(subset(means, Model=="M2" & parameter=="alpha2")$mean, 2)}
+\\
+Outdegree & $\beta_1$ & $s_{i1}(x) = \sum\limits_{j=1}^n x_{ij}$ & \Sexpr{round(subset(means, Model=="M1" & parameter=="beta1")$mean, 2)} &
+\Sexpr{round(subset(means, Model=="M2" & parameter=="beta1")$mean, 2)}
+\\
+Reciprocity & $\beta_2$ & $s_{i2}(x) = \sum\limits_{j=1}^n x_{ij}x_{ji}$ & \Sexpr{round(subset(means, Model=="M1" & parameter=="beta2")$mean, 2)} &
+\Sexpr{round(subset(means, Model=="M2" & parameter=="beta2")$mean, 2)}
+\\
+Jumping Transitive Triplets & $\beta_3$ & $s_{i3}(x) = \sum\limits_{\forall j\neq h} x_{ij}x_{ih}x_{hj} \mathbb{I}(v_i = v_h \neq v_j)$ & -- & \Sexpr{round(subset(means, Model=="M2" & parameter=="beta3")$mean, 2)}
+\end{tabular}}
+\end{table}
+
+\begin{figure}
+\centering
+<<hist-estimates, dependson='read-simu', echo=FALSE, fig.width=8, fig.height=5, out.width='.8\\linewidth'>>=
 qplot(data=simu, estimate, fill=Model, geom="density", alpha=I(0.75)) + facet_wrap(~parameter, scales="free") + theme_bw()

### 2016-04-19 Code/fixed_coef_sims.R
Create lineups for fixed models and push lineup data

@@ -1,0 +1,234 @@
+# write a new function for lineup creation. 
+# WANT:
+# 1. All lineup images come from same instance of estimated parameters
+# Doesn't happen now as it is written, due to the create_smfriends_lineup fun.
+# idea: fit the models 100 times and average the effect estimates to get values
+# to use in all 
+# 2. Split the function into two: 
+#  a. create and save the data into a csv file 
+#  b. get the data and put it into a network
+
+# 
+
+library(ggplot2)
+library(geomnet)
+library(dplyr)
+library(RSiena)
+library(grid)
+library(gridSVG)
+
+# get data 
+null_model_dist <- read.csv("../Data/distribution_null_model.csv")
+null_model_dist <- null_model_dist[,-1]
+alt_model_dist <- read.csv("../Data/distribution_jumpTT_model.csv")
+alt_model_dist <- alt_model_dist[,-1]
+alt_model2_dist <- read.csv("../Data/distribution_dblpairs_model.csv")
+alt_model2_dist <- alt_model2_dist[,-1]
+
+null_mod_sv <- as.numeric(null_model_dist %>% summarise_each(funs(mean)))
+alt_mod_sv <- as.numeric(alt_model_dist %>% summarise_each(funs(mean)))
+alt_mod2_sv <- as.numeric(alt_model2_dist %>% summarise_each(funs(mean)))
+
+eff_models_smallFriends <- readRDS("Data/eff_models_smallFriends.RDS")
+
+
+# function 1: create and save data:
+
+create_lu_dat_xy <- function(null_model, null_model_svs, alt_model, alt_model_svs, m, my_dat = mysmalldata, sd, ...){
+  require(RSiena)
+  # null model setup
+  null_model$initialValue[null_model$include] <- null_model_svs
+  # alt model setup
+  alt_model$initialValue[alt_model$include] <- alt_model_svs
+  # algorithm setup
+  sim_algorithm <- sienaAlgorithmCreate( projname = "sim_model", cond = FALSE,
+                                         useStdInits = FALSE, nsub = 0 , simOnly = TRUE)
+  # get model sims
+  set.seed(sd) # for replication
+  null_sims    <-  siena07(sim_algorithm, data = my_dat, effects = null_model, 
+                          returnDeps = TRUE, batch=TRUE, verbose = FALSE, 
[…]

### 2016-04-19 Code/get_coeff_dists.R

@@ -59,7 +59,7 @@ alt_model2_ests <- get_eff_val_dist(B = 1000, eff_struct = eff_models_smallFrien
 
 null_model_dist <- read.csv("../Data/distribution_null_model.csv")
 alt_model_dist <- read.csv("../Data/distribution_jumpTT_model.csv")
-alt_model2_dist <- read.csv("Data/distribtution_dblpairs_model.csv")
+alt_model2_dist <- read.csv("../Data/distribution_dblpairs_model.csv")
 
 #null_model_dist <- as.data.frame(null_model_dist)
 rownames(null_model_dist) <- 1:nrow(null_model_dist)

### 2016-04-19 Code/fixed_coef_sims.R
Whoops. Wrote over some files. Fixed it.

@@ -180,17 +180,23 @@ savepdf <- function(file, width=16, height=10)
 create_lu <- function(null_model, null_model_svs, alt_model, alt_model_svs, M, my_dat = mysmalldata, reps, yx = FALSE,...){
   for (r in 1:reps){
     if (yx == FALSE){
-      lu_dat <- create_lu_dat_xy(null_model, null_model_svs, alt_model, alt_model_svs, m=M, my_dat, sd = r)
-    } else lu_dat <- create_lu_dat_yx(null_model, null_model_svs, alt_model, alt_model_svs,m=M, my_dat, sd = r)
+      lu_dat <- create_lu_dat_xy(null_model, null_model_svs, alt_model, alt_model_svs, m=M, my_dat, sd = (r + M))
+    } else lu_dat <- create_lu_dat_yx(null_model, null_model_svs, alt_model, alt_model_svs,m=M, my_dat, sd = (r + M))
     lu_dat$plot_order <- rep(sample(M), as.vector(table(lu_dat$count)))
     lu_dat$X1 <- as.factor(lu_dat$X1)
     lu_dat$X2 <- as.factor(lu_dat$X2)
-    filename <- paste0("Data/lineupdata/smallfriends-m-", M, '-rep-', r, ".csv")
+    if (yx == FALSE){
+      filename <- paste0("Data/lineupdata/smallfriends-m-", M, '-rep-', r, ".csv")
+      filename2 <- sprintf("Lineup-Images/pdfs/smallfriends-m-%s-rep-%s",M,r)
+    } else {
+      filename <- paste0("Data/lineupdata/smallfriends-rev-m-", M, '-rep-', r, ".csv")
+      filename2 <- sprintf("Lineup-Images/pdfs/smallfriends-rev-m-%s-rep-%s",M,r)
+    }
     write.csv(lu_dat, file = filename)
     lu_plot <- ggplot(data = lu_dat, aes(from_id = X1, to_id = X2)) + 
       geom_net(fiteach = TRUE, directed = F, size = 1, arrowsize = .5) +
       facet_wrap(~plot_order) + theme_net() + theme(panel.background = element_rect(fill = "white", color = 'black'))
-    savepdf(file = sprintf("Lineup-Images/pdfs/smallfriends-m-%s-rep-%s",M,r), width = 8*2.54, height = 8*2.54)
+    savepdf(file = filename2, width = 8*2.54, height = 8*2.54)
[…]

### 2016-04-19 Code/fixed_coef_sims.R
Lineups with another significant effect. Description in paper to follow

@@ -177,7 +177,70 @@ savepdf <- function(file, width=16, height=10)
 }
 
 
-create_lu <- function(null_model, null_model_svs, alt_model, alt_model_svs, M, my_dat = mysmalldata, reps, yx = FALSE,...){
+# create_lu <- function(null_model, null_model_svs, alt_model, alt_model_svs, M, my_dat = mysmalldata, reps, yx = FALSE,...){
+#   for (r in 1:reps){
+#     if (yx == FALSE){
+#       lu_dat <- create_lu_dat_xy(null_model, null_model_svs, alt_model, alt_model_svs, m=M, my_dat, sd = (r + M))
+#     } else lu_dat <- create_lu_dat_yx(null_model, null_model_svs, alt_model, alt_model_svs,m=M, my_dat, sd = (r + M))
+#     lu_dat$plot_order <- rep(sample(M), as.vector(table(lu_dat$count)))
+#     lu_dat$X1 <- as.factor(lu_dat$X1)
+#     lu_dat$X2 <- as.factor(lu_dat$X2)
+#     if (yx == FALSE){
+#       filename <- paste0("Data/lineupdata/smallfriends-m-", M, '-rep-', r, ".csv")
+#       filename2 <- sprintf("Lineup-Images/pdfs/smallfriends-m-%s-rep-%s",M,r)
+#     } else {
+#       filename <- paste0("Data/lineupdata/smallfriends-rev-m-", M, '-rep-', r, ".csv")
+#       filename2 <- sprintf("Lineup-Images/pdfs/smallfriends-rev-m-%s-rep-%s",M,r)
+#     }
+#     write.csv(lu_dat, file = filename)
+#     lu_plot <- ggplot(data = lu_dat, aes(from_id = X1, to_id = X2)) + 
+#       geom_net(fiteach = TRUE, directed = F, size = 1, arrowsize = .5) +
+#       facet_wrap(~plot_order) + theme_net() + theme(panel.background = element_rect(fill = "white", color = 'black'))
+#     savepdf(file = filename2, width = 8*2.54, height = 8*2.54)
+#     print(lu_plot)
+#     dev.off()
+#     tmpfile <- sprintf("%s.svg",tempfile(tmpdir="Lineup-Images/svgs"))
+#     make_interactive(filename2 = tmpfile, script=scriptURL,  
+#                      high="#d5d5d5",  background="#ffffff")
+#     }
+# }
+
+# create_lu(null_model = null_model_eff2, null_model_svs = null_mod_sv,
[…]

### 2016-04-19 lineups.Rnw
model X is model M_1 and model Y is model M_2
add third model. not done yet.

@@
 
-We fit two different stochastic actor-oriented models to a subset of the 50 actor dataset from the ``Teenage Friends and Lifestyle Study" that is provided on the RSiena webpage.\footnote{\url{http://www.stats.ox.ac.uk/~snijders/siena/s50_data.htm}} We chose to subset the data to decrease the cognitive load on our experiment's subjects.
+We fit \st{three} different stochastic actor-oriented models to a subset of the 50 actor dataset from the ``Teenage Friends and Lifestyle Study" that is provided on the RSiena webpage.\footnote{\url{http://www.stats.ox.ac.uk/~snijders/siena/s50_data.htm}} We chose to subset the data to decrease the cognitive load on our experiment's subjects.
 The subset contained actors 20 through 35 and the ties between them, as well as the drinking behavior of each actor at each of the three waves.
@@
 
-The two models we fit are a ``null model" (model $M_1$) and an ``alternative model" (model $M_2$).
+The two models we fit are a ``null model" (model $M_1$) and two ``alternative models" (model $M_2$ and $M_3$).
 The null model only includes the default parameters that are included in the RSiena estimation.
-The alternative model includes one additional covariate parameter that was determined to be significant by the Wald test in the RSienaTest package.
+The alternative model $M_2$ includes one additional covariate parameter that was determined to be significant by the Wald test in the RSienaTest package, while the alternative model $M_3$ includes one additional structural parameter whose significance was determined in the same way.
 Details on these effects are given in Table~\ref{tab:models}.
@@
 )
+library(dplyr)
+null_model_dist <- read.csv("Data/distribution_null_model.csv")
+null_model_dist <- null_model_dist[,-1]
+alt_model_dist <- read.csv("Data/distribution_jumpTT_model.csv")
+alt_model_dist <- alt_model_dist[,-1]
+alt_model2_dist <- read.csv("Data/distribution_dblpairs_model.csv")
+alt_model2_dist <- alt_model2_dist[,-1]
+
+null_mod_sv <- as.numeric(null_model_dist %>% summarise_each(funs(mean)))
+alt_mod_sv <- as.numeric(alt_model_dist %>% summarise_each(funs(mean)))
+alt_mod2_sv <- as.numeric(alt_model2_dist %>% summarise_each(funs(mean)))
 @
@@
 \scalebox{0.8}{
-\begin{tabular}{lccrr}
-Effect name & Parameter & Corresponding Statistic & $M_1$ Estimate & $M_2$ Estimate\\
+\begin{tabular}{lccrrr}
+Effect name & Parameter & Corresponding Statistic & $M_1$ Estimate & $M_2$ Estimate & $M_3$ Estimate\\
 \hline
@@
 Rate 1 (wave 1 $\rightarrow$ 2) & $\alpha_1$ & $\sum\limits_{i,j = 1 i\neq j}^n (x_{ij}(t_2) - x_{ij}(t_1))^2 $ & 
-\Sexpr{round(subset(means, Model=="M1" & parameter=="alpha1")$mean, 2)} &
-\Sexpr{round(subset(means, Model=="M2" & parameter=="alpha1")$mean, 2)}
+\Sexpr{round(null_mod_sv[1], 2)} &
+\Sexpr{round(alt_mod_sv[1], 2)} & 
+\Sexpr{round(alt_mod2_sv[1], 2)} 
 \\
 Rate 2 (wave 2 $\rightarrow$ 3) & $\alpha_2$ & $\sum\limits_{i,j = 1 i\neq j}^n (x_{ij}(t_3) - x_{ij}(t_2))^2 $ & 
-\Sexpr{round(subset(means, Model=="M1" & parameter=="alpha2")$mean, 2)} &
-\Sexpr{round(subset(means, Model=="M2" & parameter=="alpha2")$mean, 2)}
+\Sexpr{round(null_mod_sv[2], 2)} &
+\Sexpr{round(alt_mod_sv[2], 2)} & 
+\Sexpr{round(alt_mod2_sv[2], 2)}
 \\
-Outdegree & $\beta_1$ & $s_{i1}(x) = \sum\limits_{j=1}^n x_{ij}$ & \Sexpr{round(subset(means, Model=="M1" & parameter=="beta1")$mean, 2)} &
-\Sexpr{round(subset(means, Model=="M2" & parameter=="beta1")$mean, 2)}
+Outdegree & $\beta_1$ & $s_{i1}(x) = \sum\limits_{j=1}^n x_{ij}$ & 
+\Sexpr{round(null_mod_sv[3], 2)} &
+\Sexpr{round(alt_mod_sv[3], 2)} & 
+\Sexpr{round(alt_mod2_sv[3], 2)}
 \\
-Reciprocity & $\beta_2$ & $s_{i2}(x) = \sum\limits_{j=1}^n x_{ij}x_{ji}$ & \Sexpr{round(subset(means, Model=="M1" & parameter=="beta2")$mean, 2)} &
-\Sexpr{round(subset(means, Model=="M2" & parameter=="beta2")$mean, 2)}
+Reciprocity & $\beta_2$ & $s_{i2}(x) = \sum\limits_{j=1}^n x_{ij}x_{ji}$ & \Sexpr{round(null_mod_sv[4], 2)} &
+\Sexpr{round(alt_mod_sv[4], 2)} & 
+\Sexpr{round(alt_mod2_sv[4], 2)}
 \\
-Jumping Transitive Triplets & $\beta_3$ & $s_{i3}(x) = \sum\limits_{\forall j\neq h} x_{ij}x_{ih}x_{hj} \mathbb{I}(v_i = v_h \neq v_j)$ & -- & \Sexpr{round(subset(means, Model=="M2" & parameter=="beta3")$mean, 2)}
+Jumping Transitive Triplets & $\beta_3$ & $s_{i3}(x) = \sum\limits_{\forall j\neq h} x_{ij}x_{ih}x_{hj} \mathbb{I}(v_i = v_h \neq v_j)$ & -- & 
+\Sexpr{round(alt_mod_sv[5], 2)} & -- \\
+\# doubly achieved distances & $\beta_4$ & $s_{i4}(x) = |\{j : x_{ij} = 0, \sum\limits_h x_{ih}x_{hj} \geq 2\}|$ & -- & -- &
+\Sexpr{round(alt_mod2_sv[5], 2)}
 \end{tabular}}
@@
   f^{M_1}_{i}(x) & = \hat{\beta}_1^{M_1}s_{i1}(x) +  \hat{\beta}_2^{M_1}s_{i2}(x) \\
-  f^{M_2}_{i}(x) & = \hat{\beta}_1^{M_2}s_{i1}(x) +  \hat{\beta}_2^{M_2}s_{i2}(x) + \hat{\beta}_3^{M_2}s_{i3}(x) 
+  f^{M_2}_{i}(x) & = \hat{\beta}_1^{M_2}s_{i1}(x) +  \hat{\beta}_2^{M_2}s_{i2}(x) + \hat{\beta}_3^{M_2}s_{i3}(x) \\
+  f^{M_3}_{i}(x) & = \hat{\beta}_1^{M_3}s_{i1}(x) +  \hat{\beta}_2^{M_3}s_{i2}(x) + \hat{\beta}_4^{M_3}s_{i4}(x) 
   \end{align*}
@@
 \hh{XXX Sam, when you are creating these lineups, could you split the functions into two, and create and save the data into a csv file in a first function, and then in a second function get the data and put it into a network?
-- That way we don't have to create the data again, if/when we change the design.}
+- That way we don't have to create the data again, if/when we change the design.} \st{done!}
 
@@
 
-\section{Model X versus Y lineup size 3}
+\section{Model $M_1$ versus $M_2$ lineup size 3}
 \newcounter{k}
@@
 
-\section{Model X versus Y lineup size 6}
+\section{Model $M_1$ versus $M_2$ lineup size 6}
 \newcounter{k1}
@@
 
-\section{Model X versus Y lineup size 9}
+\section{Model $M_1$ versus $M_2$ lineup size 9}
 \newcounter{k2}
@@
 
-\section{Model X versus Y lineup size 12}
+\section{Model $M_1$ versus $M_2$ lineup size 12}
 \newcounter{k3}
@@
 
-\section{Model X versus Y lineup size 16}
+\section{Model $M_1$ versus $M_2$ lineup size 16}
 \newcounter{k4}
@@
 
-\section{Model Y versus X lineup size 3}
+\section{Model $M_2$ versus $M_1$ lineup size 3}
 \newcounter{k5}
@@
 
-\section{Model Y versus X lineup size 6}
+\section{Model $M_2$ versus $M_1$ lineup size 6}
 \newcounter{k6}
@@
 
-\section{Model Y versus X lineup size 9}
+\section{Model $M_2$ versus $M_1$ lineup size 9}
 \newcounter{k7}
@@
 
-\section{Model Y versus X lineup size 12}
+\section{Model $M_2$ versus $M_1$ lineup size 12}
 \newcounter{k8}
@@
 
-\section{Model Y versus X lineup size 16}
+\section{Model $M_2$ versus $M_1$ lineup size 16}
 \newcounter{k9}

### 2016-04-19 lineups.Rnw
add 3rd model plots. pdf very long!

@@
 \newcounter{k5}
-\forloop{k5}{1}{\value{k5} < 21}{smallfriendsrev-m-3-rep-\arabic{k5}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriendsrev-m-3-rep-\arabic{k5}} \\ }
+\forloop{k5}{1}{\value{k5} < 21}{smallfriends-rev-m-3-rep-\arabic{k5}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-rev-m-3-rep-\arabic{k5}} \\ }
 
@@
 \newcounter{k6}
-\forloop{k6}{1}{\value{k6} < 21}{smallfriendsrev-m-6-rep-\arabic{k6}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriendsrev-m-6-rep-\arabic{k6}} \\ }
+\forloop{k6}{1}{\value{k6} < 21}{smallfriends-rev-m-6-rep-\arabic{k6}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-rev-m-6-rep-\arabic{k6}} \\ }
 
@@
 \newcounter{k7}
-\forloop{k7}{1}{\value{k7} < 21}{smallfriendsrev-m-9-rep-\arabic{k7}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriendsrev-m-9-rep-\arabic{k7}} \\ }
+\forloop{k7}{1}{\value{k7} < 21}{smallfriends-rev-m-9-rep-\arabic{k7}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-rev-m-9-rep-\arabic{k7}} \\ }
 
@@
 \newcounter{k8}
-\forloop{k8}{1}{\value{k8} < 21}{smallfriendsrev-m-12-rep-\arabic{k8}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriendsrev-m-12-rep-\arabic{k8}} \\ }
+\forloop{k8}{1}{\value{k8} < 21}{smallfriends-rev-m-12-rep-\arabic{k8}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-rev-m-12-rep-\arabic{k8}} \\ }
 
@@
 \newcounter{k9}
-\forloop{k9}{1}{\value{k9} < 21}{smallfriendsrev-m-16-rep-\arabic{k9}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriendsrev-m-16-rep-\arabic{k9}} \\ }
+\forloop{k9}{1}{\value{k9} < 21}{smallfriends-rev-m-16-rep-\arabic{k9}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-rev-m-16-rep-\arabic{k9}} \\ }
+
+\section{Model $M_1$ versus $M_3$ lineup size 3}
+\newcounter{k10}
+\forloop{k10}{1}{\value{k10} < 21}{smallfriends-eff2-m-3-rep-\arabic{k10}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-3-rep-\arabic{k10}} \\ }
+
+\section{Model $M_1$ versus $M_3$ lineup size 6}
+\newcounter{k11}
+\forloop{k11}{1}{\value{k11} < 21}{smallfriends-eff2-m-6-rep-\arabic{k11}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-6-rep-\arabic{k11}} \\ }
+
+\section{Model $M_1$ versus $M_3$ lineup size 9}
+\newcounter{k12}
+\forloop{k12}{1}{\value{k12} < 21}{smallfriends-eff2-m-9-rep-\arabic{k12}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-9-rep-\arabic{k12}} \\ }
+
+\section{Model $M_1$ versus $M_3$ lineup size 12}
+\newcounter{k13}
+\forloop{k13}{1}{\value{k13} < 21}{smallfriends-eff2-m-12-rep-\arabic{k13}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-12-rep-\arabic{k13}} \\ }
+
+\section{Model $M_1$ versus $M_3$ lineup size 16}
+\newcounter{k14}
+\forloop{k14}{1}{\value{k14} < 21}{smallfriends-eff2-m-16-rep-\arabic{k14}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-16-rep-\arabic{k14}} \\ }
+
+\section{Model $M_3$ versus $M_1$ lineup size 3}
+\newcounter{k15}
+\forloop{k15}{1}{\value{k15} < 21}{smallfriends-eff2-rev-m-3-rep-\arabic{k15}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-3-rep-\arabic{k15}} \\ }
+
+\section{Model $M_3$ versus $M_1$ lineup size 6}
+\newcounter{k16}
+\forloop{k16}{1}{\value{k16} < 21}{smallfriends-eff2-rev-m-6-rep-\arabic{k16}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-6-rep-\arabic{k16}} \\ }
+
+\section{Model $M_3$ versus $M_1$ lineup size 9}
+\newcounter{k17}
+\forloop{k17}{1}{\value{k17} < 21}{smallfriends-eff2-rev-m-9-rep-\arabic{k17}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-9-rep-\arabic{k17}} \\ }
+
+\section{Model $M_3$ versus $M_1$ lineup size 12}
+\newcounter{k18}
+\forloop{k18}{1}{\value{k18} < 21}{smallfriends-eff2-rev-m-12-rep-\arabic{k18}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-12-rep-\arabic{k18}} \\ }
+
+\section{Model $M_3$ versus $M_1$ lineup size 16}
+\newcounter{k19}
+\forloop{k19}{1}{\value{k19} < 21}{smallfriends-eff2-rev-m-16-rep-\arabic{k19}\newline
+\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-16-rep-\arabic{k19}} \\ }
+
 \end{document}

### 2016-04-20 lineups.Rnw
add third model

@@
 # 
+# alt2nd <- read.csv("Data/distribution_dblpairs_model.csv")
+# names(alt2nd)[-1] <- c("alpha1", "alpha2", "beta1", "beta2", "beta4")
+# alt2nd$Sim <- 1:nrow(alt2nd)
+# alt2nd$Model <- "M3"
+# 
 # alt2 <- gather(alt, parameter, estimate, 2:6)
 # null2 <- gather(nulls, parameter, estimate, 2:5)
+# alt2nd2 <- gather(alt2nd, parameter, estimate, 2:6)
 # 
@@
 
-simu <- read.csv("Data/simulation-1000-M1-M2.csv")
-means <- simu %>% group_by(Model, parameter) %>% summarize(
+simu2 <- read.csv("Data/simulation-1000-M1-M2-M3.csv")
+means <- simu2 %>% group_by(Model, parameter) %>% summarize(
   mean = mean(estimate)
@@
 <<hist-estimates, dependson='read-simu', echo=FALSE, fig.width=8, fig.height=5, out.width='.8\\linewidth'>>=
-qplot(data=simu, estimate, fill=Model, geom="density", alpha=I(0.75)) + facet_wrap(~parameter, scales="free") + theme_bw()
+qplot(data=simu2, estimate, fill=Model, geom="density", alpha=I(0.65)) + facet_wrap(~parameter, scales="free") + theme_bw()
 @
@@
 Model parameter $\beta_3$ for jumping transitive triplets in model $M_2$ is significantly different from zero, but its inclusion also leads to significant changes in the  other model parameters of model $M_1$.
-}
+The parameter $\beta_4$ for doubly acheived distances is also significantly different from zero, but has larger variance.
+The inclusion of $\beta_4$ also changes the estimates of the other model parameters, but not as much as the inclusion of $\beta_3$.}
 \end{figure}

### 2016-04-20 lineups.Rnw
add more model description

@@
 <<readsimu, echo=FALSE>>=
-# nulls <- read.csv("Data/distribution_null_model.csv")
+ nulls <- read.csv("Data/distribution_null_model.csv")
 # names(nulls)[-1] <- c("alpha1", "alpha2", "beta1", "beta2")
@@
 # 
-# alt <- read.csv("Data/distribution_jumpTT_model.csv")
+ alt <- read.csv("Data/distribution_jumpTT_model.csv")
 # names(alt)[-1] <- c("alpha1", "alpha2", "beta1", "beta2", "beta3")
@@
 # 
-# alt2nd <- read.csv("Data/distribution_dblpairs_model.csv")
+ alt2nd <- read.csv("Data/distribution_dblpairs_model.csv")
 # names(alt2nd)[-1] <- c("alpha1", "alpha2", "beta1", "beta2", "beta4")
@@
 library(dplyr)
-null_model_dist <- read.csv("Data/distribution_null_model.csv")
-null_model_dist <- null_model_dist[,-1]
-alt_model_dist <- read.csv("Data/distribution_jumpTT_model.csv")
-alt_model_dist <- alt_model_dist[,-1]
-alt_model2_dist <- read.csv("Data/distribution_dblpairs_model.csv")
-alt_model2_dist <- alt_model2_dist[,-1]
-
-null_mod_sv <- as.numeric(null_model_dist %>% summarise_each(funs(mean)))
-alt_mod_sv <- as.numeric(alt_model_dist %>% summarise_each(funs(mean)))
-alt_mod2_sv <- as.numeric(alt_model2_dist %>% summarise_each(funs(mean)))
+null_mod_sv <- as.numeric(nulls[,-c(1,6,7)] %>% summarise_each(funs(mean)))
+alt_mod_sv <- as.numeric(alt[,-c(1, 7,8)] %>% summarise_each(funs(mean)))
+alt_mod2_sv <- as.numeric(alt2nd[,-c(1,7,8)] %>% summarise_each(funs(mean)))
 @
@@
 \begin{table}[h]
-\caption{\label{tab:models}Parameters and estimates of models $M_1$ and $M_2$.
-\hh{Estimates stem from one of the simulations.
-XXX We probably should change that to the average of the 1000 simulations, and then simulate the lineups from those, like you suggested.
-}}
+\caption{\label{tab:models}Parameters and estimates of models $M_1$, $M_2$, and $M_3$.
+Estimates are the mean of 1000 iterations of the model estimates.
+The lineups that follow are simulated from models using these values.}
 \centering
@@
 The rate parameters, $\alpha_1$ and $\alpha_2$ represent how many opportunities for change each actor gets when moving from wave 1 to 2 and from wave 2 to 3, respectively.
+The outdegree parameter, $\beta_1$, represents how likely an actor is to change outgoing ties.
+If the estimate, $\hat{\beta}_1$, is positive, the actor is more likely to create outgoing ties, while a negative estimate leads the actor to deleting outgoing ties.
+This effect is highly correlated with the reciprocity parameter, $\beta_2$.
+A negative estimate of this parameter implies that the actor is discouraged from reciprocating its incoming ties, while a positive estimate implies that the actor is encouraged to reciprocate all ties.
+The additional parameter in $M_2$, $\beta_3$, is a covariate parameter.
+The covariate in this model is the drinking behaviour of the 16 students in our data.
+The possible values are 1, 2, 3, and 4, which means the student drinks never, once or twice a year, once a month, or once a week.
+This jumping transitive triplet effect impacts the transitive closure of actors from different groups.
+Thus, a positive estimate encourages transitive closure when one of three actors is in a different covariate group than the other two, while a negative estimate discourages this behavior.
+An example of this type of closure is given in Figure~\ref{fig:jtt}.
+Finally, the doubly acheived distances effect is defined by the number of actors to whom actor $i$ is not directly tied, and tied through twopaths via at least two intermediaries.
+This is a structural effect, like the density and repciprocity effects.
+A positive value encourages indirect ties, while a negative value discourages the formation of indirect ties.
 
-\hh{XXX Sam, when you are creating these lineups, could you split the functions into two, and create and save the data into a csv file in a first function, and then in a second function get the data and put it into a network?
-- That way we don't have to create the data again, if/when we change the design.} \st{done!}
+<<jtt, echo=FALSE, fig.width=2, fig.height=2, message = FALSE, warning = FALSE, fig.align='center', fig.cap="A realization of a jumping transitive triplet, where $i$ is the focal actor, $j$ is the target actor, and $h$ is the intermediary.
+The group of the actors is represented by the shape of the node.">>=
+jTTe <- data.frame(from = c('i', 'i', 'h'), to = c('h', 'j', 'j'))
+jTTn <- data.frame(id = letters[8:10], group = c(1,1,2))
 
-\hh{XXX question for the meeting: right now the difference between models X and Y are the Jumping Transitive Triplets, which we might (or not) see in the graphs.
-Is there another effect that is significant at this stage as well?
-It would be good to have the same kind of lineups for a different effect.
-}
+jTT <- merge(jTTe, jTTn, by.x = 'from', by.y = "id", all = T)
 
-\hh{XXX another question for tomorrow: in the first four reps of the reverse model it looks like one graph is more complex than the other two models ...
-which is not what I'd expect from this scenario, because ehere, we are looking for the simplest model, right?
-}
-
-\hh{XXX would a size 15 (= 5x3) lineup be more `square' than a size 16 one? }
+set.seed(12345)
+ggplot(data = jTT, aes(from_id = from, to_id = to)) + 
+  geom_net(aes(shape = as.factor(group)), directed = T, label = T, labelcolour='blue',vjust = -1, hjust =2, arrowgap = .05, colour = 'black') + 
+  theme_net() + theme(legend.position = "none")
+@
 

### 2016-04-27 lineups.Rnw
done in purrr. figuring out how to extract estimates now

@@
 
-set.seed(12345)
+set.seed(12345) # make shapes bigger, put labels in center. 
 ggplot(data = jTT, aes(from_id = from, to_id = to)) + 

### 2016-04-27 Code/ests_from_lineups.R

@@ -1,0 +1,140 @@
+# Read in lineup data and re-fit the model and store the estimates
+# in function: null_mod_eff is the effects structure w/4 effects and 
+# alt_mod_eff is the effects structure w/5 effects (the 4 + additional)
+# give lineupname either "smallfriends", "smallfriends-rev", "smallfriends-eff2" or "smallfriends-eff2-rev" 
+
+#TO DO:
+# 1. get original first wave of data + alcohol covariate
+# 2. create siena data object for each lineup data plot (M data objs)
+# 3. fit appropriate siena model to each of the M data objects
+# 4. store effects in a purrr data frame
+
+lineups <- expand.grid(lineupname = c("smallfriends", "smallfriends-rev", "smallfriends-eff2", "smallfriends-eff2-rev" ),
+                       M = c(3,6,9,12,16), 
+                       rep = 1:20)
+friend.data.w1 <- as.matrix(read.table("Data/s50_data/s50-network1.dat"))
+friend.data.w3 <- as.matrix(read.table("Data/s50_data/s50-network3.dat"))
+fd2.w1 <- friend.data.w1[20:35,20:35]
+fd2.w3 <- friend.data.w3[20:35,20:35]
+
+drink <- as.matrix(read.table("Data/s50_data/s50-alcohol.dat"))
+drink2 <- drink[20:35,1]
+alcohol2 <- coCovar( as.vector(drink2) )
+mysmalldata <- sienaDataCreate( friend2, alcohol2)
+
+estimates_from_lineup <- function(dat, alt_eff_name, covariate = "alcohol2"){
+  lineupname <- dat$lineupname
+  M <- dat$M
+  rep <- dat$rep
+  require(RSiena)
+  require(dplyr)
+#   if (length(grep("rev", lineupname)) == 0){
+#     null_model_effects <- null_mod_eff # which effects to use for estimation. 
+#     # "null" means the model with M-1 plots in lineup, while "alt" is the "test plot" 
+#     alt_model_effects <- alt_mod_eff
+#   } else{
+#     null_model_effects <- alt_mod_eff # for the "rev" lineups
+#     alt_model_effects <- null_mod_eff
+#   }
+  
+  # model setup
+  myalgorithm_null <- sienaAlgorithmCreate( projname = 's50null' , n3 = 1000)
+  myalgorithm_alt <- sienaAlgorithmCreate( projname = 's50null' , n3 = 20000)
[…]

### 2016-04-27 Code/ests_from_lineups.R
all estimates gathered by parameter type.

@@ -130,11 +130,75 @@ nest_lu <- lineups %>% arrange(lineupname, M, rep) %>% nest(-lineupid)
 lineups %>% arrange(lineupname, M, rep) %>% nest(-lineupid) %>%
   mutate(fits = map(data, safely(estimates_from_lineup),alt_eff_name = "jumpXTransTrip")) -> lu_ests_eric
 
+#lu_ests_eric <- readRDS("Data/ests_from_lineup.RDS")
+
 lu_ests_eric %>% unnest(data) -> unnest_ests 
 library(plyr)
 
 lu_ests_eric %>% unnest(data) %>% 
-  select(fits) %>% flatten()
+  select(fits) %>% flatten() -> listRes 
 
 
-for (i in 1:nrow(unnest_ests))
+erridx <- NULL
+for (i in 1:length(listRes)){
+  if(!is.null(listRes[[i]]$error)){
+    erridx <- c(erridx, i)
+  } else {erridx <- erridx}
+}
+
+erridx
+
+listRes[-erridx,] 
+
+allres <- NULL
+
+for (i in (1:length(listRes))[-erridx]){
+  resi <- listRes[[i]]$result
+  resi$lineupid <- unnest_ests$lineupid[i]
+  resi$lineupname <- unnest_ests$lineupname[i]
+  resi$M <- unnest_ests$M[i]
+  resi$rep <- unnest_ests$rep[i]
+  resi$idx <- i
+  allres <- rbind(allres, resi)
+  print(i)
+}
+
+head(allres)
+
+if (allres$M[i] == allres$plot[i]){
+  
+}
+
+allres %>% group_by(lineupid, plot) %>% arrange(lineupname, M, rep)
+
+allres$effectname <- 1:nrow(allres) 
+
+for (i in unique(allres$idx)){ 
+  idx2 <- which(allres$idx == i)
+  if (unique(allres$lineupname[idx2]) == 'smallfriends'){
+    allres[idx2,]$effectname <- c(rep(c("alpha1", "beta1", "beta2"), (unique(allres$M[idx2])-1)), 
+                                  c("alpha1", "beta1", "beta2",'beta3'))
+  } else if (unique(allres$lineupname[idx2]) == 'smallfriends-rev'){
+    allres[idx2,]$effectname <- c(rep(c("alpha1", "beta1", "beta2", 'beta3'), (unique(allres$M[idx2])-1)), 
+                                  c("alpha1", "beta1", "beta2"))
[…]

### 2016-04-27 Code/ests_from_lineups.R
make plots like the density estimate plots

@@ -191,14 +191,50 @@ for (i in unique(allres$idx)){
   print(i)
 }     
 
+allres$model <- 1:nrow(allres)
+
+for (i in unique(allres$idx)){ 
+  idx2 <- which(allres$idx == i)
+  if (unique(allres$lineupname[idx2]) == 'smallfriends'){
+    allres[idx2,]$model <- c(rep("M1", 3*(unique(allres$M[idx2])-1)), rep("M2", 4))
+  } else if (unique(allres$lineupname[idx2]) == 'smallfriends-rev'){
+    allres[idx2,]$model <- c(rep("M2", 4*(unique(allres$M[idx2])-1)),rep("M1", 3))
+  } else if (unique(allres$lineupname[idx2]) == 'smallfriends-eff2'){
+    allres[idx2,]$model <- c(rep("M1", 3*(unique(allres$M[idx2])-1)), 
+                                  rep("M3", 4))
+  } else{
+    allres[idx2,]$model <- c(rep("M3", 4*(unique(allres$M[idx2])-1)), 
+                                  rep("M1",3))
+  } 
+  print(i)
+} 
 head(allres)
 
 ggplot(data = allres, aes(x = idx, y = effects, color = plot)) +
   geom_point() + facet_wrap(~effectname, scales = 'free')
 
 ggplot(data = allres, aes(x = M, y = effects, color = plot, fill = plot, group = M)) +
-  geom_boxplot() + facet_wrap(~effectname, scales = 'free')
+  geom_boxplot() + ylim(c(-15,15)) +  facet_wrap(~effectname, scales = 'free')
+
+#very interesting
+ggplot(data = allres, aes(x = effects)) +
+  geom_density(aes(fill = model), alpha = .4) + xlim(c(-10,10)) + 
+  facet_wrap(~effectname, scales = 'free')  
+
+ggplot()+ 
+geom_density(data = effects_dist, 
+             aes(x = estimate, fill = model), alpha = .4) + 
+  facet_wrap(~effectname, scales = 'free')  
+
+
+effects_dist <- read.csv("Data/simulation-1000-M1-M2-M3.csv")
+names(effects_dist)[3] <- "effectname"
+names(effects_dist)[2] <- 'model'
+ggplot() + 
+  geom_density(data = effects_dist, aes(x = estimate, fill = Model), alpha = .4)+
+  facet_wrap(~effectname, scales = 'free')
   
+
 
[…]

### 2016-04-27 Code/ests_from_lineups.R
fix typo

@@ -234,7 +234,3 @@ ggplot() +
   geom_density(data = effects_dist, aes(x = estimate, fill = Model), alpha = .4)+
   facet_wrap(~effectname, scales = 'free')
   
-
-
-idx2 <- which(allres$idx == 1)
-allres[idx2,]  c(rep(c("alpha1", "beta1", "beta2"), (unique(allres$M[idx2])-1)), c("alpha1", "beta1", "beta2",'beta3'))

### 2016-04-29 Code/ests_from_lineups.R
complete the estimation from lineups and compile results of experiment

@@ -22,10 +22,15 @@ drink2 <- drink[20:35,1]
 alcohol2 <- coCovar( as.vector(drink2) )
 mysmalldata <- sienaDataCreate( friend2, alcohol2)
 
+#get starting values for the simulations from the distributions
+effects_dist <- read.csv("Data/simulation-1000-M1-M2-M3.csv")
+effects_dist %>% group_by(model, effectname) %>% dplyr::summarise(bhat = mean(estimate))
+
 estimates_from_lineup <- function(dat, alt_eff_name, covariate = "alcohol2"){
   lineupname <- dat$lineupname
   M <- dat$M
   rep <- dat$rep
+  
   require(RSiena)
   require(dplyr)
 #   if (length(grep("rev", lineupname)) == 0){

### 2016-04-29 Code/ests_from_lineups2.R

@@ -1,0 +1,101 @@
+effects_dist <- read.csv("Data/simulation-1000-M1-M2-M3.csv")
+effects_dist %>% group_by(model, effectname) %>% 
+  dplyr::summarise(bhat = mean(estimate)) %>%
+  filter(effectname !="alpha2") -> starting_values
+
+lineups <- expand.grid(lineupname = c("smallfriends", "smallfriends-rev", "smallfriends-eff2", "smallfriends-eff2-rev" ),
+                       M = c(3,6,9,12,16), 
+                       rep = 1:20)
+friend.data.w1 <- as.matrix(read.table("Data/s50_data/s50-network1.dat"))
+friend.data.w3 <- as.matrix(read.table("Data/s50_data/s50-network3.dat"))
+fd2.w1 <- friend.data.w1[20:35,20:35]
+fd2.w3 <- friend.data.w3[20:35,20:35]
+
+as.numeric(data.frame(starting_values %>% filter(model == "M1") %>% ungroup() %>%
+  select_("bhat"))[,1]) -> sv_M1
+as.numeric(data.frame(starting_values %>% filter(model == "M2") %>% ungroup() %>%
+                        select_("bhat"))[,1]) -> sv_M2
+as.numeric(data.frame(starting_values %>% filter(model == "M3") %>% ungroup() %>%
+                        select_("bhat"))[,1]) -> sv_M3
+
+estimates_from_lineup2 <- function(dat){
+  lineupname <- dat$lineupname
+  M <- dat$M
+  rep <- dat$rep
+  null_model_name <- "M1"
+  null_model_eff_names <- c('alpha1', 'beta1', 'beta2')
+  if (length(grep("eff2", lineupname)) == 0){
+    covariate <- "alcohol2"
+    alt_eff_name <- "jumpXTransTrip"
+    inits <- sv_M2
+    alt_model_name <- "M2"
+    alt_model_eff_names <- c("alpha1", 'beta1', 'beta2', 'beta3')
+  } else{
+    covariate <- ""
+    alt_eff_name <- "nbrDist2twice"
+    inits <- sv_M3
+    alt_model_name <- "M3"
+    alt_model_eff_names <- c("alpha1", 'beta1', 'beta2', 'beta4')
+  }
+  require(RSiena)
+  require(dplyr)
+  
+  # model setup
+  myalgorithm <- sienaAlgorithmCreate( projname = 's50null' , n3 = 5000)
[…]

### 2016-04-29 Code/small_friends.R

@@ -51,7 +51,7 @@ mysmalldata <- sienaDataCreate( friend2, alcohol2)
 #test_model_eff2 <- includeEffects( null_model_eff2, egoX, altX, egoXaltX, interaction1 = "alcohol2")
 #test_model_eff2 <- includeEffects( null_model_eff2, sameXTransTrip, interaction1 = "alcohol2")
 #myalgorithm2 <- sienaAlgorithmCreate( projname = 's50' , n3 = 1000)
-#ests_null <- siena07( myalgorithm2, data = mysmalldata, returnDeps = TRUE, effects = null_model_eff2, batch=TRUE, verbose = FALSE)
+ests_null <- siena07( myalgorithm2, data = mysmalldata, returnDeps = TRUE, effects = null_model_eff2, batch=TRUE, verbose = FALSE, silent = TRUE)
 #ests_test <- siena07( myalgorithm2, data = mysmalldata, returnDeps = TRUE, effects = test_model_eff2, batch=TRUE, verbose = FALSE)
 
 # pull out the last 10 networks for periods 1 (wave 1->2) and 2 (wave 2->3)

### 2016-04-29 GGExperimentApr28/GGpresentationApr28.Rmd
add mini-experiment materials

---
title: "Network Lineups"
author: "Sam Tyner"
date: "April 28, 2016"
output: pdf_document
---

```{r, global_options, include=FALSE}
    library(knitr)
    library(formatR)
    opts_chunk$set(
      echo=FALSE,
      warning=FALSE,
      message=FALSE,
      fig.width = 8,
      fig.align = 'left', cache = FALSE)
```

## Intro

For each of the following lineups, write down the number of the panel that contains the graph you feel least resembles the graphs in the other panels.

For example, for the lineup below, you should pick 2.

```{r,echo = FALSE, fig.height=3}
datx <- data.frame(x = rnorm(30,2), y = rnorm(30,2), plot = 1)
daty <- data.frame(x = rnorm(30,2), y = rnorm(30,2), plot = 3)
datz <- data.frame(x = rnorm(30, 2), plot = 2)
datz$y = datz$x + rnorm(30, sd = .1)
dattt <- rbind(datx, daty, datz)
library(ggplot2)
ggplot(data = dattt) + geom_point(aes(x = x, y= y)) + theme(aspect.ratio = 1) + facet_wrap(~plot)
```

\newpage

```{r, echo=FALSE, warning=FALSE, message=FALSE}
setwd("~/Desktop/NetworksResearch/NetworksVizInference/GGExperimentApr28/")
for_gg <- data.frame(lineupname = rep(c("smallfriends", "smallfriends-rev", 
                                        "smallfriends-eff2", "smallfriends-eff2-rev"), each = 5) ,
                     M = rep(c(3,6,9,12,16), 4))
for_gg$rep <- c(17,6,8,4,16, 7,18,3,14,3, 10,5,7,11,4, 4,17,5,7,14)

for_gg$id <- 1:20

set.seed(138818)
for_gg$pres_order <- sample(LETTERS[1:20])

for_gg$filename <- paste0(for_gg$lineupname, "-m-", for_gg$M, "-rep-", for_gg$rep, ".csv")

# setwd("~/Desktop/NetworksResearch/NetworksVizInference/Data/lineupdata/")
# for (i in 1:nrow(for_gg)){
#   assign(x = paste0("dat",for_gg$pres_order[i]), value = read.csv(for_gg$filename[i]))  
# }
# for_gg$cat <- "CAT"
# for_gg$cat[c(1,5,9,13,17)] <- "noa_noc"
# for_gg$cat[(c(1,5,9,13,17) + 1)] <- "noa_c"
# for_gg$cat[(c(1,5,9,13,17) + 2)] <- "a_c"
# for_gg$cat[(c(1,5,9,13,17) + 3)] <- "a_noc"
# 
# all_dat <- NULL
# for(i in 1:nrow(for_gg)){
#   all_dat[[i]] <- read.csv(for_gg$filename[i])
# }
# 
# drink <- as.matrix(read.table("../s50_data/s50-alcohol.dat"))[,2]
# 
# dr_beh <- data.frame(friendid = 1:16, drink = drink[20:35])
# dr_beh$covariate <- "AH"
# dr_beh$covariate[which(dr_beh$drink %in% c(1,2))] <- "level A"
# dr_beh$covariate[which(dr_beh$drink %in% c(3,4))] <- "level B"

# library(ggplot2)
# library(geomnet)
# gg_plots <- NULL
# for (i in 1:length(all_dat)){
#     if (for_gg$cat[i] == "noa_noc"){
#       gg_plots[[i]] <- ggplot(data = all_dat[[i]], 
#                               aes(from_id = X1, to_id = X2)) +
#         geom_net(fiteach = T, size = 1, linewidth = .5) + 
#         facet_wrap(~plot_order) + theme_net() + 
#         theme(panel.background=element_rect(colour = 'black'))
#     } else if(for_gg$cat[i] == "a_noc"){
#       gg_plots[[i]] <- ggplot(data = all_dat[[i]], 
#                               aes(from_id = X1, to_id = X2)) +
#         geom_net(fiteach = T, size = 1, directed = T,  linewidth = .5, arrowsize = .1,
#                  arrowgap = .01) + 
#         facet_wrap(~plot_order) + theme_net() + 
#         theme(panel.background=element_rect(colour = 'black'))
#     } else if(for_gg$cat[i] == "noa_c"){
#       dat <- all_dat[[i]]
#       dat2 <- merge(dat, dr_beh, by.x = "X1", by.y = "friendid")
#       gg_plots[[i]] <- ggplot(data = dat2, aes(from_id = X1, to_id = X2)) + 
#         geom_net(fiteach = T, size = 1, linewidth = .5, aes(colour = covariate)) + 
#         facet_wrap(~plot_order) + theme_net() + 
#         theme(panel.background=element_rect(colour = 'black'))
#     } else{
#       dat <- all_dat[[i]]
#       dat2 <- merge(dat, dr_beh, by.x = "X1", by.y = "friendid")
#       gg_plots[[i]] <- ggplot(data = dat2, aes(from_id = X1, to_id = X2)) + 
#         geom_net(fiteach = T, size = 1, directed = T, arrowsize = .1, linewidth = .5,
#                  arrowgap = .01, aes(colour = covariate)) + 
#         facet_wrap(~plot_order) + theme_net() + 
#         theme(panel.background=element_rect(colour = 'black'))
#     }
# }
# savepdf <- function(file, width=16, height=10){
#   fname <- paste(file,".pdf",sep="")
#   pdf(fname, width=width/2.54, height=height/2.54,
#       pointsize=10)
#   par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(3.3,3.5,1.1,1))
# }
# for (i in 1:20){
#   savepdf(file = paste(for_gg$pres_order[i], sub(".csv", "", for_gg$filename[i])), 
#           width = 8*2.54, height = 8*2.54)
#   print(gg_plots[[i]])
#   dev.off()
# }
```

## A
![]("A smallfriends-eff2-rev-m-16-rep-14")
\newpage

## B
![]("B smallfriends-rev-m-6-rep-18")
\newpage

## C
![]("C smallfriends-eff2-m-3-rep-10")
\newpage

## D
![]("D smallfriends-eff2-rev-m-3-rep-4")
\newpage

## E
![]("E smallfriends-eff2-rev-m-6-rep-17")
\newpage

## F
![]("F smallfriends-rev-m-3-rep-7")
\newpage

## G
![]("G smallfriends-rev-m-16-rep-3")
\newpage

## H
![]("H smallfriends-eff2-m-16-rep-4")
\newpage

## I
![]("I smallfriends-rev-m-12-rep-14")
\newpage

## J
![]("J smallfriends-m-16-rep-16")
\newpage

## K
![]("K smallfriends-eff2-m-6-rep-5")
\newpage

## L
![]("L smallfriends-m-12-rep-4")
\newpage

## M
![]("M smallfriends-m-3-rep-17")
\newpage

## N
![]("N smallfriends-m-9-rep-8")
\newpage

## O
![]("O smallfriends-eff2-m-9-rep-7")
\newpage

## P
![]("P smallfriends-m-6-rep-6")
\newpage

## Q
![]("Q smallfriends-rev-m-9-rep-3")
\newpage

## R
![]("R smallfriends-eff2-rev-m-9-rep-5")
\newpage

## S
![]("S smallfriends-eff2-m-12-rep-11")
\newpage

## T
![]("T smallfriends-eff2-rev-m-12-rep-7")
\newpage

### 2016-04-29 GGExperimentApr28/experiment_res_analysis.R

@@ -1,0 +1,21 @@
+# analyzing results of mini-experiment 
+
+setwd("GGExperimentApr28")
+
+library(readxl)
+library(dplyr)
+library(magrittr)
+response <- read_excel("responses_GGExpApr28.xlsx")
+response %<>% mutate(correct = as.numeric(Answer == ChosenLU))
+response %>% group_by(Respondent) %>% 
+  dplyr::summarise(perc_right = sum(na.omit(correct))/20)
+
+library(ggplot2)
+
+ggplot(data = response, aes(x = as.factor(ChosenLU), fill = as.factor(correct))) + 
+  geom_bar() +
+  facet_wrap(~Lineup, scales = 'free')
+
+response %>% group_by(Lineup) %>%
+  dplyr::summarise(perc_guessed = sum(na.omit(correct))/11, tot_resp = length(unique(ChosenLU)))
+

### 2016-05-05 lineups.Rnw
shape and size of JTT example

--- 2016-04-20 lineups.Rnw
@@
 ggplot(data = jTT, aes(from_id = from, to_id = to)) + 
-  geom_net(aes(shape = as.factor(group)), directed = T, label = T, labelcolour='blue',vjust = -1, hjust =2, arrowgap = .05, colour = 'black') + 
-  theme_net() + theme(legend.position = "none")
+  geom_net(aes(shape = as.factor(group)), directed = T, label = T, labelcolour='grey80',vjust = 0, hjust =0, arrowgap = .15, colour = 'black', size=10) + 
+  expand_limits(x=c(-0.1,1.1), y=c(-0.1,1.1)) +
+  theme_net() +
+  theme(legend.position = "none")
 @

### 2016-05-05 lineups.Rnw
structural sketch for doubly achieved distance

@@
 \usepackage{amsmath,amssymb,amsfonts, mathrsfs,fancyhdr,syntonly,lastpage,hyperref,enumitem,graphicx, lscape, forloop, url}
+\usepackage[font=small,skip=5pt]{caption}
+\usepackage{subcaption}
 
@@
 An example of this type of closure is given in Figure~\ref{fig:jtt}.
-Finally, the doubly acheived distances effect is defined by the number of actors to whom actor $i$ is not directly tied, and tied through twopaths via at least two intermediaries.
-This is a structural effect, like the density and repciprocity effects.
+\hh{XXX Could you color the edge that forms the closer?
+That would be an edge of the form 'I become a friend of my friend's friend', right?
+XXX With the directed edges we also distinguish between 'i likes j' and 'j likes i'.
+} Finally, the doubly achieved distances effect is defined by the number of actors to whom actor $i$ is not directly tied, and tied through two paths via at least two intermediaries \hh{XXX could you give an example for that, too?}.
+This is a structural effect, like the density and reciprocity effects.
 A positive value encourages indirect ties, while a negative value discourages the formation of indirect ties.
 
-<<jtt, echo=FALSE, fig.width=2, fig.height=2, message = FALSE, warning = FALSE, fig.align='center', fig.cap="A realization of a jumping transitive triplet, where $i$ is the focal actor, $j$ is the target actor, and $h$ is the intermediary.
-The group of the actors is represented by the shape of the node.">>=
+\begin{figure}
+\begin{subfigure}[t]{.45\textwidth}
+\caption{Realization of a jumping transitive triplet, where $i$ is the focal actor, $j$ is the target actor, and $h$ is the intermediary.
+The group of the actors is represented by the shape of the node.}
+<<jtt, echo=FALSE, fig.width=2, fig.height=2, message = FALSE, warning = FALSE, fig.align='center'>>=
 jTTe <- data.frame(from = c('i', 'i', 'h'), to = c('h', 'j', 'j'))
@@
 @
+\end{subfigure}\hfill
+\begin{subfigure}[t]{.45\textwidth}
+\caption{Doubly achieved distance between actors $i$ and $k$.}
+<<dab, echo=FALSE, fig.width=2, fig.height=2, message = FALSE, warning = FALSE, fig.align='center'>>=
+dade <- data.frame(from = c('i', 'i', 'h', 'j'), to = c('h', 'j', 'k', 'k'))
+dadn <- data.frame(id = letters[8:11], group = c(1,1,1,1))
+
+dad <- merge(dade, dadn, by.x = 'from', by.y = "id", all = T)
+
+set.seed(12345) 
+ggplot(data = dad, aes(from_id = from, to_id = to)) + 
+  geom_net(aes(shape = as.factor(group)), directed = T, label = T, labelcolour='grey80',vjust = 0, hjust =0, arrowgap = .15, colour = 'black', size=10) + 
+  expand_limits(x=c(-0.1,1.1), y=c(-0.1,1.1)) +
+  theme_net() +
+  theme(legend.position = "none")
+@
+\end{subfigure}
+\caption{\label{fig:structures}Structural newtwork effects. On the left, a jumping transitive triplet (JTT). On the right, a doubly achieved distance between $i$ and $k$.}
+\end{figure}
 

### 2016-05-11 lineups.Rnw
figures

@@
 
+\begin{figure}
 <<get_sm_friends, echo=FALSE, fig.width=8, fig.height=3, message = FALSE, warning = FALSE>>=
 source("Code/small_friends.R")
+view_rs2$wave <- factor(view_rs2$wave)
+levels(view_rs2$wave) <- c("Wave 1", "Wave 2", "Wave 3")
 ggplot(data= view_rs2, aes(x = x, y=y)) + 
-  geom_tile(aes(fill = as.factor(value))) + scale_fill_manual(values = c('white', 'black')) +
+  geom_tile(aes(fill = as.factor(value))) + 
+  scale_fill_manual("X likes Y", labels=c("no", "yes"), values = c('white', 'black')) +
   geom_rect(data= NULL, inherit.aes = FALSE, 
@@
   theme(aspect.ratio=1) 
+@
+\caption{\label{fig:smallfriends}Representation of the edge matrix in waves 1,2, and 3 of the friendship study \hh{XXX cite XXX}. \hh{x and y axes labels are not particularly informative.}}
+\end{figure}
 
-@
+The first wave of the network, which is conditioned on in estimation, is given in Figure~\ref{fig:wave1}.
 
-The first wave of the network, which is conditioned on in estimation, is given below.
-
-<<wave1, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3>>=
+\begin{figure}
+\centering
+<<wave1, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, out.width='.6\\linewidth'>>=
 library(sna)
@@
 library(geomnet)
+actual1$behaviour <- factor(actual1$drink)
+levels(actual1$behaviour) <- c("None", "Once or twice a year", "Once a month", "Once a week")
 ggplot(data = actual1, aes(from_id = X1, to_id = X2)) + 
-  geom_net(label = TRUE, hjust = -.5, labelcolour = 'black', fiteach = T,
-           aes(color = as.factor(drink))) + 
+  geom_net(label = TRUE, hjust = 0.5, vjust=0.5, size=10,  
+           fiteach = T, labelcolour = 'black',
+           aes(colour = behaviour)) + 
+  scale_colour_brewer("Drinking behavior", palette="YlOrRd") +
   theme_net() 
@@
 @
+\caption{\label{fig:wave1} Network of friendships of wave 1 of the subset of students marked in Figure~\ref{fig:smallfriends}. }
+\end{figure}
 
@@
 \begin{tabular}{lccrrr}
-Effect name & Parameter & Corresponding Statistic & $M_1$ Estimate & $M_2$ Estimate & $M_3$ Estimate\\
+Effect name & Parameter & Corresponding Statistic & $M_1$  & $M_2$  & $M_3$ \\
 \hline

### 2016-05-23 lineups.Rnw
couple of comments/questions

@@
 \par Between any two discrete time points, $t_{m}$ and $t_{m-1}$ for $m = 2, \dots, M$, each actor gets the opportunity to change one of their ties at a rate of $\alpha_{m-1}$.
-This is due to the properties of the Markov chain, and it also assumes that the rate of change is constant for each actor between $t_{m-1}$ and $t_{m}$.
+This is due to the properties of the Markov chain, and it also assumes that the rate of change is constant \hh{and the same?} for each actor between $t_{m-1}$ and $t_{m}$.
 When an actor is given an opportunity to change, it tries to maximize its objective function, $f_i(\beta, x)$, where $x$ is the potential new network state that actor $i$ can reach by changing one of its ties, $x_{ij}$.
 In model estimation, the parameters to be estimated are $\alpha_1, \dots, \alpha_{M-1}$ and the (possibly vector-valued) $\beta$
+
+\hh{what is the objective function? XXX}
 
@@
 @
-\caption{\label{fig:smallfriends}Representation of the edge matrix in waves 1,2, and 3 of the friendship study \hh{XXX cite XXX}. \hh{x and y axes labels are not particularly informative.}}
+\caption{\label{fig:smallfriends}Adjacency matrices describing friendship relations between 50 students  in waves 1,2, and 3 of the friendship study \hh{XXX cite XXX}.
+The red squares identify the subset we focus on in the remainder.
+\hh{x and y axes labels are not particularly informative.}}
 \end{figure}
@@
   geom_net(label = TRUE, hjust = 0.5, vjust=0.5, size=10,  
-           fiteach = T, labelcolour = 'black',
+           fiteach = T, labelcolour = "grey20", 
            aes(colour = behaviour)) + 
@@
   theme_net()
+
 #ggplot(data = waves, aes(from_id = X1, to_id = X2)) + 

### 2016-06-01 lineups.Rnw
fiddling with the node labelling

@@
 ggplot(data = jTT, aes(from_id = from, to_id = to)) + 
-  geom_net(aes(shape = as.factor(group)), directed = T, label = T, labelcolour='grey80',vjust = 0, hjust =0, arrowgap = .15, colour = 'black', size=10) + 
+  geom_net(aes(shape = as.factor(group)), directed = T, label = T, labelcolour='grey80',vjust = 0.5, hjust =0.5, arrowgap = .15, colour = 'black', size=10) + 
   expand_limits(x=c(-0.1,1.1), y=c(-0.1,1.1)) +
@@
 ggplot(data = dad, aes(from_id = from, to_id = to)) + 
-  geom_net(aes(shape = as.factor(group)), directed = T, label = T, labelcolour='grey80',vjust = 0, hjust =0, arrowgap = .15, colour = 'black', size=10) + 
+  geom_net(aes(shape = as.factor(group)), directed = T, label = T, labelcolour='grey80',vjust = 0.5, hjust =0.5, arrowgap = .15, colour = 'black', size=10) + 
   expand_limits(x=c(-0.1,1.1), y=c(-0.1,1.1)) +

### 2016-06-25 lineups.Rnw
Start writing up and add refs

@@
 \documentclass[12pt]{article}
-\usepackage{amsmath,amssymb,amsfonts, mathrsfs,fancyhdr,syntonly,lastpage,hyperref,enumitem,graphicx, lscape, forloop, url}
+\usepackage{amsmath,amssymb,amsfonts, mathrsfs,fancyhdr,syntonly,lastpage,hyperref,enumitem,graphicx, lscape, forloop, url, natbib}
 \usepackage[font=small,skip=5pt]{caption}
@@
 \par Between any two discrete time points, $t_{m}$ and $t_{m-1}$ for $m = 2, \dots, M$, each actor gets the opportunity to change one of their ties at a rate of $\alpha_{m-1}$.
-This is due to the properties of the Markov chain, and it also assumes that the rate of change is constant \hh{and the same?} for each actor between $t_{m-1}$ and $t_{m}$.
+This is due to the properties of the Markov chain, and it also assumes that the rate of change is identical and constant for each actor between $t_{m-1}$ and $t_{m}$.
 When an actor is given an opportunity to change, it tries to maximize its objective function, $f_i(\beta, x)$, where $x$ is the potential new network state that actor $i$ can reach by changing one of its ties, $x_{ij}$.
-In model estimation, the parameters to be estimated are $\alpha_1, \dots, \alpha_{M-1}$ and the (possibly vector-valued) $\beta$
+In model estimation, the parameters to be estimated are $\alpha_1, \dots, \alpha_{M-1}$ and the (possibly vector-valued) $\beta$.
+The objective function takes the form
 
-\hh{what is the objective function? XXX}
+$$ f_i(\beta, x) = \sum_k \beta_k s_{ki}(x)$$
+
+where $k = 1, \dots, K$, $K$ the number of non-rate parameters in the models and $s{ki}(x)$ is the corresponding statistic.
+There are several parameters that can be chosen for use in the objective function.
+Examples of these parameters and their corresponding statistics are given in Table~\ref{tab:models}.
 
@@
 <<get_sm_friends, echo=FALSE, fig.width=8, fig.height=3, message = FALSE, warning = FALSE>>=
-source("Code/small_friends.R")
+source("Code/00e_small_friends.R")
 view_rs2$wave <- factor(view_rs2$wave)
@@
   geom_tile(aes(fill = as.factor(value))) + 
-  scale_fill_manual("X likes Y", labels=c("no", "yes"), values = c('white', 'black')) +
-  geom_rect(data= NULL, inherit.aes = FALSE, 
-            aes(xmin = 20, xmax = 35, ymin = 20, ymax = 35, fill =NA), color = 'red') + 
+  scale_fill_manual("X likes Y", labels=c("no", "yes"), 
+                    values = c('white', 'black')) +
+  geom_rect(data= NULL, inherit.aes = FALSE, color = 'red',
+            aes(xmin = 20, xmax = 35, ymin = 20, ymax = 35, fill =NA)) + 
   facet_wrap(~wave) + theme_bw() +
-  theme(aspect.ratio=1) 
+  theme(aspect.ratio=1, axis.text = element_blank(), axis.ticks = element_blank()) 
 @
-\caption{\label{fig:smallfriends}Adjacency matrices describing friendship relations between 50 students  in waves 1,2, and 3 of the friendship study \hh{XXX cite XXX}.
-The red squares identify the subset we focus on in the remainder.
-\hh{x and y axes labels are not particularly informative.}}
+\caption{\label{fig:smallfriends}Adjacency matrices describing friendship relations between 50 students in waves 1,2, and 3 of the friendship study \citep{friendsdata}.
+The red squares identify the subset we focus on in the remainder.}
 \end{figure}
@@
 <<modelfit, echo=FALSE, results = 'hide', message=FALSE, warning = FALSE, cache = T>>=
-source("Code/complete_lineup_creation.R")
+source("Code/03c_complete_lineup_creation.R")
 lineup1 <- create_smfriend_lu(null_eff_struct = null_model_eff2, test_eff_struct = eff_models_smallFriends[[39]], M = 3)
@@
 The two models we fit are a ``null model" (model $M_1$) and two ``alternative models" (model $M_2$ and $M_3$).
-The null model only includes the default parameters that are included in the RSiena estimation.
+The null model only includes the default parameters that are included in the RSiena estimation: the reciprocity and outdegree parameters.
 The alternative model $M_2$ includes one additional covariate parameter that was determined to be significant by the Wald test in the RSienaTest package, while the alternative model $M_3$ includes one additional structural parameter whose significance was determined in the same way.
@@
 An example of this type of closure is given in Figure~\ref{fig:jtt}.
-\hh{XXX Could you color the edge that forms the closer?
-That would be an edge of the form 'I become a friend of my friend's friend', right?
-XXX With the directed edges we also distinguish between 'i likes j' and 'j likes i'.
-} Finally, the doubly achieved distances effect is defined by the number of actors to whom actor $i$ is not directly tied, and tied through two paths via at least two intermediaries \hh{XXX could you give an example for that, too?}.
+With the directed edges we also distinguish between 'i likes j' and 'j likes i'.
+Finally, the doubly achieved distances effect is defined by the number of actors to whom actor $i$ is not directly tied, and tied through two paths via at least two intermediaries.
 This is a structural effect, like the density and reciprocity effects.
-A positive value encourages indirect ties, while a negative value discourages the formation of indirect ties.
+A positive coefficient value encourages indirect ties, while a negative value discourages the formation of indirect ties.
 
@@
 ggplot(data = jTT, aes(from_id = from, to_id = to)) + 
-  geom_net(aes(shape = as.factor(group)), directed = T, label = T, labelcolour='grey80',vjust = 0.5, hjust =0.5, arrowgap = .15, colour = 'black', size=10) + 
+  geom_net(aes(shape = as.factor(group)), directed = T, label = T, 
+           labelcolour='grey80',vjust = 0.5, hjust =0.5, arrowgap = .15, 
+           colour = 'black', size=10, 
+           ecolour = c("red", "grey40", "grey40", "grey40")) + 
   expand_limits(x=c(-0.1,1.1), y=c(-0.1,1.1)) +
@@
 \end{figure}
+
+\section{Lineup Simulation}
+
+To create the lineups that will be used in our experiment, we used the values given in Table~\ref{tab:models} as starting values in simulation.
+For each lineup simulated, the same default RSiena algorithm was used as was used to generate the fitted models with the exception that the algorithm only simulated from the given set of parameters.
+Each lineup and plot within lineup was generated idependent of all the others.
+
+\section{Parameter Estimation from Lineups}
+
+After the lineups were created, we also re-fit each of our three models to each graph in each lineup.
 
@@
 
-\section{Model $M_1$ versus $M_3$ lineup size 3}
-\newcounter{k10}
-\forloop{k10}{1}{\value{k10} < 21}{smallfriends-eff2-m-3-rep-\arabic{k10}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-3-rep-\arabic{k10}} \\ }
+% \section{Model $M_1$ versus $M_3$ lineup size 3}
+% \newcounter{k10}
+% \forloop{k10}{1}{\value{k10} < 21}{smallfriends-eff2-m-3-rep-\arabic{k10}\newline
+% \includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-3-rep-\arabic{k10}} \\ }
+% 
+% \section{Model $M_1$ versus $M_3$ lineup size 6}
+% \newcounter{k11}
+% \forloop{k11}{1}{\value{k11} < 21}{smallfriends-eff2-m-6-rep-\arabic{k11}\newline
+% \includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-6-rep-\arabic{k11}} \\ }
+% 
+% \section{Model $M_1$ versus $M_3$ lineup size 9}
+% \newcounter{k12}
+% \forloop{k12}{1}{\value{k12} < 21}{smallfriends-eff2-m-9-rep-\arabic{k12}\newline
+% \includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-9-rep-\arabic{k12}} \\ }
+% 
+% \section{Model $M_1$ versus $M_3$ lineup size 12}
+% \newcounter{k13}
+% \forloop{k13}{1}{\value{k13} < 21}{smallfriends-eff2-m-12-rep-\arabic{k13}\newline
+% \includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-12-rep-\arabic{k13}} \\ }
+% 
+% \section{Model $M_1$ versus $M_3$ lineup size 16}
+% \newcounter{k14}
+% \forloop{k14}{1}{\value{k14} < 21}{smallfriends-eff2-m-16-rep-\arabic{k14}\newline
+% \includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-16-rep-\arabic{k14}} \\ }
+% 
+% \section{Model $M_3$ versus $M_1$ lineup size 3}
+% \newcounter{k15}
+% \forloop{k15}{1}{\value{k15} < 21}{smallfriends-eff2-rev-m-3-rep-\arabic{k15}\newline
+% \includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-3-rep-\arabic{k15}} \\ }
+% 
+% \section{Model $M_3$ versus $M_1$ lineup size 6}
+% \newcounter{k16}
+% \forloop{k16}{1}{\value{k16} < 21}{smallfriends-eff2-rev-m-6-rep-\arabic{k16}\newline
+% \includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-6-rep-\arabic{k16}} \\ }
+% 
+% \section{Model $M_3$ versus $M_1$ lineup size 9}
+% \newcounter{k17}
+% \forloop{k17}{1}{\value{k17} < 21}{smallfriends-eff2-rev-m-9-rep-\arabic{k17}\newline
+% \includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-9-rep-\arabic{k17}} \\ }
+% 
+% \section{Model $M_3$ versus $M_1$ lineup size 12}
+% \newcounter{k18}
+% \forloop{k18}{1}{\value{k18} < 21}{smallfriends-eff2-rev-m-12-rep-\arabic{k18}\newline
+% \includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-12-rep-\arabic{k18}} \\ }
 
-\section{Model $M_1$ versus $M_3$ lineup size 6}
-\newcounter{k11}
-\forloop{k11}{1}{\value{k11} < 21}{smallfriends-eff2-m-6-rep-\arabic{k11}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-6-rep-\arabic{k11}} \\ }
+% \section{Model $M_3$ versus $M_1$ lineup size 16}
+% \newcounter{k19}
+% \forloop{k19}{1}{\value{k19} < 21}{smallfriends-eff2-rev-m-16-rep-\arabic{k19}\newline
+% \includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-16-rep-\arabic{k19}} \\ }
 
-\section{Model $M_1$ versus $M_3$ lineup size 9}
-\newcounter{k12}
-\forloop{k12}{1}{\value{k12} < 21}{smallfriends-eff2-m-9-rep-\arabic{k12}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-9-rep-\arabic{k12}} \\ }
-
-\section{Model $M_1$ versus $M_3$ lineup size 12}
-\newcounter{k13}
-\forloop{k13}{1}{\value{k13} < 21}{smallfriends-eff2-m-12-rep-\arabic{k13}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-12-rep-\arabic{k13}} \\ }
-
-\section{Model $M_1$ versus $M_3$ lineup size 16}
-\newcounter{k14}
-\forloop{k14}{1}{\value{k14} < 21}{smallfriends-eff2-m-16-rep-\arabic{k14}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-16-rep-\arabic{k14}} \\ }
-
-\section{Model $M_3$ versus $M_1$ lineup size 3}
-\newcounter{k15}
-\forloop{k15}{1}{\value{k15} < 21}{smallfriends-eff2-rev-m-3-rep-\arabic{k15}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-3-rep-\arabic{k15}} \\ }
-
-\section{Model $M_3$ versus $M_1$ lineup size 6}
-\newcounter{k16}
-\forloop{k16}{1}{\value{k16} < 21}{smallfriends-eff2-rev-m-6-rep-\arabic{k16}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-6-rep-\arabic{k16}} \\ }
-
-\section{Model $M_3$ versus $M_1$ lineup size 9}
-\newcounter{k17}
-\forloop{k17}{1}{\value{k17} < 21}{smallfriends-eff2-rev-m-9-rep-\arabic{k17}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-9-rep-\arabic{k17}} \\ }
-
-\section{Model $M_3$ versus $M_1$ lineup size 12}
-\newcounter{k18}
-\forloop{k18}{1}{\value{k18} < 21}{smallfriends-eff2-rev-m-12-rep-\arabic{k18}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-12-rep-\arabic{k18}} \\ }
-
-\section{Model $M_3$ versus $M_1$ lineup size 16}
-\newcounter{k19}
-\forloop{k19}{1}{\value{k19} < 21}{smallfriends-eff2-rev-m-16-rep-\arabic{k19}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-16-rep-\arabic{k19}} \\ }
+\bibliographystyle{apalike}
+\bibliography{lineuppaperbib}
 

### 2016-06-27 lineups.Rnw
couple of questions

@@
 
+\hh{Needs more specifics: how many lineups were created for each model?}
+
+\hh{At the moment we are just considering models M2 and M3 for an inclusion in the lineup study. We are trying to nail down what to include and what not to include in the experiment.  }
+
 To create the lineups that will be used in our experiment, we used the values given in Table~\ref{tab:models} as starting values in simulation.
@@
 
-After the lineups were created, we also re-fit each of our three models to each graph in each lineup.
+After the lineups were created, we re-fit each of our three models to each graph in each lineup.
+
+\hh{Questions that should be answered in this section:
+
+1. Why do we want to get the parameters for each model?
+
+2. How is the fitting done, exactly?
+
+4. What parameters are in the output?
+
+5. What are the results?
+
+6. How do the results influence our decision for the lineups?
+
+}
+
+\bibliographystyle{apalike}
+\bibliography{lineuppaperbib}
 
@@
 
-\bibliographystyle{apalike}
-\bibliography{lineuppaperbib}
-
 \end{document}

### 2016-06-28 lineups.Rnw
couple quick paragraphs

@@
 
+<<results1, echo = FALSE>>=
+lus_res_long <- read.csv("Data/June_estimates_from_lineups.csv", stringsAsFactors = FALSE)
+
+@
+
 6. How do the results influence our decision for the lineups?
+
+The strong influence of the inclusion of $\beta_4$ in model 3 has made any comparison between model 1 and model 3 impossible.
+The $\beta_4$ estimate is always significantly greater than zero in fitted models that converged.
+Thus, $\beta_4$ should have been included from the beginning.
+% at end of param est we know that model m3 is useless. m3 messes with structure of m1. should have been in cluded in all of the models. nice & important but doesn't help for lineup study
+
+%  results from the the pilot study.
+how do estimates combine with pilot study ?
+have 10 non-m3 ones that are good.
+also want to see that m1 v m3 is way worse than m1 v m2.
+now have a reason for m1 v m3 is so bad.
+m1 v m3 ids should be NO GOOD.
 

### 2016-06-28 lineups.Rnw
added a bit to the section

@@
 
+\section{Results from the pilot study}
+
+\hh{Can we derive measures from the model estimates or the visual representation (ie.
+the network structure) that help us determine which lineups are more difficult than others, i.e.
+based on these estimates, how relaibly can we predict which panel will be picked from a lineup?
+XXX For that, we could also calculate the number of JTTs in each network - use `jtt` for that.}
+
 \bibliographystyle{apalike}

### 2016-06-29 Code/05b_merge_estimates_wtruth.R
save lus_ests_truth

@@ -85,6 +85,11 @@ library(dplyr)
 library(purrr)
 library(magrittr)
 
+lineups <- expand.grid(lineupname = c("smallfriends", "smallfriends-rev", "smallfriends-eff2", "smallfriends-eff2-rev" ),
+                       M = c(3,6,9,12,16), 
+                       rep = 1:20)
+
+lineups$lineupid <- paste(lineups$lineupname, lineups$M, lineups$rep,sep = '-')
 lineups$lineupname %<>% as.character()
 lineups %>% nest(-lineupid) %>% mutate(matching = map(data, get_truth)) -> matching_lineups 
 
@@ -111,6 +116,7 @@ lus_ests_truth <- merge(lus_res_long, panels_and_truth)
 head(lus_ests_truth)
 
 lus_ests_truth %<>% arrange(lineupname, M, rep, model, panel_num, param_name)
+write.csv(lus_ests_truth, file="Data/lus_ests_truth.csv", row.names=FALSE)
 
 library(ggplot2)
 ggplot(data = lus_ests_truth) + geom_density(alpha = .6, aes(x = param_est, fill = true_model)) + 

### 2016-06-29 lineups.Rnw
figure added, some discussion

@@
 5. What are the results?
+}
+\begin{figure}
+<<results1, fig.width=8, fig.height=6, out.width='\\linewidth', echo = FALSE, warning = FALSE>>=
+lus_ests_truth <- read.csv("Data/lus_ests_truth.csv", stringsAsFactors = FALSE)
+lus_ests_truth$param_name <- 
+  factor(lus_ests_truth$param_name,
+    levels = c("rate", "outdegree (density)", "reciprocity",
+               "transitive triplets jumping alcohol2",
+               "number pairs at doubly achieved distance 2"))                                
+l1 <- levels(lus_ests_truth$param_name)
+l1 <- stringr::str_trim(gsub("2","", l1))
+l1[5] <- "#pairs at doubly achieved distance"
 
-<<results1, echo = FALSE>>=
-lus_res_long <- read.csv("Data/June_estimates_from_lineups.csv", stringsAsFactors = FALSE)
+ggplot(data = lus_ests_truth) + 
+  geom_vline(aes(xintercept = 0), colour="grey50") +
+  geom_density(alpha = .5, aes(x = param_est, fill = param_name)) + 
+  facet_grid(model~true_model, scales = 'free', 
+             labeller = "label_both") + xlim(c(-25,25)) +
+  theme_bw() + 
+  scale_fill_brewer("Parameter", palette="Dark2",
+     labels=c(bquote(paste(alpha,": ", .(l1[1]), sep="")), 
+              bquote(paste(beta[1],": ", .(l1[2]), sep="")),
+              bquote(paste(beta[2],": ", .(l1[3]), sep="")),
+              bquote(paste(beta[3],": ", .(l1[4]), sep="")),
+              bquote(paste(beta[4],": ", .(l1[5]), sep="")))) +
+  theme(legend.position = "bottom") +
+  xlab("Parameter Estimate") + 
+  guides(fill = guide_legend(nrow = 3, byrow = TRUE))
+@
+\caption{\label{fig:comparison} Comparison of model estimates under all three models under investigation.}
+\end{figure}
 
-@
+Figure~\ref{fig:comparison} shows an overview of model estimates for each simulated data set.
+What we expect to see is that estimates do not change values (much) if they are estimated under a model different from the one they are generated from.
+This is true for all data sets estimated under model $M_1$ independently of which model they were generated from (top row of Figure~\ref{fig:comparison}).
+For data fitted under model $M_2$ we see that parameter $\beta_3$ (jumping transitive triplets) are estimated to be about zero, if the data is generated from models $M_1$ and $M_3$.
+For model $M_2$, parameter $\beta_3$ is estimated to be significantly different from zero in almost all cases \hh{(XXX how many exactly? XXX)}. This coincides with our expectation.
 
+However, the bottom row of Figure~\ref{fig:comparison} shows that independently of which model data is generated from, $\beta_4$, the number of pairs at doubly achieved distance, is estimated to be significantly different from zero in a large number of cases (about half in data generated from model $M_2$ and more than that in model $M_1$).
+While $\beta_4$ is a highly significant parameter in model $M_3$, this questions the way that parameters are fitted and tells us that we are not likely to be able to visually distinguish between data generated from model $M_3$ and data generated from models $M_1$ and $M_2$.
+We might, however, be able to distinguish between data sets for which $\beta_4$ is estimated to be significantly different from zero and non-significant ones.
+\hh{XXX can we see differences in the pilot study?}
+
+\hh{
 6. How do the results influence our decision for the lineups?
-
+}
 The strong influence of the inclusion of $\beta_4$ in model 3 has made any comparison between model 1 and model 3 impossible.
@@
 m1 v m3 ids should be NO GOOD.
-
-}
 

### 2016-06-29 lineups.Rnw
adding jtts for each lineup

@@
 <<results1, fig.width=8, fig.height=6, out.width='\\linewidth', echo = FALSE, warning = FALSE>>=
-lus_ests_truth <- read.csv("Data/lus_ests_truth.csv", stringsAsFactors = FALSE)
+load("Data/lus_ests_truth.rda")
 lus_ests_truth$param_name <- 
@@
   facet_grid(model~true_model, scales = 'free', 
-             labeller = "label_both") + xlim(c(-25,25)) +
+             labeller = "label_both") + xlim(c(-20,20)) +
   theme_bw() + 

### 2016-06-29 Code/06_lus_parms.R

@@ -1,0 +1,35 @@
+library(magrittr)
+library(dplyr)
+dir <- "Data/lineupdata/"
+lineupdata <- dir(dir, pattern="small")
+
+djtt <- function(X1, X2) {
+  data <- data.frame(X1=X1, X2=X2)
+  data <- na.omit(data)
+  if (nrow(data) == 0) return(0)
+  geomnet::jtt(data, from_id="X1", to_id="X2")
+}
+
+
+dframe <- lineupdata %>% purrr::map_df(.f = function(x) {
+  net = read.csv(file.path(dir=dir, x))
+  net %>% group_by(plot_order) %>% summarize(
+    jtt = djtt(X1, X2))
+}, .id="filename")
+
+dframe <- data.frame(dframe)
+dframe$filename <- lineupdata[as.numeric(dframe$filename)]
+
+dframe$m <- gsub(".*-m-([0-9]*)-.*", "\\1", dframe$filename)
+dframe$rep <- gsub(".*-rep-([0-9]*).csv", "\\1", dframe$filename)
+dframe$model <- gsub("(.*)-m.*", "\\1", dframe$filename)
+
+library(ggplot2)
+qplot(data=subset(dframe, model=="smallfriends"), 
+      x=plot_order, y=jtt, colour=factor(rep)) + facet_wrap(~filename)
+
+qplot(data=subset(dframe, model=="smallfriends-rev"), 
+      x=plot_order, y=jtt, colour=factor(rep)) + facet_wrap(~filename)
+
+
+write.csv(dframe, file="Data/lineups-jtt.csv", row.names=FALSE)

### 2016-06-30 lineups.Rnw
beginning to answer question 1

@@
 1. Why do we want to get the parameters for each model? 
+\st{We fit all three models to every plot in the lineups.
+Fitting all models to all lineup plots allows us to gauge tha ability of the parameter estimates to provide a measure of lineup identification difficulty.
+XXX ?
+not quite sure ?
+XXX For instance, when comparing models 1 and 2 in a lineup, the estimates from fitting model 2 to plots which were simulated from model 2 should be significantly different from the model 2 estimates from plots simulated from model 1.
+The smaller the difference between these estimates, the harder it should be to identify the different model in the lineup.
+XXX IDEA: should we consider some sort of distance metric comparing all estimates from the models at once in addition to the differences in the $\beta_3$/$\beta_4$ values?
+XXX}
 
 2. How is the fitting done, exactly?
+\st{XXX get into nitty gritty from RSiena model XXX }
 

### 2016-07-01 Code/07pilot-estimates-merge.R
initial evaluation of pilot study results

@@ -1,0 +1,29 @@
+pdffiles <- dir("../GGExperimentApr28/", pattern="pdf")
+pdffiles <- gsub(".pdf","",pdffiles)
+pdffiles <- gsub("-m","",pdffiles)
+pdffiles <- gsub("-rep","",pdffiles)
+
+dframe <- strsplit(pdffiles, split=" ") %>% plyr::ldply(function(x) x)
+names(dframe) <- c("stimulus", "lineupid")
+
+results <- read.csv("../GGExperimentApr28/responses_GGExpApr28.csv")
+lus <- results %>% group_by(Lineup, ChosenLU) %>% summarize(
+  tally = n(),
+  data = Answer[1],
+  reason = paste(Reasoning, collapse="|")
+)
+names(lus)[2] <- "panel"
+
+dframe_res <- merge(dframe, lus, all=TRUE, by.x="stimulus", by.y="Lineup")
+dframe_res$data_pick <- with(dframe_res, data==panel)
+dframe_res$model <- gsub("(.*)-[0-9]*-.*", "\\1", dframe_res$lineupid)
+dframe_res$data_pick <- factor(dframe_res$data_pick, levels=c("TRUE", "FALSE"))
+dframe_res$res <- factor(dframe_res$data_pick, levels=c("TRUE", "FALSE"))
+qplot(model, weight=tally, fill=data_pick, data=dframe_res) 
+qplot(lineupid, weight=tally, fill=data_pick, data=dframe_res) +facet_grid(.~model, scales="free", space="free") +
+  theme(axis.text.x = element_text(angle=90))
+
+
+load("../Data/lus_ests_truth.rda")
+# spread the lus_ests_truth data by parameter. There should be one line for each panel in each lineup
+lus_ests_pilot <- merge(lus_ests_truth, dframe, all.x=TRUE, by="lineupid") # 

### 2016-07-01 lineups.Rnw
barchart uploaded

@@
 
-\hh{Can we derive measures from the model estimates or the visual representation (ie.
+\hh{Big goal: Can we derive measures from the model estimates or the visual representation (ie.
 the network structure) that help us determine which lineups are more difficult than others, i.e.
@@
 XXX For that, we could also calculate the number of JTTs in each network - use `jtt` for that.}
+
+\hh{Littler goal: evaluate pilot study and see whether the results line up with the conjectures made in the previous section.}
+
+\hh{The pilot study consisted of an evaluation of a set of 20 lineups by 11 volunteers.
+For each of the model situations ($M_1$ vs $M_2$, $M_2$ vs $M_1$, $M_1$ vs $M_3$ and $M_3$ vs $M_1$) on lineup of size $m = 3, 6, 9, 12$ was used.}
+\paragraph{Suitability of $M_3$ in lineups:}
+\begin{figure}
+\centering
+<<data-picks, echo=FALSE, fig.width=10, fig.height=5, out.width='0.85\\linewidth'>>=
+pdffiles <- dir("GGExperimentApr28/", pattern="pdf")
+pdffiles <- gsub(".pdf","",pdffiles)
+#pdffiles <- gsub("-m","",pdffiles)
+#pdffiles <- gsub("-rep","",pdffiles)
+
+dframe <- strsplit(pdffiles, split=" ") %>% plyr::ldply(function(x) x)
+names(dframe) <- c("stimulus", "lineupid")
+
+results <- read.csv("GGExperimentApr28/responses_GGExpApr28.csv")
+lus <- results %>% group_by(Lineup, ChosenLU) %>% summarize(
+  tally = n(),
+  data = Answer[1],
+  reason = paste(Reasoning, collapse="|")
+)
+names(lus)[2] <- "panel"
+
+dframe_res <- merge(dframe, lus, all=TRUE, by.x="stimulus", by.y="Lineup")
+dframe_res$data_pick <- with(dframe_res, data==panel)
+dframe_res$model <- factor(gsub("(.*)-m.*", "\\1", dframe_res$lineupid))
+levels(dframe_res$model) <- c("M1 vs M2",  "M1 vs M3", "M3 vs M1", "M2 vs M1")
+dframe_res$model <- factor(dframe_res$model, levels = c("M1 vs M2",  "M2 vs M1", "M1 vs M3", "M3 vs M1"))
+dframe_res$label <- gsub(".*-(m.*)", "\\1", dframe_res$lineupid)
+
+dframe_res$data_pick <- factor(dframe_res$data_pick, levels=c("TRUE", "FALSE"))
+dframe_res$res <- factor(dframe_res$data_pick, levels=c("TRUE", "FALSE"))
+qplot(label, weight=tally, fill=data_pick, data=dframe_res) +
+  facet_grid(.~model, scales="free", space="free") + theme_bw() +
+  theme(axis.text.x = element_text(angle=45, hjust=1)) + 
+  scale_fill_brewer("Data panel chosen", palette="Set1", na.value="grey80") 
+@
+\caption{\label{fig:data-picks}Barchart summarizing the number of responses from the pilot study by model and lineup.
+Color shows the number of times the data panel was chosen from the lineup.
+Clearly, the first two sets of lineups ($M_1$ vs $M_2$ and $M_2$ vs $M_1$) have on average higher number of data identifications.
+}
+\end{figure}
+
+\hh{Figure~\ref{fig:data-picks} shows barcharts of responses from the pilot study distinguishing the}
+\hh{It is more difficult to identify the data plot in lineups of graphs based on data from models $M_1$ and $M_3$ than  in lineups of graphs based on data from models $M_1$ and $M_2$.
+}
 

### 2016-07-01 lineups.Rnw
cosmetic changes to chart

@@
 \hh{The pilot study consisted of an evaluation of a set of 20 lineups by 11 volunteers.
-For each of the model situations ($M_1$ vs $M_2$, $M_2$ vs $M_1$, $M_1$ vs $M_3$ and $M_3$ vs $M_1$) on lineup of size $m = 3, 6, 9, 12$ was used.}
+For each of the model situations ($M_1$ vs $M_2$, $M_2$ vs $M_1$, $M_1$ vs $M_3$ and $M_3$ vs $M_1$) one lineup of size $m = 3, 6, 9, 12$ was used.
+Overall, the number of data identifications in the lineups was very low (34 out of 220 evaluations).
+Participants identified the data plot in two to four of the lineups they evaluated.
+The mode was three data identifications.
+}
 \paragraph{Suitability of $M_3$ in lineups:}
@@
   theme(axis.text.x = element_text(angle=45, hjust=1)) + 
-  scale_fill_brewer("Data panel chosen", palette="Set1", na.value="grey80") 
+  scale_fill_brewer("Data panel chosen", palette="Set1", na.value="grey80") + xlab("Lineup")
 @
@@
 
-\hh{Figure~\ref{fig:data-picks} shows barcharts of responses from the pilot study distinguishing the}
+\hh{Figure~\ref{fig:data-picks} shows barcharts of responses from the pilot study detailing the number of data identifications in each lineup. }
 \hh{It is more difficult to identify the data plot in lineups of graphs based on data from models $M_1$ and $M_3$ than  in lineups of graphs based on data from models $M_1$ and $M_2$.

### 2016-07-02 Code/05b_merge_estimates_wtruth.R
file to merge the experiment with sim data

@@ -166,7 +166,7 @@ for (i in 1:nrow(lus_ests_truth)){
     } else { lus_ests_truth$convergence[i] <- "Did not converge" }
   }
 }
-
+saveRDS(lus_ests_truth,"~/Desktop/NetworksResearch/NetworksVizInference/Data/lus_ests_truth.rda")
 table(lus_ests_truth$convergence, useNA = 'ifany')
 
 ggplot(data = lus_ests_truth) + geom_density(alpha = .8, aes(x = param_est, fill = true_model)) + 

### 2016-07-02 Code/07_merge_psdata_simres.R

@@ -1,0 +1,28 @@
+# merge the lineup pilot study info with the lus_ests_truth
+
+load("~/Desktop/NetworksResearch/NetworksVizInference/Data/lus_ests_truth.rda")
+lus_ests_truth$convergence <- NA
+
+for (i in 1:nrow(lus_ests_truth)){
+  if (is.na(lus_ests_truth$converg_stat[i])){
+    lus_ests_truth$convergence[i] <- NA
+  } else if (lus_ests_truth$param_name[i] != "rate"){
+    if (lus_ests_truth$converg_stat[i] <= 0.1 & lus_ests_truth$converg_stat[i] >= -.1){
+      lus_ests_truth$convergence[i] <- "Converged"
+    } else{  lus_ests_truth$convergence[i] <- "Did not converge"} 
+  } else {
+    if (lus_ests_truth$converg_stat[i] <= 0.3){
+      lus_ests_truth$convergence[i] <- "Converged"
+    } else { lus_ests_truth$convergence[i] <- "Did not converge" }
+  }
+}
+
+
+
+
+names(lus_ests_truth)
+summary(lus_ests_truth$convergence)
+
+
+# read.csv("~/Desktop/NetworksResearch/NetworksVizInference/GGExperimentApr28/data_for_experiment.csv")
+readxl::read_excel("~/Desktop/NetworksResearch/NetworksVizInference/GGExperimentApr28/responses_GGExpApr28.xlsx")

### 2016-07-02 Code/05b_merge_estimates_wtruth.R
merge pilot study results (lus_truth_expres)

@@ -152,7 +152,8 @@ ggplot(data = dplyr::filter(others, model == "M3")) +
 # good convergence overall <.3, great convergence <.2
 
 names(lus_ests_truth2)
-lus_ests_truth2$convergence <- NA
+lus_ests_truth <- lus_ests_truth2
+lus_ests_truth$convergence <- NA
 
 for (i in 1:nrow(lus_ests_truth)){
   if (is.na(lus_ests_truth$converg_stat[i])){
@@ -168,7 +169,7 @@ for (i in 1:nrow(lus_ests_truth)){
   }
 }
 
-saveRDS(lus_ests_truth,"~/Desktop/NetworksResearch/NetworksVizInference/Data/lus_ests_truth.rda")
+save(lus_ests_truth, file = "~/Desktop/NetworksResearch/NetworksVizInference/Data/lus_ests_truth.rda")
 table(lus_ests_truth$convergence, useNA = 'ifany')
 
 ggplot(data = lus_ests_truth) + geom_density(alpha = .8, aes(x = param_est, fill = true_model)) + 

### 2016-07-02 Code/07_merge_psdata_simres.R

@@ -1,5 +1,4 @@
 # merge the lineup pilot study info with the lus_ests_truth
-
 load("~/Desktop/NetworksResearch/NetworksVizInference/Data/lus_ests_truth.rda")
 # lus_ests_truth$convergence <- NA
 # 
@@ -38,4 +37,5 @@ head(lus_ests_truth)
 # ggexp_datres: lineupname  M rep 
 head(ggexp_datres)
 
-left_join(lus_ests_truth, ggexp_datres, by = c("lineupname" = "lineupname", "M" = "M", "rep" = "rep"))
+lus_truth_expres <- left_join(lus_ests_truth, ggexp_datres, by = c("lineupname" = "lineupname", "M" = "M", "rep" = "rep", "panel_num" = "ChosenLU"))
+save(lus_truth_expres, file = "~/Desktop/NetworksResearch/NetworksVizInference/Data/lus_truth_expres.RDA")

### 2016-07-03 lineups.Rnw
notes from mtg today
barchart with convergence pattern

@@
 not quite sure ?
-XXX For instance, when comparing models 1 and 2 in a lineup, the estimates from fitting model 2 to plots which were simulated from model 2 should be significantly different from the model 2 estimates from plots simulated from model 1.
+XXX For instance, when comparing models 1 and 2 in a lineup, the estimates from fitting model 2 to plots which were simulated from model 2 should be significantly different from the model 2 estimates from plots simulated from model
+0. Convergence issues
+
+\begin{figure}
+<<results0, fig.width=8, fig.height=3, out.width='\\linewidth', echo = FALSE, warning = FALSE>>=
+load("Data/lus_ests_truth.rda")
+lus_ests_truth$param_name <- 
+  factor(lus_ests_truth$param_name,
+    levels = c("rate", "outdegree (density)", "reciprocity",
+               "transitive triplets jumping alcohol2",
+               "number pairs at doubly achieved distance 2"))
+
+ggplot(data=lus_ests_truth) +
+  geom_bar(aes(x=model, fill=convergence), position="fill") + 
+  facet_grid(.~true_model, labeller="label_both") +
+  theme_bw() +
+  theme(legend.position="bottom") + ylab("Ratio") +
+  scale_fill_brewer("Convergence", palette="Set1") +
+  xlab("Estimated Model")
+@
+\caption{\label{fig:convergence}Pattern of convergence.
+A total of 67.7\% of all lineup data converged in 5,000 iterations.
+The simplest model, $M_1$ has the highest rate of convergence.
+For models $M_2$ and $M_3$ data generated from the model converged at a higher rate than data generated from the other model.}
+\end{figure}
+
+1.
 The smaller the difference between these estimates, the harder it should be to identify the different model in the lineup.
@@
 4. What parameters are in the output?
-
+% ST: add how convergence is calculated. from RSiena manual. 
 5. What are the results?
@@
 <<results1, fig.width=8, fig.height=6, out.width='\\linewidth', echo = FALSE, warning = FALSE>>=
-load("Data/lus_ests_truth.rda")
-lus_ests_truth$param_name <- 
-  factor(lus_ests_truth$param_name,
-    levels = c("rate", "outdegree (density)", "reciprocity",
-               "transitive triplets jumping alcohol2",
-               "number pairs at doubly achieved distance 2"))                                
 l1 <- levels(lus_ests_truth$param_name)
@@
 
-ggplot(data = lus_ests_truth) + 
+# subset on ones that are actually converged. 
+ggplot(data = subset(lus_ests_truth, convergence=="Converged")) + 
   geom_vline(aes(xintercept = 0), colour="grey50") +
@@
 
-However, the bottom row of Figure~\ref{fig:comparison} shows that independently of which model data is generated from, $\beta_4$, the number of pairs at doubly achieved distance, is estimated to be significantly different from zero in a large number of cases (about half in data generated from model $M_2$ and more than that in model $M_1$).
+However, the bottom row of Figure~\ref{fig:comparison} shows that independently of which model data is generated from, $\beta_4$, the number of pairs at doubly achieved distance, is estimated to be significantly different from zero in a large number of cases (about two thirds of the data generated from model $M_2$ and more than that in model $M_1$).
 While $\beta_4$ is a highly significant parameter in model $M_3$, this questions the way that parameters are fitted and tells us that we are not likely to be able to visually distinguish between data generated from model $M_3$ and data generated from models $M_1$ and $M_2$.

### 2016-07-03 lineups.Rnw
convergence description in write up

@@
 4. What parameters are in the output?
+\st{In RSiena, the convergence of a non-rate parameter is determined through the simulated values from Phase 3 of the SienaFit algorithm.
+XXX more detail will be above later XXX  The simulations are compared to the observed values of the statistics observed in the data.
+The values from the simulations should be fairly close to the observed values, but because the fitting is done stochastically, deviations from statistics will not be exactly zero.
+Checking for convergence is based on a t-ratio of the average of these deviations to the standard deviation of these deviations.
+RSiena also performs an overall maximum convergence check by finding the maximum t-ratio value for any linear combination of the observed statistics.
+According to the RSiena Manual, ``convergence is excellent when the overall maximum convergence ratio is less than 0.2", and for the non-rate parameters, the threshold for ``reasonable" convergence is set at 0.3 with excellent convergence when the t-ratio is less than or equal to 0.1 in absolute value.
+(\cite{RSiena}).
+}
+
 % ST: add how convergence is calculated. from RSiena manual. 

### 2016-07-05 lineups.Rnw
start writing up model fitting procedure.

@@
 2. How is the fitting done, exactly?
-\st{XXX get into nitty gritty from RSiena model XXX }
+\st{XXX get into nitty gritty from RSiena model XXX}
+
+\st{The model fitting in RSiena is done in three phases.
+Briefly, in the initial phase, sensitivity of the statistics to the parameter values is determined, then in the second phase, parameter values are fit iteratively.
+Finally, in the third phase, networks are simulated from the fitted models and the model is checked for convergence.
+}
 
@@
 }
-
-% ST: add how convergence is calculated. from RSiena manual. 
 5. What are the results?

### 2016-07-06 lineups.Rnw
going back on showing only converged parameter estimates - not useful

@@
 
-# subset on ones that are actually converged. 
-ggplot(data = subset(lus_ests_truth, convergence=="Converged")) + 
+# subset on ones that are actually converged. XXX No - that kills the zeroes in estimates that should be zero
+ggplot(data = subset(lus_ests_truth)) + #, convergence=="Converged")) + 
   geom_vline(aes(xintercept = 0), colour="grey50") +
@@
 @
-\caption{\label{fig:comparison} Comparison of model estimates under all three models under investigation.}
+\caption{\label{fig:comparison} Comparison of model estimates under all three models under investigation. }
 \end{figure}

### 2016-07-06 Code/06b_lus_parameters.R
deriving first distance between data and null plots

@@ -1,0 +1,54 @@
+load("Data/lus_ests_truth.rda")
+
+lus_ests_truth$true_panel <- lus_ests_truth$true_model == lus_ests_truth$model
+lus_spread <- tidyr::spread(lus_ests_truth[,c("lineupid","model", "true_model", "panel_num", "param_name", "param_est")], model, param_est)
+
+# parameters should be most different between data plot and null plots
+
+# merge in true values:
+truth <- unique(subset(lus_ests_truth, true_panel==TRUE)[,c("lineupid", "true_model","param_name","true_value")])
+dim(lus_spread)
+lus_spread <- merge(lus_spread, truth, by=c("lineupid","true_model", "param_name"), all.x=TRUE)
+lus_spread$true_value[is.na(lus_spread$true_value)] <- 0
+qplot(M2, true_value, facets=~param_name, data=lus_spread, colour=true_model)
+qplot(M3, true_value, facets=~param_name, data=lus_spread, colour=true_model)
+
+# compute a distance between the data panel and the closest null panel
+lus_spread$part <- gsub("(.*)-[0-9]*-[0-9]*", '\\1', lus_spread$lineupid)
+
+lus_summary_2 <- lus_spread %>% group_by(part, lineupid, param_name) %>%
+  summarize(
+    data_n = sum(true_model=="M2"),
+    data_est = mean(M2[true_model=="M2"]),
+    data_sd = sd(M2[true_model=="M2"], na.rm=TRUE),
+    null_est = list(M2[true_model=="M1"]),
+    data_dist = data_est - max(unlist(null_est)),
+    upper_qu = sum(data_est > unlist(null_est))/(n()-1)
+  )
+lus_summary_2$m <- gsub(".*-([0-9]*)-.*","\\1",lus_summary$lineupid)
+
+qplot(lineupid, upper_qu, data = subset(lus_summary_2, !is.na(upper_qu) & param_name=="transitive triplets jumping alcohol2")) +
+  facet_grid(.~part, scale="free_x", space="free")
+
+qplot(upper_qu, geom="histogram", data = subset(lus_summary_2, !is.na(upper_qu) & param_name=="transitive triplets jumping alcohol2")) +
+  facet_grid(.~part)
+
+qplot(data_dist, upper_qu, colour=m,  data = subset(lus_summary_2, !is.na(upper_qu) & param_name=="transitive triplets jumping alcohol2")) +
[…]

### 2016-07-06 lineups.Rnw
first figure on differences between estimates of data and null

@@
 
-However, the bottom row of Figure~\ref{fig:comparison} shows that independently of which model data is generated from, $\beta_4$, the number of pairs at doubly achieved distance, is estimated to be significantly different from zero in a large number of cases (about two thirds of the data generated from model $M_2$ and more than that in model $M_1$).
+However, the bottom row of Figure~\ref{fig:comparison} shows that independently of which model data is generated from, $\beta_4$, the number of pairs at doubly achieved distance, is estimated to be significantly different from zero in a large number of cases (about half of the data generated from model $M_2$ and more than that in model $M_1$).
 While $\beta_4$ is a highly significant parameter in model $M_3$, this questions the way that parameters are fitted and tells us that we are not likely to be able to visually distinguish between data generated from model $M_3$ and data generated from models $M_1$ and $M_2$.
@@
 \hh{XXX can we see differences in the pilot study?}
+
+\begin{figure}
+\centering
+<<echo=FALSE, fig.with=8, fig.height = 8, out.width='.6\\linewidth'>>=
+load("Data/lus_ests_truth.rda")
+
+lus_ests_truth$true_panel <- lus_ests_truth$true_model == lus_ests_truth$model
+lus_spread <- tidyr::spread(lus_ests_truth[,c("lineupid","model", "true_model", "panel_num", "param_name", "param_est")], model, param_est)
+
+# parameters should be most different between data plot and null plots
+
+# merge in true values:
+truth <- unique(subset(lus_ests_truth, true_panel==TRUE)[,c("lineupid", "true_model","param_name","true_value")])
+#dim(lus_spread)
+lus_spread <- merge(lus_spread, truth, by=c("lineupid","true_model", "param_name"), all.x=TRUE)
+lus_spread$true_value[is.na(lus_spread$true_value)] <- 0
+#qplot(M2, true_value, facets=~param_name, data=lus_spread, colour=true_model)
+#qplot(M3, true_value, facets=~param_name, data=lus_spread, colour=true_model)
+
+# compute a distance between the data panel and the closest null panel
+lus_spread$part <- gsub("(.*)-[0-9]*-[0-9]*", '\\1', lus_spread$lineupid)
+
+lus_summary_2 <- lus_spread %>% group_by(part, lineupid, param_name) %>%
+  summarize(
+    data_n = sum(true_model=="M2"),
+    data_est = mean(M2[true_model=="M2"]),
+    data_sd = sd(M2[true_model=="M2"], na.rm=TRUE),
+    null_est = list(M2[true_model=="M1"]),
+    data_dist = data_est - max(unlist(null_est)),
+    upper_qu = sum(data_est > unlist(null_est))/(n()-1)
+  )
+
+p1 <- qplot(data_dist, upper_qu, data = subset(lus_summary_2, !is.na(upper_qu) & param_name=="transitive triplets jumping alcohol2")) +
+  facet_grid(.~part)
+
+lus_summary_3 <- lus_spread %>% group_by(part, lineupid, param_name) %>%
+  summarize(
+    data_n = sum(true_model=="M3"),
+    data_est = mean(M3[true_model=="M3"]),
+    data_sd = sd(M3[true_model=="M3"], na.rm=TRUE),
+    null_est = list(M3[true_model=="M1"]),
+    data_dist = min(unlist(null_est)) - data_est,
+    lower_qu = sum(data_est < unlist(null_est))/(n()-1)
+  )
+
+p2 <- qplot(data_dist, lower_qu,  data = subset(lus_summary_3, !is.na(lower_qu) & param_name=="number pairs at doubly achieved distance 2")) +
+  facet_grid(.~part)
+
+gridExtra::grid.arrange(p1,p2)
+@
+\caption{\label{fig:lusparms}}
+\end{figure}
 

### 2016-07-06 lineups.Rnw
why dplyr, whyyyyyy???? :'(

@@
     null_est = list(M2[true_model=="M1"]),
-    data_dist = data_est - max(unlist(null_est)),
-    upper_qu = sum(data_est > unlist(null_est))/(n()-1)
+    #data_dist = data_est - max(unlist(null_est)),
+    #upper_qu = sum(data_est > unlist(null_est))/(n()-1)
   )
+# XXX The lines commented out above crash my RStudio. I haven't figured out why yet! :( XXX
 

### 2016-07-07 lineups.Rnw
still cannot figure out why rstudio keeps crashing

--- 2016-07-06 lineups.Rnw
@@
 
-lus_summary_2 <- lus_spread %>% group_by(part, lineupid, param_name) %>%
-  summarize(
+lus_summary_2 <- lus_spread %>% dplyr::group_by(part, lineupid, param_name) %>%
+  dplyr::summarize(
     data_n = sum(true_model=="M2"),
@@
     null_est = list(M2[true_model=="M1"]),
-    data_dist = data_est - max(unlist(null_est)),
-    upper_qu = sum(data_est > unlist(null_est))/(n()-1)
-  )
+    data_dist = data_est - max(unlist(null_est))),
+  upper_qu = sum(data_est > unlist(null_est)) / (n()-1) )
+
+# XXX The lines commented out above crash my RStudio. I haven't figured out why yet! :( XXX
 

### 2016-07-07 lineups.Rnw
caption added

@@
 p1 <- qplot(data_dist, upper_qu, data = subset(lus_summary_2, !is.na(upper_qu) & param_name=="transitive triplets jumping alcohol2")) +
-  facet_grid(.~part)
+  facet_grid(.~part) + xlab("Difference in M2 estimate of data panel and maximum of the null panels") +
+  ylab("Ratio of null panels with smaller M2 estimate") +
+  ggtitle("jumping transitive triplets")
 
@@
 p2 <- qplot(data_dist, lower_qu,  data = subset(lus_summary_3, !is.na(lower_qu) & param_name=="number pairs at doubly achieved distance 2")) +
-  facet_grid(.~part)
+  facet_grid(.~part) +
+  xlab("Difference in M3 estimate of data panel and minimum of the null panels") +
+  ylab("Ratio of null panels with larger M3 estimate") +
+  ggtitle("doubly achieved distance")
 
@@
 @
-\caption{\label{fig:lusparms}}
+\caption{\label{fig:lusparms} Scatterplots }
 \end{figure}

### 2016-07-07 lineups.Rnw
does it still crash? - I'm using dplyr_0.5.0

@@
 \begin{subfigure}[t]{.45\textwidth}
-\caption{Realization of a jumping transitive triplet, where $i$ is the focal actor, $j$ is the target actor, and $h$ is the intermediary.
+\caption{\label{fig:jtt}Realization of a jumping transitive triplet, where $i$ is the focal actor, $j$ is the target actor, and $h$ is the intermediary.
 The group of the actors is represented by the shape of the node.}
@@
 \centering
-<<echo=FALSE, fig.with=8, fig.height = 8, out.width='.6\\linewidth'>>=
+<<echo=FALSE, fig.width=8, fig.height = 8, out.width='.6\\linewidth', warning = FALSE>>=
 load("Data/lus_ests_truth.rda")
@@
 truth <- unique(subset(lus_ests_truth, true_panel==TRUE)[,c("lineupid", "true_model","param_name","true_value")])
-#dim(lus_spread)
+
 lus_spread <- merge(lus_spread, truth, by=c("lineupid","true_model", "param_name"), all.x=TRUE)
 lus_spread$true_value[is.na(lus_spread$true_value)] <- 0
-#qplot(M2, true_value, facets=~param_name, data=lus_spread, colour=true_model)
-#qplot(M3, true_value, facets=~param_name, data=lus_spread, colour=true_model)
 
@@
     data_sd = sd(M2[true_model=="M2"], na.rm=TRUE),
-    null_est = list(M2[true_model=="M1"]),
-    data_dist = data_est - max(unlist(null_est))),
-  upper_qu = sum(data_est > unlist(null_est)) / (n()-1) )
-
-# XXX The lines commented out above crash my RStudio. I haven't figured out why yet! :( XXX
+    data_dist = data_est - max(M2[true_model=="M1"]),
+    upper_qu = sum(data_est > M2[true_model=="M1"]) / (n()-1) 
+)
 
@@
 
+lus_summary_2b <- lus_spread %>% dplyr::group_by(part, lineupid, param_name) %>%
+  dplyr::summarize(
+    data_n = sum(true_model=="M1"),
+    data_est = mean(M2[true_model=="M1"]),
+    data_sd = sd(M2[true_model=="M1"], na.rm=TRUE),
+    data_dist = data_est - min(M2[true_model=="M2"]),
+    lower_qu = sum(data_est < M2[true_model=="M2"]) / (n()-1) 
+)
+
+p2 <- qplot(data_dist, lower_qu, data = subset(lus_summary_2b, !is.na(data_dist) & param_name=="transitive triplets jumping alcohol2" & part %in% c("smallfriends", "smallfriends-rev"))) +
+  facet_grid(.~part) + xlab("Difference in M2 estimate of data panel and minimum of the null panels") +
+  ylab("Ratio of null panels with larger M2 estimate") +
+  ggtitle("jumping transitive triplets")
+
+gridExtra::grid.arrange(p1,p2)
+@
+\caption{\label{fig:lusparms} Scatterplots of smallfriends and smallfriends-rev: comparing the ratio of null plots with smaller estimates of $\beta_3$ in the nulls than in the data panel (top row) and larger estimates of $\beta_3$ in the nulls than in the data panel.
+In lineups with a ratio of 1 the data should be 'easy' to identify in both scenarios.}
+\end{figure}
+
+\begin{figure}
+\centering
+<<echo=FALSE, fig.width=8, fig.height = 8, out.width='.6\\linewidth'>>=
 lus_summary_3 <- lus_spread %>% group_by(part, lineupid, param_name) %>%
@@
     data_sd = sd(M3[true_model=="M3"], na.rm=TRUE),
-    null_est = list(M3[true_model=="M1"]),
-    data_dist = min(unlist(null_est)) - data_est,
-    lower_qu = sum(data_est < unlist(null_est))/(n()-1)
+    data_dist = min(M3[true_model=="M1"]) - data_est,
+    lower_qu = sum(data_est < M3[true_model=="M1"])/(n()-1)
   )
 
-p2 <- qplot(data_dist, lower_qu,  data = subset(lus_summary_3, !is.na(lower_qu) & param_name=="number pairs at doubly achieved distance 2")) +
+p1 <- qplot(data_dist, lower_qu,  data = subset(lus_summary_3, !is.na(lower_qu) & param_name=="number pairs at doubly achieved distance 2")) +
   facet_grid(.~part) +
@@
 
-gridExtra::grid.arrange(p1,p2)
+#gridExtra::grid.arrange(p1,p2)
+p1
 @
-\caption{\label{fig:lusparms} Scatterplots }
+\caption{\label{fig:lusparms-eff2} }
 \end{figure}

### 2016-07-07 lineups.Rnw
fix dplyr issue and calculate significance of beta3s

--- 2016-07-07 lineups.Rnw
@@
 
+<<est>>=
+#For model $M_2$, parameter $\beta_3$ is estimated to be significantly different from zero in almost all cases \hh{(XXX how many exactly? XXX)}.
+
+# get beta3s where true model and fit model are M2
+beta3sig <- lus_ests_truth %>% 
+  filter(true_model == "M2", model == "M2", 
+         param_name == "transitive triplets jumping alcohol2")
+convergebeta3 <- length(which(beta3sig$convergence == "Converged")) / nrow(beta3sig)
+
+# get ones that converged.
+beta3sig2 <- beta3sig %>% 
+  filter(convergence == "Converged") %>%
+  mutate(tstat = param_est/param_est_se,
+         ttestpval = 2*dt(tstat, df = 1))
+beta3sig05 <- sum(beta3sig2$ttestpval <= 0.05) / nrow(beta3sig2)
+beta3sig10 <- sum(beta3sig2$ttestpval <= 0.10) / nrow(beta3sig2)
+@
 Figure~\ref{fig:comparison} shows an overview of model estimates for each simulated data set.
@@
 For data fitted under model $M_2$ we see that parameter $\beta_3$ (jumping transitive triplets) are estimated to be about zero, if the data is generated from models $M_1$ and $M_3$.
-For model $M_2$, parameter $\beta_3$ is estimated to be significantly different from zero in almost all cases \hh{(XXX how many exactly? XXX)}. This coincides with our expectation.
+For model $M_2$, parameter $\beta_3$ is estimated to be significantly different from zero in \st{\Sexpr{round(100 * beta3sig10, 1)}\% of cases}. This coincides with our expectation.
 
@@
     null_est = list(M2[true_model=="M1"]),
-    data_dist = data_est - max(unlist(null_est))),
-  upper_qu = sum(data_est > unlist(null_est)) / (n()-1) )
+    data_dist = data_est - max(unlist(null_est)),
+    upper_qu = sum(data_est > unlist(null_est)) / (n()-1) 
+  )
 
-# XXX The lines commented out above crash my RStudio. I haven't figured out why yet! :( XXX
+# XXX okay, i updated dplyr and it works just fine. ugh. sorry for wasting time.
 

### 2016-07-07 lineups.Rnw

--- 2016-07-07 lineups.Rnw
@@
 
-p2 <- qplot(data_dist, lower_qu, data = subset(lus_summary_2b, !is.na(data_dist) & param_name=="transitive triplets jumping alcohol2" & part %in% c("smallfriends", "smallfriends-rev"))) +
-  facet_grid(.~part) + xlab("Difference in M2 estimate of data panel and minimum of the null panels") +
-  ylab("Ratio of null panels with larger M2 estimate") +
-  ggtitle("jumping transitive triplets")
-
-gridExtra::grid.arrange(p1,p2)
 @

### 2016-07-07 lineups.Rnw
add file for calculating difference in parameter estimates.

@@
 
-<<est>>=
+<<est, echo = FALSE>>=
 #For model $M_2$, parameter $\beta_3$ is estimated to be significantly different from zero in almost all cases \hh{(XXX how many exactly? XXX)}.

### 2016-07-07 Code/08_diff_in_lu_ests.R

@@ -1,0 +1,17 @@
+# computing differences between the model estimates in m1 v m2 situations
+
+load("Data/lus_ests_truth.rda")
+head(lus_ests_truth)
+m1m2 <- unique(lus_ests_truth$lineupname)[1:2]
+onlym1m2 <- lus_ests_truth %>% filter(lineupname %in% m1m2) %>% filter(model != "M3")
+
+onlym1m2wide <- spread(onlym1m2[,c("lineupid","model", "true_model", "panel_num", "param_name", "param_est", "convergence")], model, param_est)
+onlym1m2wide %>% mutate(diff = M2 - ifelse(is.na(M1), 0, M1)) -> onlym1m2wide
+
+ggplot(data = filter(onlym1m2wide, param_name == "transitive triplets jumping alcohol2")) + 
+  geom_density(aes(x = diff, fill = true_model), alpha = .7) + 
+  scale_fill_brewer(palette = "Set3") + 
+  facet_wrap(~convergence, scales = 'free') 
+
+# okay. jtt panel shows that when m2 is true model there is a mode around 3 for jtt coef but when true mod
+# is m1 mode is at zero for jtt. woot. 

### 2016-07-07 lineups.Rnw
added figure 7 output back in

@@
 
+p2 <- qplot(data_dist, lower_qu, data = subset(lus_summary_2b, part %in% c("smallfriends", "smallfriends-rev") & param_name=="transitive triplets jumping alcohol2")) +
+  facet_grid(.~part) + xlab("Difference in M2 estimate of data panel and minimum of the null panels") +
+  ylab("Ratio of null panels with larger M2 estimate") +
+  ggtitle("jumping transitive triplets")
+
+gridExtra::grid.arrange(p1,p2)
 @

### 2016-07-09 Code/09_checking_easy_lus.R
write some code to quickly check "easy" lineups.

@@ -1,0 +1,62 @@
+# "easy" lineups
+load("Data/lus_ests_truth.rda")
+
+lus_ests_truth$true_panel <- lus_ests_truth$true_model == lus_ests_truth$model
+lus_spread <- tidyr::spread(lus_ests_truth[,c("lineupid","model", "true_model", "panel_num", "param_name", "param_est")], model, param_est)
+
+# parameters should be most different between data plot and null plots
+
+# merge in true values:
+truth <- unique(subset(lus_ests_truth, true_panel==TRUE)[,c("lineupid", "true_model","param_name","true_value")])
+
+lus_spread <- merge(lus_spread, truth, by=c("lineupid","true_model", "param_name"), all.x=TRUE)
+lus_spread$true_value[is.na(lus_spread$true_value)] <- 0
+
+# compute a distance between the data panel and the closest null panel
+lus_spread$part <- gsub("(.*)-[0-9]*-[0-9]*", '\\1', lus_spread$lineupid)
+
+lus_summary_2 <- lus_spread %>% dplyr::group_by(part, lineupid, param_name) %>%
+  dplyr::summarize(
+    data_n = sum(true_model=="M2"),
+    data_est = mean(M2[true_model=="M2"]),
+    data_sd = sd(M2[true_model=="M2"], na.rm=TRUE),
+    data_dist = data_est - max(M2[true_model=="M1"]),
+    upper_qu = sum(data_est > M2[true_model=="M1"]) / (n()-1) 
+  )
+
+p1 <- qplot(data_dist, upper_qu, data = subset(lus_summary_2, !is.na(upper_qu) & param_name=="transitive triplets jumping alcohol2")) +
+  facet_grid(.~part) + xlab("Difference in M2 estimate of data panel and maximum of the null panels") +
+  ylab("Ratio of null panels with smaller M2 estimate") +
+  ggtitle("jumping transitive triplets")
+
+
+lus_summary_2b <- lus_spread %>% dplyr::group_by(part, lineupid, param_name) %>%
+  dplyr::summarize(
+    data_n = sum(true_model=="M1"),
+    data_est = mean(M2[true_model=="M1"]),
+    data_sd = sd(M2[true_model=="M1"], na.rm=TRUE),
+    data_dist = data_est - min(M2[true_model=="M2"]),
+    lower_qu = sum(data_est < M2[true_model=="M2"]) / (n()-1) 
[…]

### 2016-07-09 lineups.Rnw

--- 2016-07-07 lineups.Rnw
@@
 p1 <- qplot(data_dist, upper_qu, data = subset(lus_summary_2, !is.na(upper_qu) & param_name=="transitive triplets jumping alcohol2")) +
-  facet_grid(.~part) + xlab("Difference in M2 estimate of data panel and maximum of the null panels") +
+  facet_grid(.~part) + xlab("Difference in M2 estimate of data panel and maximum of the null panels (from M1)") +
   ylab("Ratio of null panels with smaller M2 estimate") +
@@
 
+p2 <- qplot(data_dist, lower_qu, data = subset(lus_summary_2b, part %in% c("smallfriends", "smallfriends-rev") & param_name=="transitive triplets jumping alcohol2")) +
+  facet_grid(.~part) + xlab("Difference in M2 estimate of data panel and minimum of the null panels (from M2)") +
+  ylab("Ratio of null panels with larger M2 estimate") +
+  ggtitle("jumping transitive triplets")
+
+gridExtra::grid.arrange(p1,p2)
 @

### 2016-07-10 JSM/VizInfSlides.Rmd
skeleton slides for JSM

---
title: "Visual Inference for Networks"
author: "Sam Tyner, Heike Hofmann"
date: "3 August 2016"
output: ioslides_presentation
---

## Outline

- Introduction
  * Visual Inference
  * Stochactic Actor-Oriented Models for Social Networks
- Simulation
- Experiment
- Results
- Future Work

# Introduction

## Visual Inference

## SAOMs

# Simulation

## RSiena

## ? SOMETHING ELSE ?

# Experiment

## Pilot Study (?)

## Procedure

# Results

## Summary

## Significance (?)

# Conclusion

## What we've learned

## Future Work

### 2016-07-11 lineups.Rnw
not feeling good today but still committed to the streak!
small amount more about estimation

@@
 
+\st{Because the likelihood function for these complicated models is intractable, RSiena implements a Monte Carlo simulation to obtain method of moments estimates of the parameter values.
+This fitting procedure was first introduced in ~\cite{saompaper}.
+}
+
 \st{The model fitting in RSiena is done in three phases.
@@
 Participants identified the data plot in two to four of the lineups they evaluated.
-The mode was three data identifications.
+The mode was three data identifications out of twenty per participant.
 }

### 2016-07-11 lineups.Rnw
doubly achieved distance figure

@@
 \centering
-<<echo=FALSE, fig.width=8, fig.height = 8, out.width='.6\\linewidth'>>=
+<<echo=FALSE, fig.width=8, fig.height = 8, out.width='.6\\linewidth', warning=FALSE>>=
 lus_summary_3 <- lus_spread %>% group_by(part, lineupid, param_name) %>%
@@
   facet_grid(.~part) +
-  xlab("Difference in M3 estimate of data panel and minimum of the null panels") +
+  xlab("Difference in M3 estimate of data panel and minimum of the null panels (M1 nulls)") +
   ylab("Ratio of null panels with larger M3 estimate") +
@@
 
-#gridExtra::grid.arrange(p1,p2)
-p1
+lus_summary_3b <- lus_spread %>% group_by(part, lineupid, param_name) %>%
+  summarize(
+    data_n = sum(true_model=="M1"),
+    data_est = mean(M3[true_model=="M1"]),
+    data_sd = sd(M3[true_model=="M1"], na.rm=TRUE),
+    data_dist = max(M3[true_model=="M3"]) - data_est,
+    upper_qu = sum(data_est > M3[true_model=="M3"])/(n()-1)
+  )
+
+p2 <- qplot(data_dist, upper_qu,  data = subset(lus_summary_3b, !is.na(upper_qu) & param_name=="number pairs at doubly achieved distance 2" &
+part %in% c("smallfriends-eff2", "smallfriends-eff2-rev"))) +
+  facet_grid(.~part) +
+  xlab("Difference in M3 estimate of data panel and minimum of the null panels (M3 nulls)") +
+  ylab("Ratio of null panels with smaller M3 estimate") +
+  ggtitle("doubly achieved distance")
+
+gridExtra::grid.arrange(p1,p2)
+#p1
 @

### 2016-07-12 JSM/VizInfSlides.Rmd
small amount more about estimation
start slides

@@
 
+## Lineup (?)
+
 ## SAOMs
+
+Goal: modeling dynamic networks using
+
+- Network effects (e.g. outdegree, reciprocity)
+- Actor covariate effects
+
+## Model "Rules"
+
+Observed networks treated as discrete observations of a continuous time Markov Chain.
+
+1. At each (unobserved) time point in the CTMC one actor can change one of its ties.
+2. Between network observations (say from $t_1$ to $t_2$) all actors are given the opportunity to change a tie at the same rate.
+3. Once an actor can change, it tries to maximize its objective functon, $f_i(x)$
+Parameters:
 
@@
 ## Future Work
+
+## References
+
+## Thank you!

### 2016-07-12 JSM/VizInfSlides.Rmd
more sliding

@@
 
-## Lineup (?)
+## Lineups
+
+$H_0$: "boring model" XX how specific? XX
+
+$H_1$: "alternate model consisting of one additional statistically significant parameter" XX how specific? XX
+
+Want: In a lineup, networks simulated from the $H_1$ model should appear so different from a series of models simulated from $H_0$ that an observer can pick out the different, alternative model.
 
@@
 2. Between network observations (say from $t_1$ to $t_2$) all actors are given the opportunity to change a tie at the same rate.
-3. Once an actor can change, it tries to maximize its objective functon, $f_i(x)$
-Parameters:
+3.
+Once an actor can change, it tries to maximize its objective functon, $f_i(\mathbf{\beta},x) = \sum_k \beta_k s_{ki}(x)$, where $\beta_k$ are network or actor parameters, and $s_{ki}(x)$ are corresponding statistics.
 
 # Simulation
+
+Using a subset of the friendship network data from Michell & Amos we:
+
+1. Fit three (XX two? XX) models 1000 times to obtain a distribution of parameter estimates
+2. Simulated the second wave from each of our models to form lineups. Means from 1 were used as starting parameter values.
+3. Estimated the parameters of each of our three (XX two? XX) models using the simulated second waves.
 

### 2016-07-14 JSM/VizInfSlides.Rmd
more sliding
model slides

@@
 
+## Model Parameters
+
+Effect name | Parameter | Corresponding Statistic
+------------|-----------|-------------------------
+Rate 1 (wave 1 $\rightarrow$ 2) | $\alpha_1$ | $\sum\limits_{i,j = 1 i \neq j}^n (x_{ij}(t_2) - x_{ij}(t_1))^2$  
+Rate 2 (wave 2 $\rightarrow$ 3) | $\alpha_2$ | $\sum\limits_{i,j = 1 i\neq j}^n (x_{ij}(t_3) - x_{ij}(t_2))^2$
+Outdegree | $\beta_1$ | $s_{i1}(x) = \sum\limits_{j=1}^n x_{ij}$ 
+Reciprocity | $\beta_2$ | $s_{i2}(x) = \sum\limits_{j=1}^n x_{ij}x_{ji}$ 
+Jumping Transitive Triplets | $\beta_3$ | $s_{i3}(x) = \sum\limits_{\forall j\neq h} x_{ij}x_{ih}x_{hj} \mathbb{I}(v_i = v_h \neq v_j)$
+
 # Simulation
+
+## RSiena
+
+R Simulation Investigation for Empirical Network Analysis
+
+Method of moment Monte Carlo estimation of the parameters in models
+
+## Procedure
 
@@
 
-## RSiena
-
 ## ? SOMETHING ELSE ?
@@
 ## Pilot Study (?)
+
+Help to find a measure of difficulty of a lineup
 

### 2016-07-14 lineups.Rnw

@@
 
+<<setup, fig.keep='all', cache=FALSE, echo=FALSE, eval=TRUE, message=F, warning=F>>=
+#rm(list=ls())
+#wd <- getwd()
+library(knitr)
+
+options(replace.assign=TRUE,scipen=3, digits=2)
+
+library(ggplot2)
+library(RColorBrewer)
+library(dplyr)
+library(gridExtra)
+theme_set(theme_bw())
+@
 \section{Introduction}
@@
 ggplot(data=lus_ests_truth) +
-  geom_bar(aes(x=model, fill=convergence), position="fill") + 
+  geom_bar(aes(x=model, fill=convergence), position="fill", alpha = 0.6) + 
   facet_grid(.~true_model, labeller="label_both") +
@@
     data_dist = data_est - max(M2[true_model=="M1"]),
-    upper_qu = sum(data_est > M2[true_model=="M1"]) / (n()-1) 
+    upper_qu = sum(data_est > M2[true_model=="M1"]) / (n()-1),
+    data_panel = mean(panel_num[true_model=="M2"])
 )
@@
 \end{figure}
+
+<<>>=
+jtts <- read.csv("Data/lineups-jtt.csv")
+jtts$lineupid <- with(jtts, paste(model, m, rep, sep="-"))
+jtts$data_panel <- jtts$plot_order
+lus_summary_2 <- merge(lus_summary_2, jtts[, c("lineupid", "jtt", "data_panel")], by=c("lineupid","data_panel"), all.x=TRUE)
+qplot(data=subset(lus_summary_2, param_name=="transitive triplets jumping alcohol2" & part=="smallfriends"), data_est, jtt, geom="jitter") + geom_label(data=subset(lus_summary_2, param_name=="transitive triplets jumping alcohol2" & jtt > 20 & part=="smallfriends"), aes(label=lineupid), alpha=.6)
+@
 

### 2016-07-20 Code/05b_merge_estimates_wtruth.R
fix the lus_ests_truth

@@ -109,13 +109,13 @@ true_params <- data.frame(model = rep(c("M1", "M2", "M3"), c(3,4,4)),
                                          "rate","outdegree (density)","reciprocity", "transitive triplets jumping alcohol2",
                                          "rate","outdegree (density)","reciprocity", "number pairs at doubly achieved distance 2"), 
                           true_value = starting_values[[3]])
-
+names(true_params)[1] <- "true_model"
 panels_and_truth <- merge(match_lu_long, true_params)
-names(panels_and_truth)[5] <- "panel_num"
+names(panels_and_truth)[6] <- "panel_num"
 names(lus_res_long)
-lus_ests_truth <- merge(lus_res_long, panels_and_truth)
+lus_ests_truth <- merge(lus_res_long, panels_and_truth, by = intersect(names(lus_res_long), names(panels_and_truth)))
 head(lus_ests_truth)
-
+lus_ests_truth <-lus_ests_truth[, -which(names(lus_ests_truth) == "X")]
 lus_ests_truth %<>% arrange(lineupname, M, rep, model, panel_num, param_name)
 write.csv(lus_ests_truth, file="Data/lus_ests_truth.csv", row.names=FALSE)
 

### 2016-07-20 lineups.Rnw
start on figure X

@@
 For models $M_2$ and $M_3$ data generated from the model converged at a higher rate than data generated from the other model.}
+\end{figure}
+
+\begin{figure}
+<<results0b, fig.width=8, fig.height=3, out.width='\\linewidth', echo = FALSE, warning = FALSE>>=
+load("Data/lus_ests_truth.rda")
+lus_ests_truth$param_name <- 
+  factor(lus_ests_truth$param_name,
+    levels = c("rate", "outdegree (density)", "reciprocity",
+               "transitive triplets jumping alcohol2",
+               "number pairs at doubly achieved distance 2"))        
+ggplot(data=lus_ests_truth) +
+  geom_density(aes(x=param_est-true_value, fill=convergence), alpha= 0.6) +
+  facet_grid(model+param_name~true_model, labeller="label_both") +
+  theme_bw() +
+  theme(legend.position="bottom") + 
+  scale_colour_brewer("Convergence", palette="Set1") +
+  xlab("Parameter Estimate") + xlim(c(-25,25))
+@
+\caption{\label{fig:convergence-ests}}
 \end{figure}

### 2016-07-20 Code/05b_merge_estimates_wtruth.R
whoops. uploaded the wrong filetype

@@ -169,7 +169,7 @@ for (i in 1:nrow(lus_ests_truth)){
   }
 }
 
-save(lus_ests_truth, file = "~/Desktop/NetworksResearch/NetworksVizInference/Data/lus_ests_truth.rda")
+save(lus_ests_truth, file = "~/Desktop/NetworksResearch/NetworksVizInference/Data/lus_ests_truth.RDA")
 table(lus_ests_truth$convergence, useNA = 'ifany')
 
 ggplot(data = lus_ests_truth) + geom_density(alpha = .8, aes(x = param_est, fill = true_model)) + 

### 2016-07-20 Code/05b_merge_estimates_wtruth.R
code to add zeros

@@ -110,6 +110,9 @@ true_params <- data.frame(model = rep(c("M1", "M2", "M3"), c(3,4,4)),
                                          "rate","outdegree (density)","reciprocity", "number pairs at doubly achieved distance 2"), 
                           true_value = starting_values[[3]])
 names(true_params)[1] <- "true_model"
+true_params <- rbind(true_params, data.frame(true_model = c("M1", "M1", "M2", "M3"), 
+                                             param_name = c("transitive triplets jumping alcohol2", "number pairs at doubly achieved distance 2", "number pairs at doubly achieved distance 2", "transitive triplets jumping alcohol2"), 
+                                             true_value = rep(0,4)))
 panels_and_truth <- merge(match_lu_long, true_params)
 names(panels_and_truth)[6] <- "panel_num"
 names(lus_res_long)
@@ -175,3 +178,7 @@ table(lus_ests_truth$convergence, useNA = 'ifany')
 ggplot(data = lus_ests_truth) + geom_density(alpha = .8, aes(x = param_est, fill = true_model)) + 
   facet_grid(convergence~param_name, scales = 'free', labeller = "label_both") 
 
+# create zeros for the parameters that don't exist. 
+head(data.frame(lus_ests_truth %>% spread(model, param_est)), 100)
+
+head(data.frame(lus_ests_truth %>% filter(model == "M1" & true_model == "M2")))

### 2016-07-20 lineups.Rnw
new figure 6

--- 2016-07-14 lineups.Rnw
@@
 For models $M_2$ and $M_3$ data generated from the model converged at a higher rate than data generated from the other model.}
+\end{figure}
+
+\begin{figure}
+<<results0b, fig.width=8, fig.height=10, out.width='\\linewidth', echo = FALSE, warning = FALSE>>=
+load("Data/lus_ests_truth.rda")
+lus_ests_truth$param_name <- 
+  factor(lus_ests_truth$param_name,
+    levels = c("rate", "outdegree (density)", "reciprocity",
+               "transitive triplets jumping alcohol2",
+               "number pairs at doubly achieved distance 2"))        
+lus_ests_truth$model_label <- lus_ests_truth$model
+lus_ests_truth$true_model_label <- sprintf("True~Model:~%s", lus_ests_truth$true_model)
+lus_ests_truth$param_label <- c("alpha", "beta[1]", "beta[2]", "beta[3]", "beta[4]")[as.numeric(lus_ests_truth$param_name)]
+ggplot(data=lus_ests_truth %>% filter(!is.na(convergence))) +
+  geom_rect(xmin=-25, xmax=25, ymin=-0.5, ymax = 1, fill=rgb(.25,.25,.25,alpha=.25), data=unique(subset(lus_ests_truth, (true_model==model) & (!is.na(param_est)) )[,c("true_model_label", "model_label", "param_label")]) ) +
+  geom_vline(xintercept = 0, colour="grey50") +
+  geom_density(aes(x=param_est-true_value, fill=convergence), 
+               alpha= 0.6) +
+  facet_grid(model_label+param_label~true_model_label, drop=TRUE, labeller=label_parsed) +
+  theme_bw() +
+  theme(legend.position="bottom", 
+        strip.text.y = element_text(angle=0)) + 
+  scale_fill_brewer("Convergence", palette="Set1") +
+  xlab("Difference between parameter estimate and true value") + 
+    xlim(c(-15,15))
+@
+\caption{\label{fig:convergence-ests} Difference between parameter estimate and true value.
+Densities are drawn for both converged and non-converged data.
+Panels with a light grey background show model fits with data sampled from the same model.
+}
 \end{figure}

### 2016-07-20 lineups.Rnw
added some explanation

--- 2016-07-14 lineups.Rnw
@@
 For models $M_2$ and $M_3$ data generated from the model converged at a higher rate than data generated from the other model.}
+\end{figure}
+
+\begin{figure}
+<<results0b, fig.width=8, fig.height=8, out.width='\\linewidth', echo = FALSE, warning = FALSE>>=
+load("Data/lus_ests_truth.rda")
+lus_ests_truth$param_name <- 
+  factor(lus_ests_truth$param_name,
+    levels = c("rate", "outdegree (density)", "reciprocity",
+               "transitive triplets jumping alcohol2",
+               "number pairs at doubly achieved distance 2"))        
+lus_ests_truth$model_label <- lus_ests_truth$model
+lus_ests_truth$true_model_label <- sprintf("True~Model:~%s", lus_ests_truth$true_model)
+lus_ests_truth$param_label <- c("alpha", "beta[1]", "beta[2]", "beta[3]", "beta[4]")[as.numeric(lus_ests_truth$param_name)]
+ggplot(data=lus_ests_truth %>% 
+         filter(
+           !is.na(convergence),
+           (true_model=="M3" & model != "M2") | (true_model=="M1") |
+             (true_model=="M2" & model != "M3")
+           )) +
+  geom_rect(xmin=-25, xmax=25, ymin=-0.5, ymax = 1, fill=rgb(.25,.25,.25,alpha=.25), data=unique(subset(lus_ests_truth, (true_model==model) & (!is.na(param_est)) )[,c("true_model_label", "model_label", "param_label")]) ) +
+  geom_vline(xintercept = 0, colour="grey50") +
+  geom_density(aes(x=param_est-true_value, fill=convergence), 
+               alpha= 0.6) +
+  facet_grid(model_label+param_label~true_model_label, drop=TRUE, labeller=label_parsed) +
+  theme_bw() +
+  theme(legend.position="bottom", 
+        strip.text.y = element_text(angle=0)) + 
+  scale_fill_brewer("Convergence", palette="Set1") +
+  xlab("Difference between parameter estimate and true value") + 
+    xlim(c(-15,15))
+@
+\caption{\label{fig:convergence-ests} Difference between parameter estimate and true value. 
+Panels with a light grey background show model fits with data sampled from the same model.
+Densities are drawn for both converged and non-converged data.
+The red-filled densities should have a mode in zero, indicating that the model converged in the correct value.
+For data from model $M_1$ estimates for parameters $\beta_3$ and $\beta_4$ converge to a wrong value.
+}
 \end{figure}

### 2016-08-02 JSM/VizInfSlides.Rmd
not sure what i changed.
slides

--- 2016-07-12 JSM/VizInfSlides.Rmd
@@
 - Introduction
-  * Visual Inference
-  * Stochactic Actor-Oriented Models for Social Networks
-- Simulation
-- Experiment
+- Stochactic Actor-Oriented Models for Social Networks
+- Visual Inference Experiment
 - Results
@@
 
-# Introduction
+## Friendship Network
 
-## Visual Inference
+```{r, echo = FALSE, message=FALSE, warning=FALSE}
+library(sna)
+library(network)
+friend.data.w1 <- as.matrix(read.table("../Data/s50_data/s50-network1.dat"))
+friend.data.w2 <- as.matrix(read.table("../Data/s50_data/s50-network2.dat"))
+fd2.w1 <- friend.data.w1[20:35,20:35]
+fd2.w2 <- friend.data.w2[20:35,20:35]
+drink <- as.matrix(read.table("../Data/s50_data/s50-alcohol.dat"))
+drink2 <- drink[20:35,]
+actual1 <- merge(data.frame(as.edgelist(as.network(fd2.w1))), 
+                 data.frame(id = 1:16, drink = drink2[,1]), 
+                 by.x = "X1", by.y = "id", all = T)
+for (j in 1:nrow(actual1)){
+      if (!(actual1$X1[j] %in% actual1$X2) & is.na(actual1$X2[j])){
+        actual1$X2[j] <- actual1$X1[j]
+      } else {actual1$X2[j] <- actual1$X2[j]}
+}
+actual2 <- merge(data.frame(as.edgelist(as.network(fd2.w2))), 
+                 data.frame(id = 1:16, drink = drink2[,2]), 
+                 by.x = "X1", by.y = "id", all = T)
+for (j in 1:nrow(actual2)){
+      if (!(actual2$X1[j] %in% actual2$X2) & is.na(actual2$X2[j])){
+        actual2$X2[j] <- actual2$X1[j]
+      } else {actual2$X2[j] <- actual2$X2[j]}
+    }
+actual1$wave <- 1
+actual2$wave <- 2
 
-## Lineup (?)
+waves <- rbind(actual1, actual2)
+library(geomnet)
+actual1$behaviour <- factor(actual1$drink)
+levels(actual1$behaviour) <- c("None", "Once or twice a year", "Once a month", "Once a week")
+set.seed(34567)
+ggplot(data = actual1, aes(from_id = X1, to_id = X2)) + 
+  geom_net(label = TRUE, hjust = 0.5, vjust=0.5, size=9,  
+          labelcolour = "grey20", directed = T, arrowgap = .04,
+           aes(colour = behaviour)) + 
+  scale_colour_brewer("Drinking behavior", palette="YlOrRd") +
+  theme_net()
 
-## SAOMs
+#ggplot(data = waves, aes(from_id = X1, to_id = X2)) + 
+#  geom_net(label = TRUE, hjust = -.5, labelcolour = 'black', fiteach = T,
+#           aes(color = as.factor(drink))) + 
+#  theme_net() + theme(panel.background = element_rect(fill = "white", color = 'black')) + facet_wrap(~wave)
+```
 
-Goal: modeling dynamic networks using
+## Stochastic Actor Oriented Models
+
+Goal: model dynamic networks with
 
@@
 
-## Model "Rules"
-
 Observed networks treated as discrete observations of a continuous time Markov Chain.
 
-1. At each (unobserved) time point in the CTMC one actor can change one of its ties.
-2. Between network observations (say from $t_1$ to $t_2$) all actors are given the opportunity to change a tie at the same rate.
-3. Once an actor can change, it tries to maximize its objective functon, $f_i(x)$
-Parameters:
+1. At each (unobserved) time point 1 actor can change 1 tie
+2. Between observations all actors are given the opportunity to change at same rate
+3.
+When an actor can change, aims to maximize its objective functon, $f_i(\mathbf{\beta},x) = \sum_k \beta_k s_{ki}(x)$, where $\beta_k$ are network or actor parameters, and $s_{ki}(x)$ are corresponding statistics.
 
-# Simulation
+## Model Parameters
 
-## RSiena
+1. Rate - how often actors change ties 
+2. Outdegree - measure of activeness
+3. Reciprocity - how often ties are reciprocated
+4. Jumping Transitive Triplets
+
+```{r, echo = FALSE}
+jTTe <- data.frame(from = c('i', 'i', 'h'), to = c('h', 'j', 'j'))
+jTTn <- data.frame(id = letters[8:10], group = c(1,1,2))
+
+jTT <- merge(jTTe, jTTn, by.x = 'from', by.y = "id", all = T)
+
+set.seed(12345) 
+ggplot(data = jTT, aes(from_id = from, to_id = to)) + 
+  geom_net(aes(shape = as.factor(group)), directed = T, label = T, 
+           labelcolour='grey80',vjust = 0.5, hjust =0.5, arrowgap = .15, 
+           colour = 'black', size=10, 
+           ecolour = c("red", "grey40", "grey40", "grey40")) + 
+  expand_limits(x=c(-0.1,1.1), y=c(-0.1,1.1)) +
+  theme_net() +
+  theme(legend.position = "none")
+```
+
+## Model Parameters {.smaller}
+
+```{r echo=FALSE, eval=FALSE}
+A<- c(1, rep(0,4))
+
+library(RSienaTest)
+Multipar.RSiena(runs_models_smallFriends[[1]], 1)
+Multipar.RSiena(runs_models_smallFriends[[1]], 2)
+```
+
+Effect name | Parameter | Corresponding Statistic | M1 | M2 | p-value M1 | p-value M2 |
+:------------:|:-----------:|:-------------------------:|:----:|:----:|:--------:|:----:
+Rate 1 (wave 1 $\rightarrow$ 2) | $\alpha_1$ | $\sum\limits_{i,j = 1 i \neq j}^n (x_{ij}(t_2) - x_{ij}(t_1))^2$  | **X** | **X** | -- | --
+Rate 2 (wave 2 $\rightarrow$ 3) | $\alpha_2$ | $\sum\limits_{i,j = 1 i\neq j}^n (x_{ij}(t_3) - x_{ij}(t_2))^2$  | **X** | **X** | -- | --
+Outdegree | $\beta_1$ | $s_{i1}(x) = \sum\limits_{j=1}^n x_{ij}$ | **X** | **X** | $5.67 \times 10^{-8}$ | $4.66 \times 10^{-10}$
+Reciprocity | $\beta_2$ | $s_{i2}(x) = \sum\limits_{j=1}^n x_{ij}x_{ji}$ | **X** | **X** | $1.10 \times 10^{-5}$ | $4.42 \times 10^{-7}$
+Jumping Transitive Triplets | $\beta_3$ | $s_{i3}(x) = \sum\limits_{\forall j\neq h} x_{ij}x_{ih}x_{hj} \mathbb{I}(v_i = v_h \neq v_j)$ | | **X** | -- | $0.00036$
+
+## Can you see the difference?
+
+## Procedure
+
+Using a subset of the friendship network data from Michell & Amos we:
+
+1. Fit three (XX two? XX) models 1000 times to obtain a distribution of parameter estimates
+2. Simulated the second wave from each of our models to form lineups. Means from 1 were used as starting parameter values.
+3. Estimated the parameters of each of our three (XX two? XX) models using the simulated second waves.
 
@@
 ## Pilot Study (?)
+
+Help us find a measure of difficulty of a lineup
 

### 2016-08-02 Code/03c_complete_lineup_creation.R

@@ -70,6 +70,8 @@ create_smfriend_lu <- function(null_eff_struct, test_eff_struct, M, my_dat=mysma
   data_plot_id <- unique(to_plot$plot_order[which(to_plot$count == M)])
   return(list(data = to_plot, lineup = plot, test_id = data_plot_id, null_mod_ests = c(null_net$rate,null_net$theta), alt_mod_estimates = c(test_net$rate,test_net$theta)))
 }
+
+
 # lu1 <- create_smfriend_lu(null_eff_struct = null_model_eff2, test_eff_struct = eff_models_smallFriends[[39]], M = 6)
 # savepdf(file = "Lineup-Images/pdfs/testcropping", width = 7.2*2.54, height = 4.5*2.54)
 # print(lu1$lineup)
@@ -133,13 +135,16 @@ make_interactive_lineups <- function(M, reps, model.effects, null.effects){
     #ggsave(lineup$lineup, file=sprintf("Lineup-Images/pdfs/smallfriends-m-%s-rep-%s.pdf",M,r), width = 7.2, height = 4.5, units = "in")
     savepdf(file = sprintf("Lineup-Images/pdfs/smallfriends-m-%s-rep-%s",M,r), width = 7.2*2.54, height = 4.5*2.54)
     print(lineup$lineup)
-    dev.off()
     tmpfile <- sprintf("%s.svg",tempfile(tmpdir="Lineup-Images/svgs"))
     make_interactive(filename2 = tmpfile, script=scriptURL,  
                      high="#d5d5d5",  background="#ffffff")
+    dev.off()
   }
   return(list(lineup = lineup))
 }
+
+#make_interactive_lineups(M = 4, reps=3, model.effects = null_model_eff2, null.effects = eff_models_smallFriends[[39]])
+
 
 #make_interactive_lineups(M = 3, reps=20, model.effects = null_model_eff2, null.effects = eff_models_smallFriends[[39]])
 

### 2016-08-02 Code/09_checking_easy_lus.R

@@ -24,7 +24,24 @@ lus_summary_2 <- lus_spread %>% dplyr::group_by(part, lineupid, param_name) %>%
     upper_qu = sum(data_est > M2[true_model=="M1"]) / (n()-1) 
   )
 
-p1 <- qplot(data_dist, upper_qu, data = subset(lus_summary_2, !is.na(upper_qu) & param_name=="transitive triplets jumping alcohol2")) +
+# smallfriends rev 
+
+data.frame(lus_summary_2 %>% 
+  filter(part == "smallfriends-rev", param_name == "transitive triplets jumping alcohol2", upper_qu == 0) %>% 
+  dplyr::select(part, lineupid, param_name, data_dist, upper_qu))
+data.frame(lus_summary_2 %>% 
+             filter(part == "smallfriends-rev", param_name == "transitive triplets jumping alcohol2", upper_qu > 0, upper_qu<.25, data_dist < 1) %>% 
+             dplyr::select(part, lineupid, param_name, data_dist, upper_qu))
+
+data.frame(lus_summary_2 %>% 
+             filter(part == "smallfriends-eff2", param_name == "number pairs at doubly achieved distance 2", upper_qu == 1) %>% 
+             dplyr::select(part, lineupid, param_name, data_dist, upper_qu))
+data.frame(lus_summary_2 %>% 
+             filter(part == "smallfriends-eff2", param_name == "transitive triplets jumping alcohol2", upper_qu > 0, upper_qu<.25, data_dist < 1) %>% 
+             dplyr::select(part, lineupid, param_name, data_dist, upper_qu))
+
+
+p1 <- qplot(data_dist, upper_qu, data = subset(lus_summary_2, !is.na(upper_qu) & param_name=="transitive triplets jumping alcohol2"), color = upper_qu) +
   facet_grid(.~part) + xlab("Difference in M2 estimate of data panel and maximum of the null panels") +
   ylab("Ratio of null panels with smaller M2 estimate") +
   ggtitle("jumping transitive triplets")
@@ -41,11 +58,23 @@ lus_summary_2b <- lus_spread %>% dplyr::group_by(part, lineupid, param_name) %>%
 
 
 # which are 1 ? should be easy! 
+easy_ids <- filter(lus_summary_2, !is.na(upper_qu) & 
[…]

### 2016-08-02 Code/10_weird_models.R

@@ -137,28 +137,15 @@ get_fit_info <- function(dat){
 }
 
 weirdones3 %>% mutate(fit_info = map(data, get_fit_info)) -> weirdones_refit
-
-
+weirdones_refit[weirdones_refit$lineupid == "smallfriends-6-17","fit_info"][[1]]
+lus_ests_truth[lus_ests_truth$lineupid == "smallfriends-6-17",]
 # okay run the code from HH on the weird ones refit. 
 # 
-library(geomnet)
 
-get_jtts <- function(dat){
-  lineupname <- dat$lineupname
-  M <- dat$M
-  rep <- dat$rep
-  filename <- paste0("Data/lineupdata/", lineupname, "-m-", M, "-rep-",rep, ".csv")
-  lu_dat <- read.csv(filename)
-  jtts <- data.frame(plot_order = unique(lu_dat$plot_order), jtt = NA)
-  for(i in 1:M){
-    sub_dat <- subset(lu_dat, plot_order == i)
-    my_jtt <- jtt(data = sub_dat, from_id = "X1", to_id = "X2")
-    jtts[i,"jtt"] <- my_jtt
-  }
-  return(jtts)
-}
-# won't work.......... :'(
-# Error in if (n < 0) stop("Network objects cannot be of negative order.") : 
-# missing value where TRUE/FALSE needed
+library(geomnet)
+# redo jtts 
+dframe <- data.frame(dframe)
+dframe$filename <- lineupdata[as.numeric(dframe$filename)]
+
 
 weirdones_refit %>% mutate(jtts = map(data, get_jtts)) -> weirdones_refit

### 2016-08-02 JSM/VizInfSlides.Rmd
push changes

@@
 
-1. Rate - how often actors change ties 
+1. Rate - how often girls change friendships 
 2. Outdegree - measure of activeness
-3. Reciprocity - how often ties are reciprocated
-4. Jumping Transitive Triplets
+3. Reciprocity - how often friendships are reciprocated
+4. Jumping Transitive Triplets - friends of the same person become friends
 
-```{r, echo = FALSE}
+```{r, echo = FALSE, fig.height=2}
 jTTe <- data.frame(from = c('i', 'i', 'h'), to = c('h', 'j', 'j'))
@@
 
-## Procedure
+```{r echo=FALSE}
+dat <- readr::read_csv("../Lineup-Images/ForExperimentTraining/datafiles/smallfriends-rev-m-12-rep-c.csv")
+ggplot(data = dat %>% filter(plot_order2 %in% c(11,12)) %>% mutate(plot = plot_order2 - 10) ) + 
+  geom_net(aes(from_id = X1, to_id = X2), size = 2, fiteach = T, directed = T, 
+           colour = 'black', arrowsize = .25, arrowgap = .02, ecolour = 'grey60') + 
+  facet_wrap(~plot) + 
+  theme_net()
+```
 
-Using a subset of the friendship network data from Michell & Amos we:
+## Can you see the difference?
 
-1. Fit three (XX two? XX) models 1000 times to obtain a distribution of parameter estimates
-2. Simulated the second wave from each of our models to form lineups. Means from 1 were used as starting parameter values.
-3. Estimated the parameters of each of our three (XX two? XX) models using the simulated second waves.
+```{r echo=FALSE}
+dat <- readr::read_csv("../Lineup-Images/ForExperimentTraining/datafiles/smallfriends-m-12-rep-1.csv")
+ggplot(data = dat %>% filter(plot_order %in% c(2,7)) %>% mutate(plot = ifelse(plot_order == 2, 1, 2))) + 
+  geom_net(aes(from_id = X1, to_id = X2), size = 2, fiteach = T, directed = T, 
+           colour = 'black', arrowsize = .25, arrowgap = .02, ecolour = 'grey60') + 
+  facet_wrap(~plot) + 
+  theme_net()
+```
 
-## ? SOMETHING ELSE ?
+## Question
 
-# Experiment
+Can we *see* a significant parameter when visualizing networks simulated from two different models?
 
-## Pilot Study (?)
+- Idea: a very statistically significant difference in models should look different to us
 
-Help us find a measure of difficulty of a lineup
+## Experiment
 
-## Procedure
-
-# Results
+- Small Amazon Turk experiment
 

### 2016-08-03 JSM/VizInfSlides.Rmd
2 more slides

@@
 
-Goal: model dynamic networks with
+Observed networks: discrete observations of continuous time Markov Chain.
 
-- Network effects (e.g. outdegree, reciprocity)
-- Actor covariate effects
-
-Observed networks treated as discrete observations of a continuous time Markov Chain.
-
-1. At each (unobserved) time point 1 actor can change 1 tie
-2. Between observations all actors are given the opportunity to change at same rate
-3.
-When an actor can change, aims to maximize its objective functon, $f_i(\mathbf{\beta},x) = \sum_k \beta_k s_{ki}(x)$, where $\beta_k$ are network or actor parameters, and $s_{ki}(x)$ are corresponding statistics.
+1. Between observations:  actors act (i.e. change ties with others) at same rate
+2. Actor aims to maximize  objective functon, $f_i(\mathbf{\beta},x) = \sum_k \beta_k s_{ki}(x)$.
 
@@
 
-## Model Parameters {.smaller}
+## Two Models {.smaller}
 
@@
 dat <- readr::read_csv("../Lineup-Images/ForExperimentTraining/datafiles/smallfriends-rev-m-12-rep-c.csv")
-ggplot(data = dat %>% filter(plot_order2 %in% c(11,12)) %>% mutate(plot = plot_order2 - 10) ) + 
+ggplot(data = dat %>% filter(plot_order2 %in% c(11,12)) %>% mutate(Model = plot_order2 - 10) ) + 
   geom_net(aes(from_id = X1, to_id = X2), size = 2, fiteach = T, directed = T, 
            colour = 'black', arrowsize = .25, arrowgap = .02, ecolour = 'grey60') + 
-  facet_wrap(~plot) + 
-  theme_net()
+  facet_wrap(~Model, labeller = "label_both") + 
+  theme_net() + theme(panel.background = element_rect(color = 'black'))
 ```
@@
 dat <- readr::read_csv("../Lineup-Images/ForExperimentTraining/datafiles/smallfriends-m-12-rep-1.csv")
-ggplot(data = dat %>% filter(plot_order %in% c(2,7)) %>% mutate(plot = ifelse(plot_order == 2, 1, 2))) + 
+ggplot(data = dat %>% filter(plot_order %in% c(2,7)) %>% mutate(Model = ifelse(plot_order == 2, 2, 1))) + 
   geom_net(aes(from_id = X1, to_id = X2), size = 2, fiteach = T, directed = T, 
            colour = 'black', arrowsize = .25, arrowgap = .02, ecolour = 'grey60') + 
-  facet_wrap(~plot) + 
-  theme_net()
+  facet_wrap(~Model, labeller = "label_both") + 
+  theme_net() + theme(panel.background = element_rect(color = 'black'))
 ```
@@
 - Idea: a very statistically significant difference in models should look different to us
+- Method: visual inference - use lineups with graphs from models 1 & 2 to test this idea
+
+## Lineup
+
+- Show $m$ plots to people
+- Ask to identify most different plot
+- Null ($m-1$ plots) vs. Alternative ($1$ plot) models
+- Identification of the alternative plot is evidence against null
+- Probability to pick alternative plot randomly is $\frac{1}{m}$
 
@@
 
-- Small Amazon Turk experiment
+- M1 v. M2 (most complex structure)
+- M2 v. M1 (least complex structure)
+- five lineups each
+- each participant randomly sees ten lineups
 
-## Summary
+## Most complex?
 
-## Significance (?)
+```{r, echo=FALSE, fig.align='center'}
+dat <- readr::read_csv("../Lineup-Images/ForExperimentTraining/datafiles/smallfriends-m-12-rep-10.csv")
+ggplot(data = dat) +
+  geom_net(aes(from_id = X1, to_id = X2), size = 1, fiteach = T, directed = T, 
+           colour = 'black', arrowsize = .15, arrowgap = .02, ecolour = 'grey60') + 
+  facet_wrap(~plot_order) + 
+  theme_net() + theme(panel.background = element_rect(color = 'black'))
+```
 
-# Conclusion
+## Least Complex?
 
-## What we've learned
+```{r echo=FALSE, fig.align='center'}
+dat <- readr::read_csv("../Lineup-Images/ForExperimentTraining/datafiles/smallfriends-rev-m-12-rep-b.csv")
+ggplot(data = dat) +
+  geom_net(aes(from_id = X1, to_id = X2), size = 1, fiteach = T, directed = T, 
+           colour = 'black', arrowsize = .15, arrowgap = .02, ecolour = 'grey60') + 
+  facet_wrap(~plot_order2) + 
+  theme_net() + theme(panel.background = element_rect(color = 'black'))
+```
 
-## Future Work
+## Small Amazon Turk experiment
 
-## References
+<img src="screenshot.png" width="800px" height="400px" />
 
-## Thank you!
+## Results 1
+
+``` {r, echo=FALSE}
+tab <- readr::read_csv("../Results/turk21_users.csv")
+res <- readr::read_csv("../Results/turk21_feedback.csv")
+```
+
+- barchart with filled correct
+
+## Results 2
+
+- scatterplot of %correct by beta 3
+- scatterplot of %correct by #jtt
+
+## Conclusion
+
+- Doesn't work for networks
+- Need another way! (layout?)

### 2016-08-23 Results/TurkExp1WriteUp.Rnw
write-up of experiment.

--- 2016-07-20 lineups.Rnw
@@
 \begin{document}
-\lhead{Graph Lineups}
+\lhead{Amazon Turk Experiment 1}
 \chead{}
@@
 
-<<setup, fig.keep='all', cache=FALSE, echo=FALSE, eval=TRUE, message=F, warning=F>>=
-#rm(list=ls())
-#wd <- getwd()
-library(knitr)
-
-options(replace.assign=TRUE,scipen=3, digits=2)
-
-library(ggplot2)
-library(RColorBrewer)
-library(dplyr)
-library(gridExtra)
-theme_set(theme_bw())
-@
 \section{Introduction}
 
+% ALTERNATE OPENER: Stochastic actor-oriented models (SAOMs) are used to model dynamic social networks.
+In these models, there are three different sets of parameters: rate, structural, and covariate.
+The rate parameters dictate how often the actors in the network change ties; the structural parameters dictate how ties change based solely on the network structure; the covariate parameters dictate how ties change based on actor characteristics.
+The many parameters in the model
+
+% UNDERLYING THEORY
 Stochastic actor-oriented models (SAOMs) are a set of network models that consider both network and actor effects when determining how and why a network changes over time.
@@
 
-where $k = 1, \dots, K$, $K$ the number of non-rate parameters in the models and $s{ki}(x)$ is the corresponding statistic.
+where $k = 1, \dots, K$, $K$ the number of non-rate parameters in the models and $s_{ki}(x)$ is the corresponding statistic.
 There are several parameters that can be chosen for use in the objective function.
@@
 
-\section{Models}
+% BACKGROUND & PURPOSE
+The parameters in the objective function are tested for significance using $t$-tests.
+The test statistic is the ratio of the parameter estimate to its standard error.
+If this value is larger than 2 in absolute value, then the parameter is significant at the $\alpha = 0.05$ level and should be included in the model.
+This is a fairly simple statistical test, so we wanted to test whether this significance can be detected visually just as simply as the statistical test detects it.
+If visualizations of simulated networks from two nested models have a much different appearance when placed side-by-side, then the difference in appearance can be attributed to the additional parameter in one.
+If, however, there is no visually detectable difference, then the additional parameter does not appear to have changed the network structure all that much.
+Because model selection and diagnostic in network analysis are less developed areas of the theory, testing network parameters in this visual way could lead to additional methods of model selection for networks.
 
-We fit \st{three} different stochastic actor-oriented models to a subset of the 50 actor dataset from the ``Teenage Friends and Lifestyle Study" that is provided on the RSiena webpage.\footnote{\url{http://www.stats.ox.ac.uk/~snijders/siena/s50_data.htm}} We chose to subset the data to decrease the cognitive load on our experiment's subjects.
-The subset contained actors 20 through 35 and the ties between them, as well as the drinking behavior of each actor at each of the three waves.
-This specific subset was chosen because it showed somewhat higher connectivity than other subsets, as we've emphasized in the visualizations of the three network adjacency matrices below.
-For model fitting, we condition on wave 1 and estimate the parameters of our models from the second and third waves.
+\section{Methods}
+In order to test our hypothesis, we set up an amazon turk experiment using the lineup protocal of {\color{green} Buja et all (2009). FIX CITE}.
 
-\begin{figure}
-<<get_sm_friends, echo=FALSE, fig.width=8, fig.height=3, message = FALSE, warning = FALSE>>=
-source("Code/00e_small_friends.R")
-view_rs2$wave <- factor(view_rs2$wave)
-levels(view_rs2$wave) <- c("Wave 1", "Wave 2", "Wave 3")
-ggplot(data= view_rs2, aes(x = x, y=y)) + 
-  geom_tile(aes(fill = as.factor(value))) + 
-  scale_fill_manual("X likes Y", labels=c("no", "yes"), 
-                    values = c('white', 'black')) +
-  geom_rect(data= NULL, inherit.aes = FALSE, color = 'red',
-            aes(xmin = 20, xmax = 35, ymin = 20, ymax = 35, fill =NA)) + 
-  facet_wrap(~wave) + theme_bw() +
-  theme(aspect.ratio=1, axis.text = element_blank(), axis.ticks = element_blank()) 
-@
-\caption{\label{fig:smallfriends}Adjacency matrices describing friendship relations between 50 students in waves 1,2, and 3 of the friendship study \citep{friendsdata}.
-The red squares identify the subset we focus on in the remainder.}
-\end{figure}
-
-The first wave of the network, which is conditioned on in estimation, is given in Figure~\ref{fig:wave1}.
-
-\begin{figure}
-\centering
-<<wave1, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, out.width='.6\\linewidth'>>=
-library(sna)
-library(network)
-actual1 <- merge(data.frame(as.edgelist(as.network(fd2.w1))), 
-                 data.frame(id = 1:16, drink = drink2[,1]), 
-                 by.x = "X1", by.y = "id", all = T)
-for (j in 1:nrow(actual1)){
-      if (!(actual1$X1[j] %in% actual1$X2) & is.na(actual1$X2[j])){
-        actual1$X2[j] <- actual1$X1[j]
-      } else {actual1$X2[j] <- actual1$X2[j]}
-}
-actual2 <- merge(data.frame(as.edgelist(as.network(fd2.w2))), 
-                 data.frame(id = 1:16, drink = drink2[,2]), 
-                 by.x = "X1", by.y = "id", all = T)
-for (j in 1:nrow(actual2)){
-      if (!(actual2$X1[j] %in% actual2$X2) & is.na(actual2$X2[j])){
-        actual2$X2[j] <- actual2$X1[j]
-      } else {actual2$X2[j] <- actual2$X2[j]}
-    }
-actual1$wave <- 1
-actual2$wave <- 2
-
-waves <- rbind(actual1, actual2)
-library(geomnet)
-actual1$behaviour <- factor(actual1$drink)
-levels(actual1$behaviour) <- c("None", "Once or twice a year", "Once a month", "Once a week")
-ggplot(data = actual1, aes(from_id = X1, to_id = X2)) + 
-  geom_net(label = TRUE, hjust = 0.5, vjust=0.5, size=10,  
-           fiteach = T, labelcolour = "grey20", 
-           aes(colour = behaviour)) + 
-  scale_colour_brewer("Drinking behavior", palette="YlOrRd") +
-  theme_net()
-
-#ggplot(data = waves, aes(from_id = X1, to_id = X2)) + 
-#  geom_net(label = TRUE, hjust = -.5, labelcolour = 'black', fiteach = T,
-#           aes(color = as.factor(drink))) + 
-#  theme_net() + theme(panel.background = element_rect(fill = "white", color = 'black')) + facet_wrap(~wave)
-@
-\caption{\label{fig:wave1} Network of friendships of wave 1 of the subset of students marked in Figure~\ref{fig:smallfriends}. }
-\end{figure}
-
-<<modelfit, echo=FALSE, results = 'hide', message=FALSE, warning = FALSE, cache = T>>=
-source("Code/03c_complete_lineup_creation.R")
-lineup1 <- create_smfriend_lu(null_eff_struct = null_model_eff2, test_eff_struct = eff_models_smallFriends[[39]], M = 3)
-@
-
-The two models we fit are a ``null model" (model $M_1$) and two ``alternative models" (model $M_2$ and $M_3$).
-The null model only includes the default parameters that are included in the RSiena estimation: the reciprocity and outdegree parameters.
-The alternative model $M_2$ includes one additional covariate parameter that was determined to be significant by the Wald test in the RSienaTest package, while the alternative model $M_3$ includes one additional structural parameter whose significance was determined in the same way.
-Details on these effects are given in Table~\ref{tab:models}.
-
-<<readsimu, echo=FALSE>>=
- nulls <- read.csv("Data/distribution_null_model.csv")
+<<readsim, echo=FALSE, message=FALSE, warning=FALSE>>=
+setwd("~/Desktop/NetworksResearch/NetworksVizInference/Results/")
+nulls <- read.csv("../Data/distribution_null_model.csv")
 # names(nulls)[-1] <- c("alpha1", "alpha2", "beta1", "beta2")
@@
 # 
- alt <- read.csv("Data/distribution_jumpTT_model.csv")
+alt <- read.csv("../Data/distribution_jumpTT_model.csv")
 # names(alt)[-1] <- c("alpha1", "alpha2", "beta1", "beta2", "beta3")
@@
 # 
- alt2nd <- read.csv("Data/distribution_dblpairs_model.csv")
-# names(alt2nd)[-1] <- c("alpha1", "alpha2", "beta1", "beta2", "beta4")
-# alt2nd$Sim <- 1:nrow(alt2nd)
-# alt2nd$Model <- "M3"
-# 
-# alt2 <- gather(alt, parameter, estimate, 2:6)
-# null2 <- gather(nulls, parameter, estimate, 2:5)
-# alt2nd2 <- gather(alt2nd, parameter, estimate, 2:6)
-# 
-# simu <- rbind(alt2[,-1], null2[,-1])
-# write.csv(simu, "Data/simulation-1000-M1-M2.csv", row.names=FALSE)
-
-simu2 <- read.csv("Data/simulation-1000-M1-M2-M3.csv")
-means <- simu2 %>% group_by(Model, parameter) %>% summarize(
-  mean = mean(estimate)
-)
+alt2nd <- read.csv("../Data/distribution_dblpairs_model.csv")
 library(dplyr)
@@
 \begin{table}[h]
-\caption{\label{tab:models}Parameters and estimates of models $M_1$, $M_2$, and $M_3$.
-Estimates are the mean of 1000 iterations of the model estimates.
-The lineups that follow are simulated from models using these values.}
 \centering
@@
 \end{tabular}}
+\caption{\label{tab:models}Parameters and estimates of models $M_1$, $M_2$, and $M_3$.
+Estimates are the mean of 1000 iterations of the model estimates.
+The lineups that follow are simulated from models using these values.}
 \end{table}
 
-\begin{figure}
-\centering
-<<hist-estimates, dependson='read-simu', echo=FALSE, fig.width=8, fig.height=5, out.width='.8\\linewidth'>>=
-qplot(data=simu2, estimate, fill=Model, geom="density", alpha=I(0.65)) + facet_wrap(~parameter, scales="free") + theme_bw()
-@
-\caption{\label{fig:hist-simulations}Histogram of the distribution of model parameters based on 1,000 simulation runs.
-Model parameter $\beta_3$ for jumping transitive triplets in model $M_2$ is significantly different from zero, but its inclusion also leads to significant changes in the  other model parameters of model $M_1$.
-The parameter $\beta_4$ for doubly acheived distances is also significantly different from zero, but has larger variance.
-The inclusion of $\beta_4$ also changes the estimates of the other model parameters, but not as much as the inclusion of $\beta_3$.}
-\end{figure}
-
-After estimating the parameters in each of these models, we simulate from them to obtain realizations of each of the models. The objective function for each actor in each model is given below. 
-  \begin{align*}
-  f^{M_1}_{i}(x) & = \hat{\beta}_1^{M_1}s_{i1}(x) +  \hat{\beta}_2^{M_1}s_{i2}(x) \\
-  f^{M_2}_{i}(x) & = \hat{\beta}_1^{M_2}s_{i1}(x) +  \hat{\beta}_2^{M_2}s_{i2}(x) + \hat{\beta}_3^{M_2}s_{i3}(x) \\
-  f^{M_3}_{i}(x) & = \hat{\beta}_1^{M_3}s_{i1}(x) +  \hat{\beta}_2^{M_3}s_{i2}(x) + \hat{\beta}_4^{M_3}s_{i4}(x) 
-  \end{align*}
-The rate parameters, $\alpha_1$ and $\alpha_2$ represent how many opportunities for change each actor gets when moving from wave 1 to 2 and from wave 2 to 3, respectively.
-The outdegree parameter, $\beta_1$, represents how likely an actor is to change outgoing ties.
-If the estimate, $\hat{\beta}_1$, is positive, the actor is more likely to create outgoing ties, while a negative estimate leads the actor to deleting outgoing ties.
-This effect is highly correlated with the reciprocity parameter, $\beta_2$.
-A negative estimate of this parameter implies that the actor is discouraged from reciprocating its incoming ties, while a positive estimate implies that the actor is encouraged to reciprocate all ties.
-The additional parameter in $M_2$, $\beta_3$, is a covariate parameter.
-The covariate in this model is the drinking behaviour of the 16 students in our data.
-The possible values are 1, 2, 3, and 4, which means the student drinks never, once or twice a year, once a month, or once a week.
-This jumping transitive triplet effect impacts the transitive closure of actors from different groups.
-Thus, a positive estimate encourages transitive closure when one of three actors is in a different covariate group than the other two, while a negative estimate discourages this behavior.
-An example of this type of closure is given in Figure~\ref{fig:jtt}.
-With the directed edges we also distinguish between 'i likes j' and 'j likes i'.
-Finally, the doubly achieved distances effect is defined by the number of actors to whom actor $i$ is not directly tied, and tied through two paths via at least two intermediaries.
-This is a structural effect, like the density and reciprocity effects.
-A positive coefficient value encourages indirect ties, while a negative value discourages the formation of indirect ties.
-
-\begin{figure}
-\begin{subfigure}[t]{.45\textwidth}
-\caption{\label{fig:jtt}Realization of a jumping transitive triplet, where $i$ is the focal actor, $j$ is the target actor, and $h$ is the intermediary.
-The group of the actors is represented by the shape of the node.}
-<<jtt, echo=FALSE, fig.width=2, fig.height=2, message = FALSE, warning = FALSE, fig.align='center'>>=
-jTTe <- data.frame(from = c('i', 'i', 'h'), to = c('h', 'j', 'j'))
-jTTn <- data.frame(id = letters[8:10], group = c(1,1,2))
-
-jTT <- merge(jTTe, jTTn, by.x = 'from', by.y = "id", all = T)
-
-set.seed(12345) 
-ggplot(data = jTT, aes(from_id = from, to_id = to)) + 
-  geom_net(aes(shape = as.factor(group)), directed = T, label = T, 
-           labelcolour='grey80',vjust = 0.5, hjust =0.5, arrowgap = .15, 
-           colour = 'black', size=10, 
-           ecolour = c("red", "grey40", "grey40", "grey40")) + 
-  expand_limits(x=c(-0.1,1.1), y=c(-0.1,1.1)) +
-  theme_net() +
-  theme(legend.position = "none")
-@
-\end{subfigure}\hfill
-\begin{subfigure}[t]{.45\textwidth}
-\caption{Doubly achieved distance between actors $i$ and $k$.}
-<<dab, echo=FALSE, fig.width=2, fig.height=2, message = FALSE, warning = FALSE, fig.align='center'>>=
-dade <- data.frame(from = c('i', 'i', 'h', 'j'), to = c('h', 'j', 'k', 'k'))
-dadn <- data.frame(id = letters[8:11], group = c(1,1,1,1))
-
-dad <- merge(dade, dadn, by.x = 'from', by.y = "id", all = T)
-
-set.seed(12345) 
-ggplot(data = dad, aes(from_id = from, to_id = to)) + 
-  geom_net(aes(shape = as.factor(group)), directed = T, label = T, labelcolour='grey80',vjust = 0.5, hjust =0.5, arrowgap = .15, colour = 'black', size=10) + 
-  expand_limits(x=c(-0.1,1.1), y=c(-0.1,1.1)) +
-  theme_net() +
-  theme(legend.position = "none")
-@
-\end{subfigure}
-\caption{\label{fig:structures}Structural newtwork effects. On the left, a jumping transitive triplet (JTT). On the right, a doubly achieved distance between $i$ and $k$.}
-\end{figure}
-
-\section{Lineup Simulation}
-
-\hh{Needs more specifics: how many lineups were created for each model?}
-
-\hh{At the moment we are just considering models M2 and M3 for an inclusion in the lineup study. We are trying to nail down what to include and what not to include in the experiment.  }
-
-To create the lineups that will be used in our experiment, we used the values given in Table~\ref{tab:models} as starting values in simulation.
-For each lineup simulated, the same default RSiena algorithm was used as was used to generate the fitted models with the exception that the algorithm only simulated from the given set of parameters.
-Each lineup and plot within lineup was generated idependent of all the others.
-
-\section{Parameter Estimation from Lineups}
-
-After the lineups were created, we re-fit each of our three models to each graph in each lineup.
-
-\hh{Questions that should be answered in this section:
-
-1. Why do we want to get the parameters for each model? 
-\st{We fit all three models to every plot in the lineups.
-Fitting all models to all lineup plots allows us to gauge tha ability of the parameter estimates to provide a measure of lineup identification difficulty.
-XXX ?
-not quite sure ?
-XXX For instance, when comparing models 1 and 2 in a lineup, the estimates from fitting model 2 to plots which were simulated from model 2 should be significantly different from the model 2 estimates from plots simulated from model
-0. Convergence issues
-
-\begin{figure}
-<<results0, fig.width=8, fig.height=3, out.width='\\linewidth', echo = FALSE, warning = FALSE>>=
-load("Data/lus_ests_truth.rda")
-lus_ests_truth$param_name <- 
-  factor(lus_ests_truth$param_name,
-    levels = c("rate", "outdegree (density)", "reciprocity",
-               "transitive triplets jumping alcohol2",
-               "number pairs at doubly achieved distance 2"))
-
-ggplot(data=lus_ests_truth) +
-  geom_bar(aes(x=model, fill=convergence), position="fill", alpha = 0.6) + 
-  facet_grid(.~true_model, labeller="label_both") +
-  theme_bw() +
-  theme(legend.position="bottom") + ylab("Ratio") +
-  scale_fill_brewer("Convergence", palette="Set1") +
-  xlab("Estimated Model")
-@
-\caption{\label{fig:convergence}Pattern of convergence.
-A total of 67.7\% of all lineup data converged in 5,000 iterations.
-The simplest model, $M_1$ has the highest rate of convergence.
-For models $M_2$ and $M_3$ data generated from the model converged at a higher rate than data generated from the other model.}
-\end{figure}
-
-\begin{figure}
-<<results0b, fig.width=8, fig.height=3, out.width='\\linewidth', echo = FALSE, warning = FALSE>>=
-load("Data/lus_ests_truth.rda")
-lus_ests_truth$param_name <- 
-  factor(lus_ests_truth$param_name,
-    levels = c("rate", "outdegree (density)", "reciprocity",
-               "transitive triplets jumping alcohol2",
-               "number pairs at doubly achieved distance 2"))        
-ggplot(data=lus_ests_truth) +
-  geom_density(aes(x=param_est-true_value, fill=convergence), alpha= 0.6) +
-  facet_grid(model+param_name~true_model, labeller="label_both") +
-  theme_bw() +
-  theme(legend.position="bottom") + 
-  scale_colour_brewer("Convergence", palette="Set1") +
-  xlab("Parameter Estimate") + xlim(c(-25,25))
-@
-\caption{\label{fig:convergence-ests}}
-\end{figure}
-
-1.
-The smaller the difference between these estimates, the harder it should be to identify the different model in the lineup.
-XXX IDEA: should we consider some sort of distance metric comparing all estimates from the models at once in addition to the differences in the $\beta_3$/$\beta_4$ values?
-XXX}
-
-2. How is the fitting done, exactly?
-\st{XXX get into nitty gritty from RSiena model XXX}
-
-\st{Because the likelihood function for these complicated models is intractable, RSiena implements a Monte Carlo simulation to obtain method of moments estimates of the parameter values.
-This fitting procedure was first introduced in ~\cite{saompaper}.
-}
-
-\st{The model fitting in RSiena is done in three phases.
-Briefly, in the initial phase, sensitivity of the statistics to the parameter values is determined, then in the second phase, parameter values are fit iteratively.
-Finally, in the third phase, networks are simulated from the fitted models and the model is checked for convergence.
-}
-
-4. What parameters are in the output?
-\st{In RSiena, the convergence of a non-rate parameter is determined through the simulated values from Phase 3 of the SienaFit algorithm.
-XXX more detail will be above later XXX  The simulations are compared to the observed values of the statistics observed in the data.
-The values from the simulations should be fairly close to the observed values, but because the fitting is done stochastically, deviations from statistics will not be exactly zero.
-Checking for convergence is based on a t-ratio of the average of these deviations to the standard deviation of these deviations.
-RSiena also performs an overall maximum convergence check by finding the maximum t-ratio value for any linear combination of the observed statistics.
-According to the RSiena Manual, ``convergence is excellent when the overall maximum convergence ratio is less than 0.2", and for the non-rate parameters, the threshold for ``reasonable" convergence is set at 0.3 with excellent convergence when the t-ratio is less than or equal to 0.1 in absolute value.
-(\cite{RSiena}).
-}
-5. What are the results?
-}
-\begin{figure}
-<<results1, fig.width=8, fig.height=6, out.width='\\linewidth', echo = FALSE, warning = FALSE>>=
-l1 <- levels(lus_ests_truth$param_name)
-l1 <- stringr::str_trim(gsub("2","", l1))
-l1[5] <- "#pairs at doubly achieved distance"
-
-# subset on ones that are actually converged. XXX No - that kills the zeroes in estimates that should be zero
-ggplot(data = subset(lus_ests_truth)) + #, convergence=="Converged")) + 
-  geom_vline(aes(xintercept = 0), colour="grey50") +
-  geom_density(alpha = .5, aes(x = param_est, fill = param_name)) + 
-  facet_grid(model~true_model, scales = 'free', 
-             labeller = "label_both") + xlim(c(-20,20)) +
-  theme_bw() + 
-  scale_fill_brewer("Parameter", palette="Dark2",
-     labels=c(bquote(paste(alpha,": ", .(l1[1]), sep="")), 
-              bquote(paste(beta[1],": ", .(l1[2]), sep="")),
-              bquote(paste(beta[2],": ", .(l1[3]), sep="")),
-              bquote(paste(beta[3],": ", .(l1[4]), sep="")),
-              bquote(paste(beta[4],": ", .(l1[5]), sep="")))) +
-  theme(legend.position = "bottom") +
-  xlab("Parameter Estimate") + 
-  guides(fill = guide_legend(nrow = 3, byrow = TRUE))
-@
-\caption{\label{fig:comparison} Comparison of model estimates under all three models under investigation. }
-\end{figure}
-
-<<est, echo = FALSE>>=
-#For model $M_2$, parameter $\beta_3$ is estimated to be significantly different from zero in almost all cases \hh{(XXX how many exactly? XXX)}.
-
-# get beta3s where true model and fit model are M2
-beta3sig <- lus_ests_truth %>% 
-  filter(true_model == "M2", model == "M2", 
-         param_name == "transitive triplets jumping alcohol2")
-convergebeta3 <- length(which(beta3sig$convergence == "Converged")) / nrow(beta3sig)
-
-# get ones that converged.
-beta3sig2 <- beta3sig %>% 
-  filter(convergence == "Converged") %>%
-  mutate(tstat = param_est/param_est_se,
-         ttestpval = 2*dt(tstat, df = 1))
-beta3sig05 <- sum(beta3sig2$ttestpval <= 0.05) / nrow(beta3sig2)
-beta3sig10 <- sum(beta3sig2$ttestpval <= 0.10) / nrow(beta3sig2)
-@
-Figure~\ref{fig:comparison} shows an overview of model estimates for each simulated data set.
-What we expect to see is that estimates do not change values (much) if they are estimated under a model different from the one they are generated from.
-This is true for all data sets estimated under model $M_1$ independently of which model they were generated from (top row of Figure~\ref{fig:comparison}).
-For data fitted under model $M_2$ we see that parameter $\beta_3$ (jumping transitive triplets) are estimated to be about zero, if the data is generated from models $M_1$ and $M_3$.
-For model $M_2$, parameter $\beta_3$ is estimated to be significantly different from zero in \st{\Sexpr{round(100 * beta3sig10, 1)}\% of cases}. This coincides with our expectation.
-
-However, the bottom row of Figure~\ref{fig:comparison} shows that independently of which model data is generated from, $\beta_4$, the number of pairs at doubly achieved distance, is estimated to be significantly different from zero in a large number of cases (about half of the data generated from model $M_2$ and more than that in model $M_1$).
-While $\beta_4$ is a highly significant parameter in model $M_3$, this questions the way that parameters are fitted and tells us that we are not likely to be able to visually distinguish between data generated from model $M_3$ and data generated from models $M_1$ and $M_2$.
-We might, however, be able to distinguish between data sets for which $\beta_4$ is estimated to be significantly different from zero and non-significant ones.
-\hh{XXX can we see differences in the pilot study?}
-
-\begin{figure}
-\centering
-<<echo=FALSE, fig.width=8, fig.height = 8, out.width='.6\\linewidth', warning = FALSE>>=
-load("Data/lus_ests_truth.rda")
-
-lus_ests_truth$true_panel <- lus_ests_truth$true_model == lus_ests_truth$model
-lus_spread <- tidyr::spread(lus_ests_truth[,c("lineupid","model", "true_model", "panel_num", "param_name", "param_est")], model, param_est)
-
-# parameters should be most different between data plot and null plots
-
-# merge in true values:
-truth <- unique(subset(lus_ests_truth, true_panel==TRUE)[,c("lineupid", "true_model","param_name","true_value")])
-
-lus_spread <- merge(lus_spread, truth, by=c("lineupid","true_model", "param_name"), all.x=TRUE)
-lus_spread$true_value[is.na(lus_spread$true_value)] <- 0
-
-# compute a distance between the data panel and the closest null panel
-lus_spread$part <- gsub("(.*)-[0-9]*-[0-9]*", '\\1', lus_spread$lineupid)
-
-lus_summary_2 <- lus_spread %>% dplyr::group_by(part, lineupid, param_name) %>%
-  dplyr::summarize(
-    data_n = sum(true_model=="M2"),
-    data_est = mean(M2[true_model=="M2"]),
-    data_sd = sd(M2[true_model=="M2"], na.rm=TRUE),
-    data_dist = data_est - max(M2[true_model=="M1"]),
-    upper_qu = sum(data_est > M2[true_model=="M1"]) / (n()-1),
-    data_panel = mean(panel_num[true_model=="M2"])
-)
-
-p1 <- qplot(data_dist, upper_qu, data = subset(lus_summary_2, !is.na(upper_qu) & param_name=="transitive triplets jumping alcohol2")) +
-  facet_grid(.~part) + xlab("Difference in M2 estimate of data panel and maximum of the null panels (from M1)") +
-  ylab("Ratio of null panels with smaller M2 estimate") +
-  ggtitle("jumping transitive triplets")
-
-lus_summary_2b <- lus_spread %>% dplyr::group_by(part, lineupid, param_name) %>%
-  dplyr::summarize(
-    data_n = sum(true_model=="M1"),
-    data_est = mean(M2[true_model=="M1"]),
-    data_sd = sd(M2[true_model=="M1"], na.rm=TRUE),
-    data_dist = data_est - min(M2[true_model=="M2"]),
-    lower_qu = sum(data_est < M2[true_model=="M2"]) / (n()-1) 
-)
-
-p2 <- qplot(data_dist, lower_qu, data = subset(lus_summary_2b, part %in% c("smallfriends", "smallfriends-rev") & param_name=="transitive triplets jumping alcohol2")) +
-  facet_grid(.~part) + xlab("Difference in M2 estimate of data panel and minimum of the null panels (from M2)") +
-  ylab("Ratio of null panels with larger M2 estimate") +
-  ggtitle("jumping transitive triplets")
-
-gridExtra::grid.arrange(p1,p2)
-@
-\caption{\label{fig:lusparms} Scatterplots of smallfriends and smallfriends-rev: comparing the ratio of null plots with smaller estimates of $\beta_3$ in the nulls than in the data panel (top row) and larger estimates of $\beta_3$ in the nulls than in the data panel.
-In lineups with a ratio of 1 the data should be 'easy' to identify in both scenarios.}
-\end{figure}
-
-<<>>=
-jtts <- read.csv("Data/lineups-jtt.csv")
-jtts$lineupid <- with(jtts, paste(model, m, rep, sep="-"))
-jtts$data_panel <- jtts$plot_order
-lus_summary_2 <- merge(lus_summary_2, jtts[, c("lineupid", "jtt", "data_panel")], by=c("lineupid","data_panel"), all.x=TRUE)
-qplot(data=subset(lus_summary_2, param_name=="transitive triplets jumping alcohol2" & part=="smallfriends"), data_est, jtt, geom="jitter") + geom_label(data=subset(lus_summary_2, param_name=="transitive triplets jumping alcohol2" & jtt > 20 & part=="smallfriends"), aes(label=lineupid), alpha=.6)
-@
-
-\begin{figure}
-\centering
-<<echo=FALSE, fig.width=8, fig.height = 8, out.width='.6\\linewidth', warning=FALSE>>=
-lus_summary_3 <- lus_spread %>% group_by(part, lineupid, param_name) %>%
-  summarize(
-    data_n = sum(true_model=="M3"),
-    data_est = mean(M3[true_model=="M3"]),
-    data_sd = sd(M3[true_model=="M3"], na.rm=TRUE),
-    data_dist = min(M3[true_model=="M1"]) - data_est,
-    lower_qu = sum(data_est < M3[true_model=="M1"])/(n()-1)
-  )
-
-p1 <- qplot(data_dist, lower_qu,  data = subset(lus_summary_3, !is.na(lower_qu) & param_name=="number pairs at doubly achieved distance 2")) +
-  facet_grid(.~part) +
-  xlab("Difference in M3 estimate of data panel and minimum of the null panels (M1 nulls)") +
-  ylab("Ratio of null panels with larger M3 estimate") +
-  ggtitle("doubly achieved distance")
-
-lus_summary_3b <- lus_spread %>% group_by(part, lineupid, param_name) %>%
-  summarize(
-    data_n = sum(true_model=="M1"),
-    data_est = mean(M3[true_model=="M1"]),
-    data_sd = sd(M3[true_model=="M1"], na.rm=TRUE),
-    data_dist = max(M3[true_model=="M3"]) - data_est,
-    upper_qu = sum(data_est > M3[true_model=="M3"])/(n()-1)
-  )
-
-p2 <- qplot(data_dist, upper_qu,  data = subset(lus_summary_3b, !is.na(upper_qu) & param_name=="number pairs at doubly achieved distance 2" &
-part %in% c("smallfriends-eff2", "smallfriends-eff2-rev"))) +
-  facet_grid(.~part) +
-  xlab("Difference in M3 estimate of data panel and minimum of the null panels (M3 nulls)") +
-  ylab("Ratio of null panels with smaller M3 estimate") +
-  ggtitle("doubly achieved distance")
-
-gridExtra::grid.arrange(p1,p2)
-#p1
-@
-\caption{\label{fig:lusparms-eff2} }
-\end{figure}
-
-\hh{
-6. How do the results influence our decision for the lineups?
-}
-The strong influence of the inclusion of $\beta_4$ in model 3 has made any comparison between model 1 and model 3 impossible.
-The $\beta_4$ estimate is always significantly greater than zero in fitted models that converged.
-Thus, $\beta_4$ should have been included from the beginning.
-% at end of param est we know that model m3 is useless. m3 messes with structure of m1. should have been in cluded in all of the models. nice & important but doesn't help for lineup study
-
-%  results from the the pilot study.
-how do estimates combine with pilot study ?
-have 10 non-m3 ones that are good.
-also want to see that m1 v m3 is way worse than m1 v m2.
-now have a reason for m1 v m3 is so bad.
-m1 v m3 ids should be NO GOOD.
-
-\section{Results from the pilot study}
-
-\hh{Big goal: Can we derive measures from the model estimates or the visual representation (ie.
-the network structure) that help us determine which lineups are more difficult than others, i.e.
-based on these estimates, how relaibly can we predict which panel will be picked from a lineup?
-XXX For that, we could also calculate the number of JTTs in each network - use `jtt` for that.}
-
-\hh{Littler goal: evaluate pilot study and see whether the results line up with the conjectures made in the previous section.}
-
-\hh{The pilot study consisted of an evaluation of a set of 20 lineups by 11 volunteers.
-For each of the model situations ($M_1$ vs $M_2$, $M_2$ vs $M_1$, $M_1$ vs $M_3$ and $M_3$ vs $M_1$) one lineup of size $m = 3, 6, 9, 12$ was used.
-Overall, the number of data identifications in the lineups was very low (34 out of 220 evaluations).
-Participants identified the data plot in two to four of the lineups they evaluated.
-The mode was three data identifications out of twenty per participant.
-}
-\paragraph{Suitability of $M_3$ in lineups:}
-\begin{figure}
-\centering
-<<data-picks, echo=FALSE, fig.width=10, fig.height=5, out.width='0.85\\linewidth'>>=
-pdffiles <- dir("GGExperimentApr28/", pattern="pdf")
-pdffiles <- gsub(".pdf","",pdffiles)
-#pdffiles <- gsub("-m","",pdffiles)
-#pdffiles <- gsub("-rep","",pdffiles)
-
-dframe <- strsplit(pdffiles, split=" ") %>% plyr::ldply(function(x) x)
-names(dframe) <- c("stimulus", "lineupid")
-
-results <- read.csv("GGExperimentApr28/responses_GGExpApr28.csv")
-lus <- results %>% group_by(Lineup, ChosenLU) %>% summarize(
-  tally = n(),
-  data = Answer[1],
-  reason = paste(Reasoning, collapse="|")
-)
-names(lus)[2] <- "panel"
-
-dframe_res <- merge(dframe, lus, all=TRUE, by.x="stimulus", by.y="Lineup")
-dframe_res$data_pick <- with(dframe_res, data==panel)
-dframe_res$model <- factor(gsub("(.*)-m.*", "\\1", dframe_res$lineupid))
-levels(dframe_res$model) <- c("M1 vs M2",  "M1 vs M3", "M3 vs M1", "M2 vs M1")
-dframe_res$model <- factor(dframe_res$model, levels = c("M1 vs M2",  "M2 vs M1", "M1 vs M3", "M3 vs M1"))
-dframe_res$label <- gsub(".*-(m.*)", "\\1", dframe_res$lineupid)
-
-dframe_res$data_pick <- factor(dframe_res$data_pick, levels=c("TRUE", "FALSE"))
-dframe_res$res <- factor(dframe_res$data_pick, levels=c("TRUE", "FALSE"))
-qplot(label, weight=tally, fill=data_pick, data=dframe_res) +
-  facet_grid(.~model, scales="free", space="free") + theme_bw() +
-  theme(axis.text.x = element_text(angle=45, hjust=1)) + 
-  scale_fill_brewer("Data panel chosen", palette="Set1", na.value="grey80") + xlab("Lineup")
-@
-\caption{\label{fig:data-picks}Barchart summarizing the number of responses from the pilot study by model and lineup.
-Color shows the number of times the data panel was chosen from the lineup.
-Clearly, the first two sets of lineups ($M_1$ vs $M_2$ and $M_2$ vs $M_1$) have on average higher number of data identifications.
-}
-\end{figure}
-
-\hh{Figure~\ref{fig:data-picks} shows barcharts of responses from the pilot study detailing the number of data identifications in each lineup. }
-\hh{It is more difficult to identify the data plot in lineups of graphs based on data from models $M_1$ and $M_3$ than  in lineups of graphs based on data from models $M_1$ and $M_2$.
-}
-
-\bibliographystyle{apalike}
-\bibliography{lineuppaperbib}
-
-\section{Model $M_1$ versus $M_2$ lineup size 3}
-\newcounter{k}
-\forloop{k}{1}{\value{k} < 21}{smallfriends-m-3-rep-\arabic{k}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-m-3-rep-\arabic{k}} \\ }
-
-\section{Model $M_1$ versus $M_2$ lineup size 6}
-\newcounter{k1}
-\forloop{k1}{1}{\value{k1} < 21}{
-Lineup-Images/pdfs/smallfriends-m-6-rep-\arabic{k1}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-m-6-rep-\arabic{k1}} \\ }
-
-\section{Model $M_1$ versus $M_2$ lineup size 9}
-\newcounter{k2}
-\forloop{k2}{1}{\value{k2} < 21}{Lineup-Images/pdfs/smallfriends-m-9-rep-\arabic{k2}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-m-9-rep-\arabic{k2}} \\ }
-
-\section{Model $M_1$ versus $M_2$ lineup size 12}
-\newcounter{k3}
-\forloop{k3}{1}{\value{k3} < 21}{Lineup-Images/pdfs/smallfriends-m-12-rep-\arabic{k3}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-m-12-rep-\arabic{k3}} \\ }
-
-\section{Model $M_1$ versus $M_2$ lineup size 16}
-\newcounter{k4}
-\forloop{k4}{1}{\value{k4} < 21}{smallfriends-m-16-rep-\arabic{k4}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-m-16-rep-\arabic{k4}} \\ }
-
-\section{Model $M_2$ versus $M_1$ lineup size 3}
-\newcounter{k5}
-\forloop{k5}{1}{\value{k5} < 21}{smallfriends-rev-m-3-rep-\arabic{k5}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-rev-m-3-rep-\arabic{k5}} \\ }
-
-\section{Model $M_2$ versus $M_1$ lineup size 6}
-\newcounter{k6}
-\forloop{k6}{1}{\value{k6} < 21}{smallfriends-rev-m-6-rep-\arabic{k6}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-rev-m-6-rep-\arabic{k6}} \\ }
-
-\section{Model $M_2$ versus $M_1$ lineup size 9}
-\newcounter{k7}
-\forloop{k7}{1}{\value{k7} < 21}{smallfriends-rev-m-9-rep-\arabic{k7}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-rev-m-9-rep-\arabic{k7}} \\ }
-
-\section{Model $M_2$ versus $M_1$ lineup size 12}
-\newcounter{k8}
-\forloop{k8}{1}{\value{k8} < 21}{smallfriends-rev-m-12-rep-\arabic{k8}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-rev-m-12-rep-\arabic{k8}} \\ }
-
-\section{Model $M_2$ versus $M_1$ lineup size 16}
-\newcounter{k9}
-\forloop{k9}{1}{\value{k9} < 21}{smallfriends-rev-m-16-rep-\arabic{k9}\newline
-\includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-rev-m-16-rep-\arabic{k9}} \\ }
-
-% \section{Model $M_1$ versus $M_3$ lineup size 3}
-% \newcounter{k10}
-% \forloop{k10}{1}{\value{k10} < 21}{smallfriends-eff2-m-3-rep-\arabic{k10}\newline
-% \includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-3-rep-\arabic{k10}} \\ }
-% 
-% \section{Model $M_1$ versus $M_3$ lineup size 6}
-% \newcounter{k11}
-% \forloop{k11}{1}{\value{k11} < 21}{smallfriends-eff2-m-6-rep-\arabic{k11}\newline
-% \includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-6-rep-\arabic{k11}} \\ }
-% 
-% \section{Model $M_1$ versus $M_3$ lineup size 9}
-% \newcounter{k12}
-% \forloop{k12}{1}{\value{k12} < 21}{smallfriends-eff2-m-9-rep-\arabic{k12}\newline
-% \includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-9-rep-\arabic{k12}} \\ }
-% 
-% \section{Model $M_1$ versus $M_3$ lineup size 12}
-% \newcounter{k13}
-% \forloop{k13}{1}{\value{k13} < 21}{smallfriends-eff2-m-12-rep-\arabic{k13}\newline
-% \includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-12-rep-\arabic{k13}} \\ }
-% 
-% \section{Model $M_1$ versus $M_3$ lineup size 16}
-% \newcounter{k14}
-% \forloop{k14}{1}{\value{k14} < 21}{smallfriends-eff2-m-16-rep-\arabic{k14}\newline
-% \includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-m-16-rep-\arabic{k14}} \\ }
-% 
-% \section{Model $M_3$ versus $M_1$ lineup size 3}
-% \newcounter{k15}
-% \forloop{k15}{1}{\value{k15} < 21}{smallfriends-eff2-rev-m-3-rep-\arabic{k15}\newline
-% \includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-3-rep-\arabic{k15}} \\ }
-% 
-% \section{Model $M_3$ versus $M_1$ lineup size 6}
-% \newcounter{k16}
-% \forloop{k16}{1}{\value{k16} < 21}{smallfriends-eff2-rev-m-6-rep-\arabic{k16}\newline
-% \includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-6-rep-\arabic{k16}} \\ }
-% 
-% \section{Model $M_3$ versus $M_1$ lineup size 9}
-% \newcounter{k17}
-% \forloop{k17}{1}{\value{k17} < 21}{smallfriends-eff2-rev-m-9-rep-\arabic{k17}\newline
-% \includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-9-rep-\arabic{k17}} \\ }
-% 
-% \section{Model $M_3$ versus $M_1$ lineup size 12}
-% \newcounter{k18}
-% \forloop{k18}{1}{\value{k18} < 21}{smallfriends-eff2-rev-m-12-rep-\arabic{k18}\newline
-% \includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-12-rep-\arabic{k18}} \\ }
-
-% \section{Model $M_3$ versus $M_1$ lineup size 16}
-% \newcounter{k19}
-% \forloop{k19}{1}{\value{k19} < 21}{smallfriends-eff2-rev-m-16-rep-\arabic{k19}\newline
-% \includegraphics[width = \textwidth]{Lineup-Images/pdfs/smallfriends-eff2-rev-m-16-rep-\arabic{k19}} \\ }
-
 \end{document}

### 2016-08-23 Results/TurkExp1WriteUp.Rnw
add more description of models

@@
 
-% BACKGROUND & PURPOSE
-The parameters in the objective function are tested for significance using $t$-tests.
-The test statistic is the ratio of the parameter estimate to its standard error.
-If this value is larger than 2 in absolute value, then the parameter is significant at the $\alpha = 0.05$ level and should be included in the model.
-This is a fairly simple statistical test, so we wanted to test whether this significance can be detected visually just as simply as the statistical test detects it.
-If visualizations of simulated networks from two nested models have a much different appearance when placed side-by-side, then the difference in appearance can be attributed to the additional parameter in one.
-If, however, there is no visually detectable difference, then the additional parameter does not appear to have changed the network structure all that much.
-Because model selection and diagnostic in network analysis are less developed areas of the theory, testing network parameters in this visual way could lead to additional methods of model selection for networks.
-
-\section{Methods}
-In order to test our hypothesis, we set up an amazon turk experiment using the lineup protocal of {\color{green} Buja et all (2009). FIX CITE}.
-
 <<readsim, echo=FALSE, message=FALSE, warning=FALSE>>=
@@
 \Sexpr{round(alt_mod_sv[5], 2)} & -- \\
-\# doubly achieved distances & $\beta_4$ & $s_{i4}(x) = |\{j : x_{ij} = 0, \sum\limits_h x_{ih}x_{hj} \geq 2\}|$ & -- & -- &
+\# doubly achieved distances 2 effect & $\beta_4$ & $s_{i4}(x) = |\{j : x_{ij} = 0, \sum\limits_h x_{ih}x_{hj} \geq 2\}|$ & -- & -- &
 \Sexpr{round(alt_mod2_sv[5], 2)}
@@
 
+The outdegree and reciprocity effects are the most important effects to any SAOMs.
+The outdegree effect ``always must be included," and the reciprocity effect ``practically always must be included" {\color{green} (CITE Rsiena Manual, p 41)}.
+The outdegree of a node in a network is a measure of how active that node is within the network, while the reciprocity parameter measures of how often a node's outgoing ties are reciprocated.
+The additional effects, jumping transitive triplets and doubly achieved distance, were deemed very significant by a $t$-test.
+
+\begin{figure}
+\begin{subfigure}[t]{.45\textwidth}
+\caption{\label{fig:jtt}Realization of a jumping transitive triplet, where $i$ is the focal actor, $j$ is the target actor, and $h$ is the intermediary.
+The group of the actors is represented by the shape of the node.}
+<<jtt, echo=FALSE, fig.width=2, fig.height=2, message = FALSE, warning = FALSE, fig.align='center'>>=
+jTTe <- data.frame(from = c('i', 'i', 'h'), to = c('h', 'j', 'j'))
+jTTn <- data.frame(id = letters[8:10], group = c(1,1,2))
+
+jTT <- merge(jTTe, jTTn, by.x = 'from', by.y = "id", all = T)
+
+set.seed(12345) 
+ggplot(data = jTT, aes(from_id = from, to_id = to)) + 
+  geom_net(aes(shape = as.factor(group)), directed = T, label = T, 
+           labelcolour='grey80',vjust = 0.5, hjust =0.5, arrowgap = .15, 
+           colour = 'black', size=10, 
+           ecolour = c("red", "grey40", "grey40", "grey40")) + 
+  expand_limits(x=c(-0.1,1.1), y=c(-0.1,1.1)) +
+  theme_net() +
+  theme(legend.position = "none")
+@
+\end{subfigure}\hfill
+\begin{subfigure}[t]{.45\textwidth}
+\caption{Doubly achieved distance between actors $i$ and $k$.}
+<<dab, echo=FALSE, fig.width=2, fig.height=2, message = FALSE, warning = FALSE, fig.align='center'>>=
+dade <- data.frame(from = c('i', 'i', 'h', 'j'), to = c('h', 'j', 'k', 'k'))
+dadn <- data.frame(id = letters[8:11], group = c(1,1,1,1))
+
+dad <- merge(dade, dadn, by.x = 'from', by.y = "id", all = T)
+
+set.seed(12345) 
+ggplot(data = dad, aes(from_id = from, to_id = to)) + 
+  geom_net(aes(shape = as.factor(group)), directed = T, label = T, labelcolour='grey80',vjust = 0.5, hjust =0.5, arrowgap = .15, colour = 'black', size=10) + 
+  expand_limits(x=c(-0.1,1.1), y=c(-0.1,1.1)) +
+  theme_net() +
+  theme(legend.position = "none")
+@
+\end{subfigure}
+\caption{\label{fig:structures}Structural newtwork effects. On the left, a jumping transitive triplet (JTT). On the right, a doubly achieved distance between $i$ and $k$.}
+\end{figure}
+
+% BACKGROUND AND PURPOSE
+The parameters in the objective function are tested for significance using $t$-tests.
+The test statistic is the ratio of the parameter estimate to its standard error.
+If this value is larger than 2 in absolute value, then the parameter is said to be significant at the $\alpha = 0.05$ level and should be included in the model.
+This is a fairly simple statistical test, so we wanted to test whether this significance can be detected visually just as simply as the statistical test detects it.
+If visualizations of simulated networks from two nested models have a much different appearance when placed side-by-side, then the difference in appearance can be attributed to the additional parameter in one.
+If, however, there is no visually detectable difference, then the additional parameter does not appear to have changed the network structure all that much.
+Because model selection and diagnostics for network models are less developed areas of the theory, testing network parameters in this visual way could lead to additional methods of model selection for networks.
+
+\section{Methods}
+In order to test our hypothesis, we set up an amazon turk experiment using the lineup protocal of {\color{green} Buja et all (2009).
+FIX CITE}.
+We presented four types of lineups: M1 v.
+M2, M2 v.
+M1, M1 v.
+M3, and M3 v.
+M1, where M1, M2, and M3 have the objective functions givein in ~\ref{tab:models}.
+For each lineup, there were 12 plots shown: 11 of the plots were simulated from the first model (the ``null" model) and 1 was simulated from the second model (the ``alternative" model).
+The order of the plots in the lineups was randomly assigned.
+
+In order to become a subject in our experiment, the Amazon turk user had to first read through some introductory material and prove they could identify the correct plot in two test lineups, one for M1 v.
+M2 and one for M2 v.
+M1.
+These two lineups were constructed to be very simple in order to train the turkers.
+Once the turker made it into the experiment, they looked at 10 lineups selected at random from a pool of 25.
+There were five lineups for each of the four types with an additional five lineups of the type M1 v.
+M2 which were chosen for their high counts of jumping transitive triplets in the alternative model network.
+
+\section{Appendix}
+
 \end{document}

### 2016-08-23 JSM/VizInfSlides.Rmd
add screen grabs from experiment

@@
 
-## Results 1
+## Results
 
-``` {r, echo=FALSE}
+``` {r, echo=FALSE, fig.align = 'center'}
+ids <- readr::read_csv("https://raw.githubusercontent.com/erichare/lineups/f19635a00507210013ba9e631761ad0c89a6564d/experiments/turk21/details/picture-details.csv")
 tab <- readr::read_csv("../Results/turk21_users.csv")
 res <- readr::read_csv("../Results/turk21_feedback.csv")
+res2 <- left_join(res, ids, by = 'pic_id')
+res3 <- res2 %>% filter(test_param %in% c("smallfriends", "smallfriends-rev")) 
+res3$correct <- res3$response_no == res3$obs_plot_location
+ggplot(data = res3) + 
+  geom_bar(aes(x = as.factor(response_no), fill = correct)) + 
+  facet_wrap(~data_name) + coord_flip() + 
+  labs(x = "response", y = "count") + 
+  theme(axis.text.y = element_blank())
 ```
-
-- barchart with filled correct
 
@@
 
-- scatterplot of %correct by beta 3
-- scatterplot of %correct by #jtt
+```{r, echo=FALSE}
+lus_jtts <- readr::read_csv("../Data/lineups-jtt.csv")
+load("../Data/lus_ests_truth.rda")
+
+res4 <-left_join(res3, lus_jtts, by = c("data_name" = "filename", "obs_plot_location" = "plot_order"))
+res4.2 <- res4 %>% group_by(data_name, jtt) %>% summarise(perc_corr = sum(correct)/n())
+p1 <- ggplot(data = res4.2) +geom_point(aes(x = jtt, y = perc_corr)) + labs(x = "Number of jumping transitive triplets", y = "Percent of Participants Choosing Correctly")
+res3a <- res3
+res3a$data_name <- sub(".csv", "", res3a$data_name)
+res3a$data_name <- sub("-m-", "-", res3a$data_name)
+res3a$data_name <- sub("-rep-", "-", res3a$data_name)
+exp_names <- unique(res3a$data_name)
+lus_truth2 <- lus_ests_truth %>% 
+  filter(param_name == "transitive triplets jumping alcohol2" &
+           lineupid %in% exp_names & model == "M2")
+res5 <- left_join(res3a, lus_truth2, by = c("data_name" = "lineupid",
+                                            "obs_plot_location" = "panel_num"))
+res5.2 <- res5 %>% group_by(data_name, param_est) %>% 
+  summarise(perc_corr = sum(correct)/n())
+p2 <- ggplot(data = res5.2) + geom_point(aes(x = param_est, y = perc_corr)) + labs(x = "Parameter estimate", y = "Percent of Participants Choosing Correctly")
+gridExtra::grid.arrange(p1, p2, nrow = 1)
+```
 
@@
 - Doesn't work for networks
-- Need another way! (layout?)
+- Need another way (layout?)

### 2016-08-23 Results/TurkExp1WriteUp.Rnw
add screen grabs from experiment

@@
 The outdegree of a node in a network is a measure of how active that node is within the network, while the reciprocity parameter measures of how often a node's outgoing ties are reciprocated.
-The additional effects, jumping transitive triplets and doubly achieved distance, were deemed very significant by a $t$-test.
+The additional effects, jumping transitive triplets and doubly achieved distance, were deemed very significant by a $t$-test, which was why we chose them for testing.
 
@@
 There were five lineups for each of the four types with an additional five lineups of the type M1 v.
-M2 which were chosen for their high counts of jumping transitive triplets in the alternative model network.
+M2 which were chosen for their high counts of jumping transitive triplets in the alternative model network in an attempt to gauge how important this statistic is to the visualization of the network.
+If there are many jumping transitive triplets in the alternative model plot and more turkers correctly choose that plot, that would be evidence that the jumping transitive triplets are very noticeable to the user.
+\st{????}
 
-\section{Appendix}
+\section{Procedure}
+
+Each Amazon Turk user was greeted with the following message: ``In this survey a series of similar looking charts will be presented. We would like you to respond to the following questions.
+\begin{enumerate}
+\item Pick the plot based on the survey question
+\item Provide reasons for choice
+\item How certain are you?
+\end{enumerate}
+Finally we would like to collect some information about you.
+(age category, education and gender)."  Then, they were led through a training page about how to identify the correct plot in a lineup, seen in Figure~\ref{fig:lineupex}.
+Then, they saw two trial plots that they had to get correct in order to proceed to the rest of the experiment.
+They chose the plot that looked the most different from the others, why they chose is (most complex, least complex, or other), and how certain they were that they had chosen correctly (Very Uncertain, Uncertain, Neutral, Certain, Very Certain).
+These trial plots are shown in Figure~\ref{fig:lineuptrial}
+
+\begin{figure}
+\begin{subfigure}[t]{.45\textwidth}
+\caption{Example 1}
+\includegraphics[width=\textwidth]{Ex1}
+\end{subfigure}\hfill
+\begin{subfigure}[t]{.45\textwidth}
+\caption{Example 2}
+\includegraphics[width=\textwidth]{Ex2}
+\end{subfigure}
+\caption{\label{fig:lineupex} The first training page in the Amazon Turk experiment.}
+\end{figure}
+
+\begin{figure}
+\begin{subfigure}[t]{.45\textwidth}
+\caption{Trial 1}
+\includegraphics[width=\textwidth]{Trial1}
+\end{subfigure}\hfill
+\begin{subfigure}[t]{.45\textwidth}
+\caption{Trial 2}
+\includegraphics[width=\textwidth]{Trial2}
+\end{subfigure}
+\caption{\label{fig:lineupex} The trial plots that users had to get correct in order to participate in the Amazon Turk experiment.}
+\end{figure}
 

### 2016-08-23 Results/TurkExp1WriteUp.Rnw
almost done with write up. have a few more ideas of what to add

@@
 \documentclass[12pt]{article}
-\usepackage{amsmath,amssymb,amsfonts, mathrsfs,fancyhdr,syntonly,lastpage,hyperref,enumitem,graphicx, lscape, forloop, url, natbib}
+\usepackage{amsmath,amssymb,amsfonts, mathrsfs,fancyhdr,syntonly,lastpage,hyperref,enumitem,graphicx, lscape, forloop, url, natbib, blindtext}
 \usepackage[font=small,skip=5pt]{caption}
@@
 
+\st{XXX Most of this I copied over from the other paper so it might not be super coherent right now. XXX}
 % ALTERNATE OPENER: Stochastic actor-oriented models (SAOMs) are used to model dynamic social networks.
@@
 jTT <- merge(jTTe, jTTn, by.x = 'from', by.y = "id", all = T)
-
+library(geomnet)
 set.seed(12345) 
@@
 \end{subfigure}
-\caption{\label{fig:lineupex} The trial plots that users had to get correct in order to participate in the Amazon Turk experiment.}
+\caption{\label{fig:lineuptrial} The trial plots that users had to get correct in order to participate in the Amazon Turk experiment.}
 \end{figure}
 
+Once the turk user chose the correct plot in the two trial plots, the experiment proceeded with the same interface and users selecting which plot they thought was most different, why they thought that, and how certain they were for 10 plots chosen at random.
+
+\section{Results}
+
+<<get_res, echo=FALSE, warning=FALSE, message=FALSE, fig.align="center", fig.cap="Boxplot showing the percent of users correctly choosing the alternative model in all 25 lineups.
+In 15 out of 25 lineups, no user picked the correct plot.
+The maximum was 64.00\\%, and the second highest was 35.29\\% correct.", fig.height=3>>=
+ids <- readr::read_csv("https://raw.githubusercontent.com/erichare/lineups/f19635a00507210013ba9e631761ad0c89a6564d/experiments/turk21/details/picture-details.csv")
+tab <- readr::read_csv("../Results/turk21_users.csv")
+tab2 <- unique(tab)
+#length(unique(tab2$nick_name))
+#head(sort(table(tab2$nick_name),decreasing = T))
+#tab2 %>% filter(nick_name == "A3U3L1XQVR362A")
+#tab2 %>% filter(nick_name == "ADTIO3A6TM9CG")
+#tab2 %>% filter(nick_name %in% names(table(res$nick_name))[table(res$nick_name) != 10])
+res <- readr::read_csv("../Results/turk21_feedback.csv")
+# remove people who only completed 1 plot
+res <- res %>% filter(!(nick_name %in% names(table(res$nick_name))[table(res$nick_name) != 10]))
+#intersect(names(ids), names(res))
+res2 <- left_join(res, ids, by = 'pic_id')
+res2$correct <- res2$response_no == res2$obs_plot_location
+res2$time <- res2$end_time - res2$start_time
+
+res2$lineup_name <- as.factor(paste(res2$test_param, res2$param_value))
+levels(res2$lineup_name) <- c(paste("M1 v M2 rep", 1:10), paste("M1 v M3 rep", 1:5),
+                              paste("M3 v M1 rep", 1:5), paste("M2 v M1 rep", 1:5))
+
+res2 %>% group_by(lineup_name) %>% 
+  summarize(total = n(), tot_correct = sum(correct), perc_correct = tot_correct/total) %>% 
+  arrange(desc(perc_correct)) -> lineup_by_perc_correct
+#lineup_by_perc_correct[which.max(lineup_by_perc_correct$total),]
+#lineup_by_perc_correct[which.min(lineup_by_perc_correct$total),]
+# box plot
+ggplot(data = lineup_by_perc_correct, aes(x = 1, y = perc_correct)) + geom_boxplot() + 
+  scale_y_continuous(breaks = seq(0, 1, .10), labels = paste0(seq(0, 1, .10)*100, "%")) + 
+  coord_flip() + theme_classic() + 
+  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) + 
+  labs(x = "", title = "Percent of Users Correctly Identifying Alternate Model", y = "")
+@
+
+We collected 10 lineups each from 77 Amazon Turk users on July 28, 2016 from 19:11:03 GMT to 22:17:47 GMT. A table of demographic information is in
+
+\begin{figure}
+\begin{subfigure}[c]{.45\textwidth}
+\textbf{Gender:} \\
+<<tab_demo, echo=FALSE, results='asis'>>=
+t1 <- t(data.frame(table(tab2$gender)))
+rownames(t1) <- c("Gender:", "Count:")
+colnames(t1) <- t1[1,]
+t1 <- t1[2,]
+kable(t1, format = 'latex')
+@
+\vspace{.5cm}
+\textbf{Education Level:} \\
+<<tab_demo3, echo=FALSE, results='asis'>>=
+t3 <- t(data.frame(table(tab2$academic_study)))
+rownames(t3) <- c("Highest Education:", "Count:")
+colnames(t3) <- t3[1,]
+t3 <- t3[2,]
+kable(t3, format = 'latex')
+@
+\end{subfigure}\hfill
+\begin{subfigure}[c]{.45\textwidth}
+\centering
+\textbf{Age Group:} \\
+<<tab_demo2, echo=FALSE, results='asis'>>=
+t2 <- t(data.frame(table(tab2$age)))
+rownames(t2) <- c("Age:", "Count:")
+colnames(t2) <- t2[1,]
+t2 <- t2[2,]
+kable(t2, format = 'latex')
+@
+\end{subfigure}
+\caption{\label{fig:demos} Demographic information on the Amazon Turk users participating in the experiment}
+\end{figure}
+
+Because of the random assignment of the 25 plots to users, almost every lineup was seen by a different number of people.
+Repetition 3 of M1 v.
+M3 was seen by 47 different users while repetition 2 of the same type was seen by only 6 different users.
+The mean number of times seens was 30.8 and the median was 31.
+In 15 of the 25 lineups, no user chose the correct plot, and in the remaining 10 lineups, the users correctly identified the alternative plot a minimum of 4.35\% of the time and a maximum of 64\% of the time.
+
+<<user_demos, echo=FALSE, warning=FALSE, eval=FALSE>>=
+makeplot_mosaic <- function(data, x, y, ...){
+  xvar <- deparse(substitute(x))
+  yvar <- deparse(substitute(y))
+  mydata <- data[c(xvar, yvar)];
+  mytable <- table(mydata);
+  widths <- c(0, cumsum(apply(mytable, 1, sum)));
+  heights <- apply(mytable, 1, function(x){c(0, cumsum(x/sum(x)))});
+
+  alldata <- data.frame();
+  allnames <- data.frame();
+  for(i in 1:nrow(mytable)){
+    for(j in 1:ncol(mytable)){
+      alldata <- rbind(alldata, c(widths[i], widths[i+1], heights[j, i], heights[j+1, i]));
+    }
+  }
+  colnames(alldata) <- c("xmin", "xmax", "ymin", "ymax")
+
+  alldata[[xvar]] <- rep(dimnames(mytable)[[1]],rep(ncol(mytable), nrow(mytable)));
+  alldata[[yvar]] <- rep(dimnames(mytable)[[2]],nrow(mytable));
+
+  ggplot(alldata, aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax)) + 
+    geom_rect(color="black", aes_string(fill=yvar)) +
+    xlab(paste(xvar, "(count)")) + ylab(paste(yvar, "(proportion)"));
+}
+
+makeplot_mosaic(tab2, x = age, y = academic_study)
+@
+
+<<facets_correct, echo=FALSE, fig.align="center", fig.cap="Plots chosen for each of the 25 lineups in the experiment.
+The x-axis ticks do not have a label because the label is less important than the number of different plots users chose and whether or not they chose the correct plot, shown in the coloring of the bars.", fig.height=9>>=
+ggplot(data = res2, aes(x = as.factor(response_no), fill = correct)) + 
+  geom_bar(color = 'black') + 
+  scale_fill_manual(values = c("white", "grey30"), name = "Correct?") + 
+  facet_wrap(~lineup_name, scales = "free", nrow = 5) +
+  theme_bw() + 
+  theme(axis.text.x = element_blank(), legend.position = 'bottom') + 
+  labs(x = "User Response" , y = "Number of Responses")
+@
+
 \end{document}

### 2016-08-24 Results/TurkExp1WriteUp.Rnw
describing facetted results plot

@@
 
-We collected 10 lineups each from 77 Amazon Turk users on July 28, 2016 from 19:11:03 GMT to 22:17:47 GMT. A table of demographic information is in
+We collected 10 lineups each from 77 Amazon Turk users on July 28, 2016 from 19:11:03 GMT to 22:17:47 GMT.
+A table of demographic information collected on the participants in the experiment is in Table~\ref{fig:demos}.
 
@@
 In 15 of the 25 lineups, no user chose the correct plot, and in the remaining 10 lineups, the users correctly identified the alternative plot a minimum of 4.35\% of the time and a maximum of 64\% of the time.
-
-<<user_demos, echo=FALSE, warning=FALSE, eval=FALSE>>=
-makeplot_mosaic <- function(data, x, y, ...){
-  xvar <- deparse(substitute(x))
-  yvar <- deparse(substitute(y))
-  mydata <- data[c(xvar, yvar)];
-  mytable <- table(mydata);
-  widths <- c(0, cumsum(apply(mytable, 1, sum)));
-  heights <- apply(mytable, 1, function(x){c(0, cumsum(x/sum(x)))});
-
-  alldata <- data.frame();
-  allnames <- data.frame();
-  for(i in 1:nrow(mytable)){
-    for(j in 1:ncol(mytable)){
-      alldata <- rbind(alldata, c(widths[i], widths[i+1], heights[j, i], heights[j+1, i]));
-    }
-  }
-  colnames(alldata) <- c("xmin", "xmax", "ymin", "ymax")
-
-  alldata[[xvar]] <- rep(dimnames(mytable)[[1]],rep(ncol(mytable), nrow(mytable)));
-  alldata[[yvar]] <- rep(dimnames(mytable)[[2]],nrow(mytable));
-
-  ggplot(alldata, aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax)) + 
-    geom_rect(color="black", aes_string(fill=yvar)) +
-    xlab(paste(xvar, "(count)")) + ylab(paste(yvar, "(proportion)"));
-}
-
-makeplot_mosaic(tab2, x = age, y = academic_study)
-@
+A plot showing the number of different plots chosen, how many times they were chosen, and whether or not it was the correct choice is given in Figure~\ref{fig:facets_correct}.
 
@@
 
+This plot shows us the abysmal ability of the Turkers to identify the althernatie plot correctly.
+There are also a few other interesting results from this plot.
+First, there are several lineups where a majority of participants selected the same wrong alternative plot.
+Additionally, we see a lot of variation in the number of different plots selected at the alternate plot by the Turkers.
+At most, there were 10 different plots selected, compared to 2 at the opposite end.
+
 \end{document}

### 2016-08-25 Results/TurkExp1WriteUp.Rnw
more results write-up

@@
 
-This plot shows us the abysmal ability of the Turkers to identify the althernatie plot correctly.
+This plot shows us the abysmal ability of the Turkers to identify the alternative plot correctly.
 There are also a few other interesting results from this plot.
@@
 At most, there were 10 different plots selected, compared to 2 at the opposite end.
+\st{XXX is it worth exploring these ``interesting things" more in-depth?
+XXX}
+
+Next, we investigate the confidence of the experiment's participants in selecting the most different plot from the others.
+Overall, in 40\% of the responses, the respondent was certain they were correct.
+User confidence in the remaining categories, neutral, uncertain, very certain, and very uncertain, was 26.88\%, 16.88\%, 9.48\%, and 6.76\%, respectively.
+These results are summarized by lineup in Figure~\ref{fig:cert_plot}.
+
+<<cert_plot, echo=FALSE, fig.align='center', fig.height=4, fig.cap="All responses for all lineups, separated by whether the plot selected was the true alternative plot (TRUE) or not (FALSE) and colored by the respondent's uncertainty levels.
+We see here that most respondents were certain in their answer whether or not they were correct.
+">>=
+res2$conf_level <- ordered(res2$conf_level, levels = c("Very Uncertain", "Uncertain", "Neutral", "Certain", "Very Certain"))
+
+ggplot(data = res2, aes(x = lineup_name)) + 
+  geom_bar(aes(fill = conf_level), color = 'grey60') + 
+  scale_fill_brewer(palette = "RdBu") + 
+  coord_flip() + 
+  facet_wrap(~correct) + 
+  theme_bw() + 
+  theme(legend.position = 'bottom')
+
+prop.table(table(res2$conf_level, res2$correct))
+
+@
+
+We next investigate the length of time that users' took to respond to each lineup.
+The minimum was 3.762 seconds and the maximum was 578.8 seconds (nearly 10 minutes).
+The median was 13.01 seconds and the mean was 20.32 seconds.
+
+<<time_plot, echo=FALSE>>=
+#qplot(x = res2$time, binwidth = 2)
+#ggplot(data = res2, aes(x = conf_level, y = time)) + geom_boxplot() + coord_flip()
+ggplot(data = res2, aes(x = lineup_name, y = time)) + 
+  geom_boxplot() + 
+#  facet_wrap(~conf_level) + 
+  coord_flip() + 
+  theme_bw() + 
+  labs(x = "Time (in seconds)", y = "Lineup Name")
+@
 

### 2016-08-25 Results/TurkExp1WriteUp.Rnw
proportion test - certainty and correctness

@@
 These results are summarized by lineup in Figure~\ref{fig:cert_plot}.
+We also performed a 5-sample test for equality of proportions in order to test the null hypothesis that the proportion of correct responses is the same in all 5 certainty categories.
+This test resulted in a $p$-value of 0.5881, so there is no evidence that the certainty of a user's response affects whether or not they choose the correct plot.
+This provides further evidence that detecting a network simulated from a different model is an extremely difficult problem.
 
@@
 
-prop.table(table(res2$conf_level, res2$correct))
-
+summary.conf <- data.frame(table(res2$conf_level, res2$correct))
+summary.conf %>% tidyr::spread(Var2, Freq) -> summary.conf
+rownames(summary.conf) <- as.character(summary.conf[,1])
+names(summary.conf) <- c("Var1", "Failures", "Successes")
+summary.conf <- summary.conf[,-1]
+summary.conf <- summary.conf[,c(2,1)]
+prop_test_for_conf_corr <- prop.test(as.matrix(summary.conf))
 @
@@
 #qplot(x = res2$time, binwidth = 2)
-#ggplot(data = res2, aes(x = conf_level, y = time)) + geom_boxplot() + coord_flip()
+ggplot(data = res2, aes(x = conf_level, y = time)) + geom_boxplot() + coord_flip()
 ggplot(data = res2, aes(x = lineup_name, y = time)) + 

### 2016-08-25 Results/TurkExp1WriteUp.Rnw
ks test for equality of distributions of time taken to respond

@@
 The median was 13.01 seconds and the mean was 20.32 seconds.
+First, a 2-sample Kolmogorov-Smirnov test for equality of distribution was done to see if the distribution of times for correct plots chosen is the same as the distribution of times for incorrect plots chosen.
+This two-sided test resulted in a $p$-value of 1, so there is no evidence that the distribution of times is different whether or not the response was correct.
 
@@
 ggplot(data = res2, aes(x = conf_level, y = time)) + geom_boxplot() + coord_flip()
+ggplot(data = res2, aes(x = correct, y = time)) + geom_boxplot() + coord_flip() + theme_bw()
+ggplot(data = res2, aes(x = time)) + geom_histogram(binwidth = 5, fill ='white', color = 'black') + facet_wrap(~correct, nrow = 2)
 ggplot(data = res2, aes(x = lineup_name, y = time)) + 
@@
   labs(x = "Time (in seconds)", y = "Lineup Name")
+# 2-sample kolmogorov-smirnov test for equality of distributions
+# ks.test(x = summary(res2$time[res2$correct]), y = summary(res2$time[!res2$correct]))
+# fail to reject null
 @

### 2016-08-25 Results/TurkExp1WriteUp.Rnw
results are obvious and yet still interesting!

@@
 
+\section{Discussion/Future Work}
+
+This small experiment suggests that detecting a network model difference from just one simulation from one model and 11 from the other model is just about impossible.
+This result is fairly unsurprising: drawing a single random point from, say, a $\Chi^2_1$ distribution and placing it in a lineup with 11 separate draws from a standard normal distribution would likely have similar results.
+Every once in a while, a chi-square value would appear that would be too large to belong with draws from a standard normal distribution, but values near 1, the mean, would probably not appear that different from some random standard normal draws.
+This means that we need to develop a different way to compare two network models.
+We need a way to visualize many samples from a network model in one panel.
+
 \end{document}

### 2016-08-29 Results/TurkExp1WriteUp.Rnw
fix global file path

@@
 <<readsim, echo=FALSE, message=FALSE, warning=FALSE>>=
-setwd("~/Desktop/NetworksResearch/NetworksVizInference/Results/")
+setwd("../Results/")
 nulls <- read.csv("../Data/distribution_null_model.csv")
@@
 
-<<cert_plot, echo=FALSE, fig.align='center', fig.height=4, fig.cap="All responses for all lineups, separated by whether the plot selected was the true alternative plot (TRUE) or not (FALSE) and colored by the respondent's uncertainty levels.
-We see here that most respondents were certain in their answer whether or not they were correct.
-">>=
+<<cert_plot, echo=FALSE, warning = FALSE, message = FALSE, fig.align='center', fig.height=4, fig.cap="All responses for all lineups, separated by whether the plot selected was the true alternative plot (TRUE) or not (FALSE) and colored by the respondent's uncertainty levels.
+We see here that most respondents were certain in their answer whether or not they were correct.">>=
 res2$conf_level <- ordered(res2$conf_level, levels = c("Very Uncertain", "Uncertain", "Neutral", "Certain", "Very Certain"))
@@
   theme_bw() + 
-  labs(x = "Time (in seconds)", y = "Lineup Name")
+  labs(y = "Time (in seconds)", x = "Lineup Name")
 # 2-sample kolmogorov-smirnov test for equality of distributions
@@
 This small experiment suggests that detecting a network model difference from just one simulation from one model and 11 from the other model is just about impossible.
-This result is fairly unsurprising: drawing a single random point from, say, a $\Chi^2_1$ distribution and placing it in a lineup with 11 separate draws from a standard normal distribution would likely have similar results.
+This result is fairly unsurprising: drawing a single random point from, say, a $\chi^2_1$ distribution and placing it in a lineup with 11 separate draws from a standard normal distribution would likely have similar results.
 Every once in a while, a chi-square value would appear that would be too large to belong with draws from a standard normal distribution, but values near 1, the mean, would probably not appear that different from some random standard normal draws.

### 2016-08-30 Results/TurkExp1WriteUp.Rnw
copy-paste the turk write-up into the lineups.rnw

@@
 ids <- readr::read_csv("https://raw.githubusercontent.com/erichare/lineups/f19635a00507210013ba9e631761ad0c89a6564d/experiments/turk21/details/picture-details.csv")
-tab <- readr::read_csv("../Results/turk21_users.csv")
+tab <- read.csv("../Results/turk21_users.csv")
 tab2 <- unique(tab)
@@
 #tab2 %>% filter(nick_name %in% names(table(res$nick_name))[table(res$nick_name) != 10])
-res <- readr::read_csv("../Results/turk21_feedback.csv")
+res <- read.csv("../Results/turk21_feedback.csv")
 # remove people who only completed 1 plot

### 2016-09-01 lineups.Rnw
plot of results ordered by grouping and JTTs :) :) :) :)

@@
 \end{figure}
+
+The parameters in the objective function are tested for significance using $t$-tests.
+The test statistic is the ratio of the parameter estimate to its standard error.
+If this value is larger than 2 in absolute value, then the parameter is said to be significant at the $\alpha = 0.05$ level and should be included in the model.
+This is a fairly simple statistical test, so we wanted to test whether this significance can be detected visually just as simply as the statistical test detects it.
+If visualizations of simulated networks from two nested models have a much different appearance when placed side-by-side, then the difference in appearance can be attributed to the additional parameter in one.
+If, however, there is no visually detectable difference, then the additional parameter does not appear to have changed the network structure all that much.
+Because model selection and diagnostics for network models are less developed areas of the theory, testing network parameters in this visual way could lead to additional methods of model selection for networks.
 
@@
 
+\section{Amazon Turk Experiment}
+
+\subsection{Methods}
+In order to test our hypothesis, we set up an amazon turk experiment using the lineup protocal of {\color{green} Buja et all (2009).
+FIX CITE}.
+We presented four types of lineups: M1 v.
+M2, M2 v.
+M1, M1 v.
+M3, and M3 v.
+M1, where M1, M2, and M3 have the objective functions givein in ~\ref{tab:models}.
+For each lineup, there were 12 plots shown: 11 of the plots were simulated from the first model (the ``null" model) and 1 was simulated from the second model (the ``alternative" model).
+The order of the plots in the lineups was randomly assigned.
+
+In order to become a subject in our experiment, the Amazon turk user had to first read through some introductory material and prove they could identify the correct plot in two test lineups, one for M1 v.
+M2 and one for M2 v.
+M1.
+These two lineups were constructed to be very simple in order to train the turkers.
+Once the turker made it into the experiment, they looked at 10 lineups selected at random from a pool of 25.
+There were five lineups for each of the four types with an additional five lineups of the type M1 v.
+M2 which were chosen for their high counts of jumping transitive triplets in the alternative model network in an attempt to gauge how important this statistic is to the visualization of the network.
+If there are many jumping transitive triplets in the alternative model plot and more turkers correctly choose that plot, that would be evidence that the jumping transitive triplets are very noticeable to the user.
+\st{????}
+
+\subsection{Procedure}
+
+Each Amazon Turk user was greeted with the following message: ``In this survey a series of similar looking charts will be presented. We would like you to respond to the following questions.
+\begin{enumerate}
+\item Pick the plot based on the survey question
+\item Provide reasons for choice
+\item How certain are you?
+\end{enumerate}
+Finally we would like to collect some information about you.
+(age category, education and gender)."  Then, they were led through a training page about how to identify the correct plot in a lineup, seen in Figure~\ref{fig:lineupex}.
+Then, they saw two trial plots that they had to get correct in order to proceed to the rest of the experiment.
+They chose the plot that looked the most different from the others, why they chose is (most complex, least complex, or other), and how certain they were that they had chosen correctly (Very Uncertain, Uncertain, Neutral, Certain, Very Certain).
+These trial plots are shown in Figure~\ref{fig:lineuptrial}
+
+\begin{figure}
+\begin{subfigure}[t]{.45\textwidth}
+\caption{Example 1}
+\includegraphics[width=\textwidth]{Ex1}
+\end{subfigure}\hfill
+\begin{subfigure}[t]{.45\textwidth}
+\caption{Example 2}
+\includegraphics[width=\textwidth]{Ex2}
+\end{subfigure}
+\caption{\label{fig:lineupex} The first training page in the Amazon Turk experiment.}
+\end{figure}
+
+\begin{figure}
+\begin{subfigure}[t]{.45\textwidth}
+\caption{Trial 1}
+\includegraphics[width=\textwidth]{Trial1}
+\end{subfigure}\hfill
+\begin{subfigure}[t]{.45\textwidth}
+\caption{Trial 2}
+\includegraphics[width=\textwidth]{Trial2}
+\end{subfigure}
+\caption{\label{fig:lineuptrial} The trial plots that users had to get correct in order to participate in the Amazon Turk experiment.}
+\end{figure}
+
+Once the turk user chose the correct plot in the two trial plots, the experiment proceeded with the same interface and users selecting which plot they thought was most different, why they thought that, and how certain they were for 10 plots chosen at random.
+
+\subsection{Results}
+
+<<get_res, echo=FALSE, warning=FALSE, message=FALSE, fig.align="center", fig.cap="Boxplot showing the percent of users correctly choosing the alternative model in all 25 lineups.
+In 15 out of 25 lineups, no user picked the correct plot.
+The maximum was 64.00\\%, and the second highest was 35.29\\% correct.", fig.height=3>>=
+ids <- readr::read_csv("https://raw.githubusercontent.com/erichare/lineups/f19635a00507210013ba9e631761ad0c89a6564d/experiments/turk21/details/picture-details.csv")
+tab <- read.csv("../Results/turk21_users.csv")
+tab2 <- unique(tab)
+#length(unique(tab2$nick_name))
+#head(sort(table(tab2$nick_name),decreasing = T))
+#tab2 %>% filter(nick_name == "A3U3L1XQVR362A")
+#tab2 %>% filter(nick_name == "ADTIO3A6TM9CG")
+#tab2 %>% filter(nick_name %in% names(table(res$nick_name))[table(res$nick_name) != 10])
+res <- read.csv("../Results/turk21_feedback.csv", stringsAsFactors = F)
+# remove people who only completed 1 plot
+res <- res %>% filter(!(nick_name %in% names(table(res$nick_name))[table(res$nick_name) != 10]))
+# need to repeat rows with weights for people who selected multiple plots.
+res_norpt <- res[-grep(",", res$response_no),]  
+res_rpt <- res[grep(",", res$response_no),]  
+res_rpt$response_no
+library(stringr)
+# count # of responses (ttr = times to repeat)
+ttr <- str_count(string = res_rpt$response_no, pattern = ",") + 1
+rpt_rows <- rep(1:nrow(res_rpt), ttr)
+rpt_rows2 <- res_rpt[rpt_rows,]
+rpt_rows2$count_response <- unlist(sapply(ttr, seq_len))
+rpt_rows2$weight <- 1/rep(ttr, ttr)
+split_rows <- strsplit(rpt_rows2$response_no, ",")
+rpt_rows2$response_no2 <- NA
+for(i in 1:nrow(rpt_rows2)){
+  rpt_rows2$response_no2[i] <- split_rows[[i]][rpt_rows2$count_response[i]]
+}
+
+res_norpt$count_response <- 1 
+res_norpt$response_no2 <- res_norpt$response_no
+res_norpt$tot_response <- 1 
+res_norpt$weight <- 1
+# res with weight column   
+new_res <- rbind(res_norpt, rpt_rows2)
+
+#intersect(names(ids), names(res))
+res2 <- left_join(new_res, ids, by = 'pic_id')
+res2$correct <- res2$response_no == res2$obs_plot_location
+res2$time <- res2$end_time - res2$start_time
+res2$lineup_name <- as.factor(paste(res2$test_param, res2$param_value))
+# with the correct labels (in order of data plot w/value most different from rest to least different)
+more_details <- read.csv("experimentdetails_Aug31.csv", stringsAsFactors = F)
+res3 <- left_join(res2, more_details, by = c("data_name" = "lineup_filename"))
+# higest jtts are: reps 18, 20, 6, 19, 3 with 
+              # jtts of 40, 24, 18,17, 9
+head(res3)
+res3$group_name <- paste0(res3$group, " #", res3$group.rep)
+
+# get the weights
+res3 %>% group_by(group_name, response_no2, correct) %>% summarise(tot_wt = sum(weight)) -> summ_res_wts
+
+# res3 %>% group_by(group_name) %>% 
+#   summarize(total = n(), tot_correct = sum(correct), perc_correct = tot_correct/total) %>% 
+#   arrange(desc(perc_correct)) -> lineup_by_perc_correct
+#lineup_by_perc_correct[which.max(lineup_by_perc_correct$total),]
+#lineup_by_perc_correct[which.min(lineup_by_perc_correct$total),]
+# box plot no - ew. 
+# ggplot(data = lineup_by_perc_correct, aes(x = 1, y = perc_correct)) + geom_boxplot() + 
+#   scale_y_continuous(breaks = seq(0, 1, .10), labels = paste0(seq(0, 1, .10)*100, "%")) + 
+#   coord_flip() + theme_classic() + 
+#   theme(axis.text.y = element_blank(), axis.ticks.y = element_blank()) + 
+#   labs(x = "", title = "Percent of Users Correctly Identifying Alternate Model", y = "")
+@
+
+We collected 10 lineups each from 77 Amazon Turk users on July 28, 2016 from 19:11:03 GMT to 22:17:47 GMT.
+A table of demographic information collected on the participants in the experiment is in Table~\ref{fig:demos}.
+
+\begin{figure}
+\begin{subfigure}[c]{.45\textwidth}
+\textbf{Gender:} \\
+<<tab_demo, echo=FALSE, results='asis'>>=
+t1 <- t(data.frame(table(tab2$gender)))
+rownames(t1) <- c("Gender:", "Count:")
+colnames(t1) <- t1[1,]
+t1 <- t1[2,]
+kable(t1, format = 'latex')
+@
+\vspace{.5cm}
+\textbf{Education Level:} \\
+<<tab_demo3, echo=FALSE, results='asis'>>=
+t3 <- t(data.frame(table(tab2$academic_study)))
+rownames(t3) <- c("Highest Education:", "Count:")
+colnames(t3) <- t3[1,]
+t3 <- t3[2,]
+kable(t3, format = 'latex')
+@
+\end{subfigure}\hfill
+\begin{subfigure}[c]{.45\textwidth}
+\centering
+\textbf{Age Group:} \\
+<<tab_demo2, echo=FALSE, results='asis'>>=
+t2 <- t(data.frame(table(tab2$age)))
+rownames(t2) <- c("Age:", "Count:")
+colnames(t2) <- t2[1,]
+t2 <- t2[2,]
+kable(t2, format = 'latex')
+@
+\end{subfigure}
+\caption{\label{fig:demos} Demographic information on the Amazon Turk users participating in the experiment}
+\end{figure}
+
+Because of the random assignment of the 25 plots to users, almost every lineup was seen by a different number of people.
+Repetition 3 of M1 v.
+M3 was seen by 47 different users while repetition 2 of the same type was seen by only 6 different users.
+The mean number of times seens was 30.8 and the median was 31.
+In 15 of the 25 lineups, no user chose the correct plot, and in the remaining 10 lineups, the users correctly identified the alternative plot a minimum of 4.35\% of the time and a maximum of 64\% of the time.
+A plot showing the number of different plots chosen, how many times they were chosen, and whether or not it was the correct choice is given in Figure~\ref{fig:facets_correct}.
+
+<<facets_correct, echo=FALSE, fig.align="center", fig.cap="Plots chosen for each of the 25 lineups in the experiment.
+The x-axis ticks do not have a label because the label is less important than the number of different plots users chose and whether or not they chose the correct plot, shown in the coloring of the bars.", fig.height=9>>=
+ggplot(data = summ_res_wts, aes(x = as.factor(response_no2), 
+                                y = tot_wt,
+                                fill = correct)) + 
+  geom_bar(color = 'black', stat = 'identity') + 
+  scale_fill_manual(values = c("white", "grey30"), 
+                    name = "Correct?") + 
+  facet_wrap(~group_name, scales = "free", nrow = 5) +
+  theme_bw() + 
+  theme(axis.text.x = element_blank(), 
+        legend.position = 'bottom') + 
+  labs(x = "User Response" , y = "Number of Responses")
+@
+
+This plot shows us the abysmal ability of the Turkers to identify the alternative plot correctly.
+There are also a few other interesting results from this plot.
+First, there are several lineups where a majority of participants selected the same wrong alternative plot.
+Additionally, we see a lot of variation in the number of different plots selected at the alternate plot by the Turkers.
+At most, there were 10 different plots selected, compared to 2 at the opposite end.
+\st{XXX is it worth exploring these ``interesting things" more in-depth?
+XXX}
+
+Next, we investigate the confidence of the experiment's participants in selecting the most different plot from the others.
+Overall, in 40\% of the responses, the respondent was certain they were correct.
+User confidence in the remaining categories, neutral, uncertain, very certain, and very uncertain, was 26.88\%, 16.88\%, 9.48\%, and 6.76\%, respectively.
+These results are summarized by lineup in Figure~\ref{fig:cert_plot}.
+We also performed a 5-sample test for equality of proportions in order to test the null hypothesis that the proportion of correct responses is the same in all 5 certainty categories.
+This test resulted in a $p$-value of 0.5881, so there is no evidence that the certainty of a user's response affects whether or not they choose the correct plot.
+This provides further evidence that detecting a network simulated from a different model is an extremely difficult problem.
+
+<<cert_plot, echo=FALSE, warning = FALSE, message = FALSE, fig.align='center', fig.height=4, fig.cap="All responses for all lineups, separated by whether the plot selected was the true alternative plot (TRUE) or not (FALSE) and colored by the respondent's uncertainty levels.
+We see here that most respondents were certain in their answer whether or not they were correct.">>=
+res2$conf_level <- ordered(res2$conf_level, levels = c("Very Uncertain", "Uncertain", "Neutral", "Certain", "Very Certain"))
+
+ggplot(data = res2, aes(x = lineup_name)) + 
+  geom_bar(aes(fill = conf_level), color = 'grey60') + 
+  scale_fill_brewer(palette = "RdBu") + 
+  coord_flip() + 
+  facet_wrap(~correct) + 
+  theme_bw() + 
+  theme(legend.position = 'bottom')
+
+summary.conf <- data.frame(table(res2$conf_level, res2$correct))
+summary.conf %>% tidyr::spread(Var2, Freq) -> summary.conf
+rownames(summary.conf) <- as.character(summary.conf[,1])
+names(summary.conf) <- c("Var1", "Failures", "Successes")
+summary.conf <- summary.conf[,-1]
+summary.conf <- summary.conf[,c(2,1)]
+prop_test_for_conf_corr <- prop.test(as.matrix(summary.conf))
+@
+
+We next investigate the length of time that users' took to respond to each lineup.
+The minimum was 3.762 seconds and the maximum was 578.8 seconds (nearly 10 minutes).
+The median was 13.01 seconds and the mean was 20.32 seconds.
+First, a 2-sample Kolmogorov-Smirnov test for equality of distribution was done to see if the distribution of times for correct plots chosen is the same as the distribution of times for incorrect plots chosen.
+This two-sided test resulted in a $p$-value of 1, so there is no evidence that the distribution of times is different whether or not the response was correct.
+
+<<time_plot, echo=FALSE>>=
+#qplot(x = res2$time, binwidth = 2)
+ggplot(data = res2, aes(x = conf_level, y = time)) + geom_boxplot() + coord_flip()
+ggplot(data = res2, aes(x = correct, y = time)) + geom_boxplot() + coord_flip() + theme_bw()
+ggplot(data = res2, aes(x = time)) + geom_histogram(binwidth = 5, fill ='white', color = 'black') + facet_wrap(~correct, nrow = 2)
+ggplot(data = res2, aes(x = lineup_name, y = time)) + 
+  geom_boxplot() + 
+#  facet_wrap(~conf_level) + 
+  coord_flip() + 
+  theme_bw() + 
+  labs(y = "Time (in seconds)", x = "Lineup Name")
+# 2-sample kolmogorov-smirnov test for equality of distributions
+# ks.test(x = summary(res2$time[res2$correct]), y = summary(res2$time[!res2$correct]))
+# fail to reject null
+@
+
+\subsection{Discussion/Future Work}
+
+This small experiment suggests that detecting a network model difference from just one simulation from one model and 11 from the other model is just about impossible.
+This result is fairly unsurprising: drawing a single random point from, say, a $\chi^2_1$ distribution and placing it in a lineup with 11 separate draws from a standard normal distribution would likely have similar results.
+Every once in a while, a chi-square value would appear that would be too large to belong with draws from a standard normal distribution, but values near 1, the mean, would probably not appear that different from some random standard normal draws.
+This means that we need to develop a different way to compare two network models.
+We need a way to visualize many samples from a network model in one panel.
+
 \bibliographystyle{apalike}

### 2016-09-01 Code/deciding_lineups.R

@@ -1,0 +1,185 @@
+# picking the lineups!!!!
+
+# M1v M2 
+
+# want size M = 12 with upper_qu = 1 /
+
+data.frame(lus_summary_2 %>% filter(part == "smallfriends" & 
+                         param_name == "transitive triplets jumping alcohol2" &  
+                         upper_qu == 1) %>% 
+  arrange(desc(data_est))) 
+
+# only M 12 rep 7
+
+# want size M = 12 with upper_qu large (>.5) and small data_dist
+data.frame(lus_summary_2 %>% filter(part == "smallfriends" & 
+                                    param_name == "transitive triplets jumping alcohol2" & 
+                                    upper_qu < 1 & upper_qu >=0.5 &
+                                    data_dist < 2) %>% 
+             arrange(data_dist)) 
+
+# M 12: rep 19, 11, 3, 4
+
+7, 19, 11, 3, 4
+
+# M2 v M1 
+
+# want size M = 12 with lower_qu = 1 
+
+data.frame(lus_summary_2b %>% filter(part == "smallfriends-rev" & 
+                                      param_name == "transitive triplets jumping alcohol2" &  
+                                      lower_qu == 1) %>% 
+             arrange(data_dist)) 
+
+# M 12: rep 18,4, 15
+
+# want size M = 12 with lower_qu large (>.5) and small data_dist
+
+data.frame(lus_summary_2b %>% filter(part == "smallfriends-rev" & 
+                                       param_name == "transitive triplets jumping alcohol2" &  
+                                       lower_qu < 1 & lower_qu >=.75 & 
+                                       data_dist < 2) %>% 
+             arrange(data_dist)) 
+# M 12: rep 2, 20
+
+18,4, 15, 2, 20 
+# M1 v M3
+
+# want size M = 12 with lower_qu = 1 
+
+data.frame(lus_summary_3 %>% filter(part == "smallfriends-eff2" & 
+                                       param_name == "number pairs at doubly achieved distance 2" &  
+                                       lower_qu == 1) %>% 
+             arrange(desc(data_dist))) 
+
+# NO M = 12
+
+# want size M = 12 with lower_qu large (>.5) and small data_dist
+
+data.frame(lus_summary_3 %>% filter(part == "smallfriends-eff2" & 
[…]

### 2016-09-01 Code/for_testing_experiment.R

@@ -1,0 +1,119 @@
+# make really easy lineups
+
+source("Code/03c_complete_lineup_creation.R")
+
+make_interactive_lineups_M1M2 <- function(M, reps, model.effects, null.effects){
+  for (r in 1:reps){
+    lineup <- create_smfriend_lu(null_eff_struct = null.effects, test_eff_struct = model.effects, M = M)
+    filename <- paste0("Data/lineupdata_experiment/smallfriends-m-", M, '-rep-', r, ".csv")
+    write.csv(lineup$data, sprintf("Data/lineupdata_experiment/smallfriends-m-%s-rep-%s.csv",M,r), row.names=FALSE)
+    #ggsave(lineup$lineup, file=sprintf("Lineup-Images/pdfs/smallfriends-m-%s-rep-%s.pdf",M,r), width = 7.2, height = 4.5, units = "in")
+    savepdf(file = sprintf("Lineup-Images/ForExperimentTraining/pdfs/smallfriends-m-%s-rep-%s",M,r), width = 7.2*2.54, height = 4.5*2.54)
+    print(lineup$lineup)
+    tmpfile <- sprintf("%s.svg",tempfile(tmpdir="Lineup-Images/ForExperimentTraining/svgs"))
+    make_interactive(filename2 = tmpfile, script=scriptURL,  
+                     high="#d5d5d5",  background="#ffffff")
+    dev.off()
+  }
+  return(list(lineup = lineup))
+}
+
+make_interactive_lineups_M1M2(M = 12, reps = 20, model.effects = null_model_eff2, null.effects = eff_models_smallFriends[[39]])
+
+make_interactive_lineups_M2M1 <- function(M, reps, model.effects, null.effects){
+  for (r in 1:reps){
+    lineup <- create_smfriend_lu(null_eff_struct = null.effects, test_eff_struct = model.effects, M = M)
+    filename <- paste0("Data/lineupdata_experiment/smallfriends-rev-m-", M, '-rep-', r, ".csv")
+    write.csv(lineup$data, sprintf("Data/lineupdata_experiment/smallfriends-rev-m-%s-rep-%s.csv",M,r), row.names=FALSE)
+    #ggsave(lineup$lineup, file=sprintf("Lineup-Images/pdfs/smallfriends-m-%s-rep-%s.pdf",M,r), width = 7.2, height = 4.5, units = "in")
[…]

### 2016-09-01 lineups.Rnw
code chunk facets_correct: reordering of bars not working? why?????

@@
 
-We collected 10 lineups each from 77 Amazon Turk users on July 28, 2016 from 19:11:03 GMT to 22:17:47 GMT.
-A table of demographic information collected on the participants in the experiment is in Table~\ref{fig:demos}.
+We collected 10 lineups each from 77 Amazon Turk users. A table of demographic information collected on the participants in the experiment is in Table~\ref{fig:demos}.
 
@@
 The x-axis ticks do not have a label because the label is less important than the number of different plots users chose and whether or not they chose the correct plot, shown in the coloring of the bars.", fig.height=9>>=
-ggplot(data = summ_res_wts, aes(x = as.factor(response_no2), 
+ggplot(data = summ_res_wts, 
+       aes(x = reorder(as.factor(response_no2), -tot_wt), 
                                 y = tot_wt,
@@
   theme_bw() + 
-  theme(axis.text.x = element_blank(), 
-        legend.position = 'bottom') + 
+  theme(legend.position = 'bottom') + 
   labs(x = "User Response" , y = "Number of Responses") 

### 2016-09-01 lineups.Rnw
anon'd tables

@@
 \caption{Example 1}
-\includegraphics[width=\textwidth]{Ex1}
+\includegraphics[width=\textwidth]{Results/Ex1}
 \end{subfigure}\hfill
@@
 \caption{Example 2}
-\includegraphics[width=\textwidth]{Ex2}
+\includegraphics[width=\textwidth]{Results/Ex2}
 \end{subfigure}
@@
 \caption{Trial 1}
-\includegraphics[width=\textwidth]{Trial1}
+\includegraphics[width=\textwidth]{Results/Trial1}
 \end{subfigure}\hfill
@@
 \caption{Trial 2}
-\includegraphics[width=\textwidth]{Trial2}
+\includegraphics[width=\textwidth]{Results/Trial2}
 \end{subfigure}
@@
 ids <- readr::read_csv("https://raw.githubusercontent.com/erichare/lineups/f19635a00507210013ba9e631761ad0c89a6564d/experiments/turk21/details/picture-details.csv")
-tab <- read.csv("../Results/turk21_users.csv")
+tab <- read.csv("Results/turk21_users.csv")
 tab2 <- unique(tab)
@@
 #tab2 %>% filter(nick_name %in% names(table(res$nick_name))[table(res$nick_name) != 10])
-res <- read.csv("../Results/turk21_feedback.csv", stringsAsFactors = F)
+res <- read.csv("Results/turk21_feedback.csv", stringsAsFactors = F)
 # remove people who only completed 1 plot
@@
 t1 <- t1[2,]
+t1[as.numeric(t1) < 5] <- "*"
 kable(t1, format = 'latex')
@@
 t3 <- t3[2,]
+t3[as.numeric(t3) < 5] <- "*"
 kable(t3, format = 'latex')
@@
 t2 <- t2[2,]
+names(t2)
+t2.2 <- t2
+t2.2 <- t2.2[1:7]
+names(t2.2)[6:7] <- c("46-55", "Over 55")
+t2.2[6:7] <- c(sum(as.numeric(t2[6:7])), sum(as.numeric(t2[8:9])))
 kable(t2, format = 'latex')
@@
 \end{subfigure}
-\caption{\label{fig:demos} Demographic information on the Amazon Turk users participating in the experiment}
+\caption{\label{fig:demos} Demographic information on the Amazon Turk users participating in the experiment. Asterisks (*) indicate less than 5 responses.}
 \end{figure}

### 2016-09-02 lineups.Rnw
remove results that were printing that shouldn't have been

@@
 res_rpt <- res[grep(",", res$response_no),]
-res_rpt$response_no
+
 library(stringr)
@@
 split_rows <- strsplit(rpt_rows2$response_no, ",")
+rpt_rows2$tot_response <- rep(ttr, ttr)
 rpt_rows2$response_no2 <- NA
@@
 res_norpt$count_response <- 1 
+res_norpt$weight <- 1
 res_norpt$response_no2 <- res_norpt$response_no
 res_norpt$tot_response <- 1
-res_norpt$weight <- 1
+
 # res with weight column   
@@
               # jtts of 40, 24, 18,17, 9
-head(res3)
 res3$group_name <- paste0(res3$group, " #", res3$group.rep)
@@
 t2 <- t2[2,]
-names(t2)
 t2.2 <- t2
@@
 t2.2[6:7] <- c(sum(as.numeric(t2[6:7])), sum(as.numeric(t2[8:9])))
-kable(t2, format = 'latex')
+kable(t2.2, format = 'latex')
 @

### 2016-09-02 lineups.Rnw
add more description of lineups and how they were chosen

@@
 \subsection{Methods}
-In order to test our hypothesis, we set up an amazon turk experiment using the lineup protocal of {\color{green} Buja et all (2009).
-FIX CITE}.
+In order to test our hypothesis, we set up an Amazon Mechanical Turk experiment (~\cite{turk}) using the lineup protocal of ~\cite{Bujaetal}.
 We presented four types of lineups: M1 v.
@@
 M3, and M3 v.
-M1, where M1, M2, and M3 have the objective functions givein in ~\ref{tab:models}.
-For each lineup, there were 12 plots shown: 11 of the plots were simulated from the first model (the ``null" model) and 1 was simulated from the second model (the ``alternative" model).
-The order of the plots in the lineups was randomly assigned.
+M1, where M1, M2, and M3 have the objective functions given in ~\ref{tab:models}.
+We created a total of 25 lineups to show to Amazon Turk users taking our experiment: 10 for M1 v.
+M2 and 5 each for the other three types of lineups.
+In each lineup, there were 12 plots shown: 11 of the plots were simulated from the first model (the ``null" model) and 1 was simulated from the second model (the ``alternative" model).
+We chose to show lineups of size 12 because we felt that showing more than 12 would be too large of a cognitive load, while showing fewer than 12 would leave too much of the experiment to random chance.
+The order of the plots within the lineups was randomly assigned.
+We selected five lineups presented from each group based on the following criteria: first, for M1 v M2 and M1 v M3, we selected the lineups where the additional parameter in the objective function, $\beta_3$ and $\beta_4$, respectively, had the largest estimated value among the 12 networks presented.
+There were not many of these in the size 12 lineups, so our other selection criteria was that the estimated parameter value was larger in the alternative panel than in at least half of the other lineups and it was close to the estimate from the panel that did have the largest value.
+For the M2 v were chosen where the smallest parameter estimate belonged to the alternative model.
+If there more lineups were needed, we next selected those which had estimates smaller than at least half of the other panels and were close to the minimum estimate in the lineup.
+For the remaining five plots of type M1 v M2, we chose the lineups where the alternative plot had the largest number of jumping transitive triplets appear in the network.
+We chose this  statistic because its value has great effect on both the estimation of the $\beta_3$ parameter and on the visual appearance of the plot.
 

### 2017-09-23 ASApaper/paper.Rmd
add senate stuffs

---
title: Visual Inference for Goodness-of-Fit for A Network Model

# to produce blinded version set to 1
blinded: 0

authors: 
- name: Sam Tyner
  thanks: The authors gratefully acknowledge ...
  affiliation: Department of Statistics, Iowa State University

- name: Heike Hofmann
  affiliation: Department of Statistics, Iowa State University

keywords:
- statistics
- social networks
- more

abstract: |
  The text of your abstract.  200 or fewer words.

bibliography: bibliography.bib
output: rticles::asa_article
---

# Introduction

# Data

# Model Definition and Choice

## Choosing appropriate models

In order to determine which models to explore for application of our novel goodness-of-fit testing method, we first fit many possible models and selected a few significant effects. To determine the significance of the effects, we followed this procedure:

1. Define the simple effects structure of the data: the rate parameters and the outdegree and reciprocity parameters. 
2. Add each possible effect one-at-a-time to the model structure, as determined by the effects documentation function.
3. Fit each model to the data and check for convergence.
    a. If the model converged, move to 4.
    b. If the model did not converge, use the previous fitted values as starting values and repeat 5 times or until convergence, whichever comes first.
4. Test the added parameter for significance using the Wald-type test.
5. Report out the estimate of the additional parameter, its standard error, Wald $p$ value, and convergence criterion.

After completing the procedure for all model effects, we selected effects to include that converged and had a Wald $p$-value of less than 0.10, and seemed to have a reasonable interpretation for our data.

# Simulation Procedure

1. Fit each model 1,000 times and save the parameter estimates, standard errors, and overall convergence criterion.
2. (From the models that converged ?) Compute the mean of the 1,000 estimates of each parameter.
3. Use the mean values from step 2 as the parameter values in the model and simulate 100 observations from the model. These simulated networks will become the "true data."
4. To each simulation, refit the model form from which it was simulated. 
5. Repeat steps 4-5 for various permutations of model forms and parameter values; e.g. double the parameter values, quadruple the parameter values, change the signs on the parameters, etc.
6. Simulate additional "alternative data" from the model in step 3 and the fitted model in step 4. 
7. Place the appropriate node-link diagram of each "true data" observation in a lineup with $M-1$ (M=12? M=9?) "alternative data" node-link diagrams. 
8. Show the lineups to participants, with no participant seeing the same "true data" or "alternative data" plot twice.

<!--
This template demonstrates some of the basic latex you'll need to know to create a ASA article.

\section{Verifications}
\label{sec:verify}
This section will be just long enough to illustrate what a full page of
text looks like, for margins and spacing.

@Campbell02 @Schubert13 [@Chi81]

The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
The quick brown fox jumped over the lazy dog.
__With this spacing we have 30 lines per page.__
-->

### 2017-09-23 Code/Senate/combineSenatedataFits.R

@@ -1,0 +1,23 @@
+# combine data from server
+library(tidyverse)
+M1sen1 <- read_csv("Data/senate/basicModelFitsSenate.csv")
+M1sen2 <- read_csv("Data/senate/basicModelFitsSenate2.csv")
+M1sen3 <- read_csv("Data/senate/basicModelFitsSenate3.csv")
+M1sen4 <- read_csv("Data/senate/basicModelFitsSenate4.csv")
+M1sen4 <- M1sen4[,-1]
+M1sen5 <- read_csv("Data/senate/basicModelFitsSenate5.csv")
+M1sen6 <- read_csv("Data/senate/basicModelFitsSenate6.csv")
+M1sen6 <- M1sen6[,-1]
+M1sen7 <- read_csv("Data/senate/basicModelFitsSenate7.csv")
+M1sen8 <- read_csv("Data/senate/basicModelFitsSenate8.csv")
+M1sen8 <- M1sen8[,-1]
+M1sen9 <- read_csv("Data/senate/basicModelFitsSenate9.csv")
+M1sen10 <- read_csv("Data/senate/basicModelFitsSenate10.csv")
+M1sen10 <- M1sen10[,-1]
+
+allM1sen <- do.call(rbind, list(M1sen1,M1sen2, M1sen3, M1sen4, M1sen5, M1sen6,
+                    M1sen7, M1sen8, M1sen9, M1sen10))
+allM1sen$sim <- 1:nrow(allM1sen)
+#write_csv(allM1sen, "Data/senate/basicModelFitsSenate1000.csv")
+
+

### 2017-09-23 Code/Senate/fittingModelstoSenatedata.R

@@ -1,0 +1,74 @@
+# repeatedly fit models to data
+library(tidyverse)
+library(RSienaTest)
+library(RSiena)
+SenateEffects <- read_csv("Data/senate/sigEffsSenate912.csv")
+sigSenEffects <- SenateEffects %>% 
+  filter(overallConverge < 0.25 & Waldpval < 0.10 & !is.na(shortName)) %>% 
+  arrange(Waldpval) %>% mutate(fullEffect = ifelse(!is.na(inter1), paste(shortName, inter1), shortName))
+
+fullEffectNames <- c("jumpXTransTrip party", "jumpXTransTrip sex", "inIsDegree", "sameX party", 
+                "simXTransTrip bills", "simRecipX party", "transTrip", "sameXTransTrip party")
+
+sigSenEffects %>% filter(fullEffect %in% fullEffectNames) -> effectsIWant
+
+load("Data/senate/senateSienaNoHRC.rda")
+
+SenBasic <- getEffects(senateSiena)
+Senjtt_p <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+Senjtt_s <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "sex", character = TRUE)
+SeninIsD <- includeEffects(SenBasic, "inIsDegree", include = TRUE, type = "eval", interaction1 = "", character = TRUE)
+Sensame_p <- includeEffects(SenBasic, "sameX", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+Senstt_b <- includeEffects(SenBasic, "simXTransTrip", include = TRUE, type = "eval", interaction1 = "bills", character = TRUE)
+Sensimre_p <- includeEffects(SenBasic, "simRecipX", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+Sentt_p <- includeEffects(SenBasic, "transTrip", include = TRUE, type = "eval", interaction1 = "", character = TRUE)
+Senstt_p <- includeEffects(SenBasic, "sameXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+
+require(RSiena)
+struct <- Senjtt_p
+myalg <- sienaAlgorithmCreate( projname = Sys.time() , n3 = 1000)
+nparm <- sum(struct$include)
[…]

### 2017-09-23 Code/Senate/scriptforServer.R

@@ -1,0 +1,52 @@
+library(RSiena)
+dat <- load("senateSienaNoHRC.rda")
+# basic effects structure
+SenBasic <- getEffects(senateSiena)
+# basic algorithm structure
+myalg <- sienaAlgorithmCreate( projname = Sys.time() , n3 = 1000)
+# additional effects structures
+Senjtt_p <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+# Senjtt_s <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "sex", character = TRUE)
+# SeninIsD <- includeEffects(SenBasic, "inIsDegree", include = TRUE, type = "eval", interaction1 = "", character = TRUE)
+# Sensame_p <- includeEffects(SenBasic, "sameX", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+# Senstt_b <- includeEffects(SenBasic, "simXTransTrip", include = TRUE, type = "eval", interaction1 = "bills", character = TRUE)
+# Sensimre_p <- includeEffects(SenBasic, "simRecipX", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+# Sentt_p <- includeEffects(SenBasic, "transTrip", include = TRUE, type = "eval", interaction1 = "", character = TRUE)
+# Senstt_p <- includeEffects(SenBasic, "sameXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+
+get_effects_dist <- function(dat, struct, N){
+  require(RSiena)
+  myalg <- sienaAlgorithmCreate( projname = Sys.time() , n3 = 1000)
+  nparm <- sum(struct$include)
+  Parmnames <- struct$shortName[struct$include]
+  Nnr <- sum(Parmnames != "Rate")
+  df.ests <- data.frame(matrix(0, nrow = N, ncol = nparm + Nnr + 1))
+  
+  names(df.ests) <- c(Parmnames,
+                      paste("se", Parmnames[which(Parmnames!="Rate")], sep = "_"),
+                      "maxConv")
+  for (i in 1:N){
+    fits <- siena07(myalg, data = dat, effects = struct, returnDeps = TRUE,
[…]

### 2017-09-23 Code/Senate/scriptforServerparallel.R

@@ -1,0 +1,40 @@
+library(parallel)
+library(RSiena)
+load("Data/senate/senateSienaNoHRC.rda")
+dat <- senateSiena
+SenBasic <- getEffects(senateSiena)
+struct <- SenBasic
+
+get_effects_dist <- function(dat, struct, N){
+  myalg <- sienaAlgorithmCreate( projname = Sys.time() , n3 = 1000)
+  no_cores <- detectCores()
+  cl <- makeCluster(no_cores, outfile = "", renice = 0)
+  clusterExport(cl, varlist = c("myalg", "dat", "struct"))
+
+  mycoefs <- parLapply(cl, 1:N, function(i, dat, struct, myalg) {
+      library(RSiena)
+    
+    fits <- siena07(myalg, data = dat, effects = struct, returnDeps = TRUE,
+              batch=TRUE, verbose = FALSE, silent = TRUE)
+
+    return(c(fits$rate, fits$theta, sqrt(diag(fits$covtheta)), fits$tconv.max))
+  })
+
+  df.ests <- as.data.frame(do.call(rbind, mycoefs))
+
+  nparm <- sum(struct$include)
+  Parmnames <- struct$shortName[struct$include] 
+  
+  names(df.ests) <- c(Parmnames,
+                      paste("se", Parmnames[which(Parmnames!="Rate")], sep = "_"),
+                      "maxConv")
+  #for (i in 1:N){
+  #  fits <- siena07(myalg, data = dat, effects = struct, returnDeps = TRUE,
+  #                  batch=TRUE, verbose = FALSE, silent = TRUE)
+  #  df.ests[i,] <- c(fits$rate, fits$theta, sqrt(diag(fits$covtheta)), fits$tconv.max)
+  #}
+  return(df.ests)
+}
+
+M1Senate <- get_effects_dist(dat = senateSiena, struct = SenBasic, N = 3)
+write.csv(M1Senate, "basicModelFitsSenate5.csv")

### 2017-09-23 Code/Senate/senate_siena_nohrc.R

@@ -70,6 +70,7 @@ rownames(nbills_au) <- senateCovars$source
 bills <- varCovar(nbills_au)
 # rsiena objects are party, sex, bills, senateData2
 senateSiena <- sienaDataCreate( senateData2, bills, sex, party )
+save(senateSiena, file = "Data/senate/senateSienaNoHRC.rda")
 senateModelEff <- getEffects(senateSiena)
 #setwd("Data/senate")
 #effectsDocumentation(senateModelEff)

### 2017-09-23 Code/Senate/significanteffects_senate.R

@@ -20,3 +20,20 @@ ggplot(data = test_res, aes(x = log(Waldpval), y = overallConverge)) +
   scale_color_brewer(palette = "YlOrRd") + 
   geom_hline(yintercept = .25, color = 'red') + 
   geom_vline(xintercept = c(log(.01),log(.05), log(.10)), color = c('blue', "green", "magenta"))
+
+test_res %>% filter(Waldpval < 0.1 & overallConverge <= .25) %>% 
+  arrange(Waldpval) %>% 
+  mutate(overallConverge = round(overallConverge,4),
+         Waldpval = ifelse(Waldpval < 0.001, "<0.001", round(Waldpval, 4))) -> sigEffsConverged 
+# write_csv(sigEffsConverged, "Data/senate/sigEffsSenateConverged.csv")
+
+ggplot(data = sigEffsConverged, aes(x = readr::parse_number(Waldpval), y = overallConverge)) + 
+  geom_line() + 
+  geom_point(aes(color = as.factor(refits))) + 
+  scale_color_brewer(palette = "YlOrRd") + 
+  geom_hline(yintercept = .25, color = 'red') + 
+  geom_vline(xintercept = c(.01,.05, .10), color = c('blue', "green", "magenta"))
+
+
+
+

### 2017-09-26 ASApaper/paper.Rmd
add data for senate example; beginning of paper as well

@@
 ---
+
+```{r pkgs, echo = FALSE}
+library(tidyverse)
+library(RSienaTest)
+library(RSiena)
+library(geomnet)
+library(GGally)
+```
 
@@
 
+# Exploratory plots
+
+```{r M1}
+M1senateEsts <- read_csv("../Data/senate/basicModelFitsSenate1000.csv")
+inisdSenateEsts <- read_csv("../Data/senate/inisdModelFitsSenate1000.csv")
+# ggpairs(M1senateEsts, columns = 1:8)
+# M1senateEsts %>% mutate(converged = (maxConv <= 0.25)) %>% 
+#   ggplot() + 
+#   geom_density(aes(x = density, color = converged), adjust = .5) 
+# M1senateEsts %>% mutate(converged = (maxConv <= 0.25)) %>% 
+#   ggplot() + 
+#   geom_histogram(aes(x = density, color = converged), binwidth = .005, fill = NA) + facet_wrap(~converged)
+
+M1senateEsts %>% 
+  select(Rate:recip, maxConv, sim) %>% 
+  gather(parameter, estimate, Rate:recip) -> melt1 
+M1senateEsts %>% 
+  select(se_density:sim) %>% 
+  gather(parameter, se, se_density:se_recip) %>% 
+  mutate(parameter = str_replace(parameter, "se_", "")) -> melt2
+
+left_join(melt1, melt2) %>% arrange(sim) -> M1SenEstsLong
+
+inisdSenateEsts %>% 
+  select(Rate:inIsDegree, maxConv, sim) %>% 
+  gather(parameter, estimate, Rate:inIsDegree) -> melt1 
+inisdSenateEsts %>% 
+  select(se_density:sim) %>% 
+  gather(parameter, se, se_density:se_inIsDegree) %>% 
+  mutate(parameter = str_replace(parameter, "se_", "")) -> melt2 
+left_join(melt1, melt2) %>% arrange(sim) -> inisdSenEstsLong
+
+```
+
 <!--

### 2017-09-26 Code/Senate/combineSenatedataFits.R

@@ -20,4 +20,49 @@ allM1sen <- do.call(rbind, list(M1sen1,M1sen2, M1sen3, M1sen4, M1sen5, M1sen6,
 allM1sen$sim <- 1:nrow(allM1sen)
 #write_csv(allM1sen, "Data/senate/basicModelFitsSenate1000.csv")
 
+inisdsen1 <- read_csv("Data/senate/inisdModelFitsSenate1-10.csv")
+inisdsen1.2 <- read_csv("Data/senate/inisdModelFitsSenate111-100.csv")
+inisdsen2 <- read_csv("Data/senate/inisdModelFitsSenate200.csv")
+inisdsen3 <- read_csv("Data/senate/inisdModelFitsSenate300.csv")
+inisdsen4 <- read_csv("Data/senate/inisdModelFitsSenate400.csv")
+inisdsen5 <- read_csv("Data/senate/inisdModelFitsSenate500.csv")
+inisdsen6 <- read_csv("Data/senate/inisdModelFitsSenate600.csv")
+inisdsen7 <- read_csv("Data/senate/inisdModelFitsSenate700.csv")
+inisdsen8 <- read_csv("Data/senate/inisdModelFitsSenate800.csv")
+inisdsen9 <- read_csv("Data/senate/inisdModelFitsSenate900.csv")
+inisdsen10 <- read_csv("Data/senate/inisdModelFitsSenate1000.csv")
 
+allinisdsen <- do.call(rbind, list(inisdsen1, inisdsen1.2, inisdsen2, inisdsen3, inisdsen4,
+                                   inisdsen5, inisdsen6, inisdsen7, inisdsen8, inisdsen9, inisdsen10))
+allinisdsen <- allinisdsen[,-1]
+allinisdsen$sim <- 1:nrow(allinisdsen)
+# write_csv(allinisdsen, "Data/senate/inisdModelFitsSenate1000.csv")
+
+jttpSen1 <- read_csv("Data/senate/jttpModelFitsSenate100.csv")
+jttpSen2 <- read_csv("Data/senate/jttpModelFitsSenate200.csv")
+jttpSen3 <- read_csv("Data/senate/jttpModelFitsSenate300.csv")
+jttpSen4 <- read_csv("Data/senate/jttpModelFitsSenate400.csv")
+jttpSen5 <- read_csv("Data/senate/jttpModelFitsSenate500.csv")
+jttpSen6 <- read_csv("Data/senate/jttpModelFitsSenate600.csv")
[…]

### 2017-09-26 Code/Senate/scriptforServer.R

@@ -1,15 +1,15 @@
 library(RSiena)
-dat <- load("senateSienaNoHRC.rda")
+load("senateSienaNoHRC.rda")
 # basic effects structure
 SenBasic <- getEffects(senateSiena)
 # basic algorithm structure
 myalg <- sienaAlgorithmCreate( projname = Sys.time() , n3 = 1000)
 # additional effects structures
-Senjtt_p <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
-# Senjtt_s <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "sex", character = TRUE)
+#Senjtt_p <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+#Senjtt_s <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "sex", character = TRUE)
 # SeninIsD <- includeEffects(SenBasic, "inIsDegree", include = TRUE, type = "eval", interaction1 = "", character = TRUE)
 # Sensame_p <- includeEffects(SenBasic, "sameX", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
-# Senstt_b <- includeEffects(SenBasic, "simXTransTrip", include = TRUE, type = "eval", interaction1 = "bills", character = TRUE)
+ Senstt_b <- includeEffects(SenBasic, "simXTransTrip", include = TRUE, type = "eval", interaction1 = "bills", character = TRUE)
 # Sensimre_p <- includeEffects(SenBasic, "simRecipX", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
 # Sentt_p <- includeEffects(SenBasic, "transTrip", include = TRUE, type = "eval", interaction1 = "", character = TRUE)
 # Senstt_p <- includeEffects(SenBasic, "sameXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
@@ -26,15 +26,56 @@ get_effects_dist <- function(dat, struct, N){
                       paste("se", Parmnames[which(Parmnames!="Rate")], sep = "_"),
                       "maxConv")
[…]

### 2017-09-26 Code/Senate/scriptforServerparallel.R

@@ -1,40 +1,73 @@
-library(parallel)
 library(RSiena)
-load("Data/senate/senateSienaNoHRC.rda")
-dat <- senateSiena
+load("senateSienaNoHRC.rda")
+# basic effects structure
 SenBasic <- getEffects(senateSiena)
-struct <- SenBasic
+# basic algorithm structure
+myalg <- sienaAlgorithmCreate( projname = Sys.time() , n3 = 1000)
+# additional effects structures
+#Senjtt_p <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+#Senjtt_s <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "sex", character = TRUE)
+# SeninIsD <- includeEffects(SenBasic, "inIsDegree", include = TRUE, type = "eval", interaction1 = "", character = TRUE)
+# Sensame_p <- includeEffects(SenBasic, "sameX", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+Senstt_b <- includeEffects(SenBasic, "simXTransTrip", include = TRUE, type = "eval", interaction1 = "bills", character = TRUE)
+# Sensimre_p <- includeEffects(SenBasic, "simRecipX", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+# Sentt_p <- includeEffects(SenBasic, "transTrip", include = TRUE, type = "eval", interaction1 = "", character = TRUE)
+# Senstt_p <- includeEffects(SenBasic, "sameXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
 
 get_effects_dist <- function(dat, struct, N){
+  require(RSiena)
   myalg <- sienaAlgorithmCreate( projname = Sys.time() , n3 = 1000)
-  no_cores <- detectCores()
-  cl <- makeCluster(no_cores, outfile = "", renice = 0)
-  clusterExport(cl, varlist = c("myalg", "dat", "struct"))
-
-  mycoefs <- parLapply(cl, 1:N, function(i, dat, struct, myalg) {
-      library(RSiena)
-    
-    fits <- siena07(myalg, data = dat, effects = struct, returnDeps = TRUE,
[…]

### 2017-09-27 ASApaper/paper.Rmd
discussion in paper

--- 2017-09-23 ASApaper/paper.Rmd
@@
 
+```{r pkgs, echo = FALSE}
+library(tidyverse)
+library(RSienaTest)
+library(RSiena)
+library(geomnet)
+library(GGally)
+```
+
 # Introduction
@@
 
+The most significant effect was the jumping transitive triplet (JTT) parameter for the party covariate, which was estimated to be about -6 with a standard error of 0.11.
+This effect relies on the number of transitive closures formed between two senators from different parties.
+Due to the divisive nature of American politics, the party effect is dominant.
+Another significant effect was the same JTT parameter for the sex covariate, with an estimate of about 3 with a standard error of 0.89.
+Another significant effect was the covariate-related similarity score-weighted transitive triplets for the number of bills authored by a senator.
+This effect was estimated at about 10 with standard error of 3.9, and the high positive effect suggests senators tend to collaborate with other senators who author about the same number of bills they do.
+
 # Simulation Procedure
@@
 8. Show the lineups to participants, with no participant seeing the same "true data" or "alternative data" plot twice.
+
+# Exploratory plots
+
+```{r M1}
+M1senateEsts <- read_csv("../Data/senate/basicModelFitsSenate1000.csv")
+inisdSenateEsts <- read_csv("../Data/senate/inisdModelFitsSenate1000.csv")
+jttpSenateEsts <- read_csv("../Data/senate/jttpModelFitsSenateAll1000.csv")
+jttsSenateEsts <- read_csv("../Data/senate/jttsModelFitsSenateAll1000.csv")
+# ggpairs(M1senateEsts, columns = 1:8)
+# M1senateEsts %>% mutate(converged = (maxConv <= 0.25)) %>% 
+#   ggplot() + 
+#   geom_density(aes(x = density, color = converged), adjust = .5) 
+# M1senateEsts %>% mutate(converged = (maxConv <= 0.25)) %>% 
+#   ggplot() + 
+#   geom_histogram(aes(x = density, color = converged), binwidth = .005, fill = NA) + facet_wrap(~converged)
+
+# make ests long - M1
+M1senateEsts %>% 
+  select(Rate:recip, maxConv, sim) %>% 
+  gather(parameter, estimate, Rate:recip) -> melt1 
+M1senateEsts %>% 
+  select(se_density:sim) %>% 
+  gather(parameter, se, se_density:se_recip) %>% 
+  mutate(parameter = str_replace(parameter, "se_", "")) -> melt2 
+left_join(melt1, melt2) %>% arrange(sim) -> M1SenEstsLong
+
+# make ests long - inisd
+inisdSenateEsts %>% 
+  select(Rate:inIsDegree, maxConv, sim) %>% 
+  gather(parameter, estimate, Rate:inIsDegree) -> melt1 
+inisdSenateEsts %>% 
+  select(se_density:sim) %>% 
+  gather(parameter, se, se_density:se_inIsDegree) %>% 
+  mutate(parameter = str_replace(parameter, "se_", "")) -> melt2 
+left_join(melt1, melt2) %>% arrange(sim) -> inisdSenEstsLong
+
+# make ests long - jttp
+jttpSenateEsts %>% 
+  select(Rate:jumpXTransTrip, maxConv, sim) %>% 
+  gather(parameter, estimate, Rate:jumpXTransTrip) -> melt1 
+jttpSenateEsts %>% 
+  select(se_density:maxConv, sim) %>% 
+  gather(parameter, se, se_density:se_jumpXTransTrip) %>% 
+  mutate(parameter = str_replace(parameter, "se_", "")) -> melt2 
+left_join(melt1, melt2) %>% arrange(sim) -> jttpSenLong
+
+# make ests long - jtts
+jttsSenateEsts %>% 
+  select(Rate:jumpXTransTrip, maxConv, sim) %>% 
+  gather(parameter, estimate, Rate:jumpXTransTrip) -> melt1 
+jttsSenateEsts %>% 
+  select(se_density:maxConv, sim) %>% 
+  gather(parameter, se, se_density:se_jumpXTransTrip) %>% 
+  mutate(parameter = str_replace(parameter, "se_", "")) -> melt2 
+left_join(melt1, melt2) %>% arrange(sim) -> jttsSenLong
+
+```
 

### 2017-09-27 Code/Senate/combineSenatedataFits.R

@@ -51,18 +51,34 @@ jttpSen10 <- read_csv("Data/senate/jttpModelFitsSenate1000.csv")
 
 alljttpsen <- do.call(rbind, list(jttpSen1, jttpSen2, jttpSen3, jttpSen4,
                       jttpSen5,jttpSen6,jttpSen7,jttpSen8,jttpSen9,jttpSen10))
-head(alljttpsen)
-alljttpsen <- alljttpsen[,-1]
+names(alljttpsen)[1] <- "sim"
+alljttpsen$sim <- 1:nrow(alljttpsen)
 #write_csv(alljttpsen, "Data/senate/jttpModelFitsSenateAll1000.csv")
 
 alljttpsen %>% mutate(converged = maxConv <= 0.25) %>% 
 ggplot() + 
   geom_density(aes(x = jumpXTransTrip, color = converged))
 
-
 alljttpsen %>% mutate(lower = jumpXTransTrip - 1.96*se_jumpXTransTrip, 
                    upper = jumpXTransTrip + 1.96*se_jumpXTransTrip,
                    issig = !(0 < upper & 0 > lower)) %>% 
 ggplot() + geom_histogram(aes(jumpXTransTrip, fill = issig), position = "fill")
   
-    ggplot() + geom_density(aes(x = jumpXTransTrip, color = issig))
+# jtts 
+
+jttsSen1 <- read_csv("Data/senate/jttsModelFitsSenate100.csv")
+jttsSen2 <- read_csv("Data/senate/jttsModelFitsSenate200.csv")
+jttsSen3 <- read_csv("Data/senate/jttsModelFitsSenate300.csv")
+jttsSen4 <- read_csv("Data/senate/jttsModelFitsSenate400.csv")
+jttsSen5 <- read_csv("Data/senate/jttsModelFitsSenate500.csv")
+jttsSen6 <- read_csv("Data/senate/jttsModelFitsSenate600.csv")
+jttsSen7 <- read_csv("Data/senate/jttsModelFitsSenate700.csv")
+jttsSen8 <- read_csv("Data/senate/jttsModelFitsSenate800.csv")
+jttsSen9 <- read_csv("Data/senate/jttsModelFitsSenate900.csv")
+jttsSen10 <- read_csv("Data/senate/jttsModelFitsSenate1000.csv")
+
+alljttssen <- do.call(rbind, list(jttsSen1, jttsSen2, jttsSen3, jttsSen4,
[…]

### 2017-09-27 Code/Senate/scriptforServer.R

@@ -1,5 +1,5 @@
 library(RSiena)
-load("senateSienaNoHRC.rda")
+load("Data/senate/senateSienaNoHRC.rda")
 # basic effects structure
 SenBasic <- getEffects(senateSiena)
 # basic algorithm structure
@@ -8,8 +8,8 @@ myalg <- sienaAlgorithmCreate( projname = Sys.time() , n3 = 1000)
 #Senjtt_p <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
 #Senjtt_s <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "sex", character = TRUE)
 # SeninIsD <- includeEffects(SenBasic, "inIsDegree", include = TRUE, type = "eval", interaction1 = "", character = TRUE)
-# Sensame_p <- includeEffects(SenBasic, "sameX", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
- Senstt_b <- includeEffects(SenBasic, "simXTransTrip", include = TRUE, type = "eval", interaction1 = "bills", character = TRUE)
+Sensame_p <- includeEffects(SenBasic, "sameX", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+# Senstt_b <- includeEffects(SenBasic, "simXTransTrip", include = TRUE, type = "eval", interaction1 = "bills", character = TRUE)
 # Sensimre_p <- includeEffects(SenBasic, "simRecipX", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
 # Sentt_p <- includeEffects(SenBasic, "transTrip", include = TRUE, type = "eval", interaction1 = "", character = TRUE)
 # Senstt_p <- includeEffects(SenBasic, "sameXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
@@ -72,8 +72,9 @@ write.csv(simttbFitsSenate, "Data/senate/simttbModelFitsSenate1000.csv")
 # write.csv(samepFitsSenate, "Data/senate/samepModelFitsSenate800.csv")
 # samepFitsSenate <- get_effects_dist(dat = senateSiena, struct = Sensame_p, N = 100)
 # write.csv(samepFitsSenate, "Data/senate/samepModelFitsSenate900.csv")
[…]

### 2017-09-28 ASApaper/paper.Rmd
add paper changes

@@
 
-```{r pkgs, echo = FALSE}
+```{r pkgs, echo = FALSE, message=FALSE,warning=FALSE}
 library(tidyverse)
@@
 
+Talk about goodness-of-fit
+
+Talk about political networks
+
 # Data
 
+The data we use are of collaboration networks in the United States Senate during the $111^{th}$ through $114^{th}$ Congresses, overlapping with Barack Obama's presidency.
+These senates began on January 6, 2009 and ended on January 3, 2017.^[Details of how this data can be downloaded are provided by François Briatte at https://github.com/briatte/congress].
+In this network, ties are directed from senator $i$ to senator $j$ when senator $i$ signs on as a cosponsor to the bill that senator $j$ authored.
+There are (somewhat surprisingly) many hundreds of ties between senators when they are connected in this way, so we simplify the network by computing a single value for each senator-senator collaboration called the *weighted propensity to cosponsor* (WPC).
+This value is defined in @senate as
+
+\begin{equation}\label{eq:sen1}
+    WPC_{ij} = \dfrac{\sum\limits_{k=1}^{n_j} \frac{Y_{ij(k)}}{c_{j(k)}}}{\sum\limits_{k=1}^{n_j} \frac{1}{c_{j(k)}}}
+\end{equation}
+
+where $n_j$ is the number of bills in a congressional session authored by senator $j$, $c_{j(k)}$ is the number of cosponsors on senator $j$'s $k^{th}$ bill, where $k \in \{1,\dots, n_j\}$, and $Y_{ij(k)}$ is a binary variable that is 1 if senator $i$ cosponsored senator $j$'s $k^{th}$ bill, and is 0 otherwise.
+This measure ranges in value from 0 to 1, where $WPC_{ij} = 1$ if senator $i$ is a cosponsor on every one of senator $j$'s bills and $WPC_{ij} = 0$ if senator $i$ is never a cosponsor any of senator $j$'s bills.
+
 # Model Definition and Choice
+
+To model family that we apply to the Senate data is the Stochastic Actor-Oriented Model (SAOM) family.
+This family of models incorporates both network structure and node-level information to describe how a network changes over time.
+This model was first introduced by Snijders in 1996 [@saompaper].
+The two titular properties of SAOMs, stochasticity and actor-orientation, are crucial to understanding networks as they exist naturally: social networks are ever-changing as relationships decay or grow in seemingly random ways, and most actors in them have characteristics that could affect how they change their ties to other nodes in the network.
+
+In cosponsorship networks, it is well known that party affiliation and reciprocity of relationships are major influences on structure [@legnet].
 
@@
 
-1. Define the simple effects structure of the data: the rate parameters and the outdegree and reciprocity parameters. 
-2. Add each possible effect one-at-a-time to the model structure, as determined by the effects documentation function.
-3. Fit each model to the data and check for convergence.
-    a. If the model converged, move to 4.
-    b. If the model did not converge, use the previous fitted values as starting values and repeat 5 times or until convergence, whichever comes first.
-4. Test the added parameter for significance using the Wald-type test.
-5. Report out the estimate of the additional parameter, its standard error, Wald $p$ value, and convergence criterion.
+\begin{enumerate}
+\item Define the simple effects structure of the data: the rate parameters and the outdegree and reciprocity parameters. 
+\item Add each possible effect one-at-a-time to the model structure, as determined by the effects documentation function.
+\item Fit each model to the data and check for convergence.
+    \begin{enumerate}
+    \item If the model converged, move to 4.
+    \item If the model did not converge, use the previous fitted values as starting values and repeat 5 times or until convergence, whichever comes first.
+    \end{enumerate}
+\item Test the added parameter for significance using the Wald-type test.
+\item Report out the estimate of the additional parameter, its standard error, Wald $p$ value, and convergence criterion.
+\end{enumerate}
 
@@
 This effect was estimated at about 10 with standard error of 3.9, and the high positive effect suggests senators tend to collaborate with other senators who author about the same number of bills they do.
+Finally, we found the same party transitive triplet effect was also significant, with a fitted value of 1.3 and standard error of 0.7.
+
+List of effects: 
+\begin{enumerate}
+\item jtt - party
+\item jtt - sex
+\item similiarity trans trip - bills
+\item same tt - party
+\end{enumerate}
 
 # Simulation Procedure
-
-1. Fit each model 1,000 times and save the parameter estimates, standard errors, and overall convergence criterion.
-2. (From the models that converged ?) Compute the mean of the 1,000 estimates of each parameter.
-3. Use the mean values from step 2 as the parameter values in the model and simulate 100 observations from the model. These simulated networks will become the "true data."
-4. To each simulation, refit the model form from which it was simulated. 
-5. Repeat steps 4-5 for various permutations of model forms and parameter values; e.g. double the parameter values, quadruple the parameter values, change the signs on the parameters, etc.
-6. Simulate additional "alternative data" from the model in step 3 and the fitted model in step 4. 
-7. Place the appropriate node-link diagram of each "true data" observation in a lineup with $M-1$ (M=12? M=9?) "alternative data" node-link diagrams. 
-8. Show the lineups to participants, with no participant seeing the same "true data" or "alternative data" plot twice.
+\begin{enumerate}
+\item Fit each model 1,000 times and save the parameter estimates, standard errors, and overall convergence criterion.
+\item (From the models that converged ?) Compute the mean of the 1,000 estimates of each parameter.
+\item Use the mean values from step 2 as the parameter values in the model and simulate 100 observations from the model. These simulated networks will become the "true data."
+\item To each simulation, refit the model form from which it was simulated. 
+\item Repeat steps 4-5 for various permutations of model forms and parameter values; e.g. double the parameter values, quadruple the parameter values, change the signs on the parameters, etc.
+\item Simulate additional "alternative data" from the model in step 3 and the fitted model in step 4. 
+\item Place the appropriate node-link diagram of each "true data" observation in a lineup with $M-1$ (M=12? M=9?) "alternative data" node-link diagrams. 
+\item Show the lineups to participants, with no participant seeing the same "true data" or "alternative data" plot twice. 
+\end{enumerate}
 
@@
 
-```{r M1}
+```{r M1, echo = FALSE, message=FALSE,warning=FALSE}
 M1senateEsts <- read_csv("../Data/senate/basicModelFitsSenate1000.csv")

### 2017-09-28 Code/Senate/combineSenatedataFits.R
add simulated data

@@ -82,3 +82,21 @@ alljttssen <- do.call(rbind, list(jttsSen1, jttsSen2, jttsSen3, jttsSen4,
 names(alljttssen)[1] <- "sim"
 alljttssen$sim <- 1:nrow(alljttssen)
 #write_csv(alljttssen, "Data/senate/jttsModelFitsSenateAll1000.csv")
+
+samepSen1 <- read_csv("Data/senate/samepModelFitsSenate100.csv")
+samepSen2 <- read_csv("Data/senate/samepModelFitsSenate200.csv")
+samepSen3 <- read_csv("Data/senate/samepModelFitsSenate300.csv")
+samepSen4 <- read_csv("Data/senate/samepModelFitsSenate400.csv")
+samepSen5 <- read_csv("Data/senate/samepModelFitsSenate500.csv")
+samepSen6 <- read_csv("Data/senate/samepModelFitsSenate600.csv")
+samepSen7 <- read_csv("Data/senate/samepModelFitsSenate700.csv")
+samepSen8 <- read_csv("Data/senate/samepModelFitsSenate800.csv")
+samepSen9 <- read_csv("Data/senate/samepModelFitsSenate900.csv")
+samepSen10 <- read_csv("Data/senate/samepModelFitsSenate1000.csv")
+
+head(samepSen5)
+
+allsamepSen <- do.call(rbind, list(samepSen1[,-1], samepSen2[,-1], samepSen3[,-1], samepSen4[,-1], samepSen5[,-1],
+                                   samepSen6[,-1], samepSen7[,-1], samepSen8[,-1], samepSen9[,-1], samepSen10))
+allsamepSen$sim <- 1:nrow(allsamepSen)
+#write_csv(allsamepSen, "Data/senate/samepModelFitsSenateall1000.csv")

### 2017-09-28 Code/Senate/scriptforServer.R

@@ -8,11 +8,11 @@ myalg <- sienaAlgorithmCreate( projname = Sys.time() , n3 = 1000)
 #Senjtt_p <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
 #Senjtt_s <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "sex", character = TRUE)
 # SeninIsD <- includeEffects(SenBasic, "inIsDegree", include = TRUE, type = "eval", interaction1 = "", character = TRUE)
-Sensame_p <- includeEffects(SenBasic, "sameX", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
-# Senstt_b <- includeEffects(SenBasic, "simXTransTrip", include = TRUE, type = "eval", interaction1 = "bills", character = TRUE)
+# Sensame_p <- includeEffects(SenBasic, "sameX", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+Senstt_b <- includeEffects(SenBasic, "simXTransTrip", include = TRUE, type = "eval", interaction1 = "bills", character = TRUE)
 # Sensimre_p <- includeEffects(SenBasic, "simRecipX", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
 # Sentt_p <- includeEffects(SenBasic, "transTrip", include = TRUE, type = "eval", interaction1 = "", character = TRUE)
-# Senstt_p <- includeEffects(SenBasic, "sameXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+Senstt_p <- includeEffects(SenBasic, "sameXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
 
 get_effects_dist <- function(dat, struct, N){
   require(RSiena)
@@ -27,33 +27,53 @@ get_effects_dist <- function(dat, struct, N){
                       "maxConv")
   for (i in 1:N){
     fits <- siena07(myalg, data = dat, effects = struct, returnDeps = FALSE,
-                    batch=TRUE, verbose = FALSE, silent = TRUE, useCluster = TRUE, nbrNodes = 3, 
[…]

### 2017-09-28 Code/Senate/simulatefrommodels.R
add code to simulate data

@@ -1,0 +1,100 @@
+# Simulating "data" from the models
+setwd("~/Desktop/Dissertation/NetworksVizInference/")
+library(RSiena)
+library(netvizinf)
+library(geomnet)
+
+load("Data/senate/senateSienaNoHRC.rda")
+# basic effects structure
+SenBasic <- getEffects(senateSiena)
+# basic algorithm structure
+myalg <- sienaAlgorithmCreate( projname = Sys.time() , n3 = 1000)
+# additional effects structures
+# jumpXTransTrip party
+Senjtt_p <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+# jumpXTransTrip sex
+Senjtt_s <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "sex", character = TRUE) 
+# indegree isolates
+# SeninIsD <- includeEffects(SenBasic, "inIsDegree", include = TRUE, type = "eval", interaction1 = "", character = TRUE)
+# sameX party
+Sensame_p <- includeEffects(SenBasic, "sameX", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+# simXTransTrip bills
+Senstt_b <- includeEffects(SenBasic, "simXTransTrip", include = TRUE, type = "eval", interaction1 = "bills", character = TRUE)
+# sim recip party
+#Sensimre_p <- includeEffects(SenBasic, "simRecipX", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+Sentt_p <- includeEffects(SenBasic, "transTrip", include = TRUE, type = "eval", interaction1 = "", character = TRUE)
+# Senstt_p <- includeEffects(SenBasic, "sameXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+
+# means of parameters to use as starting values
+M1senateEsts <- read_csv("Data/senate/basicModelFitsSenate1000.csv")
+#inisdSenateEsts <- read_csv("Data/senate/inisdModelFitsSenate1000.csv")
+jttpSenateEsts <- read_csv("Data/senate/jttpModelFitsSenateAll1000.csv")
[…]

### 2017-09-29 ASApaper/paper.Rmd
add simulated data

@@
 ```{r M1, echo = FALSE, message=FALSE,warning=FALSE}
-M1senateEsts <- read_csv("../Data/senate/basicModelFitsSenate1000.csv")
-inisdSenateEsts <- read_csv("../Data/senate/inisdModelFitsSenate1000.csv")
-jttpSenateEsts <- read_csv("../Data/senate/jttpModelFitsSenateAll1000.csv")
-jttsSenateEsts <- read_csv("../Data/senate/jttsModelFitsSenateAll1000.csv")
-# ggpairs(M1senateEsts, columns = 1:8)
-# M1senateEsts %>% mutate(converged = (maxConv <= 0.25)) %>% 
-#   ggplot() + 
-#   geom_density(aes(x = density, color = converged), adjust = .5) 
-# M1senateEsts %>% mutate(converged = (maxConv <= 0.25)) %>% 
-#   ggplot() + 
-#   geom_histogram(aes(x = density, color = converged), binwidth = .005, fill = NA) + facet_wrap(~converged)
-
-# make ests long - M1
-M1senateEsts %>% 
-  select(Rate:recip, maxConv, sim) %>% 
-  gather(parameter, estimate, Rate:recip) -> melt1 
-M1senateEsts %>% 
-  select(se_density:sim) %>% 
-  gather(parameter, se, se_density:se_recip) %>% 
-  mutate(parameter = str_replace(parameter, "se_", "")) -> melt2 
-left_join(melt1, melt2) %>% arrange(sim) -> M1SenEstsLong
-
-# make ests long - inisd
-inisdSenateEsts %>% 
-  select(Rate:inIsDegree, maxConv, sim) %>% 
-  gather(parameter, estimate, Rate:inIsDegree) -> melt1 
-inisdSenateEsts %>% 
-  select(se_density:sim) %>% 
-  gather(parameter, se, se_density:se_inIsDegree) %>% 
-  mutate(parameter = str_replace(parameter, "se_", "")) -> melt2 
-left_join(melt1, melt2) %>% arrange(sim) -> inisdSenEstsLong
-
-# make ests long - jttp
-jttpSenateEsts %>% 
-  select(Rate:jumpXTransTrip, maxConv, sim) %>% 
-  gather(parameter, estimate, Rate:jumpXTransTrip) -> melt1 
-jttpSenateEsts %>% 
-  select(se_density:maxConv, sim) %>% 
-  gather(parameter, se, se_density:se_jumpXTransTrip) %>% 
-  mutate(parameter = str_replace(parameter, "se_", "")) -> melt2 
-left_join(melt1, melt2) %>% arrange(sim) -> jttpSenLong
-
-# make ests long - jtts
-jttsSenateEsts %>% 
-  select(Rate:jumpXTransTrip, maxConv, sim) %>% 
-  gather(parameter, estimate, Rate:jumpXTransTrip) -> melt1 
-jttsSenateEsts %>% 
-  select(se_density:maxConv, sim) %>% 
-  gather(parameter, se, se_density:se_jumpXTransTrip) %>% 
-  mutate(parameter = str_replace(parameter, "se_", "")) -> melt2 
-left_join(melt1, melt2) %>% arrange(sim) -> jttsSenLong
 

### 2017-09-29 Code/Senate/combineSenatedataFits.R

@@ -100,3 +100,23 @@ allsamepSen <- do.call(rbind, list(samepSen1[,-1], samepSen2[,-1], samepSen3[,-1
                                    samepSen6[,-1], samepSen7[,-1], samepSen8[,-1], samepSen9[,-1], samepSen10))
 allsamepSen$sim <- 1:nrow(allsamepSen)
 #write_csv(allsamepSen, "Data/senate/samepModelFitsSenateall1000.csv")
+
+
+simttbSen1 <- read_csv("Data/senate/simttbModelFitsSenate100.csv")
+simttbSen2 <- read_csv("Data/senate/simttbModelFitsSenate200.csv")
+simttbSen3 <- read_csv("Data/senate/simttbModelFitsSenate300.csv")
+simttbSen4 <- read_csv("Data/senate/simttbModelFitsSenate400.csv")
+simttbSen5 <- read_csv("Data/senate/simttbModelFitsSenate500.csv")
+simttbSen6 <- read_csv("Data/senate/simttbModelFitsSenate600.csv")
+simttbSen7 <- read_csv("Data/senate/simttbModelFitsSenate700.csv")
+simttbSen8 <- read_csv("Data/senate/simttbModelFitsSenate800.csv")
+simttbSen9 <- read_csv("Data/senate/simttbModelFitsSenate900.csv")
+simttbSen10 <- read_csv("Data/senate/simttbModelFitsSenate1000.csv")
+
+allsimttbSen <- do.call(rbind, list(simttbSen1, simttbSen2, simttbSen3, simttbSen4, simttbSen5,
+                                   simttbSen6, simttbSen7, simttbSen8, simttbSen9, simttbSen10))
+names(allsimttbSen)[1] <- "sim"
+allsimttbSen$sim <- 1:nrow(allsimttbSen)
+
+head(allsimttbSen)
+#write_csv(allsimttbSen, "Data/senate/simttbModelFitsSenateAll1000.csv")

### 2017-09-29 Code/Senate/simulatefrommodels.R

@@ -31,6 +31,7 @@ M1senateEsts <- read_csv("Data/senate/basicModelFitsSenate1000.csv")
 jttpSenateEsts <- read_csv("Data/senate/jttpModelFitsSenateAll1000.csv")
 jttsSenateEsts <- read_csv("Data/senate/jttsModelFitsSenateAll1000.csv")
 samepSenateEsts <- read_csv("Data/senate/samepModelFitsSenateall1000.csv")
+simttbSenateEsts <- read_csv("Data/senate/simttbModelFitsSenateAll1000.csv")
 
 M1SenMeans <- M1senateEsts %>% filter(maxConv <=0.25) %>% summarise_at(.cols = 1:5, .funs = mean) %>% as.numeric()
 M1SenNconv <-  M1senateEsts %>% filter(maxConv <=0.25) %>% count() %>% as.numeric
@@ -40,6 +41,10 @@ jttsSenMeans <- jttsSenateEsts %>% filter(maxConv <=0.25) %>% summarise_at(.cols
 jttsSenNconv <- jttsSenateEsts %>% filter(maxConv <=0.25) %>% count() %>% as.numeric
 samepSenMeans <- samepSenateEsts %>% filter(maxConv <=0.25) %>% summarise_at(.cols = 2:7, .funs = mean) %>% as.numeric()
 samePSenNconv <- samepSenateEsts %>% filter(maxConv <=0.25) %>% count() %>% as.numeric
+simttbSenMeans <- simttbSenateEsts %>% filter(maxConv <= 0.25) %>% summarise_at(.cols = 2:7, .funs = mean) %>% as.numeric()
+simttbSenNconv <- simttbSenateEsts %>% filter(maxConv <=0.25) %>% count() %>% as.numeric
+
+
 
 ######################################################################################################
 # Dont need to do again!  
@@ -50,6 +55,8 @@ samePSenNconv <- samepSenateEsts %>% filter(maxConv <=0.25) %>% count() %>% as.n
 #jttpdat <- saom_simulate(dat = senateSiena, struct = Senjtt_p, parms = jttpSenMeans, N = 1000)
 #jttsdat <- saom_simulate(dat = senateSiena, struct = Senjtt_s, parms = jttsSenMeans, N = 1000)
 #samepdat <- saom_simulate(dat = senateSiena, struct = Sensame_p, parms = samepSenMeans, N=1000)
[…]

### 2017-09-29 Code/Senate/combineSenatedataFits.R
add more estimates

@@ -120,3 +120,23 @@ allsimttbSen$sim <- 1:nrow(allsimttbSen)
 
 head(allsimttbSen)
 #write_csv(allsimttbSen, "Data/senate/simttbModelFitsSenateAll1000.csv")
+
+
+samettpSen1 <- read_csv("Data/senate/samettpModelFitsSenate100.csv")
+samettpSen2 <- read_csv("Data/senate/samettpModelFitsSenate200.csv")
+samettpSen3 <- read_csv("Data/senate/samettpModelFitsSenate300.csv")
+samettpSen4 <- read_csv("Data/senate/samettpModelFitsSenate400.csv")
+samettpSen5 <- read_csv("Data/senate/samettpModelFitsSenate500.csv")
+samettpSen6 <- read_csv("Data/senate/samettpModelFitsSenate600.csv")
+samettpSen7 <- read_csv("Data/senate/samettpModelFitsSenate700.csv")
+samettpSen8 <- read_csv("Data/senate/samettpModelFitsSenate800.csv")
+samettpSen9 <- read_csv("Data/senate/samettpModelFitsSenate900.csv")
+samettpSen10 <- read_csv("Data/senate/samettpModelFitsSenate1000.csv")
+
+allsamettpSen <- do.call(rbind, list(samettpSen1, samettpSen2, samettpSen3, samettpSen4, samettpSen5,
+                                    samettpSen6, samettpSen7, samettpSen8, samettpSen9, samettpSen10))
+names(allsamettpSen)[1] <- "sim"
+allsamettpSen$sim <- 1:nrow(allsamettpSen)
+
+#write_csv(allsamettpSen, "Data/senate/samettpModelFitsSenateAll1000.csv")
+

### 2017-09-29 Code/Senate/exploring_model_fits.R

@@ -1,0 +1,129 @@
+# Exploratory Data analyses
+library(tidyverse)
+M1senateEsts <- read_csv("../Data/senate/basicModelFitsSenate1000.csv")
+#inisdSenateEsts <- read_csv("../Data/senate/inisdModelFitsSenate1000.csv")
+jttpSenateEsts <- read_csv("../Data/senate/jttpModelFitsSenateAll1000.csv")
+jttsSenateEsts <- read_csv("../Data/senate/jttsModelFitsSenateAll1000.csv")
+samepSenateEsts <- read_csv("../Data/senate/samepModelFitsSenateall1000.csv")
+simttbSenateEsts <- read_csv("../Data/senate/simttbModelFitsSenateAll1000.csv")
+samettpSenateEsts <- read_csv("../Data/senate/samettpModelFitsSenateAll1000.csv")
+
+# ggpairs(M1senateEsts, columns = 1:8)
+# M1senateEsts %>% mutate(converged = (maxConv <= 0.25)) %>% 
+#   ggplot() + 
+#   geom_density(aes(x = density, color = converged), adjust = .5) 
+# M1senateEsts %>% mutate(converged = (maxConv <= 0.25)) %>% 
+#   ggplot() + 
+#   geom_histogram(aes(x = density, color = converged), binwidth = .005, fill = NA) + facet_wrap(~converged)
+
+# make ests long - M1
+M1senateEsts %>% 
+  select(Rate:recip, maxConv, sim) %>% 
+  gather(parameter, estimate, Rate:recip) -> melt1 
+M1senateEsts %>% 
+  select(se_density:sim) %>% 
+  gather(parameter, se, se_density:se_recip) %>% 
+  mutate(parameter = str_replace(parameter, "se_", "")) -> melt2 
+left_join(melt1, melt2) %>% arrange(sim) -> M1SenEstsLong 
+
+# make ests long - inisd
+# inisdSenateEsts %>% 
+#   select(Rate:inIsDegree, maxConv, sim) %>% 
+#   gather(parameter, estimate, Rate:inIsDegree) -> melt1 
+# inisdSenateEsts %>% 
+#   select(se_density:sim) %>% 
+#   gather(parameter, se, se_density:se_inIsDegree) %>% 
+#   mutate(parameter = str_replace(parameter, "se_", "")) -> melt2 
+# left_join(melt1, melt2) %>% arrange(sim) -> inisdSenEstsLong
+
[…]

### 2017-10-02 Code/Senate/exploring_model_fits.R
haven't committed in a while.

@@ -127,3 +127,5 @@ ggplot(data = allModelsLong %>% filter(!str_detect(parameter, "Rate") & maxConv 
   geom_point(aes(x = se, y = estimate, color = parameter)) + 
   facet_wrap(~model, scales = 'free')
 
+
+

### 2017-10-02 Code/Senate/fitModelsFauxData.R

@@ -1,0 +1,66 @@
+# fit every model to the simulated data
+
+load("Data/senate/senateSienaNoHRC.rda")
+
+SenBasic <- getEffects(senateSiena)
+Senjtt_p <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+Senjtt_s <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "sex", character = TRUE)
+SeninIsD <- includeEffects(SenBasic, "inIsDegree", include = TRUE, type = "eval", interaction1 = "", character = TRUE)
+Sensame_p <- includeEffects(SenBasic, "sameX", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+Senstt_b <- includeEffects(SenBasic, "simXTransTrip", include = TRUE, type = "eval", interaction1 = "bills", character = TRUE)
+Sensimre_p <- includeEffects(SenBasic, "simRecipX", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+Sentt_p <- includeEffects(SenBasic, "transTrip", include = TRUE, type = "eval", interaction1 = "", character = TRUE)
+Senstt_p <- includeEffects(SenBasic, "sameXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+
+# fitBasicSaom <- function(sienaDat){
+#   SenBasic <- getEffects(sienaDat)
+#   myalg <- sienaAlgorithmCreate( projname = Sys.time() , n3 = 1000)
+#   fits <- siena07(myalg, data = sienaDat, effects = SenBasic, returnDeps = TRUE,
+#                   batch=TRUE, verbose = FALSE, silent = TRUE)
+#   return(fits)
+# }
+# 
+# fitjttpSaom <- function(sienaDat){
+#   SenBasic <- getEffects(sienaDat)
+#   myalg <- sienaAlgorithmCreate( projname = Sys.time() , n3 = 1000)
+#   fits <- siena07(myalg, data = sienaDat, effects = Senjtt_p, returnDeps = TRUE,
+#                   batch=TRUE, verbose = FALSE, silent = TRUE)
+#   return(fits)
+# }
+# 
+# fitjttsSaom <- function(sienaDat){
[…]

### 2017-10-02 Code/Senate/otherEststoSimFrom.R

@@ -1,0 +1,99 @@
+# simulate from different parameter values 
+
+M1SenMeans <- M1senateEsts %>% filter(maxConv <=0.25) %>% summarise_at(.cols = 1:5, .funs = mean) %>% as.numeric()
+M1SenNconv <-  M1senateEsts %>% filter(maxConv <=0.25) %>% count() %>% as.numeric
+jttpSenMeans <- jttpSenateEsts %>% filter(maxConv <=0.25) %>% summarise_at(.cols = 2:7, .funs = mean) %>% as.numeric()
+jttpSenNconv <- jttpSenateEsts %>% filter(maxConv <=0.25) %>% count() %>% as.numeric
+jttsSenMeans <- jttsSenateEsts %>% filter(maxConv <=0.25) %>% summarise_at(.cols = 2:7, .funs = mean) %>% as.numeric()
+jttsSenNconv <- jttsSenateEsts %>% filter(maxConv <=0.25) %>% count() %>% as.numeric
+samepSenMeans <- samepSenateEsts %>% filter(maxConv <=0.25) %>% summarise_at(.cols = 2:7, .funs = mean) %>% as.numeric()
+samePSenNconv <- samepSenateEsts %>% filter(maxConv <=0.25) %>% count() %>% as.numeric
+simttbSenMeans <- simttbSenateEsts %>% filter(maxConv <= 0.25) %>% summarise_at(.cols = 2:7, .funs = mean) %>% as.numeric()
+simttbSenNconv <- simttbSenateEsts %>% filter(maxConv <=0.25) %>% count() %>% as.numeric
+samettpSenMeans <- samettpSenateEsts %>% filter(maxConv <= 0.25) %>% summarise_at(.cols = 2:7, .funs = mean) %>% as.numeric()
+samettpSenNconv <- samettpSenateEsts %>% filter(maxConv <=0.25) %>% count() %>% as.numeric
+
+
+basicParms <- (M1SenEstsLong %>% select(parameter) %>% unique())[[1]] 
+jttpParms <- (jttpSenLong %>% select(parameter) %>% unique())[[1]] 
+jttsParms <- (jttsSenLong %>% select(parameter) %>% unique())[[1]] 
+samepParms <- (samepSenLong %>% select(parameter) %>% unique())[[1]] 
+simttbParms <- (simttbSenLong %>% select(parameter) %>% unique())[[1]] 
+samettpParms <- (samettpSenLong %>% select(parameter) %>% unique())[[1]] 
+
[…]

### 2017-10-02 Code/Senate/senate_siena_nohrc.R

@@ -45,6 +45,10 @@ se114adj <- make_se_adjmat(se_no = 114, obamaSenates = senateNet)
 
 remove_sens <- which(!(colnames(se111adj) %in% senators))
 
+allSensAMempty <- startingsemat[-remove_sens, -remove_sens]
+save(allSensAMempty, file = "Data/senate/allSensAdjMatEmpty.RDS")
+
+
 se111adj <- se111adj[-remove_sens, -remove_sens]
 se112adj <- se112adj[-remove_sens, -remove_sens]
 se113adj <- se113adj[-remove_sens, -remove_sens]

### 2017-10-02 Code/Senate/simstoSienaData.R

@@ -1,0 +1,81 @@
+# write function to take simulated data and turn it into siena data. 
+library(tidyverse)
+library(RSiena)
+load("Data/senate/se111adjmat.RDS")
+load("Data/senate/allSensAdjMatEmpty.RDS")
+senateCovars <- read_csv('Data/senate/senateCovars.csv')
+
+# turn a wave df into an adjmat 
+make_se_adjmat2 <- function(adjmat=allSensAMempty, wave){
+  N <- nrow(wave)
+  for(i in 1:N){
+    if(!(is.na(wave[i,c("to")]))){
+    k <- which(rownames(adjmat) == wave$from[i])
+    l <- which(colnames(adjmat) == wave$to[i])
+    adjmat[k, l] <- 1
+    }
+  }
+  return(adjmat)
+}
+# turn 3 waves into a 4-wave sienaData object
+dat_to_SienaDat <- function(wave1 = se111adj, Sim, covars = senateCovars){
+  wave2 <- filter(Sim, wave == 1)
+  wave2$from <- replace(wave2$from, 1:length(wave2$from), covars$source[wave2$from])
+  wave2$to <- replace(wave2$to, 1:length(wave2$to), covars$source[wave2$to])
+  wave2 <- make_se_adjmat2(wave = wave2)
+  wave3 <- filter(Sim, wave == 2)
+  wave3$from <- replace(wave3$from, 1:length(wave3$from), covars$source[wave3$from])
+  wave3$to <- replace(wave3$to, 1:length(wave3$to), covars$source[wave3$to])
+  wave3 <- make_se_adjmat2(wave = wave3)
+  wave4 <- filter(Sim, wave == 3)
+  wave4$from <- replace(wave4$from, 1:length(wave4$from), covars$source[wave4$from])
+  wave4$to <- replace(wave4$to, 1:length(wave4$to), covars$source[wave4$to])
+  wave4 <- make_se_adjmat2(wave = wave4)
+  
+  senateData <- array( c( wave1, wave2, wave3, wave4 ),
+                       dim = c( 155, 155, 4 ) )
+  senateData2 <- sienaDependent(senateData)
+  party <- coCovar(val = as.numeric(as.factor(covars$party)))
+  # party 1 is dem 2 is ind 3 is rep
+  sex <- coCovar(val = as.numeric(as.factor(covars$party)))
+  # sex 1 is F 2 is M
+  nbills_au <- senateCovars[,4:7]
[…]

### 2017-10-02 Code/Senate/simulatefrommodels.R

@@ -23,7 +23,7 @@ Senstt_b <- includeEffects(SenBasic, "simXTransTrip", include = TRUE, type = "ev
 # sim recip party
 #Sensimre_p <- includeEffects(SenBasic, "simRecipX", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
 Sentt_p <- includeEffects(SenBasic, "transTrip", include = TRUE, type = "eval", interaction1 = "", character = TRUE)
-# Senstt_p <- includeEffects(SenBasic, "sameXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+Senstt_p <- includeEffects(SenBasic, "sameXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
 
 # means of parameters to use as starting values
 M1senateEsts <- read_csv("Data/senate/basicModelFitsSenate1000.csv")
@@ -33,6 +33,7 @@ jttsSenateEsts <- read_csv("Data/senate/jttsModelFitsSenateAll1000.csv")
 samepSenateEsts <- read_csv("Data/senate/samepModelFitsSenateall1000.csv")
 simttbSenateEsts <- read_csv("Data/senate/simttbModelFitsSenateAll1000.csv")
 
+# get means 
 M1SenMeans <- M1senateEsts %>% filter(maxConv <=0.25) %>% summarise_at(.cols = 1:5, .funs = mean) %>% as.numeric()
 M1SenNconv <-  M1senateEsts %>% filter(maxConv <=0.25) %>% count() %>% as.numeric
 jttpSenMeans <- jttpSenateEsts %>% filter(maxConv <=0.25) %>% summarise_at(.cols = 2:7, .funs = mean) %>% as.numeric()
@@ -43,36 +44,76 @@ samepSenMeans <- samepSenateEsts %>% filter(maxConv <=0.25) %>% summarise_at(.co
 samePSenNconv <- samepSenateEsts %>% filter(maxConv <=0.25) %>% count() %>% as.numeric
 simttbSenMeans <- simttbSenateEsts %>% filter(maxConv <= 0.25) %>% summarise_at(.cols = 2:7, .funs = mean) %>% as.numeric()
 simttbSenNconv <- simttbSenateEsts %>% filter(maxConv <=0.25) %>% count() %>% as.numeric
[…]

### 2017-10-02 Code/Senate/exploring_model_fits.R
sims for lus

@@ -1,12 +1,12 @@
 # Exploratory Data analyses
 library(tidyverse)
-M1senateEsts <- read_csv("../Data/senate/basicModelFitsSenate1000.csv")
+M1senateEsts <- read_csv("Data/senate/basicModelFitsSenate1000.csv")
 #inisdSenateEsts <- read_csv("../Data/senate/inisdModelFitsSenate1000.csv")
-jttpSenateEsts <- read_csv("../Data/senate/jttpModelFitsSenateAll1000.csv")
-jttsSenateEsts <- read_csv("../Data/senate/jttsModelFitsSenateAll1000.csv")
-samepSenateEsts <- read_csv("../Data/senate/samepModelFitsSenateall1000.csv")
-simttbSenateEsts <- read_csv("../Data/senate/simttbModelFitsSenateAll1000.csv")
-samettpSenateEsts <- read_csv("../Data/senate/samettpModelFitsSenateAll1000.csv")
+jttpSenateEsts <- read_csv("Data/senate/jttpModelFitsSenateAll1000.csv")
+jttsSenateEsts <- read_csv("Data/senate/jttsModelFitsSenateAll1000.csv")
+samepSenateEsts <- read_csv("Data/senate/samepModelFitsSenateall1000.csv")
+simttbSenateEsts <- read_csv("Data/senate/simttbModelFitsSenateAll1000.csv")
+samettpSenateEsts <- read_csv("Data/senate/samettpModelFitsSenateAll1000.csv")
 
 # ggpairs(M1senateEsts, columns = 1:8)
 # M1senateEsts %>% mutate(converged = (maxConv <= 0.25)) %>% 

### 2017-10-02 Code/Senate/fitModelsFauxData.R

@@ -1,5 +1,6 @@
 # fit every model to the simulated data
-
+library(RSiena)
+library(tidyverse)
 load("Data/senate/senateSienaNoHRC.rda")
 
 SenBasic <- getEffects(senateSiena)

### 2017-10-02 Code/Senate/otherEststoSimFrom.R

@@ -95,5 +95,5 @@ simstoget$N <- (20-1)*10
 gotsims <- simstoget %>%   
   mutate(sims = pmap(.f = saom_simulate2, 
                      .l = list(dat = dat, struct = structs, parms = inits, N = N)))
-
+save(gotsims, file = "Data/senate/FauxTrueData/M_1_sims_for_fdata.RDS")
 

### 2017-10-09 shinyApp/server.R
add radio buttons, produce singletons.

@@ -31,7 +31,8 @@ saom_simulate2 <- function(dat, struct, parms, N) {
   return(getsims$sims)
 }
 # turn simulation lists into dfs 
-sims_to_df <- function(sims) {
+sims_to_df <- function(sims, dat) {
+  Nact <- length(dat$nodeSets$Actors)
   N <- length(sims)
   waves <- length(sims[[1]][[1]][[1]])
   simsdf <- NULL
@@ -40,7 +41,7 @@ sims_to_df <- function(sims) {
     for (j in 1:waves) {
       dat <- as.data.frame(sims[[i]][[1]][[1]][[j]])
       names(dat) <- c("from", "to", "dep.var.id")
-      ids <- unique(c(unique(dat$from), unique(dat$to)))
+      ids <- 1:Nact
       nodes <- data.frame(id = ids)
       dat2 <- merge(dat, nodes, by.x = "from", by.y = "id", 
                     all = T)
@@ -50,7 +51,7 @@ sims_to_df <- function(sims) {
       counter <- counter + 1
     }
   }
-  mydf <- plyr::rbind.fill(simsdf)
+  mydf <- dplyr::bind_rows(simsdf)
   return(mydf)
 }
 
@@ -66,13 +67,29 @@ server <- function(input, output) {
   dat_for_lu <- reactive({
     # get a null data set
     nulldat <- filter(alldat, model == input$model)
+    # needed to add singletons in later
+    ids <- 1:length(senateSiena$nodeSets$Actors)
+    #set seed 
     set.seed(input$seed)
-    if (input$reverse){
+    # get the "data"
+    if (input$whichModel == "alt"){
       whichdat <- sample(max(nulldat$sim), input$M-1)
       nulldat <- filter(nulldat, sim %in% whichdat, wave == input$wave)
+      nodes <- data.frame(id = rep(ids, input$M-1), 
+                          sim = rep(unique(nulldat$sim), each = max(ids)),
+                          wave = 1, model = unique(nulldat$model))
+      nulldat <- merge(nulldat, nodes, by.x = c("from", "wave", "sim", "model"), 
+                       by.y = c("id", "wave", "sim", "model"), all = T)
+      nulldat <- select(nulldat, from, to, dep.var.id, wave, sim, model) %>% arrange(sim, from, to)
     } else {
       whichdat <- sample(max(nulldat$sim), 1)
       nulldat <- filter(nulldat, sim == whichdat, wave == input$wave) %>% mutate(sim = 1001)
[…]

### 2017-10-09 shinyApp/ui.R

@@ -12,11 +12,17 @@ ui <- fluidPage(
     
     # Sidebar panel for inputs ----
     sidebarPanel(
-      
+      # Input: which model are you designating? The null (M-1 plots) or the alternate (1 plot)
+      radioButtons(inputId = "whichModel", 
+                    label = 'Which model are you designating?', 
+                    choices = list("Null (simulate M-1 new plots)" = "null",
+                                   "Alternative (simulate 1 new plot)" = "alt"),
+                    selected = "null"),
       # Input: pick a model
       selectInput(inputId = "model",
                   label = "Select a model:",
-                  choices = c("basic", "jttp", "jtts", "samep", "samettp", "simttb"),
+                  choices = list("basic" = "basic", "jttp (-3.45)" = "jttp", "jtts (3.34)"= "jtts", 
+                              "samep (.20)" = "samep", "samettp (1.33)" = "samettp", "simttb (10.09)" = "simttb"),
                   selected = "basic"),
       # Input: pick a wave 
       numericInput(inputId = "wave",
@@ -26,7 +32,8 @@ ui <- fluidPage(
       selectInput(inputId = "basicParm", 
                   label = "Select a parameter (density, reciprocity, or both)\n
                   for basic model multiplier. Select none for all other models:",
-                  choices = c("density", "reciprocity", "both", "none"), 
+                  choices = list("density (-4.90)"= "density", "reciprocity (4.89)" = "reciprocity", 
+                                 "both", "none"), 
                   selected = "none"),
       # Input: pick a lineup size
       numericInput(inputId = "M",
@@ -40,10 +47,6 @@ ui <- fluidPage(
       numericInput(inputId = "seed",
                    label = "Set a random seed:", 
                    value = 123456, min = 10000, max = 999999, step = 1),
-      # Input: "reverse" the lineup or not? Check box to reverse
-      checkboxInput(inputId = "reverse", 
-                    label = 'Check to "reverse" lineup (1 multiplied plot, M-1 non-multiplied)', 
-                    value = FALSE),
       # Submit button 
       submitButton(text = "Generate lineup!", icon = icon("connectdevelop"))
     ),

### 2017-10-09 ASApaper/paper.Rmd
haven't committed in a while.
couple of intro blurbs: SAOMs and visual inference

@@
 ---
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(cache=TRUE, echo = FALSE, message=FALSE,warning=FALSE,
+                      out.width = '\\linewidth')
+```
 
-```{r pkgs, echo = FALSE, message=FALSE,warning=FALSE}
+```{r pkgs}
 library(tidyverse)
@@
 library(GGally)
+load("../Data/senate/senateSienaNoHRC.rda")
+
 ```
@@
 
-Talk about goodness-of-fit
+<!--Talk about SAOMs-->
+XX cleverly worded intro sentences about social networks and ways to model them XX...
+One such model, introduced by @saompaper, is the *stochastic actor-oriented model* (SAOM).
+This model is unique among network modeling techniques because it (1) incorporates node-level information into the process in addition the usual network structural information and (2) incorporates two or more observations of a network to capture how relationships between nodes change in time.
+These unique properties allow for the fitting of some very complicated models to inherently complex data, so it can be exceedingly difficult to interpret parameters and their corresponding estimates.
+The sheer amount of possible parameters to include in the model combined with the difficulty of interpretation make model selection and goodness-of-fit testing burdensome as well.
 
-Talk about political networks
+<!--Talk about visual inference-->
+We propose to attack some of the aforementioned difficulties with SAOMs by using a technique known as *visual inference*.
+This technique was created by @Bujaetal to provide well-defined, statistical rigor to the usual exploratory data analyses and model diagnostics that are typically performed by visualizing the data as opposed to looking at it raw or gathering numerical summaries of it.
+
+<!--Talk about political networks-->
+
+<!--Talk about goodness-of-fit-->
+
+<!--Talk about significance testings-->
 
@@
 This measure ranges in value from 0 to 1, where $WPC_{ij} = 1$ if senator $i$ is a cosponsor on every one of senator $j$'s bills and $WPC_{ij} = 0$ if senator $i$ is never a cosponsor any of senator $j$'s bills.
+
+```{r senateAll, fig.cap="The four senate collaboration networks that we estimate "}
+seobama <- read_csv("../Data/senate/senateobamapres_gsw_25.csv")
+to_label <- c("Joseph R. Biden Jr.", "John S. McCain", "Ted Cruz","Marco Rubio",
+  "Lindsey O. Graham","Rand Paul", "Bernard Sanders", "Jim Webb", "Mitch McConnell",
+  "Harry M. Reid", "Hillary Rodham Clinton", "Amy Jean Klobuchar")
+seobama$label <- as.factor(ifelse(seobama$source %in% to_label, seobama$source, ""))
+levels(seobama$label) <- c("", "Amy Klobuchar", "Bernie Sanders", "Harry Reid",
+                           "Hillary Clinton", "Jim Webb", "John McCain", "Joe Biden",
+                           "Lindsey Graham", "Marco Rubio", "Mitch McConnell", 
+                           "Rand Paul", "Ted Cruz")
+se111clint <- filter(seobama, senate == 111)
+se111noclint <- filter(seobama, senate == 111)
+se111noclint$target[which(se111noclint$target == "Hillary Rodham Clinton")] <- NA
+seobama2 <- seobama %>% filter(senate != 111) %>% bind_rows(se111noclint)
+
+set.seed(56049382)
+ggplot(data = seobama2) + 
+  geom_net(directed = T, labelon=T, arrowsize = .25, singletons= T, fiteach = T, linewidth = .25, layout.alg = 'fruchtermanreingold', fontsize = 3, repel=T, 
+           aes(from_id = source, to_id = target, color = party, label = label),
+           size = 1) + 
+  theme_net() + 
+  scale_color_manual(values = c("royalblue", "forestgreen","firebrick")) + 
+  theme(legend.position = 'bottom', strip.text = element_text(size = 14)) + 
+  facet_wrap(~senate, nrow = 2, labeller = "label_both") + 
+  xlim(c(0,1.05)) + 
+  ylim(c(0,1.05))
+```
+```{r senateClinton}
+se111clint$Clinton <- 'Yes'
+se111noclint$Clinton <- 'No'
+clintonSenate <- rbind(se111clint, se111noclint)
+clintonSenate$Clinton <- as.factor(clintonSenate$Clinton)
+clintonSenate$Clinton <- ordered(clintonSenate$Clinton, levels = c("Yes", "No"))
+clintonSenate %>% filter(!(source %in% c("Roland Burris", "Bernard Sanders", "Frank R. Lautenberg", "Judd A. Gregg", "Mary L. Landrieu") & is.na(target))) %>%
+ggplot() + 
+  geom_net(directed = T, labelon=F, arrowsize = .3, singletons= F, fiteach = T, arrowgap = .01, layout.alg = 'fruchtermanreingold',
+           aes(from_id = source, to_id = target, color = party, linewidth = gsw, shape = sex,size = n_au)) + 
+  scale_size_continuous(name="# bills\nauthored") +
+  theme_net() + 
+  scale_shape_manual(values = c(17,16)) + 
+  scale_color_manual(values = c("royalblue", "forestgreen","firebrick")) + 
+  theme(legend.position = 'bottom') + 
+  facet_wrap(~Clinton, nrow = 1, labeller = 'label_both')
+```
 
@@
 \item Fit each model 1,000 times and save the parameter estimates, standard errors, and overall convergence criterion.
-\item (From the models that converged ?) Compute the mean of the 1,000 estimates of each parameter.
+\item From the models that converged, compute the mean of the 1,000 estimates of each parameter.
 \item Use the mean values from step 2 as the parameter values in the model and simulate 100 observations from the model. These simulated networks will become the "true data."
@@
 
-# Exploratory plots
+# Experiment Set-Up
 
-```{r M1, echo = FALSE, message=FALSE,warning=FALSE}
+We want to determine at which point an effect becomes noticeable in a SAOM.
+By *noticeable*, we mean that the inclusion of the effect alters the appearance of networks simulated from a model.
+An effect becomes noticeable when as simulation from a model with a noticeable effect appears different from simulations from a model where the same affect is absent.
+Because the traditional way to visualize a network is a node-link diagram, we will examine the affect of different parameter values using node-link visualizations.
 
-```
+# Results
+
+# Conclusions & Future Work
 

### 2017-10-09 Code/Senate/scriptforServer.R

@@ -9,7 +9,7 @@ myalg <- sienaAlgorithmCreate( projname = Sys.time() , n3 = 1000)
 #Senjtt_s <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "sex", character = TRUE)
 # SeninIsD <- includeEffects(SenBasic, "inIsDegree", include = TRUE, type = "eval", interaction1 = "", character = TRUE)
 # Sensame_p <- includeEffects(SenBasic, "sameX", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
-Senstt_b <- includeEffects(SenBasic, "simXTransTrip", include = TRUE, type = "eval", interaction1 = "bills", character = TRUE)
+# Senstt_b <- includeEffects(SenBasic, "simXTransTrip", include = TRUE, type = "eval", interaction1 = "bills", character = TRUE)
 # Sensimre_p <- includeEffects(SenBasic, "simRecipX", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
 # Sentt_p <- includeEffects(SenBasic, "transTrip", include = TRUE, type = "eval", interaction1 = "", character = TRUE)
 Senstt_p <- includeEffects(SenBasic, "sameXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
@@ -27,32 +27,12 @@ get_effects_dist <- function(dat, struct, N){
                       "maxConv")
   for (i in 1:N){
     fits <- siena07(myalg, data = dat, effects = struct, returnDeps = FALSE,
-                    batch=TRUE, verbose = FALSE, silent = TRUE)
+                    batch=TRUE, verbose = FALSE, silent = TRUE, useCluster = TRUE, nbrNodes = 3, 
+                    initC = TRUE, clusterIter = FALSE)
     df.ests[i,] <- c(fits$rate, fits$theta, sqrt(diag(fits$covtheta)), fits$tconv.max)
   }
   return(df.ests)
 }
-
-simttbFitsSenate <- get_effects_dist(dat = senateSiena, struct = Senstt_b, N = 100)
-write.csv(simttbFitsSenate, "simttbModelFitsSenate100.csv")
-simttbFitsSenate <- get_effects_dist(dat = senateSiena, struct = Senstt_b, N = 100)
[…]

### 2017-10-09 ASApaper/paper.Rmd
SAOMs + visinf

@@
 This technique was created by @Bujaetal to provide well-defined, statistical rigor to the usual exploratory data analyses and model diagnostics that are typically performed by visualizing the data as opposed to looking at it raw or gathering numerical summaries of it.
+In visual inference there are *null plots* and *data plots*: the null plots are visualizations of data simulated from the model according to the null hypothesis, while the data plots are visualizations of data simulated from the model according to an alternative hypothesis.
+These two data sources are visualized side-by-side using small multiples using what Buja et al called the *lineup protocol*.
+The idea behind the lineup protocol is the police lineup, where witnesses to crimes are brought to the police station to observe a group of people, one of whom is suspected to have committed the crime, and are asked to identify the perpetrator of the crime.
+If the witness identifies the suspect in the lineup, that is taken as evidence the suspect is guilty, whereas if the witness does not identify the suspect, that is taken as evidence the suspect is not guilty.
+In the visual inference lineup protocol, the *suspect* is the data plot among $M-1$ null plots.
+If the "witness" can pick the data plot out of the lineup, that is taken as evidence that the null model should be rejected, whereas failing to pick out the data plot is taken as evidence that the null model should not be rejected.
+In our application of the lineup protocol to SAOMs, we include various effects of varying sizes in the SAOMs we fit to some data and try to *see* those effects in node-link diagram corresponding to simulations from the various models.
 

### 2017-10-11 shinyApp/server.R
add color, change so that the "nulls" are being simulated too when alt is selected

@@ -3,6 +3,7 @@ library(shiny)
 library(tidyverse)
 library(RSiena)
 library(geomnet)
+library(igraph)
 # senate siena data
 load('dat/senateSienaNoHRC.rda')
 # all data for the "data plot"
@@ -75,11 +76,19 @@ server <- function(input, output) {
     set.seed(input$seed)
     # get the "data"
     if (input$whichModel == "alt"){
-      whichdat <- sample(max(nulldat$sim), input$M-1)
-      nulldat <- filter(nulldat, sim %in% whichdat, wave == input$wave)
+      # whichdat <- sample(max(nulldat$sim), input$M-1)
+      # nulldat <- filter(nulldat, sim %in% whichdat, wave == input$wave)
+      # simulate from the basic model 
+      sienaSim <- saom_simulate2(dat = senateSiena,
+                                 struct = SenBasic,
+                                 parms = modelMeanEsts$ests[modelMeanEsts$model == "basic"][[1]],
+                                 N = input$M - 1)
+      nulldat <- sims_to_df(sienaSim, senateSiena)
+      nulldat <- nulldat %>% filter(wave == input$wave)
+      nulldat$model <- "basic"
       nodes <- data.frame(id = rep(ids, input$M-1), 
                           sim = rep(unique(nulldat$sim), each = max(ids)),
-                          wave = 1, model = unique(nulldat$model))
+                          wave = 1, model = "basic")
       nulldat <- merge(nulldat, nodes, by.x = c("from", "wave", "sim", "model"), 
                        by.y = c("id", "wave", "sim", "model"), all = T)
       nulldat <- select(nulldat, from, to, dep.var.id, wave, sim, model) %>% arrange(sim, from, to)
@@ -149,7 +158,29 @@ server <- function(input, output) {
     dat_for_lu <- bind_rows(altdat, nulldat) %>%  
       mutate(from = paste0("V", from), to = ifelse(is.na(to), to, paste0("V", to))) 
     dat_for_lu$ord <- rep(sample(input$M), as.numeric(table(dat_for_lu$sim)))
-    dat_for_lu
+    
+    if (input$color){
+      df <- data.frame(expand.grid(
+        from=paste0("V", 1:155), to = paste0("V", 1:155)), stringsAsFactors = FALSE)
+      netnest <- dat_for_lu %>% group_by(sim) %>% nest()
+      netnest <- netnest %>% mutate(
+        data = data %>% purrr::map(.f = function(d) {
[…]

### 2017-10-11 shinyApp/ui.R

@@ -47,6 +47,10 @@ ui <- fluidPage(
       numericInput(inputId = "seed",
                    label = "Set a random seed (10,000-999,999):", 
                    value = 123456, min = 10000, max = 999999, step = 1),
+      # Input: check a box to color the graphs
+      checkboxInput(inputId = "color", 
+                    label = "Check box to color clusters.",
+                    value = FALSE),
       # Submit button 
       submitButton(text = "Generate lineup!", icon = icon("connectdevelop"))
     ),

### 2017-10-11 shinyApp/server.R
add ability to change layout

@@ -232,7 +232,7 @@ server <- function(input, output) {
       ggplot(data = dat, aes(from_id=from, to_id=to, colour=log(csize))) + 
         geom_net(arrow = arrow(type = 'open', length = unit(2, "points") ), 
                  linewidth = .25, singletons = T, fiteach = T, directed = T, 
-                 arrowgap = .015, arrowsize = .3, size = 1) + 
+                 arrowgap = .015, arrowsize = .3, size = 1, layout.alg = input$layoutalg) + 
         theme_net() +
         theme(panel.background = element_rect(color = 'black')) +
         scale_colour_gradient(low="#f768a1", high = "#49006a") + facet_wrap(~ord) +
@@ -243,7 +243,7 @@ server <- function(input, output) {
         geom_net(aes(from_id = from, to_id = to), 
                  arrow = arrow(type = 'open', length = unit(2, "points") ), 
                  linewidth = .25, singletons = T, fiteach = T, directed = T, 
-                 color = 'black', arrowgap = .015, arrowsize = .3, size = 1) + 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size = 1, layout.alg = input$layoutalg) + 
         theme_net() + 
         theme(panel.background = element_rect(color = 'black')) +
         facet_wrap(~ord)

### 2017-10-11 shinyApp/ui.R

@@ -14,15 +14,15 @@ ui <- fluidPage(
     sidebarPanel(
       # Input: which model are you designating? The null (M-1 plots) or the alternate (1 plot)
       radioButtons(inputId = "whichModel", 
-                    label = 'Which model are you designating?', 
-                    choices = list("Null (simulate M-1 new plots)" = "null",
-                                   "Alternative (simulate 1 new plot)" = "alt"),
-                    selected = "alt"),
+                   label = 'Which model are you designating?', 
+                   choices = list("Null (simulate M-1 new plots)" = "null",
+                                  "Alternative (simulate 1 new plot)" = "alt"),
+                   selected = "alt"),
       # Input: pick a model
       selectInput(inputId = "model",
                   label = "Select an effect to test:",
                   choices = list("basic" = "basic", "jttp (-3.45)" = "jttp", "jtts (3.34)"= "jtts", 
-                              "samep (.20)" = "samep", "samettp (1.33)" = "samettp", "simttb (10.09)" = "simttb"),
+                                 "samep (.20)" = "samep", "samettp (1.33)" = "samettp", "simttb (10.09)" = "simttb"),
                   selected = "basic"),
       # Input: pick a wave 
       numericInput(inputId = "wave",
@@ -47,6 +47,11 @@ ui <- fluidPage(
       numericInput(inputId = "seed",
                    label = "Set a random seed (10,000-999,999):", 
                    value = 123456, min = 10000, max = 999999, step = 1),
+      selectInput(inputId = "layoutalg",
+                  label = "Select a layout algorithm", 
+                  choices = list("Fruchterman-Reingold" = "fruchtermanreingold",
+                                 "Kamada-Kawai" = "kamadakawai"),
+                  selected = "kamadakawai"),
       # Input: check a box to color the graphs
       checkboxInput(inputId = "color", 
                     label = "Check box to color clusters.",
@@ -64,9 +69,9 @@ ui <- fluidPage(
                  tabPanel("Data plot", textOutput(outputId = "dataPlot")),
                  # Output: data from the lineup
                  tabPanel("Lineup data", dataTableOutput(outputId = "lineupData"),
-                          downloadButton("downloadData", "Download")
-                 )
+                          downloadButton("downloadData", "Download"))
       )
     )
   )
-  )
+)
+

### 2017-10-11 GG-10-12/slides.Rmd
slides

---
title: "Applying visual inference to a social network model"
author: "Sam Tyner"
date: "10/12/2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, out.width = '100%', message = FALSE, warning = FALSE,
                      fig.align = 'center')
```

```{r}
library(tidyverse)
library(RSiena)
library(geomnet)
```

## Training 1

*Which plot is most different from the others?*

```{r train1}
dat <- read_csv("LineupData/samepx1sd234567pickalt.csv")
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5, 
                 layout.alg = 'fruchtermanreingold') + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
```

## Training 2

*Which plot is most different from the others?*

```{r train2}
dat <- read_csv("LineupData/jttsx5sd234567picknull.csv")
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5, 
                 layout.alg = 'fruchtermanreingold') + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
```

## Trial A
<!--basic 1.3-->
```{r trialA}
dat <- read_csv("LineupData/basicx1.3sd123456pickalt.csv")
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5, 
                 layout.alg = 'fruchtermanreingold') + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
```

## Trial B
<!--jttp -1.5-->

```{r trialB}
dat <- read_csv("LineupData/jttpx-1.5sd98765pickalt.csv")
ggplot(data = dat) + 
        geom_net(aes(from_id = from, to_id = to), 
                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
        theme_net() + 
        theme(panel.background = element_rect(color = 'black')) +
        facet_wrap(~ord)
```

## Trial C
<!--jtts 2-->

## Trial D
<!--simttb 2-->

## Trial E
<!--basic-recip -5-->

## Trial F
<!--basic -1-->

## Trial G
<!--samettp 2-->

## Trial H
<!--basic-density -1-->

## Trial I
<!--basic-recip 1.5-->

## Trial J
<!--jtts -5-->

## Trial K
<!--samettp 2-->

## Trial L
<!--simttb -1-->

## Trial M
<!--jttp 4-->

## Trial N
<!--basic-density 1.2-->

## Trial O
<!--jttp 4-->

## Trial P
<!--samettp -1 -->

### 2017-10-12 GG-10-12/slides.Rmd
lineups for tomorrow

@@
 library(geomnet)
+library(igraph)
 ```
@@
 
+```{r trialC}
+dat <- read_csv("LineupData/jttsx2sd123456pickalt.csv")
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
+        theme_net() + 
+        theme(panel.background = element_rect(color = 'black')) +
+        facet_wrap(~ord)
+```
+
 ## Trial D
 <!--simttb 2-->
+
+```{r trialD}
+dat <- read_csv("LineupData/simttbx2sd123456pickalt.csv")
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
+        theme_net() + 
+        theme(panel.background = element_rect(color = 'black')) +
+        facet_wrap(~ord)
+```
 
@@
 
+```{r trialE}
+dat <- read_csv("LineupData/basicx-5sd346231pickalt.csv")
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
+        theme_net() + 
+        theme(panel.background = element_rect(color = 'black')) +
+        facet_wrap(~ord)
+```
+
 ## Trial F
 <!--basic -1-->
+
+```{r trialF}
+dat <- read_csv("LineupData/basicx-1sd346231pickalt.csv")
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5, 
+                 layout.alg = "fruchtermanreingold") + 
+        theme_net() + 
+        theme(panel.background = element_rect(color = 'black')) +
+        facet_wrap(~ord)
+```
 
@@
 
+```{r trialG}
+dat <- read_csv("LineupData/samettpx2sd23452pickalt.csv")
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5, 
+                 layout.alg = "fruchtermanreingold") + 
+        theme_net() + 
+        theme(panel.background = element_rect(color = 'black')) +
+        facet_wrap(~ord)
+```
+
 ## Trial H
 <!--basic-density -1-->
+
+```{r trialH}
+dat <- read_csv("LineupData/basicx-1sd987234pickalt.csv")
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5, 
+                 layout.alg = "fruchtermanreingold") + 
+        theme_net() + 
+        theme(panel.background = element_rect(color = 'black')) +
+        facet_wrap(~ord)
+```
 
@@
 
+```{r trialI}
+dat <- read_csv("LineupData/basicx1.5sd232342pickalt.csv")
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5, 
+                 layout.alg = "fruchtermanreingold") + 
+        theme_net() + 
+        theme(panel.background = element_rect(color = 'black')) +
+        facet_wrap(~ord)
+```
+
 ## Trial J
@@
 
+```{r trialJ}
+dat <- read_csv("LineupData/jttsx-5sd45745pickalt.csv")
+
+ggplot(data = dat, aes(from_id=from, to_id=to, colour=log(csize))) + 
+        geom_net(arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 arrowgap = .015, arrowsize = .3, size = .75, layout.alg = 'fruchtermanreingold') + 
+        theme_net() +
+        theme(panel.background = element_rect(color = 'black')) +
+        scale_colour_gradient(low="#f768a1", high = "#49006a") + facet_wrap(~ord) +
+        theme(legend.position="none")
+```
+
 ## Trial K
-<!--samettp 2-->
+<!--samettp -1-->
+
+```{r trialK}
+dat <- read_csv("LineupData/samettpx-1sd35792pickalt.csv")
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
+        theme_net() + 
+        theme(panel.background = element_rect(color = 'black')) +
+        facet_wrap(~ord)
+```
 
@@
 
+```{r trialL}
+dat <- read_csv("LineupData/simttbx-1sd234624pickalt.csv")
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
+        theme_net() + 
+        theme(panel.background = element_rect(color = 'black')) +
+        facet_wrap(~ord)
+```
+
 ## Trial M
 <!--jttp 4-->
+
+```{r trialM}
+dat <- read_csv("LineupData/jttpx4sd234624pickalt.csv")
+df <- data.frame(expand.grid(
+        from=paste0("V", 1:155), to = paste0("V", 1:155)), stringsAsFactors = FALSE)
+netnest <- dat %>% group_by(sim) %>% nest()
+netnest <- netnest %>% mutate(
+        data = data %>% purrr::map(.f = function(d) {
+          adj1 <- d %>% right_join(df) %>% 
+            xtabs(!is.na(dep.var.id)~from+to, data =.) %>% as.matrix()
+
+          ig1 <- graph_from_adjacency_matrix(adj1)
+          comps <- components(ig1)
+          clu_df <- data.frame(from=names(comps$membership), clu = comps$membership)
+          clu_df$csize=comps$csize[clu_df$clu]
+          d %>% left_join(clu_df, by="from")
+        })
+) 
+dat <- unnest(netnest, data)
+ggplot(data = dat, aes(from_id=from, to_id=to, colour=log(csize))) + 
+        geom_net(arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 arrowgap = .015, arrowsize = .3, size = .75, layout.alg = 'fruchtermanreingold') + 
+        theme_net() +
+        theme(panel.background = element_rect(color = 'black')) +
+        scale_colour_gradient(low="#f768a1", high = "#49006a") + facet_wrap(~ord) +
+        theme(legend.position="none")
+```
 
@@
 
-## Trial O
-<!--jttp 4-->
-
-## Trial P
-<!--samettp -1 -->
+```{r trialN}
+dat <- read_csv("LineupData/basicx1.2sd34231pickalt.csv")
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
+        theme_net() + 
+        theme(panel.background = element_rect(color = 'black')) +
+        facet_wrap(~ord)
+```

### 2017-10-12 ASApaper/paper.Rmd
finalize lineups for today, add table to paper of effects.

@@
 
+preamble: 
+  - \usepackage{amssymb}
 bibliography: bibliography.bib
@@
 
-```{r senateAll, fig.cap="The four senate collaboration networks that we estimate "}
+```{r senateAll, fig.cap="The four senate collaboration networks that we use as our example data to visually assess the SAOM effects.", fig.height=4}
 seobama <- read_csv("../Data/senate/senateobamapres_gsw_25.csv")
@@
 ```
-```{r senateClinton}
+
+```{r senateClinton, fig.cap="We removed Hillary Clinton's ties from the network because she had abnormally high collaboration with senators during the time she was in the 111th senate and before she left office to become Secretary of State.", fig.height=2.5}
 se111clint$Clinton <- 'Yes'
@@
 
-In cosponsorship networks, it is well known that party affiliation and reciprocity of relationships are major influences on structure [@legnet].
+In legislative cosponsorship networks, it is well known that party affiliation and reciprocity of relationships are major influences on structure [@legnet].
 
@@
 
-After completing the procedure for all model effects, we selected effects to include that converged and had a Wald $p$-value of less than 0.10, and seemed to have a reasonable interpretation for our data.
+After completing the procedure for all model effects, we selected effects whose estimates converged, had a Wald $p$-value of less than 0.10, and seemed to have a reasonable interpretation for our data according to well-known properties of legislative networks [@legnet].
 
@@
 This effect relies on the number of transitive closures formed between two senators from different parties.
-Due to the divisive nature of American politics, the party effect is dominant.
+The negative estimate is an indication that forming transitive ties between two people from different parties is discouraged, which tracks with the divisive nature of American politics, where party affilitation is dominant.
 Another significant effect was the same JTT parameter for the sex covariate, with an estimate of about 3 with a standard error of 0.89.
@@
 
+```{r geteffects, eval = FALSE}
+# used for table below
+initeff <- read_csv("../Data/senate/sigEffsSenate912.csv")
+initeff %>% filter(shortName == "sameX")
+#density & -- & $\sum_j x_{ij}$ & & & \\
+#reciprocity & -- & $\sum_j x_{ij}x_{ji}$ & & & \\
+```
+
 List of effects: 
-\begin{enumerate}
-\item jtt - party
-\item jtt - sex
-\item similiarity trans trip - bills
-\item same tt - party
-\end{enumerate}
-
+\begin{table}
+\scalebox{0.9}{
+\begin{tabular}{p{2cm}|p{2.1cm}|c|l|p{1.75cm}|p{2cm}}
+{\bf Effect name} & {\bf Interaction Variable} & {\bf Formula} & {\bf Picture} & {\bf Initial estimate} & {\bf Wald $p$-value} \\
+\hline 
+jumping transitive triplet & party & $\sum_{j\neq h} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(p_i = p_h \neq p_j)$ & \includegraphics[width=.6in]{img/jttp.png} & -5.884 & $<0.0001$\\
+jumping transitive triplet & sex & $\sum_{j\neq h} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(s_i = s_h \neq s_j)$ & \includegraphics[width=.6in]{img/jtts.png}  & 3.335 & 0.0002 \\
+similarity transitive triplet & bills & $\sum_{j} x_{ij}x_{ih}x_{hj}\cdot (sim^b_{ij} - \overline{sim}^b)^*$ & \includegraphics[width=1in]{img/simttb.png} & 9.821 & 0.0128 \\
+same transtive triplet & party & $\sum_{j} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(p_i = p_j)$ & \includegraphics[width=1in]{img/samettp.png}  & 1.306 & 0.0642 \\
+same & party & $\sum_j x_{ij}\mathbb{I}(p_i = p_j)$ & \includegraphics[width=.6in]{img/samep.png}  & 0.363 & 0.0074
+\end{tabular}
+}
+\caption{\label{tab:effects} The additional effects we used in the SAOMs fit to the senate data.
+* - $sim^b_{ij} = \frac{\max_{hk}|b_h - b_k| - |b_i - b_j|}{\max_{hk}|b_h - b_k|}$ is the similarity score between two senators based on the number of bills authored, and $\overline{sim}^b = \frac{1}{n(n-1)}\sum_{i\neq j} sim^b_{ij}$ is the average bill similarity score between any two senators.}
+\end{table}
 # Simulation Procedure

### 2017-10-12 GG-10-12/slides.Rmd
finalize lineups for today, add table to paper of effects.

@@
 
+## Trial D2
+
+```{r trialD2}
+#same as K with KK instead of FR layout
+dat <- read_csv("LineupData/samettpx-1sd35792pickalt.csv")
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
+        theme_net() + 
+        theme(panel.background = element_rect(color = 'black')) +
+        facet_wrap(~ord)
+```
+
 ## Trial E
@@
                  linewidth = .25, singletons = T, fiteach = T, directed = T, 
-                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5,
+                 layout.alg = 'fruchtermanreingold') + 
         theme_net() + 
@@
                  linewidth = .25, singletons = T, fiteach = T, directed = T, 
-                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5, 
-                 layout.alg = "fruchtermanreingold") + 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
         theme_net() + 
@@
                  linewidth = .25, singletons = T, fiteach = T, directed = T, 
-                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5, 
+                 layout.alg = 'fruchtermanreingold') + 
         theme_net() + 

### 2017-10-12 Code/Senate/creating_lus.R

@@ -196,3 +196,13 @@ lineup_details <- senate_lu_storage(details = make_lu_dat, Altsims = gotsims, Nu
 make_lu_dat[which(make_lu_dat$pic_id == 147),]
 
 senate_lu(altsims = gotsims, nullsims = datSims, M = 6, seed = make_lu_dat[which(make_lu_dat$pic_id == 151),"seed"][[1]], w = 1, mod = make_lu_dat[which(make_lu_dat$pic_id == 151),"model"][[1]], cond = make_lu_dat[which(make_lu_dat$pic_id == 151),"condition"][[1]])
+
+# final details 
+
+effectsTest <- c("basic-both", "basic-density", "basic-reciprocity", "jttp", "jtts", "simttb", "samettp")
+condition1 <- c("negative", "positive")
+condition2 <- c("easy", "medium", "hard")
+
+allconditions <- expand.grid(effectsTest, condition1, condition2)
+names(allconditions) <- c("effect", "sign", "difficulty")
+allconditions <- bind_rows(allconditions, allconditions, allconditions) %>% mutate(rep = rep(1:3, each = 42)) %>% arrange()

### 2017-10-19 ASApaper/paper.Rnw
introduction about 2/3 ish done

\documentclass[12pt]{article}
% \usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
% \geometry{letterpaper}                   % ... or a4paper or a5paper or ...
%\usepackage{graphicx}
\usepackage[font=small,skip=5pt]{caption}
\usepackage{subcaption}
\usepackage{animate}
\usepackage{afterpage}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{amsfonts}
% \usepackage{color}
\usepackage{multirow}
\usepackage{rotating}
\usepackage[dvipsnames,svgnames,table]{xcolor}
\usepackage{hyperref}
\graphicspath{{figure/}}
% \usepackage{endfloat} % Figures to the end of the document

\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
%---------------------------------------------------
%                 Editing Commands
%\newcommand{\hh}[1]{{\color{magenta} #1}}
%\newcommand{\st}[1]{{\color{orange} #1}}

%---------------------------------------------------
%                 Placing Figures
\renewcommand{\topfraction}{0.99}	% max fraction of floats at top
\renewcommand{\bottomfraction}{0.8}	% max fraction of floats at bottom
\renewcommand{\textfraction}{0.05}	% allow minimal text w. figs

%---------------------------------------------------
% Define new environment
\newtheorem{theorem}{Theorem}[section]
\newtheorem{algorithm}[theorem]{Algorithm}
%---------------------------------------------------

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\begin{document}

%\bibliographystyle{natbib}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\if0\blind
{
  \title{\bf Visual Inference for Significance and Goodness-of-Fit Testing for a Social Network Model}
  \author{Samantha Tyner\thanks{
    The authors gratefully acknowledge funding from the National Science Foundation Grant \# DMS 1007697. All data collection has been conducted with approval from the Institutional Review Board IRB 10-347}\hspace{.2cm}\\
    Department of Statistics and Statistical Laboratory, Iowa State University\\
    and \\
    Heike Hofmann\\
    Department of Statistics and Statistical Laboratory, Iowa State University}
  \maketitle
} \fi

\if1\blind
{
  \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Visual Inference for Significance and Goodness-of-Fit Testing for a Social Network Model}
\end{center}
  \medskip
} \fi

\bigskip
\begin{abstract}
Abstract goes here.
\end{abstract}

\noindent%
{\it Keywords:} social network analysis, visual inference, dynamic networks, network visualization, network mapping, goodness-of-fit, hypothesis testing
\vfill

\newpage
\spacingset{1.45} % DON'T change the spacing!

\tableofcontents
\newpage

<<child-demo, child='paper-child.Rnw'>>=
@

\bibliographystyle{asa}
\bibliography{bibliography}

\end{document}

### 2017-10-21 ASApaper/paper-child.Rnw
more paper stuffs - mostly saom intro, fix error in fig 2

\newcommand{\st}[1]{{\color{orange} #1}}
\newcommand{\hh}[1]{{\color{magenta} #1}}

<<setup, echo = FALSE, message = FALSE, warning = FALSE>>=
options(replace.assign=TRUE,width=70, digits=3)
require(knitr)
opts_chunk$set(fig.path='figure/', cache.path='cache/', fig.align='center', fig.pos='h', out.width='.99\\textwidth', par=TRUE, cache=FALSE, concordance=TRUE, autodep=TRUE, message=F, warning=F, echo = FALSE, dev="cairo_pdf", fig.width = 6, fig.height = 6)#, root.dir = "~/Desktop/Dissertation/SAOM-removing-blindfold/")
@

<<pkgs>>=
library(tidyverse)
library(RSienaTest)
library(RSiena)
library(geomnet)
library(GGally)
library(extrafont)
loadfonts(quiet = T)
load("../Data/senate/senateSienaNoHRC.rda")
ThemeNoNet <- theme_bw() %+replace% 
            theme(plot.title = element_text(size = 12,
                                            face = 'plain', 
                                            angle = 0, 
                                            family="Times New Roman"), 
                  axis.title.x = element_text(size = 12,
                                            face = 'plain', 
                                            angle = 0,
                                            family="Times New Roman"),
                  axis.title.y = element_text(size = 12,
                                            face = 'plain', 
                                            angle = 90,
                                            family="Times New Roman"),
                  axis.text.x.top = element_text(size = 12,
                                            face = 'plain', 
                                            angle = 0,
                                            family="Times New Roman"), 
                  axis.text.x.bottom = element_text(size = 12,
                                            face = 'plain', 
                                            angle = 0,
                                            family="Times New Roman"), 
                  axis.text.y.left = element_text(size = 12,
                                            face = 'plain', 
                                            angle = 0,
                                            family="Times New Roman"),
                  axis.text.y.right = element_text(size = 12,
                                            face = 'plain', 
                                            angle = 0,
                                            family="Times New Roman"),
                  strip.text.x = element_text(size = 12,
                                            face = 'plain', 
                                            angle = 0, 
                                            family="Times New Roman",
                                            margin = margin(t = 3, r = 0, b = 3, l = 0, unit = "pt")),
                  strip.text.y =element_text(size = 12,
                                            face = 'plain', 
                                            angle = 90,
                                            family="Times New Roman",
                                            margin = margin(t = 3, r = 3, b = 3, l = 3, unit = "pt")),
                  strip.background = element_rect(colour = "black", fill = "white")
                    )
ThemeNet <- theme_net() %+replace% 
            theme(plot.title = element_text(size = 12,
                                            face = 'plain', angle = 0, family = "Times New Roman"), 
                  strip.text.x = element_text(size = 12,
                                            face = 'plain', angle = 0, family = "Times New Roman", 
                                            margin = margin(t = 3, r = 0, b = 3, l = 0, unit = "pt")),
                  strip.text.y =element_text(size = 12,
                                            face = 'plain', angle = 0, family = "Times New Roman",margin = margin(t = 3, r = 0, b = 3, l = 0, unit = "pt")),
                  panel.border = element_rect(fill = NA, color = 'black'),
                  strip.background = element_rect(colour = "black", fill = "white"))
@

\section{Introduction}
\st{
\begin{enumerate}
\item significance testing: what is it and why is it important
\item goodness-of-fit testing: what is it and why is it important
\item difficulty of these two for sna, why do vis inf
\end{enumerate}
}
Two of the most imporant pieces of statistical modeling are significance testing of model parameters and goodness-of-fit measures and tests. In the former, the data are usually assumed to come from a simple model under the null hypothesis, and additional parameters are added and tests done to determine if the additional should be included in the model. In the latter, the model of interest is examined to determine how well it fits the data. Both of these aspects of statistical modeling increase greatly in difficulty as the complexity of the model of interest increases. The more complicated the model, the harder it is to determine fit or to include or exclude a parameter.

Some particularly complicated sets of models are those designed to model network change. A \textit{network} is any set of things, such as people, computers, or neurons, that are connected in some way, through social relations, internet connection, or electrical impulses in the brain. We refer to the ``things" in the network as \textit{nodes}, or \textit{actors} in a social network, and the connections as \textit{edges}, or \textit{ties} in a social network. Network objects are particularly difficult to model due to the dependencies inherent to the data.

XX smart stuff with more refs XX

The paper is outlined as follows: Section 2 gives a basic overview of visual inference and the lineup protocal. Section 3 provides an introduction to the our models of interest, stochastic actor-oriented models. Sections 4 and 5 discuss problems with traditional significance tests and goodness-of-fit measures for these models and how visual inference methods can be used as solutions. Section 6 outlines our experimental protocol, and we close with a discussion in Section 7.

\section{Visual Inference}

Data visualizations are an important component of data analysis, providing a mechanism for discovering patterns in data. Pioneering research by \citet{gelman:2004}, \citet{Bujaetal} and \citet{majumder:2011} provide methods to quantify the significance of discoveries made from visualizations. \citet{Bujaetal} introduced two protocols, the Rorschach and the lineup protocol, which bridge the gulf between traditional statistical inference and exploratory data analysis. %The Rorschach protocol consists of a set of $m$ (usually, $m=20$) plots (called the {\it null plots}) rendered from data that is consistent with a given null model. The Rorschach protocol helps to understand the extent of randomness in the null model. 
We use line lineup protocol. Under this protocol, a plot of the observed data is placed randomly among a set of $m-1$ null plots, and human observers are then asked to examine the lineup and to identify the most different plot. If observers identify the data plot, this is quantifiable evidence against the null hypothesis.

The lineup protocol places a plot firmly in the framework of hypothesis tests: a plot of the data is considered to be the test statistic, which is compared against the sampling distribution under the null hypothesis represented by the null plots. Obviously, the null generating mechanism, i.e.\ the method of obtaining the data for null plots, is crucial for both the lineup and the Rorschach protocol, as the null hypothesis directly affects the choice of null generating method. Null generating methods are typically based on (a) simulation, if the null hypothesis allows us to directly specify a parametric model, (b) sampling, as for example in the case of large data sets, or (c) permutation of the original data \citep[see e.g.\ ][]{Good05}, which allows for non-parametric testing that preserves marginal distributions  while ensuring independence in higher dimensions. The model of interest here allows us to simulate directly from a parameteric model for dynamic social network data. 
%In the experimental data that we analyzed the null generating methods used were permutation methods and direct simulation from a null model.

The lineup protocol was formally tested in a head-to-head comparison with the equivalent conventional test in \citet{majumder:2011}. The experiment utilized human subjects from Amazon's Mechanical Turk \citep{turk} and used simulation to control conditions. The results suggest that  visual inference is comparable to conventional tests in a controlled conventional setting. This provides support for its appropriateness for testing in real exploratory situations where no conventional test exists. %Interestingly, the power of a visual test increases with the number of observers engaged to evaluate lineups, and the pattern in results suggests that the power will provide results consistent with practical significance \citep{kirk:1996}.

\section{Stochastic Actor-Oriented Models}

%To model family that we apply to the Senate data is the Stochastic Actor-Oriented Model (SAOM) family. 
This family of models for dynamic network data, which were first introduced by \citet{saompaper}, incorporates both network structure and node-level information to describe how a network observed on two or more occasions changes in time. The two titular properties of SAOMs, stochasticity and actor-orientation, are crucial to understanding networks as they exist naturally: social networks are ever-changing as relationships decay or grow in seemingly random ways, and most actors in them have characteristics that could affect how they change their ties to other nodes in the network. 
% Talk about SAOMs
These unique properties allow for the fitting of some very complicated models to inherently complex data, so it can be exceedingly difficult to interpret parameters and their corresponding estimates. The sheer amount of possible parameters to include in the model combined with the difficulty of interpretation make parameter selection and goodness-of-fit testing burdensome as well.

Broadly, a SAOM takes network structure and node covariate information into account in two ways: first through a rate function that describes \textit{when} changes in the network occur, and second through an objective function that describes \textit{what} those changes are. As in many other network models, the variables of interest, $x_{ij}$, are the binary edges of the network. For any two nodes $i \neq j$ in the network, $i,j \in \{1, 2, \dots, n=\text{the number of nodes}\}$, 
  \begin{equation}\label{eq:edgevars}
  x_{ij} =
  \begin{cases}
                                   1 & \text{if an edge from $i$ to $j$ exists} \\
                                   0 & \text{otherwise}
  \end{cases}
  \end{equation}
The edges are also treated as \textit{directed}, so it is not required that $x_{ij} = x_{ji}$, and self-referencing edges or loops, $x_{ii}$, are not allowed. The entire network at time point $m = 1, 2, \dots M$ for $M$ the number of network observations, is denoted $x(t_m)$. More detail on SAOMs follows, and additional details can be found in \citet{saompaper, snijders01, snijders2010, snijdersetall:2007, snijders:2010, snijders:2017},

\subsection{Rate Function}

All changes in SAOMs are treated as changes made by the nodes, or \textit{actors}, in the network. So, each actor, $i$, gets a chance to make a change according to the rate function, typically denoted $\lambda_i$, which dictates when relationships between nodes in the network can change. In general, the rate function can take the network structure e.g. outdegree of node $i$, and the node covariates into account, but we use the simple rate function, which is constant over all nodes in a given time period. We denote the rate from $t_{m}$ to $t_{m+1}$ as $\alpha_m$ for $m = 1, \dots, M-1$. Using this notation, the waiting time to the next chance for actor $i$ to make a change is exponentially distributed with expected value $\alpha_m^{-1}$. Since the rate is the same for all actors, the waiting time for \textit{any} actor to get the chance to change is exponentially distributed with expected value $(n\alpha_m)^{-1}$.

\subsection{Objective Function}

After actor $i$ has been given the opportunity to change, it probabilistically chooses one of its current ties, $x_{ij}$, to change. The probability that actor $i$ changes its current tie to actor $j$ is determined by the \textit{objective function} of the model and a random component, $U$. The random component is assumed to have log-Weibull distribution with location parameter $\mu = 0$ and scale parameter $\sigma = 0$ (see \citet{modelsSnijders}), so the probability density function of $U$ is
\begin{equation}\label{eq:logweibull}
g(u) = \exp\{-(u + \exp\{-u\})\}.
\end{equation}
The objective function, $f_i$, is the driving force of a SAOM. Actor $i$ always wants to maximize its objective function given the current state of the network, $x$ and the node-level covariates, $\mathbf{Z}$:% which has the form
\begin{equation}\label{eq:objective}
f_i(x, \boldsymbol{\beta}, \mathbf{Z}) = \sum_{k = 1}^K \beta_k s_{ik}(x, \mathbf{Z}),
\end{equation}
where $\boldsymbol{\beta} = (\beta_1, \dots, \beta_K)$ are additional model parameters, each associated with some network statistics calculated with respect to actor $i$, $s_{ik}(x, \mathbf{Z})$. The network statistics range from the simple outdegree, $s_i(x) = \sum_{i\neq j} x_{ij}$, to the more complicated transitive triplets jumping to different covariate, $s_i(x, \mathbf{Z}) = \sum_{i \neq j \neq h} x_{ij}x_{ih}x_{hj} \cdot \mathbb{I}(z_i = z_h \neq z_j)$, plus many more. At last count, in the software we use to fit these models to network data, \texttt{RSiena}, there are over 80 possible effects that can be included in the objective function \citep{RSiena, RSienamanual}.

\subsection{Example Data}

The data we use are of collaboration networks in the United States Senate during the $111^{th}$ through $114^{th}$ Congresses, overlapping with Barack Obama's presidency. These senates began on January 6, 2009 and ended on January 3, 2017.\footnote{Details of how this data can be downloaded are provided by François Briatte at \url{https://github.com/briatte/congress}}. In this network, ties are directed from senator $i$ to senator $j$ when senator $i$ signs on as a cosponsor to the bill that senator $j$ authored. There are (somewhat surprisingly) many hundreds of ties between senators when they are connected in this way, so we simplify the network by computing a single value for each senator-senator collaboration called the \textit{weighted propensity to cosponsor} (WPC). This value is defined in \citet{senate} as

\begin{equation}\label{eq:sen1}
    WPC_{ij} = \dfrac{\sum\limits_{k=1}^{n_j} \frac{Y_{ij(k)}}{c_{j(k)}}}{\sum\limits_{k=1}^{n_j} \frac{1}{c_{j(k)}}}
\end{equation}

where $n_j$ is the number of bills in a congressional session authored by senator $j$, $c_{j(k)}$ is the number of cosponsors on senator $j$'s $k^{th}$ bill, where $k \in \{1,\dots, n_j\}$, and $Y_{ij(k)}$ is a binary variable that is 1 if senator $i$ cosponsored senator $j$'s $k^{th}$ bill, and is 0 otherwise. This measure ranges in value from 0 to 1, where $WPC_{ij} = 1$ if senator $i$ is a cosponsor on every one of senator $j$'s bills and $WPC_{ij} = 0$ if senator $i$ is never a cosponsor any of senator $j$'s bills.

<<senateAll, fig.cap="The four senate collaboration networks that we use as our example data to visually assess the SAOM effects.">>=
seobama <- read_csv("../Data/senate/senateobamapres_gsw_25.csv")
to_label <- c("Joseph R. Biden Jr.", "John S. McCain", "Ted Cruz","Marco Rubio",
  "Lindsey O. Graham","Rand Paul", "Bernard Sanders", "Jim Webb", "Mitch McConnell",
  "Harry M. Reid", "Hillary Rodham Clinton", "Amy Jean Klobuchar")
seobama$label <- as.factor(ifelse(seobama$source %in% to_label, seobama$source, ""))
levels(seobama$label) <- c("", "Amy Klobuchar", "Bernie Sanders", "Harry Reid",
                           "Hillary Clinton", "Jim Webb", "John McCain", "Joe Biden",
                           "Lindsey Graham", "Marco Rubio", "Mitch McConnell", 
                           "Rand Paul", "Ted Cruz")
se111clint <- filter(seobama, senate == 111)
se111noclint <- filter(seobama, senate == 111)
se111noclint$target[which(se111noclint$target == "Hillary Rodham Clinton")] <- NA
seobama2 <- seobama %>% filter(senate != 111) %>% bind_rows(se111noclint)

set.seed(56049382)
ggplot(data = seobama2) + 
  geom_net(directed = T, labelon=T, arrowsize = .25, singletons= T, fiteach = T, linewidth = .25, layout.alg = 'fruchtermanreingold', fontsize = 2, repel=T, 
           aes(from_id = source, to_id = target, color = party, label = label),
           size = 1) + 
  ThemeNet + 
  scale_color_manual(values = c("royalblue", "forestgreen","firebrick")) + 
  theme(legend.position = 'bottom', strip.text = element_text(size = 8)) + 
  facet_wrap(~senate, nrow = 2, labeller = "label_both") + 
  xlim(c(0,1.05)) + 
  ylim(c(0,1.05))
@

For each senate session, we have the WPC value between any two senators in the session, the party affiliation of each senator, the number of bills they authored in each session, their age, and their sex. We explored each of these covariates in the model to determine if they affect the change of the network structure.

<<senateClinton, fig.cap="We removed Hillary Clinton's ties from the network because she had abnormally high collaboration with senators during the time she was in the 111th senate and before she left office to become Secretary of State.", fig.height=3>>=
se111clint$Clinton <- 'Yes'
se111noclint$Clinton <- 'No'
clintonSenate <- rbind(se111clint, se111noclint)
clintonSenate$Clinton <- as.factor(clintonSenate$Clinton)
clintonSenate$Clinton <- ordered(clintonSenate$Clinton, levels = c("Yes", "No"))
clintonSenate %>% filter(!(source %in% c("Roland Burris", "Bernard Sanders", "Frank R. Lautenberg", "Mary L. Landrieu") & is.na(target))) %>%
ggplot() + 
  geom_net(directed = T, labelon=F, arrowsize = .3, singletons= F, fiteach = T, arrowgap = .01, layout.alg = 'fruchtermanreingold',
           aes(from_id = source, to_id = target, color = party, linewidth = gsw, shape = sex,size = n_au)) + 
  scale_size_continuous(name="# bills\nauthored") +
  ThemeNet + 
  scale_shape_manual(values = c(17,16)) + 
  scale_color_manual(values = c("royalblue", "forestgreen","firebrick")) + 
  theme(legend.position = 'bottom') + 
  facet_wrap(~Clinton, nrow = 1, labeller = 'label_both')
@

In legislative cosponsorship networks, it is well known that party affiliation and reciprocity of relationships are major influences on structure \citep{legnet}.

\subsection{Models of Interest}

In order to determine which models to explore for application of our novel goodness-of-fit testing method, we first fit many possible models and selected a few significant effects. To determine the significance of the effects, we followed this procedure:

\begin{enumerate}
\item Define the simple effects structure of the data: the rate parameters and the outdegree and reciprocity parameters. 
\item Add each possible effect one-at-a-time to the model structure, as determined by the effects documentation function.
\item Fit each model to the data and check for convergence.
    \begin{enumerate}
    \item If the model converged, move to 4.
    \item If the model did not converge, use the previous fitted values as starting values and repeat 5 times or until convergence, whichever comes first.
    \end{enumerate}
\item Test the added parameter for significance using the Wald-type test.
\item Report out the estimate of the additional parameter, its standard error, Wald $p$ value, and convergence criterion.
\end{enumerate}

After completing the procedure for all model effects, we selected effects whose estimates converged, had a Wald $p$-value of less than 0.10, and seemed to have a reasonable interpretation for our data according to well-known properties of legislative networks \citep{legnet}.

The most significant effect was the jumping transitive triplet (JTT) parameter for the party covariate, which was estimated to be about -6 with a standard error of 0.11. This effect relies on the number of transitive closures formed between two senators from different parties. The negative estimate is an indication that forming transitive ties between two people from different parties is discouraged, which tracks with the divisive nature of American politics, where party affilitation is dominant. Another significant effect was the same JTT parameter for the sex covariate, with an estimate of about 3 with a standard error of 0.89. Another significant effect was the covariate-related similarity score-weighted transitive triplets for the number of bills authored by a senator. This effect was estimated at about 10 with standard error of 3.9, and the high positive effect suggests senators tend to collaborate with other senators who author about the same number of bills they do. Finally, we found the same party transitive triplet effect was also significant, with a fitted value of 1.3 and standard error of 0.7.

<<geteffects, eval = FALSE>>=
# used for table below
initeff <- read_csv("../Data/senate/sigEffsSenate912.csv")
initeff %>% filter(shortName == "sameX")
#density & -- & $\sum_j x_{ij}$ & & & \\
#reciprocity & -- & $\sum_j x_{ij}x_{ji}$ & & & \\
@

\begin{table}
\scalebox{0.9}{
\begin{tabular}{p{2cm}|p{2.1cm}|c|l|p{1.75cm}|p{2cm}}
{\bf Effect name} & {\bf Interaction Variable} & {\bf Formula} & {\bf Picture} & {\bf Initial estimate} & {\bf Wald $p$-value} \\
\hline 
jumping transitive triplet & party & $\sum_{j\neq h} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(p_i = p_h \neq p_j)$ & \includegraphics[width=.6in]{img/jttp.png} & -5.884 & $<0.0001$\\
jumping transitive triplet & sex & $\sum_{j\neq h} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(s_i = s_h \neq s_j)$ & \includegraphics[width=.6in]{img/jtts.png}  & 3.335 & 0.0002 \\
similarity transitive triplet & bills & $\sum_{j} x_{ij}x_{ih}x_{hj}\cdot (sim^b_{ij} - \overline{sim}^b)^*$ & \includegraphics[width=1in]{img/simttb.png} & 9.821 & 0.0128 \\
same transtive triplet & party & $\sum_{j} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(p_i = p_j)$ & \includegraphics[width=1in]{img/samettp.png}  & 1.306 & 0.0642 \\
same & party & $\sum_j x_{ij}\mathbb{I}(p_i = p_j)$ & \includegraphics[width=.6in]{img/samep.png}  & 0.363 & 0.0074
\end{tabular}
}
\caption{\label{tab:effects} The additional effects we used in the SAOMs fit to the senate data. * - $sim^b_{ij} = \frac{\max_{hk}|b_h - b_k| - |b_i - b_j|}{\max_{hk}|b_h - b_k|}$ is the similarity score between two senators based on the number of bills authored, and $\overline{sim}^b = \frac{1}{n(n-1)}\sum_{i\neq j} sim^b_{ij}$ is the average bill similarity score between any two senators.}
\end{table}

\section{Significance Testing}

We propose to attack some of the aforementioned difficulties with SAOMs through visual inference.

\begin{enumerate}
\item Fit each model 1,000 times and save the parameter estimates, standard errors, and overall convergence criterion.
\item From the models that converged, compute the mean of the 1,000 estimates of each parameter.
\end{enumerate}

We want to determine at which point an effect becomes noticeable in a SAOM. By \textit{noticeable}, we mean that the inclusion of the effect alters the appearance of networks simulated from a model. An effect becomes noticeable when as simulation from a model with a noticeable effect appears different from simulations from a model where the same affect is absent. Because the traditional way to visualize a network is a node-link diagram, we will examine the affect of different parameter values using node-link visualizations.

\section{Goodness-of-Fit Testing}

\section{Experiment Protocol}

\begin{enumerate}
\item Use the mean values from step 2 as the parameter values in the model and simulate 100 observations from the model. These simulated networks will become the "true data."
\item To each simulation, refit the model form from which it was simulated. 
\item Repeat steps 4-5 for various permutations of model forms and parameter values; e.g. double the parameter values, quadruple the parameter values, change the signs on the parameters, etc.
\item Simulate additional "alternative data" from the model in step 3 and the fitted model in step 4. 
\item Place the appropriate node-link diagram of each "true data" observation in a lineup with $M-1$ ($M=6$) "alternative data" node-link diagrams. 
\item Show the lineups to participants, with no participant seeing the same "true data" or "alternative data" plot twice. 
\end{enumerate}

\section{Discussion}

% Talk about visual inference
%We propose to attack some of the aforementioned difficulties with SAOMs by using a technique known as \textit{visual inference}. This technique was created by \citet{Bujaetal} to provide well-defined, statistical rigor to the usual exploratory data analyses and model diagnostics that are typically performed by visualizing the data as opposed to looking at it raw or gathering numerical summaries of it. In visual inference there are \textit{null plots} and \textit{data plots}: the null plots are visualizations of data simulated from the model according to the null hypothesis, while the data plots are visualizations of data simulated from the model according to an alternative hypothesis. These two data sources are visualized side-by-side using small multiples using what Buja et al called the \textit{lineup protocol}. The idea behind the lineup protocol is the police lineup, where witnesses to crimes are brought to the police station to observe a group of people, one of whom is suspected to have committed the crime, and are asked to identify the perpetrator of the crime. If the witness identifies the suspect in the lineup, that is taken as evidence the suspect is guilty, whereas if the witness does not identify the suspect, that is taken as evidence the suspect is not guilty. In the visual inference lineup protocol, the \textit{suspect} is the data plot among $M-1$ null plots. If the "witness" can pick the data plot out of the lineup, that is taken as evidence that the null model should be rejected, whereas failing to pick out the data plot is taken as evidence that the null model should not be rejected. In our application of the lineup protocol to SAOMs, we include various effects of varying sizes in the SAOMs we fit to some data and try to \textit{see} those effects in node-link diagram corresponding to simulations from the various models.

%-Talk about political networks-

%-Talk about goodness-of-fit-

%-Talk about significance testings-

### 2017-10-21 ASApaper/paper-child.Rnw
write write write write write

@@
 At last count, in the software we use to fit these models to network data, \texttt{RSiena}, there are over 80 possible effects that can be included in the objective function \citep{RSiena, RSienamanual}.
+We discuss these statistics in more detail in Section~\ref{sec:models}.
+
+The objective function $f_{i}(x, \boldsymbol{\beta}, \mathbf{Z})$ and the random component $U$ combine to form a \textit{transition probability}, $p_{ij}$, of the network changing from its current state $x$ to the state with only the tie $x_{ij}$ changed, which is denoted $x(i \leadsto j)$:
+\begin{equation}\label{eq:transprob}
+p_{ij} = \dfrac{\exp\{f_i(x(i \leadsto j), \boldsymbol{\beta}, \mathbf{Z})\}}{\sum_{h} \exp\{f_i(x(i \leadsto h), \boldsymbol{\beta}, \mathbf{Z})\}}
+\end{equation}
 
@@
 
-The data we use are of collaboration networks in the United States Senate during the $111^{th}$ through $114^{th}$ Congresses, overlapping with Barack Obama's presidency.
-These senates began on January 6, 2009 and ended on January 3, 2017.\footnote{Details of how this data can be downloaded are provided by François Briatte at \url{https://github.com/briatte/congress}}.
+The data we use are collaboration networks in the United States Senate during the $111^{th}$ through $114^{th}$ Congresses, overlapping with Barack Obama's presidency.
+These senates began on January 6, 2009 and ended on January 3, 2017\footnote{Details of how this data can be downloaded are provided by François Briatte at \url{https://github.com/briatte/congress}}.
+There are three legislative ways that senators can show support for legislation: they can author a bill, cosponsor a bill, and vote for a bill.
+We use cosponsorship as a metric because it results in a network that is unimodal (all nodes are senators) and directed.
 In this network, ties are directed from senator $i$ to senator $j$ when senator $i$ signs on as a cosponsor to the bill that senator $j$ authored.
-There are (somewhat surprisingly) many hundreds of ties between senators when they are connected in this way, so we simplify the network by computing a single value for each senator-senator collaboration called the \textit{weighted propensity to cosponsor} (WPC).
+There are many hundreds of ties between senators when they are connected in this way, so we simplify the network by computing a single value for each senator-senator collaboration called the \textit{weighted propensity to cosponsor} (WPC).
 This value is defined in \citet{senate} as
@@
 This measure ranges in value from 0 to 1, where $WPC_{ij} = 1$ if senator $i$ is a cosponsor on every one of senator $j$'s bills and $WPC_{ij} = 0$ if senator $i$ is never a cosponsor any of senator $j$'s bills.
+Because SAOMs require binary edges, we construct the edges as follows:
+ \begin{equation}\label{eq:edgewpc}
+  x_{ij} =
+\begin{cases}
+                                   1 & WPC_{ij} > 0.25 \\
+                                   0 & WPC_{ij} \leq 0.25
+\end{cases}
+\end{equation}
+For each of the four senate sessions, we have the WPC value between any two senators in the session, the party affiliation of each senator, the number of bills they authored in each session, and their gender.
+We explored each of these covariates in the model to determine if they affect the overall network structure and how ties are formed between senators.
+The node-link diagram representations of the data we use for modelling are shown in Figure~\ref{fig:senateAll}.
+We have labelled some of the nodes in these networks whose names will be familiar to US readers, because they are leaders in their party or they have run for president.
+The size of the nodes represent how many bills the senator authored in a session, the color represents party affiliation, and the shape represent gender.
+In each of the four sessions, there is one very large connected component tying many of the prominent senators together, with many smaller groups of two to ten senators surrounding the larger component.
+In each senate, the structure changes slightly as new senators arrive or come to prominence.
 
-<<senateAll, fig.cap="The four senate collaboration networks that we use as our example data to visually assess the SAOM effects.">>=
+<<senateAll, fig.cap="The four senate collaboration networks that we use as our example data to visually assess the SAOM effects.
+Color represents party, shape represents gender, and size represents number of bills authored in a session.
+The Frucherman-Reingold layout is shown.
+">>=
 seobama <- read_csv("../Data/senate/senateobamapres_gsw_25.csv")
@@
   "Lindsey O. Graham","Rand Paul", "Bernard Sanders", "Jim Webb", "Mitch McConnell",
-  "Harry M. Reid", "Hillary Rodham Clinton", "Amy Jean Klobuchar")
+  "Harry M. Reid", "Hillary Rodham Clinton", "Amy Jean Klobuchar", "Elizabeth Warren")
 seobama$label <- as.factor(ifelse(seobama$source %in% to_label, seobama$source, ""))
-levels(seobama$label) <- c("", "Amy Klobuchar", "Bernie Sanders", "Harry Reid",
-                           "Hillary Clinton", "Jim Webb", "John McCain", "Joe Biden",
-                           "Lindsey Graham", "Marco Rubio", "Mitch McConnell", 
+levels(seobama$label) <- c("", "Amy Klobuchar", "Bernie Sanders", "Elizabeth Warren",
+                           "Harry Reid", "Hillary Clinton", "Jim Webb", "John McCain", 
+                           "Joe Biden","Lindsey Graham", "Marco Rubio", "Mitch McConnell", 
                            "Rand Paul", "Ted Cruz")
@@
   geom_net(directed = T, labelon=T, arrowsize = .25, singletons= T, fiteach = T, linewidth = .25, layout.alg = 'fruchtermanreingold', fontsize = 2, repel=T, 
-           aes(from_id = source, to_id = target, color = party, label = label),
-           size = 1) + 
+           aes(from_id = source, to_id = target, color = party, 
+               label = label,shape = sex,size = n_au)) + 
   ThemeNet + 
   scale_color_manual(values = c("royalblue", "forestgreen","firebrick")) + 
+  scale_shape_manual(values = c(17,16)) + 
+  scale_size_continuous(name = "Bills\nauthored", range = c(.75, 3)) + 
   theme(legend.position = 'bottom', strip.text = element_text(size = 8)) + 
@@
 
-For each senate session, we have the WPC value between any two senators in the session, the party affiliation of each senator, the number of bills they authored in each session, their age, and their sex.
-We explored each of these covariates in the model to determine if they affect the change of the network structure.
+For Senate 111, for instance, we see Hillary Clinton, serving out her second term in the senate until she became Secretary of State.
+She is isolated in Figure~\ref{fig:senateAll}, but in actuality, she had many cosponsors on two pieces of legislation she authored in that short time, as is shown in Figure~\ref{fig:senateClinton}.
+We chose to remove Clinton and her edges from the network because they make the overall structure look so different from the other three senates, showing that the pattern is not typical of a senate in any other year.
+We suspect that because Hillary Clinton had just been appointed Secretary of State, the cosponsorships were largely symbolic, so the $111^{th}$ Senate without Hillary Clinton is more typical than the $111^{th}$ Senate with her.
 
@@
            aes(from_id = source, to_id = target, color = party, linewidth = gsw, shape = sex,size = n_au)) + 
-  scale_size_continuous(name="# bills\nauthored") +
+  scale_size_continuous(name = "Bills\nauthored", range = c(.75, 3)) + 
   ThemeNet + 
@@
 In legislative cosponsorship networks, it is well known that party affiliation and reciprocity of relationships are major influences on structure \citep{legnet}.
+We focus on these two covariates when choosing which SAO models to fit to the data.
 
-\subsection{Models of Interest}
+\subsection{Models of Interest}\label{sec:models}
 
-In order to determine which models to explore for application of our novel goodness-of-fit testing method, we first fit many possible models and selected a few significant effects.
-To determine the significance of the effects, we followed this procedure:
+In addition to considering already well-known effects in legislative networks for application of our significance and goodness-of-fit methods, we first fit many other possible models and selected a few significant effects.
+To determine the effects that we would move forward with, we followed this procedure:
 
@@
 \item Define the simple effects structure of the data: the rate parameters and the outdegree and reciprocity parameters. 
-\item Add each possible effect one-at-a-time to the model structure, as determined by the effects documentation function.
+\item Add each additional possible evaluation effect in \texttt{RSiena} one-at-a-time to the model structure, as determined by the effects documentation function \citep{RSiena}.
 \item Fit each model to the data and check for convergence.
@@
     \end{enumerate}
-\item Test the added parameter for significance using the Wald-type test.
+\item Test the added parameter for significance using a Wald-type test.
 \item Report out the estimate of the additional parameter, its standard error, Wald $p$ value, and convergence criterion.
@@
 
-The most significant effect was the jumping transitive triplet (JTT) parameter for the party covariate, which was estimated to be about -6 with a standard error of 0.11.
-This effect relies on the number of transitive closures formed between two senators from different parties.
+The parameters we use for the remainder of the paper are detailed in Table~\ref{tab:effects}.
+The most significant effect was the jumping transitive triplet (JTT) parameter for the party covariate, which was estimated to be about -6 with a standard error of 0.11, resulting in a $p$-value of less than 0.0001.
+This estimate of the parameter associated with this statistic relies on the number of transitive closures formed between two senators from different parties.
 The negative estimate is an indication that forming transitive ties between two people from different parties is discouraged, which tracks with the divisive nature of American politics, where party affilitation is dominant.
 Another significant effect was the same JTT parameter for the sex covariate, with an estimate of about 3 with a standard error of 0.89.
-Another significant effect was the covariate-related similarity score-weighted transitive triplets for the number of bills authored by a senator.
+The covariate-related similarity score-weighted transitive triplets parameter estimate for the number of bills authored by a senator was also significant.
 This effect was estimated at about 10 with standard error of 3.9, and the high positive effect suggests senators tend to collaborate with other senators who author about the same number of bills they do.
-Finally, we found the same party transitive triplet effect was also significant, with a fitted value of 1.3 and standard error of 0.7.
+This tendency of senators to cosponsor bills written by senators who are similarly ``prolific" corresponds to another well-known property of the U.S.
+Senate structure: the tendency of senators to be either ``workhorses" or ``showhorses".
+Senators known as workhorses author many pieces of legislation in a session, and largley stay out of the public arena.
+The showhorse senators, on the other hand, author relatively few pieces of legislation, and tend to appear on television, radio, and other media a great deal.
+Finally, we found the same party transitive triplet effect was also significant, with a fitted value of 1.3 and standard error of 0.7, meaning that transitive relationships between senators tend to form when they are from the same party.
 

### 2017-10-23 SenateLineupDataRefitModels/redos/redo_lus.R
add rest of refits

@@ -1,0 +1,173 @@
+# redoing some lineups 10-19-17
+library(tidyverse)
+library(netvizinf)
+library(geomnet)
+library(grid)
+library(gridSVG)
+
+# for the svgs 
+make_interactive <- function(filename2, script, toggle="toggle", high = "#c5c5c5", background="#ffffff") {
+  require(gridSVG)
+  grid.force()  
+  grobs <- grid.ls()
+  
+  idx <- grep("panel-", grobs$name)
+  for (d in idx) { 
+    grid.garnish(grobs$name[d],
+                 onmouseover=paste("frame('",grobs$name[d+2], ".1.1')", sep=""),
+                 onmouseout=paste("deframe('",grobs$name[d+2], ".1.1')", sep=""), 
+                 onmousedown=sprintf("%shigh(evt, '%s.1.1', '%s', '%s')", toggle, grobs$name[d+2], high, background)
+    )
+  }
+  # use script on server to get locally executable javascript code
+  # or use inline option (default)
+  grid.script(filename=script)
+  grid.export(filename2)
+}
+# for the pdfs 
+savepdf <- function(file, width=16, height=10){
+  fname <- paste(file,".pdf",sep="")
+  pdf(fname, width=width/2.54, height=height/2.54,
+      pointsize=10)
+  par(mgp=c(2.2,0.45,0), tcl=-0.4, mar=c(3.3,3.5,1.1,1))
+}
+
+# get all the data
+redo_dets <- data_frame(data_file = list.files("SenateLineupData/redos/data"))
+# create information necessary for making svgs, pdfs, pic details file
+redo_dets %>% mutate(data_name = data_file) %>% 
+  mutate(data_file = str_replace(data_file, ".csv", "")) %>%
+  separate(data_file, c("model", "sign", "difficulty", "rep"), "_") -> redo_dets
+redo_dets$model <- as.factor(redo_dets$model)
+levels(redo_dets$model) <- c(3,4,2,5)
+redo_dets$model <- as.integer(as.character(redo_dets$model))
+redo_dets$sign <- as.factor(redo_dets$sign)
+levels(redo_dets$sign) <- c(2,1)
+redo_dets$sign <- as.integer(as.character(redo_dets$sign))
+redo_dets$difficulty <- as.factor(redo_dets$difficulty)
+levels(redo_dets$difficulty) <- c(1,3,2)
[…]

### 2017-10-23 SenateLineupDataRefitModels/refit_models_server.R

@@ -1,0 +1,78 @@
+# refit-models csafe-01 server
+library(RSiena)
+library(dplyr)
+library(tibble)
+library(purrr)
+
+load("Senate_lineup_data_Siena.RDA")
+load("allModelMeans.RDS")
+modelMeanEsts <- modelMeanEsts %>% add_row(model = "bigmod", ests = NA)
+modelMeanEsts$ests <- list(modelMeanEsts$ests[[1]], modelMeanEsts$ests[[2]],
+                           modelMeanEsts$ests[[3]], modelMeanEsts$ests[[4]], 
+                           modelMeanEsts$ests[[5]],modelMeanEsts$ests[[6]],
+                           c(2.4405048,2.4594403,2.2098176,-4.9232775,4.8916183,2.3743720,0.2047038,6.9661589))
+
+
+refit_models <- function(dat, mod){
+  SenBasic <- getEffects(dat)
+  myalg <- sienaAlgorithmCreate( projname = Sys.time() , n3 = 1000, firstg = .02)
+  if (mod == "basic"){
+    myeffects <- SenBasic
+    myeffects$initialValue[myeffects$include] <- modelMeanEsts[modelMeanEsts$model == "basic","ests"][[1]][[1]]
+  } else if (mod == "jttp"){
+    myeffects <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+    myeffects$initialValue[myeffects$include] <- modelMeanEsts[modelMeanEsts$model == "jttp","ests"][[1]][[1]]
+  } else if (mod == "jtts"){
+    myeffects <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "sex", character = TRUE)
+    myeffects$initialValue[myeffects$include] <- modelMeanEsts[modelMeanEsts$model == "jtts","ests"][[1]][[1]]
+  } else if (mod == "simttb"){
+    myeffects <- includeEffects(SenBasic, "simXTransTrip", include = TRUE, type = "eval", interaction1 = "bills", character = TRUE)
+    myeffects$initialValue[myeffects$include] <- modelMeanEsts[modelMeanEsts$model == "simttb","ests"][[1]][[1]]
+  } else if (mod == "samettp"){   
[…]

### 2017-10-23 SenateLineupDataRefitModels/refitmodels.R

@@ -1,0 +1,168 @@
+# refit models to data for lineup experiment.
+library(RSiena)
+library(tidyverse)
+setwd("../SenateLineupDataRefitModels/")
+# senate data
+load("../Data/senate/se111adjmat.RDS")
+load("../Data/senate/se112adjmat.RDS")
+load("../Data/senate/se113adjmat.RDS")
+load("../Data/senate/se114adjmat.RDS")
+senateCovars <- read_csv('../Data/senate/senateCovars.csv')
+party <- coCovar(val = as.numeric(as.factor(senateCovars$party)))
+# party 1 is dem 2 is ind 3 is rep
+sex <- coCovar(val = as.numeric(as.factor(senateCovars$party)))
+# sex 1 is F 2 is M
+nbills_au <- senateCovars[,4:7]
+names(nbills_au) <- paste0("Sen",111:114) 
+nbills_au <- as.matrix(nbills_au)
+nbills_au[which(is.na(nbills_au))] <- 0
+rownames(nbills_au) <- senateCovars$source
+bills <- varCovar(nbills_au)
+
+# initial mean values to use as starting vals 
+load("../shinyApp/dat/allModelMeans.RDS")
+modelMeanEsts <- modelMeanEsts %>% add_row(model = "bigmod", ests = NA)
+modelMeanEsts$ests <- list(modelMeanEsts$ests[[1]], modelMeanEsts$ests[[2]],
+                           modelMeanEsts$ests[[3]], modelMeanEsts$ests[[4]], 
+                           modelMeanEsts$ests[[5]],modelMeanEsts$ests[[6]],
+                           c(2.4405048,2.4594403,2.2098176,-4.9232775,4.8916183,2.3743720,0.2047038,6.9661589))
+modelMeanEsts
+
+## necessary functions
+# take a df and make an adjmat
+make_adjmat <- function(d) {
+  d %>% right_join(df) %>% 
+    mutate(from = parse_number(from), to = parse_number(to)) %>% 
+    xtabs(!is.na(dep.var.id)~from+to, data =.) %>% as.matrix()
+}
+# take a lineup and make all data adjmats
+make_all_adjmats <- function(dat){
+  dat %>% nest(-sim) %>% mutate(adjmats = map(data, make_adjmat))
+}
+# take a wave and put it in senate siena data
+make_new_senate_siena <- function(wave2){
+  wave1 <- se111adj
+  wave3 <- se113adj
+  wave4 <- se114adj
[…]

### 2017-10-23 ASApaper/paper-child.Rnw
add most recent paper changes

@@
 
-Broadly, a SAOM takes network structure and node covariate information into account in two ways: first through a rate function that describes \textit{when} changes in the network occur, and second through an objective function that describes \textit{what} those changes are.
+Broadly, a SAOM takes network structure and node covariate information into account in two ways and models the network changes as a continuous time Markov chain (CTMC).
+First, the rate of change between states is dictated by a rate function that describes \textit{when} changes in the network occur, and secondy, the objective function describes \textit{what} those state changes are.
 As in many other network models, the variables of interest, $x_{ij}$, are the binary edges of the network.
@@
 \end{equation}
+
+This probability dictates which edge change is made by the acting node. The acting node can also choose to \textit{not} change at all, in which case $j \equiv i$.
+
+As advised by \citet{RSienamanual}, at least two parameters must be included in the objective function: the density and the reciprocity.
+We denote the density, or out-degree, parameter by $\beta_1$ and the associated statistic as $s_{i1}(x) = \sum_{j} x_{ij}$.
+Similarly, we denote the reciprocity parameter by $\beta_2$ and the associated statistic as $s_{i2}(x) = \sum_{j} x_{ij}x_{ji}$.
+We refer to the model with only these two parameters in the objective function as M1.
 
@@
 \begin{table}
-\scalebox{0.9}{
-\begin{tabular}{p{2cm}|p{2.1cm}|c|l|p{1.75cm}|p{2cm}}
-{\bf Effect name} & {\bf Interaction Variable} & {\bf Formula} & {\bf Picture} & {\bf Initial estimate} & {\bf Wald $p$-value} \\
+\scalebox{0.8}{
+\begin{tabular}{c|p{2cm}|p{2.1cm}|c|l|p{1.75cm}|p{2cm}}
+$\beta_k$ & {\bf Effect name} & {\bf Interaction Variable} & {\bf Formula} & {\bf Picture} & {\bf Initial estimate} & {\bf Wald $p$-value} \\
 \hline 
-jumping transitive triplet & party & $\sum_{j\neq h} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(p_i = p_h \neq p_j)$ & \includegraphics[width=.6in]{img/jttp.png} & -5.884 & $<0.0001$\\
-jumping transitive triplet & sex & $\sum_{j\neq h} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(s_i = s_h \neq s_j)$ & \includegraphics[width=.6in]{img/jtts.png}  & 3.335 & 0.0002 \\
-similarity transitive triplet & bills & $\sum_{j} x_{ij}x_{ih}x_{hj}\cdot (sim^b_{ij} - \overline{sim}^b)^*$ & \includegraphics[width=1in]{img/simttb.png} & 9.821 & 0.0128 \\
-same transtive triplet & party & $\sum_{j} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(p_i = p_j)$ & \includegraphics[width=1in]{img/samettp.png}  & 1.306 & 0.0642 \\
-same & party & $\sum_j x_{ij}\mathbb{I}(p_i = p_j)$ & \includegraphics[width=.6in]{img/samep.png}  & 0.363 & 0.0074
+$\beta_3$ & jumping transitive triplet & party & $s_{i3}(x, \mathbf{p}) = \sum_{j\neq h} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(p_i = p_h \neq p_j)$ & \includegraphics[width=.6in]{img/jttp.png} & -5.884 & $<0.0001$\\
+$\beta_4$ & jumping transitive triplet & sex & $s_{i4}(x, \mathbf{s}) = \sum_{j\neq h} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(s_i = s_h \neq s_j)$ & \includegraphics[width=.6in]{img/jtts.png}  & 3.335 & 0.0002 \\
+$\beta_5$ & similarity transitive triplet & bills & $s_{i5}(x, \mathbf{b}) = \sum_{j} x_{ij}x_{ih}x_{hj}\cdot (sim^b_{ij} - \overline{sim}^b)^*$ & \includegraphics[width=1in]{img/simttb.png} & 9.821 & 0.0128 \\
+$\beta_6$ & same transtive triplet & party & $s_{i6}(x, \mathbf{p}) =\sum_{j} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(p_i = p_j)$ & \includegraphics[width=1in]{img/samettp.png}  & 1.306 & 0.0642 %\\
+%$\beta_7$ & same & party & $s_{i7}(x, \mathbf{p}) = \sum_j x_{ij}\mathbb{I}(p_i = p_j)$ & \includegraphics[width=.6in]{img/samep.png}  & 0.363 & 0.0074
 \end{tabular}
@@
 
-\section{Significance Testing}
-
-We propose to attack some of the aforementioned difficulties with SAOMs through visual inference.
-
+We examine a total of six models, each identified by its objective function: 
 \begin{enumerate}
-\item Fit each model 1,000 times and save the parameter estimates, standard errors, and overall convergence criterion.
-\item From the models that converged, compute the mean of the 1,000 estimates of each parameter.
+\item Model M1: $f_{i}(x, \boldsymbol{\beta}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x)$
+\item Model M2: $f_{i}(x, \boldsymbol{\beta}, \mathbf{p}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_3 s_{i3}(x, \mathbf{p})$
+\item Model M3: $f_{i}(x, \boldsymbol{\beta}, \mathbf{s}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_4 s_{i4}(x, \mathbf{s})$
+\item Model M4: $f_{i}(x, \boldsymbol{\beta}, \mathbf{b}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_5 s_{i5}(x, \mathbf{b})$
+\item Model M5: $f_{i}(x, \boldsymbol{\beta}, \mathbf{p}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_6 s_{i6}(x, \mathbf{p})$
+\item Model M6: $f_{i}(x, \boldsymbol{\beta}, \mathbf{p}, \mathbf{b}, \mathbf{s}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_4 s_{i4}(x, \mathbf{s}) + \beta_5 s_{i5}(x, \mathbf{b}) + \beta_6 s_{i6}(x, \mathbf{p})$
 \end{enumerate}
 
-We want to determine at which point an effect becomes noticeable in a SAOM.
+\section{Significance Testing}\label{sec:sigtest}
+
+<<sigtestWald, eval = FALSE, echo = FALSE>>=
+SenBasic <- getEffects(senateSiena)
+Senjtt_p <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+myalg <- sienaAlgorithmCreate( projname = Sys.time() , n3 = 1000)
+fits <- siena07(myalg, data = senateSiena, effects = Senjtt_p, returnDeps = TRUE,
+                                      batch=TRUE, verbose = FALSE, silent = TRUE)
+thet <- fits$theta
+sig <- fits$covtheta
+Wald.RSiena(A = c(0,0,1), fits)
+th <- c(0,0,1) %*% thet
+covmat <- c(0,0,1) %*% sig %*% c(0,0,1)
+csq <- drop(th %*% solve(covmat) %*% th)
+1 - pchisq(csq, 1)
+@
+
+For a SAOM, there are two ways a significance test of the parameters can be performed.
+In \texttt{RSiena}, there are $t$-type tests and Wald-type test for a single parameter and for multiple parameters.
+The $t$-type test statistic is simply the parameter estimate divided by its standard error, and compared to a standard normal distribution.
+The Wald-type test statistic for a single parameter, $\beta_k$ is
+\begin{equation}\label{eq:wald1}
+\frac{(\hat{\beta_k})^2}{var(\hat{\beta_k})},
+\end{equation}
+which is compared to a Chi-square distribution with one degree of freedom \citep{RSienamanual}.
+Testing the significance of multiple parameters depends on the hypothesis we wish to test, and a $P \times K$ matrix, $A$, must be appropriately designed to test the $P$ hypotheses of interest.
+The null hypothesis is that $A\boldsymbol{\beta} = \mathbf{0}$, and the test statistic is
+\begin{equation}\label{eq:wald2}
+(A\boldsymbol{\hat{\beta}})' \hat{\Sigma}^{-1} A\boldsymbol{\hat{\beta}},
+\end{equation}
+where $\hat{\Sigma}$ is the estimated covariance matrix of $\boldsymbol{\beta}$. This statistic is then compared to a Chi-square distribution with $P$ degrees of freedom.
+
+We added each parameter to the basic model M1 to create models M2-M5, and then combined three of the four effects to create model M7.
+We fit these models in \texttt{RSiena} using Markov Chain Monte Carlo (MCMC) methods to approximate the method of moments estimates of the parameters.
+Because the estimation is done through MCMC simulation, we fit each model to the data 1,000 times to get a better estimate of the true value of $\boldsymbol{\beta}$.
+From the simulations that converged, which made up over 90\% of the fits for each model, we computed the mean of the 1,000 estimates of each parameter to get our values of $\hat{\boldsymbol{\beta}}$ from which to simulate data.
+
+Through visual inference, we want to determine at which point an effect becomes noticeable in a SAOM.
 By \textit{noticeable}, we mean that the inclusion of the effect alters the appearance of networks simulated from a model.
-An effect becomes noticeable when as simulation from a model with a noticeable effect appears different from simulations from a model where the same affect is absent.
+An effect becomes noticeable when a simulation from a model with a noticeable effect appears different from simulations from a model where the same affect is absent.
 Because the traditional way to visualize a network is a node-link diagram, we will examine the affect of different parameter values using node-link visualizations.
 
+For each lineup, we consider two models: model M1 is the null model, and another model, from M2-M5, which is chosen to be the alternative model.
+To construct a lineup, we simulate five networks from the null model and one network from the alternative model.
+We chose to expose our participants to only six plots at a time in order to show the node-link diagrams in more detail and to lower their cognitive load.
+Typically, lineup experiments are done with sets of 20 plots at a time c.f.~\citet{loy:2015, vanderplas:2016}, but we determined that not enough structure could be shown in each plot for 20 node-link diagrams\footnote{If you would like to explore the lineups we use in further detail, please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/}}.
+
+To determine the threshold at which the effect becomes noticeable, we examined six different levels of the effect, three negative and three positive.
+We hypothesized that the perception of the effect by our participants would look something like Figure~\ref{fig:hypothesis}.
+The higher in absolute value the parameter is, the more likely participants are to choose the alternative model out of the lineup.
+
+<<hypothesis, fig.height=2, fig.cap="We hypothesize that as the parameter value of interest increases in absolute value, more viewers of the lineup will pick the alternative data out of a lineup.">>=
+x <- seq(-10, 10, .05)
+N <- length(x)
+qplot(x = 1:N, y = pt(x, 1), geom = 'line') + geom_line(aes(x = -(1:N), pt(x,1))) + 
+    scale_x_continuous(name = "Parameter Value", labels = c("very negative", "negative", "0", "positive", "very positive")) + 
+    scale_y_continuous(name = "% detecting the plot from\nthe alternative model", labels = paste0(c(0,25,50,75,100), "%"))
+@
+
 \section{Goodness-of-Fit Testing}
+
+Goodness-of-fit testing for network models is notoriously difficult.
+Most network models, other than the most simple, lack the asymptotics required to develop the goodness-of-fit methods required \citep{goldenberg09}.
+Some methods have been developed based on what \citeauthor{RSienamanual} call ``auxiliary statistics" such as the indegree or outdegree distribution on the nodes.
+In \texttt{RSiena}, the \texttt{sienaGOF} function performs goodness-of-fit testing as follows:
+\begin{enumerate}
+\item Auxiliary statistics are computed on the observed data ($\mathbf{u}_d$) and on $N$ simulated observations from the model ($\mathbf{u}_1 \dots \mathbf{u}_N$). (Usually, $N=1000$) 
+\item The mean vector, $\overline{\mathbf{u}}$ and covariance matrix, $\mathbf{S}$ of the statistics on the simulations from the model are computed, and the Mahalanobis distance, $d_M(\mathbf{u})$ from the observed statistics to the distribution of the simulated statistics is computed:
+\begin{equation}\label{eq:mahal}
+d_M(\mathbf{u}) = \sqrt{(\mathbf{u} - \overline{\mathbf{u}})' S^{-1} (\mathbf{u} - \overline{\mathbf{u}})}
+\end{equation}
+\item The Mahalanobis distance for each of the $N$ simulations is calculated and $d_M(\mathbf{u}_d)$ is compared to this distribution of distances.
+\item An empirical $p$-value is found by computing the proportion of simulated distances found in step 4 that are as large or larger than $d_M(\mathbf{u}_d)$.
+A SAOM is thus considered a good fit to the data if $p$ is large.
+A plot comparing the data to the simulations is also considered, and a similar plot is shown in Figure~\ref{fig:gofsiena}
+\end{enumerate}
+
+<<gofsiena, fig.height=3, fig.cap="An example of what a goodness-of-fit plot from \\texttt{RSiena} looks like.
+The overlaid boxplots and violin plots show the distribution of each of the outdegree values on the simulated networks, and the red points and lines are the observed data values.", cache = TRUE>>=
+load("data/ansnullpaper.rda")
+library(lattice)
+gof1 <- sienaGOF(ansnull, OutdegreeDistribution, varName ="friendship", join = FALSE)
+sims1 <- data.frame(gof1$`Period 1`$Simulations)
+dat1 <- gof1$`Period 1`$Observations
+dat1 %>% data.frame() %>% gather(outdegree, val) %>%
+  mutate(outdegree = parse_number(outdegree)-1) -> dat1
+p1 <- sims1 %>%
+  mutate(sim = row_number()) %>%
+  gather(outdegree, val, X1:X9) %>%
+  mutate(outdegree = parse_number(outdegree) - 1) %>%
+  filter(outdegree <= 6) %>%
+  ggplot(aes(x = outdegree, y = val)) + 
+  geom_boxplot(aes(group = outdegree), size = .5, outlier.shape = "x", outlier.size = 5) + 
+  geom_violin(aes(group = outdegree), bw = 2, fill = NA) + 
+  geom_point(data = dat1, color = 'red') + 
+  geom_line(data= dat1, color = 'red') + 
+  geom_text(data = dat1, aes(label = val), hjust = -.5) + 
+  labs(x = "Outdegree (p = 0.154)", y = "Statistic", 
+       title = "Goodness-of-Fit: Outdegree distribution period 1") + 
+  ThemeNoNet
+p1
+@
+
+The \texttt{RSiena} software also provides a Rao score-type test for goodness-of-fit for assessing one or more parameters, the test statistic of which is compared to a Chi-square distribution with $P$ degrees of freedom, where $P$ has the same definition as in Section~\ref{sec:sigtest}.
+For full detail on the score-type test, see \citet{scoretest}.
+
+These methods are similar in that they are both restriced: the \texttt{sienaGOF} method only considers one measure on the data and simulations from the model, while the score-type tests only consider subsets of parameters, ``nuisance parameters" in \citet{scoretest}, not the entire set of parameters.
+By using visual inference instead of more traditional statistical methods, we hope to perform a more holistic goodness-of-fit test.
+
+Using the lineup protocol, we show Amazon Mechanical Turk workers the data once, in a lineup with five other plots of simulated data from one of the models we chose.
 

### 2017-10-24 ASApaper/paper-child.Rnw
data and updates for paper. forgot to push last night/this morning

@@
 library(tidyverse)
+library(xtable)
 library(RSienaTest)
@@
 library(GGally)
+library(vinference)
 library(extrafont)
@@
 \section{Introduction}
-\st{
-\begin{enumerate}
-\item significance testing: what is it and why is it important
-\item goodness-of-fit testing: what is it and why is it important
-\item difficulty of these two for sna, why do vis inf
-\end{enumerate}
-}
+% \st{
+% \begin{enumerate}
+% \item significance testing: what is it and why is it important
+% \item goodness-of-fit testing: what is it and why is it important
+% \item difficulty of these two for sna, why do vis inf
+% \end{enumerate}
+% }
 Two of the most imporant pieces of statistical modeling are significance testing of model parameters and goodness-of-fit measures and tests.
@@
 Network objects are particularly difficult to model due to the dependencies inherent to the data.
+This complexity can increases when, instead of looking at one instance of a network, we consider the dynamics of network change in between those singular observed instances.
+This type of modelling for dynamic networks is often performed on social network data, such as for friendship networks among students or the spread of HIV in communities of men sleeping with other men.
+These models lack the asymptotics required to perform many well-known goodness-of-fit tests, and the maximum likelihood estimation of parameters is so difficult that it can make significance testing significant as well \citep{goldenberg09}.
 
-XX smart stuff with more refs XX
+We propose new methods for significance and goodness-of-fit testing for a set of social network models, stochastic actor-oriented models for dynamic network data \citep{saompaper}.
+These new methods use \textit{visual inference} in place of traditional statistical methods for social network models, such as Wald tests for significance of parameters and in- and outdegree distribution metrics for determining goodness-of-fit.
+By using the visual inference methodology of \citet{Bujaetal}, we can look at the entire dataset simulated from a network model as opposed to a single metric on the network such as outdegree, or a $p$-value for a single parameter in the model.
 
@@
 The higher in absolute value the parameter is, the more likely participants are to choose the alternative model out of the lineup.
+To determine the six levels of the effects we want to test, started with the estimates of the parameter at hand as determined by the procedure described in Section~\ref{sec:models}, and multiply it or changing its sign to determine what \textit{we} notice about the effect of the parameter in simulations from the changed models.
 
@@
     scale_x_continuous(name = "Parameter Value", labels = c("very negative", "negative", "0", "positive", "very positive")) + 
-    scale_y_continuous(name = "% detecting the plot from\nthe alternative model", labels = paste0(c(0,25,50,75,100), "%"))
+    scale_y_continuous(name = "% detecting plot from\nthe alternative model", labels = paste0(c(0,25,50,75,100), "%"))
+@
+
+We showed participants from Amazon Mechanical Turk a series of lineups containing data simulated from M1 through M5.
+In model M1, with only two parameters in the objective function, we varied both the density and reciprocity parameteter values one at a time.
+In models M2 through M5, we vary the additional parameter, $\beta_3$ through $\beta_6$.
+In the case of both $\beta_4$ and $\beta_6$, we could not determine a value for negative effects that made the data simulated from M3 and M5 look different than null model simulations from model M1.
+Because we could not detect these effects at these values, we decided that the less experienced Turkers would also not be able to.
+So instead of testing the negative values of these effects, we examined a different scenario: we placed 5 simulations from positive values of the parameter with one simulation from model M1 in a lineup.
+We did this to determine if the perception of the effect size would be symmetric: if an effect is noticed $x$\% of the time at value $\beta = 5$ when a simulation from the model is placed among null plots, then when five simulations from the model are put together with one simulation from model M1, the plot from the simpler model should be noticed about $x$\% of the time as well.
+A summary of the results from our experiment is shown in Figure~\ref{fig:readturk22}.
+On the $x$ axis, we plot the value of the parameter of interest, and on the $y$ axis, the proportion of times the plot of data simulated from the alternative model was picked out.
+The results are split into groups based on the sign of the parameter.
+We can see clear patterns in all of the groups: as the parameter value approaches 0, fewer participants identified the alternamtive plot.
+
+<<readturk22, fig.height=6, fig.cap="A summary of some results from the Amazon mechanical turk study.
+On the $x$-axis, the value of the parameter changed in the model of interest, and on the $y$-axis, the proprotion of times turkers identified the different data plot.
+The pattern as the parameter value approaches zero is clear in all conditions.">>=
+turk22 <- read_csv("data/turk22-sig.csv")
+turk22 %>% mutate(type2 = ifelse(sign == 0, sign(initialEst), sign),
+                       type2 = ifelse(type == "one", type2, -1)) %>% 
+  group_by(pic_id, size, test_param, sign, type2,alt_model) %>% 
+  summarize(datapick = mean(datapick)) %>%
+  filter(pic_id != 3132) %>% 
+  ggplot(aes(x = size, y = datapick, colour=factor(type2))) + 
+  geom_point() +
+  scale_color_brewer(palette = "Dark2", name = "Parameter Code") + 
+  facet_wrap(~test_param, scales="free", labeller = 'label_both') +
+  geom_smooth(se=FALSE, method="lm") + theme(legend.position = 'bottom')
 @
@@
 
-Using the lineup protocol, we show Amazon Mechanical Turk workers the data once, in a lineup with five other plots of simulated data from one of the models we chose.
+Using the lineup protocol, we show each Amazon Mechanical Turk worker the data once, in a lineup with five other plots of simulated data from one of the models we chose.
+We examined four different models, M2, M3, M4, and M6, and examined three repititons of each, for a total of 12 goodness-of-fit lineups.
+In each lineup, the ``null model" is one of the four models and the ``alternative" model is the true, unknown model that generated the senate network data The hypotheses for our goodness-of-fit tests are:
+\begin{description}
+\item $H_0$:  The senate network data come from the null model.
+\item $H_A$: The senate network data do not come from the null model. 
+\end{description}
+If a lineup viewer picks out the data among the five simulations from the null model, that is evidence in favor of the alternative hypothesis.
+On the contrary, if the lineup viewer picks one of the null plots, that is evidence against the alternative hypothsis.
+Because the size of the lineups is small, the probability of picking the data by chance is high, $\frac{1}{6}$, but if \textit{many} independent viewers pick out the data from the nulls, the evidence in favor of the alternative hypothesis becomes stronger.
+Results from our MTurk goodness-of-fit plots are provided in Table~\ref{tab:gofstats}.
+
+<<goftab, cache = TRUE, results='asis'>>=
+turk22_gof <- read_csv("data/turk22-gof.csv")
+turk22_gof_stats <- turk22_gof %>% group_by(pic_id) %>% 
+  summarize(npickdata = sum(datapick), total = n(), 
+          pvinf = map2_dbl(npickdata, total, pV, m = 6, scenario = 3))
+turk22_gof_stats %>% separate(pic_id, into = c("discard", "model", "rep"), sep = c(2,3)) -> print_gof_stats
+print_gof_stats %>% select(-discard) %>% mutate(model = paste0("M", model))-> print_gof_stats
+names(print_gof_stats) <- c("Model", "Replicate", "Data Picks", "Total Viewers", "p-value")
+print_gof_stats$`p-value` <- ifelse(print_gof_stats$`p-value` < .0001, "< 0.0001", as.character(round(print_gof_stats$`p-value`,4)))
+print(xtable::xtable(print_gof_stats, label = "tab:gofstats", caption = "An overview of the results from our 12 goodness-of-fit lineup tests.", digits = 4), caption.placement = "bottom", include.rownames = FALSE)
+@
+
+The $p$-values were calcluated using the \texttt{vinference} package by \citet{vinference}.
+This package contains methods to calculate \textit{Visual distributions} for lineup experiment data.
+The distribution depends on the number of evaluations of a plot, $K$, the size of the lineup, $m$, and the lineup scenario, which here is that each lineup containing the same data and the same set of null plots is shown to $K$ independent observers.
+Using these $p$-values, all but one lineup results in a rejection of the null hypothesis at Type-I error rate of $\alpha = 0.05$.
+The lineup that resulted in a failure to reject the null hypothesis is shown in Figure~\ref{fig:failgof}.
+The null model in this lineup is M4, and the senate data is shown in panel number $3^2 - 7$, but the panel most chose was number four, and the most common reasoning for that choice was that it had the most simple structure.
+
+<<failgof, fig.height=3, fig.cap="The goodness-of-fit lineup that failed to reject the null hypothesis.
+The null model for this lineup is M4.
+Only 7 of 21 viewers of this lineup selected the data plot as the most different from the others.">>=
+load("data/se112adjmat.RDS")
+wave2 <- data.frame(se112adj)
+senators <- data.frame(id = colnames(wave2), number = 1:155)
+wave2$from <- colnames(wave2)
+wave2 <- wave2 %>% gather(to, val, Alan.Stuart.Franken:William.Cowan) %>% filter(val > 0)
+wave2 <- merge(wave2, senators, by.x = "from", by.y = "id", all = T)
+wave2$from <- senators$number[match(wave2$from, senators$id)]
+wave2$to <- senators$number[match(wave2$to, senators$id)]
+wave2$from <- paste0("V", wave2$from)
+wave2$to <- ifelse(is.na(wave2$to), NA, paste0("V", wave2$to))
+wave2 <- wave2 %>% mutate(sim = 1001, model = "data", wave = 1) %>% select(-c(val, number))
+dat <- read_csv(paste0("data/jtts_gof_9_2.csv"))
+datplot <- unique(dat$ord[dat$sim == 1001])
+dat <- dat %>% filter(sim != 1001) %>% 
+      bind_rows(wave2) %>% 
+      mutate(ord = ifelse(is.na(ord), datplot, ord))
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
+        theme_net() + 
+        theme(panel.background = element_rect(color = 'black')) +
+        facet_wrap(~ord)
+@
+
+The smallest $p$-value for one of the goodness-of-fit lineups was for the third replicate of the null model M4.
+This contradicts our previous finding that the only lineup to fail to reject the null was also when the null model was M4.
+The plot of this lineup is shown in Figure~\ref{fig:mostsiggof}.
+In the remaining replicate of M4 as the null model, 13 of 16 viewers identified the data plot, corresponding to a $p$-values of less than $0.0001$, just like the third replicate.
+This leads us to believe that the high $p$-value observed in the second replicate is more of a fluke.
+%This demonstrates some of the problems inherent with network models, such as model degeneracy, as discussed in \citet{degeneracy}.
+Network models
+Furthermore, for the first lineup replicate with model M5 as the null model, only 9 of 21 participants were able to pick out the data, but the Visual distribution $p$-value is less than 0.05.
+This is much higher than the corresponding Binomial probability of $Pr(X = 9|n = 21, p = \frac{1}{6}) = \binom{21}{9}(\frac{1}{6})^9(\frac{5}{6})^{21-9} \approx 0.003$, but we not sure that having fewer than half of our participants identify the data plot is a truly significant result.
+We could move the threshold from $\alpha = 0.05$ to $\alpha = 0.01$ or $\alpha = 0.001$, but more work needs to be done to determine which, if any of these, is appropriate.
+
+<<mostsiggof, fig.height = 3, fig.cap="The lineup resulting in the smallest $p$-value rejecting the null hypothesis. Surprisingly, this another repetition for M4 as the null model.">>=
+dat <- read_csv(paste0("data/jtts_gof_9_3.csv"))
+datplot <- unique(dat$ord[dat$sim == 1001])
+dat <- dat %>% filter(sim != 1001) %>% 
+      bind_rows(wave2) %>% 
+      mutate(ord = ifelse(is.na(ord), datplot, ord))
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
+        theme_net() + 
+        theme(panel.background = element_rect(color = 'black')) +
+        facet_wrap(~ord)
+@
+
+We believe this goodness-of-fit testing method holds promise for the future of social network analysis.
+The participants in our experiments are very good overall at picking out the data when it is noticeably different from the null plots in the lineups.
+In addition, as in replicate three for null model M4, when the null plots contain similarly sized structures as the data plot, our participants have a hard time distinguishing the data.
+We believe that running these tests multiple times using several different sets of null models to adequately explore the possible structures generated by the models is a step in the right direction for a more comprehensive goodness-of-fit test for network models.
 

### 2017-10-24 ASApaper/paper.Rnw
data and updates for paper. forgot to push last night/this morning

@@
 \begin{abstract}
-Abstract goes here.
+Two of the most imporant pieces of statistical modeling are significance testing of model parameters and goodness-of-fit measures and tests.
+The more complicated the model, the harder it is to determine fit or whether to include or exclude a parameter.
+Some particularly complicated sets of models are those designed to model network change.
+By using the visual inference methodology of \citet{Bujaetal}, we can look at the entire dataset simulated from a network model as opposed to a single metric on the network such as outdegree, or a $p$-value for a single parameter in the model.
 \end{abstract}

### 2017-10-24 ASApaper/code/modeling_data.R

@@ -1,0 +1,61 @@
+# modeling data
+
+turkbf2_sig %>% mutate(type2 = ifelse(sign == 0, sign(initialEst), sign),
+                       type2 = ifelse(type == "one", type2, -1)) %>% 
+  group_by(pic_id, size, test_param, sign, type2,alt_model) %>% summarize(
+    datapick = mean(datapick)
+  ) %>%
+  filter(pic_id != 3132) %>% 
+  ggplot(aes(x = size, y = datapick, colour=factor(type2))) + 
+  geom_point() +
+  facet_grid(~test_param, scales="free", labeller = 'label_both') +
+  geom_smooth(se=FALSE, method="lm")
+
+modelData_sig <- turkbf2_sig %>% mutate(type2 = ifelse(sign == 0, sign(initialEst), sign),
+                                        type2 = ifelse(type == "one", type2, -1)) %>% filter(pic_id != 3132) 
+
+model1 <- glmer(datapick ~ -1 + test_param:size:as.factor(type2) + (1|pic_id) + (1|nick_name), 
+                data = modelData_sig, family = 'binomial')
+summary(model1)
+
+library(vinference)
+
+pV(x = 16, K = 30, m = 6, scenario = 3)
+
+vinfturk <- turkfb2 %>% group_by(pic_id) %>% summarise(npick = sum(datapick), tot = n(), 
+                                                       pv = map2_dbl(npick, tot, pV, m = 6, scenario = 3))
+
+
+# is difficulty balanced?
+ggplot(data = turkfb2) + 
+  geom_bar(aes(x = difficulty, fill = datapick), position = 'fill')
+
+turkfb2 %>% group_by(pic_id, size, test_param, sign, type,alt_model) %>% summarize(
+  datapick = mean(datapick)
+) %>% ggplot(aes(x = size, y = datapick, colour=factor(sign), shape=type)) + 
+  geom_point() +
+  facet_grid(alt_model~test_param, scales="free", labeller = 'label_both') +
+  geom_smooth(se=FALSE, method="lm")
+
+library(lme4)
+
+tryglmm <- glmer(formula = as.numeric(datapick) ~ test_param + size:test_param:sign:type + (1|pic_id) + (1|nick_name), family = binomial, data = turkfb2 %>% filter(alt_model != "data"))
+summary(tryglmm)
+
+# separate out type 
+control <- glmerControl(optimizer = 'bobyqa')
[…]

### 2017-10-24 ASApaper/code/raw_to_data.R

@@ -1,0 +1,98 @@
+# looking at experiment data
+library(tidyverse)
+library(RSQLite)
+
+con <- dbConnect(SQLite(), "~/Desktop/turk.db")
+#dbListTables(con)
+
+fb <- dbReadTable(con, "feedback")
+#head(fb)
+
+fb <- filter(fb, description == "turk22")
+#head(fb)
+#fb$nick_name %>% unique %>% sort
+#c("Amy Crawford", "hh", "sam", "fdgdfg", "Nick Berry", "sdfds")
+#fb %>% filter(nick_name == "campaign_2165")
+#fb %>% filter(nick_name == "ivasanthi@gmail.com")
+
+turkfb <- filter(fb, !(nick_name %in% c("Amy Crawford", "hh", "sam", "fdgdfg", "Nick Berry", "sdfds")))
+
+turkfb <- turkfb %>% mutate(response = map(response_no, 
+                                 .f = function(x){parse_number(unlist(str_split(x, pattern=",")))}))
+
+turkfb <- turkfb %>% mutate(weight = map_int(response, length)) 
+
+turkfb <- turkfb %>% mutate(weight = 1/weight) 
+
+turkfb <- turkfb %>% unnest(response)
+
+pic_details <- read_csv("data/picture_details.csv")
+
+#head(pic_details)
+
+#pic_details <- filter(pic_details, experiment == "turk22")
+
+turkfb2 <- left_join(turkfb, pic_details, by = "pic_id")
+
+turkfb2 <- turkfb2 %>% mutate(datapick = response == obs_plot_location)
+#summary(turkfb2$datapick)
+
+# turkfb2 %>% group_by(test_param, param_value, pic_id) %>% 
+#   summarize(datamean = mean(datapick, na.rm=T), n = n()) %>% 
+#   ungroup() %>% 
+#   mutate(param_value = as.numeric(param_value)) %>% 
+#   filter(test_param == "recip") %>% 
+#   ggplot(aes(x = param_value, y = datamean)) + 
+#   geom_point() + 
+#   xlim(c(0, .01)) + 
+#   scale_x_log10()
+#   #facet_wrap(~test_param, scales = 'free')
+  
+turkfb2 <- filter(turkfb2, !is.na(pic_id))
+#write_csv(turkfb2, "data/turk22.csv")
+
+
+# get parameter values of lineups 
+load("data/allModelMeans.RDS")
[…]

### 2017-10-25 ASApaper/paper-child.Rnw
add things left behind

@@
 
-<<setup, echo = FALSE, message = FALSE, warning = FALSE>>=
+<<setup-ch-2, echo = FALSE, message = FALSE, warning = FALSE>>=
 options(replace.assign=TRUE,width=70, digits=3)
@@
 library(vinference)
+library(lme4)
 library(extrafont)
@@
 The network statistics range from the simple outdegree, $s_i(x) = \sum_{i\neq j} x_{ij}$, to the more complicated transitive triplets jumping to different covariate, $s_i(x, \mathbf{Z}) = \sum_{i \neq j \neq h} x_{ij}x_{ih}x_{hj} \cdot \mathbb{I}(z_i = z_h \neq z_j)$, plus many more.
-At last count, in the software we use to fit these models to network data, \texttt{RSiena}, there are over 80 possible effects that can be included in the objective function \citep{RSiena, RSienamanual}.
+At last count, in the software we use to fit these models to network data, \texttt{RSiena}, there are over 80 possible effects that can be included in the objective function \citep{RSiena}.
 We discuss these statistics in more detail in Section~\ref{sec:models}.
@@
 
-As advised by \citet{RSienamanual}, at least two parameters must be included in the objective function: the density and the reciprocity.
+According to Snijders et al in the \texttt{RSiena} manual, at least two parameters must be included in the objective function: the density and the reciprocity.
 We denote the density, or out-degree, parameter by $\beta_1$ and the associated statistic as $s_{i1}(x) = \sum_{j} x_{ij}$.
@@
 
-\subsection{Example Data}
+\subsection{Example Data}\label{sec:senatedata}
 
@@
 \end{equation}
-which is compared to a Chi-square distribution with one degree of freedom \citep{RSienamanual}.
+which is compared to a Chi-square distribution with one degree of freedom.
 Testing the significance of multiple parameters depends on the hypothesis we wish to test, and a $P \times K$ matrix, $A$, must be appropriately designed to test the $P$ hypotheses of interest.
@@
 In models M2 through M5, we vary the additional parameter, $\beta_3$ through $\beta_6$.
+Thus, we have six different parameters of interest to us: $\beta_1, \dots, \beta_6$.
+We want to determine how the size of these parameters affects the overall structure of the network data simulated from the models M1 through M5, so we also vary the value of the parameters in two directions: positive and negative.
+
+In order to determine which specific values of the parameters we wanted to examine, we constructed an online application that created the lineup protocol for us to be the guinea pigs of our own experiment \citep{shiny}.
+A screen grab of the app is shown in Figure~\ref{fig:shiny}.
+On the left side of the screen, the user\footnote{Please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/} to create lineups for yourself.} can input the information necessary for creating a lineup of the models M1 through M6 for the data in Section~\ref{sec:senatedata}: first, choose to simulate only one plot from the specified model (analogous to changing the alternative model in the lineup protocol) or to simulated $M-1$ plots from the specified model (analogous to changing the null model in the lineup protocol); the model of interest; the wave of the data to examine; if model M1 is selected, whether to alter the density or the reciprocity parameter; the size of the lineup; the amount by which to multiply the effect selected; a random seed for replicability; and a layout algorithm to use for the node-link diagrams.
+The plot(s) that appear(s) that are \textit{not} from the model specified using the other options are simulations from model M1 with the estimates of the rate parameters, $\beta_1$ and $\beta_2$ as calculated by the procedure described in Section~\ref{sec:models}.
+
+\begin{figure}
+\includegraphics[width=.9\textwidth]{img/shinyappscreen.png}
+\caption{\label{fig:shiny} A screen grab of the web application we created to design our lineup experiment.
+In this view, model M4 is selected and one estimate from model M4 with $\beta_5$ set to twice its estimated value given in Table~\ref{tab:effects} is placed at random among five observations simulated from model M1.
+The other two tabs in the right side of the screen tell you which plot comes from the alternative model, and allow you to view and download the data that is plotted in the lineup.}
+\end{figure}
+
+Using the ``Picking Lineups" web application, we settled on six parameter values to test for each of our six effects, $\beta_1, \dots \beta_6$.
+The complete details of the parameters tested using the lineup protocol is given in Table~\ref{tab:experdetail}.
 In the case of both $\beta_4$ and $\beta_6$, we could not determine a value for negative effects that made the data simulated from M3 and M5 look different than null model simulations from model M1.
@@
 We did this to determine if the perception of the effect size would be symmetric: if an effect is noticed $x$\% of the time at value $\beta = 5$ when a simulation from the model is placed among null plots, then when five simulations from the model are put together with one simulation from model M1, the plot from the simpler model should be noticed about $x$\% of the time as well.
+
+<<getexperdetail, results='asis'>>=
+turk22 <- read_csv("data/turk22-sig.csv")
+modelData_sig <- turk22 %>% 
+  mutate(type2 = ifelse(sign == 0, sign(initialEst), sign),
+         type2 = ifelse(type == "one", type2, -1)) %>% 
+  filter(pic_id != 3132) 
+modelData_sig %>% group_by(test_param, type2, difficulty, size) %>%
+  count() %>% ungroup() %>%
+  spread(difficulty, size) -> step1
+step1 %>% select(1:4) %>% filter(!is.na(`1`)) -> step2a
+step1 %>% select(c(1:3,5)) %>% filter(!is.na(`2`)) -> step2b
+step1 %>% select(c(1:3,6)) %>% filter(!is.na(`3`)) -> step2c
+experdet <- step2a %>% mutate(easy = `1`, medium = step2b$`2`, hard = step2c$`3`) %>% select(-c(`1`, n))
+experdet$test_param <- as.factor(experdet$test_param)
+levels(experdet$test_param) <- paste0("beta", c(1,3,4,2,6,5))
+experdet$test_param <- as.character(experdet$test_param)
+names(experdet) <- c("Parameter", "Lineup Type", "Easy Value", "Medium Value", "Hard Value")
+
+print(xtable::xtable(experdet, caption = "All conditions in our MTurk experiment.
+Note that for $\\beta_4$ and $\\beta_6$, the negative type experiments are the reverse lineups, where 5 plots were simulated from the model with the parameter value given in the table, while 1 plot was simulated from M1.", label = "tab:experdetail", digits = c(0,0,0,3,4,3)))
+@
+
 A summary of the results from our experiment is shown in Figure~\ref{fig:readturk22}.
@@
 The results are split into groups based on the sign of the parameter.
-We can see clear patterns in all of the groups: as the parameter value approaches 0, fewer participants identified the alternamtive plot.
+We can see clear patterns in all of the groups: as the parameter value approaches 0, fewer participants identified the alternative plot.
 
@@
 On the $x$-axis, the value of the parameter changed in the model of interest, and on the $y$-axis, the proprotion of times turkers identified the different data plot.
+A simple linear regression is fit to each group.
 The pattern as the parameter value approaches zero is clear in all conditions.">>=
-turk22 <- read_csv("data/turk22-sig.csv")
 turk22 %>% mutate(type2 = ifelse(sign == 0, sign(initialEst), sign),
@@
   geom_point() +
-  scale_color_brewer(palette = "Dark2", name = "Parameter Code") + 
+  scale_color_brewer(palette = "Dark2", name = "Lineup Type") + 
   facet_wrap(~test_param, scales="free", labeller = 'label_both') +
-  geom_smooth(se=FALSE, method="lm") + theme(legend.position = 'bottom')
+  geom_smooth(se=FALSE, method="lm") + 
+  ThemeNoNet + 
+  theme(legend.position = 'bottom')
+@
+
+We futher explore this relationship between identification of the alternative data in the lineup and the parameter, effect size, and lineuptype with a generalized linear mixed model.
+The response variable, $Y_{ijk\ell m}$, is binary, indicating whether participant $m$ picked the alternative data plot in lineup $\ell$.
+The covariate $x_{\ell 1}$ takes on the value -1 or 1 according to the lineup type code in Table~\ref{tab:experdetail}.
+In most cases, -1 indicates the parameter value is less than original estimate, while 1 indicates the parameter value is at or above the original estimate.
+The continuous covariate $x_{\ell 2}$ is the centered and scaled size of the effect of interest from which the alternative data were simulated, the values of which are labeled ``easy", ``medium", and ``hard" in Table~\ref{tab:experdetail} according to how difficult we thought the Turk participants would find each lineup.
+In Equation~\ref{eq:glmm}, $i \in \{1, 2, 3, 4, 5\}$ corresponding to the effects $\beta_3, \beta_4, \beta_2, \beta_6, \beta_5$, respectively.
+We include random effects in the model for each lineup, $\delta_{\ell}$ and each participant, $\epsilon_m$, and fit a hierarchical model as follows:
+\begin{align}\label{eq:glmm}
+Y_{i\ell m} &\sim \text{Bernoulli}(\pi_{i\ell m}) \\
+\text{logit}(\pi_{i\ell m}) & = \mu + \alpha_{i} + \theta\mathbb{I}(x_{\ell 1} = 1) + \gamma x_{\ell 2} + (\alpha\theta)_i \mathbb{I}(x_{\ell 1} = 1) + (\alpha\gamma)_i x_{\ell 2} + (\theta\gamma)\mathbb{I}(x_{\ell 1} = 1)x_{\ell 2} + (\alpha\theta\gamma)_i\mathbb{I}(x_{\ell 1} = 1) x_{\ell 2} + \delta_{\ell} + \epsilon_{m} \\
+\delta_{\ell} & \sim N(0, \sigma^2_{\delta}) \\
+\epsilon_{m} & \sim N(0, \sigma^2_{\epsilon}) \\
+\end{align}
+The results of fitting this model using \texttt{glmer} from the \texttt{lme4} package are summarized in Table~\ref{tab:glmmests} \citep{lme4}.
+
+<<glmmres>>=
+load("data/finalglmm.RDA")
+mod <- model0randomscalesize
+res <- summary(mod)[[10]]
+ests <- as.numeric(round(res[,1], 4))
+odds <- round(exp(ests), 4)
+se <- as.numeric(round(res[,2], 4))
+pval <- as.numeric(round(res[,4],4))
+@
+
+\begin{table}
+\centering
+\begin{tabular}{lcccc}
+Parameter & Estimate & Std Error & $p$-value & Odds Multiplier \\
+$\mu$ & \Sexpr{ests[1]} & \Sexpr{se[1]} & \Sexpr{pval[1]} & \Sexpr{odds[1]}\\
+$\alpha_1$ & \Sexpr{ests[2]} & \Sexpr{se[2]} & \Sexpr{pval[2]} & \Sexpr{odds[2]} \\
+$\alpha_2$ & \Sexpr{ests[3]} & \Sexpr{se[3]} & \Sexpr{pval[3]} & \Sexpr{odds[3]}  \\
+$\alpha_3$  & \Sexpr{ests[4]} & \Sexpr{se[4]} & \Sexpr{pval[4]} & \Sexpr{odds[4]} \\
+$\alpha_4$ & \Sexpr{ests[5]} & \Sexpr{se[5]} & \Sexpr{pval[5]} & \Sexpr{odds[5]}  \\
+$\alpha_5$  & \Sexpr{ests[6]} & \Sexpr{se[6]} & \Sexpr{pval[6]} & \Sexpr{odds[6]} \\
+$\theta$  & \Sexpr{ests[7]} & \Sexpr{se[7]} & \Sexpr{pval[7]} & \Sexpr{odds[7]} \\
+$\gamma$  & \Sexpr{ests[8]} & \Sexpr{se[8]} & \Sexpr{pval[8]} & \Sexpr{odds[8]} \\ 
+$(\alpha\theta)_1$  & \Sexpr{ests[9]} & \Sexpr{se[9]} & \Sexpr{pval[9]} & \Sexpr{odds[9]} \\
+$(\alpha\theta)_2$  & \Sexpr{ests[10]} & \Sexpr{se[10]} & \Sexpr{pval[10]} & \Sexpr{odds[10]} \\
+$(\alpha\theta)_3$  & \Sexpr{ests[11]} & \Sexpr{se[11]} & \Sexpr{pval[11]} & \Sexpr{odds[11]} \\
+$(\alpha\theta)_4$  & \Sexpr{ests[12]} & \Sexpr{se[12]} & \Sexpr{pval[12]} & \Sexpr{odds[12]} \\
+$(\alpha\theta)_5$  & \Sexpr{ests[13]} & \Sexpr{se[13]} & \Sexpr{pval[13]} & \Sexpr{odds[13]} \\
+$(\alpha\gamma)_1$  & \Sexpr{ests[14]} & \Sexpr{se[14]} & \Sexpr{pval[14]} & \Sexpr{odds[14]} \\
+$(\alpha\gamma)_2$  & \Sexpr{ests[15]} & \Sexpr{se[15]} & \Sexpr{pval[15]} & \Sexpr{odds[15]} \\
+$(\alpha\gamma)_3$  & \Sexpr{ests[16]} & \Sexpr{se[16]} & \Sexpr{pval[16]} & \Sexpr{odds[16]} \\
+$(\alpha\gamma)_4$  & \Sexpr{ests[17]} & \Sexpr{se[17]} & \Sexpr{pval[17]} & \Sexpr{odds[17]} \\
+$(\alpha\gamma)_5$  & \Sexpr{ests[18]} & \Sexpr{se[18]} & \Sexpr{pval[18]} & \Sexpr{odds[18]} \\
+$\theta\gamma$  & \Sexpr{ests[19]} & \Sexpr{se[19]} & \Sexpr{pval[19]} & \Sexpr{odds[19]} \\
+$(\alpha\theta\gamma)_1$ & \Sexpr{ests[20]} & \Sexpr{se[20]} & \Sexpr{pval[20]} & \Sexpr{odds[20]}  \\
+$(\alpha\theta\gamma)_2$  & \Sexpr{ests[21]} & \Sexpr{se[21]} & \Sexpr{pval[21]} & \Sexpr{odds[21]} \\
+$(\alpha\theta\gamma)_3$  & \Sexpr{ests[22]} & \Sexpr{se[22]} & \Sexpr{pval[22]} & \Sexpr{odds[22]} \\
+$(\alpha\theta\gamma)_4$  & \Sexpr{ests[23]} & \Sexpr{se[23]} & \Sexpr{pval[23]} & \Sexpr{odds[23]} \\
+$(\alpha\theta\gamma)_5$  & \Sexpr{ests[24]} & \Sexpr{se[24]} & \Sexpr{pval[24]} & \Sexpr{odds[24]} \\ 
+$\sigma^2_{\delta}$ & 0.5638 \\
+$\sigma^2_{\epsilon}$ & 0.3416
+\end{tabular}
+\caption{\label{tab:glmmests}Summary of the results from fitting the model given in Equation~\ref{eq:glmm}}
+\end{table}
+
+<<predictglmm, fig.height=6, fig.cap="Predictions from our generalized linear mixed effects model given in Equation~\ref{eq:glmm}.
+The lines show the expected probability of detecting the alternative data in a lineup of size 6 for new observers of new lineups is plotted on the $y$-axis, and the size of the parameter of interest is on the $x$-axis.
+The proportions detected by our Turk participants for each lineup group are shown by the points.">>=
+newdata <- read_csv("data/newdata_pred_glmm.csv")
+plotdata <- turk22 %>% mutate(type2 = ifelse(sign == 0, sign(initialEst), sign),
+                       type2 = ifelse(type == "one", type2, -1)) %>% 
+  group_by(pic_id, size, test_param, sign, type2,alt_model) %>% summarize(
+    datapick = mean(datapick)
+)
+
+ggplot() + 
+  geom_line(data = newdata, aes(x = size, y = predictfinal, color = as.factor(type2))) + 
+  geom_point(data = plotdata, aes(x = size, y = datapick, color = as.factor(type2))) + 
+  scale_color_brewer(palette = "Dark2", name = "Lineup Type") + 
+  facet_wrap(~test_param, scales = 'free') + 
+  labs(x = "Parameter value", y = "Expected probability of detection in a lineup of size 6") + 
+  ThemeNoNet + 
+  theme(legend.position = 'bottom')
+
+#vals <- predict(model1, type = "response")
+#modelData_sig$prediction <- vals
+#ggplot(data = modelData_sig) + 
+#  geom_point(aes(x = size, y = prediction, color = as.factor(type2))) + 
+#  facet_wrap(~test_param, scales = "free")
 @
@@
 Most network models, other than the most simple, lack the asymptotics required to develop the goodness-of-fit methods required \citep{goldenberg09}.
-Some methods have been developed based on what \citeauthor{RSienamanual} call ``auxiliary statistics" such as the indegree or outdegree distribution on the nodes.
+Some methods have been developed based on what Snijders et al call ``auxiliary statistics" such as the indegree or outdegree distribution on the nodes.
 In \texttt{RSiena}, the \texttt{sienaGOF} function performs goodness-of-fit testing as follows:
@@
 Furthermore, for the first lineup replicate with model M5 as the null model, only 9 of 21 participants were able to pick out the data, but the Visual distribution $p$-value is less than 0.05.
-This is much higher than the corresponding Binomial probability of $Pr(X = 9|n = 21, p = \frac{1}{6}) = \binom{21}{9}(\frac{1}{6})^9(\frac{5}{6})^{21-9} \approx 0.003$, but we not sure that having fewer than half of our participants identify the data plot is a truly significant result.
+This is much higher than the corresponding Binomial probability of $Pr(X = 9|n = 21, p = \frac{1}{6}) = \binom{21}{9}(\frac{1}{6})^9(\frac{5}{6})^{21-9} \approx 0.003$, but we are not sure that having fewer than half of our participants identify the data plot is a truly significant result.
 We could move the threshold from $\alpha = 0.05$ to $\alpha = 0.01$ or $\alpha = 0.001$, but more work needs to be done to determine which, if any of these, is appropriate.
@@
 
-\begin{enumerate}
-\item Use the mean values from step 2 as the parameter values in the model and simulate 100 observations from the model. These simulated networks will become the "true data."
-\item To each simulation, refit the model form from which it was simulated. 
-\item Repeat steps 4-5 for various permutations of model forms and parameter values; e.g. double the parameter values, quadruple the parameter values, change the signs on the parameters, etc.
-\item Simulate additional "alternative data" from the model in step 3 and the fitted model in step 4. 
-\item Place the appropriate node-link diagram of each "true data" observation in a lineup with $M-1$ ($M=6$) "alternative data" node-link diagrams. 
-\item Show the lineups to participants, with no participant seeing the same "true data" or "alternative data" plot twice. 
-\end{enumerate}
+We recruited 250 participants for our experiment through Amazon Mechanical Turk.
+Each participant read over some brief training material before beginning the experiment.
+After reading this material and agreeing to participate, the participants saw two trial plots, one where they were required to select the plot that was most different from the others due to its complex structure, and one where they were required to select the plot that was most different from the others due to its very simple structure.
+Upon getting the trial plots correct, they moved into the experiment.
+Each participant was randomly assigned 13 lineups to look at.
+The were asked to select one or more plots that they perceived as ``most different" from the others, and provide a reasoning for their choice.
+They could select from ``Most simple overall structure," ``Most complex overall structure," or ``Other" and provide their own text description of their reasoning.
+One of the 13 lineups the participants saw was the true data from the 112th senate shown in Section~\ref{sec:senatedata}.
+Each participant only saw the data one in order to avoid bias.
 
 \section{Discussion}
+
+By using visual inference methods, we have developed new ways to perform significance and goodness-of-fit testing for a complicated and intractable set of statistical models for social network data.
+These new methods can be used to supplement traditional methods and check our assumptions about network models.
+The traditional methods only look at one piece or one measure of a network model, but our methods look at the models holistically for a broader sense of what it means for a parameter to be significant or a model to be a good fit.
+By looking at an entire network simulated from a SAOM side-by-side with other instances of networks simulated from another model, we develop an idea of the model in terms of the data itself, instead of in terms of statistical summaries of the data.
+
+We hope to apply these methods further for different types of network data and different types of network models.
+We accept the limitations of this type of network data visualization, in that even in small instances, the cognitive load of looking at a lineup is very high for the average observer.
+We would therefore like to explore larger datasets, different layout algorithms, and different ways of visualizing network data, such as adjacency matrix visualizations, through visual inference to see if similar patterns emerge.
 

### 2017-10-25 ASApaper/code/modeling_data.R

@@ -14,9 +14,57 @@ turkbf2_sig %>% mutate(type2 = ifelse(sign == 0, sign(initialEst), sign),
 modelData_sig <- turkbf2_sig %>% mutate(type2 = ifelse(sign == 0, sign(initialEst), sign),
                                         type2 = ifelse(type == "one", type2, -1)) %>% filter(pic_id != 3132) 
 
+modelData_sig$centersize <- as.numeric(scale(modelData_sig$size))
+sub <- attr(scale(modelData_sig$size), "scaled:center")
+div <- attr(scale(modelData_sig$size), "scaled:scale")
+
 model1 <- glmer(datapick ~ -1 + test_param:size:as.factor(type2) + (1|pic_id) + (1|nick_name), 
-                data = modelData_sig, family = 'binomial')
+                data = modelData_sig, family = 'binomial', control = control)
 summary(model1)
+model1cent <- glmer(datapick ~ -1 + test_param:centersize:as.factor(type2) + (1|pic_id) + (1|nick_name), 
+                data = modelData_sig, family = 'binomial')
+summary(model1cent)
+
+control <- glmerControl(optimizer = 'bobyqa', optCtrl = list(maxfun = 1e06))
+model1mult <- glmer(datapick ~ test_param:param_value:as.factor(type2) + (1|pic_id) + (1|nick_name), 
+                    data = modelData_sig, family = 'binomial', control = control)
+summary(model1mult)
+model2mult <- glmer(datapick ~ -1 + test_param + test_param:param_value:as.factor(type2) + (1|pic_id) + (1|nick_name), 
+                    data = modelData_sig, family = 'binomial', control = control)
+summary(model2mult)
+save(model1mult, file = "fitglmm.RDA")
+model2mult <- glmer(datapick ~ test_param:poly(param_value,2):as.factor(type2) + (1|pic_id) + (1|nick_name), 
+                    data = modelData_sig, family = 'binomial', control = control)
+summary(model2mult)
+modelData_sig <- mutate(modelData_sig, interaction = interaction(test_param, as.factor(type2)))
+model3mult <- glmer(datapick ~ -1 + param_value:interaction + (1|pic_id) + (1|nick_name), 
+                data = modelData_sig, family = 'binomial', control = control)
+summary(model3mult)
+model2 <- glmer(datapick ~ -1 + (test_param:as.factor(type2)):poly(size,2) + (1|pic_id) + (1|nick_name), 
[…]

### 2017-10-25 ASApaper/code/prediction_glmm.R

@@ -1,0 +1,66 @@
+# new data for prediction from model 
+
+turk22 <- read_csv("data/turk22-sig.csv")
+turk22 <- turk22 %>% mutate(type2 = ifelse(sign == 0, sign(initialEst), sign),
+       type2 = ifelse(type == "one", type2, -1))
+parmsRange <- turk22 %>% group_by(test_param, type2) %>% summarize(minparm = min(size), maxparm = max(size)) 
+
+load("data/allModelMeans.RDS")
+get_mult <- function(mod, size, values = modelValues){
+  if (mod %in% c("dens", "recip")){
+    est <- filter(values, model == "basic", param == mod) %>% select(ests) %>% as.numeric()
+  } else if (mod == "bigmod"){
+    est <- 1
+  } else{
+    est <- filter(values, model == mod, param == mod) %>% select(ests) %>% as.numeric()
+  }
+  size/est
+}
+
+newdata <- parmsRange %>% 
+  mutate(newsizes = map2(minparm, maxparm, function(x,y){seq(x,y, (y-x)/1000)})) %>% 
+  unnest %>% select(test_param, type2, size = newsizes)
+newdata$type2 <- as.factor(newdata$type2)
+newdata$pic_id <- sample(1001:1050, replace = T, size = nrow(newdata))
+newdata$nick_name <- sample(LETTERS, replace = T, size = nrow(newdata))
+newdata$centersize <- (newdata$size - sub)/div
+newdata <- newdata %>% mutate(param_value = map2_dbl(test_param, size, get_mult))
+
+newdata <- newdata %>% mutate(interaction = interaction(test_param, as.factor(type2)))
+
+testnew <- predict(model1, newdata = newdata, allow.new.levels = T, type = 'response')
+testnewcent <- predict(model1cent, newdata = newdata, allow.new.levels = T, type = 'response')
+testnewmult <- predict(model1mult, newdata = newdata, allow.new.levels = T, type = 'response')
+testnewmult2 <- predict(model2mult, newdata = newdata, allow.new.levels = T, type = 'response')
+testfinal <- predict(finalGLMM, newdata = newdata, allow.new.levels = T, type = 'response')
+
+testnew3 <- predict(model3, newdata = newdata, allow.new.levels = T, type = 'response')
+
+testfixnew <- predict(modelfix, newdata = newdata, allow.new.levels = T, type = 'response')
+
+newdata$prediction <- testnew
[…]

### 2017-10-25 Code/Senate/big_model.R

@@ -1,0 +1,129 @@
+# simulate some observations from the "big model" 
+library(RSiena)
+bigModEsts <- read_csv("Data/senate/bigModFitsSenate.csv")
+
+bigModEsts %>% summarise_all(mean) %>% select(Rate:simXTransTrip) %>% as.numeric -> bigModMeans
+load("Data/senate/senateSienaNoHRC.rda")
+SenBasic <- getEffects(senateSiena)
+Sen_all_nojttp <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "sex", character = TRUE)
+Sen_all_nojttp <- includeEffects(Sen_all_nojttp, "simXTransTrip", include = TRUE, type = "eval", interaction1 = "bills", character = TRUE)
+Sen_all_nojttp <- includeEffects(Sen_all_nojttp, "sameXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+
+saom_simulate2 <- function(dat, struct, parms, N) {
+  require(RSiena)
+  struct$initialValue[struct$include] <- parms
+  myalgorithm <- sienaAlgorithmCreate(projname = Sys.time(), 
+                                      useStdInits = FALSE, cond = TRUE, nsub = 0, simOnly = TRUE, 
+                                      n3 = N)
+  getsims <- siena07(myalgorithm, data = dat, returnDeps = TRUE, 
+                     effects = struct, batch = TRUE, verbose = FALSE, silent = TRUE)
+  return(getsims$sims)
+}
+# turn simulation lists into dfs 
+sims_to_df <- function(sims, dat) {
+  Nact <- length(dat$nodeSets$Actors)
+  N <- length(sims)
+  waves <- length(sims[[1]][[1]][[1]])
+  simsdf <- NULL
+  counter <- 1
+  for (i in 1:N) {
+    for (j in 1:waves) {
+      dat <- as.data.frame(sims[[i]][[1]][[1]][[j]])
+      names(dat) <- c("from", "to", "dep.var.id")
+      ids <- 1:Nact
+      nodes <- data.frame(id = ids)
+      dat2 <- merge(dat, nodes, by.x = "from", by.y = "id", 
+                    all = T)
+      dat2$wave = j
+      dat2$sim = i
+      simsdf[[counter]] <- dat2
[…]

### 2017-10-25 Code/Senate/colorclusters.R

@@ -1,0 +1,61 @@
+# make an svg with only 1 plot
+dat <- read_csv("~/Desktop/Dissertation/NetworksVizInference/SenateLineupData/trials/bigmod_1_1.csv")
+
+dat4 <- filter(dat, ord == 4)
+
+p <- ggplot(data = dat) + 
+  geom_net(aes(from_id = from, to_id = to), 
+           arrow = arrow(type = 'open', length = unit(2, "points") ), 
+           linewidth = .25, singletons = T, fiteach = T, directed = T, 
+           color = 'black', arrowgap = .015, arrowsize = .3, size = 1) + 
+  theme_net() + 
+  theme(panel.background = element_rect(color = 'black')) +
+  facet_wrap(~ord)
+make_interactive("testinggridsvg.svg", script = scriptURL)
+ggplot(data = dat4) + 
+  geom_net(aes(from_id = from, to_id = to), 
+           arrow = arrow(type = 'open', length = unit(2, "points") ), 
+           linewidth = .25, singletons = T, fiteach = T, directed = T, 
+           color = 'black', arrowgap = .015, arrowsize = .3, size = 1) + 
+  theme_net() + 
+  theme(panel.background = element_rect(color = 'black')) +
+  facet_wrap(~ord)
+make_interactive("testinggridsvg4.svg", script = scriptURL)
+
+# test coloring to add to the graphs
+
+library(tidyverse)
+library(igraph)
+net <- read.csv("~/Downloads/simttbx3sd1392457revTRUE.csv", stringsAsFactors = FALSE)
+df <- data.frame(expand.grid(
+  from=paste0("V", 1:155), to = paste0("V", 1:155)), stringsAsFactors = FALSE)
+
+
+netnest <- net %>% group_by(sim) %>% nest()
+netnest <- netnest %>% mutate(
+  data = data %>% purrr::map(.f = function(d) {
+    adj1 <- d %>% right_join(df) %>% 
+      xtabs(!is.na(dep.var.id)~from+to, data =.) %>% as.matrix()
+    
+    ig1 <- graph_from_adjacency_matrix(adj1)
+    comps <- components(ig1)
+    clu_df <- data.frame(from=names(comps$membership), clu = comps$membership)
+    clu_df$csize=comps$csize[clu_df$clu]
+    d %>% left_join(clu_df, by="from")
+  })
+)
+
+net1 <- unnest(netnest, data)
[…]

### 2017-10-25 Code/Senate/creating_lus.R

@@ -174,7 +174,6 @@ senate_lu_storage <- function(details, Altsims, Nullsims, m, Wave, url = "http:/
       
     }
   return(picture_details)
-  
 }
 
 library(grid)

### 2017-10-25 Code/Senate/creating_lus2.R

@@ -1,0 +1,235 @@
+# create lineups for experiment
+library(tidyverse)
+library(netvizinf)
+library(geomnet)
+library(grid)
+library(gridSVG)
+details <- data_frame(pic_id = integer(0), data_file = character(0), obs_plot_location = integer(0))
+
+# get all the data
+details <- data_frame(data_file = c(list.files("SenateLineupData/Non-GoF/"), list.files("SenateLineupData/GoF/")))
+details %>% mutate(data_name = data_file) %>% 
+  mutate(data_file = str_replace(data_file, ".csv", "")) %>%
+  separate(data_file, c("model", "sign", "difficulty", "rep"), "_") -> details
+details$model <- as.factor(details$model)
+levels(details$model) <- c(1,3,4,2,6,5)
+details$model <- as.integer(as.character(details$model))
+details$sign <- as.factor(details$sign)
+levels(details$sign) <- c(9,2,1)
+details$sign <- as.integer(as.character(details$sign))
+details$difficulty <- as.factor(details$difficulty)
+levels(details$difficulty) <- c(1,3,2)
+details$difficulty <- as.integer(as.character(details$difficulty))
+details$difficulty[is.na(details$difficulty)] <- 9
+details$rep[is.na(details$rep)] <- 1
+get_dat_plot <- function(sign, data_name){
+  if(sign == 9){
+    dat <- read_csv(paste0("SenateLineupData/GoF/", data_name))
+  } else {
+    dat <- read_csv(paste0("SenateLineupData/Non-GoF/", data_name))
+  }
+  sims <- unique(dat$sim)
+  datplot <- setdiff(sims, 1:5)
+  unique(dat$ord[which(dat$sim == datplot)])
+}
+
+details %>% mutate(obs_plot_location = map2_int(sign, data_name, get_dat_plot)) -> details
+details %>% mutate(pic_id = as.integer(paste0(difficulty, sign,model,rep))) -> details
+
+# get test_param, test_value
+library(readxl)
+expDet <- read_excel("SenateLineupData/effects_for_lineups.xlsx")
+expDet <- expDet %>% mutate(test_param = Model)
+expDet$test_param <- as.factor(expDet$test_param)
+levels(expDet$test_param) <- c(1,3,4,2,6,5)
+expDet$test_param <- as.integer(as.character(expDet$test_param))
+expDet$model <- expDet$test_param
+expDet$Sign <- as.factor(expDet$Sign)
+levels(expDet$Sign) <- c(2,1,9)
[…]

### 2017-10-25 Code/Senate/fit_bigger_models.R

@@ -1,0 +1,66 @@
+setwd("~/Desktop/Dissertation/NetworksVizInference")
+
+load("Data/senate/senateSienaNoHRC.rda")
+library(RSiena)
+SenBasic <- getEffects(senateSiena)
+Sen_all <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+Sen_all <- includeEffects(Sen_all, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "sex", character = TRUE)
+Sen_all <- includeEffects(Sen_all, "simXTransTrip", include = TRUE, type = "eval", interaction1 = "bills", character = TRUE)
+Sen_all <- includeEffects(Sen_all, "sameXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+Sen_all_nojttp <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "sex", character = TRUE)
+Sen_all_nojttp <- includeEffects(Sen_all_nojttp, "simXTransTrip", include = TRUE, type = "eval", interaction1 = "bills", character = TRUE)
+Sen_all_nojttp <- includeEffects(Sen_all_nojttp, "sameXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+
+myalg <- sienaAlgorithmCreate( projname = Sys.time() , n3 = 1000)
+
+p = Sys.time()
+big_model <- siena07(myalg, data = senateSiena, effects = Sen_all, returnDeps = TRUE,
+                    batch=TRUE, verbose = TRUE, silent = FALSE)
+Sys.time() - p       
+
+p = Sys.time()
+big_model2 <- siena07(myalg, data = senateSiena, effects = Sen_all, prevAns = big_model, returnDeps = TRUE,
+                    batch=TRUE, verbose = TRUE, silent = FALSE)
+Sys.time() - p  
+
+p = Sys.time()
+big_model3 <- siena07(myalg, data = senateSiena, effects = Sen_all, prevAns = big_model2, returnDeps = TRUE,
+                    batch=TRUE, verbose = TRUE, silent = FALSE)
+Sys.time() - p  
+
+p = Sys.time()
[…]

### 2017-10-25 Code/Senate/trial_lus_exper.R

@@ -1,0 +1,64 @@
+# create trials for experiment
+
+# 10 total
+# 5 most complex
+# 5 most simple 
+
+details_tri <- data_frame(data_file = list.files("SenateLineupData/trials/"))
+
+details_tri <- details_tri %>% mutate(data_name = data_file) %>% 
+  mutate(data_file = str_replace(data_file, ".csv", "")) %>%
+  separate(data_file, c("model", "type", "rep"), "_") 
+details_tri
+details_tri$model <- as.factor(details_tri$model)
+levels(details_tri$model) <- c(7,3,4,6,5)
+details_tri$model <- as.integer(as.character(details_tri$model))
+details_tri$type <- as.factor(details_tri$type)
+levels(details_tri$type) <- c(1, 5)
+details_tri$type <- as.integer(as.character(details_tri$type))
+
+get_dat_plot_tri <- function(data_name){
+  dat <- read_csv(paste0("SenateLineupData/trials/", data_name))
+  sims <- unique(dat$sim)
+  datplot <- setdiff(sims, 1:5)
+  unique(dat$ord[which(dat$sim == datplot)])
+}
+
+details_tri %>% mutate(obs_plot_location = map_int(data_name, get_dat_plot_tri)) -> details_tri
+details_tri %>% mutate(pic_id = as.integer(paste0(model,type,rep))) -> details_tri
+
+senate_lu_storage3 <- function(details, url = "http://www.hofroe.net/examples/lineup/action-back.js"){
+  #  browser()
+  picture_details <- data_frame(pic_id = integer(0), sample_size = integer(0), test_param = character(0),
+                                param_value = character(0), p_value = numeric(0), obs_plot_location = integer(0),
+                                pic_name = character(0), experiment = character(0), difficulty = character(0), data_name = character(0))
+  
+  for(i in 1:nrow(details)){
+    dat <- read_csv(paste0("SenateLineupData/trials/", details$data_name[i]))
+    p <-ggplot(data = dat) + 
+      geom_net(aes(from_id = from, to_id = to), 
+               arrow = arrow(type = 'open', length = unit(2, "points") ), 
+               linewidth = .25, singletons = T, fiteach = T, directed = T, 
+               color = 'black', arrowgap = .015, arrowsize = .3, size = .3) + 
+      theme_net() + 
+      theme(panel.background = element_rect(color = 'black')) +
+      facet_wrap(~ord)
[…]

### 2017-10-25 ASApaper/paper-child.Rnw
couple of comments - sorry have to go blind with the Latex

@@
 % }
-Two of the most imporant pieces of statistical modeling are significance testing of model parameters and goodness-of-fit measures and tests.
-In the former, the data are usually assumed to come from a simple model under the null hypothesis, and additional parameters are added and tests done to determine if the additional should be included in the model.
+Two of the most important pieces of statistical modeling are significance testing of model parameters and goodness-of-fit tests.
+In the former, the data are usually assumed to come from a simple model under the null hypothesis, and additional parameters are tested whether they significantly contribute to explaining variability in the data.
 In the latter, the model of interest is examined to determine how well it fits the data.
 Both of these aspects of statistical modeling increase greatly in difficulty as the complexity of the model of interest increases.
-The more complicated the model, the harder it is to determine fit or to include or exclude a parameter.
+%The more complicated the model, the harder it is to determine fit or to include or exclude a parameter.
 
@@
 We refer to the ``things" in the network as \textit{nodes}, or \textit{actors} in a social network, and the connections as \textit{edges}, or \textit{ties} in a social network.
-Network objects are particularly difficult to model due to the dependencies inherent to the data.
-This complexity can increases when, instead of looking at one instance of a network, we consider the dynamics of network change in between those singular observed instances.
-This type of modelling for dynamic networks is often performed on social network data, such as for friendship networks among students or the spread of HIV in communities of men sleeping with other men.
+Dependencies inherent to the data make network objects particularly difficult to model.
+Even more challenging is the situation when we go beyond single instances of a network and consider the dynamics of network change between observed instances.
+This type of modelling for dynamic networks is often performed on social network data, such as for friendship networks among students or the spread of HIV in  sexually active communities.% of men sleeping with other men.
 These models lack the asymptotics required to perform many well-known goodness-of-fit tests, and the maximum likelihood estimation of parameters is so difficult that it can make significance testing significant as well \citep{goldenberg09}.
+\hh{is the previous sentence a quote? I'm not sure that I follow.}
 
 We propose new methods for significance and goodness-of-fit testing for a set of social network models, stochastic actor-oriented models for dynamic network data \citep{saompaper}.
-These new methods use \textit{visual inference} in place of traditional statistical methods for social network models, such as Wald tests for significance of parameters and in- and outdegree distribution metrics for determining goodness-of-fit.
-By using the visual inference methodology of \citet{Bujaetal}, we can look at the entire dataset simulated from a network model as opposed to a single metric on the network such as outdegree, or a $p$-value for a single parameter in the model.
+Specifically, we are using \textit{visual inference} in place of traditional statistical methods for social network models, such as Wald tests for significance of parameters and in- and outdegree distribution metrics for determining goodness-of-fit.
+Visual inference, introduced by \citet{Bujaetal}, allows us to look at the entire dataset simulated from a network model as opposed to a (set of) usually one-dimensional metric(s) derived from the network such as outdegree, or a $p$-value for a single parameter in the model.
 
-The paper is outlined as follows: Section 2 gives a basic overview of visual inference and the lineup protocal.
+The paper is outlined as follows: Section~\ref{vi} gives a basic overview of visual inference and the lineup protocol.
 Section 3 provides an introduction to the our models of interest, stochastic actor-oriented models.
@@
 
-\section{Visual Inference}
+\hh{where do the results come in?}
+
+\section{Visual Inference}\label{vi}
 
@@
 The Rorschach protocol helps to understand the extent of randomness in the null model.
-We use line lineup protocol.
-Under this protocol, a plot of the observed data is placed randomly among a set of $m-1$ null plots, and human observers are then asked to examine the lineup and to identify the most different plot.
-If observers identify the data plot, this is quantifiable evidence against the null hypothesis.
+Here, we use the lineup protocol.
+Under this protocol, a plot of the observed data is placed randomly among a set of $m-1$ null plots (where $m =20$, usually), and human observers are then asked to examine the lineup and to identify the most different plot.
+If an observer identifies the data plot, this is quantifiable evidence against the null hypothesis.
+\hh{Similarly, an observer has a chance of 1 in $m$ to pick the data plot from the lineup by simply guessing, ie in a situation where the data plot is virtually indistinguishable from the null plots.}
 
@@
 %To model family that we apply to the Senate data is the Stochastic Actor-Oriented Model (SAOM) family. 
-This family of models for dynamic network data, which were first introduced by \citet{saompaper}, incorporates both network structure and node-level information to describe how a network observed on two or more occasions changes in time.
+Stochastic Actor-Oriented Models (SAOMs) are a family of models for dynamic network data \citep{saompaper} that incorporate both network structure and node-level information to describe how a network observed on two or more occasions changes over time.
 The two titular properties of SAOMs, stochasticity and actor-orientation, are crucial to understanding networks as they exist naturally: social networks are ever-changing as relationships decay or grow in seemingly random ways, and most actors in them have characteristics that could affect how they change their ties to other nodes in the network.
@@
 Broadly, a SAOM takes network structure and node covariate information into account in two ways and models the network changes as a continuous time Markov chain (CTMC).
-First, the rate of change between states is dictated by a rate function that describes \textit{when} changes in the network occur, and secondy, the objective function describes \textit{what} those state changes are.
-As in many other network models, the variables of interest, $x_{ij}$, are the binary edges of the network.
-For any two nodes $i \neq j$ in the network, $i,j \in \{1, 2, \dots, n=\text{the number of nodes}\}$,
+First, the rate of change between states is dictated by a rate function that describes \textit{how often} changes in the network occur, and secondly, the objective function describes \textit{what} those state changes are.
+As in many other network models, the variables of interest,  are the binary edges of the network.
+Let $x_{ij}$ denote the edge between nodes $i$ and $j$, where $i,j \in \{1, 2, \dots, n=\text{the number of nodes}\}$.
+$x_{ij}$ is modelled as a binary variable, i.e.
   \begin{equation}\label{eq:edgevars}
@@
   \end{equation}
-The edges are also treated as \textit{directed}, so it is not required that $x_{ij} = x_{ji}$, and self-referencing edges or loops, $x_{ii}$, are not allowed.
-The entire network at time point $m = 1, 2, \dots M$ for $M$ the number of network observations, is denoted $x(t_m)$.
-More detail on SAOMs follows, and additional details can be found in \citet{saompaper, snijders01, snijders2010, snijdersetall:2007, snijders:2010, snijders:2017},
+Edges are treated as \textit{directed}, i.e. in general $x_{ij} \neq x_{ji}$, and self-referencing edges or loops are not allowed, i.e.  $x_{ii} = 0$ for all $i$. 
+Assume, the network is observed $M$ times at time points  $t_1 < t_2 < ...
+< t_M$, then the  entire network at time point $t_m$ is denoted as $x(t_m)$.
+In sections~\ref{rf} and~\ref{of} we discuss the rate and objective functions of a SOAM in more depth.
+Additional details on SOAMs can be found in \citet{saompaper, snijders01, snijders2010, snijdersetall:2007, snijders:2010, snijders:2017},
 
-\subsection{Rate Function}
+\subsection{Rate Function}\label{rf}
 
-All changes in SAOMs are treated as changes made by the nodes, or \textit{actors}, in the network.
-So, each actor, $i$, gets a chance to make a change according to the rate function, typically denoted $\lambda_i$, which dictates when relationships between nodes in the network can change.
+All changes in SAOMs are treated as changes made by the nodes, or \textit{actors}, in the network, i.e.\ each actor, $i$, gets a chance to make a change according to the rate function, typically denoted $\lambda_i$, which dictates when relationships between nodes in the network can change.
 In general, the rate function can take the network structure e.g.
@@
 
-\subsection{Objective Function}
+\subsection{Objective Function}\label{of}
 

### 2017-10-25 ASApaper/paper-child.Rnw
couple more comments

@@
 The probability that actor $i$ changes its current tie to actor $j$ is determined by the \textit{objective function} of the model and a random component, $U$.
-The random component is assumed to have log-Weibull distribution with location parameter $\mu = 0$ and scale parameter $\sigma = 0$ (see \citet{modelsSnijders}), so the probability density function of $U$ is
-\begin{equation}\label{eq:logweibull}
-g(u) = \exp\{-(u + \exp\{-u\})\}.
-\end{equation}
-The objective function, $f_i$, is the driving force of a SAOM.
-Actor $i$ always wants to maximize its objective function given the current state of the network, $x$ and the node-level covariates, $\mathbf{Z}$:% which has the form
+\hh{XXX can you paraphrase the purpose of U in half a sentence?}
+%The random component is assumed to have log-Weibull distribution with location parameter $\mu = 0$ and scale parameter $\sigma = 0$ (see \citet{modelsSnijders}), so the probability density function of $U$ is
+%\begin{equation}\label{eq:logweibull}
+%g(u) = \exp\{-(u + \exp\{-u\})\}.
+%\end{equation}
+%The objective function, $f_i$, is the driving force of a SAOM. 
+Actor $i$ is aiming to maximize the objective function $f_i$ given the current state of the network, $x$ and the node-level covariates, $\mathbf{Z}$, given as:% which has the form
 \begin{equation}\label{eq:objective}
@@
 \end{equation}
-where $\boldsymbol{\beta} = (\beta_1, \dots, \beta_K)$ are additional model parameters, each associated with some network statistics calculated with respect to actor $i$, $s_{ik}(x, \mathbf{Z})$.
-The network statistics range from the simple outdegree, $s_i(x) = \sum_{i\neq j} x_{ij}$, to the more complicated transitive triplets jumping to different covariate, $s_i(x, \mathbf{Z}) = \sum_{i \neq j \neq h} x_{ij}x_{ih}x_{hj} \cdot \mathbb{I}(z_i = z_h \neq z_j)$, plus many more.
-At last count, in the software we use to fit these models to network data, \texttt{RSiena}, there are over 80 possible effects that can be included in the objective function \citep{RSiena}.
+where $\boldsymbol{\beta} = (\beta_1, \dots, \beta_K)$ are additional model parameters, each associated with some network statistics, $s_{ik}(x, \mathbf{Z})$, $s_{ik}(x, \mathbf{Z})$, calculated with respect to actor $i$.
+Network statistics range from the simple outdegree, $s_i(x) = \sum_{i\neq j} x_{ij}$, to the more complicated {\it transitive triplets jumping to different covariate}, $s_i(x, \mathbf{Z}) = \sum_{i \neq j \neq h} x_{ij}x_{ih}x_{hj} \cdot \mathbb{I}(z_i = z_h \neq z_j)$.
+%, plus many more.
+At last count, in the software we use to fit these models to network data,
+Version 1.2-3 of \texttt{RSiena} \citep{RSiena}, the software used to fit the models here, provides  over 80 possible effects that can be included in the objective function.
 We discuss these statistics in more detail in Section~\ref{sec:models}.
 
-The objective function $f_{i}(x, \boldsymbol{\beta}, \mathbf{Z})$ and the random component $U$ combine to form a \textit{transition probability}, $p_{ij}$, of the network changing from its current state $x$ to the state with only the tie $x_{ij}$ changed, which is denoted $x(i \leadsto j)$:
+Objective function $f_{i}(x, \boldsymbol{\beta}, \mathbf{Z})$ and  random component $U$ are combined to form the \textit{transition probability}, $p_{ij}$, of the network changing from its current state $x$ to the state with changed tie $x_{ij}$, denoted as $x(i \leadsto j)$:
 \begin{equation}\label{eq:transprob}
@@
 This probability dictates which edge change is made by the acting node. The acting node can also choose to \textit{not} change at all, in which case $j \equiv i$.
+
+\hh{XXX how is the probability computed in the case that no change is made? Why is the notation $j \equiv i$ used for no change?}
 

### 2017-10-26 ASApaper/paper-child.Rnw
edit file paths to data

@@
 ">>=
-seobama <- read_csv("../Data/senate/senateobamapres_gsw_25.csv")
+seobama <- read_csv("data/senateobamapres_gsw_25.csv")
 to_label <- c("Joseph R. Biden Jr.", "John S. McCain", "Ted Cruz","Marco Rubio",
@@
 # used for table below
-initeff <- read_csv("../Data/senate/sigEffsSenate912.csv")
+initeff <- read_csv("data/sigEffsSenate912.csv")
 initeff %>% filter(shortName == "sameX")

### 2017-10-27 ASApaper/paper-child.Rnw
reorganize, add chang correlation reference to bib (not used yet)

@@
 
-\section{Significance Testing}\label{sec:sigtest}
+We fit models M1 through M6 in \texttt{RSiena} using Markov Chain Monte Carlo (MCMC) methods to approximate the method of moments estimates of the parameters.
+Because the estimation is done through MCMC simulation, we fit each model to the data 1,000 times to get a better estimate of the true value of $\boldsymbol{\beta}$.
+From the simulations that converged, which made up over 90\% of the fits for each model, we computed the mean of the 1,000 estimates of each parameter to get final estimates of  $\hat{\boldsymbol{\beta}}$ for each model, which are given in Table~\ref{tab:fittedvalues}.
 
-<<sigtestWald, eval = FALSE, echo = FALSE>>=
-SenBasic <- getEffects(senateSiena)
-Senjtt_p <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
-myalg <- sienaAlgorithmCreate( projname = Sys.time() , n3 = 1000)
-fits <- siena07(myalg, data = senateSiena, effects = Senjtt_p, returnDeps = TRUE,
-                                      batch=TRUE, verbose = FALSE, silent = TRUE)
-thet <- fits$theta
-sig <- fits$covtheta
-Wald.RSiena(A = c(0,0,1), fits)
-th <- c(0,0,1) %*% thet
-covmat <- c(0,0,1) %*% sig %*% c(0,0,1)
-csq <- drop(th %*% solve(covmat) %*% th)
-1 - pchisq(csq, 1)
+<<modelestimates>>=
+load("data/allModelMeans.RDS")
+modelMeanEsts %>% unnest(ests) %>% mutate(param = "rate") -> modelmeans
+modelmeans$param[c(1,6,12, 18,24, 30)] <- "alpha1"
+modelmeans$param[c(1,6,12, 18,24, 30)+1] <- "alpha2"
+modelmeans$param[c(1,6,12, 18,24, 30)+2] <- "alpha3"
+modelmeans$param[c(1,6,12, 18,24, 30)+3] <- "beta1"
+modelmeans$param[c(1,6,12, 18,24, 30)+4] <- "beta2"
+modelmeans$param[modelmeans$param == "rate"] <- c("beta3", 'beta4', 'samep', 'beta6', 'beta5')
+modelmeans %>% filter(model != "samep") %>% spread(param, ests) -> modelmeans
+M6 <- c(2.4405048,2.4594403,2.2098176,-4.9232775,4.8916183,2.3743720,0.2047038,6.9661589)
 @
 
-For a SAOM, there are two ways a significance test of the parameters can be performed.
-In \texttt{RSiena}, there are $t$-type tests and Wald-type test for a single parameter and for multiple parameters.
-The $t$-type test statistic is simply the parameter estimate divided by its standard error, and compared to a standard normal distribution.
-The Wald-type test statistic for a single parameter, $\beta_k$ is
-\begin{equation}\label{eq:wald1}
-\frac{(\hat{\beta_k})^2}{var(\hat{\beta_k})},
-\end{equation}
-which is compared to a Chi-square distribution with one degree of freedom.
-Testing the significance of multiple parameters depends on the hypothesis we wish to test, and a $P \times K$ matrix, $A$, must be appropriately designed to test the $P$ hypotheses of interest.
-The null hypothesis is that $A\boldsymbol{\beta} = \mathbf{0}$, and the test statistic is
-\begin{equation}\label{eq:wald2}
-(A\boldsymbol{\hat{\beta}})' \hat{\Sigma}^{-1} A\boldsymbol{\hat{\beta}},
-\end{equation}
-where $\hat{\Sigma}$ is the estimated covariance matrix of $\boldsymbol{\beta}$. This statistic is then compared to a Chi-square distribution with $P$ degrees of freedom.
+\begin{table}
+\centering
+\begin{tabular}{|l|ccccccccc|}
+\hline
+Model & $\alpha_1$ & $\alpha_2$ & $\alpha_3$ & $\beta_1$ & $\beta_2$ & $\beta_3$ & $\beta_4$ & $\beta_5$ & $\beta_6$ \\
+\hline
+\hline
+M1 & \Sexpr{round(modelmeans$alpha1[1], 3)} & \Sexpr{round(modelmeans$alpha2[1], 3)} & \Sexpr{round(modelmeans$alpha3[1], 3)} & \Sexpr{round(modelmeans$beta1[1], 3)} & \Sexpr{round(modelmeans$beta2[1], 3)} & -- & -- & -- & --\\
+M2 & \Sexpr{round(modelmeans$alpha1[2], 3)} & \Sexpr{round(modelmeans$alpha2[2], 3)} & \Sexpr{round(modelmeans$alpha3[2], 3)} & \Sexpr{round(modelmeans$beta1[2], 3)} & \Sexpr{round(modelmeans$beta2[2], 3)} & \Sexpr{round(modelmeans$beta3[2], 3)} & -- & -- & --\\
+M3 & \Sexpr{round(modelmeans$alpha1[3], 3)} & \Sexpr{round(modelmeans$alpha2[3], 3)} & \Sexpr{round(modelmeans$alpha3[3], 3)} & \Sexpr{round(modelmeans$beta1[3], 3)} & \Sexpr{round(modelmeans$beta2[3], 3)} & -- & \Sexpr{round(modelmeans$beta4[3], 3)} & -- & --\\
+M4 & \Sexpr{round(modelmeans$alpha1[5], 3)} & \Sexpr{round(modelmeans$alpha2[5], 3)} & \Sexpr{round(modelmeans$alpha3[5], 3)} & \Sexpr{round(modelmeans$beta1[5], 3)}& \Sexpr{round(modelmeans$beta2[5], 3)} & -- & -- & \Sexpr{round(modelmeans$beta5[5], 3)} & -- \\
+M5 & \Sexpr{round(modelmeans$alpha1[4], 3)} & \Sexpr{round(modelmeans$alpha2[4], 3)} & \Sexpr{round(modelmeans$alpha3[4], 3)} & \Sexpr{round(modelmeans$beta1[4], 3)} & \Sexpr{round(modelmeans$beta2[4], 3)} & -- & -- & -- & \Sexpr{round(modelmeans$beta6[4], 3)}\\
+M6 & \Sexpr{round(M6[1], 3)} & \Sexpr{round(M6[2], 3)} & \Sexpr{round(M6[3], 3)} & \Sexpr{round(M6[4], 3)} & \Sexpr{round(M6[5], 3)} & -- & \Sexpr{round(M6[6], 3)} & \Sexpr{round(M6[8], 3)} & \Sexpr{round(M6[7], 3)} \\
+\hline
+\end{tabular}
+\caption{\label{tab:fittedvalues} The final estimates from repeated estimation of models M1 through M6.}
+\end{table}
 
-We added each parameter to the basic model M1 to create models M2-M5, and then combined three of the four effects to create model M7.
-We fit these models in \texttt{RSiena} using Markov Chain Monte Carlo (MCMC) methods to approximate the method of moments estimates of the parameters.
-Because the estimation is done through MCMC simulation, we fit each model to the data 1,000 times to get a better estimate of the true value of $\boldsymbol{\beta}$.
-From the simulations that converged, which made up over 90\% of the fits for each model, we computed the mean of the 1,000 estimates of each parameter to get our values of $\hat{\boldsymbol{\beta}}$ from which to simulate data.
+We use the estimates given in Table~\ref{tab:fittedvalues} to simulate from models M1 through M6. We discuss the simulation procedure and how we use the simulations in Section~\ref{sec:expersetup}.
+
+\section{Experiment Set-Up}\label{sec:expersetup}
+
+We want to explore three different aspects of the SAOM models using the lineup protocol: significance of parameters, goodness-of-fit of a model, and visual detection of additional parameters.
+Each on of these requires a different setup, which we describe in detail, but we use the lineup protocol to examine each aspect.
+In each lineup, there are two models shown: the null model and the alternative model.
+The definition of the null and alternative model will vary according to which aspects of the SAOMs we are exploring.
+%For each lineup, we consider two models: model M1 is the null model, and another model, from M2-M5, which is chosen to be the alternative model.
+
+To construct a lineup, we simulate five networks from the null model and one network from the alternative model.
+We chose to expose our participants to only six plots at a time in order to show the node-link diagrams in more detail and to lower their cognitive load.
+Typically, lineup experiments are done with sets of 20 plots at a time c.f.~\citet{loy:2015, vanderplas:2016}, but we determined that not enough structure could be shown in each plot for 20 node-link diagrams.
+An example of a lineup like those shown to our participants is shown on the right side of the image  in Figure~\ref{fig:shiny}.
+\footnote{If you would like to explore the lineups we use in further detail, please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/}}.
+
+\begin{figure}
+\includegraphics[width=.9\textwidth]{img/shinyappscreen.png}
+\caption{\label{fig:shiny} A screen grab of the web application we created to design our lineup experiment.
+In this view, model M4 is selected and one estimate from model M4 with $\beta_5$ set to twice its estimated value given in Table~\ref{tab:effects} is placed at random among five observations simulated from model M1.
+The other two tabs in the right side of the screen tell you which plot comes from the alternative model, and allow you to view and download the data that is plotted in the lineup.}
+\end{figure}
+
+In model M1, with only two parameters in the objective function, we varied both the density and reciprocity parameteter values one at a time.
+In models M2 through M5, we vary the additional parameter, $\beta_3$ through $\beta_6$.
+Thus, we have six different parameters of interest to us: $\beta_1, \dots, \beta_6$.
+We want to determine how the size of these parameters affects the overall structure of the network data simulated from the models M1 through M5, so we also vary the value of the parameters in two directions: positive and negative.
+
+\subsection{Significance Testing}
+
+\subsection{Goodness-of-Fit}
+\subsection{Visual Detection}
 
@@
 An effect becomes noticeable when a simulation from a model with a noticeable effect appears different from simulations from a model where the same affect is absent.
-Because the traditional way to visualize a network is a node-link diagram, we will examine the affect of different parameter values using node-link visualizations.
-
-For each lineup, we consider two models: model M1 is the null model, and another model, from M2-M5, which is chosen to be the alternative model.
-To construct a lineup, we simulate five networks from the null model and one network from the alternative model.
-We chose to expose our participants to only six plots at a time in order to show the node-link diagrams in more detail and to lower their cognitive load.
-Typically, lineup experiments are done with sets of 20 plots at a time c.f.~\citet{loy:2015, vanderplas:2016}, but we determined that not enough structure could be shown in each plot for 20 node-link diagrams\footnote{If you would like to explore the lineups we use in further detail, please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/}}.
-
-To determine the threshold at which the effect becomes noticeable, we examined six different levels of the effect, three negative and three positive.
-We hypothesized that the perception of the effect by our participants would look something like Figure~\ref{fig:hypothesis}.
-The higher in absolute value the parameter is, the more likely participants are to choose the alternative model out of the lineup.
-To determine the six levels of the effects we want to test, started with the estimates of the parameter at hand as determined by the procedure described in Section~\ref{sec:models}, and multiply it or changing its sign to determine what \textit{we} notice about the effect of the parameter in simulations from the changed models.
+%Because the traditional way to visualize a network is a node-link diagram, we will examine the effect of different parameter values using node-link visualizations.
 
@@
 
-We showed participants from Amazon Mechanical Turk a series of lineups containing data simulated from M1 through M5.
-In model M1, with only two parameters in the objective function, we varied both the density and reciprocity parameteter values one at a time.
-In models M2 through M5, we vary the additional parameter, $\beta_3$ through $\beta_6$.
-Thus, we have six different parameters of interest to us: $\beta_1, \dots, \beta_6$.
-We want to determine how the size of these parameters affects the overall structure of the network data simulated from the models M1 through M5, so we also vary the value of the parameters in two directions: positive and negative.
+To determine the threshold at which the effect becomes noticeable, we examined six different levels of the effect, three negative and three positive.
+We hypothesized that the perception of the effect by our participants would look something like Figure~\ref{fig:hypothesis}.
+The higher in absolute value the parameter is, the more likely participants are to choose the alternative model out of the lineup.
+To determine the six levels of the effects we want to test, started with the estimates of the parameter at hand as determined by the procedure described in Section~\ref{sec:models}, and multiply it or changing its sign to determine what \textit{we} notice about the effect of the parameter in simulations from the changed models.
 
@@
 The plot(s) that appear(s) that are \textit{not} from the model specified using the other options are simulations from model M1 with the estimates of the rate parameters, $\beta_1$ and $\beta_2$ as calculated by the procedure described in Section~\ref{sec:models}.
-
-\begin{figure}
-\includegraphics[width=.9\textwidth]{img/shinyappscreen.png}
-\caption{\label{fig:shiny} A screen grab of the web application we created to design our lineup experiment.
-In this view, model M4 is selected and one estimate from model M4 with $\beta_5$ set to twice its estimated value given in Table~\ref{tab:effects} is placed at random among five observations simulated from model M1.
-The other two tabs in the right side of the screen tell you which plot comes from the alternative model, and allow you to view and download the data that is plotted in the lineup.}
-\end{figure}
 
@@
 
-A summary of the results from our experiment is shown in Figure~\ref{fig:readturk22}.
-On the $x$ axis, we plot the value of the parameter of interest, and on the $y$ axis, the proportion of times the plot of data simulated from the alternative model was picked out.
-The results are split into groups based on the sign of the parameter.
-We can see clear patterns in all of the groups: as the parameter value approaches 0, fewer participants identified the alternative plot.
+\section{Experiment Results}
 
-<<readturk22, fig.height=6, fig.cap="A summary of some results from the Amazon mechanical turk study.
-On the $x$-axis, the value of the parameter changed in the model of interest, and on the $y$-axis, the proprotion of times turkers identified the different data plot.
-A simple linear regression is fit to each group.
-The pattern as the parameter value approaches zero is clear in all conditions.">>=
-turk22 %>% mutate(type2 = ifelse(sign == 0, sign(initialEst), sign),
-                       type2 = ifelse(type == "one", type2, -1)) %>% 
-  group_by(pic_id, size, test_param, sign, type2,alt_model) %>% 
-  summarize(datapick = mean(datapick)) %>%
-  filter(pic_id != 3132) %>% 
-  ggplot(aes(x = size, y = datapick, colour=factor(type2))) + 
-  geom_point() +
-  scale_color_brewer(palette = "Dark2", name = "Lineup Type") + 
-  facet_wrap(~test_param, scales="free", labeller = 'label_both') +
-  geom_smooth(se=FALSE, method="lm") + 
-  ThemeNoNet + 
-  theme(legend.position = 'bottom')
+We showed participants from Amazon Mechanical Turk a series of lineups containing data simulated from M1 through M5.
+
+\subsection{Significance Testing}\label{sec:sigtest}
+
+<<sigtestWald, eval = FALSE, echo = FALSE>>=
+SenBasic <- getEffects(senateSiena)
+Senjtt_p <- includeEffects(SenBasic, "jumpXTransTrip", include = TRUE, type = "eval", interaction1 = "party", character = TRUE)
+myalg <- sienaAlgorithmCreate( projname = Sys.time() , n3 = 1000)
+fits <- siena07(myalg, data = senateSiena, effects = Senjtt_p, returnDeps = TRUE,
+                                      batch=TRUE, verbose = FALSE, silent = TRUE)
+thet <- fits$theta
+sig <- fits$covtheta
+Wald.RSiena(A = c(0,0,1), fits)
+th <- c(0,0,1) %*% thet
+covmat <- c(0,0,1) %*% sig %*% c(0,0,1)
+csq <- drop(th %*% solve(covmat) %*% th)
+1 - pchisq(csq, 1)
 @
 
-We futher explore this relationship between identification of the alternative data in the lineup and the parameter, effect size, and lineuptype with a generalized linear mixed model.
-The response variable, $Y_{ijk\ell m}$, is binary, indicating whether participant $m$ picked the alternative data plot in lineup $\ell$.
-The covariate $x_{\ell 1}$ takes on the value -1 or 1 according to the lineup type code in Table~\ref{tab:experdetail}.
-In most cases, -1 indicates the parameter value is less than original estimate, while 1 indicates the parameter value is at or above the original estimate.
-The continuous covariate $x_{\ell 2}$ is the centered and scaled size of the effect of interest from which the alternative data were simulated, the values of which are labeled ``easy", ``medium", and ``hard" in Table~\ref{tab:experdetail} according to how difficult we thought the Turk participants would find each lineup.
-In Equation~\ref{eq:glmm}, $i \in \{1, 2, 3, 4, 5\}$ corresponding to the effects $\beta_3, \beta_4, \beta_2, \beta_6, \beta_5$, respectively.
-We include random effects in the model for each lineup, $\delta_{\ell}$ and each participant, $\epsilon_m$, and fit a hierarchical model as follows:
-\begin{align}\label{eq:glmm}
-Y_{i\ell m} &\sim \text{Bernoulli}(\pi_{i\ell m}) \\
-\text{logit}(\pi_{i\ell m}) & = \mu + \alpha_{i} + \theta\mathbb{I}(x_{\ell 1} = 1) + \gamma x_{\ell 2} + (\alpha\theta)_i \mathbb{I}(x_{\ell 1} = 1) + (\alpha\gamma)_i x_{\ell 2} + (\theta\gamma)\mathbb{I}(x_{\ell 1} = 1)x_{\ell 2} + (\alpha\theta\gamma)_i\mathbb{I}(x_{\ell 1} = 1) x_{\ell 2} + \delta_{\ell} + \epsilon_{m} \\
-\delta_{\ell} & \sim N(0, \sigma^2_{\delta}) \\
-\epsilon_{m} & \sim N(0, \sigma^2_{\epsilon}) \\
-\end{align}
-The results of fitting this model using \texttt{glmer} from the \texttt{lme4} package are summarized in Table~\ref{tab:glmmests} \citep{lme4}.
+For a SAOM, there are two ways a significance test of the parameters can be performed.
+In \texttt{RSiena}, there are $t$-type tests and Wald-type test for a single parameter and for multiple parameters.
+The $t$-type test statistic is simply the parameter estimate divided by its standard error, and compared to a standard normal distribution.
+The Wald-type test statistic for a single parameter, $\beta_k$ is
+\begin{equation}\label{eq:wald1}
+\frac{(\hat{\beta_k})^2}{var(\hat{\beta_k})} \sim \chi^2_1,
+\end{equation}
+which is compared to a Chi-square distribution with one degree of freedom.
+Testing the significance of multiple parameters depends on the hypothesis we wish to test, and a $P \times K$ matrix, $A$, must be appropriately designed to test the $P$ hypotheses of interest.
+The null hypothesis is that $A\boldsymbol{\beta} = \mathbf{0}$, and the test statistic is
+\begin{equation}\label{eq:wald2}
+(A\boldsymbol{\hat{\beta}})' \hat{\Sigma}^{-1} A\boldsymbol{\hat{\beta}} \sim \chi^2_p,
+\end{equation}
+where $\hat{\Sigma}$ is the estimated covariance matrix of $\boldsymbol{\beta}$. This statistic is then compared to a Chi-square distribution with $P$ degrees of freedom.
 
-<<glmmres>>=
-load("data/finalglmm.RDA")
-mod <- model0randomscalesize
-res <- summary(mod)[[10]]
-ests <- as.numeric(round(res[,1], 4))
-odds <- round(exp(ests), 4)
-se <- as.numeric(round(res[,2], 4))
-pval <- as.numeric(round(res[,4],4))
-@
-
-\begin{table}
-\centering
-\begin{tabular}{lcccc}
-Parameter & Estimate & Std Error & $p$-value & Odds Multiplier \\
-$\mu$ & \Sexpr{ests[1]} & \Sexpr{se[1]} & \Sexpr{pval[1]} & \Sexpr{odds[1]}\\
-$\alpha_1$ & \Sexpr{ests[2]} & \Sexpr{se[2]} & \Sexpr{pval[2]} & \Sexpr{odds[2]} \\
-$\alpha_2$ & \Sexpr{ests[3]} & \Sexpr{se[3]} & \Sexpr{pval[3]} & \Sexpr{odds[3]}  \\
-$\alpha_3$  & \Sexpr{ests[4]} & \Sexpr{se[4]} & \Sexpr{pval[4]} & \Sexpr{odds[4]} \\
-$\alpha_4$ & \Sexpr{ests[5]} & \Sexpr{se[5]} & \Sexpr{pval[5]} & \Sexpr{odds[5]}  \\
-$\alpha_5$  & \Sexpr{ests[6]} & \Sexpr{se[6]} & \Sexpr{pval[6]} & \Sexpr{odds[6]} \\
-$\theta$  & \Sexpr{ests[7]} & \Sexpr{se[7]} & \Sexpr{pval[7]} & \Sexpr{odds[7]} \\
-$\gamma$  & \Sexpr{ests[8]} & \Sexpr{se[8]} & \Sexpr{pval[8]} & \Sexpr{odds[8]} \\ 
-$(\alpha\theta)_1$  & \Sexpr{ests[9]} & \Sexpr{se[9]} & \Sexpr{pval[9]} & \Sexpr{odds[9]} \\
-$(\alpha\theta)_2$  & \Sexpr{ests[10]} & \Sexpr{se[10]} & \Sexpr{pval[10]} & \Sexpr{odds[10]} \\
-$(\alpha\theta)_3$  & \Sexpr{ests[11]} & \Sexpr{se[11]} & \Sexpr{pval[11]} & \Sexpr{odds[11]} \\
-$(\alpha\theta)_4$  & \Sexpr{ests[12]} & \Sexpr{se[12]} & \Sexpr{pval[12]} & \Sexpr{odds[12]} \\
-$(\alpha\theta)_5$  & \Sexpr{ests[13]} & \Sexpr{se[13]} & \Sexpr{pval[13]} & \Sexpr{odds[13]} \\
-$(\alpha\gamma)_1$  & \Sexpr{ests[14]} & \Sexpr{se[14]} & \Sexpr{pval[14]} & \Sexpr{odds[14]} \\
-$(\alpha\gamma)_2$  & \Sexpr{ests[15]} & \Sexpr{se[15]} & \Sexpr{pval[15]} & \Sexpr{odds[15]} \\
-$(\alpha\gamma)_3$  & \Sexpr{ests[16]} & \Sexpr{se[16]} & \Sexpr{pval[16]} & \Sexpr{odds[16]} \\
-$(\alpha\gamma)_4$  & \Sexpr{ests[17]} & \Sexpr{se[17]} & \Sexpr{pval[17]} & \Sexpr{odds[17]} \\
-$(\alpha\gamma)_5$  & \Sexpr{ests[18]} & \Sexpr{se[18]} & \Sexpr{pval[18]} & \Sexpr{odds[18]} \\
-$\theta\gamma$  & \Sexpr{ests[19]} & \Sexpr{se[19]} & \Sexpr{pval[19]} & \Sexpr{odds[19]} \\
-$(\alpha\theta\gamma)_1$ & \Sexpr{ests[20]} & \Sexpr{se[20]} & \Sexpr{pval[20]} & \Sexpr{odds[20]}  \\
-$(\alpha\theta\gamma)_2$  & \Sexpr{ests[21]} & \Sexpr{se[21]} & \Sexpr{pval[21]} & \Sexpr{odds[21]} \\
-$(\alpha\theta\gamma)_3$  & \Sexpr{ests[22]} & \Sexpr{se[22]} & \Sexpr{pval[22]} & \Sexpr{odds[22]} \\
-$(\alpha\theta\gamma)_4$  & \Sexpr{ests[23]} & \Sexpr{se[23]} & \Sexpr{pval[23]} & \Sexpr{odds[23]} \\
-$(\alpha\theta\gamma)_5$  & \Sexpr{ests[24]} & \Sexpr{se[24]} & \Sexpr{pval[24]} & \Sexpr{odds[24]} \\ 
-$\sigma^2_{\delta}$ & 0.5638 \\
-$\sigma^2_{\epsilon}$ & 0.3416
-\end{tabular}
-\caption{\label{tab:glmmests}Summary of the results from fitting the model given in Equation~\ref{eq:glmm}}
-\end{table}
-
-<<predictglmm, fig.height=6, fig.cap="Predictions from our generalized linear mixed effects model given in Equation~\ref{eq:glmm}.
-The lines show the expected probability of detecting the alternative data in a lineup of size 6 for new observers of new lineups is plotted on the $y$-axis, and the size of the parameter of interest is on the $x$-axis.
-The proportions detected by our Turk participants for each lineup group are shown by the points.">>=
-newdata <- read_csv("data/newdata_pred_glmm.csv")
-plotdata <- turk22 %>% mutate(type2 = ifelse(sign == 0, sign(initialEst), sign),
-                       type2 = ifelse(type == "one", type2, -1)) %>% 
-  group_by(pic_id, size, test_param, sign, type2,alt_model) %>% summarize(
-    datapick = mean(datapick)
-)
-
-ggplot() + 
-  geom_line(data = newdata, aes(x = size, y = predictfinal, color = as.factor(type2))) + 
-  geom_point(data = plotdata, aes(x = size, y = datapick, color = as.factor(type2))) + 
-  scale_color_brewer(palette = "Dark2", name = "Lineup Type") + 
-  facet_wrap(~test_param, scales = 'free') + 
-  labs(x = "Parameter value", y = "Expected probability of detection in a lineup of size 6") + 
-  ThemeNoNet + 
-  theme(legend.position = 'bottom')
-
-#vals <- predict(model1, type = "response")
-#modelData_sig$prediction <- vals
-#ggplot(data = modelData_sig) + 
-#  geom_point(aes(x = size, y = prediction, color = as.factor(type2))) + 
-#  facet_wrap(~test_param, scales = "free")
-@
-
-\section{Goodness-of-Fit Testing}
+\subsection{Goodness-of-Fit Testing}
 
@@
 
+\subsection{Visual Detection}
+
+A summary of the results from our experiment is shown in Figure~\ref{fig:predictglmm}.
+On the $x$ axis, we plot the value of the parameter of interest, and on the $y$ axis, the proportion of times the plot of data simulated from the alternative model was picked out for each lineup.
+The results are split into groups based on the sign of the parameter.
+We can see clear patterns in all of the groups: as the parameter value approaches 0, fewer participants identified the alternative plot.
+
+<<readturk22, fig.height=6, fig.cap="A summary of some results from the Amazon mechanical turk study.
+On the $x$-axis, the value of the parameter changed in the model of interest, and on the $y$-axis, the proprotion of times turkers identified the different data plot.
+A simple linear regression is fit to each group.
+The pattern as the parameter value approaches zero is clear in all conditions.", eval = FALSE>>=
+turk22 %>% mutate(type2 = ifelse(sign == 0, sign(initialEst), sign),
+                       type2 = ifelse(type == "one", type2, -1)) %>% 
+  group_by(pic_id, size, test_param, sign, type2,alt_model) %>% 
+  summarize(datapick = mean(datapick)) %>%
+  filter(pic_id != 3132) %>% 
+  ggplot(aes(x = size, y = datapick, colour=factor(type2))) + 
+  geom_point() +
+  scale_color_brewer(palette = "Dark2", name = "Lineup Type") + 
+  facet_wrap(~test_param, scales="free", labeller = 'label_both') +
+  geom_smooth(se=FALSE, method="lm") + 
+  ThemeNoNet + 
+  theme(legend.position = 'bottom')
+@
+
+We futher explore this relationship between identification of the alternative data in the lineup and the parameter, effect size, and lineuptype with a generalized linear mixed model.
+The response variable, $Y_{ijk\ell m}$, is binary, indicating whether participant $m$ picked the alternative data plot in lineup $\ell$.
+The covariate $x_{\ell 1}$ takes on the value -1 or 1 according to the lineup type code in Table~\ref{tab:experdetail}.
+In most cases, -1 indicates the parameter value is less than original estimate, while 1 indicates the parameter value is at or above the original estimate.
+The continuous covariate $x_{\ell 2}$ is the centered and scaled size of the effect of interest from which the alternative data were simulated, the values of which are labeled ``easy", ``medium", and ``hard" in Table~\ref{tab:experdetail} according to how difficult we thought the Turk participants would find each lineup.
+In Equation~\ref{eq:glmm}, $i \in \{1, 2, 3, 4, 5\}$ corresponding to the effects $\beta_3, \beta_4, \beta_2, \beta_6, \beta_5$, respectively.
+We include random effects in the model for each lineup, $\delta_{\ell}$ and each participant, $\epsilon_m$, and fit a hierarchical model as follows:
+\begin{align}\label{eq:glmm}
+Y_{i\ell m} &\sim \text{Bernoulli}(\pi_{i\ell m}) \\
+\text{logit}(\pi_{i\ell m}) & = \mu + \alpha_{i} + \theta\mathbb{I}(x_{\ell 1} = 1) + \gamma x_{\ell 2} + \\ & (\alpha\theta)_i \mathbb{I}(x_{\ell 1} = 1) + (\alpha\gamma)_i x_{\ell 2} + (\theta\gamma)\mathbb{I}(x_{\ell 1} = 1)x_{\ell 2} + \\ & (\alpha\theta\gamma)_i\mathbb{I}(x_{\ell 1} = 1) x_{\ell 2} + \delta_{\ell} + \epsilon_{m} \\
+\delta_{\ell} & \sim N(0, \sigma^2_{\delta}) \\
+\epsilon_{m} & \sim N(0, \sigma^2_{\epsilon}) \\
+\end{align}
+The results of fitting this model using \texttt{glmer} from the \texttt{lme4} package are summarized in Table~\ref{tab:glmmests} \citep{lme4}.
+
+<<glmmres>>=
+load("data/finalglmm.RDA")
+mod <- model0randomscalesize
+res <- summary(mod)[[10]]
+ests <- as.numeric(round(res[,1], 4))
+odds <- round(exp(ests), 4)
+se <- as.numeric(round(res[,2], 4))
+pval <- as.numeric(round(res[,4],4))
+@
+
+\begin{table}
+\centering
+\begin{tabular}{lcccc}
+Parameter & Estimate & Std Error & $p$-value & Odds Multiplier \\
+$\mu$ & \Sexpr{ests[1]} & \Sexpr{se[1]} & \Sexpr{pval[1]} & \Sexpr{odds[1]}\\
+$\alpha_1$ & \Sexpr{ests[2]} & \Sexpr{se[2]} & \Sexpr{pval[2]} & \Sexpr{odds[2]} \\
+$\alpha_2$ & \Sexpr{ests[3]} & \Sexpr{se[3]} & \Sexpr{pval[3]} & \Sexpr{odds[3]}  \\
+$\alpha_3$  & \Sexpr{ests[4]} & \Sexpr{se[4]} & \Sexpr{pval[4]} & \Sexpr{odds[4]} \\
+$\alpha_4$ & \Sexpr{ests[5]} & \Sexpr{se[5]} & \Sexpr{pval[5]} & \Sexpr{odds[5]}  \\
+$\alpha_5$  & \Sexpr{ests[6]} & \Sexpr{se[6]} & \Sexpr{pval[6]} & \Sexpr{odds[6]} \\
+$\theta$  & \Sexpr{ests[7]} & \Sexpr{se[7]} & \Sexpr{pval[7]} & \Sexpr{odds[7]} \\
+$\gamma$  & \Sexpr{ests[8]} & \Sexpr{se[8]} & \Sexpr{pval[8]} & \Sexpr{odds[8]} \\ 
+$(\alpha\theta)_1$  & \Sexpr{ests[9]} & \Sexpr{se[9]} & \Sexpr{pval[9]} & \Sexpr{odds[9]} \\
+$(\alpha\theta)_2$  & \Sexpr{ests[10]} & \Sexpr{se[10]} & \Sexpr{pval[10]} & \Sexpr{odds[10]} \\
+$(\alpha\theta)_3$  & \Sexpr{ests[11]} & \Sexpr{se[11]} & \Sexpr{pval[11]} & \Sexpr{odds[11]} \\
+$(\alpha\theta)_4$  & \Sexpr{ests[12]} & \Sexpr{se[12]} & \Sexpr{pval[12]} & \Sexpr{odds[12]} \\
+$(\alpha\theta)_5$  & \Sexpr{ests[13]} & \Sexpr{se[13]} & \Sexpr{pval[13]} & \Sexpr{odds[13]} \\
+$(\alpha\gamma)_1$  & \Sexpr{ests[14]} & \Sexpr{se[14]} & \Sexpr{pval[14]} & \Sexpr{odds[14]} \\
+$(\alpha\gamma)_2$  & \Sexpr{ests[15]} & \Sexpr{se[15]} & \Sexpr{pval[15]} & \Sexpr{odds[15]} \\
+$(\alpha\gamma)_3$  & \Sexpr{ests[16]} & \Sexpr{se[16]} & \Sexpr{pval[16]} & \Sexpr{odds[16]} \\
+$(\alpha\gamma)_4$  & \Sexpr{ests[17]} & \Sexpr{se[17]} & \Sexpr{pval[17]} & \Sexpr{odds[17]} \\
+$(\alpha\gamma)_5$  & \Sexpr{ests[18]} & \Sexpr{se[18]} & \Sexpr{pval[18]} & \Sexpr{odds[18]} \\
+$\theta\gamma$  & \Sexpr{ests[19]} & \Sexpr{se[19]} & \Sexpr{pval[19]} & \Sexpr{odds[19]} \\
+$(\alpha\theta\gamma)_1$ & \Sexpr{ests[20]} & \Sexpr{se[20]} & \Sexpr{pval[20]} & \Sexpr{odds[20]}  \\
+$(\alpha\theta\gamma)_2$  & \Sexpr{ests[21]} & \Sexpr{se[21]} & \Sexpr{pval[21]} & \Sexpr{odds[21]} \\
+$(\alpha\theta\gamma)_3$  & \Sexpr{ests[22]} & \Sexpr{se[22]} & \Sexpr{pval[22]} & \Sexpr{odds[22]} \\
+$(\alpha\theta\gamma)_4$  & \Sexpr{ests[23]} & \Sexpr{se[23]} & \Sexpr{pval[23]} & \Sexpr{odds[23]} \\
+$(\alpha\theta\gamma)_5$  & \Sexpr{ests[24]} & \Sexpr{se[24]} & \Sexpr{pval[24]} & \Sexpr{odds[24]} \\ 
+$\sigma^2_{\delta}$ & 0.5638 \\
+$\sigma^2_{\epsilon}$ & 0.3416
+\end{tabular}
+\caption{\label{tab:glmmests}Summary of the results from fitting the model given in Equation~\ref{eq:glmm}}
+\end{table}
+
+<<predictglmm, fig.height=6, fig.cap="Predictions from our generalized linear mixed effects model given in Equation~\ref{eq:glmm}.
+The lines show the expected probability of detecting the alternative data in a lineup of size 6 for new observers of new lineups is plotted on the $y$-axis, and the size of the parameter of interest is on the $x$-axis.
+The proportions detected by our Turk participants for each lineup group are shown by the points.">>=
+newdata <- read_csv("data/newdata_pred_glmm.csv")
+plotdata <- turk22 %>% mutate(type2 = ifelse(sign == 0, sign(initialEst), sign),
+                       type2 = ifelse(type == "one", type2, -1)) %>% 
+  group_by(pic_id, size, test_param, sign, type2,alt_model) %>% summarize(
+    datapick = mean(datapick)
+)
+
+ggplot() + 
+  geom_line(data = newdata, aes(x = size, y = predictfinal, color = as.factor(type2))) + 
+  geom_point(data = plotdata, aes(x = size, y = datapick, color = as.factor(type2))) + 
+  scale_color_brewer(palette = "Dark2", name = "Lineup Type") + 
+  facet_wrap(~test_param, scales = 'free') + 
+  labs(x = "Parameter value", y = "Expected probability of detection in a lineup of size 6") + 
+  ThemeNoNet + 
+  theme(legend.position = 'bottom')
+
+#vals <- predict(model1, type = "response")
+#modelData_sig$prediction <- vals
+#ggplot(data = modelData_sig) + 
+#  geom_point(aes(x = size, y = prediction, color = as.factor(type2))) + 
+#  facet_wrap(~test_param, scales = "free")
+@
+
 \section{Experiment Protocol}

### 2017-10-27 ASApaper/paper-child.Rnw
move content from experiment protocol section & delete that section

@@
 An example of a lineup like those shown to our participants is shown on the right side of the image  in Figure~\ref{fig:shiny}.
-\footnote{If you would like to explore the lineups we use in further detail, please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/}}.
+In this lineup, model M4 is the alternative model, and model M1 is the null model.
+To simulate lineups from the models, we set the parameters to the values given in Table~\ref{tab:fitted values} for all parameters within the respective models, with the exception of $\beta_5$.
+For $\beta_5$, twice the estimated value was used.
+More detail on why we use twice the fitted value is provided in Section~\ref{sec:setupvd}.\footnote{If you would like to explore the kinds of lineups we use in further detail, please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/}}.
+To get the simulations, we used the \tettt{siena07} function in \texttt{RSiena} \citep{RSiena}.
 
@@
 \caption{\label{fig:shiny} A screen grab of the web application we created to design our lineup experiment.
-In this view, model M4 is selected and one estimate from model M4 with $\beta_5$ set to twice its estimated value given in Table~\ref{tab:effects} is placed at random among five observations simulated from model M1.
-The other two tabs in the right side of the screen tell you which plot comes from the alternative model, and allow you to view and download the data that is plotted in the lineup.}
+More details about this application are given in Section~\ref{sec:setupvd}.
+In the lineup, M4 is the alternative model and one estimate from model M4 with $\beta_5$ set to twice its estimated value given in Table~\ref{tab:fittedvalues} is placed at random among five observations simulated from the null model, M1.
+Which plot comes from model M4?
+%The other two tabs in the right side of the screen tell you which plot comes from the alternative model, and allow you to view and download the data that is plotted in the lineup.
+}
 \end{figure}
 
+We want to explore the role of each of these parameters in the objective functions for each model.
+
+\subsection{Significance Testing}
+
+\subsection{Goodness-of-Fit}
+\subsection{Visual Detection}\label{sec:setupvd}
+
+Through visual inference, we want to determine at which point an effect becomes noticeable in a SAOM.
+By \textit{noticeable}, we mean that the inclusion of the effect alters the appearance of networks simulated from a model.
+An effect becomes noticeable when a simulation from a model with a noticeable effect appears different from simulations from a model where the same affect is absent.
 In model M1, with only two parameters in the objective function, we varied both the density and reciprocity parameteter values one at a time.
@@
 
-\subsection{Significance Testing}
-
-\subsection{Goodness-of-Fit}
-\subsection{Visual Detection}
-
-Through visual inference, we want to determine at which point an effect becomes noticeable in a SAOM.
-By \textit{noticeable}, we mean that the inclusion of the effect alters the appearance of networks simulated from a model.
-An effect becomes noticeable when a simulation from a model with a noticeable effect appears different from simulations from a model where the same affect is absent.
 %Because the traditional way to visualize a network is a node-link diagram, we will examine the effect of different parameter values using node-link visualizations.
@@
 
-We showed participants from Amazon Mechanical Turk a series of lineups containing data simulated from M1 through M5.
+We recruited 250 participants for our experiment through Amazon Mechanical Turk.
+Each participant read over some brief training material before beginning the experiment.
+After reading this material and agreeing to participate, the participants saw two trial plots, one where they were required to select the plot that was most different from the others due to its complex structure, and one where they were required to select the plot that was most different from the others due to its very simple structure.
+Upon getting the trial plots correct, they moved into the experiment.
+Each participant was randomly assigned 13 lineups to look at.
+The were asked to select one or more plots that they perceived as ``most different" from the others, and provide a reasoning for their choice.
+They could select from ``Most simple overall structure," ``Most complex overall structure," or ``Other" and provide their own text description of their reasoning.
+One of the 13 lineups the participants saw was the true data from the 112th senate shown in Section~\ref{sec:senatedata}.
+Each participant only saw the data one in order to avoid bias.
 
@@
 
-\section{Experiment Protocol}
-
-We recruited 250 participants for our experiment through Amazon Mechanical Turk.
-Each participant read over some brief training material before beginning the experiment.
-After reading this material and agreeing to participate, the participants saw two trial plots, one where they were required to select the plot that was most different from the others due to its complex structure, and one where they were required to select the plot that was most different from the others due to its very simple structure.
-Upon getting the trial plots correct, they moved into the experiment.
-Each participant was randomly assigned 13 lineups to look at.
-The were asked to select one or more plots that they perceived as ``most different" from the others, and provide a reasoning for their choice.
-They could select from ``Most simple overall structure," ``Most complex overall structure," or ``Other" and provide their own text description of their reasoning.
-One of the 13 lineups the participants saw was the true data from the 112th senate shown in Section~\ref{sec:senatedata}.
-Each participant only saw the data one in order to avoid bias.
-
 \section{Discussion}

### 2017-10-28 ASApaper/paper-child.Rnw
add pdf
finish section exper setup for sig testing

@@
 
-We use the estimates given in Table~\ref{tab:fittedvalues} to simulate from models M1 through M6. We discuss the simulation procedure and how we use the simulations in Section~\ref{sec:expersetup}.
+We want to explore the role of each of these parameters in the objective functions for each model.
+So, we use the estimates given in Table~\ref{tab:fittedvalues} to simulate from models M1 through M6.
+We discuss the simulation procedure and how we use the simulations in Section~\ref{sec:expersetup}.
 
@@
 More detail on why we use twice the fitted value is provided in Section~\ref{sec:setupvd}.\footnote{If you would like to explore the kinds of lineups we use in further detail, please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/}}.
-To get the simulations, we used the \tettt{siena07} function in \texttt{RSiena} \citep{RSiena}.
+To get the simulations, we used the \texttt{siena07} function in \texttt{RSiena} \citep{RSiena}.
 
@@
 
-We want to explore the role of each of these parameters in the objective functions for each model.
+To explore stochastic actor-oriented models through visual inference, we show a series of lineups to independent observers recruited through Amazon Mechanical Turk.
+We provide more details on the Turk setup in Section~\ref{sec:res}.
 
 \subsection{Significance Testing}
+
+In the significance testing protocol, a parameter of interest is selected to test, say $\beta_k$. The hypotheses we use to generate lineups are:
+
+\begin{description}
+\item $H_0$: $\beta_k = 0 $
+\item $H_A$: $\beta_k \neq $
+\end{description}
+
+Under the null hypothesis, we assume that the model that generated the network data is the simplest model presented in Section~\ref{sec:models}, model M1.
+Thus, the five null plots in the lineup are simulations from M1 with $\beta_1, \beta_2$ set to the estimates given in Table~\ref{tab:fittedvalues} for M1.
+The alternate model is the model with $\beta_1, \beta_2$, and $\beta_k$ in the objective function, with the remaining plot simulated from the appropriate model.
+
+The lineup generated under this scenario is shown to independent viewers.
+If the viewers pick out the alternative data plot, that is evidence in favor of the null hypothesis, while picking one of the null plots is evidence against the null hypothesis.
+The significance tests that we perform in our experiment are for $\beta_3$ and $\beta_4$, making the alternative models M3 and M4, respectively.
 
@@
 
-\section{Experiment Results}
+\section{Experiment Results}\label{sec:res}
 

### 2017-10-28 ASApaper/paper-child.Rnw
add rest of setup for today.

@@
 In this lineup, model M4 is the alternative model, and model M1 is the null model.
-To simulate lineups from the models, we set the parameters to the values given in Table~\ref{tab:fitted values} for all parameters within the respective models, with the exception of $\beta_5$.
+To simulate lineups from the models, we set the parameters to the values given in Table~\ref{tab:fittedvalues} for all parameters within the respective models, with the exception of $\beta_5$.
 For $\beta_5$, twice the estimated value was used.
@@
 
-\subsection{Significance Testing}
+\subsection{Significance Testing}\label{sec:setupst}
 
@@
 If the viewers pick out the alternative data plot, that is evidence in favor of the null hypothesis, while picking one of the null plots is evidence against the null hypothesis.
-The significance tests that we perform in our experiment are for $\beta_3$ and $\beta_4$, making the alternative models M3 and M4, respectively.
+The significance tests that we perform in our experiment are for $\beta_3$ and $\beta_4$, making the alternative models M2 and M3, respectively.
 
 \subsection{Goodness-of-Fit}
+
+For the goodness-of-fit tests, we compare one model of interest, say M\textit{i} to the data. The hypothesis we use to generate lineups are 
+\begin{description}
+\item $H_0$: The data come from model M\textit{i}
+\item $H_A$: The data come from some other, unknown model
+\end{description}
+
+To generate the null plots, we simulate five networks from model M\textit{i} using the corresponding parameters in Table~\ref{tab:fittedvalues}.
+We pick a wave to focus on, wave two, which is the first simulated network, and among these five plots, we place the true second wave data.
+We cannot show the data more than once to each participant, so we examine several different models in our Amazon Mechanical Turk experiment, each participant never seeing the true data wave twice.
+The models we chose for goodness-of-fit testing are M2, M3, M4, and M6.
+
 \subsection{Visual Detection}\label{sec:setupvd}
 
+Visual detection does not fit into the typical statistical testing framework as easily as the significance and goodness-of-fit tests we perform do.
 Through visual inference, we want to determine at which point an effect becomes noticeable in a SAOM.
 By \textit{noticeable}, we mean that the inclusion of the effect alters the appearance of networks simulated from a model.
-An effect becomes noticeable when a simulation from a model with a noticeable effect appears different from simulations from a model where the same affect is absent.
+An effect becomes noticeable when a simulation from a model with a noticeable effect appears different from simulations from a model where the same affect is absent when they simulations are placed in a lineup.
+We explore all parameters in the objective function, $\beta_1, \dots, \beta_6$ in this way.
+
 In model M1, with only two parameters in the objective function, we varied both the density and reciprocity parameteter values one at a time.
@@
 
-<<hypothesis, fig.height=2, fig.cap="We hypothesize that as the parameter value of interest increases in absolute value, more viewers of the lineup will pick the alternative data out of a lineup.">>=
+<<hypothesis, fig.height=2, fig.cap="We hypothesize that as the parameter value of interest increases in absolute value, more viewers of the lineup will pick the alternative data out of a lineup.
+Note that the significance test we construct in Section~\\ref{sec:setupst} is just one point on the line below, represented by the vertical red line.">>=
 x <- seq(-10, 10, .05)
@@
     scale_x_continuous(name = "Parameter Value", labels = c("very negative", "negative", "0", "positive", "very positive")) + 
+  #geom_hline(yintercept = 1/6) +
+    geom_vline(xintercept = 150, color = 'red', linetype = 'dotdash') + 
     scale_y_continuous(name = "% detecting plot from\nthe alternative model", labels = paste0(c(0,25,50,75,100), "%"))
@@
 A screen grab of the app is shown in Figure~\ref{fig:shiny}.
-On the left side of the screen, the user\footnote{Please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/} to create lineups for yourself.} can input the information necessary for creating a lineup of the models M1 through M6 for the data in Section~\ref{sec:senatedata}: first, choose to simulate only one plot from the specified model (analogous to changing the alternative model in the lineup protocol) or to simulated $M-1$ plots from the specified model (analogous to changing the null model in the lineup protocol); the model of interest; the wave of the data to examine; if model M1 is selected, whether to alter the density or the reciprocity parameter; the size of the lineup; the amount by which to multiply the effect selected; a random seed for replicability; and a layout algorithm to use for the node-link diagrams.
-The plot(s) that appear(s) that are \textit{not} from the model specified using the other options are simulations from model M1 with the estimates of the rate parameters, $\beta_1$ and $\beta_2$ as calculated by the procedure described in Section~\ref{sec:models}.
+On the left side of the screen, the user\footnote{Please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/} to create lineups constructed from the models we present for this data for yourself.} can input the information necessary for creating a lineup of the models M1 through M6 for the data in Section~\ref{sec:senatedata}: first, choose to simulate only one plot from the specified model (analogous to changing the alternative model in the lineup protocol) or to simulated $M-1$ plots from the specified model (analogous to changing the null model in the lineup protocol); the model of interest; the wave of the data to examine; if model M1 is selected, whether to alter the density or the reciprocity parameter; the size of the lineup; the amount by which to multiply the effect selected; a random seed for replicability; and a layout algorithm to use for the node-link diagrams.
+There is also a checkbox if the user wishes the nodes to be colored by the size of the connected component to which they belong.
+The plots that appear that are \textit{not} from the model specified using the other options are simulations from model M1 with the estimates of the rate parameters, $\beta_1$ and $\beta_2$ given in Table~\ref{tab:fittedvalues}.
 
@@
 In the case of both $\beta_4$ and $\beta_6$, we could not determine a value for negative effects that made the data simulated from M3 and M5 look different than null model simulations from model M1.
-Because we could not detect these effects at these values, we decided that the less experienced Turkers would also not be able to.
+Because we could not detect these effects at these values, we decided that the less experienced participants in our experiment would also not be able to.
 So instead of testing the negative values of these effects, we examined a different scenario: we placed 5 simulations from positive values of the parameter with one simulation from model M1 in a lineup.
-We did this to determine if the perception of the effect size would be symmetric: if an effect is noticed $x$\% of the time at value $\beta = 5$ when a simulation from the model is placed among null plots, then when five simulations from the model are put together with one simulation from model M1, the plot from the simpler model should be noticed about $x$\% of the time as well.
+We did this to determine if the perception of the effect size would be symmetric: if an effect is noticed $x$\% of the time at value $\beta_k = \beta_{k_0}$ when one simulation from the corresponding model is placed among five null plots from model M1, then when five simulations from the model with $\beta_k = \beta_{k_0}$ are put in a lineup with one simulation from model M1, the plot from the simpler model should be noticed about $x$\% of the time as well.
 
@@
 One of the 13 lineups the participants saw was the true data from the 112th senate shown in Section~\ref{sec:senatedata}.
-Each participant only saw the data one in order to avoid bias.
+Each participant only saw the data one time in order to avoid bias.
+Upon completion of the 13 lineups, each participant was paid \$1.75.
 
@@
 
-For a SAOM, there are two ways a significance test of the parameters can be performed.
+For a SAOM, there are two ways a conventional significance test of the parameters can be performed.
 In \texttt{RSiena}, there are $t$-type tests and Wald-type test for a single parameter and for multiple parameters.

### 2017-10-29 ASApaper/paper-child.Rnw
mini changes

@@
 
-We want to explore three different aspects of the SAOM models using the lineup protocol: significance of parameters, goodness-of-fit of a model, and visual detection of additional parameters.
-Each on of these requires a different setup, which we describe in detail, but we use the lineup protocol to examine each aspect.
-In each lineup, there are two models shown: the null model and the alternative model.
-The definition of the null and alternative model will vary according to which aspects of the SAOMs we are exploring.
+We want to explore three different aspects of the SAOM models using the lineup protocol: (1) significance of parameters, (2) goodness-of-fit of a model, and (3) visual detection of parameters. 
+Each one of these situations requires a different setup, which we describe in detail, but we make ue of the lineup protocol for all of these aspects.
+
+In each lineup, we include plots from two models: the null model and an alternative model. The definition of the null and alternative model varies with the aspect of the SAOMs we are exploring. 
 %For each lineup, we consider two models: model M1 is the null model, and another model, from M2-M5, which is chosen to be the alternative model.
 
+ Typically, a lineup shows sets of 20 plots at a time c.f.~\citet{loy:2015, vanderplas:2016}, but we determined that not enough structure could be shown in each plot for 20 node-link diagrams.
+We chose to expose our participants to only six plots at a time in order to show the node-link diagrams in more detail and to lower cognitive load for participants.
 To construct a lineup, we simulate five networks from the null model and one network from the alternative model.
-We chose to expose our participants to only six plots at a time in order to show the node-link diagrams in more detail and to lower their cognitive load.
-Typically, lineup experiments are done with sets of 20 plots at a time c.f.~\citet{loy:2015, vanderplas:2016}, but we determined that not enough structure could be shown in each plot for 20 node-link diagrams.
-An example of a lineup like those shown to our participants is shown on the right side of the image  in Figure~\ref{fig:shiny}.
+An example of a lineup like those shown to our participants is given on the right side of the image  in Figure~\ref{fig:shiny}.
 In this lineup, model M4 is the alternative model, and model M1 is the null model.
+
 To simulate lineups from the models, we set the parameters to the values given in Table~\ref{tab:fittedvalues} for all parameters within the respective models, with the exception of $\beta_5$.
@@
 \includegraphics[width=.9\textwidth]{img/shinyappscreen.png}
-\caption{\label{fig:shiny} A screen grab of the web application we created to design our lineup experiment.
+\caption{\label{fig:shiny} A screen shot of the web application we created to design our lineup experiment.
 More details about this application are given in Section~\ref{sec:setupvd}.
-In the lineup, M4 is the alternative model and one estimate from model M4 with $\beta_5$ set to twice its estimated value given in Table~\ref{tab:fittedvalues} is placed at random among five observations simulated from the null model, M1.
-Which plot comes from model M4?
+In the lineup, M4 is the alternative model with $\beta_5$ set to twice its estimated value given in Table~\ref{tab:fittedvalues}.
+One plot simulated from this model is placed at random among five observations simulated from the null model, M1.
+Participants of the study are asked to identify the most different plot.
 %The other two tabs in the right side of the screen tell you which plot comes from the alternative model, and allow you to view and download the data that is plotted in the lineup.
@@
 \item $H_0$: $\beta_k = 0 $
-\item $H_A$: $\beta_k \neq $
+\item $H_A$: $\beta_k \neq 0$
 \end{description}
@@
 
-The lineup generated under this scenario is shown to independent viewers.
-If the viewers pick out the alternative data plot, that is evidence in favor of the null hypothesis, while picking one of the null plots is evidence against the null hypothesis.
+The lineup generated under this scenario is shown to a number of independent viewers.
+If an observer picks out the alternative data plot, that is evidence in favor of the null hypothesis, while picking one of the null plots is evidence against the null hypothesis.
 The significance tests that we perform in our experiment are for $\beta_3$ and $\beta_4$, making the alternative models M2 and M3, respectively.
@@
 To generate the null plots, we simulate five networks from model M\textit{i} using the corresponding parameters in Table~\ref{tab:fittedvalues}.
-We pick a wave to focus on, wave two, which is the first simulated network, and among these five plots, we place the true second wave data.
+We pick a wave to focus on, wave two, which is the first simulated network, and among these five plots, we place a node-link diagram of the true second wave data.
 We cannot show the data more than once to each participant, so we examine several different models in our Amazon Mechanical Turk experiment, each participant never seeing the true data wave twice.
@@
 By \textit{noticeable}, we mean that the inclusion of the effect alters the appearance of networks simulated from a model.
-An effect becomes noticeable when a simulation from a model with a noticeable effect appears different from simulations from a model where the same affect is absent when they simulations are placed in a lineup.
+We define an effect to be noticeable, if viewers are able to reliably pick out a node-link diagram rendered from data simulated from a model with the effect from a lineup of plots without the effect.
 We explore all parameters in the objective function, $\beta_1, \dots, \beta_6$ in this way.
@@
 
-To determine the threshold at which the effect becomes noticeable, we examined six different levels of the effect, three negative and three positive.
+To determine the threshold at which the effect becomes noticeable, we examined six different levels of the effect, three negative and three positive ones.
 We hypothesized that the perception of the effect by our participants would look something like Figure~\ref{fig:hypothesis}.
@@
 In order to determine which specific values of the parameters we wanted to examine, we constructed an online application that created the lineup protocol for us to be the guinea pigs of our own experiment \citep{shiny}.
-A screen grab of the app is shown in Figure~\ref{fig:shiny}.
+A screen shot of the app is shown in Figure~\ref{fig:shiny}.
 On the left side of the screen, the user\footnote{Please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/} to create lineups constructed from the models we present for this data for yourself.} can input the information necessary for creating a lineup of the models M1 through M6 for the data in Section~\ref{sec:senatedata}: first, choose to simulate only one plot from the specified model (analogous to changing the alternative model in the lineup protocol) or to simulated $M-1$ plots from the specified model (analogous to changing the null model in the lineup protocol); the model of interest; the wave of the data to examine; if model M1 is selected, whether to alter the density or the reciprocity parameter; the size of the lineup; the amount by which to multiply the effect selected; a random seed for replicability; and a layout algorithm to use for the node-link diagrams.

### 2017-10-30 ASApaper/paper-child.Rnw
commit changes from earlier in the day

--- 2017-10-28 ASApaper/paper-child.Rnw
@@
 So instead of testing the negative values of these effects, we examined a different scenario: we placed 5 simulations from positive values of the parameter with one simulation from model M1 in a lineup.
-We did this to determine if the perception of the effect size would be symmetric: if an effect is noticed $x$\% of the time at value $\beta_k = \beta_{k_0}$ when one simulation from the corresponding model is placed among five null plots from model M1, then when five simulations from the model with $\beta_k = \beta_{k_0}$ are put in a lineup with one simulation from model M1, the plot from the simpler model should be noticed about $x$\% of the time as well.
+We refer to this later on as the ``reverse" scenario.
+We used the reverse scenario to determine if the perception of the effect size would be symmetric: if an effect is noticed $x$\% of the time at value $\beta_k = \beta_{k_0}$ when one simulation from the corresponding model is placed among five null plots from model M1, then when five simulations from the model with $\beta_k = \beta_{k_0}$ are put in a lineup with one simulation from model M1, the plot from the simpler model should be noticed about $x$\% of the time as well.
 
@@
 where $\hat{\Sigma}$ is the estimated covariance matrix of $\boldsymbol{\beta}$. This statistic is then compared to a Chi-square distribution with $P$ degrees of freedom.
+
+All of the parameters we test for significance using the lineup protocol, $\beta_3$ and $\beta_4$, were determined to be statistically significant using Equation~\ref{eq:wald1}.
+The results from the significance tests we performed using the lineup protocol are given in Table~\ref{tab:sigtesting}
+
+<<sigtesting, results='asis'>>=
+turk22_sig <- turk22 %>% 
+  mutate(type2 = ifelse(sign == 0, sign(initialEst), sign),
+         type2 = ifelse(type == "one", type2, -1)) %>% 
+  filter(param_value == 1)
+summ_sig <- turk22_sig %>% group_by(pic_id, test_param) %>% 
+  summarize(npick = sum(datapick), n = n(), 
+            pv = map2_dbl(npick, n, pV, m=6, scenario=3))
+names(summ_sig) <- c("Lineup ID", "parameter", "# Alt. Model Picks", "Total Views", "p-value")
+summ_sig$parameter <- as.factor(summ_sig$parameter)
+levels(summ_sig$parameter) <- c("beta3", "beta4")
+print(xtable::xtable(summ_sig, caption = "Experiment results for the two parameters for which we performed significance tests.
+There were three lineups for each parameter, so there are three results for each plot.", label = "tab:sigtesting", digits = c(0,0,0,0,0,5)), include.rownames = FALSE)
+@
 

### 2017-10-30 ASApaper/paper-child.Rnw
respond to hh's comments/questions

@@
 Even more challenging is the situation when we go beyond single instances of a network and consider the dynamics of network change between observed instances.
-This type of modelling for dynamic networks is often performed on social network data, such as for friendship networks among students or the spread of HIV in  sexually active communities.% of men sleeping with other men.
-These models lack the asymptotics required to perform many well-known goodness-of-fit tests, and the maximum likelihood estimation of parameters is so difficult that it can make significance testing significant as well \citep{goldenberg09}.
-\hh{is the previous sentence a quote? I'm not sure that I follow.}
+This type of modelling for dynamic networks is often performed on social network data, such as friendship networks among students or the spread of HIV in \st{drug users sharing needles.} %sexually active communities.% of men sleeping with other men.
+X ST a bit less jarring hopefully.
+X
+These models lack the asymptotics required to perform many well-known goodness-of-fit tests, and the maximum likelihood estimation of parameters is so difficult that it can make significance testing difficult as well \citep{goldenberg09}.
+\hh{is the previous sentence a quote? I'm not sure that I follow.} \st{it's a paraphrase... shall I make a quote? Also I definitely botched a word. Hopefully makes more sense now? }
 
@@
 The paper is outlined as follows: Section~\ref{vi} gives a basic overview of visual inference and the lineup protocol.
-Section 3 provides an introduction to the our models of interest, stochastic actor-oriented models.
-Sections 4 and 5 discuss problems with traditional significance tests and goodness-of-fit measures for these models and how visual inference methods can be used as solutions.
-Section 6 outlines our experimental protocol, and we close with a discussion in Section 7.
+Section~\ref{sec:saom} provides an introduction to the our models of interest, stochastic actor-oriented models.
+Section~\ref{sec:expersetup} details how we define significance testing and goodness of fit procedures for SAOMs through visual inference, and Section~\ref{sec:res} details the results of a visual inference survey of Amazon Mechanical Turk workers.
+We close with a discussion in Section~\ref{sec:concl}.
 
-\hh{where do the results come in?}
+\hh{where do the results come in?}\st{whoops, forgot to updat this with the reorg. }
 
@@
 If an observer identifies the data plot, this is quantifiable evidence against the null hypothesis.
-\hh{Similarly, an observer has a chance of 1 in $m$ to pick the data plot from the lineup by simply guessing, ie in a situation where the data plot is virtually indistinguishable from the null plots.}
+\hh{Since an observer has a chance of 1 in $m$ to pick the data plot from the lineup by simply guessing, i.e.
+in a situation where the data plot is virtually indistinguishable from the null plots,} \st{the evidence grows in strength with the number of independent observers identifying the data plot.}
 
@@
 
-\section{Stochastic Actor-Oriented Models}
+\section{Stochastic Actor-Oriented Models}\label{sec:saom}
 
@@
 After actor $i$ has been given the opportunity to change, it probabilistically chooses one of its current ties, $x_{ij}$, to change.
-The probability that actor $i$ changes its current tie to actor $j$ is determined by the \textit{objective function} of the model and a random component, $U$.
-\hh{XXX can you paraphrase the purpose of U in half a sentence?}
+The probability that actor $i$ changes its current tie to actor $j$ is determined by the \textit{objective function} of the model and a random component, $U$, \st{which can be thought of as encompassing any other factors that may be influencing the changes the node makes not accounted for by the parameters in the model.} \hh{XXX can you paraphrase the purpose of U in half a sentence?}
 %The random component is assumed to have log-Weibull distribution with location parameter $\mu = 0$ and scale parameter $\sigma = 0$ (see \citet{modelsSnijders}), so the probability density function of $U$ is
@@
 
-This probability dictates which edge change is made by the acting node. The acting node can also choose to \textit{not} change at all, in which case $j \equiv i$.
+This probability dictates which edge change is made by the acting node.
+The acting node can also choose to \textit{not} change at all.
+This occurs when the numerator, as calculated for the current state of the network, is larger than for any changes that could be made.
 
-\hh{XXX how is the probability computed in the case that no change is made? Why is the notation $j \equiv i$ used for no change?}
+\hh{XXX how is the probability computed in the case that no change is made? Why is the notation $j \equiv i$ used for no change?}\st{that's more for the computation. I'll adjust accordingly}
 
@@
 
-\section{Discussion}
+\section{Discussion}\label{sec:concl}
 

### 2017-10-30 ASApaper/paper-child.Rnw
add expanded recip figure, add expected values for glmm

@@
 We futher explore this relationship between identification of the alternative data in the lineup and the parameter, effect size, and lineuptype with a generalized linear mixed model.
-The response variable, $Y_{ijk\ell m}$, is binary, indicating whether participant $m$ picked the alternative data plot in lineup $\ell$.
+The response variable, $Y_{i\ell m}$, is binary, indicating whether participant $m$ picked the alternative data plot in lineup $\ell$.
 The covariate $x_{\ell 1}$ takes on the value -1 or 1 according to the lineup type code in Table~\ref{tab:experdetail}.
@@
 We include random effects in the model for each lineup, $\delta_{\ell}$ and each participant, $\epsilon_m$, and fit a hierarchical model as follows:
-\begin{align}\label{eq:glmm}
+\begin{align}
+\begin{split}
 Y_{i\ell m} &\sim \text{Bernoulli}(\pi_{i\ell m}) \\
-\text{logit}(\pi_{i\ell m}) & = \mu + \alpha_{i} + \theta\mathbb{I}(x_{\ell 1} = 1) + \gamma x_{\ell 2} + \\ & (\alpha\theta)_i \mathbb{I}(x_{\ell 1} = 1) + (\alpha\gamma)_i x_{\ell 2} + (\theta\gamma)\mathbb{I}(x_{\ell 1} = 1)x_{\ell 2} + \\ & (\alpha\theta\gamma)_i\mathbb{I}(x_{\ell 1} = 1) x_{\ell 2} + \delta_{\ell} + \epsilon_{m} \\
+\text{logit}(\pi_{i\ell m}) & = \mu + \alpha_{i} + \theta\mathbb{I}(x_{\ell 1} = 1) + \gamma x_{\ell 2} + \\ & (\alpha\theta)_i \mathbb{I}(x_{\ell 1} = 1) + (\alpha\gamma)_i x_{\ell 2} + (\theta\gamma)\mathbb{I}(x_{\ell 1} = 1)x_{\ell 2} + \\ & (\alpha\theta\gamma)_i\mathbb{I}(x_{\ell 1} = 1) x_{\ell 2} + \delta_{\ell} + \epsilon_{m} \label{eq:glmm} \\
 \delta_{\ell} & \sim N(0, \sigma^2_{\delta}) \\
-\epsilon_{m} & \sim N(0, \sigma^2_{\epsilon}) \\
+\epsilon_{m} & \sim N(0, \sigma^2_{\epsilon})
+\end{split}
 \end{align}
 The results of fitting this model using \texttt{glmer} from the \texttt{lme4} package are summarized in Table~\ref{tab:glmmests} \citep{lme4}.
+The baseline condition for this model against which all other experimental conditions are compared is the first condition in Table~\ref{tab:experdetail}, for parameter $\beta_1$ and lineup type -1.
+The expected value of the link function for this scenario with $\beta_1 = 0$ is $\mu$, and the expected probability that a new observer will pick out the data plot in a new lineup generated from this scenario is $\frac{\exp(\mu)}{1+\exp{\mu}}$.
+Similarly, for a new lineup generated from the tenth condition in in Table~\ref{tab:experdetail}, with $\beta_6 = 1$, and lineup type 1, the expected value of the link function is $\mu + \alpha_{4} + \theta + \gamma + (\alpha\theta)_4 + (\alpha\gamma)_4  + (\theta\gamma) + (\alpha\theta\gamma)_4$.
+The corresponding expected probability that a new observer viewing a new lineup from this scenario is:
+\begin{equation}
+\frac{\exp\{\mu + \alpha_{4} + \theta + \gamma + (\alpha\theta)_4 + (\alpha\gamma)_4  + (\theta\gamma) + (\alpha\theta\gamma)_4\}}{1 + \exp\{\mu + \alpha_{4} + \theta + \gamma + (\alpha\theta)_4 + (\alpha\gamma)_4  + (\theta\gamma) + (\alpha\theta\gamma)_4\}}
+\end{equation}
 
@@
 
-<<predictglmm, fig.height=6, fig.cap="Predictions from our generalized linear mixed effects model given in Equation~\ref{eq:glmm}.
+<<predictglmm, fig.height=6, fig.cap='Predictions from our generalized linear mixed effects model given in Equation~\ref{eq:glmm}.
 The lines show the expected probability of detecting the alternative data in a lineup of size 6 for new observers of new lineups is plotted on the $y$-axis, and the size of the parameter of interest is on the $x$-axis.
-The proportions detected by our Turk participants for each lineup group are shown by the points.">>=
+The proportions detected by our Turk participants for each lineup group are shown by the points.
+The lineup marked as ``outlier" was removed from modeling.
+The ``recip" panel is also presented in Figure~\\ref{fig:recipzoom} in more detail.'>>=
 newdata <- read_csv("data/newdata_pred_glmm.csv")
@@
 
+labdat <- data_frame(test_param = "jttp", type2 = -1, x = -5, y = .9, label = "outlier")
+
 ggplot() + 
@@
   geom_point(data = plotdata, aes(x = size, y = datapick, color = as.factor(type2))) + 
+  geom_text(data = labdat, aes(x = x, y=y, label = label, color = as.factor(type2)),
+            show.legend = F) + 
   scale_color_brewer(palette = "Dark2", name = "Lineup Type") + 
-  facet_wrap(~test_param, scales = 'free') + 
+  facet_wrap(~test_param, scales = 'free_x') + 
   labs(x = "Parameter value", y = "Expected probability of detection in a lineup of size 6") + 
@@
 @
+
+<<recipzoom, fig.height=2.5, fig.cap='The bottom left panel of Figure~\\ref{fig:predictglmm} expanded to show greater detail.
+For this parameter, as its value approaches zero, the probability of identifying the alternate data model decreases, then increases, which is noticeably different from the pattern exhibited by the others.'>>=
+ggplot() + 
+  geom_line(data = newdata %>% filter(test_param == "recip"), aes(x = size, y = predictfinal, color = as.factor(type2))) + 
+  geom_point(data = plotdata %>% filter(test_param == "recip"), aes(x = size, y = datapick, color = as.factor(type2))) + 
+  scale_color_brewer(palette = "Dark2", name = "Lineup Type") + 
+  facet_wrap(~type2, scales = 'free_x') + 
+  labs(x = "Parameter value (recip)", y = "Expected probability of detection in a lineup of size 6") + 
+  ThemeNoNet + 
+  theme(legend.position = 'bottom')
+@
+
+In Figure~\ref{fig:predictglmm}
 

### 2017-10-30 ASApaper/paper-child.Rnw
couple more mini changes

@@
 
-To explore stochastic actor-oriented models through visual inference, we show a series of lineups to independent observers recruited through Amazon Mechanical Turk.
-We provide more details on the Turk setup in Section~\ref{sec:res}.
+A series of these lineups is shown to independent observers recruited through Amazon Mechanical Turk for feedback (more details on the Turk setup in Section~\ref{sec:res}).
 
@@
 
-\begin{description}
-\item $H_0$: $\beta_k = 0 $
-\item $H_A$: $\beta_k \neq 0$
-\end{description}
+\begin{equation}\label{eq:sig}
+H_0: \beta_k = 0 \ \ \ \text{ versus } \ \ \ H_A: \beta_k \neq 0
+\end{equation}
+%\begin{description}
+%\item $H_0$: $\beta_k = 0 $
+%\item $H_A$: $\beta_k \neq 0$
+%\end{description}
 
-Under the null hypothesis, we assume that the model that generated the network data is the simplest model presented in Section~\ref{sec:models}, model M1.
+Under the null hypothesis, we assume that the model that generated the network data is $M_1$, the simplest model presented in Section~\ref{sec:models}.
 Thus, the five null plots in the lineup are simulations from M1 with $\beta_1, \beta_2$ set to the estimates given in Table~\ref{tab:fittedvalues} for M1.

### 2017-10-30 ASApaper/paper-child.Rnw
start of experiment results

@@
 Through visual inference, we want to determine at which point an effect becomes noticeable in a SAOM.
-By \textit{noticeable}, we mean that the inclusion of the effect alters the appearance of networks simulated from a model.
-We define an effect to be noticeable, if viewers are able to reliably pick out a node-link diagram rendered from data simulated from a model with the effect from a lineup of plots without the effect.
+By \textit{noticeable}, we mean that the inclusion of the effect alters the appearance of networks simulated from a model to a degree that viewers are able to reliably pick out a node-link diagram rendered from data simulated from a model with the effect from a lineup of plots without the effect.
 We explore all parameters in the objective function, $\beta_1, \dots, \beta_6$ in this way.
 
-In model M1, with only two parameters in the objective function, we varied both the density and reciprocity parameteter values one at a time.
+In model M1, with only two parameters in the objective function, we varied both the density and reciprocity parameter values one at a time.
 In models M2 through M5, we vary the additional parameter, $\beta_3$ through $\beta_6$.
 Thus, we have six different parameters of interest to us: $\beta_1, \dots, \beta_6$.
-We want to determine how the size of these parameters affects the overall structure of the network data simulated from the models M1 through M5, so we also vary the value of the parameters in two directions: positive and negative.
+We want to determine how the size of these parameters affects the overall structure of the network data simulated from the models M1 through M5, so we also vary the value of the parameters in both negative and positive directions.
 
@@
 
-To determine the threshold at which the effect becomes noticeable, we examined six different levels of the effect, three negative and three positive ones.
-We hypothesized that the perception of the effect by our participants would look something like Figure~\ref{fig:hypothesis}.
-The higher in absolute value the parameter is, the more likely participants are to choose the alternative model out of the lineup.
-To determine the six levels of the effects we want to test, started with the estimates of the parameter at hand as determined by the procedure described in Section~\ref{sec:models}, and multiply it or changing its sign to determine what \textit{we} notice about the effect of the parameter in simulations from the changed models.
+To determine the threshold at which an effect becomes noticeable, we examine six different levels of the effect, three negative and three positive ones.
+Figure~\ref{fig:hypothesis} shows a sketch of what the detection probability by participants' looks like hypothetically with varying effect size: the higher in absolute value the parameter is, the more likely participants are to choose the alternative model out of the lineup.
+To determine the exact values of the six levels we want to test for each effect, we started with the estimates of the parameter at hand (see  Section~\ref{sec:models}), and used small negative and positive factors to determine at what point \textit{we} noticed the effect of the parameter in simulations from the changed models.
 
-In order to determine which specific values of the parameters we wanted to examine, we constructed an online application that created the lineup protocol for us to be the guinea pigs of our own experiment \citep{shiny}.
+For this we constructed an online application \citep{shiny} that created lineups to test ourselves as guinea pigs of our own experiment \hh{XXX a reference for the quantified self would be good here}.
 A screen shot of the app is shown in Figure~\ref{fig:shiny}.
@@
 The complete details of the parameters tested using the lineup protocol is given in Table~\ref{tab:experdetail}.
-In the case of both $\beta_4$ and $\beta_6$, we could not determine a value for negative effects that made the data simulated from M3 and M5 look different than null model simulations from model M1.
-Because we could not detect these effects at these values, we decided that the less experienced participants in our experiment would also not be able to.
-So instead of testing the negative values of these effects, we examined a different scenario: we placed 5 simulations from positive values of the parameter with one simulation from model M1 in a lineup.
-We refer to this later on as the ``reverse" scenario.
-We used the reverse scenario to determine if the perception of the effect size would be symmetric: if an effect is noticed $x$\% of the time at value $\beta_k = \beta_{k_0}$ when one simulation from the corresponding model is placed among five null plots from model M1, then when five simulations from the model with $\beta_k = \beta_{k_0}$ are put in a lineup with one simulation from model M1, the plot from the simpler model should be noticed about $x$\% of the time as well.
+In the case of both $\beta_4$ and $\beta_6$, we could not determine any values for negative effects that made the data simulated from M3 and M5 look different than null model simulations from model M1.
+Therefore we decided that the lesser experienced participants in our experiment would also not be able to.
+Instead of testing the negative values of these effects, we are examining a different scenario: we placed 5 simulations from positive values of the parameter with one simulation from model M1 in a lineup.
+We refer to this later on as the ``reverse" lineup scenario.
+We used the reverse scenario to determine if the perception of the effect size is symmetric: if an effect is noticed $x$\% of the time at value $\beta_k = \beta_{k_0}$ when one simulation from the corresponding model is placed among five null plots from model M1, then when five simulations from the model with $\beta_k = \beta_{k_0}$ are put in a lineup with one simulation from model M1, the plot from the simpler model should be noticed about $x$\% of the time as well.
 
@@
 We recruited 250 participants for our experiment through Amazon Mechanical Turk.
-Each participant read over some brief training material before beginning the experiment.
-After reading this material and agreeing to participate, the participants saw two trial plots, one where they were required to select the plot that was most different from the others due to its complex structure, and one where they were required to select the plot that was most different from the others due to its very simple structure.
-Upon getting the trial plots correct, they moved into the experiment.
+Each participant was peresented with some brief training material before beginning the experiment.
+After agreeing to participate, the participants were shown two trial plots, one where the data plot was the most different from the others due to its complex structure, while the other trial included a data plot that was most different from the others due to its very simple structure.
+Only when participants were able to correctly identify the data plot from the trial lineups, they were allowed to begin the experiment.
 Each participant was randomly assigned 13 lineups to look at.
@@
 They could select from ``Most simple overall structure," ``Most complex overall structure," or ``Other" and provide their own text description of their reasoning.
+\hh{what about the first 12 plots? 6 parameters, 3 levels, 2 directions}
 One of the 13 lineups the participants saw was the true data from the 112th senate shown in Section~\ref{sec:senatedata}.

### 2017-10-30 ASApaper/paper-child.Rnw
updates from this morning

--- 2017-10-30 ASApaper/paper-child.Rnw
@@
 
-In order to determine which specific values of the parameters we wanted to examine, we constructed an online application that created the lineup protocol for us to be the guinea pigs of our own experiment \citep{shiny}.
-A screen shot of the app is shown in Figure~\ref{fig:shiny}.
+In order to determine which specific values of the parameters we wanted to examine, we constructed an online application that created the lineup protocol for us to be the guinea pigs of our own experiment \citep{theqs}.
+A screen shot of the app wee created with the \texttt{shiny} package is shown in Figure~\ref{fig:shiny} \citep{shiny}.
 On the left side of the screen, the user\footnote{Please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/} to create lineups constructed from the models we present for this data for yourself.} can input the information necessary for creating a lineup of the models M1 through M6 for the data in Section~\ref{sec:senatedata}: first, choose to simulate only one plot from the specified model (analogous to changing the alternative model in the lineup protocol) or to simulated $M-1$ plots from the specified model (analogous to changing the null model in the lineup protocol); the model of interest; the wave of the data to examine; if model M1 is selected, whether to alter the density or the reciprocity parameter; the size of the lineup; the amount by which to multiply the effect selected; a random seed for replicability; and a layout algorithm to use for the node-link diagrams.
@@
 newdata <- read_csv("data/newdata_pred_glmm.csv")
+newdata$param <- newdata$test_param
+newdata$param <- as.factor(newdata$test_param)
+levels(newdata$param) <- c("beta[1]", "beta[3]", "beta[4]", "beta[2]", "beta[6]", "beta[5]")
+newdata$param <-as.character(newdata$param)
 plotdata <- turk22 %>% mutate(type2 = ifelse(sign == 0, sign(initialEst), sign),
@@
 )
-
-labdat <- data_frame(test_param = "jttp", type2 = -1, x = -5, y = .9, label = "outlier")
+plotdata$param <- plotdata$test_param
+plotdata$param <- as.factor(plotdata$param)
+levels(plotdata$param) <- c("beta[1]", "beta[3]", "beta[4]", "beta[2]", "beta[6]", "beta[5]")
+plotdata$param <- as.character(plotdata$param)
+labdat <- data_frame(test_param = "jttp", param = "beta[3]", type2 = -1, x = -5, y = .9, label = "outlier")
 
@@
   geom_point(data = plotdata, aes(x = size, y = datapick, color = as.factor(type2))) + 
+  #geom_text(data = plotdata, aes(x = size, y = datapick, color = as.factor(type2), label = pic_id)) +
   geom_text(data = labdat, aes(x = x, y=y, label = label, color = as.factor(type2)),
             show.legend = F) +
+
   scale_color_brewer(palette = "Dark2", name = "Lineup Type") + 
-  facet_wrap(~test_param, scales = 'free_x') + 
+  facet_wrap(~param, scales = 'free_x', labeller = label_parsed) + 
   labs(x = "Parameter value", y = "Expected probability of detection in a lineup of size 6") + 
@@
 
-<<recipzoom, fig.height=2.5, fig.cap='The bottom left panel of Figure~\\ref{fig:predictglmm} expanded to show greater detail.
+<<recipzoom, fig.height=2.5, fig.cap='The top middle panel of Figure~\\ref{fig:predictglmm} expanded to show greater detail.
 For this parameter, as its value approaches zero, the probability of identifying the alternate data model decreases, then increases, which is noticeably different from the pattern exhibited by the others.'>>=
@@
   facet_wrap(~type2, scales = 'free_x') + 
-  labs(x = "Parameter value (recip)", y = "Expected probability of detection in a lineup of size 6") + 
+  labs(x = "Parameter value (recip)", y = "Expected probability of detection\nin a lineup of size 6") + 
   ThemeNoNet + 
@@
 
-In Figure~\ref{fig:predictglmm}
+In Figure~\ref{fig:predictglmm}, we see a clear trend in all parameters except $\beta_1$ and $\beta_2$ that as the parameter value approaches zero from either side, the probability of picking the data plot in a lineup of size six descreases.
+For $\beta_3$ and $\beta_5$, the slope of the fitted line is much steeper for positive values of the parameter than form negative values, meaning that our participants perceived differences more often for postitive parameter values than for negative parameter values.
+This finding is similar to that of \citet{corrviz}, who found that people detect positive correlations sooner and better than negative correlations.
+
+<<beta41, fig.height = 3, fig.cap='In our experiment, 52.8\\% of viewers of this plot selected the plot from the alternative model, M4.
+The ``reverse" of this lineup is given in Figure~\\ref{fig:beta4neg1}, where 41.4\\% of viewers selected the plot from the alternative model, M1.
+Here, the alternative plot is $\\sqrt{25} - 3$.'>>=
+dat <- read_csv("data/jtts_pos_easy_3.csv")
+# "answer" is 2
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
+        theme_net() + 
+        theme(panel.background = element_rect(color = 'black')) +
+        facet_wrap(~ord)
+@
+
+<<beta4neg1,fig.height=3, fig.cap='In our experiment, 41.4\\% of viewers of this plot selected the plot from the alternative model, M1.
+The ``reverse" of this lineup is given in Figure~\\ref{fig:beta41}, where 52.8\\% of viewers selected the plot from the alternative model, M1.
+Here, the alternative plot is $\\sqrt{25} - 1$.' >>=
+dat <- read_csv("data/jtts_neg_med_2.csv")
+# 'answer' is 4
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
+        theme_net() + 
+        theme(panel.background = element_rect(color = 'black')) +
+        facet_wrap(~ord)
+@
+
+For $\beta_4$ and $\beta_6$, where one plot simulated from M1 was placed among five plots from the corresponding model, we see that the predictions for the reverse lineup type (-1), are less than the standard lineup type (1) for all values of the parameter that we have.
+This contradicts our hypothesis for this scenario, which was that these two scenarios would perform similarly.
+One of the lineups for the $\beta_4 = 6.681$, lineup type 1 scenario is given in Figure~\ref{fig:beta41}, and a corresponding lineup for the lineup type -1 scenario is given in Figure~\ref{fig:beta4neg1}.
+For identical values of the parameter, viewers had a harder time identifying the different plot when they were selected the most ``simple" structure, detecting M1 in five plots from the more complicated model, than they did identifying the most ``complex" structure, the plot from the more complicated model, from the five plots from M1.
 
@@
 The traditional methods only look at one piece or one measure of a network model, but our methods look at the models holistically for a broader sense of what it means for a parameter to be significant or a model to be a good fit.
-By looking at an entire network simulated from a SAOM side-by-side with other instances of networks simulated from another model, we develop an idea of the model in terms of the data itself, instead of in terms of statistical summaries of the data.
+By looking at an entire network simulated from a SAOM side-by-side with other instances of networks simulated from another model, instead of one-dimensional derived features, we develop an idea of the model in terms of the data itself, instead of in terms of statistical summaries of the data.
 

### 2017-10-30 ASApaper/paper-child.Rnw

@@
 
-To determine the threshold at which the effect becomes noticeable, we examined six different levels of the effect, three negative and three positive ones.
-We hypothesized that the perception of the effect by our participants would look something like Figure~\ref{fig:hypothesis}.
-The higher in absolute value the parameter is, the more likely participants are to choose the alternative model out of the lineup.
-To determine the six levels of the effects we want to test, started with the estimates of the parameter at hand as determined by the procedure described in Section~\ref{sec:models}, and multiply it or changing its sign to determine what \textit{we} notice about the effect of the parameter in simulations from the changed models.
+To determine the threshold at which an effect becomes noticeable, we examine six different levels of the effect, three negative and three positive ones.
+Figure~\ref{fig:hypothesis} shows a sketch of what the detection probability by participants' looks like hypothetically with varying effect size: the higher in absolute value the parameter is, the more likely participants are to choose the alternative model out of the lineup.
+To determine the exact values of the six levels we want to test for each effect, we started with the estimates of the parameter at hand (see  Section~\ref{sec:models}), and used small negative and positive factors to determine at what point \textit{we} noticed the effect of the parameter in simulations from the changed models.
 
-In order to determine which specific values of the parameters we wanted to examine, we constructed an online application that created the lineup protocol for us to be the guinea pigs of our own experiment \citep{theqs}.
-A screen shot of the app wee created with the \texttt{shiny} package is shown in Figure~\ref{fig:shiny} \citep{shiny}.
+For this we constructed an online application that created the lineup protocol for us to be the guinea pigs of our own experiment \citep{theqs}.
+A screen shot of the app we created with the \texttt{shiny} package is shown in Figure~\ref{fig:shiny} \citep{shiny}.
 On the left side of the screen, the user\footnote{Please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/} to create lineups constructed from the models we present for this data for yourself.} can input the information necessary for creating a lineup of the models M1 through M6 for the data in Section~\ref{sec:senatedata}: first, choose to simulate only one plot from the specified model (analogous to changing the alternative model in the lineup protocol) or to simulated $M-1$ plots from the specified model (analogous to changing the null model in the lineup protocol); the model of interest; the wave of the data to examine; if model M1 is selected, whether to alter the density or the reciprocity parameter; the size of the lineup; the amount by which to multiply the effect selected; a random seed for replicability; and a layout algorithm to use for the node-link diagrams.

### 2017-10-30 ASApaper/paper-child.Rnw
fix bib reference, add sentence to intro.

@@
 Visual inference, introduced by \citet{Bujaetal}, allows us to look at the entire dataset simulated from a network model as opposed to a (set of) usually one-dimensional metric(s) derived from the network such as outdegree, or a $p$-value for a single parameter in the model.
+In addition, we use visual inference to better understand some parameters in a social network model, by looking at how different values for parameters affect the networks simulated from these models.
 
@@
 The acting node can also choose to \textit{not} change at all.
-This occurs when the numerator, as calculated for the current state of the network, is larger than for any changes that could be made.
+This occurs when the numerator, as calculated for the current state of the network, is larger than for any changes $x(i \leadsto j)$ that could be made.
 
@@
 
-According to Snijders et al in the \texttt{RSiena} manual, at least two parameters must be included in the objective function: the density and the reciprocity.
+According to \citet{RSienamanual}, at least two parameters must be included in the objective function: the density and the reciprocity.
 We denote the density, or out-degree, parameter by $\beta_1$ and the associated statistic as $s_{i1}(x) = \sum_{j} x_{ij}$.

### 2017-10-30 ASApaper/paper-child.Rnw
add 1/6 line to plots

@@
 The lines show the expected probability of detecting the alternative data in a lineup of size 6 for new observers of new lineups is plotted on the $y$-axis, and the size of the parameter of interest is on the $x$-axis.
-The proportions detected by our Turk participants for each lineup group are shown by the points.
+The proportions detected by our Turk participants for each lineup group are shown by the points, with the probability of picking out the data plot at random shown by a horizontal line at 1/6.
 The lineup marked as ``outlier" was removed from modeling.
-The ``recip" panel is also presented in Figure~\\ref{fig:recipzoom} in more detail.'>>=
+The panel for the reciprocity parameter, $\\beta_2$ is also presented in Figure~\\ref{fig:recipzoom} in more detail.'>>=
 newdata <- read_csv("data/newdata_pred_glmm.csv")
@@
             show.legend = F) + 
-
+  geom_hline(yintercept = 1/6, linetype='dashed') +
   scale_color_brewer(palette = "Dark2", name = "Lineup Type") + 
@@
 <<recipzoom, fig.height=2.5, fig.cap='The top middle panel of Figure~\\ref{fig:predictglmm} expanded to show greater detail.
-For this parameter, as its value approaches zero, the probability of identifying the alternate data model decreases, then increases, which is noticeably different from the pattern exhibited by the others.'>>=
+For this parameter, as its value approaches zero, the probability of identifying the alternate data model decreases, then increases, which is noticeably different from the pattern exhibited by the others.
+Again, a horizontal line is drawn at 1/6, the chance of selecting the data plot at random.'>>=
 ggplot() + 
@@
   scale_color_brewer(palette = "Dark2", name = "Lineup Type") + 
+  geom_hline(yintercept = 1/6, linetype='dashed') + 
   facet_wrap(~type2, scales = 'free_x') + 

### 2017-10-30 ASApaper/paper-child.Rnw
more mini comments

@@
 All of the parameters we test for significance using the lineup protocol, $\beta_3$ and $\beta_4$, were determined to be statistically significant using Equation~\ref{eq:wald1}.
-The results from the significance tests we performed using the lineup protocol are given in Table~\ref{tab:sigtesting}
+The results from the significance tests we performed using the lineup protocol are given in Table~\ref{tab:sigtesting}.
+\hh{XXX how do we interpret these  visual results?
+Include the six lineups in an appendix and refer to them from here.}
 
@@
 Using the lineup protocol, we show each Amazon Mechanical Turk worker the data once, in a lineup with five other plots of simulated data from one of the models we chose.
-We examined four different models, M2, M3, M4, and M6, and examined three repititons of each, for a total of 12 goodness-of-fit lineups.
-In each lineup, the ``null model" is one of the four models and the ``alternative" model is the true, unknown model that generated the senate network data The hypotheses for our goodness-of-fit tests are:
+We examined four different models, M2, M3, M4, and M6, and examined three repetitions of each, for a total of 12 goodness-of-fit lineups.
+In each lineup, the ``null model" is one of the four models and the ``alternative" model is the true, unknown model that generated the senate network data.
+The hypotheses for our goodness-of-fit tests are:
 \begin{description}
-\item $H_0$:  The senate network data come from the null model.
+\item $H_0$:  The senate network data come from (or could have come from) the null model.
 \item $H_A$: The senate network data do not come from the null model. 
 \end{description}
-If a lineup viewer picks out the data among the five simulations from the null model, that is evidence in favor of the alternative hypothesis.
+If a lineup viewer picks out the data among the five simulations from the null model, it is evidence in favor of the alternative hypothesis.
 On the contrary, if the lineup viewer picks one of the null plots, that is evidence against the alternative hypothsis.
@@
 
-The $p$-values were calcluated using the \texttt{vinference} package by \citet{vinference}.
+The $p$-values were calculated using the \texttt{vinference} package by \citet{vinference}.
 This package contains methods to calculate \textit{Visual distributions} for lineup experiment data.
@@
 The lineup that resulted in a failure to reject the null hypothesis is shown in Figure~\ref{fig:failgof}.
-The null model in this lineup is M4, and the senate data is shown in panel number $3^2 - 7$, but the panel most chose was number four, and the most common reasoning for that choice was that it had the most simple structure.
+The null model in this lineup is M4, and the senate data is shown in panel number $3^2 - 7$.
+However, the panel most participants chose was number four, and the most common reasoning for that choice was that it had the most simple structure.
 
@@
 @
+
+\hh{XXX we need to go throught the logic of the argument below.}
 

### 2017-10-30 ASApaper/paper-child.Rnw
visual detection is statistical power

@@
 
+\hh{this is the power of the visual test XXX}
 We futher explore this relationship between identification of the alternative data in the lineup and the parameter, effect size, and lineuptype with a generalized linear mixed model.

### 2017-10-30 ASApaper/paper-child.Rnw
couple of points for discussion

@@
 
+\hh{Best results: visual detection/ power analysis, a lot of variability in goodness of fit and significance tests.
+BUT: visualizations show that a decision of significance/ goodness of fit are not as clear cut as the traditional tests want us to believe.
+
+Limitations: 
+\begin{itemize}
+\item number of null plots: null plots are representatives of the null distribution of all possible node-link diagrams of data sampled from the null model.
+Here, the number of null plots is  reduced to five, which increases the variability seen from a single lineup dramatically and unfortunately leads to different conclusions.
+\item ability to generalize: the lineups are made for only one test case - it is not clear, whether and how far, power results transfer to other situations with different number of actors and different edge density.
+\end{itemize}}
+
 % Talk about visual inference

### 2017-10-30 ASApaper/paper-child.Rnw
beef up the discussion sections in 5.

--- 2017-10-30 ASApaper/paper-child.Rnw
@@
 All of the parameters we test for significance using the lineup protocol, $\beta_3$ and $\beta_4$, were determined to be statistically significant using Equation~\ref{eq:wald1}.
-The results from the significance tests we performed using the lineup protocol are given in Table~\ref{tab:sigtesting}.
-\hh{XXX how do we interpret these  visual results?
-Include the six lineups in an appendix and refer to them from here.}
+The results from the significance tests we performed using the lineup protocol are given in Table~\ref{tab:sigtesting}
 
@@
 @
+
+We see that the $p$-values from visual inference, which are calculated using the \texttt{vinference} \texttt{R} package by \citet{vinference}), are variable.
+This variability is introduced through the null plots generated from M1, since not all simulations look alike.
+In addition, the necessarily small number of null plots do not give the viewer as complete of a view of the null model as the usual 19 null plots would.
+The results of the significance tests for $\beta_3$ and $\beta_4$ are not clear cut.
+Thus, unlike the Wald-type tests described at the beginning of this section, there is no way to flat out reject or to fail to reject the null hypothesis that the parameter value is 0.
+We include all of the lineups shown to our participants in the appendix.
 
@@
 Using the lineup protocol, we show each Amazon Mechanical Turk worker the data once, in a lineup with five other plots of simulated data from one of the models we chose.
-We examined four different models, M2, M3, M4, and M6, and examined three repetitions of each, for a total of 12 goodness-of-fit lineups.
-In each lineup, the ``null model" is one of the four models and the ``alternative" model is the true, unknown model that generated the senate network data.
-The hypotheses for our goodness-of-fit tests are:
+We examined four different models, M2, M3, M4, and M6, and examined three repititons of each, for a total of 12 goodness-of-fit lineups.
+In each lineup, the ``null model" is one of the four models and the ``alternative" model is the true, unknown model that generated the senate network data The hypotheses for our goodness-of-fit tests are:
 \begin{description}
-\item $H_0$:  The senate network data come from (or could have come from) the null model.
+\item $H_0$:  The senate network data come from the null model.
 \item $H_A$: The senate network data do not come from the null model. 
 \end{description}
-If a lineup viewer picks out the data among the five simulations from the null model, it is evidence in favor of the alternative hypothesis.
+If a lineup viewer picks out the data among the five simulations from the null model, that is evidence in favor of the alternative hypothesis.
 On the contrary, if the lineup viewer picks one of the null plots, that is evidence against the alternative hypothsis.
@@
 turk22_gof_stats %>% separate(pic_id, into = c("discard", "model", "rep"), sep = c(2,3)) -> print_gof_stats
-print_gof_stats %>% select(-discard) %>% mutate(model = paste0("M", model))-> print_gof_stats
+print_gof_stats %>% select(-discard) %>% mutate(model = paste0("M", (as.integer(model)-1)))-> print_gof_stats
 names(print_gof_stats) <- c("Model", "Replicate", "Data Picks", "Total Viewers", "p-value")
@@
 
-The $p$-values were calculated using the \texttt{vinference} package by \citet{vinference}.
+The $p$-values were calcluated using the \texttt{vinference} package by \citet{vinference}.
 This package contains methods to calculate \textit{Visual distributions} for lineup experiment data.
@@
 The lineup that resulted in a failure to reject the null hypothesis is shown in Figure~\ref{fig:failgof}.
-The null model in this lineup is M4, and the senate data is shown in panel number $3^2 - 7$.
-However, the panel most participants chose was number four, and the most common reasoning for that choice was that it had the most simple structure.
+The null model in this lineup is M4, and the senate data is shown in panel number $3^2 - 7$, but the panel most chose was number four, and the most common reasoning for that choice was that it had the most simple structure.
 
@@
 The null model for this lineup is M4.
-Only 7 of 21 viewers of this lineup selected the data plot as the most different from the others.">>=
+Only 7 of 20 viewers of this lineup selected the data plot as the most different from the others.">>=
 load("data/se112adjmat.RDS")
@@
 
-\hh{XXX we need to go throught the logic of the argument below.}
-
 The smallest $p$-value for one of the goodness-of-fit lineups was for the third replicate of the null model M4.
@@
 In the remaining replicate of M4 as the null model, 13 of 16 viewers identified the data plot, corresponding to a $p$-values of less than $0.0001$, just like the third replicate.
-This leads us to believe that the high $p$-value observed in the second replicate is more of a fluke.
+This variability in results is similar to the variability we found in Section~\ref{sec:sigtest}.
+This variability is again introduced through the plots simulated from null model, and does not provide us with a clear cut decision resulting the hypothesis test.
+For model M4, we can neither reject nor fail to reject the null hypothesis that the data come from model M4.
+This is evidence that the goodness-of-fit of network models cannot always be determined by one dimensional derived features, such as $p$-value shown on the $x$-axis in Figure~\ref{fig:gofsiena}.
+%This leads us to believe that the high $p$-value observed in the second replicate is more of a fluke.
 %This demonstrates some of the problems inherent with network models, such as model degeneracy, as discussed in \citet{degeneracy}.
 Network models
-Furthermore, for the first lineup replicate with model M5 as the null model, only 9 of 21 participants were able to pick out the data, but the Visual distribution $p$-value is less than 0.05.
-This is much higher than the corresponding Binomial probability of $Pr(X = 9|n = 21, p = \frac{1}{6}) = \binom{21}{9}(\frac{1}{6})^9(\frac{5}{6})^{21-9} \approx 0.003$, but we are not sure that having fewer than half of our participants identify the data plot is a truly significant result.
+%Furthermore, for the first lineup replicate with model M5 as the null model, only 9 of 21 participants were able to pick out the data, but the Visual distribution $p$-value is less than 0.05.
+%This is much higher than the corresponding Binomial probability of $Pr(X = 9|n = 21, p = \frac{1}{6}) = \binom{21}{9}(\frac{1}{6})^9(\frac{5}{6})^{21-9} \approx 0.003$, but we are not sure that having fewer than half of our participants identify the data plot is a truly significant result.
 We could move the threshold from $\alpha = 0.05$ to $\alpha = 0.01$ or $\alpha = 0.001$, but more work needs to be done to determine which, if any of these, is appropriate.
+For the other models for which we tested goodness-of-fit, however, we do have significant evidence from all three replicates to reject the null hypothesis that the null model generated the data.
+All of the goodness-of-fit lineups are provided in the appendix.
 

### 2017-10-30 ASApaper/paper-child.Rnw

@@
 
+\st{I reformulated the discussion per our convo today.  }
+\hh{XXX we need to go throught the logic of the argument below.}
+
 <<mostsiggof, fig.height = 3, fig.cap="The lineup resulting in the smallest $p$-value rejecting the null hypothesis. Surprisingly, this another repetition for M4 as the null model.">>=

### 2017-10-30 ASApaper/paper-child.Rnw
respond to hh comments; add to the power discussion

@@
 % }
-Two of the most important pieces of statistical modeling are significance testing of model parameters and goodness-of-fit tests.
-In the former, the data are usually assumed to come from a simple model under the null hypothesis, and additional parameters are tested whether they significantly contribute to explaining variability in the data.
-In the latter, the model of interest is examined to determine how well it fits the data.
-Both of these aspects of statistical modeling increase greatly in difficulty as the complexity of the model of interest increases.
+\st{Three} of the most important pieces of statistical modeling are significance testing of model parameters, goodness-of-fit tests, and power of a test.
+In the first, the data are usually assumed to come from a simple model under the null hypothesis, and additional parameters are tested whether they significantly contribute to explaining variability in the data.
+In the second, the model of interest is examined to determine how well it fits the data.
+\st{In the final, the ability of the  hypothesis test in question to detect the difference between the null and alternative hypothesis is determined.}  All three of these aspects of statistical modeling increase greatly in difficulty as the complexity of the model of interest increases.
 %The more complicated the model, the harder it is to determine fit or to include or exclude a parameter.
@@
 
-We propose new methods for significance and goodness-of-fit testing for a set of social network models, stochastic actor-oriented models for dynamic network data \citep{saompaper}.
+We propose new methods for significance and goodness-of-fit testing and power calculation of these for a set of social network models, stochastic actor-oriented models for dynamic network data \citep{saompaper}.
 Specifically, we are using \textit{visual inference} in place of traditional statistical methods for social network models, such as Wald tests for significance of parameters and in- and outdegree distribution metrics for determining goodness-of-fit.
 Visual inference, introduced by \citet{Bujaetal}, allows us to look at the entire dataset simulated from a network model as opposed to a (set of) usually one-dimensional metric(s) derived from the network such as outdegree, or a $p$-value for a single parameter in the model.
-In addition, we use visual inference to better understand some parameters in a social network model, by looking at how different values for parameters affect the networks simulated from these models.
+%In addition, we use visual inference to better understand some parameters in a social network model, by looking at how different values for parameters affect the networks simulated from these models.
 
@@
 For $\beta_5$, twice the estimated value was used.
-More detail on why we use twice the fitted value is provided in Section~\ref{sec:setupvd}.\footnote{If you would like to explore the kinds of lineups we use in further detail, please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/}}.
+More detail on why we use twice the fitted value is provided in Section~\ref{sec:setupvp}.\footnote{If you would like to explore the kinds of lineups we use in further detail, please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/}}.
 To get the simulations, we used the \texttt{siena07} function in \texttt{RSiena} \citep{RSiena}.
@@
 \caption{\label{fig:shiny} A screen shot of the web application we created to design our lineup experiment.
-More details about this application are given in Section~\ref{sec:setupvd}.
+More details about this application are given in Section~\ref{sec:setupvp}.
 In the lineup, M4 is the alternative model with $\beta_5$ set to twice its estimated value given in Table~\ref{tab:fittedvalues}.
@@
 
-\subsection{Visual Detection}\label{sec:setupvd}
+\subsection{Visual Power}\label{sec:setupvp}
 
-Visual detection does not fit into the typical statistical testing framework as easily as the significance and goodness-of-fit tests we perform do.
+%Visual detection does not fit into the typical statistical testing framework as easily as the significance and goodness-of-fit tests we perform do. 
 Through visual inference, we want to determine at which point an effect becomes noticeable in a SAOM.
 By \textit{noticeable}, we mean that the inclusion of the effect alters the appearance of networks simulated from a model to a degree that viewers are able to reliably pick out a node-link diagram rendered from data simulated from a model with the effect from a lineup of plots without the effect.
-We explore all parameters in the objective function, $\beta_1, \dots, \beta_6$ in this way.
+\st{This is a way to determine the power of the visual test.}  We explore all parameters in the objective function, $\beta_1, \dots, \beta_6$ in this way.
 
@@
 They could select from ``Most simple overall structure," ``Most complex overall structure," or ``Other" and provide their own text description of their reasoning.
-\hh{what about the first 12 plots? 6 parameters, 3 levels, 2 directions}
+\st{Twelve of the 13 lineups that the participants saw were used for the significance testing and the visual power methods discussed in Sections~\ref{sec:setupst} and \ref{sec:setupvp}.
+The six parameters, $\beta_1, \dots, \beta_6$ were set to three different values according to how difficult we thought picking the data plot from the lineup would be for our participants, and were also set to be less than or greater than the initial estimate, creating the lineup type variable.}\hh{what about the first 12 plots?
+6 parameters, 3 levels, 2 directions}
 One of the 13 lineups the participants saw was the true data from the 112th senate shown in Section~\ref{sec:senatedata}.
@@
 \hh{XXX how do we interpret these  visual results?
-Include the six lineups in an appendix and refer to them from here.}
+Include the six lineups in an appendix and refer to them from here.} \st{XXX should be taken care of in next paragraph now.}
 
@@
 
-We see that the $p$-values from visual inference, which are calculated using the \texttt{vinference} \texttt{R} package by \citet{vinference}), are variable.
+We see that the $p$-values from visual inference, which are calculated using the \texttt{vinference} \texttt{R} package by \citet{vinference}), are highly variable.
 This variability is introduced through the null plots generated from M1, since not all simulations look alike.
@@
 The distribution depends on the number of evaluations of a plot, $K$, the size of the lineup, $m$, and the lineup scenario, which here is that each lineup containing the same data and the same set of null plots is shown to $K$ independent observers.
-Using these $p$-values, all but one lineup results in a rejection of the null hypothesis at Type-I error rate of $\alpha = 0.05$.
+\st{The visual inference family of distributions is similar to the binomial distribution, but takes the dependency among the $m$ plots in a signle lineup shown to multiple viewers into account.
+} Using these $p$-values, all but one lineup results in a rejection of the null hypothesis at Type-I error rate of $\alpha = 0.05$.
 The lineup that resulted in a failure to reject the null hypothesis is shown in Figure~\ref{fig:failgof}.
@@
 
-<<failgof, fig.height=3, fig.cap="The goodness-of-fit lineup that failed to reject the null hypothesis.
+<<failgof, fig.height=4, fig.cap="The goodness-of-fit lineup that failed to reject the null hypothesis.
 The null model for this lineup is M4.
@@
 
-\st{I reformulated the discussion per our convo today.  }
-\hh{XXX we need to go throught the logic of the argument below.}
+\hh{XXX we need to go throught the logic of the argument below.}\st{ XXX I reformulated the discussion per our convo today.  }
 
-<<mostsiggof, fig.height = 3, fig.cap="The lineup resulting in the smallest $p$-value rejecting the null hypothesis. Surprisingly, this another repetition for M4 as the null model.">>=
+<<mostsiggof, fig.height = 4, fig.cap="The lineup resulting in the smallest $p$-value rejecting the null hypothesis. Surprisingly, this another repetition for M4 as the null model.">>=
 dat <- read_csv(paste0("data/jtts_gof_9_3.csv"))
@@
 
-\subsection{Visual Detection}
+\subsection{Visual Power}
 
@@
 On the $x$ axis, we plot the value of the parameter of interest, and on the $y$ axis, the proportion of times the plot of data simulated from the alternative model was picked out for each lineup.
-The results are split into groups based on the sign of the parameter.
-We can see clear patterns in all of the groups: as the parameter value approaches 0, fewer participants identified the alternative plot.
+The results are split into groups based on the value of the parameter and the lineup type.
+We can see clear patterns in nearly all of the groups: as the parameter value approaches 0, fewer participants identified the alternative plot.
 
@@
 \hh{this is the power of the visual test XXX}
-We futher explore this relationship between identification of the alternative data in the lineup and the parameter, effect size, and lineuptype with a generalized linear mixed model.
-The response variable, $Y_{i\ell m}$, is binary, indicating whether participant $m$ picked the alternative data plot in lineup $\ell$.
+We futher explore this relationship between identification of the alternative data in the lineup and the parameter, effect size, and lineuptype with a generalized linear mixed model \st{that provides us with an estimate of the power of the visual test.} The response variable, $Y_{i\ell m}$, is binary, indicating whether participant $m$ picked the alternative data plot in lineup $\ell$.
 The covariate $x_{\ell 1}$ takes on the value -1 or 1 according to the lineup type code in Table~\ref{tab:experdetail}.
@@
 plotdata$param <- as.character(plotdata$param)
-labdat <- data_frame(test_param = "jttp", param = "beta[3]", type2 = -1, x = -5, y = .9, label = "outlier")
+labdat <- data_frame(test_param = "jttp", param = "beta[3]", type2 = -1, x = -5, y = .9, label = "jttp_neg_hard_2")
 
@@
 <<recipzoom, fig.height=2.5, fig.cap='The top middle panel of Figure~\\ref{fig:predictglmm} expanded to show greater detail.
+The square root of the estimate is shown on the $x$-axis.
 For this parameter, as its value approaches zero, the probability of identifying the alternate data model decreases, then increases, which is noticeably different from the pattern exhibited by the others.
@@
 ggplot() + 
-  geom_line(data = newdata %>% filter(test_param == "recip"), aes(x = size, y = predictfinal, color = as.factor(type2))) + 
-  geom_point(data = plotdata %>% filter(test_param == "recip"), aes(x = size, y = datapick, color = as.factor(type2))) + 
+  geom_line(data = newdata %>% filter(test_param == "recip"), aes(x = sqrt(size), y = predictfinal, color = as.factor(type2))) + 
+  geom_point(data = plotdata %>% filter(test_param == "recip"), aes(x = sqrt(size), y = datapick, color = as.factor(type2))) + 
   scale_color_brewer(palette = "Dark2", name = "Lineup Type") + 
@@
   facet_wrap(~type2, scales = 'free_x') + 
-  labs(x = "Parameter value (recip)", y = "Expected probability of detection\nin a lineup of size 6") + 
+  labs(x = "Square root of parameter value (recip)", y = "Expected probability of detection\nin a lineup of size 6") + 
   ThemeNoNet + 
@@
 
-<<beta41, fig.height = 3, fig.cap='In our experiment, 52.8\\% of viewers of this plot selected the plot from the alternative model, M4.
+<<beta41, fig.height = 4, fig.cap='In our experiment, 52.8\\% of viewers of this plot selected the plot from the alternative model, M4.
 The ``reverse" of this lineup is given in Figure~\\ref{fig:beta4neg1}, where 41.4\\% of viewers selected the plot from the alternative model, M1.
@@
 
-<<beta4neg1,fig.height=3, fig.cap='In our experiment, 41.4\\% of viewers of this plot selected the plot from the alternative model, M1.
+<<beta4neg1,fig.height=4, fig.cap='In our experiment, 41.4\\% of viewers of this plot selected the plot from the alternative model, M1.
 The ``reverse" of this lineup is given in Figure~\\ref{fig:beta41}, where 52.8\\% of viewers selected the plot from the alternative model, M1.
@@
 By looking at an entire network simulated from a SAOM side-by-side with other instances of networks simulated from another model, instead of one-dimensional derived features, we develop an idea of the model in terms of the data itself, instead of in terms of statistical summaries of the data.
+
+We have found the visual power of some effects in the object function of a SAOM for this particular senate data example,  and we have shown that, for the same effects, there is a lot of variability in results from significance and goodness of fit tests.
+Because the visual tests we performed show a great deal of variability, we can see that the decisions with respect to the significance of a parameter or the goodness of fit of a model to data are not as cut-and-dried as the traditional methods would have us believe.
+
+These results do not come without limitations.
+In visual inference, the null plots are supposed to play the role of good representatives of the null model.
+Here, the number of null plots is reduced to five, which increases the variability seen from a single lineup dramatically, and can unfortunately lead to different conclusions for the same lineup scenario.
+Furthermore, these results do not generalize yet.
+The lineups shown are made for only one test case, and it is not clear whether the power results transfer, nor is it clear to what degree if they do transfer, to other situations with different number of actors and different edge density.
 

### 2017-10-31 ASApaper/paper-child.Rnw
parameters need to be defined

@@
 
-\hh{this is the power of the visual test XXX}
-We futher explore this relationship between identification of the alternative data in the lineup and the parameter, effect size, and lineuptype with a generalized linear mixed model \st{that provides us with an estimate of the power of the visual test.} The response variable, $Y_{i\ell m}$, is binary, indicating whether participant $m$ picked the alternative data plot in lineup $\ell$.
+We futther explore this relationship between identification of the alternative data in the lineup and the parameter, effect size, and lineup type with a generalized linear mixed model that provides us with an estimate of the power of the visual test.
+The response variable, $Y_{i\ell m}$, is binary, indicating whether participant $m$ picked the alternative data plot in lineup $\ell$ for effect $i$.
 The covariate $x_{\ell 1}$ takes on the value -1 or 1 according to the lineup type code in Table~\ref{tab:experdetail}.
@@
 \end{align}
+
+\hh{XXX define all of the parameters above XXX}
+
 The results of fitting this model using \texttt{glmer} from the \texttt{lme4} package are summarized in Table~\ref{tab:glmmests} \citep{lme4}.

### 2017-10-31 ASApaper/paper-child.Rnw
model, simplified
label_bquote; modify themes to match with ggplot2 cran version

--- 2017-10-30 ASApaper/paper-child.Rnw
@@
                                             family="Times New Roman"),
-                  axis.text.x.top = element_text(size = 12,
+                  axis.text.x = element_text(size = 12,
                                             face = 'plain', 
@@
                                             family="Times New Roman"), 
-                  axis.text.x.bottom = element_text(size = 12,
+                  # axis.text.x.top = element_text(size = 12,
+                  #                           face = 'plain', 
+                  #                           angle = 0,
+                  #                           family="Times New Roman"), 
+                  # axis.text.x.bottom = element_text(size = 12,
+                  #                           face = 'plain', 
+                  #                           angle = 0,
+                  #                           family="Times New Roman"), 
+                  axis.text.y = element_text(size = 12,
                                             face = 'plain', 
@@
                                             family="Times New Roman"),
-                  axis.text.y.left = element_text(size = 12,
-                                            face = 'plain', 
-                                            angle = 0,
-                                            family="Times New Roman"),
-                  axis.text.y.right = element_text(size = 12,
-                                            face = 'plain', 
-                                            angle = 0,
-                                            family="Times New Roman"),
+                  # axis.text.y.left = element_text(size = 12,
+                  #                           face = 'plain', 
+                  #                           angle = 0,
+                  #                           family="Times New Roman"),
+                  # axis.text.y.right = element_text(size = 12,
+                  #                           face = 'plain', 
+                  #                           angle = 0,
+                  #                           family="Times New Roman"),
                   strip.text.x = element_text(size = 12,
@@
 newdata$param <- as.factor(newdata$test_param)
-levels(newdata$param) <- c("beta[1]", "beta[3]", "beta[4]", "beta[2]", "beta[6]", "beta[5]")
-newdata$param <-as.character(newdata$param)
+levels(newdata$param) <-  c(1,3,4,2,6,5) 
+newdata$param <- as.integer(as.character(newdata$param))
 plotdata <- turk22 %>% mutate(type2 = ifelse(sign == 0, sign(initialEst), sign),
@@
 plotdata$param <- as.factor(plotdata$param)
-levels(plotdata$param) <- c("beta[1]", "beta[3]", "beta[4]", "beta[2]", "beta[6]", "beta[5]")
-plotdata$param <- as.character(plotdata$param)
-labdat <- data_frame(test_param = "jttp", param = "beta[3]", type2 = -1, x = -5, y = .9, label = "jttp_neg_hard_2")
+levels(plotdata$param) <- c(1,3,4,2,6,5) 
+plotdata$param <- as.integer(as.character(plotdata$param))
+labdat <- data_frame(test_param = "jttp", param = 3, type2 = -1, x = -5, y = .9, label = "jttp_neg_hard_2")
 
@@
   scale_color_brewer(palette = "Dark2", name = "Lineup Type") + 
-  facet_wrap(~param, scales = 'free_x', labeller = label_parsed) + 
+  facet_wrap(~param, scales = 'free_x', 
+             labeller = label_bquote(beta [.(param)])) + 
   labs(x = "Parameter value", y = "Expected probability of detection in a lineup of size 6") + 

### 2017-10-31 ASApaper/paper-child.Rnw

@@
 We include random effects in the model for each lineup, $\delta_{\ell}$ and each participant, $\epsilon_m$, and fit a hierarchical model as follows:
+
+\begin{align}
+\begin{split}
+Y_{\ell m} &\sim \text{Bernoulli}(\pi_{\ell m}) \\
+\text{logit}(\pi_{\ell m}) & = \alpha_{\ell} + \beta_{\ell} x + \delta_{\ell} + \epsilon_{m} \label{eq:glmm} \\
+\delta_{\ell} & \sim N(0, \sigma^2_{\delta}) \\
+\epsilon_{m} & \sim N(0, \sigma^2_{\epsilon})
+\end{split}
+\end{align}
+
+\hh{
+$\ell$ encompasses all 12 combinations of parameter and positive/negative direction of the effect size.
+}
+
 \begin{align}

### 2017-10-31 ASApaper/paper-child.Rnw
changes

@@
 
-We see that the $p$-values from visual inference, which are calculated using the \texttt{vinference} \texttt{R} package by \citet{vinference}), are highly variable.
+We see that the $p$-values from visual inference, which are calculated using the \texttt{vinference} \texttt{R} package by \citet{vinference}, are highly variable.
 This variability is introduced through the null plots generated from M1, since not all simulations look alike.
@@
 
-We futther explore this relationship between identification of the alternative data in the lineup and the parameter, effect size, and lineup type with a generalized linear mixed model that provides us with an estimate of the power of the visual test.
-The response variable, $Y_{i\ell m}$, is binary, indicating whether participant $m$ picked the alternative data plot in lineup $\ell$ for effect $i$.
-The covariate $x_{\ell 1}$ takes on the value -1 or 1 according to the lineup type code in Table~\ref{tab:experdetail}.
-In most cases, -1 indicates the parameter value is less than original estimate, while 1 indicates the parameter value is at or above the original estimate.
-The continuous covariate $x_{\ell 2}$ is the centered and scaled size of the effect of interest from which the alternative data were simulated, the values of which are labeled ``easy", ``medium", and ``hard" in Table~\ref{tab:experdetail} according to how difficult we thought the Turk participants would find each lineup.
-In Equation~\ref{eq:glmm}, $i \in \{1, 2, 3, 4, 5\}$ corresponding to the effects $\beta_3, \beta_4, \beta_2, \beta_6, \beta_5$, respectively.
-We include random effects in the model for each lineup, $\delta_{\ell}$ and each participant, $\epsilon_m$, and fit a hierarchical model as follows:
+We further explore this relationship between identification of the alternative data in the lineup and the parameter, effect size, and lineup type with a generalized linear mixed model that provides us with an estimate of the power of the visual test.
+The response variable, $Y_{ijkm}$, is binary, indicating whether participant $m$ picked the alternative data plot in lineup $j$, rep $k$, for effect $i$.
+The continuous covariate $x$ is the centered and scaled size of the effect of interest from which the alternative data were simulated, the values of which are labeled ``easy", ``medium", and ``hard" in Table~\ref{tab:experdetail} according to how difficult we thought the Turk participants would find each lineup.
+In Equation~\ref{eq:glmm}, $i \in \{1, 2, 3, 4, 6\}$ corresponds to the effects $\beta_1, \beta_3, \beta_4, \beta_2, \beta_6, \beta_5$, respectively.
+We include random effects in the model for each lineup, $\delta_{ijk}$ and each participant, $\epsilon_m$, and fit a hierarchical model as follows:
 
@@
 \begin{split}
-Y_{\ell m} &\sim \text{Bernoulli}(\pi_{\ell m}) \\
-\text{logit}(\pi_{\ell m}) & = \alpha_{\ell} + \beta_{\ell} x + \delta_{\ell} + \epsilon_{m} \label{eq:glmm} \\
-\delta_{\ell} & \sim N(0, \sigma^2_{\delta}) \\
-\epsilon_{m} & \sim N(0, \sigma^2_{\epsilon})
+Y_{ijkm} &\sim \text{Bernoulli}(\pi_{ijkm}) \\
+\text{logit}(\pi_{ijkm}) & = \alpha_{ij} + \beta_{ij} x + \delta_{ijk} + \epsilon_{m} \label{eq:glmm} \\
+\delta_{ijk} & \stackrel{iid}{\sim} N(0, \sigma^2_{\delta}) \\
+\epsilon_{m} & \stackrel{iid}{\sim} N(0, \sigma^2_{\epsilon})
 \end{split}
@@
 
-\hh{
-$\ell$ encompasses all 12 combinations of parameter and positive/negative direction of the effect size.
-}
-
-\begin{align}
-\begin{split}
-Y_{i\ell m} &\sim \text{Bernoulli}(\pi_{i\ell m}) \\
-\text{logit}(\pi_{i\ell m}) & = \mu + \alpha_{i} + \theta\mathbb{I}(x_{\ell 1} = 1) + \gamma x_{\ell 2} + \\ & (\alpha\theta)_i \mathbb{I}(x_{\ell 1} = 1) + (\alpha\gamma)_i x_{\ell 2} + (\theta\gamma)\mathbb{I}(x_{\ell 1} = 1)x_{\ell 2} + \\ & (\alpha\theta\gamma)_i\mathbb{I}(x_{\ell 1} = 1) x_{\ell 2} + \delta_{\ell} + \epsilon_{m} \label{eq:glmm} \\
-\delta_{\ell} & \sim N(0, \sigma^2_{\delta}) \\
-\epsilon_{m} & \sim N(0, \sigma^2_{\epsilon})
-\end{split}
-\end{align}
-
-\hh{XXX define all of the parameters above XXX}
+% \hh{
+% $\ell$ encompasses all 12 combinations of parameter and positive/negative direction of the effect size.
+% }
+
+% \begin{align}
+% \begin{split}
+% Y_{i\ell m} &\sim \text{Bernoulli}(\pi_{i\ell m}) \\
+% \text{logit}(\pi_{i\ell m}) & = \mu + \alpha_{i} + \theta\mathbb{I}(x_{\ell 1} = 1) + \gamma x_{\ell 2} + \\ & (\alpha\theta)_i \mathbb{I}(x_{\ell 1} = 1) + (\alpha\gamma)_i x_{\ell 2} + (\theta\gamma)\mathbb{I}(x_{\ell 1} = 1)x_{\ell 2} + \\ & (\alpha\theta\gamma)_i\mathbb{I}(x_{\ell 1} = 1) x_{\ell 2} + \delta_{\ell} + \epsilon_{m} \label{eq:glmm} \\
+% \delta_{\ell} & \sim N(0, \sigma^2_{\delta}) \\
+% \epsilon_{m} & \sim N(0, \sigma^2_{\epsilon})
+% \end{split}
+% \end{align}
+
+%\hh{XXX define all of the parameters above XXX}
+%\st{
+%This model estimates, as a function of the parameter size, the expected value of $\text{logit}(\pi_{ijkm})$ for each of the 12 lineup conditions, repitions, and participants.
+The inverse logit of the estimates are the prediction lines drawn in Figure~\ref{fig:predictglmm}.
 
@@
 The baseline condition for this model against which all other experimental conditions are compared is the first condition in Table~\ref{tab:experdetail}, for parameter $\beta_1$ and lineup type -1.
-The expected value of the link function for this scenario with $\beta_1 = 0$ is $\mu$, and the expected probability that a new observer will pick out the data plot in a new lineup generated from this scenario is $\frac{\exp(\mu)}{1+\exp{\mu}}$.
-Similarly, for a new lineup generated from the tenth condition in in Table~\ref{tab:experdetail}, with $\beta_6 = 1$, and lineup type 1, the expected value of the link function is $\mu + \alpha_{4} + \theta + \gamma + (\alpha\theta)_4 + (\alpha\gamma)_4  + (\theta\gamma) + (\alpha\theta\gamma)_4$.
+The expected value of the link function for this scenario with $\beta_1 = x_{12}$ is $\mu + \gamma x_{12}$, and the expected probability that a new observer will pick out the data plot in a new lineup generated from this scenario is $\frac{\exp(\mu + \gamma x_{12})}{1+\exp{\mu + \gamma x_{12}}}$.
+Similarly, for a new lineup generated from the second condition in in Table~\ref{tab:experdetail}, with $\beta_1 = x_{12}$, and lineup type 1, the expected value of the link function is $\mu + \theta + (\gamma + (\theta\gamma))x_{12}$ and corresponding probability a new observer will pick out the data plot in a new lineup is $\exp\{\mu + \theta + (\gamma + (\theta\gamma))x_{12}\} / 1+\exp\{\mu + \theta + (\gamma + (\theta\gamma))x_{12}\}$.
+
+For any of the remaining parameters, $\beta_2, \dots, \beta_6$, with lineup type -1, the expectation of the link function as a function of the parameter value, $x_{\ell 2}$ is: 
+\begin{equation}\label{eq:predneg1}
+\mu + \alpha_{i} +  (\gamma + (\alpha\gamma)_i)x_{\ell 2}  
+\end{equation}
+where $i \in \{1,2,3,4,5\}$ corresponds to $\beta_3, \beta_4, \beta_2, \beta_6, \beta_5$, respectively.
+Similarly, for lineup type 1, the expectation of the link function as a function of the parameter value is: \begin{equation}\label{eq:predpos1}
+\mu + \alpha_{i} + \theta + (\alpha\theta)_i  +  (\gamma + (\alpha\gamma)_i + (\gamma\theta) + (\alpha\theta\gamma)_i)x_{\ell 2}  
+\end{equation}
+Thus, the first 24 parameters
+
 The corresponding expected probability that a new observer viewing a new lineup from this scenario is: 
@@
 <<glmmres>>=
-load("data/finalglmm.RDA")
-mod <- model0randomscalesize
+load("data/finalglmm31.RDA")
+mod <- model1randomscalesize
 res <- summary(mod)[[10]]
@@
 
-\begin{table}
-\centering
-\begin{tabular}{lcccc}
-Parameter & Estimate & Std Error & $p$-value & Odds Multiplier \\
-$\mu$ & \Sexpr{ests[1]} & \Sexpr{se[1]} & \Sexpr{pval[1]} & \Sexpr{odds[1]}\\
-$\alpha_1$ & \Sexpr{ests[2]} & \Sexpr{se[2]} & \Sexpr{pval[2]} & \Sexpr{odds[2]} \\
-$\alpha_2$ & \Sexpr{ests[3]} & \Sexpr{se[3]} & \Sexpr{pval[3]} & \Sexpr{odds[3]}  \\
-$\alpha_3$  & \Sexpr{ests[4]} & \Sexpr{se[4]} & \Sexpr{pval[4]} & \Sexpr{odds[4]} \\
-$\alpha_4$ & \Sexpr{ests[5]} & \Sexpr{se[5]} & \Sexpr{pval[5]} & \Sexpr{odds[5]}  \\
-$\alpha_5$  & \Sexpr{ests[6]} & \Sexpr{se[6]} & \Sexpr{pval[6]} & \Sexpr{odds[6]} \\
-$\theta$  & \Sexpr{ests[7]} & \Sexpr{se[7]} & \Sexpr{pval[7]} & \Sexpr{odds[7]} \\
-$\gamma$  & \Sexpr{ests[8]} & \Sexpr{se[8]} & \Sexpr{pval[8]} & \Sexpr{odds[8]} \\ 
-$(\alpha\theta)_1$  & \Sexpr{ests[9]} & \Sexpr{se[9]} & \Sexpr{pval[9]} & \Sexpr{odds[9]} \\
-$(\alpha\theta)_2$  & \Sexpr{ests[10]} & \Sexpr{se[10]} & \Sexpr{pval[10]} & \Sexpr{odds[10]} \\
-$(\alpha\theta)_3$  & \Sexpr{ests[11]} & \Sexpr{se[11]} & \Sexpr{pval[11]} & \Sexpr{odds[11]} \\
-$(\alpha\theta)_4$  & \Sexpr{ests[12]} & \Sexpr{se[12]} & \Sexpr{pval[12]} & \Sexpr{odds[12]} \\
-$(\alpha\theta)_5$  & \Sexpr{ests[13]} & \Sexpr{se[13]} & \Sexpr{pval[13]} & \Sexpr{odds[13]} \\
-$(\alpha\gamma)_1$  & \Sexpr{ests[14]} & \Sexpr{se[14]} & \Sexpr{pval[14]} & \Sexpr{odds[14]} \\
-$(\alpha\gamma)_2$  & \Sexpr{ests[15]} & \Sexpr{se[15]} & \Sexpr{pval[15]} & \Sexpr{odds[15]} \\
-$(\alpha\gamma)_3$  & \Sexpr{ests[16]} & \Sexpr{se[16]} & \Sexpr{pval[16]} & \Sexpr{odds[16]} \\
-$(\alpha\gamma)_4$  & \Sexpr{ests[17]} & \Sexpr{se[17]} & \Sexpr{pval[17]} & \Sexpr{odds[17]} \\
-$(\alpha\gamma)_5$  & \Sexpr{ests[18]} & \Sexpr{se[18]} & \Sexpr{pval[18]} & \Sexpr{odds[18]} \\
-$\theta\gamma$  & \Sexpr{ests[19]} & \Sexpr{se[19]} & \Sexpr{pval[19]} & \Sexpr{odds[19]} \\
-$(\alpha\theta\gamma)_1$ & \Sexpr{ests[20]} & \Sexpr{se[20]} & \Sexpr{pval[20]} & \Sexpr{odds[20]}  \\
-$(\alpha\theta\gamma)_2$  & \Sexpr{ests[21]} & \Sexpr{se[21]} & \Sexpr{pval[21]} & \Sexpr{odds[21]} \\
-$(\alpha\theta\gamma)_3$  & \Sexpr{ests[22]} & \Sexpr{se[22]} & \Sexpr{pval[22]} & \Sexpr{odds[22]} \\
-$(\alpha\theta\gamma)_4$  & \Sexpr{ests[23]} & \Sexpr{se[23]} & \Sexpr{pval[23]} & \Sexpr{odds[23]} \\
-$(\alpha\theta\gamma)_5$  & \Sexpr{ests[24]} & \Sexpr{se[24]} & \Sexpr{pval[24]} & \Sexpr{odds[24]} \\ 
-$\sigma^2_{\delta}$ & 0.5638 \\
-$\sigma^2_{\epsilon}$ & 0.3416
-\end{tabular}
-\caption{\label{tab:glmmests}Summary of the results from fitting the model given in Equation~\ref{eq:glmm}}
-\end{table}
+% \begin{table}
+% \centering
+% \begin{tabular}{lcccc}
+% Parameter & Estimate & Std Error & $p$-value & Odds Multiplier \\
+% $\mu$ & \Sexpr{ests[1]} & \Sexpr{se[1]} & \Sexpr{pval[1]} & \Sexpr{odds[1]}\\
+% $\alpha_1$ & \Sexpr{ests[2]} & \Sexpr{se[2]} & \Sexpr{pval[2]} & \Sexpr{odds[2]} \\
+% $\alpha_2$ & \Sexpr{ests[3]} & \Sexpr{se[3]} & \Sexpr{pval[3]} & \Sexpr{odds[3]}  \\
+% $\alpha_3$  & \Sexpr{ests[4]} & \Sexpr{se[4]} & \Sexpr{pval[4]} & \Sexpr{odds[4]} \\
+% $\alpha_4$ & \Sexpr{ests[5]} & \Sexpr{se[5]} & \Sexpr{pval[5]} & \Sexpr{odds[5]}  \\
+% $\alpha_5$  & \Sexpr{ests[6]} & \Sexpr{se[6]} & \Sexpr{pval[6]} & \Sexpr{odds[6]} \\
+% $\theta$  & \Sexpr{ests[7]} & \Sexpr{se[7]} & \Sexpr{pval[7]} & \Sexpr{odds[7]} \\
+% $\gamma$  & \Sexpr{ests[8]} & \Sexpr{se[8]} & \Sexpr{pval[8]} & \Sexpr{odds[8]} \\ 
+% $(\alpha\theta)_1$  & \Sexpr{ests[9]} & \Sexpr{se[9]} & \Sexpr{pval[9]} & \Sexpr{odds[9]} \\
+% $(\alpha\theta)_2$  & \Sexpr{ests[10]} & \Sexpr{se[10]} & \Sexpr{pval[10]} & \Sexpr{odds[10]} \\
+% $(\alpha\theta)_3$  & \Sexpr{ests[11]} & \Sexpr{se[11]} & \Sexpr{pval[11]} & \Sexpr{odds[11]} \\
+% $(\alpha\theta)_4$  & \Sexpr{ests[12]} & \Sexpr{se[12]} & \Sexpr{pval[12]} & \Sexpr{odds[12]} \\
+% $(\alpha\theta)_5$  & \Sexpr{ests[13]} & \Sexpr{se[13]} & \Sexpr{pval[13]} & \Sexpr{odds[13]} \\
+% $(\alpha\gamma)_1$  & \Sexpr{ests[14]} & \Sexpr{se[14]} & \Sexpr{pval[14]} & \Sexpr{odds[14]} \\
+% $(\alpha\gamma)_2$  & \Sexpr{ests[15]} & \Sexpr{se[15]} & \Sexpr{pval[15]} & \Sexpr{odds[15]} \\
+% $(\alpha\gamma)_3$  & \Sexpr{ests[16]} & \Sexpr{se[16]} & \Sexpr{pval[16]} & \Sexpr{odds[16]} \\
+% $(\alpha\gamma)_4$  & \Sexpr{ests[17]} & \Sexpr{se[17]} & \Sexpr{pval[17]} & \Sexpr{odds[17]} \\
+% $(\alpha\gamma)_5$  & \Sexpr{ests[18]} & \Sexpr{se[18]} & \Sexpr{pval[18]} & \Sexpr{odds[18]} \\
+% $\theta\gamma$  & \Sexpr{ests[19]} & \Sexpr{se[19]} & \Sexpr{pval[19]} & \Sexpr{odds[19]} \\
+% $(\alpha\theta\gamma)_1$ & \Sexpr{ests[20]} & \Sexpr{se[20]} & \Sexpr{pval[20]} & \Sexpr{odds[20]}  \\
+% $(\alpha\theta\gamma)_2$  & \Sexpr{ests[21]} & \Sexpr{se[21]} & \Sexpr{pval[21]} & \Sexpr{odds[21]} \\
+% $(\alpha\theta\gamma)_3$  & \Sexpr{ests[22]} & \Sexpr{se[22]} & \Sexpr{pval[22]} & \Sexpr{odds[22]} \\
+% $(\alpha\theta\gamma)_4$  & \Sexpr{ests[23]} & \Sexpr{se[23]} & \Sexpr{pval[23]} & \Sexpr{odds[23]} \\
+% $(\alpha\theta\gamma)_5$  & \Sexpr{ests[24]} & \Sexpr{se[24]} & \Sexpr{pval[24]} & \Sexpr{odds[24]} \\ 
+% $\sigma^2_{\delta}$ & 0.5638 \\
+% $\sigma^2_{\epsilon}$ & 0.3416
+% \end{tabular}
+% \caption{\label{tab:glmmests}Summary of the results from fitting the model given in Equation~\ref{eq:glmm}}
+% \end{table}
 
@@
 Furthermore, these results do not generalize yet.
-The lineups shown are made for only one test case, and it is not clear whether the power results transfer, nor is it clear to what degree if they do transfer, to other situations with different number of actors and different edge density.
+The lineups shown are made for only one set of data, and it is not clear whether the power results transfer, nor is it clear to what degree if they do transfer, to other situations with different number of actors, different edge densities, or different layout algorithm of the node-link diagram.
 
@@
 We accept the limitations of this type of network data visualization, in that even in small instances, the cognitive load of looking at a lineup is very high for the average observer.
-We would therefore like to explore larger datasets, different layout algorithms, and different ways of visualizing network data, such as adjacency matrix visualizations, through visual inference to see if similar patterns emerge.
+We would therefore like to explore larger datasets, different layout algorithms, and different ways of visualizing network data, such as changing the layout algorithm or using adjacency matrix visualizations, through visual inference to see if similar patterns emerge.
 

### 2017-10-31 ASApaper/paper-child.Rnw
model change; remove comments in prep for thesis

@@
-\newcommand{\st}[1]{{\color{orange} #1}}
-\newcommand{\hh}[1]{{\color{magenta} #1}}
+%\newcommand{\st}[1]{{\color{orange} #1}}
+%\newcommand{\hh}[1]{{\color{magenta} #1}}
 
@@
 % }
-\st{Three} of the most important pieces of statistical modeling are significance testing of model parameters, goodness-of-fit tests, and power of a test.
+Three of the most important pieces of statistical modeling are significance testing of model parameters, goodness-of-fit tests, and power of a test.
 In the first, the data are usually assumed to come from a simple model under the null hypothesis, and additional parameters are tested whether they significantly contribute to explaining variability in the data.
 In the second, the model of interest is examined to determine how well it fits the data.
-\st{In the final, the ability of the  hypothesis test in question to detect the difference between the null and alternative hypothesis is determined.}  All three of these aspects of statistical modeling increase greatly in difficulty as the complexity of the model of interest increases.
+In the final, the ability of the hypothesis test in question to detect the difference between the null and alternative hypothesis is determined.
+All three of these aspects of statistical modeling increase greatly in difficulty as the complexity of the model of interest increases.
 %The more complicated the model, the harder it is to determine fit or to include or exclude a parameter.
@@
 Even more challenging is the situation when we go beyond single instances of a network and consider the dynamics of network change between observed instances.
-This type of modelling for dynamic networks is often performed on social network data, such as friendship networks among students or the spread of HIV in \st{drug users sharing needles.} %sexually active communities.% of men sleeping with other men.
+This type of modelling for dynamic networks is often performed on social network data, such as friendship networks among students or the spread of HIV in drug users sharing needles.
+%sexually active communities.% of men sleeping with other men.
 X ST a bit less jarring hopefully.
@@
 These models lack the asymptotics required to perform many well-known goodness-of-fit tests, and the maximum likelihood estimation of parameters is so difficult that it can make significance testing difficult as well \citep{goldenberg09}.
-\hh{is the previous sentence a quote? I'm not sure that I follow.} \st{it's a paraphrase... shall I make a quote? Also I definitely botched a word. Hopefully makes more sense now? }
+%\hh{is the previous sentence a quote? I'm not sure that I follow.} %\st{it's a paraphrase... shall I make a quote? Also I definitely botched a word. Hopefully makes more sense now? }
 
@@
 
-\hh{where do the results come in?}\st{whoops, forgot to updat this with the reorg. }
+%\hh{where do the results come in?}\st{whoops, forgot to updat this with the reorg. }
 
@@
 If an observer identifies the data plot, this is quantifiable evidence against the null hypothesis.
-\hh{Since an observer has a chance of 1 in $m$ to pick the data plot from the lineup by simply guessing, i.e.
-in a situation where the data plot is virtually indistinguishable from the null plots,} \st{the evidence grows in strength with the number of independent observers identifying the data plot.}
+Since an observer has a chance of 1 in $m$ to pick the data plot from the lineup by simply guessing, i.e.
+in a situation where the data plot is virtually indistinguishable from the null plots, the evidence grows in strength with the number of independent observers identifying the data plot.
 
@@
 After actor $i$ has been given the opportunity to change, it probabilistically chooses one of its current ties, $x_{ij}$, to change.
-The probability that actor $i$ changes its current tie to actor $j$ is determined by the \textit{objective function} of the model and a random component, $U$, \st{which can be thought of as encompassing any other factors that may be influencing the changes the node makes not accounted for by the parameters in the model.} \hh{XXX can you paraphrase the purpose of U in half a sentence?}
+The probability that actor $i$ changes its current tie to actor $j$ is determined by the \textit{objective function} of the model and a random component, $U$, which can be thought of as encompassing any other factors that may be influencing the changes the node makes not accounted for by the parameters in the model.
+%\hh{XXX can you paraphrase the purpose of U in half a sentence?}
 %The random component is assumed to have log-Weibull distribution with location parameter $\mu = 0$ and scale parameter $\sigma = 0$ (see \citet{modelsSnijders}), so the probability density function of $U$ is
@@
 
-\hh{XXX how is the probability computed in the case that no change is made? Why is the notation $j \equiv i$ used for no change?}\st{that's more for the computation. I'll adjust accordingly}
+%\hh{XXX how is the probability computed in the case that no change is made? Why is the notation $j \equiv i$ used for no change?}\st{that's more for the computation. I'll adjust accordingly}
 
@@
 By \textit{noticeable}, we mean that the inclusion of the effect alters the appearance of networks simulated from a model to a degree that viewers are able to reliably pick out a node-link diagram rendered from data simulated from a model with the effect from a lineup of plots without the effect.
-\st{This is a way to determine the power of the visual test.}  We explore all parameters in the objective function, $\beta_1, \dots, \beta_6$ in this way.
+This is a way to determine the power of the visual test.
+We explore all parameters in the objective function, $\beta_1, \dots, \beta_6$ in this way.
 
@@
 They could select from ``Most simple overall structure," ``Most complex overall structure," or ``Other" and provide their own text description of their reasoning.
-\st{Twelve of the 13 lineups that the participants saw were used for the significance testing and the visual power methods discussed in Sections~\ref{sec:setupst} and \ref{sec:setupvp}.
-The six parameters, $\beta_1, \dots, \beta_6$ were set to three different values according to how difficult we thought picking the data plot from the lineup would be for our participants, and were also set to be less than or greater than the initial estimate, creating the lineup type variable.}\hh{what about the first 12 plots?
+
+Twelve of the 13 lineups that the participants saw were used for the significance testing and the visual power methods discussed in Sections~\ref{sec:setupst} and \ref{sec:setupvp}.
+The six parameters, $\beta_1, \dots, \beta_6$ were set to three different values according to how difficult we thought picking the data plot from the lineup would be for our participants, and were also set to be less than or greater than the initial estimate, creating the lineup type variable.%\hh{what about the first 12 plots?
 6 parameters, 3 levels, 2 directions}
@@
 The results from the significance tests we performed using the lineup protocol are given in Table~\ref{tab:sigtesting}.
-\hh{XXX how do we interpret these  visual results?
+%\hh{XXX how do we interpret these  visual results?
 Include the six lineups in an appendix and refer to them from here.} \st{XXX should be taken care of in next paragraph now.}
@@
 The distribution depends on the number of evaluations of a plot, $K$, the size of the lineup, $m$, and the lineup scenario, which here is that each lineup containing the same data and the same set of null plots is shown to $K$ independent observers.
-\st{The visual inference family of distributions is similar to the binomial distribution, but takes the dependency among the $m$ plots in a signle lineup shown to multiple viewers into account.
-} Using these $p$-values, all but one lineup results in a rejection of the null hypothesis at Type-I error rate of $\alpha = 0.05$.
+The visual inference family of distributions is similar to the binomial distribution, but takes the dependency among the $m$ plots in a signle lineup shown to multiple viewers into account.
+Using these $p$-values, all but one lineup results in a rejection of the null hypothesis at Type-I error rate of $\alpha = 0.05$.
 The lineup that resulted in a failure to reject the null hypothesis is shown in Figure~\ref{fig:failgof}.
@@
 
-\hh{XXX we need to go throught the logic of the argument below.}\st{ XXX I reformulated the discussion per our convo today.  }
+%\hh{XXX we need to go throught the logic of the argument below.}\st{ XXX I reformulated the discussion per our convo today.  }
 
@@
 We further explore this relationship between identification of the alternative data in the lineup and the parameter, effect size, and lineup type with a generalized linear mixed model that provides us with an estimate of the power of the visual test.
-The response variable, $Y_{ijkm}$, is binary, indicating whether participant $m$ picked the alternative data plot in lineup $j$, rep $k$, for effect $i$.
+The response variable, $Y_{ijkm}$, is binary, indicating whether participant $m$ picked the alternative data plot in lineup type $j$, rep $k$, for effect $i$.
 The continuous covariate $x$ is the centered and scaled size of the effect of interest from which the alternative data were simulated, the values of which are labeled ``easy", ``medium", and ``hard" in Table~\ref{tab:experdetail} according to how difficult we thought the Turk participants would find each lineup.
-In Equation~\ref{eq:glmm}, $i \in \{1, 2, 3, 4, 6\}$ corresponds to the effects $\beta_1, \beta_3, \beta_4, \beta_2, \beta_6, \beta_5$, respectively.
+In Equation~\ref{eq:glmm}, $i \in \{1, 2, 3, 4, 6\}$ corresponds to the effects $\beta_1, \beta_3, \beta_4, \beta_2, \beta_6, \beta_5$, respectively, $j \in \{-1,1\}$, and $k \in \{1,2,3\}$.
 We include random effects in the model for each lineup, $\delta_{ijk}$ and each participant, $\epsilon_m$, and fit a hierarchical model as follows:
@@
 
-The results of fitting this model using \texttt{glmer} from the \texttt{lme4} package are summarized in Table~\ref{tab:glmmests} \citep{lme4}.
-The baseline condition for this model against which all other experimental conditions are compared is the first condition in Table~\ref{tab:experdetail}, for parameter $\beta_1$ and lineup type -1.
+The results of fitting this model, including estimates of parameters, standard errors, $p$-values, and the odds ratio multipliers, using \texttt{glmer} from the \texttt{lme4} package are summarized in Table~\ref{tab:glmmests} \citep{lme4}.
+For each combination of parameter and lineup type, the expected value of the link function for a new lineup and a new observer with parameter value $x$ is
+\begin{equation}
+E[\text{logit}(\pi_{ij})] = \alpha_{ij} + \beta_{ij}x
+\end{equation}
+and the corresponding probability of picking out the alternative data plot is 
+\begin{equation}
+\pi_{ij} = \frac{\exp\{\alpha_{ij} + \beta_{ij}x\}}{1 + \exp\{\alpha_{ij} + \beta_{ij}x\}}
+\end{equation}
+
+%The baseline condition for this model against which all other experimental conditions are compared is the first condition in Table~\ref{tab:experdetail}, for parameter $\beta_1$ and lineup type -1.
 The expected value of the link function for this scenario with $\beta_1 = x_{12}$ is $\mu + \gamma x_{12}$, and the expected probability that a new observer will pick out the data plot in a new lineup generated from this scenario is $\frac{\exp(\mu + \gamma x_{12})}{1+\exp{\mu + \gamma x_{12}}}$.
@@
 
-For any of the remaining parameters, $\beta_2, \dots, \beta_6$, with lineup type -1, the expectation of the link function as a function of the parameter value, $x_{\ell 2}$ is: 
-\begin{equation}\label{eq:predneg1}
-\mu + \alpha_{i} +  (\gamma + (\alpha\gamma)_i)x_{\ell 2}  
-\end{equation}
-where $i \in \{1,2,3,4,5\}$ corresponds to $\beta_3, \beta_4, \beta_2, \beta_6, \beta_5$, respectively.
+% For any of the remaining parameters, $\beta_2, \dots, \beta_6$, with lineup type -1, the expectation of the link function as a function of the parameter value, $x_{\ell 2}$ is: 
+% \begin{equation}\label{eq:predneg1}
+% \mu + \alpha_{i} +  (\gamma + (\alpha\gamma)_i)x_{\ell 2}  
+% \end{equation}
+% where $i \in \{1,2,3,4,5\}$ corresponds to $\beta_3, \beta_4, \beta_2, \beta_6, \beta_5$, respectively.
 Similarly, for lineup type 1, the expectation of the link function as a function of the parameter value is: \begin{equation}\label{eq:predpos1}
-\mu + \alpha_{i} + \theta + (\alpha\theta)_i  +  (\gamma + (\alpha\gamma)_i + (\gamma\theta) + (\alpha\theta\gamma)_i)x_{\ell 2}  
-\end{equation}
-Thus, the first 24 parameters
-
-The corresponding expected probability that a new observer viewing a new lineup from this scenario is: 
-\begin{equation}
-\frac{\exp\{\mu + \alpha_{4} + \theta + \gamma + (\alpha\theta)_4 + (\alpha\gamma)_4  + (\theta\gamma) + (\alpha\theta\gamma)_4\}}{1 + \exp\{\mu + \alpha_{4} + \theta + \gamma + (\alpha\theta)_4 + (\alpha\gamma)_4  + (\theta\gamma) + (\alpha\theta\gamma)_4\}}
-\end{equation}
+% \mu + \alpha_{i} + \theta + (\alpha\theta)_i  +  (\gamma + (\alpha\gamma)_i + (\gamma\theta) + (\alpha\theta\gamma)_i)x_{\ell 2}  
+% \end{equation}
+% 
+% The corresponding expected probability that a new observer viewing a new lineup from this scenario is: 
+% \begin{equation}
+% \frac{\exp\{\mu + \alpha_{4} + \theta + \gamma + (\alpha\theta)_4 + (\alpha\gamma)_4  + (\theta\gamma) + (\alpha\theta\gamma)_4\}}{1 + \exp\{\mu + \alpha_{4} + \theta + \gamma + (\alpha\theta)_4 + (\alpha\gamma)_4  + (\theta\gamma) + (\alpha\theta\gamma)_4\}}
+% \end{equation}
 
 <<glmmres>>=
-load("data/finalglmm31.RDA")
-mod <- model1randomscalesize
+load("data/finalglmm31-2.RDA")
+mod <- model2randomscalesize
 res <- summary(mod)[[10]]
 ests <- as.numeric(round(res[,1], 4))
+oddsidx <- which(exp(ests) < .0001)
+oddsidx2 <- which(exp(ests) > 1000000)
 odds <- round(exp(ests), 4)
+odds[oddsidx] <- "$<$0.0001"
+odds[oddsidx2] <- "$>$1e+06"
 se <- as.numeric(round(res[,2], 4))
+pvalidx <- which(res[,4] < .0001)
 pval <- as.numeric(round(res[,4],4))
-@
+pval[pvalidx] <- "$<$0.0001"
+sigma_epsilon <- summary(model2randomscalesize)[[13]]$nick_name[[1]]
+sigma_delta <- summary(model2randomscalesize)[[13]]$pic_id[[1]]
+@
+
+\begin{table}
+\centering
+\begin{tabular}{lrrrr}
+Parameter & Estimate & Std Error & $p$-value & Odds Multiplier \\
+$\alpha_{1-1}$ & \Sexpr{ests[1]} & \Sexpr{se[1]} & \Sexpr{pval[1]} & \Sexpr{odds[1]}\\
+$\alpha_{2-1}$ & \Sexpr{ests[2]} & \Sexpr{se[2]} & \Sexpr{pval[2]} & \Sexpr{odds[2]} \\
+$\alpha_{3-1}$ & \Sexpr{ests[3]} & \Sexpr{se[3]} & \Sexpr{pval[3]} & \Sexpr{odds[3]}  \\
+$\alpha_{4-1}$  & \Sexpr{ests[4]} & \Sexpr{se[4]} & \Sexpr{pval[4]} & \Sexpr{odds[4]} \\
+$\alpha_{5-1}$ & \Sexpr{ests[5]} & \Sexpr{se[5]} & \Sexpr{pval[5]} & \Sexpr{odds[5]}  \\
+$\alpha_{6-1}$  & \Sexpr{ests[6]} & \Sexpr{se[6]} & \Sexpr{pval[6]} & \Sexpr{odds[6]} \\
+$\alpha_{11}$ & \Sexpr{ests[7]} & \Sexpr{se[7]} & \Sexpr{pval[7]} & \Sexpr{odds[7]} \\
+$\alpha_{21}$  & \Sexpr{ests[8]} & \Sexpr{se[8]} & \Sexpr{pval[8]} & \Sexpr{odds[8]} \\ 
+$\alpha_{31}$ & \Sexpr{ests[9]} & \Sexpr{se[9]} & \Sexpr{pval[9]} & \Sexpr{odds[9]} \\
+$\alpha_{41}$  & \Sexpr{ests[10]} & \Sexpr{se[10]} & \Sexpr{pval[10]} & \Sexpr{odds[10]} \\
+$\alpha_{51}$  & \Sexpr{ests[11]} & \Sexpr{se[11]} & \Sexpr{pval[11]} & \Sexpr{odds[11]} \\
+$\alpha_{61}$  & \Sexpr{ests[12]} & \Sexpr{se[12]} & \Sexpr{pval[12]} & \Sexpr{odds[12]} \\
+$\beta_{1-1}$  & \Sexpr{ests[13]} & \Sexpr{se[13]} & \Sexpr{pval[13]} & \Sexpr{odds[13]} \\
+$\beta_{2-1}$   & \Sexpr{ests[14]} & \Sexpr{se[14]} & \Sexpr{pval[14]} & \Sexpr{odds[14]} \\
+$\beta_{3-1}$  & \Sexpr{ests[15]} & \Sexpr{se[15]} & \Sexpr{pval[15]} & \Sexpr{odds[15]} \\
+$\beta_{4-1}$  & \Sexpr{ests[16]} & \Sexpr{se[16]} & \Sexpr{pval[16]} & \Sexpr{odds[16]} \\
+$\beta_{5-1}$ & \Sexpr{ests[17]} & \Sexpr{se[17]} & \Sexpr{pval[17]} & \Sexpr{odds[17]} \\
+$\beta_{6-1}$ & \Sexpr{ests[18]} & \Sexpr{se[18]} & \Sexpr{pval[18]} & \Sexpr{odds[18]} \\
+$\beta_{11}$ & \Sexpr{ests[19]} & \Sexpr{se[19]} & \Sexpr{pval[19]} & \Sexpr{odds[19]} \\
+$\beta_{21}$ & \Sexpr{ests[20]} & \Sexpr{se[20]} & \Sexpr{pval[20]} & \Sexpr{odds[20]}  \\
+$\beta_{31}$ & \Sexpr{ests[21]} & \Sexpr{se[21]} & \Sexpr{pval[21]} & \Sexpr{odds[21]} \\
+$\beta_{41}$ & \Sexpr{ests[22]} & \Sexpr{se[22]} & \Sexpr{pval[22]} & \Sexpr{odds[22]} \\
+$\beta_{51}$ & \Sexpr{ests[23]} & \Sexpr{se[23]} & \Sexpr{pval[23]} & \Sexpr{odds[23]} \\
+$\beta_{61}$ & \Sexpr{ests[24]} & \Sexpr{se[24]} & \Sexpr{pval[24]} & \Sexpr{odds[24]} \\ 
+$\sigma^2_{\delta}$ & \Sexpr{round(sigma_delta, 4)} & -- & -- & --\\
+$\sigma^2_{\epsilon}$ &  \Sexpr{round(sigma_epsilon, 4)}  & -- & -- & --
+\end{tabular}
+\caption{\label{tab:glmmests}Summary of the results from fitting the model given in Equation~\ref{eq:glmm}}
+\end{table}
 
@@
 
-\hh{Best results: visual detection/ power analysis, a lot of variability in goodness of fit and significance tests.
+%\hh{Best results: visual detection/ power analysis, a lot of variability in goodness of fit and significance tests.
 BUT: visualizations show that a decision of significance/ goodness of fit are not as clear cut as the traditional tests want us to believe.
 
-Limitations: 
-\begin{itemize}
-\item number of null plots: null plots are representatives of the null distribution of all possible node-link diagrams of data sampled from the null model.
-Here, the number of null plots is  reduced to five, which increases the variability seen from a single lineup dramatically and unfortunately leads to different conclusions.
-\item ability to generalize: the lineups are made for only one test case - it is not clear, whether and how far, power results transfer to other situations with different number of actors and different edge density.
-\end{itemize}}
+%Limitations: 
+%\begin{itemize}
+%\item number of null plots: null plots are representatives of the null %distribution of all possible node-link diagrams of data sampled from the %null model.
+Here, the number of null plots is  reduced to five, which %increases the variability seen from a single lineup dramatically and %unfortunately leads to different conclusions.
+%\item ability to generalize: the lineups are made for only one test case %- it is not clear, whether and how far, power results transfer to other %situations with different number of actors and different edge density.
+%\end{itemize}}
 

### 2017-11-01 ASApaper/paper-child.Rnw
minor formatting

@@
 names(print_gof_stats) <- c("Model", "Replicate", "Data Picks", "Total Viewers", "p-value")
-print_gof_stats$`p-value` <- ifelse(print_gof_stats$`p-value` < .0001, "< 0.0001", as.character(round(print_gof_stats$`p-value`,4)))
-print(xtable::xtable(print_gof_stats, label = "tab:gofstats", caption = "An overview of the results from our 12 goodness-of-fit lineup tests.", digits = 4), caption.placement = "bottom", include.rownames = FALSE)
+print_gof_stats$`p-value` <- ifelse(print_gof_stats$`p-value` < .0001, "< 0.0001", sprintf("%.4f", print_gof_stats$`p-value`))
+print(xtable::xtable(print_gof_stats, label = "tab:gofstats", caption = "An overview of the results from the 12 goodness-of-fit lineup tests.", digits = 4,
+                      align="rllrrr"), caption.placement = "bottom", include.rownames = FALSE)
 @

### 2017-11-02 ASApaper/paper-child.Rnw
formatting changes

--- 2017-10-31 ASApaper/paper-child.Rnw
@@
 names(print_gof_stats) <- c("Model", "Replicate", "Data Picks", "Total Viewers", "p-value")
-print_gof_stats$`p-value` <- ifelse(print_gof_stats$`p-value` < .0001, "< 0.0001", as.character(round(print_gof_stats$`p-value`,4)))
-print(xtable::xtable(print_gof_stats, label = "tab:gofstats", caption = "An overview of the results from our 12 goodness-of-fit lineup tests.", digits = 4), caption.placement = "bottom", include.rownames = FALSE)
+print_gof_stats$Model[c(2:3,5:6,8:9,11:12)] <- ""
+print_gof_stats$Model[2] <- "jtt party"
+print_gof_stats$Model[5] <- "jtt sex"
+print_gof_stats$Model[8] <- "stt party"
+print_gof_stats$`p-value` <- ifelse(print_gof_stats$`p-value` < .0001, "< 0.0001", sprintf("%.4f", print_gof_stats$`p-value`))
+print(xtable::xtable(print_gof_stats, label = "tab:gofstats", caption = "An overview of the results from the 12 goodness-of-fit lineup tests.", digits = 4,
+                      align="rlrrrr"), caption.placement = "bottom", include.rownames = FALSE, hline.after = c(-1,0,0,3,6,9,nrow(print_gof_stats)))
 @

### 2017-11-02 ASApaper/paper-child.Rnw
some more formatting

@@
 levels(experdet$test_param) <- paste0("beta", c(1,3,4,2,6,5))
+experdet$test_param <- factor(experdet$test_param, levels=paste0("beta", c(1,2,3,5,4,6)))
+experdet <- experdet %>% arrange(test_param)
 experdet$test_param <- as.character(experdet$test_param)
-names(experdet) <- c("Parameter", "Lineup Type", "Easy Value", "Medium Value", "Hard Value")
-
-print(xtable::xtable(experdet, caption = "All conditions in our MTurk experiment.
-Note that for $\\beta_4$ and $\\beta_6$, the negative type experiments are the reverse lineups, where 5 plots were simulated from the model with the parameter value given in the table, while 1 plot was simulated from M1.", label = "tab:experdetail", digits = c(0,0,0,3,4,3)))
+experdet$test_param[c(2,4,6,8,10,12)] <- NA
+experdet$type2 <- c(rep(c("neg", "pos"), 4), rep(c("regular", "reverse"), 2))
+names(experdet) <- c("Parameter", "Condition", "Easy Value", "Medium Value", "Hard Value")
+
+print(xtable::xtable(experdet, caption = "All conditions used for the MTurk experiment.
+For parameters $\beta_1, \beta_2, \beta_3,$ and $\beta_5$ M1 served as null model.
+For $\\beta_4$ and $\\beta_6$, null model M1 and the alternative model switch roles in the reversed lineups, i.e.
+five plots show data simluated from the laternative model and only one plot shows data from M1.", label = "tab:experdetail", digits = c(0,0,0,3,4,3)), include.rownames=FALSE,
+      hline.after = c(-1,0,0,8,8,12))
 @
@@
 pvalidx <- which(res[,4] < .0001)
-pval <- as.numeric(round(res[,4],4))
+pval <- sprintf("%.4f",round(res[,4],4))
 pval[pvalidx] <- "$<$0.0001"
@@
 
+\npdecimalsign{.}
+%\nprounddigits{3}
+
 \begin{table}
 \centering
-\begin{tabular}{lrrrr}
-Parameter & Estimate & Std Error & $p$-value & Odds Multiplier \\
-$\alpha_{1-1}$ & \Sexpr{ests[1]} & \Sexpr{se[1]} & \Sexpr{pval[1]} & \Sexpr{odds[1]}\\
-$\alpha_{2-1}$ & \Sexpr{ests[2]} & \Sexpr{se[2]} & \Sexpr{pval[2]} & \Sexpr{odds[2]} \\
-$\alpha_{3-1}$ & \Sexpr{ests[3]} & \Sexpr{se[3]} & \Sexpr{pval[3]} & \Sexpr{odds[3]}  \\
-$\alpha_{4-1}$  & \Sexpr{ests[4]} & \Sexpr{se[4]} & \Sexpr{pval[4]} & \Sexpr{odds[4]} \\
-$\alpha_{5-1}$ & \Sexpr{ests[5]} & \Sexpr{se[5]} & \Sexpr{pval[5]} & \Sexpr{odds[5]}  \\
-$\alpha_{6-1}$  & \Sexpr{ests[6]} & \Sexpr{se[6]} & \Sexpr{pval[6]} & \Sexpr{odds[6]} \\
-$\alpha_{11}$ & \Sexpr{ests[7]} & \Sexpr{se[7]} & \Sexpr{pval[7]} & \Sexpr{odds[7]} \\
-$\alpha_{21}$  & \Sexpr{ests[8]} & \Sexpr{se[8]} & \Sexpr{pval[8]} & \Sexpr{odds[8]} \\ 
-$\alpha_{31}$ & \Sexpr{ests[9]} & \Sexpr{se[9]} & \Sexpr{pval[9]} & \Sexpr{odds[9]} \\
-$\alpha_{41}$  & \Sexpr{ests[10]} & \Sexpr{se[10]} & \Sexpr{pval[10]} & \Sexpr{odds[10]} \\
-$\alpha_{51}$  & \Sexpr{ests[11]} & \Sexpr{se[11]} & \Sexpr{pval[11]} & \Sexpr{odds[11]} \\
-$\alpha_{61}$  & \Sexpr{ests[12]} & \Sexpr{se[12]} & \Sexpr{pval[12]} & \Sexpr{odds[12]} \\
-$\beta_{1-1}$  & \Sexpr{ests[13]} & \Sexpr{se[13]} & \Sexpr{pval[13]} & \Sexpr{odds[13]} \\
-$\beta_{2-1}$   & \Sexpr{ests[14]} & \Sexpr{se[14]} & \Sexpr{pval[14]} & \Sexpr{odds[14]} \\
-$\beta_{3-1}$  & \Sexpr{ests[15]} & \Sexpr{se[15]} & \Sexpr{pval[15]} & \Sexpr{odds[15]} \\
-$\beta_{4-1}$  & \Sexpr{ests[16]} & \Sexpr{se[16]} & \Sexpr{pval[16]} & \Sexpr{odds[16]} \\
-$\beta_{5-1}$ & \Sexpr{ests[17]} & \Sexpr{se[17]} & \Sexpr{pval[17]} & \Sexpr{odds[17]} \\
-$\beta_{6-1}$ & \Sexpr{ests[18]} & \Sexpr{se[18]} & \Sexpr{pval[18]} & \Sexpr{odds[18]} \\
-$\beta_{11}$ & \Sexpr{ests[19]} & \Sexpr{se[19]} & \Sexpr{pval[19]} & \Sexpr{odds[19]} \\
-$\beta_{21}$ & \Sexpr{ests[20]} & \Sexpr{se[20]} & \Sexpr{pval[20]} & \Sexpr{odds[20]}  \\
-$\beta_{31}$ & \Sexpr{ests[21]} & \Sexpr{se[21]} & \Sexpr{pval[21]} & \Sexpr{odds[21]} \\
-$\beta_{41}$ & \Sexpr{ests[22]} & \Sexpr{se[22]} & \Sexpr{pval[22]} & \Sexpr{odds[22]} \\
-$\beta_{51}$ & \Sexpr{ests[23]} & \Sexpr{se[23]} & \Sexpr{pval[23]} & \Sexpr{odds[23]} \\
-$\beta_{61}$ & \Sexpr{ests[24]} & \Sexpr{se[24]} & \Sexpr{pval[24]} & \Sexpr{odds[24]} \\ 
-$\sigma^2_{\delta}$ & \Sexpr{round(sigma_delta, 4)} & -- & -- & --\\
-$\sigma^2_{\epsilon}$ &  \Sexpr{round(sigma_epsilon, 4)}  & -- & -- & --
+\begin{tabular}{cn{5}{2}n{5}{3}rr}\hline
+Parameter & \multicolumn{1}{r}{Estimate} & \multicolumn{1}{r}{Std Error} & $p$-value & Odds Multiplier \\ \hline\hline
+$\alpha_{1+}$ & \Sexpr{ests[7]} & \Sexpr{se[7]} & \Sexpr{pval[7]} & \Sexpr{odds[7]} \\
+$\alpha_{1-}$ & \Sexpr{ests[1]} & \Sexpr{se[1]} & \Sexpr{pval[1]} & \Sexpr{odds[1]}\\
+$\beta_{1+}$ & \Sexpr{ests[19]} & \Sexpr{se[19]} & \Sexpr{pval[19]} & \Sexpr{odds[19]} \\
+$\beta_{1-}$  & \Sexpr{ests[13]} & \Sexpr{se[13]} & \Sexpr{pval[13]} & \Sexpr{odds[13]} \\ \hline
+$\alpha_{2+}$  & \Sexpr{ests[8]} & \Sexpr{se[8]} & \Sexpr{pval[8]} & \Sexpr{odds[8]} \\ 
+$\alpha_{2-}$ & \Sexpr{ests[2]} & \Sexpr{se[2]} & \Sexpr{pval[2]} & \Sexpr{odds[2]} \\
+$\beta_{2+}$ & \Sexpr{ests[20]} & \Sexpr{se[20]} & \Sexpr{pval[20]} & \Sexpr{odds[20]}  \\
+$\beta_{2-}$   & \Sexpr{ests[14]} & \Sexpr{se[14]} & \Sexpr{pval[14]} & \Sexpr{odds[14]} \\ \hline
+$\alpha_{3+}$ & \Sexpr{ests[9]} & \Sexpr{se[9]} & \Sexpr{pval[9]} & \Sexpr{odds[9]} \\
+$\alpha_{3-}$ & \Sexpr{ests[3]} & \Sexpr{se[3]} & \Sexpr{pval[3]} & \Sexpr{odds[3]}  \\
+$\beta_{3+}$ & \Sexpr{ests[21]} & \Sexpr{se[21]} & \Sexpr{pval[21]} & \Sexpr{odds[21]} \\ 
+$\beta_{3-}$  & \Sexpr{ests[15]} & \Sexpr{se[15]} & \Sexpr{pval[15]} & \Sexpr{odds[15]} \\ \hline\hline
+$\alpha_{5+}$  & \Sexpr{ests[11]} & \Sexpr{se[11]} & \Sexpr{pval[11]} & \Sexpr{odds[11]} \\
+$\alpha_{5-}$ & \Sexpr{ests[5]} & \Sexpr{se[5]} & \Sexpr{pval[5]} & \Sexpr{odds[5]}  \\
+$\beta_{5+}$ & \Sexpr{ests[23]} & \Sexpr{se[23]} & \Sexpr{pval[23]} & \Sexpr{odds[23]} \\
+$\beta_{5-}$ & \Sexpr{ests[17]} & \Sexpr{se[17]} & \Sexpr{pval[17]} & \Sexpr{odds[17]} \\ \hline
+$\alpha_{4+}$  & \Sexpr{ests[10]} & \Sexpr{se[10]} & \Sexpr{pval[10]} & \Sexpr{odds[10]} \\
+$\alpha_{4-}$  & \Sexpr{ests[4]} & \Sexpr{se[4]} & \Sexpr{pval[4]} & \Sexpr{odds[4]} \\
+$\beta_{4+}$ & \Sexpr{ests[22]} & \Sexpr{se[22]} & \Sexpr{pval[22]} & \Sexpr{odds[22]} \\
+$\beta_{4-}$  & \Sexpr{ests[16]} & \Sexpr{se[16]} & \Sexpr{pval[16]} & \Sexpr{odds[16]} \\ \hline
+$\alpha_{6+}$  & \Sexpr{ests[12]} & \Sexpr{se[12]} & \Sexpr{pval[12]} & \Sexpr{odds[12]} \\
+$\alpha_{6-}$  & \Sexpr{ests[6]} & \Sexpr{se[6]} & \Sexpr{pval[6]} & \Sexpr{odds[6]} \\
+$\beta_{6+}$ & \Sexpr{ests[24]} & \Sexpr{se[24]} & \Sexpr{pval[24]} & \Sexpr{odds[24]} \\ 
+$\beta_{6-}$ & \Sexpr{ests[18]} & \Sexpr{se[18]} & \Sexpr{pval[18]} & \Sexpr{odds[18]} \\
+\hline\hline
+$\sigma^2_{\delta}$ & \Sexpr{round(sigma_delta, 4)} & \multicolumn{1}{r}{--} & -- & --\\
+$\sigma^2_{\epsilon}$ &  \Sexpr{round(sigma_epsilon, 4)}  & \multicolumn{1}{r}{--} & -- & --\\
+\hline 
 \end{tabular}

### 2017-11-15 ASApaper/paper-child.Rnw
make some changes (\st)

--- 2017-11-01 ASApaper/paper-child.Rnw
@@
-%\newcommand{\st}[1]{{\color{orange} #1}}
-%\newcommand{\hh}[1]{{\color{magenta} #1}}
+% \newcommand{\st}[1]{{\color{orange} #1}}
+% \newcommand{\hh}[1]{{\color{magenta} #1}}
 
@@
 % }
-Three of the most important pieces of statistical modeling are significance testing of model parameters, goodness-of-fit tests, and power of a test.
-In the first, the data are usually assumed to come from a simple model under the null hypothesis, and additional parameters are tested whether they significantly contribute to explaining variability in the data.
+Three of the most important pieces of statistical modeling are significance testing of model parameters, \st{model goodness-of-fit tests, and determining} power of a \st{hypothesis} test.
+In the first, the data are usually assumed to come from a simple model under the null hypothesis, and additional parameters are tested \st{to determine} whether they significantly contribute to explaining the variability in the data.
 In the second, the model of interest is examined to determine how well it fits the data.
@@
 We refer to the ``things" in the network as \textit{nodes}, or \textit{actors} in a social network, and the connections as \textit{edges}, or \textit{ties} in a social network.
-Dependencies inherent to the data make network objects particularly difficult to model.
-Even more challenging is the situation when we go beyond single instances of a network and consider the dynamics of network change between observed instances.
+Dependencies inherent to the data make network objects particularly difficult to model. %Even more challenging is the situation 
+\st{THe difficulty of modeling increases} when we go beyond single instances of a network and consider the dynamics of network change between observed instances.
 This type of modelling for dynamic networks is often performed on social network data, such as friendship networks among students or the spread of HIV in drug users sharing needles.
@@
 X
-These models lack the asymptotics required to perform many well-known goodness-of-fit tests, and the maximum likelihood estimation of parameters is so difficult that it can make significance testing difficult as well \citep{goldenberg09}.
+%These 
+\st{It is known that most network} models lack the asymptotics required to perform many well-known goodness-of-fit tests, and the maximum likelihood estimation of parameters is so difficult that it can make significance testing difficult as well \citep{goldenberg09}.
 %\hh{is the previous sentence a quote? I'm not sure that I follow.} %\st{it's a paraphrase... shall I make a quote? Also I definitely botched a word. Hopefully makes more sense now? }
 
-We propose new methods for significance and goodness-of-fit testing and power calculation of these for a set of social network models, stochastic actor-oriented models for dynamic network data \citep{saompaper}.
-Specifically, we are using \textit{visual inference} in place of traditional statistical methods for social network models, such as Wald tests for significance of parameters and in- and outdegree distribution metrics for determining goodness-of-fit.
-Visual inference, introduced by \citet{Bujaetal}, allows us to look at the entire dataset simulated from a network model as opposed to a (set of) usually one-dimensional metric(s) derived from the network such as outdegree, or a $p$-value for a single parameter in the model.
-%In addition, we use visual inference to better understand some parameters in a social network model, by looking at how different values for parameters affect the networks simulated from these models.
+\st{In order to circumvent these difficulties,} we propose new methods for significance \st{testing of parameters}, goodness-of-fit testing, and power calculations %of these 
+for a set of social network models: stochastic actor-oriented models for dynamic network data \citep{saompaper}.
+Specifically, we are using \textit{visual inference} in place of traditional statistical methods for social network models, such as Wald or $t$-tests for significance of parameters and in- and outdegree distribution metrics for determining goodness-of-fit.
+Visual inference, introduced by \citet{Bujaetal}, allows us to look at the \textit{entire} dataset simulated from a network model as opposed to a (set of) usually one-dimensional metric(s) derived from the network such as outdegree, or a $p$-value for a single parameter in the model.
+\st{By using visual inference to supplement traditional statistical tests, we gain a new understanding of parameters in SAOMs and fit of SAOMs to data.}%In addition, we use visual inference to better understand some parameters in a social network model, by looking at how different values for parameters affect the networks simulated from these models.
 
@@
 Here, we use the lineup protocol.
-Under this protocol, a plot of the observed data is placed randomly among a set of $m-1$ null plots (where $m =20$, usually), and human observers are then asked to examine the lineup and to identify the most different plot.
+\st{Under this protocol, we begin with a data set of interest to us, such as a network, and a visualization of this data, such as a node-link diagram.
+We would like to know the model that generated this data.
+There are two hypothesis considered: a null hypothesis which states the model that is assumed to have generated the data, and an alternative hypothesis that the data were not generated under this model.} Then, the plot of the data of interest is placed randomly among a set of $m-1$ null plots, which are visualizations of data simulated from the null model(where $m =20$, usually).
+Human observers are then asked to examine the lineup and to identify the \st{plots that looks} most different \st{from the others}.
 If an observer identifies the data plot, this is quantifiable evidence against the null hypothesis.
@@
 The lineup protocol places a plot firmly in the framework of hypothesis tests: a plot of the data is considered to be the test statistic, which is compared against the sampling distribution under the null hypothesis represented by the null plots.
-Obviously, the null generating mechanism, i.e.\ the method of obtaining the data for null plots, is crucial for both the lineup and the Rorschach protocol, as the null hypothesis directly affects the choice of null generating method.
+Obviously, the null generating mechanism, i.e.\ the method of obtaining the data for null plots, is crucial for the lineup protocol, as the null hypothesis directly affects the choice of null generating method.
 Null generating methods are typically based on (a) simulation, if the null hypothesis allows us to directly specify a parametric model, (b) sampling, as for example in the case of large data sets, or (c) permutation of the original data \citep[see e.g.\ ][]{Good05}, which allows for non-parametric testing that preserves marginal distributions  while ensuring independence in higher dimensions.
-The model of interest here allows us to simulate directly from a parameteric model for dynamic social network data.
+The model of interest here allows us to simulate directly from a parametric model for dynamic social network data.
 %In the experimental data that we analyzed the null generating methods used were permutation methods and direct simulation from a null model.
@@
 The results suggest that  visual inference is comparable to conventional tests in a controlled conventional setting.
-This provides support for its appropriateness for testing in real exploratory situations where no conventional test exists.
+This provides support for its appropriateness for testing in real exploratory situations where no conventional test exists or is difficult.
 %Interestingly, the power of a visual test increases with the number of observers engaged to evaluate lineups, and the pattern in results suggests that the power will provide results consistent with practical significance \citep{kirk:1996}.
@@
 %To model family that we apply to the Senate data is the Stochastic Actor-Oriented Model (SAOM) family. 
-Stochastic Actor-Oriented Models (SAOMs) are a family of models for dynamic network data \citep{saompaper} that incorporate both network structure and node-level information to describe how a network observed on two or more occasions changes over time.
-The two titular properties of SAOMs, stochasticity and actor-orientation, are crucial to understanding networks as they exist naturally: social networks are ever-changing as relationships decay or grow in seemingly random ways, and most actors in them have characteristics that could affect how they change their ties to other nodes in the network.
+Stochastic Actor-Oriented Models (SAOMs) are a family of models for dynamic network data that incorporate both network structure and node-level information to describe how a network observed on two or more occasions changes over time \citep{saompaper}.
+The two titular properties of SAOMs, stochasticity and actor-orientation, are crucial to understanding networks as they exist naturally: social networks are ever-changing as relationships decay or grow in seemingly random ways, and the actors in them have characteristics that could affect how they change their ties to other nodes in the network.
 % Talk about SAOMs
 These unique properties allow for the fitting of some very complicated models to inherently complex data, so it can be exceedingly difficult to interpret parameters and their corresponding estimates.
-The sheer amount of possible parameters to include in the model combined with the difficulty of interpretation make parameter selection and goodness-of-fit testing burdensome as well.
-
-Broadly, a SAOM takes network structure and node covariate information into account in two ways and models the network changes as a continuous time Markov chain (CTMC).
+The sheer amount of available parameters to include in the model combined with the difficulty of interpretation make parameter selection and goodness-of-fit testing burdensome as well.
+
+Broadly, a SAOM takes network structure and node covariate information into account in two ways and models the network changes \st{one-at-a-time} as a continuous time Markov chain (CTMC).
 First, the rate of change between states is dictated by a rate function that describes \textit{how often} changes in the network occur, and secondly, the objective function describes \textit{what} those state changes are.
-As in many other network models, the variables of interest,  are the binary edges of the network.
-Let $x_{ij}$ denote the edge between nodes $i$ and $j$, where $i,j \in \{1, 2, \dots, n=\text{the number of nodes}\}$.
+As in most other network models, the variables of interest are the edges of the network, $x_{ij}$, where $x_{ij}$ denotes the edge between nodes $i$ and $j$, where $i,j \in \{1, 2, \dots, n=\text{the number of nodes}\}$.
 $x_{ij}$ is modelled as a binary variable, i.e.
@@
 Edges are treated as \textit{directed}, i.e. in general $x_{ij} \neq x_{ji}$, and self-referencing edges or loops are not allowed, i.e.  $x_{ii} = 0$ for all $i$. 
-Assume, the network is observed $M$ times at time points  $t_1 < t_2 < ...
-< t_M$, then the  entire network at time point $t_m$ is denoted as $x(t_m)$.
+The network is observed $M \geq 2$ times at time points  $t_1 < t_2 < ...
+< t_M$, and the entire network at time point $t_m$ is denoted as $x(t_m)$.
 In sections~\ref{rf} and~\ref{of} we discuss the rate and objective functions of a SOAM in more depth.
-Additional details on SOAMs can be found in \citet{saompaper, snijders01, snijders2010, snijdersetall:2007, snijders:2010, snijders:2017},
+Additional details on SOAMs can be found in \citet{saompaper, snijders01, snijders:2010, snijders2010, snijdersetall:2007, snijders:2017},
 
@@
 In general, the rate function can take the network structure e.g.
-outdegree of node $i$, and the node covariates into account, but we use the simple rate function, which is constant over all nodes in a given time period.
+outdegree of node $i$, and the node covariates into account, but we use the simple rate function, which is constant over all nodes in a given time period, so that $\lambda_i \equiv \alpha_m$ $\forall i \in \{1, 2, \dots, n\}$.
 We denote the rate from $t_{m}$ to $t_{m+1}$ as $\alpha_m$ for $m = 1, \dots, M-1$.
-Using this notation, the waiting time to the next chance for actor $i$ to make a change is exponentially distributed with expected value $\alpha_m^{-1}$.
-Since the rate is the same for all actors, the waiting time for \textit{any} actor to get the chance to change is exponentially distributed with expected value $(n\alpha_m)^{-1}$.
+Using this notation, the \textit{waiting time} to the next chance for actor $i$ to make a change is exponentially distributed with expected value $\alpha_m^{-1}$.
+Since the rate is the same for all actors, the waiting time for \textit{any} actor to get the \st{oppurtunity} to change \st{its set of ties} is exponentially distributed with expected value $(n\alpha_m)^{-1}$.
 
@@
 %The objective function, $f_i$, is the driving force of a SAOM. 
-Actor $i$ is aiming to maximize the objective function $f_i$ given the current state of the network, $x$ and the node-level covariates, $\mathbf{Z}$, given as:% which has the form
+Actor $i$ is aiming to maximize \st{its corresponding} objective function $f_i$ given the current state of the network, $x$ and the node-level covariates, $\mathbf{Z}$, given as:% which has the form
 \begin{equation}\label{eq:objective}
-f_i(x, \boldsymbol{\beta}, \mathbf{Z}) = \sum_{k = 1}^K \beta_k s_{ik}(x, \mathbf{Z}),
+f_i(x, \boldsymbol{\beta}, \mathbf{Z}) = \sum_{k = 1}^K \beta_k s_{ik}(x, \mathbf{Z}) + U,
 \end{equation}
-where $\boldsymbol{\beta} = (\beta_1, \dots, \beta_K)$ are additional model parameters, each associated with some network statistics, $s_{ik}(x, \mathbf{Z})$, $s_{ik}(x, \mathbf{Z})$, calculated with respect to actor $i$.
+where $\boldsymbol{\beta} = (\beta_1, \dots, \beta_K)$ are additional model parameters, each associated with some network statistics, $s_{ik}(x, \mathbf{Z})$, $s_{ik}(x, \mathbf{Z})$, calculated with respect to actor $i$ \st{at the current network state $x$}.
 Network statistics range from the simple outdegree, $s_i(x) = \sum_{i\neq j} x_{ij}$, to the more complicated {\it transitive triplets jumping to different covariate}, $s_i(x, \mathbf{Z}) = \sum_{i \neq j \neq h} x_{ij}x_{ih}x_{hj} \cdot \mathbb{I}(z_i = z_h \neq z_j)$.
@@
 
-Objective function $f_{i}(x, \boldsymbol{\beta}, \mathbf{Z})$ and  random component $U$ are combined to form the \textit{transition probability}, $p_{ij}$, of the network changing from its current state $x$ to the state with changed tie $x_{ij}$, denoted as $x(i \leadsto j)$:
+Objective function $f_{i}(x, \boldsymbol{\beta}, \mathbf{Z})$ and  random component $U$ combine to form the \textit{transition probability}, $p_{ij}$, of the network changing from its current state $x$ to the state with changed tie $x_{ij}$, denoted as $x(i \leadsto j)$:
 \begin{equation}\label{eq:transprob}
@@
 The acting node can also choose to \textit{not} change at all.
-This occurs when the numerator, as calculated for the current state of the network, is larger than for any changes $x(i \leadsto j)$ that could be made.
+This \st{is most likely to occur} when the numerator, as calculated for the current state of the network, is larger than for any changes $x(i \leadsto j)$ that could be made.
 
@@
 Similarly, we denote the reciprocity parameter by $\beta_2$ and the associated statistic as $s_{i2}(x) = \sum_{j} x_{ij}x_{ji}$.
-We refer to the model with only these two parameters in the objective function as M1.
+We \st{will} refer to the \st{very simple} model with only these two parameters in the objective function as M1.
 
@@
 names(print_gof_stats) <- c("Model", "Replicate", "Data Picks", "Total Viewers", "p-value")
-print_gof_stats$`p-value` <- ifelse(print_gof_stats$`p-value` < .0001, "< 0.0001", sprintf("%.4f", print_gof_stats$`p-value`))
-print(xtable::xtable(print_gof_stats, label = "tab:gofstats", caption = "An overview of the results from the 12 goodness-of-fit lineup tests.", digits = 4,
-                      align="rllrrr"), caption.placement = "bottom", include.rownames = FALSE)
+print_gof_stats$`p-value` <- ifelse(print_gof_stats$`p-value` < .0001, "< 0.0001", as.character(round(print_gof_stats$`p-value`,4)))
+print(xtable::xtable(print_gof_stats, label = "tab:gofstats", caption = "An overview of the results from our 12 goodness-of-fit lineup tests.", digits = 4), caption.placement = "bottom", include.rownames = FALSE)
 @

### 2017-11-15 ASApaper/paper.Rnw
minor formatting
some more formatting
make some changes (\st)

@@
 % \usepackage{endfloat} % Figures to the end of the document
+\newcommand{\st}[1]{{\color{orange} #1}}
+\newcommand{\hh}[1]{{\color{magenta} #1}}
 
@@
 \begin{abstract}
-Two of the most imporant pieces of statistical modeling are significance testing of model parameters and goodness-of-fit measures and tests.
-The more complicated the model, the harder it is to determine fit or whether to include or exclude a parameter.
+\st{
+Three of the most important pieces of statistical modeling are significance testing of model parameters, goodness-of-fit tests, and power of a test.
+All three of these aspects of statistical modeling increase greatly in difficulty as the complexity of the model of interest increases.
 Some particularly complicated sets of models are those designed to model network change.
+We propose new methods for significance and goodness-of-fit testing, and power calculation of these for a set of social network models, stochastic actor-oriented models for dynamic network data \citep{saompaper}.
 By using the visual inference methodology of \citet{Bujaetal}, we can look at the entire dataset simulated from a network model as opposed to a single metric on the network such as outdegree, or a $p$-value for a single parameter in the model.
+}
 \end{abstract}

### 2017-11-17 ASApaper/paper-child.Rnw
more edits

@@
 \begin{equation}\label{eq:objective}
-f_i(x, \boldsymbol{\beta}, \mathbf{Z}) = \sum_{k = 1}^K \beta_k s_{ik}(x, \mathbf{Z}) + U,
+f_i(x, \boldsymbol{\beta}, \mathbf{Z}) = \sum_{k = 1}^K \beta_k s_{ik}(x, \mathbf{Z}),
 \end{equation}
@@
 The data we use are collaboration networks in the United States Senate during the $111^{th}$ through $114^{th}$ Congresses, overlapping with Barack Obama's presidency.
-These senates began on January 6, 2009 and ended on January 3, 2017\footnote{Details of how this data can be downloaded are provided by François Briatte at \url{https://github.com/briatte/congress}}.
-There are three legislative ways that senators can show support for legislation: they can author a bill, cosponsor a bill, and vote for a bill.
+These senates began on January 6, 2009, the start date of the $111^{th}$, and ended on January 3, 2017, the last date of the $114^{th}$\footnote{Details of how this data can be downloaded are provided by François Briatte at \url{https://github.com/briatte/congress}.}.
+In the US Senate, there are three ways that senators can show support for a piece of legislation: they can author the bill, cosponsor the bill, and vote for the bill.
 We use cosponsorship as a metric because it results in a network that is unimodal (all nodes are senators) and directed.
@@
 \end{equation}
-For each of the four senate sessions, we have the WPC value between any two senators in the session, the party affiliation of each senator, the number of bills they authored in each session, and their gender.
+For each of the four senate sessions, \st{in addition to the} WPC value between any two senators in the session, we have three node covariates: the party affiliation of each senator, the number of bills they authored in each session, and their gender.
 We explored each of these covariates in the model to determine if they affect the overall network structure and how ties are formed between senators.
@@
 She is isolated in Figure~\ref{fig:senateAll}, but in actuality, she had many cosponsors on two pieces of legislation she authored in that short time, as is shown in Figure~\ref{fig:senateClinton}.
-We chose to remove Clinton and her edges from the network because they make the overall structure look so different from the other three senates, showing that the pattern is not typical of a senate in any other year.
-We suspect that because Hillary Clinton had just been appointed Secretary of State, the cosponsorships were largely symbolic, so the $111^{th}$ Senate without Hillary Clinton is more typical than the $111^{th}$ Senate with her.
+We chose to remove Clinton and her edges from the network because they make the overall structure look so different from the other three senates, \st{where} the pattern is not typical of a senate in any other year.
+We suspect that because Hillary Clinton had \st{recently run for President had just} been \st{selected as Obama's} Secretary of State, the cosponsorships were largely symbolic, so the $111^{th}$ Senate without Hillary Clinton is more typical than the $111^{th}$ Senate with her.
 
@@
 
-In legislative cosponsorship networks, it is well known that party affiliation and reciprocity of relationships are major influences on structure \citep{legnet}.
-We focus on these two covariates when choosing which SAO models to fit to the data.
+In legislative cosponsorship networks, it is well known that party affiliation, reciprocity of relationships, and whether senators are "workhorses" who author many bills or "showhorses" who author few bills, are major influences on structure \citep{legnet}.
+We focus on these covariates, \st{plus the additional sex covariate} when choosing which SAO models to fit to the data.
 
@@
 \item Define the simple effects structure of the data: the rate parameters and the outdegree and reciprocity parameters. 
-\item Add each additional possible evaluation effect in \texttt{RSiena} one-at-a-time to the model structure, as determined by the effects documentation function \citep{RSiena}.
+\item Add each additional possible effect, \st{as determined by the effects documentation function}, in \texttt{RSiena} one-at-a-time to the model\st{'s objective function}  \citep{RSiena}.
 \item Fit each model to the data and check for convergence.
@@
 The most significant effect was the jumping transitive triplet (JTT) parameter for the party covariate, which was estimated to be about -6 with a standard error of 0.11, resulting in a $p$-value of less than 0.0001.
-This estimate of the parameter associated with this statistic relies on the number of transitive closures formed between two senators from different parties.
-The negative estimate is an indication that forming transitive ties between two people from different parties is discouraged, which tracks with the divisive nature of American politics, where party affilitation is dominant.
+This estimate of the parameter associated with this statistic considers the number of transitive closures formed between two senators from different parties.
+The negative estimate is an indication that forming transitive ties between two people from different parties is discouraged, which tracks with the divisive nature of American politics, where party affilitation is \st{the dominant effects}.
 Another significant effect was the same JTT parameter for the sex covariate, with an estimate of about 3 with a standard error of 0.89.
-The covariate-related similarity score-weighted transitive triplets parameter estimate for the number of bills authored by a senator was also significant.
-This effect was estimated at about 10 with standard error of 3.9, and the high positive effect suggests senators tend to collaborate with other senators who author about the same number of bills they do.
+\st{This parameter also consideres transitive closures, but for senators of different genders.
+The positive value indicates that transitive ties between senators of different genders are more likley to form.} The covariate-related similarity score-weighted transitive triplets parameter estimate for the number of bills authored by a senator was also significant.
+\st{We chose to look at similarity because the number of bills authored is more continuous than gender or party.
+The similarity measure is computed as:
+\begin{equation}\label{eq:similar}
+sim^b_{ij} = \frac{\max_{hk}|b_h - b_k| - |b_i - b_j|}{\max_{hk}|b_h - b_k|} 
+\end{equation}
+where $\max_{hk}|b_h - b_k|$ is the range of number of bills authored by senators, and $b_i$, $b_j$ are the number of bills authored by senators $i,j$ repectively in the senate period.} This effect was estimated at about 10 with standard error of 3.9.
+The high positive estimate suggests senators \st{are encouraged} to collaborate with other senators who author about the same number of bills they do.
 This tendency of senators to cosponsor bills written by senators who are similarly ``prolific" corresponds to another well-known property of the U.S.
 Senate structure: the tendency of senators to be either ``workhorses" or ``showhorses".
-Senators known as workhorses author many pieces of legislation in a session, and largley stay out of the public arena.
+Senators known as workhorses author many pieces of legislation in a session, and largely stay out of the public arena.
 The showhorse senators, on the other hand, author relatively few pieces of legislation, and tend to appear on television, radio, and other media a great deal.
-Finally, we found the same party transitive triplet effect was also significant, with a fitted value of 1.3 and standard error of 0.7, meaning that transitive relationships between senators tend to form when they are from the same party.
+Finally, we found the same party transitive triplet effect was also significant, with a fitted value of 1.3 and standard error of 0.7, meaning that transitive relationships between senators tend to form when they are from the same party, \st{exactly as we would expect in a legislative body in a country with extremely entrenced partisan divides as in the US.}
 
@@
 \item Model M1: $f_{i}(x, \boldsymbol{\beta}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x)$
-\item Model M2: $f_{i}(x, \boldsymbol{\beta}, \mathbf{p}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_3 s_{i3}(x, \mathbf{p})$
-\item Model M3: $f_{i}(x, \boldsymbol{\beta}, \mathbf{s}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_4 s_{i4}(x, \mathbf{s})$
-\item Model M4: $f_{i}(x, \boldsymbol{\beta}, \mathbf{b}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_5 s_{i5}(x, \mathbf{b})$
-\item Model M5: $f_{i}(x, \boldsymbol{\beta}, \mathbf{p}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_6 s_{i6}(x, \mathbf{p})$
-\item Model M6: $f_{i}(x, \boldsymbol{\beta}, \mathbf{p}, \mathbf{b}, \mathbf{s}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_4 s_{i4}(x, \mathbf{s}) + \beta_5 s_{i5}(x, \mathbf{b}) + \beta_6 s_{i6}(x, \mathbf{p})$
+\item Model M3: $f_{i}(x, \boldsymbol{\beta}, \mathbf{p}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_3 s_{i3}(x, \mathbf{p})$
+\item Model M4: $f_{i}(x, \boldsymbol{\beta}, \mathbf{s}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_4 s_{i4}(x, \mathbf{s})$
+\item Model M5: $f_{i}(x, \boldsymbol{\beta}, \mathbf{b}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_5 s_{i5}(x, \mathbf{b})$
+\item Model M6: $f_{i}(x, \boldsymbol{\beta}, \mathbf{p}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_6 s_{i6}(x, \mathbf{p})$
+\item Model M7: $f_{i}(x, \boldsymbol{\beta}, \mathbf{p}, \mathbf{b}, \mathbf{s}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_4 s_{i4}(x, \mathbf{s}) + \beta_5 s_{i5}(x, \mathbf{b}) + \beta_6 s_{i6}(x, \mathbf{p})$
 \end{enumerate}
@@
 M1 & \Sexpr{round(modelmeans$alpha1[1], 3)} & \Sexpr{round(modelmeans$alpha2[1], 3)} & \Sexpr{round(modelmeans$alpha3[1], 3)} & \Sexpr{round(modelmeans$beta1[1], 3)} & \Sexpr{round(modelmeans$beta2[1], 3)} & -- & -- & -- & --\\
-M2 & \Sexpr{round(modelmeans$alpha1[2], 3)} & \Sexpr{round(modelmeans$alpha2[2], 3)} & \Sexpr{round(modelmeans$alpha3[2], 3)} & \Sexpr{round(modelmeans$beta1[2], 3)} & \Sexpr{round(modelmeans$beta2[2], 3)} & \Sexpr{round(modelmeans$beta3[2], 3)} & -- & -- & --\\
-M3 & \Sexpr{round(modelmeans$alpha1[3], 3)} & \Sexpr{round(modelmeans$alpha2[3], 3)} & \Sexpr{round(modelmeans$alpha3[3], 3)} & \Sexpr{round(modelmeans$beta1[3], 3)} & \Sexpr{round(modelmeans$beta2[3], 3)} & -- & \Sexpr{round(modelmeans$beta4[3], 3)} & -- & --\\
-M4 & \Sexpr{round(modelmeans$alpha1[5], 3)} & \Sexpr{round(modelmeans$alpha2[5], 3)} & \Sexpr{round(modelmeans$alpha3[5], 3)} & \Sexpr{round(modelmeans$beta1[5], 3)}& \Sexpr{round(modelmeans$beta2[5], 3)} & -- & -- & \Sexpr{round(modelmeans$beta5[5], 3)} & -- \\
-M5 & \Sexpr{round(modelmeans$alpha1[4], 3)} & \Sexpr{round(modelmeans$alpha2[4], 3)} & \Sexpr{round(modelmeans$alpha3[4], 3)} & \Sexpr{round(modelmeans$beta1[4], 3)} & \Sexpr{round(modelmeans$beta2[4], 3)} & -- & -- & -- & \Sexpr{round(modelmeans$beta6[4], 3)}\\
-M6 & \Sexpr{round(M6[1], 3)} & \Sexpr{round(M6[2], 3)} & \Sexpr{round(M6[3], 3)} & \Sexpr{round(M6[4], 3)} & \Sexpr{round(M6[5], 3)} & -- & \Sexpr{round(M6[6], 3)} & \Sexpr{round(M6[8], 3)} & \Sexpr{round(M6[7], 3)} \\
+M3 & \Sexpr{round(modelmeans$alpha1[2], 3)} & \Sexpr{round(modelmeans$alpha2[2], 3)} & \Sexpr{round(modelmeans$alpha3[2], 3)} & \Sexpr{round(modelmeans$beta1[2], 3)} & \Sexpr{round(modelmeans$beta2[2], 3)} & \Sexpr{round(modelmeans$beta3[2], 3)} & -- & -- & --\\
+M4 & \Sexpr{round(modelmeans$alpha1[3], 3)} & \Sexpr{round(modelmeans$alpha2[3], 3)} & \Sexpr{round(modelmeans$alpha3[3], 3)} & \Sexpr{round(modelmeans$beta1[3], 3)} & \Sexpr{round(modelmeans$beta2[3], 3)} & -- & \Sexpr{round(modelmeans$beta4[3], 3)} & -- & --\\
+M5 & \Sexpr{round(modelmeans$alpha1[5], 3)} & \Sexpr{round(modelmeans$alpha2[5], 3)} & \Sexpr{round(modelmeans$alpha3[5], 3)} & \Sexpr{round(modelmeans$beta1[5], 3)}& \Sexpr{round(modelmeans$beta2[5], 3)} & -- & -- & \Sexpr{round(modelmeans$beta5[5], 3)} & -- \\
+M6 & \Sexpr{round(modelmeans$alpha1[4], 3)} & \Sexpr{round(modelmeans$alpha2[4], 3)} & \Sexpr{round(modelmeans$alpha3[4], 3)} & \Sexpr{round(modelmeans$beta1[4], 3)} & \Sexpr{round(modelmeans$beta2[4], 3)} & -- & -- & -- & \Sexpr{round(modelmeans$beta6[4], 3)}\\
+M7 & \Sexpr{round(M6[1], 3)} & \Sexpr{round(M6[2], 3)} & \Sexpr{round(M6[3], 3)} & \Sexpr{round(M6[4], 3)} & \Sexpr{round(M6[5], 3)} & -- & \Sexpr{round(M6[6], 3)} & \Sexpr{round(M6[8], 3)} & \Sexpr{round(M6[7], 3)} \\
 \hline
 \end{tabular}
-\caption{\label{tab:fittedvalues} The final estimates from repeated estimation of models M1 through M6.}
+\caption{\label{tab:fittedvalues} The final estimates from repeated estimation of our models of interest.}
 \end{table}

### 2017-11-17 ASApaper/paper-child.Rnw
about 2/3 done with edits I want to make

@@
 \hline
-Model & $\alpha_1$ & $\alpha_2$ & $\alpha_3$ & $\beta_1$ & $\beta_2$ & $\beta_3$ & $\beta_4$ & $\beta_5$ & $\beta_6$ \\
+Model & $\hat{\alpha}_1$ & $\hat{\alpha}_2$ & $\hat{\alpha}_3$ & $\hat{\beta}_1$ & $\hat{\beta}_2$ & $\hat{\beta}_3$ & $\hat{\beta}_4$ & $\hat{\beta}_5$ & $\hat{\beta}_6$ \\
 \hline
@@
 \end{tabular}
-\caption{\label{tab:fittedvalues} The final estimates from repeated estimation of our models of interest.}
+\caption{\label{tab:fittedvalues} The final estimates from repeated estimation of our models of interest.
+When simulating from these models, these are the estimates that we will use unless otherwise stated.}
 \end{table}
@@
 
-We want to explore three different aspects of the SAOM models using the lineup protocol: (1) significance of parameters, (2) goodness-of-fit of a model, and (3) visual detection of parameters. 
+We want to explore three different aspects of the SAOM models using the lineup protocol: (1) significance of parameters, (2) goodness-of-fit of a model, and (3) \st{visual power of the effects.} %visual detection of parameters.
 Each one of these situations requires a different setup, which we describe in detail, but we make ue of the lineup protocol for all of these aspects.
@@
 
-To simulate lineups from the models, we set the parameters to the values given in Table~\ref{tab:fittedvalues} for all parameters within the respective models, with the exception of $\beta_5$.
+To simulate lineups from the models we used the \texttt{siena07} function in \texttt{RSiena} \citep{RSiena}.
+\st{Sections~\ref{sec:setupst} through~\ref{sec:setupvp} describe in detail how we set up the lineups, which parameter values we simulate from, and why.} Sets of the lineups we create are shown to independent observers recruited through Amazon Mechanical Turk for feedback (more details on the Turk setup in Section~\ref{sec:res}).
+%, we set the parameters to the values given in Table~\ref{tab:fittedvalues} for all parameters within the respective models, with the exception of $\beta_5$.
 For $\beta_5$, twice the estimated value was used.
 More detail on why we use twice the fitted value is provided in Section~\ref{sec:setupvp}.\footnote{If you would like to explore the kinds of lineups we use in further detail, please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/}}.
-To get the simulations, we used the \texttt{siena07} function in \texttt{RSiena} \citep{RSiena}.
+To get the simulations,
 
@@
 More details about this application are given in Section~\ref{sec:setupvp}.
-In the lineup, M4 is the alternative model with $\beta_5$ set to twice its estimated value given in Table~\ref{tab:fittedvalues}.
+In the lineup, M5 is the alternative model with $\beta_5$ set to twice its estimated value given in Table~\ref{tab:fittedvalues}.
 One plot simulated from this model is placed at random among five observations simulated from the null model, M1.
@@
 \end{figure}
-
-A series of these lineups is shown to independent observers recruited through Amazon Mechanical Turk for feedback (more details on the Turk setup in Section~\ref{sec:res}).
 
@@
 Thus, the five null plots in the lineup are simulations from M1 with $\beta_1, \beta_2$ set to the estimates given in Table~\ref{tab:fittedvalues} for M1.
-The alternate model is the model with $\beta_1, \beta_2$, and $\beta_k$ in the objective function, with the remaining plot simulated from the appropriate model.
+The alternate model is the model with $\beta_1, \beta_2$, and $\beta_k$ in the objective function, with the remaining plot simulated from the appropriate model with values set to the estimates from Table~\ref{tab:fittedvalues}.
 
 The lineup generated under this scenario is shown to a number of independent viewers.
-If an observer picks out the alternative data plot, that is evidence in favor of the null hypothesis, while picking one of the null plots is evidence against the null hypothesis.
-The significance tests that we perform in our experiment are for $\beta_3$ and $\beta_4$, making the alternative models M2 and M3, respectively.
+If an observer picks out the alternative data plot, that is evidence \st{against} the null hypothesis, while picking one of the null plots is evidence \st{in favor of} the null hypothesis.
+The significance tests that we perform in our experiment are for $\beta_3$ and $\beta_4$, making the alternative models M3 and M4, respectively, for three repitions each.
 
@@
 We pick a wave to focus on, wave two, which is the first simulated network, and among these five plots, we place a node-link diagram of the true second wave data.
-We cannot show the data more than once to each participant, so we examine several different models in our Amazon Mechanical Turk experiment, each participant never seeing the true data wave twice.
-The models we chose for goodness-of-fit testing are M2, M3, M4, and M6.
+We cannot show the data more than once to each participant, so we examine several different models with three repitions each in our Amazon Mechanical Turk experiment, with each participant never seeing the true data wave twice.
+The models we chose for goodness-of-fit testing are M3, M4, M5, and M7.
 
@@
 Through visual inference, we want to determine at which point an effect becomes noticeable in a SAOM.
-By \textit{noticeable}, we mean that the inclusion of the effect alters the appearance of networks simulated from a model to a degree that viewers are able to reliably pick out a node-link diagram rendered from data simulated from a model with the effect from a lineup of plots without the effect.
-This is a way to determine the power of the visual test.
+By \textit{noticeable}, we mean that the inclusion of the effect alters the appearance of networks simulated from a model to a degree that viewers are able to reliably pick out a node-link diagram rendered from data simulated from a model \textit{with} the effect in a lineup among plots \textit{without} the effect.
+This is a way to determine the power of the visual \st{significance} test.
 We explore all parameters in the objective function, $\beta_1, \dots, \beta_6$ in this way.
 
-In model M1, with only two parameters in the objective function, we varied both the density and reciprocity parameter values one at a time.
-In models M2 through M5, we vary the additional parameter, $\beta_3$ through $\beta_6$.
-Thus, we have six different parameters of interest to us: $\beta_1, \dots, \beta_6$.
-We want to determine how the size of these parameters affects the overall structure of the network data simulated from the models M1 through M5, so we also vary the value of the parameters in both negative and positive directions.
+In model M1, with only two parameters in the objective function, we varied both the density and reciprocity parameter values one at a time, \st{keeping all other parameters at their fitted values given in Table~\ref{tab:fittedeffects}.
+In models M3 through M6, we vary the additional parameter, $\beta_3$ through $\beta_6$, respectively.
+Thus, we have six different parameters of interest to us in total: $\beta_1, \dots, \beta_6$.
+We want to determine how the size of these parameters affects the overall structure of the network data simulated from the models M1 through M6, so we also vary the value of the parameters \st{in order to determine at what level the effects become noticeable.} %in both negative and positive directions.
 
@@
 <<hypothesis, fig.height=2, fig.cap="We hypothesize that as the parameter value of interest increases in absolute value, more viewers of the lineup will pick the alternative data out of a lineup.
-Note that the significance test we construct in Section~\\ref{sec:setupst} is just one point on the line below, represented by the vertical red line.">>=
+Note that the significance test we construct in Section~\\ref{sec:setupst} is just one point on the line below, represented by the vertical dotted line labeled $\\hat{\\beta}$.
+The easy, medium, and hard lines represent how we determined which values of the parameters to show to our participants, and the horizontal dotted line shows the type-I error for one viewer of a lineup of size 6.">>=
 x <- seq(-10, 10, .05)
@@
   scale_x_continuous(name = "Parameter Value", labels = c("very negative", "negative", "0", "positive", "very positive")) + 
-  #geom_hline(yintercept = 1/6) +
+  geom_hline(yintercept = 1/6, linetype = 'dotted') +
+  geom_vline(xintercept = 0) + 
+  geom_label(data = NULL, aes(x = 0, y = .9, label = "Null Model = Alt. Model"), size = 7) +
   geom_vline(xintercept = 150, color = 'red', linetype = 'dotdash') + 
-    scale_y_continuous(name = "% detecting plot from\nthe alternative model", labels = paste0(c(0,25,50,75,100), "%"))
+  annotate('text', x = 160, y = .9, label = "hat(beta)",parse = TRUE, size = 2) + 
+  geom_vline(xintercept = c(-300,-230, -175), color = c('green', 'yellow', 'red')) + 
+  geom_text(data = NULL, aes(x = -300, y = .75, label = "easy"), angle = 90, vjust = -.5, size = 2) + 
+  geom_text(data = NULL, aes(x = -230, y = .5, label = "medium"), angle = 90, vjust = -.5, size = 2) +
+  geom_text(data = NULL, aes(x = -175, y = .05, label = "hard"), angle = 90, vjust = -.5, size = 2) + 
+  scale_y_continuous(name = "% detecting plot from\nthe alternative model", labels = paste0(c(0,25,50,75,100), "%")) + 
+  theme_bw() + 
+  theme(text = element_text(size = rel(2)))
 @
@@
 To determine the exact values of the six levels we want to test for each effect, we started with the estimates of the parameter at hand (see  Section~\ref{sec:models}), and used small negative and positive factors to determine at what point \textit{we} noticed the effect of the parameter in simulations from the changed models.
-
-For this we constructed an online application that created the lineup protocol for us to be the guinea pigs of our own experiment \citep{theqs}.
-A screen shot of the app we created with the \texttt{shiny} package is shown in Figure~\ref{fig:shiny} \citep{shiny}.
-On the left side of the screen, the user\footnote{Please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/} to create lineups constructed from the models we present for this data for yourself.} can input the information necessary for creating a lineup of the models M1 through M6 for the data in Section~\ref{sec:senatedata}: first, choose to simulate only one plot from the specified model (analogous to changing the alternative model in the lineup protocol) or to simulated $M-1$ plots from the specified model (analogous to changing the null model in the lineup protocol); the model of interest; the wave of the data to examine; if model M1 is selected, whether to alter the density or the reciprocity parameter; the size of the lineup; the amount by which to multiply the effect selected; a random seed for replicability; and a layout algorithm to use for the node-link diagrams.
+\st{In Figure~\ref{fig:hypothesis}, we demonstrate these values with the vertical lines labeled ``easy," ``medium," and ``hard." We expected most viewers to see the effects at the ``easy" values, and we expected very few, if any, viewers to see the effects at the ``hard" values.
+}
+
+\st{To decide on the values to use for each difficulty level,} we constructed an online application that created the lineup protocol for us to be the guinea pigs of our own experiment \citep{theqs}.
+A screen shot of the app we created with the \texttt{shiny} package by \citet{shiny} is shown in Figure~\ref{fig:shiny}.
+On the left side of the screen, the user\footnote{Please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/} to create lineups constructed from the models we present for this data for yourself.} can input the information necessary for creating a lineup of the models M1 through M7 for the data in Section~\ref{sec:senatedata}: first, choose to simulate only one plot from the specified model (analogous to changing the alternative model in the lineup protocol) or to simulate $M-1$ plots from the specified model (analogous to changing the null model in the lineup protocol); the model of interest; the wave of the data to examine; if model M1 is selected, whether to alter the density or the reciprocity parameter; the size of the lineup; the amount by which to multiply the effect selected; a random seed for replicability; and a layout algorithm to use for the node-link diagrams.
 There is also a checkbox if the user wishes the nodes to be colored by the size of the connected component to which they belong.
-The plots that appear that are \textit{not} from the model specified using the other options are simulations from model M1 with the estimates of the rate parameters, $\beta_1$ and $\beta_2$ given in Table~\ref{tab:fittedvalues}.
+The plot(s) that appear(s) that are \textit{not} from the model specified using the other options are simulated from model M1 with the estimates of the rate parameters, and $\beta_1$ and $\beta_2$ as given in Table~\ref{tab:fittedvalues}.
 
@@
 The complete details of the parameters tested using the lineup protocol is given in Table~\ref{tab:experdetail}.
-In the case of both $\beta_4$ and $\beta_6$, we could not determine any values for negative effects that made the data simulated from M3 and M5 look different than null model simulations from model M1.
-Therefore we decided that the lesser experienced participants in our experiment would also not be able to.
-Instead of testing the negative values of these effects, we are examining a different scenario: we placed 5 simulations from positive values of the parameter with one simulation from model M1 in a lineup.
-We refer to this later on as the ``reverse" lineup scenario.
-We used the reverse scenario to determine if the perception of the effect size is symmetric: if an effect is noticed $x$\% of the time at value $\beta_k = \beta_{k_0}$ when one simulation from the corresponding model is placed among five null plots from model M1, then when five simulations from the model with $\beta_k = \beta_{k_0}$ are put in a lineup with one simulation from model M1, the plot from the simpler model should be noticed about $x$\% of the time as well.
+In the case of both $\beta_4$ and $\beta_6$, we could not determine any values for negative effects that made the data simulated from M4 and M6 look different than null model simulations from model M1.
+\st{We hypothesized that this was due to the negative effects \textit{removing} visually interesting structural elements as opposed to \textit{adding} noticeable structural elements.}  Therefore we decided that the lesser experienced participants in our experiment would also not be able to.
+Instead of testing the negative values of these effects, we are examining a different scenario: we placed 5 simulations from model M4 or M6 with positive values of the parameter with one simulation from model M1 in a lineup.
+We will refer to this as the ``reverse" lineup scenario.
+We used the reverse scenario to determine if the perception of the effect size is symmetric: if an effect is noticed $x$\% of the time at value $\beta_k = \beta_{k_0}$ when \textit{one} simulation from the corresponding model is placed among \textit{five} null plots from model M1, then when \textit{five} simulations from the model with $\beta_k = \beta_{k_0}$ are put in a lineup with \textit{one} simulation from model M1, the plot from the simpler model should be noticed about $x$\% of the time as well.
 
@@
 We recruited 250 participants for our experiment through Amazon Mechanical Turk.
-Each participant was peresented with some brief training material before beginning the experiment.
-After agreeing to participate, the participants were shown two trial plots, one where the data plot was the most different from the others due to its complex structure, while the other trial included a data plot that was most different from the others due to its very simple structure.
-Only when participants were able to correctly identify the data plot from the trial lineups, they were allowed to begin the experiment.
+Each participant was presented with some brief training material before beginning the experiment.
+After agreeing to participate, the participants were shown two trial plots, one where the data plot was the most different from the others due to its relatively complex structure, while the other trial included a data plot that was most different from the others due to its comparatively simple structure.
+Only when participants were able to correctly identify the data plot from the trial lineups were they allowed to begin the experiment.
 Each participant was randomly assigned 13 lineups to look at.
 The were asked to select one or more plots that they perceived as ``most different" from the others, and provide a reasoning for their choice.
-They could select from ``Most simple overall structure," ``Most complex overall structure," or ``Other" and provide their own text description of their reasoning.
+They could select from ``Most simple overall structure" or ``Most complex overall structure", \st{corresponding to what they saw in the examples and trials, or they could choose }``Other" and provide their own text description of their reasoning.
+\st{These language in these reasons is purposefully vague: we want participants to tell us what they see.
+We do not want to tell them what they ``should" be seeing.
+In this way, we can truly determine what  effects are \textit{noticeable}.}
 

### 2017-11-27 ASApaper/paper-child.Rnw
more changes - mostly change model numbers and switch around some words.

@@
 
-\subsection{Goodness-of-Fit}
+\subsection{Goodness-of-Fit}\label{sec:gofsetup}
 
@@
 
-In model M1, with only two parameters in the objective function, we varied both the density and reciprocity parameter values one at a time, \st{keeping all other parameters at their fitted values given in Table~\ref{tab:fittedeffects}.
+In model M1, with only two parameters in the objective function, we varied both the density and reciprocity parameter values one at a time, \st{keeping all other parameters at their fitted values given in Table~\ref{tab:fittedvalues}}.
 In models M3 through M6, we vary the additional parameter, $\beta_3$ through $\beta_6$, respectively.
@@
   geom_vline(xintercept = 0) + 
-  geom_label(data = NULL, aes(x = 0, y = .9, label = "Null Model = Alt. Model"), size = 7) +
+  geom_label(data = NULL, aes(x = 0, y = .9, label = "Null Model = Alt. Model"), size = 2) +
   geom_vline(xintercept = 150, color = 'red', linetype = 'dotdash') + 
@@
   geom_vline(xintercept = c(-300,-230, -175), color = c('green', 'yellow', 'red')) + 
-  geom_text(data = NULL, aes(x = -300, y = .75, label = "easy"), angle = 90, vjust = -.5, size = 2) + 
-  geom_text(data = NULL, aes(x = -230, y = .5, label = "medium"), angle = 90, vjust = -.5, size = 2) +
-  geom_text(data = NULL, aes(x = -175, y = .05, label = "hard"), angle = 90, vjust = -.5, size = 2) + 
+  geom_text(data = NULL, aes(x = -300, y = .75, label = "easy"), angle = 90, vjust = -.5, size = 3) + 
+  geom_text(data = NULL, aes(x = -230, y = .5, label = "medium"), angle = 90, vjust = -.5, size = 3) +
+  geom_text(data = NULL, aes(x = -175, y = .05, label = "hard"), angle = 90, vjust = -.5, size = 3) + 
   scale_y_continuous(name = "% detecting plot from\nthe alternative model", labels = paste0(c(0,25,50,75,100), "%")) + 
   theme_bw() + 
-  theme(text = element_text(size = rel(2)))
+  theme(text = element_text(size = rel(3)))
 @
@@
 print(xtable::xtable(experdet, caption = "All conditions used for the MTurk experiment.
-For parameters $\beta_1, \beta_2, \beta_3,$ and $\beta_5$ M1 served as null model.
+For parameters $\\beta_1, \\beta_2, \\beta_3,$ and $\\beta_5$ M1 served as null model.
 For $\\beta_4$ and $\\beta_6$, null model M1 and the alternative model switch roles in the reversed lineups, i.e.
@@
 Twelve of the 13 lineups that the participants saw were used for the significance testing and the visual power methods discussed in Sections~\ref{sec:setupst} and \ref{sec:setupvp}.
-The six parameters, $\beta_1, \dots, \beta_6$ were set to three different values according to how difficult we thought picking the data plot from the lineup would be for our participants, and were also set to be less than or greater than the initial estimate, creating the lineup type variable.%\hh{what about the first 12 plots?
+%The six parameters, $\beta_1, \dots, \beta_6$ were set to three different values according to how difficult we thought picking the data plot from the lineup would be for our participants, and were also set to be less than or greater than the initial estimate, creating the lineup type variable.%\hh{what about the first 12 plots?
 6 parameters, 3 levels, 2 directions}
-One of the 13 lineups the participants saw was the true data from the 112th senate shown in Section~\ref{sec:senatedata}.
+The final of the 13 lineups shown to participants contained the true data from the 112th senate shown in Section~\ref{sec:senatedata} placed among five other plots from models M3, M4, M5, and M7 as discussed in Section~\ref{sec:gofsetup}.
 Each participant only saw the data one time in order to avoid bias.
@@
 For a SAOM, there are two ways a conventional significance test of the parameters can be performed.
-In \texttt{RSiena}, there are $t$-type tests and Wald-type test for a single parameter and for multiple parameters.
-The $t$-type test statistic is simply the parameter estimate divided by its standard error, and compared to a standard normal distribution.
+In \texttt{RSiena}, there are $t$-type and Wald-type tests for a single parameter and for multiple parameters.
+The $t$-type test statistic is simply the parameter estimate divided by its standard error, and compared to a standard normal distribution \citep{RSienamanual}.
 The Wald-type test statistic for a single parameter, $\beta_k$ is
@@
 \end{equation}
-which is compared to a Chi-square distribution with one degree of freedom.
+which is compared to a Chi-square distribution with one degree of freedom\citep{RSienamanual}.
 Testing the significance of multiple parameters depends on the hypothesis we wish to test, and a $P \times K$ matrix, $A$, must be appropriately designed to test the $P$ hypotheses of interest.
@@
 \begin{equation}\label{eq:wald2}
-(A\boldsymbol{\hat{\beta}})' \hat{\Sigma}^{-1} A\boldsymbol{\hat{\beta}} \sim \chi^2_p,
+(A\boldsymbol{\hat{\beta}})' \hat{\Sigma}^{-1} A\boldsymbol{\hat{\beta}} \sim \chi^2_P,
 \end{equation}
@@
 
-All of the parameters we test for significance using the lineup protocol, $\beta_3$ and $\beta_4$, were determined to be statistically significant using Equation~\ref{eq:wald1}.
-The results from the significance tests we performed using the lineup protocol are given in Table~\ref{tab:sigtesting}.
-%\hh{XXX how do we interpret these  visual results?
+Both the parameters we test for significance using the lineup protocol, $\beta_3$ and $\beta_4$, were determined to be statistically significant using Equation~\ref{eq:wald1}.
+The correspoding results from the significance tests we performed using the lineup protocol are given in Table~\ref{tab:sigtesting}.
+\st{Corresponding to traditional methods, if enough participants pick out the alternative plot to result in a $p-value less than 0.05, we reject the null hypothesis that the true value of the additional parameter, either $\beta_3$ or $\beta_4$, is equal to zero.} %\hh{XXX how do we interpret these  visual results?
 Include the six lineups in an appendix and refer to them from here.} \st{XXX should be taken care of in next paragraph now.}
@@
 
-We see that the $p$-values from visual inference, which are calculated using the \texttt{vinference} \texttt{R} package by \citet{vinference}, are highly variable.
+\st{The $p$-values were calculated using the \texttt{vinference} package by \citet{vinference}.
+This package contains methods to calculate \textit{Visual distributions} for lineup experiment data.
+The distribution depends on the number of evaluations of a plot, $K$, the size of the lineup, $m$, and the lineup scenario, which here is that each lineup containing the same data and the same set of null plots is shown to $K$ independent observers.
+The visual inference family of distributions is similar to the binomial distribution, but takes the dependency among the $m$ plots in a single lineup shown to multiple viewers into account.
+Using these $p$-values, all but one lineup results in a rejection of the null hypothesis at Type-I error rate of $\alpha = 0.05$.}.
+We see that the $p$-values for visual inference here are highly variable.
 This variability is introduced through the null plots generated from M1, since not all simulations look alike.
 In addition, the necessarily small number of null plots do not give the viewer as complete of a view of the null model as the usual 19 null plots would.
-The results of the significance tests for $\beta_3$ and $\beta_4$ are not clear cut.
-Thus, unlike the Wald-type tests described at the beginning of this section, there is no way to flat out reject or to fail to reject the null hypothesis that the parameter value is 0.
+The results of the significance tests given in Table~\ref{tab:sigtesting} for $\beta_3$ and $\beta_4$ are not \st{definitive.
+For the test of $\beta_3$, two of the three tests are not significant, while the third is highly significant.
+For the test of $\beta_4$, one test is significant, one is decidedly not significant, and the third is significant at the level of 0.10.}  Thus, unlike the Wald-type tests described at the beginning of this section, there is no way to \st{decisively} reject or to fail to reject the null hypothesis that the parameter value is 0.
 We include all of the lineups shown to our participants in the appendix.
@@
 Goodness-of-fit testing for network models is notoriously difficult.
-Most network models, other than the most simple, lack the asymptotics required to develop the goodness-of-fit methods required \citep{goldenberg09}.
-Some methods have been developed based on what Snijders et al call ``auxiliary statistics" such as the indegree or outdegree distribution on the nodes.
+Most network models, other than the most simple, lack the necessary asymptotics for developing goodness-of-fit methods \citep{goldenberg09}.
+Some methods have been developed based on what \citeauthor{RSienamanual} call ``auxiliary statistics" such as the indegree or outdegree distribution on the nodes.
 In \texttt{RSiena}, the \texttt{sienaGOF} function performs goodness-of-fit testing as follows:
 \begin{enumerate}
-\item Auxiliary statistics are computed on the observed data ($\mathbf{u}_d$) and on $N$ simulated observations from the model ($\mathbf{u}_1 \dots \mathbf{u}_N$). (Usually, $N=1000$) 
+\item Auxiliary statistics, \st{such as the cumulative outdegree counts on the nodes,} are computed on the observed data ($\mathbf{u}_d$) and on $N$ simulated observations from the model ($\mathbf{u}_1 \dots \mathbf{u}_N$).
+(Usually, $N=1000$)
 \item The mean vector, $\overline{\mathbf{u}}$ and covariance matrix, $\mathbf{S}$ of the statistics on the simulations from the model are computed, and the Mahalanobis distance, $d_M(\mathbf{u})$ from the observed statistics to the distribution of the simulated statistics is computed:
@@
 A SAOM is thus considered a good fit to the data if $p$ is large.
-A plot comparing the data to the simulations is also considered, and a similar plot is shown in Figure~\ref{fig:gofsiena}
+A plot comparing the data to the simulations is also considered, and a similar plot is shown in Figure~\ref{fig:gofsiena} \st{for the outdegree distribution of small data set, shown in the points and connected lines, with the simulated values of $\mathbf{u}_d$ shown in boxplots and overlaid violin plots.}
 \end{enumerate}
@@
 <<gofsiena, fig.height=3, fig.cap="An example of what a goodness-of-fit plot from \\texttt{RSiena} looks like.
-The overlaid boxplots and violin plots show the distribution of each of the outdegree values on the simulated networks, and the red points and lines are the observed data values.", cache = TRUE>>=
+The overlaid boxplots and violin plots show the distribution of each of the outdegree count values on the simulated networks, and the red points and lines are the observed data values.", cache = TRUE>>=
 load("data/ansnullpaper.rda")
@@
 
-These methods are similar in that they are both restriced: the \texttt{sienaGOF} method only considers one measure on the data and simulations from the model, while the score-type tests only consider subsets of parameters, ``nuisance parameters" in \citet{scoretest}, not the entire set of parameters.
-By using visual inference instead of more traditional statistical methods, we hope to perform a more holistic goodness-of-fit test.
+These methods are both restriced: the \texttt{sienaGOF} method only considers \textit{one} measure on the data and simulations from the model, while the score-type tests only consider \textit{subsets} of parameters, ``nuisance parameters" in \citet{scoretest}, not the entire set of parameters.
+By using visual inference instead of more traditional statistical methods, we hope to perform a more \textit{holistic} goodness-of-fit test.
 
 Using the lineup protocol, we show each Amazon Mechanical Turk worker the data once, in a lineup with five other plots of simulated data from one of the models we chose.
-We examined four different models, M2, M3, M4, and M6, and examined three repetitions of each, for a total of 12 goodness-of-fit lineups.
+We examined four different models, M3, M4, M5, and M7, and examined three repetitions of each, for a total of 12 goodness-of-fit lineups.
 In each lineup, the ``null model" is one of the four models and the ``alternative" model is the true, unknown model that generated the senate network data.
@@
 \begin{description}
-\item $H_0$:  The senate network data come from (or could have come from) the null model.
+\item $H_0$:  The senate network data come from (or could have come from) the null model, M$i$.
 \item $H_A$: The senate network data do not come from the null model. 
 \end{description}
-If a lineup viewer picks out the data among the five simulations from the null model, it is evidence in favor of the alternative hypothesis.
-On the contrary, if the lineup viewer picks one of the null plots, that is evidence against the alternative hypothsis.
+If a lineup viewer picks out the data among the five simulations from the null model, it is evidence against the null hypothesis.
+On the contrary, if the lineup viewer picks one of the null plots, that is evidence in favor of the null hypothesis.
 Because the size of the lineups is small, the probability of picking the data by chance is high, $\frac{1}{6}$, but if \textit{many} independent viewers pick out the data from the nulls, the evidence in favor of the alternative hypothesis becomes stronger.
@@
 turk22_gof_stats %>% separate(pic_id, into = c("discard", "model", "rep"), sep = c(2,3)) -> print_gof_stats
-print_gof_stats %>% select(-discard) %>% mutate(model = paste0("M", (as.integer(model)-1)))-> print_gof_stats
+print_gof_stats %>% select(-discard) %>% mutate(model = paste0("M", (as.integer(model))))-> print_gof_stats
 names(print_gof_stats) <- c("Model", "Replicate", "Data Picks", "Total Viewers", "p-value")
@@
 
-The $p$-values were calculated using the \texttt{vinference} package by \citet{vinference}.
-This package contains methods to calculate \textit{Visual distributions} for lineup experiment data.
-The distribution depends on the number of evaluations of a plot, $K$, the size of the lineup, $m$, and the lineup scenario, which here is that each lineup containing the same data and the same set of null plots is shown to $K$ independent observers.
-The visual inference family of distributions is similar to the binomial distribution, but takes the dependency among the $m$ plots in a signle lineup shown to multiple viewers into account.
-Using these $p$-values, all but one lineup results in a rejection of the null hypothesis at Type-I error rate of $\alpha = 0.05$.
+\st{The $p$-values were again computed using the \texttt{vinference} package by \citet{vinference}}.
 The lineup that resulted in a failure to reject the null hypothesis is shown in Figure~\ref{fig:failgof}.
-The null model in this lineup is M4, and the senate data is shown in panel number $3^2 - 7$.
+The null model in this lineup is M5, and the senate data is shown in panel number $3^2 - 7$.
 However, the panel most participants chose was number four, and the most common reasoning for that choice was that it had the most simple structure.
@@
 
-The smallest $p$-value for one of the goodness-of-fit lineups was for the third replicate of the null model M4.
-This contradicts our previous finding that the only lineup to fail to reject the null was also when the null model was M4.
-The plot of this lineup is shown in Figure~\ref{fig:mostsiggof}.
-In the remaining replicate of M4 as the null model, 13 of 16 viewers identified the data plot, corresponding to a $p$-values of less than $0.0001$, just like the third replicate.
+The smallest $p$-value for one of the goodness-of-fit lineups was for the third replicate of the null model M5.
+This contradicts our previous finding that the only lineup to fail to reject the null was also when the null model was M5.
+This lineup is shown in Figure~\ref{fig:mostsiggof}.
+In the remaining replicate of M5 as the null model, 13 of 16 viewers identified the data plot, corresponding to a $p$-values of less than $0.0001$, just like the third replicate.
 This variability in results is similar to the variability we found in Section~\ref{sec:sigtest}.
 This variability is again introduced through the plots simulated from null model, and does not provide us with a clear cut decision resulting the hypothesis test.
-For model M4, we can neither reject nor fail to reject the null hypothesis that the data come from model M4.
+For model M5, we can neither reject nor fail to reject the null hypothesis that the data come from model M5.
 This is evidence that the goodness-of-fit of network models cannot always be determined by one dimensional derived features, such as $p$-value shown on the $x$-axis in Figure~\ref{fig:gofsiena}.
@@
 We could move the threshold from $\alpha = 0.05$ to $\alpha = 0.01$ or $\alpha = 0.001$, but more work needs to be done to determine which, if any of these, is appropriate.
-For the other models for which we tested goodness-of-fit, however, we do have significant evidence from all three replicates to reject the null hypothesis that the null model generated the data.
+For the other models for which we tested goodness-of-fit, however, we \textit{do} have significant evidence from all three replicates to reject the null hypothesis that the null model generated the data.
 All of the goodness-of-fit lineups are provided in the appendix.

### 2017-11-27 ASApaper/paper-child.Rnw
just have to update discussion section and beef up results discussion

@@
 The correspoding results from the significance tests we performed using the lineup protocol are given in Table~\ref{tab:sigtesting}.
-\st{Corresponding to traditional methods, if enough participants pick out the alternative plot to result in a $p-value less than 0.05, we reject the null hypothesis that the true value of the additional parameter, either $\beta_3$ or $\beta_4$, is equal to zero.} %\hh{XXX how do we interpret these  visual results?
+\st{Corresponding to traditional methods, if enough participants pick out the alternative plot to result in a $p$-value less than 0.05, we reject the null hypothesis that the true value of the additional parameter, either $\beta_3$ or $\beta_4$, is equal to zero.} %\hh{XXX how do we interpret these  visual results?
 Include the six lineups in an appendix and refer to them from here.} \st{XXX should be taken care of in next paragraph now.}
@@
 We see that the $p$-values for visual inference here are highly variable.
-This variability is introduced through the null plots generated from M1, since not all simulations look alike.
+
+<<lineup3132, fig.height=4, fig.cap="The lineup which caused a rejection of the null hypothesis that $\\beta_3$ = 0.
+The network simulated from model M3 is found in panel $\\sqrt{16} - 1$, and the remaining panels show networks simulated from model M1.">>=
+dat <- read_csv("data/jttp_pos_hard_2.csv")
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size =1) + 
+        theme_net() + 
+        theme(panel.background = element_rect(color = 'black'),
+              strip.text = element_text(size = 20)) +
+        facet_wrap(~ord)
+@
+
+<<lineup3133, fig.height=4, fig.cap="One of the lineups which failed to reject the null hypothesis that $\\beta_3$ = 0.
+The network simulated from model M3 is found in panel $\\sqrt{25} - 4$, and the remaining panels show networks simulated from model M1.">>=
+dat <- read_csv("data/jttp_pos_hard_3.csv")
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size =1) + 
+        theme_net() + 
+        theme(panel.background = element_rect(color = 'black'),
+              strip.text = element_text(size = 20)) +
+        facet_wrap(~ord)
+@
+
+\st{The lineup for significance testing of $\beta_3$ which resulted in a very small $p$-value and rejection of the null hypothesis is shown in Figure~\ref{fig:lineup3132}.
+Another significance lineup for model M3, which resulted in failure to reject the null hypothesis, is shown in Figure~\ref{fig:lineup3133}.
+When viewing Figure~\ref{fig:lineup3132}, 26 of 31 viewers chose the plot from M3, while only 2 of 27 chose the plot from M3 when viewing Figure~\ref{fig:lineup3133}.
+The most common choice in the latter was panel two, which 16 of 27 viewers chose as the most different due to its large connected component, making it seem more complex than the others.
+In viewing these two lineups, it is evident that there is a large amount of variability.
+It is difficult to see that five of the six come from the same model when they can all look different in their own way.} Thus, the variability in results is introduced through the null plots generated from M1, as not all simulations look alike.
 In addition, the necessarily small number of null plots do not give the viewer as complete of a view of the null model as the usual 19 null plots would.
@@
 However, the panel most participants chose was number four, and the most common reasoning for that choice was that it had the most simple structure.
+\st{Some of the other panels, such as three and six, in Figure~\ref{fig:failgof} have large connected components that are similar in size to the connected component of the data plot shown in panel two.
+Thus, model M5 is sometimes capable of capturing the network structure of the senate collaboration data.}
 
 <<failgof, fig.height=4, fig.cap="The goodness-of-fit lineup that failed to reject the null hypothesis.
-The null model for this lineup is M4.
-Only 7 of 20 viewers of this lineup selected the data plot as the most different from the others.">>=
+The null model for this lineup is M5.
+Only 7 of 20 viewers of this lineup selected the data plot as the most different from the others.
+The most commonly chosen panel was number four, which has a relatively simple structure compared to panels 2, 3, and 6 especially.">>=
 load("data/se112adjmat.RDS")
@@
 The smallest $p$-value for one of the goodness-of-fit lineups was for the third replicate of the null model M5.
-This contradicts our previous finding that the only lineup to fail to reject the null was also when the null model was M5.
+This \st{result contrasts with} our previous finding that the only lineup to fail to reject the null was also when the null model was M5.
 This lineup is shown in Figure~\ref{fig:mostsiggof}.
@@
 We could move the threshold from $\alpha = 0.05$ to $\alpha = 0.01$ or $\alpha = 0.001$, but more work needs to be done to determine which, if any of these, is appropriate.
+
+<<gofm71, fig.height=4, fig.cap="One repitition of a goodness-of-fit lineup testing modle M7.
+The senate data are shown in panel two, and it is evident that none of the other five panels, which show data simulated from model M7, come close to creating the large connected component that is central to the structure of the senate data.">>=
+dat <- read_csv(paste0("data/bigmod_gof_9_1.csv"))
+datplot <- unique(dat$ord[dat$sim == 1001])
+dat <- dat %>% filter(sim != 1001) %>% 
+      bind_rows(wave2) %>% 
+      mutate(ord = ifelse(is.na(ord), datplot, ord))
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size =1) + 
+        theme_net() + 
+        theme(panel.background = element_rect(color = 'black'),
+              strip.text = element_text(size = 20)) +
+        facet_wrap(~ord)
+@
 For the other models for which we tested goodness-of-fit, however, we \textit{do} have significant evidence from all three replicates to reject the null hypothesis that the null model generated the data.
-All of the goodness-of-fit lineups are provided in the appendix.
+\st{For models M3, M5, and M7, these goodness-of-fit tests have rejected the null hypotheses that the senate data come from these models.
+We hypothesized that the model with the most effects, M7, would be the best fit.
+However, as shown in Figure~\ref{fig:gofm71}, the model does not capture the overall structure very well at all.} The rest of the goodness-of-fit lineups as shown to participants are provided in the appendix.
 
@@
 
-<<mostsiggof, fig.height = 4, fig.cap="The lineup resulting in the smallest $p$-value rejecting the null hypothesis. Surprisingly, this another repetition for M4 as the null model.">>=
+<<mostsiggof, fig.height = 4, fig.cap="The lineup resulting in the smallest $p$-value rejecting the null hypothesis. Surprisingly, this another repetition for M5 as the null model.">>=
 dat <- read_csv(paste0("data/jtts_gof_9_3.csv"))
@@
 
-A summary of the results from our experiment is shown in Figure~\ref{fig:predictglmm}.
-On the $x$ axis, we plot the value of the parameter of interest, and on the $y$ axis, the proportion of times the plot of data simulated from the alternative model was picked out for each lineup.
+A summary of the results from our experiment is shown \st{as points }in Figure~\ref{fig:predictglmm}.
+On the $x$ axis, we plot the value of the parameter of interest, and on the $y$ axis, the proportion of times the alternative data plot was picked out for each lineup.
 The results are split into groups based on the value of the parameter and the lineup type.
-We can see clear patterns in nearly all of the groups: as the parameter value approaches 0, fewer participants identified the alternative plot.
+We can see clear patterns \st{in the added parameters $\beta_3, \dots, \beta_6$ }: as the parameter value approaches 0, fewer participants identified the alternative plot.
+Similarly, as $\beta_1, \beta_2$ approach their estimated values $\hat{\beta_1}, \hat{\beta_2}$, fewer people are able to identify the alternative plot.
 
@@
 
-We further explore this relationship between identification of the alternative data in the lineup and the parameter, effect size, and lineup type with a generalized linear mixed model that provides us with an estimate of the power of the visual test.
+We further explore this relationship between identification of the alternative data in the lineup and the parameter \st{of interest}, effect size, and lineup type with a generalized linear mixed model that provides us with an estimate of the power of the visual \st{significance } test.
 The response variable, $Y_{ijkm}$, is binary, indicating whether participant $m$ picked the alternative data plot in lineup type $j$, rep $k$, for effect $i$.
-The continuous covariate $x$ is the centered and scaled size of the effect of interest from which the alternative data were simulated, the values of which are labeled ``easy", ``medium", and ``hard" in Table~\ref{tab:experdetail} according to how difficult we thought the Turk participants would find each lineup.
-In Equation~\ref{eq:glmm}, $i \in \{1, 2, 3, 4, 6\}$ corresponds to the effects $\beta_1, \beta_3, \beta_4, \beta_2, \beta_6, \beta_5$, respectively, $j \in \{-1,1\}$, and $k \in \{1,2,3\}$.
-We include random effects in the model for each lineup, $\delta_{ijk}$ and each participant, $\epsilon_m$, and fit a hierarchical model as follows:
+There is one continuous covariate $x$, which is the centered and scaled size of the effect of interest from which the alternative data were simulated, the values of which are labeled ``easy", ``medium", and ``hard" in Table~\ref{tab:experdetail} according to how difficult we thought the Turk participants would find each lineup.
+In Equation~\ref{eq:glmm}, $i \in \{1, 2, 3, 4, 6\}$ corresponds to the effects $\beta_1,\dots, \beta_6$, respectively, $j \in \{-1,1\}$, and $k \in \{1,2,3\}$.
+We also include random effects in the model: one for each lineup, $\delta_{ijk}$, and one for each participant, $\epsilon_m$, and fit a hierarchical model given in Equation~\ref{eq:glmm}.
 
@@
 Y_{ijkm} &\sim \text{Bernoulli}(\pi_{ijkm}) \\
-\text{logit}(\pi_{ijkm}) & = \alpha_{ij} + \beta_{ij} x + \delta_{ijk} + \epsilon_{m} \label{eq:glmm} \\
+\text{logit}(\pi_{ijkm}) & = \alpha_{ij} + \gamma_{ij} x + \delta_{ijk} + \epsilon_{m} \label{eq:glmm} \\
 \delta_{ijk} & \stackrel{iid}{\sim} N(0, \sigma^2_{\delta}) \\
@@
 \begin{equation}
-E[\text{logit}(\pi_{ij})] = \alpha_{ij} + \beta_{ij}x
+E[\text{logit}(\pi_{ij})] = \alpha_{ij} + \gamma_{ij}x
 \end{equation}
@@
 \begin{equation}
-\pi_{ij} = \frac{\exp\{\alpha_{ij} + \beta_{ij}x\}}{1 + \exp\{\alpha_{ij} + \beta_{ij}x\}}
+\pi_{ij} = \frac{\exp\{\alpha_{ij} + \gamma_{ij}x\}}{1 + \exp\{\alpha_{ij} + \gamma_{ij}x\}}
 \end{equation}
@@
 $\alpha_{1-}$ & \Sexpr{ests[1]} & \Sexpr{se[1]} & \Sexpr{pval[1]} & \Sexpr{odds[1]}\\
-$\beta_{1+}$ & \Sexpr{ests[19]} & \Sexpr{se[19]} & \Sexpr{pval[19]} & \Sexpr{odds[19]} \\
-$\beta_{1-}$  & \Sexpr{ests[13]} & \Sexpr{se[13]} & \Sexpr{pval[13]} & \Sexpr{odds[13]} \\ \hline
+$\gamma_{1+}$ & \Sexpr{ests[19]} & \Sexpr{se[19]} & \Sexpr{pval[19]} & \Sexpr{odds[19]} \\
+$\gamma_{1-}$  & \Sexpr{ests[13]} & \Sexpr{se[13]} & \Sexpr{pval[13]} & \Sexpr{odds[13]} \\ \hline
 $\alpha_{2+}$  & \Sexpr{ests[8]} & \Sexpr{se[8]} & \Sexpr{pval[8]} & \Sexpr{odds[8]} \\ 
 $\alpha_{2-}$ & \Sexpr{ests[2]} & \Sexpr{se[2]} & \Sexpr{pval[2]} & \Sexpr{odds[2]} \\
-$\beta_{2+}$ & \Sexpr{ests[20]} & \Sexpr{se[20]} & \Sexpr{pval[20]} & \Sexpr{odds[20]}  \\
-$\beta_{2-}$   & \Sexpr{ests[14]} & \Sexpr{se[14]} & \Sexpr{pval[14]} & \Sexpr{odds[14]} \\ \hline
+$\gamma_{2+}$ & \Sexpr{ests[20]} & \Sexpr{se[20]} & \Sexpr{pval[20]} & \Sexpr{odds[20]}  \\
+$\gamma_{2-}$   & \Sexpr{ests[14]} & \Sexpr{se[14]} & \Sexpr{pval[14]} & \Sexpr{odds[14]} \\ \hline
 $\alpha_{3+}$ & \Sexpr{ests[9]} & \Sexpr{se[9]} & \Sexpr{pval[9]} & \Sexpr{odds[9]} \\
 $\alpha_{3-}$ & \Sexpr{ests[3]} & \Sexpr{se[3]} & \Sexpr{pval[3]} & \Sexpr{odds[3]}  \\
-$\beta_{3+}$ & \Sexpr{ests[21]} & \Sexpr{se[21]} & \Sexpr{pval[21]} & \Sexpr{odds[21]} \\ 
-$\beta_{3-}$  & \Sexpr{ests[15]} & \Sexpr{se[15]} & \Sexpr{pval[15]} & \Sexpr{odds[15]} \\ \hline\hline
+$\gamma_{3+}$ & \Sexpr{ests[21]} & \Sexpr{se[21]} & \Sexpr{pval[21]} & \Sexpr{odds[21]} \\ 
+$\gamma_{3-}$  & \Sexpr{ests[15]} & \Sexpr{se[15]} & \Sexpr{pval[15]} & \Sexpr{odds[15]} \\ \hline\hline
 $\alpha_{5+}$  & \Sexpr{ests[11]} & \Sexpr{se[11]} & \Sexpr{pval[11]} & \Sexpr{odds[11]} \\
 $\alpha_{5-}$ & \Sexpr{ests[5]} & \Sexpr{se[5]} & \Sexpr{pval[5]} & \Sexpr{odds[5]}  \\
-$\beta_{5+}$ & \Sexpr{ests[23]} & \Sexpr{se[23]} & \Sexpr{pval[23]} & \Sexpr{odds[23]} \\
-$\beta_{5-}$ & \Sexpr{ests[17]} & \Sexpr{se[17]} & \Sexpr{pval[17]} & \Sexpr{odds[17]} \\ \hline
+$\gamma_{5+}$ & \Sexpr{ests[23]} & \Sexpr{se[23]} & \Sexpr{pval[23]} & \Sexpr{odds[23]} \\
+$\gamma_{5-}$ & \Sexpr{ests[17]} & \Sexpr{se[17]} & \Sexpr{pval[17]} & \Sexpr{odds[17]} \\ \hline
 $\alpha_{4+}$  & \Sexpr{ests[10]} & \Sexpr{se[10]} & \Sexpr{pval[10]} & \Sexpr{odds[10]} \\
 $\alpha_{4-}$  & \Sexpr{ests[4]} & \Sexpr{se[4]} & \Sexpr{pval[4]} & \Sexpr{odds[4]} \\
-$\beta_{4+}$ & \Sexpr{ests[22]} & \Sexpr{se[22]} & \Sexpr{pval[22]} & \Sexpr{odds[22]} \\
-$\beta_{4-}$  & \Sexpr{ests[16]} & \Sexpr{se[16]} & \Sexpr{pval[16]} & \Sexpr{odds[16]} \\ \hline
+$\gamma_{4+}$ & \Sexpr{ests[22]} & \Sexpr{se[22]} & \Sexpr{pval[22]} & \Sexpr{odds[22]} \\
+$\gamma_{4-}$  & \Sexpr{ests[16]} & \Sexpr{se[16]} & \Sexpr{pval[16]} & \Sexpr{odds[16]} \\ \hline
 $\alpha_{6+}$  & \Sexpr{ests[12]} & \Sexpr{se[12]} & \Sexpr{pval[12]} & \Sexpr{odds[12]} \\
 $\alpha_{6-}$  & \Sexpr{ests[6]} & \Sexpr{se[6]} & \Sexpr{pval[6]} & \Sexpr{odds[6]} \\
-$\beta_{6+}$ & \Sexpr{ests[24]} & \Sexpr{se[24]} & \Sexpr{pval[24]} & \Sexpr{odds[24]} \\ 
-$\beta_{6-}$ & \Sexpr{ests[18]} & \Sexpr{se[18]} & \Sexpr{pval[18]} & \Sexpr{odds[18]} \\
+$\gamma_{6+}$ & \Sexpr{ests[24]} & \Sexpr{se[24]} & \Sexpr{pval[24]} & \Sexpr{odds[24]} \\ 
+$\gamma_{6-}$ & \Sexpr{ests[18]} & \Sexpr{se[18]} & \Sexpr{pval[18]} & \Sexpr{odds[18]} \\
 \hline\hline

### 2017-11-27 ASApaper/paper-child.Rnw
reorg res table

@@
                  color = 'black', arrowgap = .015, arrowsize = .3, size =1) + 
-        theme_net() + 
-        theme(panel.background = element_rect(color = 'black'),
-              strip.text = element_text(size = 20)) +
+        ThemeNet + 
+        theme(panel.background = element_rect(color = 'black')) +
         facet_wrap(~ord)
@@
 Parameter & \multicolumn{1}{r}{Estimate} & \multicolumn{1}{r}{Std Error} & $p$-value & Odds Multiplier \\ \hline\hline
-$\alpha_{1+}$ & \Sexpr{ests[7]} & \Sexpr{se[7]} & \Sexpr{pval[7]} & \Sexpr{odds[7]} \\
-$\alpha_{1-}$ & \Sexpr{ests[1]} & \Sexpr{se[1]} & \Sexpr{pval[1]} & \Sexpr{odds[1]}\\
-$\gamma_{1+}$ & \Sexpr{ests[19]} & \Sexpr{se[19]} & \Sexpr{pval[19]} & \Sexpr{odds[19]} \\
-$\gamma_{1-}$  & \Sexpr{ests[13]} & \Sexpr{se[13]} & \Sexpr{pval[13]} & \Sexpr{odds[13]} \\ \hline
-$\alpha_{2+}$  & \Sexpr{ests[8]} & \Sexpr{se[8]} & \Sexpr{pval[8]} & \Sexpr{odds[8]} \\ 
-$\alpha_{2-}$ & \Sexpr{ests[2]} & \Sexpr{se[2]} & \Sexpr{pval[2]} & \Sexpr{odds[2]} \\
-$\gamma_{2+}$ & \Sexpr{ests[20]} & \Sexpr{se[20]} & \Sexpr{pval[20]} & \Sexpr{odds[20]}  \\
-$\gamma_{2-}$   & \Sexpr{ests[14]} & \Sexpr{se[14]} & \Sexpr{pval[14]} & \Sexpr{odds[14]} \\ \hline
-$\alpha_{3+}$ & \Sexpr{ests[9]} & \Sexpr{se[9]} & \Sexpr{pval[9]} & \Sexpr{odds[9]} \\
-$\alpha_{3-}$ & \Sexpr{ests[3]} & \Sexpr{se[3]} & \Sexpr{pval[3]} & \Sexpr{odds[3]}  \\
-$\gamma_{3+}$ & \Sexpr{ests[21]} & \Sexpr{se[21]} & \Sexpr{pval[21]} & \Sexpr{odds[21]} \\ 
-$\gamma_{3-}$  & \Sexpr{ests[15]} & \Sexpr{se[15]} & \Sexpr{pval[15]} & \Sexpr{odds[15]} \\ \hline\hline
-$\alpha_{5+}$  & \Sexpr{ests[11]} & \Sexpr{se[11]} & \Sexpr{pval[11]} & \Sexpr{odds[11]} \\
-$\alpha_{5-}$ & \Sexpr{ests[5]} & \Sexpr{se[5]} & \Sexpr{pval[5]} & \Sexpr{odds[5]}  \\
-$\gamma_{5+}$ & \Sexpr{ests[23]} & \Sexpr{se[23]} & \Sexpr{pval[23]} & \Sexpr{odds[23]} \\
-$\gamma_{5-}$ & \Sexpr{ests[17]} & \Sexpr{se[17]} & \Sexpr{pval[17]} & \Sexpr{odds[17]} \\ \hline
-$\alpha_{4+}$  & \Sexpr{ests[10]} & \Sexpr{se[10]} & \Sexpr{pval[10]} & \Sexpr{odds[10]} \\
-$\alpha_{4-}$  & \Sexpr{ests[4]} & \Sexpr{se[4]} & \Sexpr{pval[4]} & \Sexpr{odds[4]} \\
-$\gamma_{4+}$ & \Sexpr{ests[22]} & \Sexpr{se[22]} & \Sexpr{pval[22]} & \Sexpr{odds[22]} \\
-$\gamma_{4-}$  & \Sexpr{ests[16]} & \Sexpr{se[16]} & \Sexpr{pval[16]} & \Sexpr{odds[16]} \\ \hline
-$\alpha_{6+}$  & \Sexpr{ests[12]} & \Sexpr{se[12]} & \Sexpr{pval[12]} & \Sexpr{odds[12]} \\
-$\alpha_{6-}$  & \Sexpr{ests[6]} & \Sexpr{se[6]} & \Sexpr{pval[6]} & \Sexpr{odds[6]} \\
-$\gamma_{6+}$ & \Sexpr{ests[24]} & \Sexpr{se[24]} & \Sexpr{pval[24]} & \Sexpr{odds[24]} \\ 
-$\gamma_{6-}$ & \Sexpr{ests[18]} & \Sexpr{se[18]} & \Sexpr{pval[18]} & \Sexpr{odds[18]} \\
-\hline\hline
+$\alpha_{1+}$ & \Sexpr{ests[7]} & \Sexpr{se[7]} & \Sexpr{pval[7]}$^{\ddagger}$ & \Sexpr{odds[7]} \\
+$\alpha_{1-}$ & \Sexpr{ests[1]} & \Sexpr{se[1]} & \Sexpr{pval[1]}$^{\dagger}$ & \Sexpr{odds[1]}\\
+$\gamma_{1+}$ & \Sexpr{ests[19]} & \Sexpr{se[19]} & \Sexpr{pval[19]}$^{\ddagger}$ & \Sexpr{odds[19]} \\
+$\gamma_{1-}$  & \Sexpr{ests[13]} & \Sexpr{se[13]} & \Sexpr{pval[13]}$^{\dagger}$ & \Sexpr{odds[13]} \\ \hline
+$\alpha_{2+}$  & \Sexpr{ests[10]} & \Sexpr{se[10]} & \Sexpr{pval[10]} & \Sexpr{odds[10]} \\
+$\alpha_{2-}$  & \Sexpr{ests[4]} & \Sexpr{se[4]} & \Sexpr{pval[4]}$^{\ddagger}$ & \Sexpr{odds[4]} \\
+$\gamma_{2+}$ & \Sexpr{ests[22]} & \Sexpr{se[22]} & \Sexpr{pval[22]}$^{*}$ & \Sexpr{odds[22]} \\
+$\gamma_{2-}$  & \Sexpr{ests[16]} & \Sexpr{se[16]} & \Sexpr{pval[16]}$^{\ddagger}$ & \Sexpr{odds[16]} \\ \hline
+$\alpha_{3+}$  & \Sexpr{ests[8]} & \Sexpr{se[8]} & \Sexpr{pval[8]}$^{\dagger}$ & \Sexpr{odds[8]} \\ 
+$\alpha_{3-}$ & \Sexpr{ests[2]} & \Sexpr{se[2]} & \Sexpr{pval[2]}$^{\dagger}$ & \Sexpr{odds[2]} \\
+$\gamma_{3+}$ & \Sexpr{ests[20]} & \Sexpr{se[20]} & \Sexpr{pval[20]}$^{\dagger}$ & \Sexpr{odds[20]}  \\
+$\gamma_{3-}$   & \Sexpr{ests[14]} & \Sexpr{se[14]} & \Sexpr{pval[14]}$^{*}$ & \Sexpr{odds[14]} \\ \hline
+$\alpha_{4+}$ & \Sexpr{ests[9]} & \Sexpr{se[9]} & \Sexpr{pval[9]}$^{**}$ & \Sexpr{odds[9]} \\
+$\alpha_{4-}$ & \Sexpr{ests[3]} & \Sexpr{se[3]} & \Sexpr{pval[3]}$^{**}$ & \Sexpr{odds[3]}  \\
+$\gamma_{4+}$ & \Sexpr{ests[21]} & \Sexpr{se[21]} & \Sexpr{pval[21]}$^{**}$ & \Sexpr{odds[21]} \\ 
+$\gamma_{4-}$  & \Sexpr{ests[15]} & \Sexpr{se[15]} & \Sexpr{pval[15]} & \Sexpr{odds[15]} \\ \hline
+$\alpha_{5+}$  & \Sexpr{ests[12]} & \Sexpr{se[12]} & \Sexpr{pval[12]}$^{*}$ & \Sexpr{odds[12]} \\
+$\alpha_{5-}$  & \Sexpr{ests[6]} & \Sexpr{se[6]} & \Sexpr{pval[6]}$^{\ddagger}$ & \Sexpr{odds[6]} \\
+$\gamma_{5+}$ & \Sexpr{ests[24]} & \Sexpr{se[24]} & \Sexpr{pval[24]}$^{*}$ & \Sexpr{odds[24]} \\ 
+$\gamma_{5-}$ & \Sexpr{ests[18]} & \Sexpr{se[18]} & \Sexpr{pval[18]}$^{\ddagger}$ & \Sexpr{odds[18]} \\ \hline
+$\alpha_{6+}$  & \Sexpr{ests[11]} & \Sexpr{se[11]} & \Sexpr{pval[11]} & \Sexpr{odds[11]} \\
+$\alpha_{6-}$ & \Sexpr{ests[5]} & \Sexpr{se[5]} & \Sexpr{pval[5]}$^{\ddagger}$ & \Sexpr{odds[5]}  \\
+$\gamma_{6+}$ & \Sexpr{ests[23]} & \Sexpr{se[23]} & \Sexpr{pval[23]} & \Sexpr{odds[23]} \\
+$\gamma_{6-}$ & \Sexpr{ests[17]} & \Sexpr{se[17]} & \Sexpr{pval[17]}$^{\ddagger}$ & \Sexpr{odds[17]} \\ \hline
 $\sigma^2_{\delta}$ & \Sexpr{round(sigma_delta, 4)} & \multicolumn{1}{r}{--} & -- & --\\
@@
 \end{tabular}
-\caption{\label{tab:glmmests}Summary of the results from fitting the model given in Equation~\ref{eq:glmm}}
+\caption{\label{tab:glmmests}Summary of the results from fitting the model given in Equation~\ref{eq:glmm}. Significance levels: * - $< 0.10$; ** - $<0.05$; $\dagger$ - $<0.01$; $\ddagger$ - $<0.001$}
 \end{table}
@@
 
-<<recipzoom, fig.height=2.5, fig.cap='The top middle panel of Figure~\\ref{fig:predictglmm} expanded to show greater detail.
-The square root of the estimate is shown on the $x$-axis.
+In Figure~\ref{fig:predictglmm}, we see a clear trend in all parameters except $\beta_1$ and $\beta_2$ that as the parameter value approaches zero from either side, the probability of picking the data plot in a lineup of size six descreases.
+For $\beta_3$ and $\beta_5$, the slope of the fitted line is \textit{much} steeper for positive values of the parameter than for negative values, meaning that our participants perceived differences more often for postitive parameter values than for negative parameter values.
+This finding is similar to that of \citet{corrviz}, who found that people detect positive correlations sooner and better than negative correlations.
+
+<<morepredicts>>=
+alphas <- (coef(mod)[[1]][,-1] %>% unique())[1:12] 
+betas <- (coef(mod)[[1]][,-1] %>% unique())[13:24] 
+names(betas) <- str_replace(str_replace(names(betas), "lps_param", ""), ":centersize", "")
+names(betas) <- str_replace(names(betas), ".-1", "neg")
+names(betas) <- str_replace(names(betas), ".1", "pos")
+names(alphas) <- str_replace(names(alphas), "lps_param", "")
+names(alphas) <- str_replace(names(alphas), ".-1", "neg")
+names(alphas) <- str_replace(names(alphas), ".1", "pos")
+source("code/newpredictions.R")
+newdata <- newdata3
+newdata$param <- newdata$test_param
+newdata$param <- as.factor(newdata$test_param)
+#levels(newdata$param) <-  c(1,3,4,2,6,5) 
+#newdata$param <- as.integer(as.character(newdata$param))
+plotdata <- turk22 %>% mutate(type2 = ifelse(sign == 0, sign(initialEst), sign),
+                       type2 = ifelse(type == "one", type2, -1)) %>% 
+  group_by(pic_id, size, test_param, sign, type2,alt_model) %>% summarize(
+    datapick = mean(datapick)
+)
+plotdata$param <- plotdata$test_param
+plotdata$param <- as.factor(plotdata$param)
+#levels(plotdata$param) <- c(1,3,4,2,6,5) 
+#plotdata$param <- as.integer(as.character(plotdata$param))
+labdat <- data_frame(test_param = "jttp", param = 3, type2 = -1, x = -5, y = .9, label = "jttp_neg_hard_2")
+@
+
+<<recipzoom, fig.height=3, fig.cap='The top middle panel of Figure~\\ref{fig:predictglmm} expanded to show greater detail.
+The square root of the parameter value is shown on the $x$-axis.
 For this parameter, as its value approaches zero, the probability of identifying the alternate data model decreases, then increases, which is noticeably different from the pattern exhibited by the others.
@@
 ggplot() + 
-  geom_line(data = newdata %>% filter(test_param == "recip"), aes(x = sqrt(size), y = predictfinal, color = as.factor(type2))) + 
-  geom_point(data = plotdata %>% filter(test_param == "recip"), aes(x = sqrt(size), y = datapick, color = as.factor(type2))) + 
-  scale_color_brewer(palette = "Dark2", name = "Lineup Type") + 
-  geom_hline(yintercept = 1/6, linetype='dashed') + 
-  facet_wrap(~type2, scales = 'free_x') + 
-  labs(x = "Square root of parameter value (recip)", y = "Expected probability of detection\nin a lineup of size 6") + 
+  geom_line(data = newdata %>% filter(test_param == "recip", category == "inside"), aes(x = sqrt(size), y = predictfinal, color = as.factor(type2)), size = 1.25) + 
+  geom_point(data = plotdata %>% filter(test_param=="recip"), aes(x = sqrt(size), y = datapick, color = as.factor(type2)), size = 2) + 
+  geom_line(data = newdata %>% filter(test_param == "recip", category == "outside"), aes(x = sqrt(size), y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
+  geom_line(data = newdata %>% filter(test_param == "recip", category == "outside2"), aes(x = sqrt(size), y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
+  #geom_text(data = plotdata, aes(x = size, y = datapick, color = as.factor(type2), label = pic_id)) +
+  #geom_text(data = labdat, aes(x = x, y=y, label = label, color = as.factor(type2)),
+  #          show.legend = F) + 
+  geom_vline(xintercept = sqrt(4.893)) + 
+  geom_hline(yintercept = 1/6, linetype='dotted') +
+  scale_color_brewer(palette = "Dark2", name = "Lineup Type", labels = c("-", "+")) +  
+  facet_grid(~param) + 
+  labs(x = "Square root of parameter value", y = "Expected probability of detection\nin a lineup of size 6") + 
   ThemeNoNet + 
-  theme(legend.position = 'bottom')
-@
-
-In Figure~\ref{fig:predictglmm}, we see a clear trend in all parameters except $\beta_1$ and $\beta_2$ that as the parameter value approaches zero from either side, the probability of picking the data plot in a lineup of size six descreases.
-For $\beta_3$ and $\beta_5$, the slope of the fitted line is much steeper for positive values of the parameter than form negative values, meaning that our participants perceived differences more often for postitive parameter values than for negative parameter values.
-This finding is similar to that of \citet{corrviz}, who found that people detect positive correlations sooner and better than negative correlations.
+  theme(legend.position = 'bottom', strip.text = element_text(size = 10), text = element_text(size = rel(3)), legend.text = element_text(size = rel(3)), legend.key.size = unit(.25, "inches"))
+@
+
+<<beta5zoom, fig.height=3, fig.cap="The bottom middle panel of Figure~\\ref{fig:predictglmm} expanded to show greater detail.
+The parameter value is shown on the $x$-axis.
+This parameter most closely follows our hypothesis shown in Figure~\\ref{fig:hypothesis}.
+However, the result is not symmetric.
+According to the model, people will detect the effect at lower values and with greater frequency as the value increases when it is positive instead of negative.">>=
+ggplot() + 
+  geom_line(data = newdata %>% filter(test_param == "simttb", category == "inside"), aes(x = size, y = predictfinal, color = as.factor(type2)), size = 1.25) + 
+  geom_point(data = plotdata %>% filter(test_param=="simttb"), aes(x = size, y = datapick, color = as.factor(type2)), size = 2) + 
+  geom_line(data = newdata %>% filter(test_param == "simttb", category == "outside"), aes(x = size, y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
+  geom_line(data = newdata %>% filter(test_param == "simttb", category == "outside2"), aes(x = size, y = predictfinal, group = as.factor(type2)), color = "gray60",  alpha = .7) +
+  #geom_text(data = plotdata, aes(x = size, y = datapick, color = as.factor(type2), label = pic_id)) +
+  #geom_text(data = labdat, aes(x = x, y=y, label = label, color = as.factor(type2)),
+  #          show.legend = F) + 
+  geom_vline(xintercept = 10.090829) + 
+  geom_hline(yintercept = 1/6, linetype='dotted') +
+  scale_color_brewer(palette = "Dark2", name = "Lineup Type", labels = c("-", "+")) +
+  facet_grid(~param) + 
+  labs(x = "Parameter value", y = "Expected probability of detection in a lineup of size 6") + 
+  ThemeNoNet + 
+  theme(legend.position = 'bottom', strip.text = element_text(size = 10), text = element_text(size = rel(3)), legend.text = element_text(size = rel(3)), legend.key.size = unit(.25, "inches"))
+@
+
+<<beta4zoom, fig.height=3, fig.cap='The bottom left panel of Figure~\\ref{fig:predictglmm} expanded to show greater detail.
+The parameter value is shown on the $x$-axis.
+The ``reverse" lineup has a much flatter slope than the ``regular" lineup, which means the participants had a harder time detecting a more simple M1 structure among many more complex M4 structures.
+Reversing the lineup scenario was not symmetric as we hypothesized.'>>=
+ggplot() + 
+  geom_line(data = newdata %>% filter(test_param == "jtts", category == "inside"), aes(x = size, y = predictfinal, color = as.factor(type2)), size = 1.25) + 
+  geom_point(data = plotdata %>% filter(test_param=="jtts"), aes(x = size, y = datapick, color = as.factor(type2)), size = 2) + 
+  geom_line(data = newdata %>% filter(test_param == "jtts", category == "outside"), aes(x = size, y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
+  geom_line(data = newdata %>% filter(test_param == "jtts", category == "outside2"), aes(x = size, y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
+  #geom_text(data = plotdata, aes(x = size, y = datapick, color = as.factor(type2), label = pic_id)) +
+  #geom_text(data = labdat, aes(x = x, y=y, label = label, color = as.factor(type2)),
+  #          show.legend = F) + 
+  geom_vline(xintercept = 3.340302) + 
+  geom_hline(yintercept = 1/6, linetype='dotted') +
+  scale_color_brewer(palette = "Dark2", name = "Lineup Condition", labels = c("reverse", "regular")) +
+  facet_grid(~param) + 
+  labs(x = "Parameter value", y = "Expected probability of detection in a lineup of size 6") + 
+  ThemeNoNet + 
+  theme(legend.position = 'bottom', strip.text = element_text(size = 10), text = element_text(size = rel(3)), legend.text = element_text(size = rel(3)), legend.key.size = unit(.25, "inches"))
+@
+
+\st{We expand portions of Figure~\ref{fig:predictglmm} in Figures~\ref{fig:recipzoom}-\ref{fig:beta4zoom}.
+These figures show the same prediction regions as in Figure~\ref{fig:predictglmm}, plus some additional predictions outside of the data range shown in gray.
+Again, the points represent the results from the experiment.
+In all three of these figures, the lack of symmetry is apparent.
+In the reverse lineup scenario shown in Figure~\ref{fig:beta4zoom}, the probability of prediction is consistently far less than the probability of prediction in the regular lineup scenario.
+This demonstrates that the visual signal of one plot from M4 among five plots from M1 is much stronger than that of one plot from M1 among five plots from M4.
+We posit that the latter is a more difficult task because it involves noticing a \textit{lack of structure} as opposed to the presence of more structure.
+We can see a similar effect in Figure~\ref{fig:beta5zoom}.
+At a value of $\beta_5 = 20$, the model predicts a probability of about 0.60 that a new viewer of a new lineup will identify the alternative data plot.
+At a value of $\beta_5 = -20$, however, the model predicts this same probability to be about 0.40.
+This again demonstrates that the presence of structure is detected sooner and more frequently than the absence of strucutre.
+}
 
@@
                  color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
-        theme_net() + 
+        ThemeNet + 
         theme(panel.background = element_rect(color = 'black')) +
@@
 For identical values of the parameter, viewers had a harder time identifying the different plot when they were selected the most ``simple" structure, detecting M1 in five plots from the more complicated model, than they did identifying the most ``complex" structure, the plot from the more complicated model, from the five plots from M1.
+\st{This result is also similar to that of \citet{corrviz} because it emphasizes the difficulty of picking out the absence of an effect relative to picking out the presence of an effect.}
 
@@
 By using visual inference methods, we have developed new ways to perform significance and goodness-of-fit testing for a complicated and intractable set of statistical models for social network data.
-These new methods can be used to supplement traditional methods and check our assumptions about network models.
-The traditional methods only look at one piece or one measure of a network model, but our methods look at the models holistically for a broader sense of what it means for a parameter to be significant or a model to be a good fit.
-By looking at an entire network simulated from a SAOM side-by-side with other instances of networks simulated from another model, instead of one-dimensional derived features, we develop an idea of the model in terms of the data itself, instead of in terms of statistical summaries of the data.
-
-We have found the visual power of some effects in the object function of a SAOM for this particular senate data example,  and we have shown that, for the same effects, there is a lot of variability in results from significance and goodness of fit tests.
-Because the visual tests we performed show a great deal of variability, we can see that the decisions with respect to the significance of a parameter or the goodness of fit of a model to data are not as cut-and-dried as the traditional methods would have us believe.
+\st{We have also developed a way to determine the power of these new visual tests.} Our methods can be used to supplement traditional methods and check our assumptions about network models.
+The traditional methods only look at one piece or derived measure of a network model, whereas our methods look at the models holistically for a broader sense of what it means for a parameter to be significant or a model to be a good fit.
+By looking at an entire network simulated from a SAOM side-by-side with other instances of networks simulated from another model, instead of singular %one-dimensional derived
+features, we develop an idea of the model in terms of the \textit{data} itself, instead of in terms of statistical summaries of the data.
+\st{These methods place the model in the data space, instead of summarizing or compressing the data to place it in the model space.}
+
+Furthermore, we have found the visual power of some effects in the object function of a SAOM for this particular senate data example,  and we have shown that, for the same effects, there is a lot of variability in results from significance and goodness of fit tests.
+Because the visual tests we performed show a great deal of variability, we can see that the decisions with respect to the significance of a parameter or the goodness of fit of a model to data are not as cut-and-dried as the \st{more} traditional methods would have us believe.
 
@@
 In visual inference, the null plots are supposed to play the role of good representatives of the null model.
-Here, the number of null plots is reduced to five, which increases the variability seen from a single lineup dramatically, and can unfortunately lead to different conclusions for the same lineup scenario.
-Furthermore, these results do not generalize yet.
+Here, the number of null plots is reduced to five, which increases the variability seen in a single lineup dramatically, and can unfortunately lead to very different conclusions for the same lineup scenario.
+Furthermore, these results do not generalize to all SAOMs or to some subset of SAOMs.
 The lineups shown are made for only one set of data, and it is not clear whether the power results transfer, nor is it clear to what degree if they do transfer, to other situations with different number of actors, different edge densities, or different layout algorithm of the node-link diagram.
+\st{We can make some generalizations about what participants are picking up on in the lineups based on their feedback and previous research, but we cannot apply our hierarchical model directly to lineups constructed for new data or new models or parameters.}
 
@@
 We accept the limitations of this type of network data visualization, in that even in small instances, the cognitive load of looking at a lineup is very high for the average observer.
-We would therefore like to explore larger datasets, different layout algorithms, and different ways of visualizing network data, such as changing the layout algorithm or using adjacency matrix visualizations, through visual inference to see if similar patterns emerge.
+We would therefore like to explore larger datasets, different layout algorithms, and different ways of visualizing network data, such adjacency matrix visualizations, using visual inference to see if similar patterns emerge.
 

### 2019-01-25 ASApaper/paper-child.Rnw
some rearranging, re-write abstract.

@@
 
-\section{Introduction}
+\section{Background} \label{sec:bg}
 % \st{
@@
 % }
-Three of the most important pieces of statistical modeling are significance testing of model parameters, \st{model goodness-of-fit tests, and determining} power of a \st{hypothesis} test.
-In the first, the data are usually assumed to come from a simple model under the null hypothesis, and additional parameters are tested \st{to determine} whether they significantly contribute to explaining the variability in the data.
+When choosing and fitting statistical models, there are often three key evaluations of interest: significance testing of model parameters, model goodness-of-fit tests, and determining the power of those tests.
+With significance testing, the data are typically assumed to come from a simple model under the null hypothesis, and additional parameters are tested to determine whether they significantly contribute to explaining the variability in the data.
 In the second, the model of interest is examined to determine how well it fits the data.
@@
 
-Some particularly complicated sets of models are those designed to model network change.
+Some particularly complicated families of models are those designed to model network change.
 A \textit{network} is any set of things, such as people, computers, or neurons, that are connected in some way, through social relations, internet connection, or electrical impulses in the brain.
@@
 Dependencies inherent to the data make network objects particularly difficult to model. %Even more challenging is the situation 
-\st{THe difficulty of modeling increases} when we go beyond single instances of a network and consider the dynamics of network change between observed instances.
+The difficulty of modeling increases when we examine more than a single instance of a network and consider the dynamics of network change between observed instances.
 This type of modelling for dynamic networks is often performed on social network data, such as friendship networks among students or the spread of HIV in drug users sharing needles.
@@
 %These 
-\st{It is known that most network} models lack the asymptotics required to perform many well-known goodness-of-fit tests, and the maximum likelihood estimation of parameters is so difficult that it can make significance testing difficult as well \citep{goldenberg09}.
+It is known that most network models lack the asymptotics required to perform many well-known goodness-of-fit tests, and the maximum likelihood estimation of parameters is so difficult that it can make significance testing difficult as well, see for example \citet{holland81}.
 %\hh{is the previous sentence a quote? I'm not sure that I follow.} %\st{it's a paraphrase... shall I make a quote? Also I definitely botched a word. Hopefully makes more sense now? }
 
-\st{In order to circumvent these difficulties,} we propose new methods for significance \st{testing of parameters}, goodness-of-fit testing, and power calculations %of these 
-for a set of social network models: stochastic actor-oriented models for dynamic network data \citep{saompaper}.
+In order to circumvent some of these difficulties, we propose new methods for significance testing of parameters, goodness-of-fit testing, and power calculations %of these 
+for a set of social network models: stochastic actor-oriented models for dynamic network data, first introduced by \citet{saompaper}.
 Specifically, we are using \textit{visual inference} in place of traditional statistical methods for social network models, such as Wald or $t$-tests for significance of parameters and in- and outdegree distribution metrics for determining goodness-of-fit.
-Visual inference, introduced by \citet{Bujaetal}, allows us to look at the \textit{entire} dataset simulated from a network model as opposed to a (set of) usually one-dimensional metric(s) derived from the network such as outdegree, or a $p$-value for a single parameter in the model.
-\st{By using visual inference to supplement traditional statistical tests, we gain a new understanding of parameters in SAOMs and fit of SAOMs to data.}%In addition, we use visual inference to better understand some parameters in a social network model, by looking at how different values for parameters affect the networks simulated from these models.
+Visual inference, introduced by \citet{Bujaetal}, will allow us to look at the \textit{entire} dataset simulated from a network model as opposed to a (set of) usually one-dimensional metric(s) derived from the network such as outdegree distribution, or a $p$-value for a parameter in the model.
+By using visual inference to supplement traditional statistical tests, we gain a new understanding of parameters in SAOMs and fit of SAOMs to data.%In addition, we use visual inference to better understand some parameters in a social network model, by looking at how different values for parameters affect the networks simulated from these models.
 
@@
 
-\section{Visual Inference}\label{vi}
-
-Data visualizations are an important component of data analysis, providing a mechanism for discovering patterns in data.
-Pioneering research by \citet{gelman:2004}, \citet{Bujaetal} and \citet{majumder:2011} provide methods to quantify the significance of discoveries made from visualizations.
-\citet{Bujaetal} introduced two protocols, the Rorschach and the lineup protocol, which bridge the gulf between traditional statistical inference and exploratory data analysis.
-%The Rorschach protocol consists of a set of $m$ (usually, $m=20$) plots (called the {\it null plots}) rendered from data that is consistent with a given null model.
-The Rorschach protocol helps to understand the extent of randomness in the null model.
-Here, we use the lineup protocol.
-\st{Under this protocol, we begin with a data set of interest to us, such as a network, and a visualization of this data, such as a node-link diagram.
-We would like to know the model that generated this data.
-There are two hypothesis considered: a null hypothesis which states the model that is assumed to have generated the data, and an alternative hypothesis that the data were not generated under this model.} Then, the plot of the data of interest is placed randomly among a set of $m-1$ null plots, which are visualizations of data simulated from the null model(where $m =20$, usually).
-Human observers are then asked to examine the lineup and to identify the \st{plots that looks} most different \st{from the others}.
-If an observer identifies the data plot, this is quantifiable evidence against the null hypothesis.
-Since an observer has a chance of 1 in $m$ to pick the data plot from the lineup by simply guessing, i.e.
-in a situation where the data plot is virtually indistinguishable from the null plots, the evidence grows in strength with the number of independent observers identifying the data plot.
-
-The lineup protocol places a plot firmly in the framework of hypothesis tests: a plot of the data is considered to be the test statistic, which is compared against the sampling distribution under the null hypothesis represented by the null plots.
-Obviously, the null generating mechanism, i.e.\ the method of obtaining the data for null plots, is crucial for the lineup protocol, as the null hypothesis directly affects the choice of null generating method.
-Null generating methods are typically based on (a) simulation, if the null hypothesis allows us to directly specify a parametric model, (b) sampling, as for example in the case of large data sets, or (c) permutation of the original data \citep[see e.g.\ ][]{Good05}, which allows for non-parametric testing that preserves marginal distributions  while ensuring independence in higher dimensions.
-The model of interest here allows us to simulate directly from a parametric model for dynamic social network data.
-%In the experimental data that we analyzed the null generating methods used were permutation methods and direct simulation from a null model.
-
-The lineup protocol was formally tested in a head-to-head comparison with the equivalent conventional test in \citet{majumder:2011}.
-The experiment utilized human subjects from Amazon's Mechanical Turk \citep{turk} and used simulation to control conditions.
-The results suggest that  visual inference is comparable to conventional tests in a controlled conventional setting.
-This provides support for its appropriateness for testing in real exploratory situations where no conventional test exists or is difficult.
-%Interestingly, the power of a visual test increases with the number of observers engaged to evaluate lineups, and the pattern in results suggests that the power will provide results consistent with practical significance \citep{kirk:1996}.
-
-\section{Stochastic Actor-Oriented Models}\label{sec:saom}
+\subsection{Stochastic Actor-Oriented Models}\label{subsec:saom}
 
@@
 
-Broadly, a SAOM takes network structure and node covariate information into account in two ways and models the network changes \st{one-at-a-time} as a continuous time Markov chain (CTMC).
+Broadly, a SAOM takes network structure and node covariate information into account in two ways and models the network changes one-at-a-time as a continuous time Markov chain (CTMC).
 First, the rate of change between states is dictated by a rate function that describes \textit{how often} changes in the network occur, and secondly, the objective function describes \textit{what} those state changes are.
@@
   \end{equation}
-Edges are treated as \textit{directed}, i.e. in general $x_{ij} \neq x_{ji}$, and self-referencing edges or loops are not allowed, i.e.  $x_{ii} = 0$ for all $i$. 
+Edges are treated as \textit{directed}, i.e. in general $x_{ij} \neq x_{ji}$, and self-referencing edges or loops are not allowed, i.e.  $x_{ii} = 0$ for all $i=1, 2, \dots, n$. 
 The network is observed $M \geq 2$ times at time points  $t_1 < t_2 < ...
@@
 
-\subsection{Rate Function}\label{rf}
+\subsubsection{Rate Function}\label{subsubsec:rf}
 
@@
 Using this notation, the \textit{waiting time} to the next chance for actor $i$ to make a change is exponentially distributed with expected value $\alpha_m^{-1}$.
-Since the rate is the same for all actors, the waiting time for \textit{any} actor to get the \st{oppurtunity} to change \st{its set of ties} is exponentially distributed with expected value $(n\alpha_m)^{-1}$.
-
-\subsection{Objective Function}\label{of}
+Since the rate is the same for all actors, the waiting time for \textit{any} actor to get the oppurtunity to change its set of ties is exponentially distributed with expected value $(n\alpha_m)^{-1}$.
+
+\subsubsection{Objective Function}\label{subsubsec:of}
 
@@
 %The objective function, $f_i$, is the driving force of a SAOM. 
-Actor $i$ is aiming to maximize \st{its corresponding} objective function $f_i$ given the current state of the network, $x$ and the node-level covariates, $\mathbf{Z}$, given as:% which has the form
+Actor $i$ is aiming to maximize its corresponding objective function $f_i$ given the current state of the network, $x$ and the node-level covariates, $\mathbf{Z}$, given as:% which has the form
 \begin{equation}\label{eq:objective}
@@
 \end{equation}
-where $\boldsymbol{\beta} = (\beta_1, \dots, \beta_K)$ are additional model parameters, each associated with some network statistics, $s_{ik}(x, \mathbf{Z})$, $s_{ik}(x, \mathbf{Z})$, calculated with respect to actor $i$ \st{at the current network state $x$}.
+where $\boldsymbol{\beta} = (\beta_1, \dots, \beta_K)$ are additional model parameters, each associated with some network statistics, $s_{ik}(x, \mathbf{Z})$, $s_{ik}(x, \mathbf{Z})$, calculated with respect to actor $i$ at the current network state $x$.
 Network statistics range from the simple outdegree, $s_i(x) = \sum_{i\neq j} x_{ij}$, to the more complicated {\it transitive triplets jumping to different covariate}, $s_i(x, \mathbf{Z}) = \sum_{i \neq j \neq h} x_{ij}x_{ih}x_{hj} \cdot \mathbb{I}(z_i = z_h \neq z_j)$.
@@
 The acting node can also choose to \textit{not} change at all.
-This \st{is most likely to occur} when the numerator, as calculated for the current state of the network, is larger than for any changes $x(i \leadsto j)$ that could be made.
+This is most likely to occur when the numerator, as calculated for the current state of the network, is larger than for any changes $x(i \leadsto j)$ that could be made.
 
@@
 Similarly, we denote the reciprocity parameter by $\beta_2$ and the associated statistic as $s_{i2}(x) = \sum_{j} x_{ij}x_{ji}$.
-We \st{will} refer to the \st{very simple} model with only these two parameters in the objective function as M1.
-
-\subsection{Example Data}\label{sec:senatedata}
-
-The data we use are collaboration networks in the United States Senate during the $111^{th}$ through $114^{th}$ Congresses, overlapping with Barack Obama's presidency.
-These senates began on January 6, 2009, the start date of the $111^{th}$, and ended on January 3, 2017, the last date of the $114^{th}$\footnote{Details of how this data can be downloaded are provided by François Briatte at \url{https://github.com/briatte/congress}.}.
+We will refer to the very simple model with only these two parameters in the objective function as M1.
+
+\subsection{Visual Inference}\label{subsec:vi}
+
+Data visualizations are an important component of data analysis, providing a mechanism for discovering patterns in data.
+Pioneering research by \citet{gelman:2004}, \citet{Bujaetal} and \citet{majumder:2011} provide methods to quantify the significance of discoveries made from visualizations.
+\citet{Bujaetal} introduced two protocols, the Rorschach and the lineup protocol, which bridge the gulf between traditional statistical inference and exploratory data analysis.
+%The Rorschach protocol consists of a set of $m$ (usually, $m=20$) plots (called the {\it null plots}) rendered from data that is consistent with a given null model.
+The Rorschach protocol helps to understand the extent of randomness in the null model.
+Here, we use the lineup protocol.
+Under this protocol, we begin with a data set of interest to us, such as a network, and a visualization of this data, such as a node-link diagram.
+We would like to know the model that generated this data.
+There are two hypothesis considered: a null hypothesis which states the model that is assumed to have generated the data, and an alternative hypothesis that the data were not generated under this model.
+Then, the plot of the data of interest is placed randomly among a set of $m-1$ null plots, which are visualizations of data simulated from the null model (where $m =20$, usually).
+Human observers are then asked to examine the lineup and to identify the plot(s) that look(s) most different from the others.
+If an observer identifies the data plot, this is quantifiable evidence against the null hypothesis.
+Since an observer has a chance of 1 in $m$ to pick the data plot from the lineup by simply guessing, i.e.
+in a situation where the data plot is virtually indistinguishable from the null plots, the evidence grows in strength with the number of independent observers identifying the data plot.
+
+The lineup protocol places a plot firmly in the framework of hypothesis tests: a plot of the data is considered to be the test statistic, which is compared against the sampling distribution under the null hypothesis represented by the null plots.
+Obviously, the null generating mechanism, i.e.\ the method of obtaining the data for null plots, is crucial for the lineup protocol, as the null hypothesis directly affects the choice of null generating method.
+Null generating methods are typically based on (a) simulation, if the null hypothesis allows us to directly specify a parametric model, (b) sampling, as for example in the case of large data sets, or (c) permutation of the original data \citep[see e.g.\ ][]{Good05}, which allows for non-parametric testing that preserves marginal distributions  while ensuring independence in higher dimensions.
+The model of interest here allows us to simulate directly from a parametric model for dynamic social network data.
+%In the experimental data that we analyzed the null generating methods used were permutation methods and direct simulation from a null model.
+
+The lineup protocol was formally tested in a head-to-head comparison with the equivalent conventional test in \citet{majumder:2011}.
+The experiment utilized human subjects from Amazon's Mechanical Turk \citep{turk} and used simulation to control conditions.
+The results suggest that  visual inference is comparable to conventional tests in a controlled conventional setting.
+This provides support for its appropriateness for testing in situations where no conventional test exists or is difficult.
+%Interestingly, the power of a visual test increases with the number of observers engaged to evaluate lineups, and the pattern in results suggests that the power will provide results consistent with practical significance \citep{kirk:1996}.
+
+\section{Example Data and Models}\label{sec:senatedata}
+
+\noindent \textbf{Data:} The data we use are collaboration networks in the United States Senate during the $111^{th}$ through $114^{th}$ Congresses, overlapping with Barack Obama's presidency.
+These senates began on January 6, 2009, the start date of the $111^{th}$ Congress, and ended on January 3, 2017, the last date of the $114^{th}$ Congress\footnote{Details of how this data can be downloaded are provided by François Briatte at \url{https://github.com/briatte/congress}.}.
 In the US Senate, there are three ways that senators can show support for a piece of legislation: they can author the bill, cosponsor the bill, and vote for the bill.
@@
 \end{equation}
-For each of the four senate sessions, \st{in addition to the} WPC value between any two senators in the session, we have three node covariates: the party affiliation of each senator, the number of bills they authored in each session, and their gender.
+For each of the four senate sessions, in addition to the WPC value between any two senators in the session, we have three node covariates: the party affiliation of each senator, the number of bills they authored in each session, and their gender.
 We explored each of these covariates in the model to determine if they affect the overall network structure and how ties are formed between senators.
@@
 For Senate 111, for instance, we see Hillary Clinton, serving out her second term in the senate until she became Secretary of State.
-She is isolated in Figure~\ref{fig:senateAll}, but in actuality, she had many cosponsors on two pieces of legislation she authored in that short time, as is shown in Figure~\ref{fig:senateClinton}.
-We chose to remove Clinton and her edges from the network because they make the overall structure look so different from the other three senates, \st{where} the pattern is not typical of a senate in any other year.
-We suspect that because Hillary Clinton had \st{recently run for President had just} been \st{selected as Obama's} Secretary of State, the cosponsorships were largely symbolic, so the $111^{th}$ Senate without Hillary Clinton is more typical than the $111^{th}$ Senate with her.
-
-<<senateClinton, fig.cap="We removed Hillary Clinton's ties from the network because she had abnormally high collaboration with senators during the time she was in the 111th senate and before she left office to become Secretary of State.", fig.height=3>>=
+She is isolated in Figure~\ref{fig:senateAll}, but in actuality, she had many cosponsors on only two pieces of legislation she authored in that short time.%, as is shown in Figure~\ref{fig:senateClinton}.
+We chose to remove Clinton and her edges from the network because they make the overall structure look vastly more connected than the other three senates.
+We suspect that because she had just recently run for President and had just been selected as Obama's Secretary of State, the cosponsorships she garnered between the start of the new congress and becoming Secretary of State were largely symbolic, so the $111^{th}$ Senate without Hillary Clinton is more typical than the $111^{th}$ Senate with her.
+In fact, she had several co-sponsors on both of her pieces of legislation, which meant she had the maximum WPC value of one with about a dozen other senators.
+Thus, though she is in the Senate for a time, subsequent analyses will not include her, and will instead include her replacement, Kirsten Gillibrand.
+
+<<senateClinton, fig.cap="We removed Hillary Clinton's ties from the network because she had abnormally high collaboration with senators during the time she was in the 111th senate and before she left office to become Secretary of State.", fig.height=3, eval=FALSE>>=
 se111clint$Clinton <- 'Yes'
@@
 
-In legislative cosponsorship networks, it is well known that party affiliation, reciprocity of relationships, and whether senators are "workhorses" who author many bills or "showhorses" who author few bills, are major influences on structure \citep{legnet}.
-We focus on these covariates, \st{plus the additional sex covariate} when choosing which SAO models to fit to the data.
-
-\subsection{Models of Interest}\label{sec:models}
-
-In addition to considering already well-known effects in legislative networks for application of our significance and goodness-of-fit methods, we first fit many other possible models and selected a few significant effects.
-To determine the effects that we would move forward with, we followed this procedure:
-
-\begin{enumerate}
-\item Define the simple effects structure of the data: the rate parameters and the outdegree and reciprocity parameters. 
-\item Add each additional possible effect, \st{as determined by the effects documentation function}, in \texttt{RSiena} one-at-a-time to the model\st{'s objective function}  \citep{RSiena}.
-\item Fit each model to the data and check for convergence.
-    \begin{enumerate}
-    \item If the model converged, move to 4.
-    \item If the model did not converge, use the previous fitted values as starting values and repeat 5 times or until convergence, whichever comes first.
-    \end{enumerate}
-\item Test the added parameter for significance using a Wald-type test.
-\item Report out the estimate of the additional parameter, its standard error, Wald $p$ value, and convergence criterion.
-\end{enumerate}
-
-After completing the procedure for all model effects, we selected effects whose estimates converged, had a Wald $p$-value of less than 0.10, and seemed to have a reasonable interpretation for our data according to well-known properties of legislative networks \citep{legnet}.
-
-The parameters we use for the remainder of the paper are detailed in Table~\ref{tab:effects}.
-The most significant effect was the jumping transitive triplet (JTT) parameter for the party covariate, which was estimated to be about -6 with a standard error of 0.11, resulting in a $p$-value of less than 0.0001.
-This estimate of the parameter associated with this statistic considers the number of transitive closures formed between two senators from different parties.
-The negative estimate is an indication that forming transitive ties between two people from different parties is discouraged, which tracks with the divisive nature of American politics, where party affilitation is \st{the dominant effects}.
-Another significant effect was the same JTT parameter for the sex covariate, with an estimate of about 3 with a standard error of 0.89.
-\st{This parameter also consideres transitive closures, but for senators of different genders.
-The positive value indicates that transitive ties between senators of different genders are more likley to form.} The covariate-related similarity score-weighted transitive triplets parameter estimate for the number of bills authored by a senator was also significant.
-\st{We chose to look at similarity because the number of bills authored is more continuous than gender or party.
-The similarity measure is computed as:
-\begin{equation}\label{eq:similar}
-sim^b_{ij} = \frac{\max_{hk}|b_h - b_k| - |b_i - b_j|}{\max_{hk}|b_h - b_k|} 
-\end{equation}
-where $\max_{hk}|b_h - b_k|$ is the range of number of bills authored by senators, and $b_i$, $b_j$ are the number of bills authored by senators $i,j$ repectively in the senate period.} This effect was estimated at about 10 with standard error of 3.9.
-The high positive estimate suggests senators \st{are encouraged} to collaborate with other senators who author about the same number of bills they do.
-This tendency of senators to cosponsor bills written by senators who are similarly ``prolific" corresponds to another well-known property of the U.S.
-Senate structure: the tendency of senators to be either ``workhorses" or ``showhorses".
-Senators known as workhorses author many pieces of legislation in a session, and largely stay out of the public arena.
-The showhorse senators, on the other hand, author relatively few pieces of legislation, and tend to appear on television, radio, and other media a great deal.
-Finally, we found the same party transitive triplet effect was also significant, with a fitted value of 1.3 and standard error of 0.7, meaning that transitive relationships between senators tend to form when they are from the same party, \st{exactly as we would expect in a legislative body in a country with extremely entrenced partisan divides as in the US.}
-
-<<geteffects, eval = FALSE>>=
-# used for table below
-initeff <- read_csv("data/sigEffsSenate912.csv")
-initeff %>% filter(shortName == "sameX")
-#density & -- & $\sum_j x_{ij}$ & & & \\
-#reciprocity & -- & $\sum_j x_{ij}x_{ji}$ & & & \\
-@
-
-\begin{table}
-\scalebox{0.8}{
-\begin{tabular}{c|p{2cm}|p{2.1cm}|c|l|p{1.75cm}|p{2cm}}
-$\beta_k$ & {\bf Effect name} & {\bf Interaction Variable} & {\bf Formula} & {\bf Picture} & {\bf Initial estimate} & {\bf Wald $p$-value} \\
-\hline 
-$\beta_3$ & jumping transitive triplet & party & $s_{i3}(x, \mathbf{p}) = \sum_{j\neq h} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(p_i = p_h \neq p_j)$ & \includegraphics[width=.6in]{img/jttp.png} & -5.884 & $<0.0001$\\
-$\beta_4$ & jumping transitive triplet & sex & $s_{i4}(x, \mathbf{s}) = \sum_{j\neq h} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(s_i = s_h \neq s_j)$ & \includegraphics[width=.6in]{img/jtts.png}  & 3.335 & 0.0002 \\
-$\beta_5$ & similarity transitive triplet & bills & $s_{i5}(x, \mathbf{b}) = \sum_{j} x_{ij}x_{ih}x_{hj}\cdot (sim^b_{ij} - \overline{sim}^b)^*$ & \includegraphics[width=1in]{img/simttb.png} & 9.821 & 0.0128 \\
-$\beta_6$ & same transtive triplet & party & $s_{i6}(x, \mathbf{p}) =\sum_{j} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(p_i = p_j)$ & \includegraphics[width=1in]{img/samettp.png}  & 1.306 & 0.0642 %\\
-%$\beta_7$ & same & party & $s_{i7}(x, \mathbf{p}) = \sum_j x_{ij}\mathbb{I}(p_i = p_j)$ & \includegraphics[width=.6in]{img/samep.png}  & 0.363 & 0.0074
-\end{tabular}
-}
-\caption{\label{tab:effects} The additional effects we used in the SAOMs fit to the senate data.
-* - $sim^b_{ij} = \frac{\max_{hk}|b_h - b_k| - |b_i - b_j|}{\max_{hk}|b_h - b_k|}$ is the similarity score between two senators based on the number of bills authored, and $\overline{sim}^b = \frac{1}{n(n-1)}\sum_{i\neq j} sim^b_{ij}$ is the average bill similarity score between any two senators.}
-\end{table}
-
+In legislative cosponsorship networks, it is well known that party affiliation, reciprocity of relationships, and whether senators are ``workhorses" who author many bills or ``showhorses" who author few bills, are major influences on structure \citep{legnet}.
+We focus on these covariates, plus the additional gender covariate when choosing which SAO models to fit to the data.
+
+\noindent \textbf{Models:} In order to determine which models to examine, we first consider already well-known effects in legislative networks.
+We also selected some effects which were significant according to the built-in Wald-type tests in \texttt{RSiena} for application of our significance and goodness-of-fit methods.
 We examine a total of six models, each identified by its objective function:
@@
 
-We fit models M1 through M6 in \texttt{RSiena} using Markov Chain Monte Carlo (MCMC) methods to approximate the method of moments estimates of the parameters.
+The parameters in the objective function, $\beta_1, \dots, \beta_6$ are defined in Table~\ref{tab:effects}.
+
+%To determine the effects that we would move forward with, we followed this procedure:
+
+% \begin{enumerate}
+% \item Define the simple effects structure of the data: the rate parameters and the outdegree and reciprocity parameters. 
+% \item Add each additional possible effect, \st{as determined by the effects documentation function}, in \texttt{RSiena} one-at-a-time to the model\st{'s objective function}  \citep{RSiena}.
+% \item Fit each model to the data and check for convergence.
+%     \begin{enumerate}
+%     \item If the model converged, move to 4.
+%     \item If the model did not converge, use the previous fitted values as starting values and repeat 5 times or until convergence, whichever comes first.
+%     \end{enumerate}
+% \item Test the added parameter for significance using a Wald-type test.
+% \item Report out the estimate of the additional parameter, its standard error, Wald $p$ value, and convergence criterion.
+% \end{enumerate}
+%  
+% After completing the procedure for all model effects, we selected effects whose estimates converged, had a Wald $p$-value of less than 0.10, and seemed to have a reasonable interpretation for our data according to well-known properties of legislative networks \citep{legnet}.
+
+%The parameters we use for the remainder of the paper are detailed in Table~\ref{tab:effects}.
+All models had the first two parameters, $\beta_1, \beta_2$ at minimum included.
+Other effects were added one at a time so that the simplest model $M1$ is nested in each of the other models.
+The most significant additional effect was $\beta_3$, the jumping transitive triplet (JTT) parameter for the party covariate, which was estimated to be about -6 with a standard error of 0.11, resulting in a $p$-value of less than 0.0001.
+This estimate of the parameter associated with this statistic considers the number of transitive closures formed between two senators from different parties.
+The negative estimate is an indication that forming transitive ties between two people from different parties is discouraged, which corresponds to the divisive nature of American politics, where party affilitation is the dominant effect.
+Another significant effect was $\beta_4$, the same JTT parameter for the sex covariate, with an estimate of about 3 with a standard error of 0.89.
+This parameter also consideres transitive closures, but for senators of different genders.
+The positive value indicates that transitive ties between senators of different genders are more likley to form.
+Another significant effect was $\beta_6$, the covariate-related similarity score-weighted transitive triplets parameter estimate for the number of bills authored by a senator.
+We chose to look at similarity instead of covariate "jumps" because the number of bills authored is more continuous than gender or party.
+The similarity measure is computed as:
+\begin{equation}\label{eq:similar}
+sim^b_{ij} = \frac{\max_{hk}|b_h - b_k| - |b_i - b_j|}{\max_{hk}|b_h - b_k|} 
+\end{equation}
+where $\max_{hk}|b_h - b_k|$ is the range of number of bills authored by senators, and $b_i$, $b_j$ are the number of bills authored by senators $i,j$ repectively in the senate period.
+This effect was estimated at about 10 with standard error of 3.9.
+The high positive estimate suggests senators are encouraged to collaborate with other senators who author about the same number of bills they do.
+This tendency of senators to cosponsor bills written by senators who are similarly ``prolific" corresponds to another well-known property of the U.S.
+Senate structure: the tendency of senators to be either ``workhorses" or ``showhorses".
+Senators known as workhorses author many pieces of legislation in a session, and largely stay out of the public arena.
+The showhorse senators, on the other hand, author relatively few pieces of legislation, and tend to appear on television, radio, and other media a great deal.
+So, showhorses and workhorses tend to stick together.
+Finally, we found $\beta_5$, the same party transitive triplet effect was also significant, with a fitted value of 1.3 and standard error of 0.7, meaning that transitive relationships between senators tend to form when they are from the same party, exactly as we would expect in a legislative body in a country with extremely entrenched partisan divides as in the US.
+
+<<geteffects, eval = FALSE>>=
+# used for table below
+initeff <- read_csv("data/sigEffsSenate912.csv")
+initeff %>% filter(shortName == "sameX")
+#density & -- & $\sum_j x_{ij}$ & & & \\
+#reciprocity & -- & $\sum_j x_{ij}x_{ji}$ & & & \\
+@
+
+\begin{table}
+\scalebox{0.8}{
+\begin{tabular}{c|p{2cm}|p{2.1cm}|c|l|p{1.75cm}|p{2cm}}
+$\beta_k$ & {\bf Effect name} & {\bf Interaction Variable} & {\bf Formula} & {\bf Picture} & {\bf Initial estimate} & {\bf Wald $p$-value} \\
+\hline 
+$\beta_1$ & density & NA & $s_{i1}(x) = \sum_{j} x_{ij}$ & \includegraphics[width=.6in]{img/density.png} & 2.204 & NA \\ 
+$\beta_2$ & reciprocity & NA & $s_{i2}(x) = \sum_{j} x_{ij}x_{ji}$ & \includegraphics[width=.6in]{img/recip.png} & -4.903 & NA \\
+$\beta_3$ & jumping transitive triplet & party & $s_{i3}(x, \mathbf{p}) = \sum_{j\neq h} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(p_i = p_h \neq p_j)$ & \includegraphics[width=.6in]{img/jttp.png} & -5.884 & $<0.0001$\\
+$\beta_4$ & jumping transitive triplet & sex & $s_{i4}(x, \mathbf{s}) = \sum_{j\neq h} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(s_i = s_h \neq s_j)$ & \includegraphics[width=.6in]{img/jtts.png}  & 3.335 & 0.0002 \\
+$\beta_5$ & similarity transitive triplet & bills & $s_{i5}(x, \mathbf{b}) = \sum_{j} x_{ij}x_{ih}x_{hj}\cdot (sim^b_{ij} - \overline{sim}^b)^*$ & \includegraphics[width=1in]{img/simttb.png} & 9.821 & 0.0128 \\
+$\beta_6$ & same transtive triplet & party & $s_{i6}(x, \mathbf{p}) =\sum_{j} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(p_i = p_j)$ & \includegraphics[width=1in]{img/samettp.png}  & 1.306 & 0.0642 %\\
+%$\beta_7$ & same & party & $s_{i7}(x, \mathbf{p}) = \sum_j x_{ij}\mathbb{I}(p_i = p_j)$ & \includegraphics[width=.6in]{img/samep.png}  & 0.363 & 0.0074
+\end{tabular}
+}
+\caption{\label{tab:effects} The effects we used in the SAOMs fit to the senate data.
+In the picture which represents each effect, the dotted tie is encouraged to form if the estimate is positive, and discouraged to form if the estimate is negative.
+$^*$ - $sim^b_{ij} = \frac{\max_{hk}|b_h - b_k| - |b_i - b_j|}{\max_{hk}|b_h - b_k|}$ is the similarity score between two senators based on the number of bills authored, and $\overline{sim}^b = \frac{1}{n(n-1)}\sum_{i\neq j} sim^b_{ij}$ is the average bill similarity score between any two senators.}
+\end{table}
+
+We fit all six of our models in \texttt{RSiena} using Markov Chain Monte Carlo (MCMC) methods to approximate the method of moments estimates of the parameters.
 Because the estimation is done through MCMC simulation, we fit each model to the data 1,000 times to get a better estimate of the true value of $\boldsymbol{\beta}$.
-From the simulations that converged, which made up over 90\% of the fits for each model, we computed the mean of the 1,000 estimates of each parameter to get final estimates of  $\hat{\boldsymbol{\beta}}$ for each model, which are given in Table~\ref{tab:fittedvalues}.
+From the simulations in which the MC converged, which made up over 90\% of the fits for each model, we computed the mean of the 1,000 estimates of each parameter to get final estimates of  $\hat{\boldsymbol{\beta}}$ for each model, which are given in Table~\ref{tab:fittedvalues}.
 
@@
 
-We want to explore the role of each of these parameters in the objective functions for each model.
+We want to determine the role that each of these parameters plays in the objective functions for the different models.
 So, we use the estimates given in Table~\ref{tab:fittedvalues} to simulate from models M1 through M6.
@@
 
-\section{Experiment Set-Up}\label{sec:expersetup}
+\section{Experimental Design}\label{sec:expersetup}
 
 We want to explore three different aspects of the SAOM models using the lineup protocol: (1) significance of parameters, (2) goodness-of-fit of a model, and (3) \st{visual power of the effects.} %visual detection of parameters.
-Each one of these situations requires a different setup, which we describe in detail, but we make ue of the lineup protocol for all of these aspects.
+Each one of these situations requires a different setup, which we describe in detail, making use of the lineup protocol to study all of these aspects.
 
@@
 
- Typically, a lineup shows sets of 20 plots at a time c.f.~\citet{loy:2015, vanderplas:2016}, but we determined that not enough structure could be shown in each plot for 20 node-link diagrams.
-We chose to expose our participants to only six plots at a time in order to show the node-link diagrams in more detail and to lower cognitive load for participants.
+ Typically, a lineup shows sets of 20 plots at a time c.f.~\citet{loy:2015, vanderplas:2016}, but we determined that not enough structure could be shown in each plot for 20 large node-link diagrams.
+We chose to present our participants with only six plots at a time in order to show the node-link diagrams in more detail and to lower cognitive load for participants.
 To construct a lineup, we simulate five networks from the null model and one network from the alternative model.
@@
 To simulate lineups from the models we used the \texttt{siena07} function in \texttt{RSiena} \citep{RSiena}.
-\st{Sections~\ref{sec:setupst} through~\ref{sec:setupvp} describe in detail how we set up the lineups, which parameter values we simulate from, and why.} Sets of the lineups we create are shown to independent observers recruited through Amazon Mechanical Turk for feedback (more details on the Turk setup in Section~\ref{sec:res}).
+Sections~\ref{sec:setupst} through~\ref{sec:setupvp} describe in detail how we set up the lineups, which parameter values we simulate from, and why.
+Sets of the lineups we create are shown to independent observers recruited through Amazon Mechanical Turk for feedback (more details on the Turk setup in Section~\ref{sec:res}).
 %, we set the parameters to the values given in Table~\ref{tab:fittedvalues} for all parameters within the respective models, with the exception of $\beta_5$.
@@
 Thus, the five null plots in the lineup are simulations from M1 with $\beta_1, \beta_2$ set to the estimates given in Table~\ref{tab:fittedvalues} for M1.
-The alternate model is the model with $\beta_1, \beta_2$, and $\beta_k$ in the objective function, with the remaining plot simulated from the appropriate model with values set to the estimates from Table~\ref{tab:fittedvalues}.
+The alternative model is the model with $\beta_1, \beta_2$, and $\beta_k$ in the objective function.
+The alternative data plot is simulated from the appropriate model also with values set to the estimates from Table~\ref{tab:fittedvalues}.
 
 The lineup generated under this scenario is shown to a number of independent viewers.
-If an observer picks out the alternative data plot, that is evidence \st{against} the null hypothesis, while picking one of the null plots is evidence \st{in favor of} the null hypothesis.
+If an observer picks out the alternative data plot, that is evidence against the null hypothesis, while picking one of the null plots is evidence in favor of the null hypothesis.
 The significance tests that we perform in our experiment are for $\beta_3$ and $\beta_4$, making the alternative models M3 and M4, respectively, for three repitions each.
@@
 We pick a wave to focus on, wave two, which is the first simulated network, and among these five plots, we place a node-link diagram of the true second wave data.
-We cannot show the data more than once to each participant, so we examine several different models with three repitions each in our Amazon Mechanical Turk experiment, with each participant never seeing the true data wave twice.
+We cannot show the data more than once to each participant, so we examine several different models with three repitions each in our Amazon Mechanical Turk experiment, with each participant never seeing the actual data twice.
 The models we chose for goodness-of-fit testing are M3, M4, M5, and M7.
@@
 By \textit{noticeable}, we mean that the inclusion of the effect alters the appearance of networks simulated from a model to a degree that viewers are able to reliably pick out a node-link diagram rendered from data simulated from a model \textit{with} the effect in a lineup among plots \textit{without} the effect.
-This is a way to determine the power of the visual \st{significance} test.
+This is a way to determine the power of the visual significance test.
 We explore all parameters in the objective function, $\beta_1, \dots, \beta_6$ in this way.
@@
 In model M1, with only two parameters in the objective function, we varied both the density and reciprocity parameter values one at a time, \st{keeping all other parameters at their fitted values given in Table~\ref{tab:fittedvalues}}.
-In models M3 through M6, we vary the additional parameter, $\beta_3$ through $\beta_6$, respectively.
+In models M3 through M6, we vary the additional parameter, $\beta_3$ through $\beta_6$, respectively, while again holding the rate and other parameters at their value in Table~\ref{tab:fittedvalues}.
 Thus, we have six different parameters of interest to us in total: $\beta_1, \dots, \beta_6$.
-We want to determine how the size of these parameters affects the overall structure of the network data simulated from the models M1 through M6, so we also vary the value of the parameters \st{in order to determine at what level the effects become noticeable.} %in both negative and positive directions.
+We want to determine how the size of these parameters affects the overall structure of the network data simulated from the models M1 through M6, so we also vary the value of the parameters in order to determine at what level the effects become noticeable.
+%in both negative and positive directions.
 
@@
 
-Using the ``Picking Lineups" web application, we settled on six parameter values to test for each of our six effects, $\beta_1, \dots \beta_6$.
+Using a ``Picking Lineups" web application, we settled on six parameter values to test for each of our six effects, $\beta_1, \dots \beta_6$.
 The complete details of the parameters tested using the lineup protocol is given in Table~\ref{tab:experdetail}.
@@
 experdet$test_param <- as.factor(experdet$test_param)
-levels(experdet$test_param) <- paste0("beta", c(1,3,4,2,6,5))
-experdet$test_param <- factor(experdet$test_param, levels=paste0("beta", c(1,2,3,5,4,6)))
+levels(experdet$test_param) <- paste0("\\$\\beta_", c(1,2,3,5,4,6), "\\$")
+experdet$test_param <- factor(experdet$test_param, levels=paste0("\\$\\beta_", c(1,2,3,5,4,6), "\\$"))
 experdet <- experdet %>% arrange(test_param)
@@
 
-\section{Experiment Results}\label{sec:res}
+\section{Results}\label{sec:res}
 
@@
 
+% what did we do? what did we learn from it? 
 By using visual inference methods, we have developed new ways to perform significance and goodness-of-fit testing for a complicated and intractable set of statistical models for social network data.
@@
 
+% what are the limitations of these findings
+
 These results do not come without limitations.
@@
 \st{We can make some generalizations about what participants are picking up on in the lineups based on their feedback and previous research, but we cannot apply our hierarchical model directly to lineups constructed for new data or new models or parameters.}
+
+% what is the future of this research?
 

### 2019-01-25 ASApaper/paper.Rnw
some rearranging, re-write abstract.

--- 2017-10-19 ASApaper/paper.Rnw
@@
 \usepackage{subcaption}
-\usepackage{animate}
+%\usepackage{animate}
 \usepackage{afterpage}
@@
 \usepackage{amsfonts}
-% \usepackage{color}
+%\usepackage{color}
 \usepackage{multirow}
@@
 \usepackage{hyperref}
+\usepackage{numprint}
+
 \graphicspath{{figure/}}
 % \usepackage{endfloat} % Figures to the end of the document
+\newcommand{\st}[1]{{\color{orange} #1}}
+\newcommand{\hh}[1]{{\color{magenta} #1}}
 
@@
 {
-  \title{\bf Visual Inference for Significance and Goodness-of-Fit Testing for a Social Network Model}
+  \title{\bf Visual Inference for a Social Network Model}
   \author{Samantha Tyner\thanks{
@@
   \begin{center}
-    {\LARGE\bf Visual Inference for Significance and Goodness-of-Fit Testing for a Social Network Model}
+    {\LARGE\bf Visual Inference for a Social Network Model}
 \end{center}
@@
 \begin{abstract}
-Abstract goes here.
+Three of the most important assessments of a statistical model are significance tests of parameters, goodness-of-fit tests, and power calculations of the tests.
+All three tasks become more difficult as the model becomes more complex.
+We will explore the three assessments of one particularly complex set of models, continuous time Markov chain (CTMC) models, for dynamic social networks  \citep{saompaper}.
+In this paper, we propose new methods for significance and goodness-of-fit testing, as well as power calculation for CTMC models via the visual inference (VI) paradigm of \citet{Bujaetal}.
+With VI, we can look at entire datasets simulated from a model, instead of relying a single metric such as a $p$-value for a single parameter in the model or goodness-of-fit assessed by looking at outdegree.
+We conducted an experiment through Amazon Mechanical Turk to assess the significance of parameters, the fit of a model, and the visual power of parameters.
 \end{abstract}
@@
 
-\tableofcontents
+%\tableofcontents
 \newpage
 
-<<child-demo, child='paper-child.Rnw'>>=
+<<child-demo, child='paper-child-thesis-ch-2.Rnw'>>=
 @

### 2019-01-28 ASApaper/paper.Rnw
finish abstract

@@
 With VI, we can look at entire datasets simulated from a model, instead of relying a single metric such as a $p$-value for a single parameter in the model or goodness-of-fit assessed by looking at outdegree.
-We conducted an experiment through Amazon Mechanical Turk to assess the significance of parameters, the fit of a model, and the visual power of parameters.
+We conducted a VI experiment, with participants recruited via Amazon Mechanical Turk, to assess the significance of CTMC parameters, the fit of a CTMC model, and the visual power of CTMC parameters.
 \end{abstract}

### 2019-01-28 ASApaper/paper-child.Rnw
starting to make my way through.......

@@
 % }
-When choosing and fitting statistical models, there are often three key evaluations of interest: significance testing of model parameters, model goodness-of-fit tests, and determining the power of those tests.
-With significance testing, the data are typically assumed to come from a simple model under the null hypothesis, and additional parameters are tested to determine whether they significantly contribute to explaining the variability in the data.
-In the second, the model of interest is examined to determine how well it fits the data.
-In the final, the ability of the hypothesis test in question to detect the difference between the null and alternative hypothesis is determined.
-All three of these aspects of statistical modeling increase greatly in difficulty as the complexity of the model of interest increases.
+When selecting and fitting statistical models, there are typiccally three key assessments of interest: significance tests of parameters, goodness-of-fit tests, and power calculations of the tests.
+With significance testing, the null hypothesis assumes that the data come from a simple model nested within the model of interest.
+We conduct significance tests of one or more additional parameters to determine if they significantly explain the variability in the data.
+For goodness-of-fit tests, we examine one or more models of interest to assess how well they fit the data.
+Finally, power tests quantify the ability of the hypothesis test to detect the difference between the null and alternative hypothesis.
+All three of these aspects of statistical modeling are vital for inference.
 %The more complicated the model, the harder it is to determine fit or to include or exclude a parameter.
 
-Some particularly complicated families of models are those designed to model network change.
-A \textit{network} is any set of things, such as people, computers, or neurons, that are connected in some way, through social relations, internet connection, or electrical impulses in the brain.
-We refer to the ``things" in the network as \textit{nodes}, or \textit{actors} in a social network, and the connections as \textit{edges}, or \textit{ties} in a social network.
-Dependencies inherent to the data make network objects particularly difficult to model. %Even more challenging is the situation 
-The difficulty of modeling increases when we examine more than a single instance of a network and consider the dynamics of network change between observed instances.
-This type of modelling for dynamic networks is often performed on social network data, such as friendship networks among students or the spread of HIV in drug users sharing needles.
-%sexually active communities.% of men sleeping with other men.
-X ST a bit less jarring hopefully.
+The more complex the model, however, the harder it is to assess significance, power, and goodness-of-fit.
+One particularly complicated family of models are those designed to model network change.
+A \textit{network} is any set of things, such as people, or computers, that are connected in some way, through social relations or the Internet.
+In a social network, the people are the \textit{nodes}, or \textit{actors}, and the relationships between the people are \textit{edges}, or \textit{ties}.
+Models for networks are particularly complex, as
+dependencies inherent to network data make them  difficult to model. %Even more challenging is the situation 
+And this difficulty only increases when studying \textit{dynamic networks}, the same set of nodes and their changing relationships observed at many points in time.
+Trying to determine the mechanism underlying change in a network over time only makes the model more complex.
+Yet, researchers are often interested in modeling dynamic social networks, such as friendship networks among students or collaboration networks between researchers.
+% X ST a bit less jarring hopefully.
 X
-%These 
-It is known that most network models lack the asymptotics required to perform many well-known goodness-of-fit tests, and the maximum likelihood estimation of parameters is so difficult that it can make significance testing difficult as well, see for example \citet{holland81}.
+Even some of the simplest network models, however, lack the asymptotics required to perform traditional goodness-of-fit tests \citep{holland81}.
+In additional, direct maximum likelihood estimation of model parameters is frequently impossible due to the intractability of the models \citep{hummeletal}.
 %\hh{is the previous sentence a quote? I'm not sure that I follow.} %\st{it's a paraphrase... shall I make a quote? Also I definitely botched a word. Hopefully makes more sense now? }
-
+% stop 4pm on 1/28/19
 In order to circumvent some of these difficulties, we propose new methods for significance testing of parameters, goodness-of-fit testing, and power calculations %of these 

### 2019-02-26 ASApaper/paper-child.Rnw
looks like i forgot to commit stuff

@@
 % }
-When selecting and fitting statistical models, there are typiccally three key assessments of interest: significance tests of parameters, goodness-of-fit tests, and power calculations of the tests.
+When selecting and fitting statistical models, there are typically three key assessments of interest: significance tests of parameters, goodness-of-fit tests, and power calculations of the tests.
 With significance testing, the null hypothesis assumes that the data come from a simple model nested within the model of interest.
-We conduct significance tests of one or more additional parameters to determine if they significantly explain the variability in the data.
+We then conduct significance tests of one or more additional parameters to determine if they significantly explain the variability in the data.
 For goodness-of-fit tests, we examine one or more models of interest to assess how well they fit the data.
@@
 dependencies inherent to network data make them  difficult to model. %Even more challenging is the situation 
-And this difficulty only increases when studying \textit{dynamic networks}, the same set of nodes and their changing relationships observed at many points in time.
-Trying to determine the mechanism underlying change in a network over time only makes the model more complex.
+This difficulty only increases when studying \textit{dynamic networks}, the same set of nodes and their changing relationships observed at many points in time, because of the added temporal dependence.
 Yet, researchers are often interested in modeling dynamic social networks, such as friendship networks among students or collaboration networks between researchers.
@@
 In order to circumvent some of these difficulties, we propose new methods for significance testing of parameters, goodness-of-fit testing, and power calculations %of these 
-for a set of social network models: stochastic actor-oriented models for dynamic network data, first introduced by \citet{saompaper}.
+for one family social network models: continuous time Markov-chain (CTMC) models for dynamic network data, as defined in \citet{saompaper}.
 Specifically, we are using \textit{visual inference} in place of traditional statistical methods for social network models, such as Wald or $t$-tests for significance of parameters and in- and outdegree distribution metrics for determining goodness-of-fit.
-Visual inference, introduced by \citet{Bujaetal}, will allow us to look at the \textit{entire} dataset simulated from a network model as opposed to a (set of) usually one-dimensional metric(s) derived from the network such as outdegree distribution, or a $p$-value for a parameter in the model.
-By using visual inference to supplement traditional statistical tests, we gain a new understanding of parameters in SAOMs and fit of SAOMs to data.%In addition, we use visual inference to better understand some parameters in a social network model, by looking at how different values for parameters affect the networks simulated from these models.
+Visual inference (VI), introduced by \citet{Bujaetal}, will allow us to look at the \textit{entire} dataset simulated from a network model as opposed to a (set of) usually one-dimensional metric(s) derived from the network, or a $p$-value for a parameter in the model.
+By using visual inference to supplement traditional statistical tests, we gain a new understanding of the parameters in these CTMC models and ability to assess the fit of the CTMCs to dynamic network data.%In addition, we use visual inference to better understand some parameters in a social network model, by looking at how different values for parameters affect the networks simulated from these models.
 
@@
 
-\subsection{Stochastic Actor-Oriented Models}\label{subsec:saom}
+\subsection{CTMC Models for Dynamic Social Network Data}\label{subsec:saom}
 
 %To model family that we apply to the Senate data is the Stochastic Actor-Oriented Model (SAOM) family. 
-Stochastic Actor-Oriented Models (SAOMs) are a family of models for dynamic network data that incorporate both network structure and node-level information to describe how a network observed on two or more occasions changes over time \citep{saompaper}.
-The two titular properties of SAOMs, stochasticity and actor-orientation, are crucial to understanding networks as they exist naturally: social networks are ever-changing as relationships decay or grow in seemingly random ways, and the actors in them have characteristics that could affect how they change their ties to other nodes in the network.
+CTMCs are a family of models for dynamic network data that incorporate both network structure and node-level information to describe how a network observed on two or more occasions changes over time \citep{saompaper}.
+The CTMCs model two key properties: social networks are ever-changing as relationships decay or grow in seemingly random ways, and the actors in them have characteristics that could affect how they change their ties to other nodes in the network.
 % Talk about SAOMs
 These unique properties allow for the fitting of some very complicated models to inherently complex data, so it can be exceedingly difficult to interpret parameters and their corresponding estimates.
-The sheer amount of available parameters to include in the model combined with the difficulty of interpretation make parameter selection and goodness-of-fit testing burdensome as well.
-
-Broadly, a SAOM takes network structure and node covariate information into account in two ways and models the network changes one-at-a-time as a continuous time Markov chain (CTMC).
-First, the rate of change between states is dictated by a rate function that describes \textit{how often} changes in the network occur, and secondly, the objective function describes \textit{what} those state changes are.
-As in most other network models, the variables of interest are the edges of the network, $x_{ij}$, where $x_{ij}$ denotes the edge between nodes $i$ and $j$, where $i,j \in \{1, 2, \dots, n=\text{the number of nodes}\}$.
-$x_{ij}$ is modelled as a binary variable, i.e.
+The large amount of available parameters to include in the model combined with the difficulty of interpretation make parameter selection and goodness-of-fit testing burdensome as well.
+
+Broadly, a CTMC model takes network structure and node covariate information into account in two ways and models the network changes one-at-a-time.
+First, the rate of change between states is dictated by a rate function that describes \textit{how often} changes in the network occur, and second, the objective function describes \textit{what} those state changes are.
+As in most other network models, the variables of interest are the edges of the network, $x_{ij}$, where $x_{ij}$ denotes the edge between nodes $i$ and $j$, where $i \neq j \in \{1, 2, \dots, n=\text{the number of nodes}\}$.
+$x_{ij}$ is binary, i.e.
   \begin{equation}\label{eq:edgevars}
@@
   \end{equation}
-Edges are treated as \textit{directed}, i.e. in general $x_{ij} \neq x_{ji}$, and self-referencing edges or loops are not allowed, i.e.  $x_{ii} = 0$ for all $i=1, 2, \dots, n$. 
-The network is observed $M \geq 2$ times at time points  $t_1 < t_2 < ...
-< t_M$, and the entire network at time point $t_m$ is denoted as $x(t_m)$.
-In sections~\ref{rf} and~\ref{of} we discuss the rate and objective functions of a SOAM in more depth.
-Additional details on SOAMs can be found in \citet{saompaper, snijders01, snijders:2010, snijders2010, snijdersetall:2007, snijders:2017},
-
-\subsubsection{Rate Function}\label{subsubsec:rf}
-
-All changes in SAOMs are treated as changes made by the nodes, or \textit{actors}, in the network, i.e.\ each actor, $i$, gets a chance to make a change according to the rate function, typically denoted $\lambda_i$, which dictates when relationships between nodes in the network can change.
-In general, the rate function can take the network structure e.g.
-outdegree of node $i$, and the node covariates into account, but we use the simple rate function, which is constant over all nodes in a given time period, so that $\lambda_i \equiv \alpha_m$ $\forall i \in \{1, 2, \dots, n\}$.
+Edges are treated as \textit{directed}, so that in general $x_{ij} \neq x_{ji}$, and self-referencing edges or loops are not allowed, i.e.  $x_{ii} = 0$ for all $i=1, 2, \dots, n$. 
+The network is observed $M \geq 2$ times at time points  $t_1 < ...
+< t_M$, and the entire network at time point $t_m$ is denoted as $x(t_m)$ for $m \in \{1, \dots, M\}$.
+We discuss the rate and objective functions of a CTMC model in more depth below.
+Additional details on these models can be found in \citet{saompaper, snijders01, snijders:2010, snijders2010, snijdersetall:2007, snijders:2017},
+
+\textbf{Rate Function:} All changes in SAOMs are treated as changes made by the nodes, or \textit{actors}, in the network.
+Each actor, $i$, gets a chance to make a change according to the rate function, which dictates when relationships between nodes in the network can change.
+In general, the rate function can take the network structure such as the outdegree of a node, and the node covariates into account, but we use a simple rate function, which is constant over all nodes in a given time period, $\alpha_m$ $\forall i \in \{1, 2, \dots, n\}$.
 We denote the rate from $t_{m}$ to $t_{m+1}$ as $\alpha_m$ for $m = 1, \dots, M-1$.
@@
 
-\subsubsection{Objective Function}\label{subsubsec:of}
-
-After actor $i$ has been given the opportunity to change, it probabilistically chooses one of its current ties, $x_{ij}$, to change.
-The probability that actor $i$ changes its current tie to actor $j$ is determined by the \textit{objective function} of the model and a random component, $U$, which can be thought of as encompassing any other factors that may be influencing the changes the node makes not accounted for by the parameters in the model.
+\textbf{Objective Function:} After actor $i$ has been given the opportunity to change, it randomly selects one of its current ties, $x_{ij}$, to change.
+The probability that actor $i$ changes its current tie to actor $j$ is determined by the \textit{objective function} of the model and a random error term $U$ with log-Weibull distribution \citep{modelsSnijders}.
+.
 %\hh{XXX can you paraphrase the purpose of U in half a sentence?}
@@
 %The objective function, $f_i$, is the driving force of a SAOM. 
-Actor $i$ is aiming to maximize its corresponding objective function $f_i$ given the current state of the network, $x$ and the node-level covariates, $\mathbf{Z}$, given as:% which has the form
+Actor $i$ aims to maximize its corresponding objective function $f_i$ given the current state of the network, $x$ and the node-level covariates, $\mathbf{Z}$.
+This function is defined as:% which has the form
 \begin{equation}\label{eq:objective}
@@
 \end{equation}
-
 This probability dictates which edge change is made by the acting node.
 The acting node can also choose to \textit{not} change at all.
-This is most likely to occur when the numerator, as calculated for the current state of the network, is larger than for any changes $x(i \leadsto j)$ that could be made.
+This is most likely to occur when the numerator of Equation~\ref{eq:transprob}, as calculated for the current state of the network, is larger than for any changes $x(i \leadsto j)$ that could be made.
 
@@
 Similarly, we denote the reciprocity parameter by $\beta_2$ and the associated statistic as $s_{i2}(x) = \sum_{j} x_{ij}x_{ji}$.
-We will refer to the very simple model with only these two parameters in the objective function as M1.
+We will refer to the very simple model with only these two parameters in the objective function as model M1.
+We define additional models of interest in Section~\ref{sec:models}
 
@@
 
-\section{Example Data and Models}\label{sec:senatedata}
+\section{Example Data and Models}\label{sec:models}
 

### 2019-02-28 ASApaper/paper-child.Rnw
surprised how close this is to being ready

@@
 Section~\ref{sec:saom} provides an introduction to the our models of interest, stochastic actor-oriented models.
-Section~\ref{sec:expersetup} details how we define significance testing and goodness of fit procedures for SAOMs through visual inference, and Section~\ref{sec:res} details the results of a visual inference survey of Amazon Mechanical Turk workers.
+Section~\ref{sec:expersetup} details how we define significance testing and goodness of fit procedures for CTMC models through visual inference, and Section~\ref{sec:res} details the results of a visual inference survey of Amazon Mechanical Turk workers.
 We close with a discussion in Section~\ref{sec:concl}.
@@
 
-\textbf{Rate Function:} All changes in SAOMs are treated as changes made by the nodes, or \textit{actors}, in the network.
-Each actor, $i$, gets a chance to make a change according to the rate function, which dictates when relationships between nodes in the network can change.
-In general, the rate function can take the network structure such as the outdegree of a node, and the node covariates into account, but we use a simple rate function, which is constant over all nodes in a given time period, $\alpha_m$ $\forall i \in \{1, 2, \dots, n\}$.
+\noindent \textbf{Rate Function:} All changes in the network variables $x_{ij}$ are considered to be made by the nodes, or \textit{actors}, in the network.
+Each actor, $i$, gets a chance to make a change to one of the other nodes $j$ according to the rate function, which dictates when relationships between nodes in the network can change.
+In general, the rate function can take the network structure, such as the outdegree of a node, and the node covariates into account, but we use a simple rate function, which is constant over all nodes in a given time period, $\alpha_m$ $\forall i \in \{1, 2, \dots, n\}$.
 We denote the rate from $t_{m}$ to $t_{m+1}$ as $\alpha_m$ for $m = 1, \dots, M-1$.
@@
 
-\textbf{Objective Function:} After actor $i$ has been given the opportunity to change, it randomly selects one of its current ties, $x_{ij}$, to change.
+\noindent \textbf{Objective Function:} After actor $i$ has been given the opportunity to change, it randomly selects one of its current ties, $x_{ij}$, to change.
 The probability that actor $i$ changes its current tie to actor $j$ is determined by the \textit{objective function} of the model and a random error term $U$ with log-Weibull distribution \citep{modelsSnijders}.
-.
 %\hh{XXX can you paraphrase the purpose of U in half a sentence?}
@@
 The Rorschach protocol helps to understand the extent of randomness in the null model.
-Here, we use the lineup protocol.
-Under this protocol, we begin with a data set of interest to us, such as a network, and a visualization of this data, such as a node-link diagram.
-We would like to know the model that generated this data.
+Here, we use the lineup protocol to design significance, goodness-of-fit (GoF) and power tests.
+Under the protocol, we begin with a data set of interest to us, such as a network, and a visualization of this data, such as a node-link diagram.
+Suppose we have a hypothesized model that we think generated this data.
 There are two hypothesis considered: a null hypothesis which states the model that is assumed to have generated the data, and an alternative hypothesis that the data were not generated under this model.
 Then, the plot of the data of interest is placed randomly among a set of $m-1$ null plots, which are visualizations of data simulated from the null model (where $m =20$, usually).
-Human observers are then asked to examine the lineup and to identify the plot(s) that look(s) most different from the others.
+Human observers are then asked to examine the lineup and to identify the plot(s) that look most different from the others.
 If an observer identifies the data plot, this is quantifiable evidence against the null hypothesis.
@@
 
-The lineup protocol places a plot firmly in the framework of hypothesis tests: a plot of the data is considered to be the test statistic, which is compared against the sampling distribution under the null hypothesis represented by the null plots.
+The lineup protocol places a \textit{plot} in the framework of hypothesis tests: the plot of the data is the test statistic, which is compared against the sampling distribution under the null hypothesis represented by the null plots.
 Obviously, the null generating mechanism, i.e.\ the method of obtaining the data for null plots, is crucial for the lineup protocol, as the null hypothesis directly affects the choice of null generating method.
-Null generating methods are typically based on (a) simulation, if the null hypothesis allows us to directly specify a parametric model, (b) sampling, as for example in the case of large data sets, or (c) permutation of the original data \citep[see e.g.\ ][]{Good05}, which allows for non-parametric testing that preserves marginal distributions  while ensuring independence in higher dimensions.
-The model of interest here allows us to simulate directly from a parametric model for dynamic social network data.
+Null generating methods are typically based on (a) simulation, if the null hypothesis allows us to directly specify a parametric model, (b) sampling, as in the case of large data sets, or (c) permutation of the original data \citep[see e.g.\ ][]{Good05}s.
+The network model of interest here allows us to simulate directly from a parametric model for dynamic social network data.
 %In the experimental data that we analyzed the null generating methods used were permutation methods and direct simulation from a null model.
@@
 The results suggest that  visual inference is comparable to conventional tests in a controlled conventional setting.
-This provides support for its appropriateness for testing in situations where no conventional test exists or is difficult.
+This provides support for its appropriateness for testing in situations where no conventional test exists or when testing is difficult.
 %Interestingly, the power of a visual test increases with the number of observers engaged to evaluate lineups, and the pattern in results suggests that the power will provide results consistent with practical significance \citep{kirk:1996}.
@@
 
-\noindent \textbf{Data:} The data we use are collaboration networks in the United States Senate during the $111^{th}$ through $114^{th}$ Congresses, overlapping with Barack Obama's presidency.
-These senates began on January 6, 2009, the start date of the $111^{th}$ Congress, and ended on January 3, 2017, the last date of the $114^{th}$ Congress\footnote{Details of how this data can be downloaded are provided by François Briatte at \url{https://github.com/briatte/congress}.}.
-In the US Senate, there are three ways that senators can show support for a piece of legislation: they can author the bill, cosponsor the bill, and vote for the bill.
+\noindent \textbf{Data:} The data we use are collaboration networks in the United States Senate during the $111^{th}$ through $114^{th}$ Congresses.
+These senates began on January 6, 2009, the start date of the $111^{th}$ Congress, and ended on January 3, 2017, the last date of the $114^{th}$ Congress (Details of how this data can be downloaded are provided by François Briatte at \url{github.com/briatte/congress}.).
+In the US Senate, there are three ways that senators can show support for a piece of legislation: they can author the bill, cosponsor the bill, or vote for the bill.
 We use cosponsorship as a metric because it results in a network that is unimodal (all nodes are senators) and directed.
@@
 This measure ranges in value from 0 to 1, where $WPC_{ij} = 1$ if senator $i$ is a cosponsor on every one of senator $j$'s bills and $WPC_{ij} = 0$ if senator $i$ is never a cosponsor any of senator $j$'s bills.
-Because SAOMs require binary edges, we construct the edges as follows:
+Because CTMC models require binary edges, we construct the edges as follows:
  \begin{equation}\label{eq:edgewpc}
@@
 For each of the four senate sessions, in addition to the WPC value between any two senators in the session, we have three node covariates: the party affiliation of each senator, the number of bills they authored in each session, and their gender.
-We explored each of these covariates in the model to determine if they affect the overall network structure and how ties are formed between senators.
+We explore each of these covariates in CTMC models to determine if they affect the overall network structure and how ties are formed between senators.
 The node-link diagram representations of the data we use for modelling are shown in Figure~\ref{fig:senateAll}.
@@
 
-<<senateAll, fig.cap="The four senate collaboration networks that we use as our example data to visually assess the SAOM effects.
+<<senateAll, fig.cap="The four senate collaboration networks that we use as our example data to visually assess the models of interest.
 Color represents party, shape represents gender, and size represents number of bills authored in a session.
 The Frucherman-Reingold layout is shown.
-">>=
+Drawn with the \\texttt{geomnet} R package.
+\\citep{geomnet, rstats}">>=
 seobama <- read_csv("data/senateobamapres_gsw_25.csv")
@@
 We chose to remove Clinton and her edges from the network because they make the overall structure look vastly more connected than the other three senates.
-We suspect that because she had just recently run for President and had just been selected as Obama's Secretary of State, the cosponsorships she garnered between the start of the new congress and becoming Secretary of State were largely symbolic, so the $111^{th}$ Senate without Hillary Clinton is more typical than the $111^{th}$ Senate with her.
+We suspect that because she had just recently run for President and had been selected by President-Elect Obama to be Secretary of State, the cosponsorships she garnered at the start of the new congress were largely symbolic, so the $111^{th}$ Senate without Hillary Clinton is more typical than the $111^{th}$ Senate with her.
 In fact, she had several co-sponsors on both of her pieces of legislation, which meant she had the maximum WPC value of one with about a dozen other senators.
@@
 In legislative cosponsorship networks, it is well known that party affiliation, reciprocity of relationships, and whether senators are ``workhorses" who author many bills or ``showhorses" who author few bills, are major influences on structure \citep{legnet}.
-We focus on these covariates, plus the additional gender covariate when choosing which SAO models to fit to the data.
+We focus on these covariates, plus the additional gender covariate when choosing which CTMC models to fit to the data.
 
@@
 \end{enumerate}
-
 The parameters in the objective function, $\beta_1, \dots, \beta_6$ are defined in Table~\ref{tab:effects}.
@@
 Other effects were added one at a time so that the simplest model $M1$ is nested in each of the other models.
-The most significant additional effect was $\beta_3$, the jumping transitive triplet (JTT) parameter for the party covariate, which was estimated to be about -6 with a standard error of 0.11, resulting in a $p$-value of less than 0.0001.
-This estimate of the parameter associated with this statistic considers the number of transitive closures formed between two senators from different parties.
-The negative estimate is an indication that forming transitive ties between two people from different parties is discouraged, which corresponds to the divisive nature of American politics, where party affilitation is the dominant effect.
-Another significant effect was $\beta_4$, the same JTT parameter for the sex covariate, with an estimate of about 3 with a standard error of 0.89.
-This parameter also consideres transitive closures, but for senators of different genders.
+The largest model is  M7: each of the other five models is nested in it.
+The most significant additional effect is $\beta_3$, the jumping transitive triplet (JTT) parameter for the party covariate, which was estimated to be about -6 with a standard error of 0.11, resulting in a Wald $p$-value of less than 0.0001.
+This parameter associated considers the number of transitive closures formed between two senators from different parties.
+The large negative estimate is an indication that forming transitive ties between two people from different parties is strongly discouraged, which corresponds to the divisive nature of American politics, where party affilitation is the dominant effect.
+Another significant effect is $\beta_4$, the same JTT parameter for the sex covariate, with an estimate of about 3 with a standard error of 0.89.
+This parameter also considers transitive closures, but for senators of different genders.
 The positive value indicates that transitive ties between senators of different genders are more likley to form.
-Another significant effect was $\beta_6$, the covariate-related similarity score-weighted transitive triplets parameter estimate for the number of bills authored by a senator.
+Next, we consider $\beta_6$, the covariate-related similarity score-weighted transitive triplets parameter estimate for the number of bills authored by a senator.
 We chose to look at similarity instead of covariate "jumps" because the number of bills authored is more continuous than gender or party.
@@
 The high positive estimate suggests senators are encouraged to collaborate with other senators who author about the same number of bills they do.
-This tendency of senators to cosponsor bills written by senators who are similarly ``prolific" corresponds to another well-known property of the U.S.
-Senate structure: the tendency of senators to be either ``workhorses" or ``showhorses".
+This tendency of senators to cosponsor bills written by senators who are similarly ``prolific" corresponds to the tendency of senators to be either ``workhorses" or ``showhorses".
 Senators known as workhorses author many pieces of legislation in a session, and largely stay out of the public arena.
 The showhorse senators, on the other hand, author relatively few pieces of legislation, and tend to appear on television, radio, and other media a great deal.
-So, showhorses and workhorses tend to stick together.
-Finally, we found $\beta_5$, the same party transitive triplet effect was also significant, with a fitted value of 1.3 and standard error of 0.7, meaning that transitive relationships between senators tend to form when they are from the same party, exactly as we would expect in a legislative body in a country with extremely entrenched partisan divides as in the US.
+Finally, we found $\beta_5$, the same party transitive triplet effect to be significant, with a fitted value of 1.3 and standard error of 0.7, meaning that transitive relationships between senators tend to form when they are from the same party, exactly as we would expect in a legislative body in a country with extremely entrenched partisan divides as in the US.
 
@@
 \hline 
-$\beta_1$ & density & NA & $s_{i1}(x) = \sum_{j} x_{ij}$ & \includegraphics[width=.6in]{img/density.png} & 2.204 & NA \\ 
-$\beta_2$ & reciprocity & NA & $s_{i2}(x) = \sum_{j} x_{ij}x_{ji}$ & \includegraphics[width=.6in]{img/recip.png} & -4.903 & NA \\
+$\beta_1$ & density & -- & $s_{i1}(x) = \sum_{j} x_{ij}$ & \includegraphics[width=.6in]{img/density.png} & 2.204 & NA \\ 
+$\beta_2$ & reciprocity & -- & $s_{i2}(x) = \sum_{j} x_{ij}x_{ji}$ & \includegraphics[width=.6in]{img/recip.png} & -4.903 & NA \\
 $\beta_3$ & jumping transitive triplet & party & $s_{i3}(x, \mathbf{p}) = \sum_{j\neq h} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(p_i = p_h \neq p_j)$ & \includegraphics[width=.6in]{img/jttp.png} & -5.884 & $<0.0001$\\
@@
 }
-\caption{\label{tab:effects} The effects we used in the SAOMs fit to the senate data.
+\caption{\label{tab:effects} The effects we used in the CTMC models fit to the senate data.
 In the picture which represents each effect, the dotted tie is encouraged to form if the estimate is positive, and discouraged to form if the estimate is negative.
@@
 
-We want to explore three different aspects of the SAOM models using the lineup protocol: (1) significance of parameters, (2) goodness-of-fit of a model, and (3) \st{visual power of the effects.} %visual detection of parameters.
-Each one of these situations requires a different setup, which we describe in detail, making use of the lineup protocol to study all of these aspects.
-
-In each lineup, we include plots from two models: the null model and an alternative model. The definition of the null and alternative model varies with the aspect of the SAOMs we are exploring. 
+We want to explore three different assessments of the CTMC models using the lineup protocol: (1) significance of parameters, (2) goodness-of-fit of a model, and (3) visual power of the effects.
+%visual detection of parameters.
+Each one of these situations requires a different setup, which we describe in detail, making use of the lineup protocol defined in \citet{Bujaetal} to study all of these aspects.
+In each lineup, we include plots from two models: the null model and an alternative model.
+The definition of the null and alternative model varies with the models we are exploring.
 %For each lineup, we consider two models: model M1 is the null model, and another model, from M2-M5, which is chosen to be the alternative model.
@@
 Sections~\ref{sec:setupst} through~\ref{sec:setupvp} describe in detail how we set up the lineups, which parameter values we simulate from, and why.
-Sets of the lineups we create are shown to independent observers recruited through Amazon Mechanical Turk for feedback (more details on the Turk setup in Section~\ref{sec:res}).
+Lineups we created were shown to independent observers recruited through Amazon Mechanical Turk for feedback (more details on the Turk setup in Section~\ref{sec:res}).
 %, we set the parameters to the values given in Table~\ref{tab:fittedvalues} for all parameters within the respective models, with the exception of $\beta_5$.
@@
 \includegraphics[width=.9\textwidth]{img/shinyappscreen.png}
-\caption{\label{fig:shiny} A screen shot of the web application we created to design our lineup experiment.
+\caption{\label{fig:shiny} A screen shot of the web application we created to help design our lineup experiment.
 More details about this application are given in Section~\ref{sec:setupvp}.
@@
 One plot simulated from this model is placed at random among five observations simulated from the null model, M1.
-Participants of the study are asked to identify the most different plot.
+%Participants of the study are asked to identify the most different plot.
 %The other two tabs in the right side of the screen tell you which plot comes from the alternative model, and allow you to view and download the data that is plotted in the lineup.
@@
 %\end{description}
-
 Under the null hypothesis, we assume that the model that generated the network data is $M_1$, the simplest model presented in Section~\ref{sec:models}.
@@
 The alternative model is the model with $\beta_1, \beta_2$, and $\beta_k$ in the objective function.
-The alternative data plot is simulated from the appropriate model also with values set to the estimates from Table~\ref{tab:fittedvalues}.
+The alternative data plot is simulated from the appropriate model with values set to the estimates from Table~\ref{tab:fittedvalues}.
 
@@
 %Visual detection does not fit into the typical statistical testing framework as easily as the significance and goodness-of-fit tests we perform do. 
-Through visual inference, we want to determine at which point an effect becomes noticeable in a SAOM.
-By \textit{noticeable}, we mean that the inclusion of the effect alters the appearance of networks simulated from a model to a degree that viewers are able to reliably pick out a node-link diagram rendered from data simulated from a model \textit{with} the effect in a lineup among plots \textit{without} the effect.
+Through visual inference, we want to determine at which point an additional effect included in the model becomes noticeable.
+By \textit{noticeable}, we mean that the inclusion of the effect alters the appearance of networks simulated from the model so much that viewers are able to reliably pick out the one node-link diagram rendered from data simulated from the model \textit{with} the effect in a lineup among plots simulated from the model \textit{without} the effect.
 This is a way to determine the power of the visual significance test.
@@
 
-In model M1, with only two parameters in the objective function, we varied both the density and reciprocity parameter values one at a time, \st{keeping all other parameters at their fitted values given in Table~\ref{tab:fittedvalues}}.
+In model M1, with only two parameters in the objective function, we varied both the density and reciprocity parameter values one at a time, keeping all other parameters at their fitted values given in Table~\ref{tab:fittedvalues}.
 In models M3 through M6, we vary the additional parameter, $\beta_3$ through $\beta_6$, respectively, while again holding the rate and other parameters at their value in Table~\ref{tab:fittedvalues}.
 Thus, we have six different parameters of interest to us in total: $\beta_1, \dots, \beta_6$.
-We want to determine how the size of these parameters affects the overall structure of the network data simulated from the models M1 through M6, so we also vary the value of the parameters in order to determine at what level the effects become noticeable.
+We want to determine how the size of these parameters affects the overall structure of the network data simulated from the models M1 through M6, so we also vary the magnitude of the parameters in order to determine at what level the effects become noticeable.
 %in both negative and positive directions.
@@
 
-<<hypothesis, fig.height=2, fig.cap="We hypothesize that as the parameter value of interest increases in absolute value, more viewers of the lineup will pick the alternative data out of a lineup.
+<<hypothesis, fig.height=2, fig.cap="As the parameter increases in absolute value, more viewers of the lineup should pick the alternative data out of a lineup.
 Note that the significance test we construct in Section~\\ref{sec:setupst} is just one point on the line below, represented by the vertical dotted line labeled $\\hat{\\beta}$.
@@
 To determine the exact values of the six levels we want to test for each effect, we started with the estimates of the parameter at hand (see  Section~\ref{sec:models}), and used small negative and positive factors to determine at what point \textit{we} noticed the effect of the parameter in simulations from the changed models.
-\st{In Figure~\ref{fig:hypothesis}, we demonstrate these values with the vertical lines labeled ``easy," ``medium," and ``hard." We expected most viewers to see the effects at the ``easy" values, and we expected very few, if any, viewers to see the effects at the ``hard" values.
-}
-
-\st{To decide on the values to use for each difficulty level,} we constructed an online application that created the lineup protocol for us to be the guinea pigs of our own experiment \citep{theqs}.
+In Figure~\ref{fig:hypothesis}, we demonstrate these values with the vertical lines labeled ``easy," ``medium," and ``hard." We expect most viewers to see the effects at the ``easy" values, and we expect very few, if any, viewers to see the effects at the ``hard" values.
+
+To decide on the values to use for each difficulty level, we constructed an online application that simulated the lineup protocol for us to be the guinea pigs of our own experiment \citep{theqs}.
 A screen shot of the app we created with the \texttt{shiny} package by \citet{shiny} is shown in Figure~\ref{fig:shiny}.
-On the left side of the screen, the user\footnote{Please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/} to create lineups constructed from the models we present for this data for yourself.} can input the information necessary for creating a lineup of the models M1 through M7 for the data in Section~\ref{sec:senatedata}: first, choose to simulate only one plot from the specified model (analogous to changing the alternative model in the lineup protocol) or to simulate $M-1$ plots from the specified model (analogous to changing the null model in the lineup protocol); the model of interest; the wave of the data to examine; if model M1 is selected, whether to alter the density or the reciprocity parameter; the size of the lineup; the amount by which to multiply the effect selected; a random seed for replicability; and a layout algorithm to use for the node-link diagrams.
+On the left side of the screen, the user\footnote{Please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/} to create lineups constructed from the models we present for this data for yourself.} can input the information necessary for creating a lineup of the models M1 through M7 for the data in Section~\ref{sec:senatedata}: first, choose to simulate only one plot from the specified model (analogous to changing the alternative model in the lineup protocol) or to simulate $M-1$ plots from the specified model (analogous to changing the null model in the lineup protocol); then choose the model of interest and the wave of the data to examine.
+If model M1 is selected as the model of interest, select whether to alter the density or the reciprocity parameter.
+Then, pick the size of the lineup; the amount by which to multiply the effect selected; a random seed for replicability; and a layout algorithm to use for the node-link diagrams.
 There is also a checkbox if the user wishes the nodes to be colored by the size of the connected component to which they belong.
@@
 In the case of both $\beta_4$ and $\beta_6$, we could not determine any values for negative effects that made the data simulated from M4 and M6 look different than null model simulations from model M1.
-\st{We hypothesized that this was due to the negative effects \textit{removing} visually interesting structural elements as opposed to \textit{adding} noticeable structural elements.}  Therefore we decided that the lesser experienced participants in our experiment would also not be able to.
+We hypothesized that this was due to the negative effects \textit{removing} visually interesting structural elements as opposed to \textit{adding} noticeable structural elements.
+Therefore we decided that the lesser experienced participants in our experiment would also not be able to.
 Instead of testing the negative values of these effects, we are examining a different scenario: we placed 5 simulations from model M4 or M6 with positive values of the parameter with one simulation from model M1 in a lineup.
@@
 Each participant was presented with some brief training material before beginning the experiment.
-After agreeing to participate, the participants were shown two trial plots, one where the data plot was the most different from the others due to its relatively complex structure, while the other trial included a data plot that was most different from the others due to its comparatively simple structure.
+After agreeing to participate, each person was shown two trial plots, one where the data plot was the most different from the others due to its relatively \textit{complex} structure, while the other trial included a data plot that was most different from the others due to its comparatively \textit{simple} structure.
 Only when participants were able to correctly identify the data plot from the trial lineups were they allowed to begin the experiment.
 Each participant was randomly assigned 13 lineups to look at.
-The were asked to select one or more plots that they perceived as ``most different" from the others, and provide a reasoning for their choice.
-They could select from ``Most simple overall structure" or ``Most complex overall structure", \st{corresponding to what they saw in the examples and trials, or they could choose }``Other" and provide their own text description of their reasoning.
-\st{These language in these reasons is purposefully vague: we want participants to tell us what they see.
+The were asked to select one or more plots that they perceived as ``most different" from the others, and provide a reasoning for their choice: ``Most simple overall structure",``Most complex overall structure", or ``Other".
+If they selected other, they were required to describe their own text description of their reasoning.
+The language in these reasons is purposefully vague: we want participants to tell us what they see.
 We do not want to tell them what they ``should" be seeing.
-In this way, we can truly determine what  effects are \textit{noticeable}.}
+In this way, we can truly determine what  effects are \textit{noticeable}.
 
@@
 Each participant only saw the data one time in order to avoid bias.
-Upon completion of the 13 lineups, each participant was paid \$1.75.
+Upon completion of the 13 lineups, which took an average of 308 seconds per person, each participant was paid \$1.75
 
@@
 
-For a SAOM, there are two ways a conventional significance test of the parameters can be performed.
+For a CTMC model, there are two ways a conventional significance test of the parameters can be performed.
 In \texttt{RSiena}, there are $t$-type and Wald-type tests for a single parameter and for multiple parameters.
@@
 The correspoding results from the significance tests we performed using the lineup protocol are given in Table~\ref{tab:sigtesting}.
-\st{Corresponding to traditional methods, if enough participants pick out the alternative plot to result in a $p$-value less than 0.05, we reject the null hypothesis that the true value of the additional parameter, either $\beta_3$ or $\beta_4$, is equal to zero.} %\hh{XXX how do we interpret these  visual results?
+If enough participants pick out the alternative plot to result in a $p$-value less than 0.05, we reject the null hypothesis that the true value of the additional parameter, either $\beta_3$ or $\beta_4$, is equal to zero.
+%\hh{XXX how do we interpret these  visual results?
 Include the six lineups in an appendix and refer to them from here.} \st{XXX should be taken care of in next paragraph now.}
@@
 
-\st{The $p$-values were calculated using the \texttt{vinference} package by \citet{vinference}.
+The $p$-values were calculated using the \texttt{vinference} package by \citet{vinference}.
 This package contains methods to calculate \textit{Visual distributions} for lineup experiment data.
@@
 The visual inference family of distributions is similar to the binomial distribution, but takes the dependency among the $m$ plots in a single lineup shown to multiple viewers into account.
-Using these $p$-values, all but one lineup results in a rejection of the null hypothesis at Type-I error rate of $\alpha = 0.05$.}.
+Using these $p$-values, all but one lineup results in a rejection of the null hypothesis at Type-I error rate of $\alpha = 0.05$.
 We see that the $p$-values for visual inference here are highly variable.
@@
 
-\st{The lineup for significance testing of $\beta_3$ which resulted in a very small $p$-value and rejection of the null hypothesis is shown in Figure~\ref{fig:lineup3132}.
+The lineup for significance testing of $\beta_3$ which resulted in a very small $p$-value and rejection of the null hypothesis is shown in Figure~\ref{fig:lineup3132}.
 Another significance lineup for model M3, which resulted in failure to reject the null hypothesis, is shown in Figure~\ref{fig:lineup3133}.
@@
 In viewing these two lineups, it is evident that there is a large amount of variability.
-It is difficult to see that five of the six come from the same model when they can all look different in their own way.} Thus, the variability in results is introduced through the null plots generated from M1, as not all simulations look alike.
+It is difficult to see that five of the six come from the same model when they can all look different in their own way.
+Thus, the variability in results is introduced through the null plots generated from M1, as not all simulations look alike.
 In addition, the necessarily small number of null plots do not give the viewer as complete of a view of the null model as the usual 19 null plots would.
-The results of the significance tests given in Table~\ref{tab:sigtesting} for $\beta_3$ and $\beta_4$ are not \st{definitive.
+The results of the significance tests given in Table~\ref{tab:sigtesting} for $\beta_3$ and $\beta_4$ are not definitive.
 For the test of $\beta_3$, two of the three tests are not significant, while the third is highly significant.
-For the test of $\beta_4$, one test is significant, one is decidedly not significant, and the third is significant at the level of 0.10.}  Thus, unlike the Wald-type tests described at the beginning of this section, there is no way to \st{decisively} reject or to fail to reject the null hypothesis that the parameter value is 0.
+For the test of $\beta_4$, one test is significant, one is decidedly not significant, and the third is significant at the level of 0.10.
+Thus, unlike the Wald-type tests described at the beginning of this section, there is no way to decisively reject or to fail to reject the null hypothesis that the parameter value is 0.
 We include all of the lineups shown to our participants in the appendix.
@@
 \begin{enumerate}
-\item Auxiliary statistics, \st{such as the cumulative outdegree counts on the nodes,} are computed on the observed data ($\mathbf{u}_d$) and on $N$ simulated observations from the model ($\mathbf{u}_1 \dots \mathbf{u}_N$).
+\item Auxiliary statistics, such as the cumulative outdegree counts on the nodes, are computed on the observed data ($\mathbf{u}_d$) and on $N$ simulated observations from the model ($\mathbf{u}_1 \dots \mathbf{u}_N$).
 (Usually, $N=1000$)
@@
 \item An empirical $p$-value is found by computing the proportion of simulated distances found in step 4 that are as large or larger than $d_M(\mathbf{u}_d)$.
-A SAOM is thus considered a good fit to the data if $p$ is large.
-A plot comparing the data to the simulations is also considered, and a similar plot is shown in Figure~\ref{fig:gofsiena} \st{for the outdegree distribution of small data set, shown in the points and connected lines, with the simulated values of $\mathbf{u}_d$ shown in boxplots and overlaid violin plots.}
+A model is thus considered a good fit to the data if $p$ is large.
+A plot comparing the data to the simulations is also considered, and a similar plot is shown in Figure~\ref{fig:gofsiena} for the outdegree distribution of small data set, shown in the points and connected lines, with the simulated values of $\mathbf{u}_d$ shown in boxplots and overlaid violin plots.
 \end{enumerate}
@@
 
-\st{The $p$-values were again computed using the \texttt{vinference} package by \citet{vinference}}.
+The $p$-values were again computed using the \texttt{vinference} package by \citet{vinference}.
 The lineup that resulted in a failure to reject the null hypothesis is shown in Figure~\ref{fig:failgof}.
@@
 However, the panel most participants chose was number four, and the most common reasoning for that choice was that it had the most simple structure.
-\st{Some of the other panels, such as three and six, in Figure~\ref{fig:failgof} have large connected components that are similar in size to the connected component of the data plot shown in panel two.
-Thus, model M5 is sometimes capable of capturing the network structure of the senate collaboration data.}
+Some of the other panels, such as three and six, in Figure~\ref{fig:failgof} have large connected components that are similar in size to the connected component of the data plot shown in panel two.
+Thus, model M5 is sometimes capable of capturing the network structure of the senate collaboration data.
 
@@
 The smallest $p$-value for one of the goodness-of-fit lineups was for the third replicate of the null model M5.
-This \st{result contrasts with} our previous finding that the only lineup to fail to reject the null was also when the null model was M5.
+This result contrasts with our previous finding that the only lineup to fail to reject the null was also when the null model was M5.
 This lineup is shown in Figure~\ref{fig:mostsiggof}.
@@
 For the other models for which we tested goodness-of-fit, however, we \textit{do} have significant evidence from all three replicates to reject the null hypothesis that the null model generated the data.
-\st{For models M3, M5, and M7, these goodness-of-fit tests have rejected the null hypotheses that the senate data come from these models.
+For models M3, M5, and M7, these goodness-of-fit tests have rejected the null hypotheses that the senate data come from these models.
 We hypothesized that the model with the most effects, M7, would be the best fit.
-However, as shown in Figure~\ref{fig:gofm71}, the model does not capture the overall structure very well at all.} The rest of the goodness-of-fit lineups as shown to participants are provided in the appendix.
+However, as shown in Figure~\ref{fig:gofm71}, the model does not capture the overall structure very well at all.
+The rest of the goodness-of-fit lineups as shown to participants are provided in the appendix.
 
@@
 
-A summary of the results from our experiment is shown \st{as points }in Figure~\ref{fig:predictglmm}.
+A summary of the results from our experiment is shown as points in Figure~\ref{fig:predictglmm}.
 On the $x$ axis, we plot the value of the parameter of interest, and on the $y$ axis, the proportion of times the alternative data plot was picked out for each lineup.
 The results are split into groups based on the value of the parameter and the lineup type.
-We can see clear patterns \st{in the added parameters $\beta_3, \dots, \beta_6$ }: as the parameter value approaches 0, fewer participants identified the alternative plot.
+We can see clear patterns in the added parameters $\beta_3, \dots, \beta_6$: as the parameter value approaches 0, fewer participants identified the alternative plot.
 Similarly, as $\beta_1, \beta_2$ approach their estimated values $\hat{\beta_1}, \hat{\beta_2}$, fewer people are able to identify the alternative plot.
@@
 
-We further explore this relationship between identification of the alternative data in the lineup and the parameter \st{of interest}, effect size, and lineup type with a generalized linear mixed model that provides us with an estimate of the power of the visual \st{significance } test.
+We further explore this relationship between identification of the alternative data in the lineup and the parameter of interest, effect size, and lineup type with a generalized linear mixed model that provides us with an estimate of the power of the visual significance test.
 The response variable, $Y_{ijkm}$, is binary, indicating whether participant $m$ picked the alternative data plot in lineup type $j$, rep $k$, for effect $i$.
@@
 
-\st{We expand portions of Figure~\ref{fig:predictglmm} in Figures~\ref{fig:recipzoom}-\ref{fig:beta4zoom}.
+We expand portions of Figure~\ref{fig:predictglmm} in Figures~\ref{fig:recipzoom}-\ref{fig:beta4zoom}.
 These figures show the same prediction regions as in Figure~\ref{fig:predictglmm}, plus some additional predictions outside of the data range shown in gray.
@@
 This again demonstrates that the presence of structure is detected sooner and more frequently than the absence of strucutre.
-}
 
@@
 For identical values of the parameter, viewers had a harder time identifying the different plot when they were selected the most ``simple" structure, detecting M1 in five plots from the more complicated model, than they did identifying the most ``complex" structure, the plot from the more complicated model, from the five plots from M1.
-\st{This result is also similar to that of \citet{corrviz} because it emphasizes the difficulty of picking out the absence of an effect relative to picking out the presence of an effect.}
+This result is also similar to that of \citet{corrviz} because it emphasizes the difficulty of picking out the absence of an effect relative to picking out the presence of an effect.
 
@@
 By using visual inference methods, we have developed new ways to perform significance and goodness-of-fit testing for a complicated and intractable set of statistical models for social network data.
-\st{We have also developed a way to determine the power of these new visual tests.} Our methods can be used to supplement traditional methods and check our assumptions about network models.
+We have also developed a way to determine the power of these new visual tests.
+Our methods can be used to supplement traditional methods and check our assumptions about network models.
 The traditional methods only look at one piece or derived measure of a network model, whereas our methods look at the models holistically for a broader sense of what it means for a parameter to be significant or a model to be a good fit.
-By looking at an entire network simulated from a SAOM side-by-side with other instances of networks simulated from another model, instead of singular %one-dimensional derived
+By looking at an entire network simulated from a CTMC model side-by-side with other instances of networks simulated from another model, instead of singular %one-dimensional derived
 features, we develop an idea of the model in terms of the \textit{data} itself, instead of in terms of statistical summaries of the data.
-\st{These methods place the model in the data space, instead of summarizing or compressing the data to place it in the model space.}
-
-Furthermore, we have found the visual power of some effects in the object function of a SAOM for this particular senate data example,  and we have shown that, for the same effects, there is a lot of variability in results from significance and goodness of fit tests.
-Because the visual tests we performed show a great deal of variability, we can see that the decisions with respect to the significance of a parameter or the goodness of fit of a model to data are not as cut-and-dried as the \st{more} traditional methods would have us believe.
+These methods place the model in the data space, instead of summarizing or compressing the data to place it in the model space.
+
+Furthermore, we have found the visual power of some effects in the object function of a CTMC model for this particular senate data example,  and we have shown that, for the same effects, there is a lot of variability in results from significance and goodness of fit tests.
+Because the visual tests we performed show a great deal of variability, we can see that the decisions with respect to the significance of a parameter or the goodness of fit of a model to data are not as cut-and-dried as the more traditional methods would have us believe.
 
@@
 In visual inference, the null plots are supposed to play the role of good representatives of the null model.
-Here, the number of null plots is reduced to five, which increases the variability seen in a single lineup dramatically, and can unfortunately lead to very different conclusions for the same lineup scenario.
-Furthermore, these results do not generalize to all SAOMs or to some subset of SAOMs.
-The lineups shown are made for only one set of data, and it is not clear whether the power results transfer, nor is it clear to what degree if they do transfer, to other situations with different number of actors, different edge densities, or different layout algorithm of the node-link diagram.
-\st{We can make some generalizations about what participants are picking up on in the lineups based on their feedback and previous research, but we cannot apply our hierarchical model directly to lineups constructed for new data or new models or parameters.}
+Here, the number of null plots is reduced to five, which increases the variability seen in a single lineup dramatically, and can unfortunately lead to very different conclusions for the same lineup scenario, as we have seen.
+Furthermore, these results do not generalize to all CTMC models or to one particular subset of CTMC models.
+The lineups shown are made for only one set of data, and it is not clear that the visual power results will transfer to other situations with different number of actors, different edge densities, or different layout algorithm of the node-link diagram.
+We can make some generalizations about what participants are picking up on in the lineups based on their feedback and previous research, but we cannot apply our hierarchical model directly to lineups constructed for new data or new models or parameters.
 
@@
 
-We hope to apply these methods further for different types of network data and different types of network models.
-We accept the limitations of this type of network data visualization, in that even in small instances, the cognitive load of looking at a lineup is very high for the average observer.
+We these methods will be applied to different types of network data and different types of network models.
+But given the limitations of the node-link visualization, the cognitive load of looking at a lineup is very high for the average observer.
 We would therefore like to explore larger datasets, different layout algorithms, and different ways of visualizing network data, such adjacency matrix visualizations, using visual inference to see if similar patterns emerge.

### 2019-03-05 ASApaper/paper-child.Rnw
big push; almost done

@@
 With significance testing, the null hypothesis assumes that the data come from a simple model nested within the model of interest.
-We then conduct significance tests of one or more additional parameters to determine if they significantly explain the variability in the data.
+We then conduct significance tests of one or more additional parameters to determine if they can explain some of the variability in the data.
 For goodness-of-fit tests, we examine one or more models of interest to assess how well they fit the data.
-Finally, power tests quantify the ability of the hypothesis test to detect the difference between the null and alternative hypothesis.
-All three of these aspects of statistical modeling are vital for inference.
+Power then quantifies the ability of the hypothesis test to detect the difference between the null and alternative hypothesis.
+All three of these elements of statistical modeling are vital for inference.
 %The more complicated the model, the harder it is to determine fit or to include or exclude a parameter.
@@
 The more complex the model, however, the harder it is to assess significance, power, and goodness-of-fit.
-One particularly complicated family of models are those designed to model network change.
+One particularly complicated family of models are those designed to model networks.
 A \textit{network} is any set of things, such as people, or computers, that are connected in some way, through social relations or the Internet.
-In a social network, the people are the \textit{nodes}, or \textit{actors}, and the relationships between the people are \textit{edges}, or \textit{ties}.
+In a social network, the people are the \textit{nodes} or \textit{actors}, and the relationships between the people are \textit{edges} or \textit{ties}.
+In this paper we use nodes and edges interchangeably with actors and ties, respectively.
 Models for networks are particularly complex, as
@@
 This difficulty only increases when studying \textit{dynamic networks}, the same set of nodes and their changing relationships observed at many points in time, because of the added temporal dependence.
-Yet, researchers are often interested in modeling dynamic social networks, such as friendship networks among students or collaboration networks between researchers.
-% X ST a bit less jarring hopefully.
-X
+Yet, researchers are often interested in modeling dynamic social networks, such as friendship networks among students or collaboration networks between legislators.
+%
 Even some of the simplest network models, however, lack the asymptotics required to perform traditional goodness-of-fit tests \citep{holland81}.
-In additional, direct maximum likelihood estimation of model parameters is frequently impossible due to the intractability of the models \citep{hummeletal}.
+In addition, direct maximum likelihood estimation of model parameters is frequently impossible due to the intractability of the models \citep{hummeletal}.
 %\hh{is the previous sentence a quote? I'm not sure that I follow.} %\st{it's a paraphrase... shall I make a quote? Also I definitely botched a word. Hopefully makes more sense now? }
 % stop 4pm on 1/28/19
+
 In order to circumvent some of these difficulties, we propose new methods for significance testing of parameters, goodness-of-fit testing, and power calculations %of these 
-for one family social network models: continuous time Markov-chain (CTMC) models for dynamic network data, as defined in \citet{saompaper}.
-Specifically, we are using \textit{visual inference} in place of traditional statistical methods for social network models, such as Wald or $t$-tests for significance of parameters and in- and outdegree distribution metrics for determining goodness-of-fit.
-Visual inference (VI), introduced by \citet{Bujaetal}, will allow us to look at the \textit{entire} dataset simulated from a network model as opposed to a (set of) usually one-dimensional metric(s) derived from the network, or a $p$-value for a parameter in the model.
-By using visual inference to supplement traditional statistical tests, we gain a new understanding of the parameters in these CTMC models and ability to assess the fit of the CTMCs to dynamic network data.%In addition, we use visual inference to better understand some parameters in a social network model, by looking at how different values for parameters affect the networks simulated from these models.
-
-The paper is outlined as follows: Section~\ref{vi} gives a basic overview of visual inference and the lineup protocol.
-Section~\ref{sec:saom} provides an introduction to the our models of interest, stochastic actor-oriented models.
-Section~\ref{sec:expersetup} details how we define significance testing and goodness of fit procedures for CTMC models through visual inference, and Section~\ref{sec:res} details the results of a visual inference survey of Amazon Mechanical Turk workers.
+for one family social network models: continuous time Markov-chain (CTMC) models for dynamic network data, as described in \citet{saompaper}.
+We use \textit{visual inference} in place of traditional statistical methods for social network models, such as $t$-tests for significance of parameters or outdegree distributions for goodness-of-fit tests.
+Visual inference (VI), introduced by \citet{Bujaetal}, will allow us to look at the \textit{entire} dataset simulated from a network model, whereas traditional methods use one-dimensional metrics derived from the network or a $p$-value for a parameter in the model.
+By using VI to supplement traditional statistical tests, we gain insight into the role of the parameters in these CTMC models, and we gain the ability to assess the fit of the CTMCs to dynamic network data.%In addition, we use visual inference to better understand some parameters in a social network model, by looking at how different values for parameters affect the networks simulated from these models.
+
+The paper is outlined as follows: Section~\ref{subsec:saom} provides an introduction to the CTMC familty of models.
+Section~\ref{subsec:vi} gives a basic overview of visual inference and the lineup protocol.
+In Section~\ref{sec:models}, we describe our example data and define our models of interst that we used to develop our VI methods.
+Section~\ref{sec:expersetup} details how we designed a VI experiment for significance testing and goodness of fit procedures for CTMC models, and Section~\ref{sec:res} details the results of our experiment.
 We close with a discussion in Section~\ref{sec:concl}.
@@
 %To model family that we apply to the Senate data is the Stochastic Actor-Oriented Model (SAOM) family. 
-CTMCs are a family of models for dynamic network data that incorporate both network structure and node-level information to describe how a network observed on two or more occasions changes over time \citep{saompaper}.
-The CTMCs model two key properties: social networks are ever-changing as relationships decay or grow in seemingly random ways, and the actors in them have characteristics that could affect how they change their ties to other nodes in the network.
+We define CTMC models for dynamic social network data here in their barest form.
+Full details on these models can be found in \citet{saompaper, snijders01, snijders:2010, snijders2010, snijdersetall:2007, snijders:2017}.
+These CTMCs are a family of models for dynamic network data that incorporate both network structure and node-level covariates to describe how a network changes over time \citep{saompaper}.
+Traditional network models, such as exponential random graph models, usually only consider network structure.
+Social networks are ever-changing as relationships decay or grow in seemingly random ways, and the actors in them have characteristics that affect how they change their ties to other nodes in the network.
 % Talk about SAOMs
-These unique properties allow for the fitting of some very complicated models to inherently complex data, so it can be exceedingly difficult to interpret parameters and their corresponding estimates.
-The large amount of available parameters to include in the model combined with the difficulty of interpretation make parameter selection and goodness-of-fit testing burdensome as well.
-
-Broadly, a CTMC model takes network structure and node covariate information into account in two ways and models the network changes one-at-a-time.
-First, the rate of change between states is dictated by a rate function that describes \textit{how often} changes in the network occur, and second, the objective function describes \textit{what} those state changes are.
-As in most other network models, the variables of interest are the edges of the network, $x_{ij}$, where $x_{ij}$ denotes the edge between nodes $i$ and $j$, where $i \neq j \in \{1, 2, \dots, n=\text{the number of nodes}\}$.
-$x_{ij}$ is binary, i.e.
+CTMC models use the network structure and the node covariate information, and as such can be very complicated models.
+As network data are inherently complex, it can be difficult to interpret model parameters and their estimates.
+There are also many possible parameters to include in a CTMC, which makes parameter selection and goodness-of-fit testing challenging.
+
+CTMC models use network structure and node covariates to model the network change one tie at a time.
+There are two model pieces: a \textit{rate function} that dictates \textit{how often} changes in the network occur and an \textit{objective function} that determines \textit{what} those changes are.
+The variables of interest are the edges of the network, $x_{ij}$ between nodes $i$ and $j$, where $i \neq j \in \{1, 2, \dots, n\}$ and $n$ the number of nodes in the network.
+The $x_{ij}$ are directed and binary, i.e.
   \begin{equation}\label{eq:edgevars}
@@
   \end{equation}
-Edges are treated as \textit{directed}, so that in general $x_{ij} \neq x_{ji}$, and self-referencing edges or loops are not allowed, i.e.  $x_{ii} = 0$ for all $i=1, 2, \dots, n$. 
-The network is observed $M \geq 2$ times at time points  $t_1 < ...
-< t_M$, and the entire network at time point $t_m$ is denoted as $x(t_m)$ for $m \in \{1, \dots, M\}$.
-We discuss the rate and objective functions of a CTMC model in more depth below.
-Additional details on these models can be found in \citet{saompaper, snijders01, snijders:2010, snijders2010, snijdersetall:2007, snijders:2017},
-
-\noindent \textbf{Rate Function:} All changes in the network variables $x_{ij}$ are considered to be made by the nodes, or \textit{actors}, in the network.
-Each actor, $i$, gets a chance to make a change to one of the other nodes $j$ according to the rate function, which dictates when relationships between nodes in the network can change.
-In general, the rate function can take the network structure, such as the outdegree of a node, and the node covariates into account, but we use a simple rate function, which is constant over all nodes in a given time period, $\alpha_m$ $\forall i \in \{1, 2, \dots, n\}$.
+and self-referencing edges or loops are not allowed, i.e.  $x_{ii} = 0$ for all $i=1, 2, \dots, n$. 
+The network is observed $M \geq 2$ times at time points  $t_1 < ... < t_M$, and the entire network at time point $t_m$ is denoted as $x(t_m)$ for $m \in \{1, \dots, M\}$.
+
+\noindent \textbf{Rate Function:} All changes in the network variables $x_{ij}$ are treated as choices made by the actors in the network.
+The rate function dictates when changes are made and which actor can make them.
+When chosen, the actor, $i$, gets a chance to make a change to one of the other nodes $j$.
+In general, the rate function can include structural and node covariate parameters into account so that each actor has a different rate of change.
+However, we choose a simple rate function that is constant over all nodes in a given time period, because we want to focus on interpreting the parameters of the objective function.
 We denote the rate from $t_{m}$ to $t_{m+1}$ as $\alpha_m$ for $m = 1, \dots, M-1$.
 Using this notation, the \textit{waiting time} to the next chance for actor $i$ to make a change is exponentially distributed with expected value $\alpha_m^{-1}$.
-Since the rate is the same for all actors, the waiting time for \textit{any} actor to get the oppurtunity to change its set of ties is exponentially distributed with expected value $(n\alpha_m)^{-1}$.
-
-\noindent \textbf{Objective Function:} After actor $i$ has been given the opportunity to change, it randomly selects one of its current ties, $x_{ij}$, to change.
+Since the rate is the same for all actors, the waiting time for \textit{any} actor to get the opportunity to change its set of ties is exponentially distributed with expected value $(n\alpha_m)^{-1}$.
+
+\noindent \textbf{Objective Function:} After actor $i$ has been selected by the rate function to change, it randomly picks one of its current ties, $x_{ij}$, to change.
 The probability that actor $i$ changes its current tie to actor $j$ is determined by the \textit{objective function} of the model and a random error term $U$ with log-Weibull distribution \citep{modelsSnijders}.
@@
 %The objective function, $f_i$, is the driving force of a SAOM. 
-Actor $i$ aims to maximize its corresponding objective function $f_i$ given the current state of the network, $x$ and the node-level covariates, $\mathbf{Z}$.
-This function is defined as:% which has the form
+Actor $i$ aims to maximize the objective function $f_i$ given the current state of the network, $x$ and the node-level covariates, $\mathbf{Z}$. This function is defined as:% which has the form
 \begin{equation}\label{eq:objective}
@@
 \end{equation}
-where $\boldsymbol{\beta} = (\beta_1, \dots, \beta_K)$ are additional model parameters, each associated with some network statistics, $s_{ik}(x, \mathbf{Z})$, $s_{ik}(x, \mathbf{Z})$, calculated with respect to actor $i$ at the current network state $x$.
-Network statistics range from the simple outdegree, $s_i(x) = \sum_{i\neq j} x_{ij}$, to the more complicated {\it transitive triplets jumping to different covariate}, $s_i(x, \mathbf{Z}) = \sum_{i \neq j \neq h} x_{ij}x_{ih}x_{hj} \cdot \mathbb{I}(z_i = z_h \neq z_j)$.
+where $\boldsymbol{\beta} = (\beta_1, \dots, \beta_K)$ are additional model parameters, each associated with some statistics, $s_{i1}(x, \mathbf{Z})$, \dots, s_{iK}(x, \mathbf{Z})$, calculated for actor $i$ at the current network state $x$.
+The statistics can be very simple, like outdegree of the node, $s_i(x) = \sum_{i\neq j} x_{ij}$, or more complicated, like the {\it transitive triplets jumping to different covariate} statistic, $s_i(x, \mathbf{Z}) = \sum_{i \neq j \neq h} x_{ij}x_{ih}x_{hj} \cdot \mathbb{I}(z_i = z_h \neq z_j)$.
 %, plus many more.
@@
 Version 1.2-3 of \texttt{RSiena} \citep{RSiena}, the software used to fit the models here, provides over 80 possible effects that can be included in the objective function.
-We discuss these statistics in more detail in Section~\ref{sec:models}.
-
-Objective function $f_{i}(x, \boldsymbol{\beta}, \mathbf{Z})$ and  random component $U$ combine to form the \textit{transition probability}, $p_{ij}$, of the network changing from its current state $x$ to the state with changed tie $x_{ij}$, denoted as $x(i \leadsto j)$:
+We discuss these and other statistics in more detail in Section~\ref{sec:models}.
+
+The objective function $f_{i}(x, \boldsymbol{\beta}, \mathbf{Z})$ and  random component $U$ combine to form the \textit{transition probability}, $p_{ij}$ of the network changing from its current state $x$ to the state $x(i\leadstoj)$, which is identical to $x$ except for $x_{ij}$: $x_{ij}(i \leadsto j) = 1 - x_{ij}$.
+The transition probability is
 \begin{equation}\label{eq:transprob}
@@
 \end{equation}
-This probability dictates which edge change is made by the acting node.
-The acting node can also choose to \textit{not} change at all.
-This is most likely to occur when the numerator of Equation~\ref{eq:transprob}, as calculated for the current state of the network, is larger than for any changes $x(i \leadsto j)$ that could be made.
+This probability dictates which edge change is made by the node $i$.
+Node $i$ can also choose to \textit{not} change at all, which is most likely when the value of Equation~\ref{eq:objective} is larger for $x$ than for any $x(i \leadsto j)$.
 
@@
 
-According to \citet{RSienamanual}, at least two parameters must be included in the objective function: the density and the reciprocity.
+According to \citet{RSienamanual}, at least two parameters must be included in the objective function: density and reciprocity.
 We denote the density, or out-degree, parameter by $\beta_1$ and the associated statistic as $s_{i1}(x) = \sum_{j} x_{ij}$.
@@
 Here, we use the lineup protocol to design significance, goodness-of-fit (GoF) and power tests.
-Under the protocol, we begin with a data set of interest to us, such as a network, and a visualization of this data, such as a node-link diagram.
-Suppose we have a hypothesized model that we think generated this data.
-There are two hypothesis considered: a null hypothesis which states the model that is assumed to have generated the data, and an alternative hypothesis that the data were not generated under this model.
-Then, the plot of the data of interest is placed randomly among a set of $m-1$ null plots, which are visualizations of data simulated from the null model (where $m =20$, usually).
-Human observers are then asked to examine the lineup and to identify the plot(s) that look most different from the others.
-If an observer identifies the data plot, this is quantifiable evidence against the null hypothesis.
-Since an observer has a chance of 1 in $m$ to pick the data plot from the lineup by simply guessing, i.e.
-in a situation where the data plot is virtually indistinguishable from the null plots, the evidence grows in strength with the number of independent observers identifying the data plot.
-
-The lineup protocol places a \textit{plot} in the framework of hypothesis tests: the plot of the data is the test statistic, which is compared against the sampling distribution under the null hypothesis represented by the null plots.
-Obviously, the null generating mechanism, i.e.\ the method of obtaining the data for null plots, is crucial for the lineup protocol, as the null hypothesis directly affects the choice of null generating method.
+Under the protocol, we begin with a data set of interest to us, such as a network, a visualization of this data, such as a node-link diagram, and a model of interest that we think generated this data.
+There are two hypotheses considered: the null hypothes that the model of interest generated the data, and the alternative hypothesis that the data were not generated under this model.
+Then, $P-1$ sets of data are simulated from the null model, where $P$ is the lineup size.
+Each of the $P-1$ simulated datasets are visualized in the same way as the data, and the plot of the data is placed randomly among the set of $P-1$ \textit{null plots}.
+Human observers then examine the lineup and identify the plot(s) that look most different from the others.
+If an observer identifies the data plot, this is evidence against the null hypothesis.
+Since an observer has a chance of $\frac{1}{P}$ to pick the data plot from the lineup by simply guessing, the evidence grows in strength with the number of independent observers identifying the data plot.
+
+The lineup protocol places a \textit{plot} in the framework of hypothesis tests: the plot of the data is the test statistic, which is compared against the null plots, representing the sampling distribution under the null hypothesis.
+%Obviously, the null generating mechanism, i.e.\ the method of obtaining the data for null plots, is crucial for the lineup protocol, as the null hypothesis directly affects the choice of null generating method.
 Null generating methods are typically based on (a) simulation, if the null hypothesis allows us to directly specify a parametric model, (b) sampling, as in the case of large data sets, or (c) permutation of the original data \citep[see e.g.\ ][]{Good05}s.
@@
 %In the experimental data that we analyzed the null generating methods used were permutation methods and direct simulation from a null model.
-
 The lineup protocol was formally tested in a head-to-head comparison with the equivalent conventional test in \citet{majumder:2011}.
 The experiment utilized human subjects from Amazon's Mechanical Turk \citep{turk} and used simulation to control conditions.
-The results suggest that  visual inference is comparable to conventional tests in a controlled conventional setting.
-This provides support for its appropriateness for testing in situations where no conventional test exists or when testing is difficult.
+The results suggest that VI done in a controlled setting gives the same results as conventional tests.
+This is evidence that VI can be used when no conventional tests exist or when testing is difficult.
 %Interestingly, the power of a visual test increases with the number of observers engaged to evaluate lineups, and the pattern in results suggests that the power will provide results consistent with practical significance \citep{kirk:1996}.
@@
 \noindent \textbf{Data:} The data we use are collaboration networks in the United States Senate during the $111^{th}$ through $114^{th}$ Congresses.
-These senates began on January 6, 2009, the start date of the $111^{th}$ Congress, and ended on January 3, 2017, the last date of the $114^{th}$ Congress (Details of how this data can be downloaded are provided by François Briatte at \url{github.com/briatte/congress}.).
-In the US Senate, there are three ways that senators can show support for a piece of legislation: they can author the bill, cosponsor the bill, or vote for the bill.
-We use cosponsorship as a metric because it results in a network that is unimodal (all nodes are senators) and directed.
-In this network, ties are directed from senator $i$ to senator $j$ when senator $i$ signs on as a cosponsor to the bill that senator $j$ authored.
+These senates began on January 6, 2009, the start date of the $111^{th}$ Congress, and ended on January 3, 2017, the last date of the $114^{th}$ Congress Details of how this data can be downloaded are provided by François Briatte at \url{github.com/briatte/congress}.
+In the US Senate, senators often show support for a piece of legislation by cosponsoring a bill authored by one of their colleagues.
+In a cosponsorship network, ties are directed from senator $i$ to senator $j$ when senator $i$ signs on as a cosponsor to the bill that senator $j$ authored.
 There are many hundreds of ties between senators when they are connected in this way, so we simplify the network by computing a single value for each senator-senator collaboration called the \textit{weighted propensity to cosponsor} (WPC).
@@
 \begin{equation}\label{eq:sen1}
-    WPC_{ij} = \dfrac{\sum\limits_{k=1}^{n_j} \frac{Y_{ij(k)}}{c_{j(k)}}}{\sum\limits_{k=1}^{n_j} \frac{1}{c_{j(k)}}}
+    WPC_{ij} = \dfrac{\sum\limits_{b=1}^{B_j} \frac{Y_{ij(b)}}{c_{j(b)}}}{\sum\limits_{b=1}^{B_j} \frac{1}{c_{j(b)}}}
 \end{equation}
-
-where $n_j$ is the number of bills in a congressional session authored by senator $j$, $c_{j(k)}$ is the number of cosponsors on senator $j$'s $k^{th}$ bill, where $k \in \{1,\dots, n_j\}$, and $Y_{ij(k)}$ is a binary variable that is 1 if senator $i$ cosponsored senator $j$'s $k^{th}$ bill, and is 0 otherwise.
+where $B_j$ is the number of bills in a congressional session authored by senator $j$, $c_{j(b)}$ is the number of cosponsors on senator $j$'s $b^{th}$ bill, where $b \in \{1,\dots, B_j\}$, and $Y_{ij(b)}$ is an indicator variable that senator $i$ cosponsored senator $j$'s $b^{th}$ bill.
 This measure ranges in value from 0 to 1, where $WPC_{ij} = 1$ if senator $i$ is a cosponsor on every one of senator $j$'s bills and $WPC_{ij} = 0$ if senator $i$ is never a cosponsor any of senator $j$'s bills.
@@
 \end{equation}
-For each of the four senate sessions, in addition to the WPC value between any two senators in the session, we have three node covariates: the party affiliation of each senator, the number of bills they authored in each session, and their gender.
-We explore each of these covariates in CTMC models to determine if they affect the overall network structure and how ties are formed between senators.
-The node-link diagram representations of the data we use for modelling are shown in Figure~\ref{fig:senateAll}.
-We have labelled some of the nodes in these networks whose names will be familiar to US readers, because they are leaders in their party or they have run for president.
-The size of the nodes represent how many bills the senator authored in a session, the color represents party affiliation, and the shape represent gender.
-In each of the four sessions, there is one very large connected component tying many of the prominent senators together, with many smaller groups of two to ten senators surrounding the larger component.
+so that only strong relationships between senators are examined.
+For each of the four senate sessions, we have three node covariates: the party affiliation of each senator, the number of bills they authored in each session, and their gender.
+We use each of these covariates in CTMC models to try to explain how ties are formed between senators over time.
+The node-link diagram for the senate network is shown in Figure~\ref{fig:senateAll}.
+We have labelled some of the nodes in the network whose names will be familiar to US readers, either because they are leaders in their party or they have run for president.
+The size of the nodes represents how many bills the senator authored in a session, the color represents party affiliation, and the shape represents gender.
+In each of the four sessions, there is one very large connected component tying many of the prominent senators together, with many smaller connected groups surrounding the larger component.
 In each senate, the structure changes slightly as new senators arrive or come to prominence.
 
-<<senateAll, fig.cap="The four senate collaboration networks that we use as our example data to visually assess the models of interest.
+<<senateAll, fig.cap="The US senate collaboration network observed at four time points.
 Color represents party, shape represents gender, and size represents number of bills authored in a session.
-The Frucherman-Reingold layout is shown.
+The Frucherman-Reingold layout is shown \citep{fr}.
 Drawn with the \\texttt{geomnet} R package.
@@
 seobama2 <- seobama %>% filter(senate != 111) %>% bind_rows(se111noclint)
-
-set.seed(56049382)
+mynames <- data.frame(do.call(rbind, stringr::str_split(seobama2$label, " ", n = 2) ))
+names(mynames) <- c("first", "last")
+seobama2$label <- mynames$last
+
+set.seed(93654)
 ggplot(data = seobama2) + 
-  geom_net(directed = T, labelon=T, arrowsize = .25, singletons= T, fiteach = T, linewidth = .25, layout.alg = 'fruchtermanreingold', fontsize = 2, repel=T, 
+  geom_net(directed = T, labelon=T, arrowsize = .25, singletons= T, fiteach = T, linewidth = .25, layout.alg = 'fruchtermanreingold', fontsize = 2, repel=T, labelcolour = "grey40", 
            aes(from_id = source, to_id = target, color = party, 
@@
 
-For Senate 111, for instance, we see Hillary Clinton, serving out her second term in the senate until she became Secretary of State.
+%For Senate 111, for instance, we see Hillary Clinton, serving out her second term in the senate until she became Secretary of State.
 She is isolated in Figure~\ref{fig:senateAll}, but in actuality, she had many cosponsors on only two pieces of legislation she authored in that short time.%, as is shown in Figure~\ref{fig:senateClinton}.
-We chose to remove Clinton and her edges from the network because they make the overall structure look vastly more connected than the other three senates.
+%We chose to remove Clinton and her edges from the network because they make the overall structure look vastly more connected than the other three senates.
 We suspect that because she had just recently run for President and had been selected by President-Elect Obama to be Secretary of State, the cosponsorships she garnered at the start of the new congress were largely symbolic, so the $111^{th}$ Senate without Hillary Clinton is more typical than the $111^{th}$ Senate with her.
@@
 
-In legislative cosponsorship networks, it is well known that party affiliation, reciprocity of relationships, and whether senators are ``workhorses" who author many bills or ``showhorses" who author few bills, are major influences on structure \citep{legnet}.
+In legislative cosponsorship networks, it is known that party affiliation, reciprocity of relationships, and whether senators are ``workhorses" who author many bills or ``showhorses" who author few bills, are major influences on structure \citep{legnet}.
 We focus on these covariates, plus the additional gender covariate when choosing which CTMC models to fit to the data.
 
-\noindent \textbf{Models:} In order to determine which models to examine, we first consider already well-known effects in legislative networks.
+\noindent \textbf{Models:} %In order to determine which models to examine, we first consider already well-known effects in legislative networks.
 We also selected some effects which were significant according to the built-in Wald-type tests in \texttt{RSiena} for application of our significance and goodness-of-fit methods.
-We examine a total of six models, each identified by its objective function:
+We examine a total of six models, each identified by its objective function below. The parameters in the objective function, $\beta_1, \dots, \beta_6$ are defined in Table~\ref{tab:effects}.
 \begin{enumerate}
@@
 \end{enumerate}
-The parameters in the objective function, $\beta_1, \dots, \beta_6$ are defined in Table~\ref{tab:effects}.
 
@@
 %The parameters we use for the remainder of the paper are detailed in Table~\ref{tab:effects}.
-All models had the first two parameters, $\beta_1, \beta_2$ at minimum included.
+All models have the first two structural parameters, $\beta_1 and \beta_2$, included.
 Other effects were added one at a time so that the simplest model $M1$ is nested in each of the other models.
 The largest model is  M7: each of the other five models is nested in it.
-The most significant additional effect is $\beta_3$, the jumping transitive triplet (JTT) parameter for the party covariate, which was estimated to be about -6 with a standard error of 0.11, resulting in a Wald $p$-value of less than 0.0001.
-This parameter associated considers the number of transitive closures formed between two senators from different parties.
-The large negative estimate is an indication that forming transitive ties between two people from different parties is strongly discouraged, which corresponds to the divisive nature of American politics, where party affilitation is the dominant effect.
-Another significant effect is $\beta_4$, the same JTT parameter for the sex covariate, with an estimate of about 3 with a standard error of 0.89.
+We used Wald tests in \texttt{RSiena} as described in Section~\ref{sec:res} to perform significance tests on $\beta_3, \dots, \beta_6$ in models M3, ..., M6.
+The most significant effect is $\beta_3$, the jumping transitive triplet (JTT) parameter for the party covariate, which was estimated to be about -5.9 with a standard error of 0.11, resulting in a Wald $p$-value of less than 0.0001.
+This parameter considers the number of transitive closures formed between two senators from different parties.
+The large negative estimate is an indication that forming transitive ties between two people from different parties is strongly discouraged, which comports with the divisive nature of American politics.
+Another significant effect is $\beta_4$, the same JTT parameter for the sex covariate, with an estimate of about 3.3 with a standard error of 0.89.
 This parameter also considers transitive closures, but for senators of different genders.
 The positive value indicates that transitive ties between senators of different genders are more likley to form.
-Next, we consider $\beta_6$, the covariate-related similarity score-weighted transitive triplets parameter estimate for the number of bills authored by a senator.
-We chose to look at similarity instead of covariate "jumps" because the number of bills authored is more continuous than gender or party.
+Next, we consider $\beta_5$, the covariate-related similarity score-weighted transitive triplets parameter for the number of bills authored by a senator.
+We chose to look at similarity instead of raw covariate value because the number of bills authored is more continuous than gender or party.
 The similarity measure is computed as:
@@
 where $\max_{hk}|b_h - b_k|$ is the range of number of bills authored by senators, and $b_i$, $b_j$ are the number of bills authored by senators $i,j$ repectively in the senate period.
-This effect was estimated at about 10 with standard error of 3.9.
+This effect was estimated at about 9.8 with standard error of 3.9.
 The high positive estimate suggests senators are encouraged to collaborate with other senators who author about the same number of bills they do.
@@
 Senators known as workhorses author many pieces of legislation in a session, and largely stay out of the public arena.
-The showhorse senators, on the other hand, author relatively few pieces of legislation, and tend to appear on television, radio, and other media a great deal.
-Finally, we found $\beta_5$, the same party transitive triplet effect to be significant, with a fitted value of 1.3 and standard error of 0.7, meaning that transitive relationships between senators tend to form when they are from the same party, exactly as we would expect in a legislative body in a country with extremely entrenched partisan divides as in the US.
+The showhorse senators, on the other hand, author relatively few pieces of legislation, and tend to appear in the media very frequently.
+Finally, we found $\beta_6$, the same party transitive triplet effect to be significant, with a fitted value of 1.3 and standard error of 0.7, meaning that transitive relationships between senators tend to form when they are from the same party, exactly as we would expect in a legislative body in a country with deeply entrenched partisan divides.
 
@@
 In the picture which represents each effect, the dotted tie is encouraged to form if the estimate is positive, and discouraged to form if the estimate is negative.
-$^*$ - $sim^b_{ij} = \frac{\max_{hk}|b_h - b_k| - |b_i - b_j|}{\max_{hk}|b_h - b_k|}$ is the similarity score between two senators based on the number of bills authored, and $\overline{sim}^b = \frac{1}{n(n-1)}\sum_{i\neq j} sim^b_{ij}$ is the average bill similarity score between any two senators.}
+$^*$ - $sim^b_{ij} is defined in Equation~\ref{eq:similar} and $\overline{sim}^b = \frac{1}{n(n-1)}\sum_{i\neq j} sim^b_{ij}$ is the average bill similarity score between two senators.}
 \end{table}
@@
 We fit all six of our models in \texttt{RSiena} using Markov Chain Monte Carlo (MCMC) methods to approximate the method of moments estimates of the parameters.
-Because the estimation is done through MCMC simulation, we fit each model to the data 1,000 times to get a better estimate of the true value of $\boldsymbol{\beta}$.
-From the simulations in which the MC converged, which made up over 90\% of the fits for each model, we computed the mean of the 1,000 estimates of each parameter to get final estimates of  $\hat{\boldsymbol{\beta}}$ for each model, which are given in Table~\ref{tab:fittedvalues}.
+Because the estimation is done through MCMC simulation, we fit each model to the data 1,000 times.
+From the simulations in which the MC converged, which made up over 90\% of the fits for each model, we computed the mean of the the estimates of each parameter to get final estimates of $\hat{\boldsymbol{\beta}}$ for each model, which are given in Table~\ref{tab:fittedvalues}.
 
@@
 
-\section{Experimental Design}\label{sec:expersetup}
-
-We want to explore three different assessments of the CTMC models using the lineup protocol: (1) significance of parameters, (2) goodness-of-fit of a model, and (3) visual power of the effects.
+\section{Experiment Design}\label{sec:expersetup}
+
+We want to assess CTMC models for social network data using the lineup protocol in three ways: (1) significance tests of parameters, (2) goodness-of-fit tests of a model, and (3) determining visual power of the effects.
 %visual detection of parameters.
-Each one of these situations requires a different setup, which we describe in detail, making use of the lineup protocol defined in \citet{Bujaetal} to study all of these aspects.
-In each lineup, we include plots from two models: the null model and an alternative model.
-The definition of the null and alternative model varies with the models we are exploring.
+Each one of these situations requires a different setup, which we describe in detail, making use of the lineup protocol defined in \citet{Bujaetal}.
+In each lineup, we include plots from two models: a null model and an alternative model.
+The definition of the null and alternative model varies with the model and the assessment we explore.
 %For each lineup, we consider two models: model M1 is the null model, and another model, from M2-M5, which is chosen to be the alternative model.
 
-Typically, a lineup shows sets of 20 plots at a time c.f.~\citet{loy:2015, vanderplas:2016}, but we determined that not enough structure could be shown in each plot for 20 large node-link diagrams.
-We chose to present our participants with only six plots at a time in order to show the node-link diagrams in more detail and to lower cognitive load for participants.
+Typically, a lineup shows sets of 20 plots at a time, for example in \citet{loy:2015, vanderplas:2016}, but we determined that looking at 20 node-link diagrams at once is too difficult.
+We chose to present our participants with only six plots at a time in order to show the node-link diagrams in more detail and to reduce their cognitive load.
 To construct a lineup, we simulate five networks from the null model and one network from the alternative model.
-An example of a lineup like those shown to our participants is given on the right side of the image  in Figure~\ref{fig:shiny}.
-In this lineup, model M4 is the alternative model, and model M1 is the null model.
+An example of a lineup like those shown to our participants is givenå in Figure~\ref{fig:shiny}.
+In this lineup, model M4 is the alternative model (one plot), and model M1 is the null model (five plots).
 
 To simulate lineups from the models we used the \texttt{siena07} function in \texttt{RSiena} \citep{RSiena}.
-Sections~\ref{sec:setupst} through~\ref{sec:setupvp} describe in detail how we set up the lineups, which parameter values we simulate from, and why.
-Lineups we created were shown to independent observers recruited through Amazon Mechanical Turk for feedback (more details on the Turk setup in Section~\ref{sec:res}).
+For the purposes of our experiment, we focus on simulating the second "wave" of data, the $112^{th}$ Senate network, and we condition on the first wave of data, $111^{th}$ Senate network.
+Sections~\ref{sec:setupst} through~\ref{sec:setupvp} describe in detail how we constructed the lineups, which parameter values we simulate from, and why.
+Lineups we created were shown to independent observers recruited through Amazon Mechanical Turk for feedback, and we provide more detail on the Turk setup in Section~\ref{sec:res}.
 %, we set the parameters to the values given in Table~\ref{tab:fittedvalues} for all parameters within the respective models, with the exception of $\beta_5$.
@@
 
-In the significance testing protocol, a parameter of interest is selected to test, say $\beta_k$. The hypotheses we use to generate lineups are:
+In the significance testing protocol, a parameter of interest $\beta_k$ is selected to test. The hypotheses we use to generate lineups are:
 
@@
 %\end{description}
-Under the null hypothesis, we assume that the model that generated the network data is $M_1$, the simplest model presented in Section~\ref{sec:models}.
-Thus, the five null plots in the lineup are simulations from M1 with $\beta_1, \beta_2$ set to the estimates given in Table~\ref{tab:fittedvalues} for M1.
+Under the null hypothesis, we assume that the model that generated the network data is M1, the simplest model presented in Section~\ref{sec:models}.
+In the lineup, there are five null plots, constructed from  five simulations from M1 of the second wave, with $\beta_1, \beta_2$ set to the corresponding estimates given in Table~\ref{tab:fittedvalues}.
 The alternative model is the model with $\beta_1, \beta_2$, and $\beta_k$ in the objective function.
 The alternative data plot is simulated from the appropriate model with values set to the estimates from Table~\ref{tab:fittedvalues}.
-
-The lineup generated under this scenario is shown to a number of independent viewers.
 If an observer picks out the alternative data plot, that is evidence against the null hypothesis, while picking one of the null plots is evidence in favor of the null hypothesis.
-The significance tests that we perform in our experiment are for $\beta_3$ and $\beta_4$, making the alternative models M3 and M4, respectively, for three repitions each.
-
-\subsection{Goodness-of-Fit}\label{sec:gofsetup}
+To avoid exhausting our participants, we chose only two parameters to test, $\beta_3$ and $\beta_4$.
+Thus, we compare the null model M1 to the alternative models M3 and M4, using three repetions of each hypothesis test.
+
+\subsection{Goodness-of-Fit Testing}\label{sec:gofsetup}
 
@@
 We pick a wave to focus on, wave two, which is the first simulated network, and among these five plots, we place a node-link diagram of the true second wave data.
-We cannot show the data more than once to each participant, so we examine several different models with three repitions each in our Amazon Mechanical Turk experiment, with each participant never seeing the actual data twice.
-The models we chose for goodness-of-fit testing are M3, M4, M5, and M7.
-
-\subsection{Visual Power}\label{sec:setupvp}
+We cannot show the data more than once to each participant.
+The models we chose for goodness-of-fit testing are M3, M4, M5, and M7, and each participant only saw one lineup for the goodness-of-fit tests.
+
+\subsection{Visual Power Testing}\label{sec:setupvp}
 
 %Visual detection does not fit into the typical statistical testing framework as easily as the significance and goodness-of-fit tests we perform do. 
-Through visual inference, we want to determine at which point an additional effect included in the model becomes noticeable.
-By \textit{noticeable}, we mean that the inclusion of the effect alters the appearance of networks simulated from the model so much that viewers are able to reliably pick out the one node-link diagram rendered from data simulated from the model \textit{with} the effect in a lineup among plots simulated from the model \textit{without} the effect.
-This is a way to determine the power of the visual significance test.
-We explore all parameters in the objective function, $\beta_1, \dots, \beta_6$ in this way.
+Using VI, we want to determine at what value an additional effect included in the model becomes noticeable.
+By \textit{noticeable}, we mean that the inclusion of the effect alters the appearance of networks simulated from the model so much that many independent viewers are able to pick out the one node-link diagram rendered from data simulated from the model \textit{with} the effect in a lineup among five plots simulated from the model \textit{without} the effect.
+In this way, we measure the visual power of a parameter.
+We perform visual power tests for all parameters in the objective function, $\beta_1, \dots, \beta_6$.
 
 In model M1, with only two parameters in the objective function, we varied both the density and reciprocity parameter values one at a time, keeping all other parameters at their fitted values given in Table~\ref{tab:fittedvalues}.
-In models M3 through M6, we vary the additional parameter, $\beta_3$ through $\beta_6$, respectively, while again holding the rate and other parameters at their value in Table~\ref{tab:fittedvalues}.
-Thus, we have six different parameters of interest to us in total: $\beta_1, \dots, \beta_6$.
-We want to determine how the size of these parameters affects the overall structure of the network data simulated from the models M1 through M6, so we also vary the magnitude of the parameters in order to determine at what level the effects become noticeable.
+In models M3 through M6, we vary the additional parameter, $\beta_3$ through $\beta_6$, respectively, while holding all other parameters at their value in Table~\ref{tab:fittedvalues}.
+We want to determine how the size of these parameters affects the overall structure of the network data simulated from the models M1 through M6, so we also vary the magnitude of the parameters in order to determine at what value the effects become noticeable.
 %in both negative and positive directions.
@@
 
-To determine the threshold at which an effect becomes noticeable, we examine six different levels of the effect, three negative and three positive ones.
-Figure~\ref{fig:hypothesis} shows a sketch of what the detection probability by participants' looks like hypothetically with varying effect size: the higher in absolute value the parameter is, the more likely participants are to choose the alternative model out of the lineup.
+To determine when an effect becomes noticeable, we examine six different levels of the effect, three negative and three positive ones.
+Figure~\ref{fig:hypothesis} shows a sketch of what the selection probability looks like hypothetically with varying effect size: the higher in absolute value the parameter is, the more likely participants are to select the alternative model out of the lineup.
 To determine the exact values of the six levels we want to test for each effect, we started with the estimates of the parameter at hand (see  Section~\ref{sec:models}), and used small negative and positive factors to determine at what point \textit{we} noticed the effect of the parameter in simulations from the changed models.
@@
 
-To decide on the values to use for each difficulty level, we constructed an online application that simulated the lineup protocol for us to be the guinea pigs of our own experiment \citep{theqs}.
+To decide on the values to use for each difficulty level, we constructed an online application that simulated the lineup protocol for us to be the guinea pigs in our own experiment \citep{theqs}.
 A screen shot of the app we created with the \texttt{shiny} package by \citet{shiny} is shown in Figure~\ref{fig:shiny}.
-On the left side of the screen, the user\footnote{Please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/} to create lineups constructed from the models we present for this data for yourself.} can input the information necessary for creating a lineup of the models M1 through M7 for the data in Section~\ref{sec:senatedata}: first, choose to simulate only one plot from the specified model (analogous to changing the alternative model in the lineup protocol) or to simulate $M-1$ plots from the specified model (analogous to changing the null model in the lineup protocol); then choose the model of interest and the wave of the data to examine.
-If model M1 is selected as the model of interest, select whether to alter the density or the reciprocity parameter.
-Then, pick the size of the lineup; the amount by which to multiply the effect selected; a random seed for replicability; and a layout algorithm to use for the node-link diagrams.
+On the left side of the screen, the user\footnote{Please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/} to create lineups constructed from the models we present for this data for yourself.} can input the information necessary for creating a lineup from the models M1 through M7 for the data in Section~\ref{sec:models}:
+\begin{enumerate}
+\item Choose to simulate one plot to change the alternative model in the lineup, or simulate $P-1$ plots to change the null model in the lineup
+\item Choose the model from which to simulate and choose the wave of data to simulate. 
+\item If M1 is selected in (2), select whether to alter the density ($\beta_1$) or the reciprocity ($\beta_2$) parameter. 
+\item Choose the size of the lineup, $P$, the magnitude of the effect size, a random seed for reproducibility, and a layout algorithm to use for the node-link diagrams.
 There is also a checkbox if the user wishes the nodes to be colored by the size of the connected component to which they belong.
-The plot(s) that appear(s) that are \textit{not} from the model specified using the other options are simulated from model M1 with the estimates of the rate parameters, and $\beta_1$ and $\beta_2$ as given in Table~\ref{tab:fittedvalues}.
-
-Using a ``Picking Lineups" web application, we settled on six parameter values to test for each of our six effects, $\beta_1, \dots \beta_6$.
-The complete details of the parameters tested using the lineup protocol is given in Table~\ref{tab:experdetail}.
-In the case of both $\beta_4$ and $\beta_6$, we could not determine any values for negative effects that made the data simulated from M4 and M6 look different than null model simulations from model M1.
-We hypothesized that this was due to the negative effects \textit{removing} visually interesting structural elements as opposed to \textit{adding} noticeable structural elements.
-Therefore we decided that the lesser experienced participants in our experiment would also not be able to.
-Instead of testing the negative values of these effects, we are examining a different scenario: we placed 5 simulations from model M4 or M6 with positive values of the parameter with one simulation from model M1 in a lineup.
+\end{enumerate}
+All plots that are \textit{not} from the model specified by steps 1-4 above are simulated from model M1 with the estimates of the rate parameters, and $\beta_1$ and $\beta_2$ as given in Table~\ref{tab:fittedvalues}.
+
+% latex table generated in R 3.5.2 by xtable 1.8-3 package
+% Mon Mar  4 14:49:43 2019
+\begin{table}[ht]
+\centering
+\begin{tabular}{llrrr}
+  \hline
+Parameter & Condition & Easy Value & Medium Value & Hard Value \\ 
+  \hline
+\hline
+$\beta_1$ & neg & -7.354 & -6.6187 & -5.883 \\ 
+   & pos & -3.922 & -4.1674 & -4.412 \\ 
+  $\beta_2$ & neg &  0.000 & 0.0005 & 0.049 \\ % -17.249 & -10.3497 & -3.450 \\ 
+   & pos & 7.340 & 6.8504 & 6.361 \\  %10.350 & 6.8998 & 5.175 \\ 
+  $\beta_3$ & neg & -17.249 & -10.3497 & -3.450 \\ %8.351 & 6.6806 & 5.010 \\ 
+   & pos & 10.350 & 6.8998 & 5.175 \\   %6.681 & 5.0105 & 3.340 \\ 
+  $\beta_5$ & neg & -30.272 & -20.1817 & -10.091 \\ %0.000 & 0.0005 & 0.049 \\ 
+   & pos &  20.182 & 17.6590 & 16.145 \\ %7.340 & 6.8504 & 6.361 \\ 
+   \hline
+\hline
+$\beta_4$ & pos & 8.351 & 6.6806 & 5.010 \\ % 5.316 & 3.9872 & 3.323 \\ 
+   & reverse & 6.681 & 5.0105 & 3.340 \\  %5.316 & 3.9872 & 3.323 \\ 
+  $\beta_6$ & pos & 5.316 & 3.9872 & 3.323 \\  %-30.272 & -20.1817 & -10.091 \\ 
+   & reverse & 5.316 & 3.9872 & 3.323 \\ %20.182 & 17.6590 & 16.145 \\ 
+   \hline
+\end{tabular}
+\caption{All conditions used for our experiment.
+For parameters $\beta_1, \beta_2, \beta_3,$ and $\beta_5$, M1 served as the null model.
+For $\beta_4$ and $\beta_6$, null model M1 and the alternative model (M4 or M6) switch roles in the reversed lineups, i.e.
+five plots show data simluated from the laternative model and only one plot shows data from M1.}
+\label{tab:experdetail}
+\end{table}
+
+Using our Shiny application, we settled on six parameter values to test for each of our six effects, $\beta_1, \dots \beta_6$.
+All values of the parameters used in the experiment are given in Table~\ref{tab:experdetail}.
+In the case of both $\beta_4$ and $\beta_6$, we could not determine any negative parameter values that made the data simulated from M4 and M6 look different than null model simulations from model M1.
+We hypothesize that this is due to negative effects \textit{removing} visually interesting structural elements as opposed to \textit{adding} noticeable structural elements.
+Since we could not detect the effects, we decided that the participants in our experiment would also not be able to.
+So, instead of testing the negative values of these effects, we use a different scenario: we place five simulations from model M4 or M6 (with positive values of the parameter) with one simulation from model M1 in a lineup.
 We will refer to this as the ``reverse" lineup scenario.
-We used the reverse scenario to determine if the perception of the effect size is symmetric: if an effect is noticed $x$\% of the time at value $\beta_k = \beta_{k_0}$ when \textit{one} simulation from the corresponding model is placed among \textit{five} null plots from model M1, then when \textit{five} simulations from the model with $\beta_k = \beta_{k_0}$ are put in a lineup with \textit{one} simulation from model M1, the plot from the simpler model should be noticed about $x$\% of the time as well.
+We used the reverse scenario to determine if the perception of the effect size is symmetric: if an effect is noticed $p$\% of the time at value $\beta_k = \beta_{k_0}$ when \textit{one} simulation from the corresponding model is placed among \textit{five} null plots from model M1, then when \textit{five} simulations from the model with $\beta_k = \beta_{k_0}$ are put in a lineup with \textit{one} simulation from model M1, the plot from the simpler model should be noticed about $p$\% of the time as well.
 
@@
 experdet$test_param <- as.factor(experdet$test_param)
-levels(experdet$test_param) <- paste0("\\$\\beta_", c(1,2,3,5,4,6), "\\$")
-experdet$test_param <- factor(experdet$test_param, levels=paste0("\\$\\beta_", c(1,2,3,5,4,6), "\\$"))
+levels(experdet$test_param) <- paste0("$\\beta_", c(1,2,3,5,4,6), "$")
+experdet$test_param <- factor(experdet$test_param, levels=paste0("$\\beta_", c(1,2,3,5,4,6), "$"))
 experdet <- experdet %>% arrange(test_param)
@@
 
-print(xtable::xtable(experdet, caption = "All conditions used for the MTurk experiment.
+# xtable::print.xtable(xtable::xtable(experdet, caption = "All conditions used for the MTurk experiment.
 For parameters $\\beta_1, \\beta_2, \\beta_3,$ and $\\beta_5$ M1 served as null model.
@@
 five plots show data simluated from the laternative model and only one plot shows data from M1.", label = "tab:experdetail", digits = c(0,0,0,3,4,3)), include.rownames=FALSE,
-      hline.after = c(-1,0,0,8,8,12))
-@
-
-\section{Results}\label{sec:res}
+#       hline.after = c(-1,0,0,8,8,12), sanitize.text.function = function(x){x})
+@
+
+\subsection{Execution}
 
@@
 Each participant was presented with some brief training material before beginning the experiment.
-After agreeing to participate, each person was shown two trial plots, one where the data plot was the most different from the others due to its relatively \textit{complex} structure, while the other trial included a data plot that was most different from the others due to its comparatively \textit{simple} structure.
-Only when participants were able to correctly identify the data plot from the trial lineups were they allowed to begin the experiment.
+Before we presented the lineups for the hypothesis tests, each person was shown two trial plots one where the alternative plot was the most different from the others due to its relatively \textit{complex} structure, while the other trial included an alternative plot that was most different from the others due to its comparatively \textit{simple} structure.
+Only when participants were able to correctly identify the alternative plot from the trial lineups were they allowed to begin the experiment.
+
 Each participant was randomly assigned 13 lineups to look at.
-The were asked to select one or more plots that they perceived as ``most different" from the others, and provide a reasoning for their choice: ``Most simple overall structure",``Most complex overall structure", or ``Other".
-If they selected other, they were required to describe their own text description of their reasoning.
-The language in these reasons is purposefully vague: we want participants to tell us what they see.
+They were asked to select one or more plots that they perceived as ``most different" from the others, and provide a reasoning for their choice: ``Most simple overall structure",``Most complex overall structure", or ``Other".
+If they selected other, they were required to describe their reasoning.
+The language in the reasoning is purposefully vague: we want participants to tell us what they see.
 We do not want to tell them what they ``should" be seeing.
@@
 6 parameters, 3 levels, 2 directions}
-The final of the 13 lineups shown to participants contained the true data from the 112th senate shown in Section~\ref{sec:senatedata} placed among five other plots from models M3, M4, M5, and M7 as discussed in Section~\ref{sec:gofsetup}.
-Each participant only saw the data one time in order to avoid bias.
+The last lineups shown to participants contained the true data from the 112th senate shown in Section~\ref{sec:models} placed among five other plots from  one of the models M3, M4, M5, and M7 as discussed in Section~\ref{sec:gofsetup}.
 Upon completion of the 13 lineups, which took an average of 308 seconds per person, each participant was paid \$1.75
+
+\section{Results}\label{sec:res}
+
+Because the size of the lineups is small, the probability of picking the data by chance is high, $\frac{1}{6}$, but if \textit{many} independent viewers pick out the data from the nulls, the evidence in favor of the alternative hypothesis becomes stronger.
 
@@
 
-For a CTMC model, there are two ways a conventional significance test of the parameters can be performed.
-In \texttt{RSiena}, there are $t$-type and Wald-type tests for a single parameter and for multiple parameters.
+For CTMC models, significance tests of the parameters are available in \texttt{RSiena}.
+There are $t$-type and Wald-type tests for a single parameter and for multiple parameters.
 The $t$-type test statistic is simply the parameter estimate divided by its standard error, and compared to a standard normal distribution \citep{RSienamanual}.
@@
 \begin{equation}\label{eq:wald1}
-\frac{(\hat{\beta_k})^2}{var(\hat{\beta_k})} \sim \chi^2_1,
+\frac{(\hat{\beta_k})^2}{var(\hat{\beta_k})},
 \end{equation}
-which is compared to a Chi-square distribution with one degree of freedom\citep{RSienamanual}.
-Testing the significance of multiple parameters depends on the hypothesis we wish to test, and a $P \times K$ matrix, $A$, must be appropriately designed to test the $P$ hypotheses of interest.
-The null hypothesis is that $A\boldsymbol{\beta} = \mathbf{0}$, and the test statistic is
+which is compared to a Chi-square distribution with one degree of freedom \citep{RSienamanual}.
+Testing the significance of multiple parameters depends on the hypothesis we wish to test, and an $H \times K$ matrix, $\mathbf{A}$, must be appropriately designed to test the $H$ hypotheses of interest.
+The null hypothesis is that $\mathbf{A}\boldsymbol{\beta} = \mathbf{0}$, and the test statistic is
 \begin{equation}\label{eq:wald2}
-(A\boldsymbol{\hat{\beta}})' \hat{\Sigma}^{-1} A\boldsymbol{\hat{\beta}} \sim \chi^2_P,
+(\mathbf{A}\boldsymbol{\hat{\beta}})' \hat{\Sigma}^{-1} \mathbf{A}\boldsymbol{\hat{\beta}},
 \end{equation}
-where $\hat{\Sigma}$ is the estimated covariance matrix of $\boldsymbol{\beta}$. This statistic is then compared to a Chi-square distribution with $P$ degrees of freedom.
-
-Both the parameters we test for significance using the lineup protocol, $\beta_3$ and $\beta_4$, were determined to be statistically significant using Equation~\ref{eq:wald1}.
-The correspoding results from the significance tests we performed using the lineup protocol are given in Table~\ref{tab:sigtesting}.
+where $\hat{\Sigma}$ is the estimated covariance matrix of $\boldsymbol{\beta}$. This statistic is then compared to a Chi-square distribution with $H$ degrees of freedom.
+
+Both parameters we test for significance using the lineup protocol, $\beta_3$ and $\beta_4$, were determined to be statistically significant using Equation~\ref{eq:wald1}.
+The corresponding results from the significance tests we performed using the lineup protocol are given in Table~\ref{tab:sigtesting}.
 If enough participants pick out the alternative plot to result in a $p$-value less than 0.05, we reject the null hypothesis that the true value of the additional parameter, either $\beta_3$ or $\beta_4$, is equal to zero.
@@
 summ_sig$parameter <- as.factor(summ_sig$parameter)
-levels(summ_sig$parameter) <- c("beta3", "beta4")
-print(xtable::xtable(summ_sig, caption = "Experiment results for the two parameters for which we performed significance tests.
-There were three lineups for each parameter, so there are three results for each plot.", label = "tab:sigtesting", digits = c(0,0,0,0,0,5)), include.rownames = FALSE)
+levels(summ_sig$parameter) <- c("$ \\beta_3 $", "$ \\beta_4$ ")
+xtable::print.xtable(xtable::xtable(summ_sig, caption = "Experiment results for the two parameters for which we performed significance tests.
+There were three lineups for each parameter, so there are three results for each plot.", label = "tab:sigtesting", digits = c(0,0,0,0,0,5)), include.rownames = FALSE, sanitize.text.function=function(x){x})
 @
@@
 This package contains methods to calculate \textit{Visual distributions} for lineup experiment data.
-The distribution depends on the number of evaluations of a plot, $K$, the size of the lineup, $m$, and the lineup scenario, which here is that each lineup containing the same data and the same set of null plots is shown to $K$ independent observers.
-The visual inference family of distributions is similar to the binomial distribution, but takes the dependency among the $m$ plots in a single lineup shown to multiple viewers into account.
-Using these $p$-values, all but one lineup results in a rejection of the null hypothesis at Type-I error rate of $\alpha = 0.05$.
-We see that the $p$-values for visual inference here are highly variable.
-
-<<lineup3132, fig.height=4, fig.cap="The lineup which caused a rejection of the null hypothesis that $\\beta_3$ = 0.
+The distribution depends on the number of evaluations of a plot, $L$, the size of the lineup, $P$, and the lineup scenario, which here is that each lineup containing the same data and the same set of null plots is shown to $L$ independent observers.
+The visual inference family of distributions is similar to the binomial distribution, but takes the dependency among the $P$ plots in a single lineup shown to multiple viewers into account.
+%Using these $p$-values, all but one lineup results in a rejection of the null hypothesis at Type-I error rate of $\alpha = 0.05$.
+We see in Table~\ref{tab:sigtesting} that the $p$-values for visual inference here are highly variable.
+
+<<lineup3132, fig.height=4, fig.cap="Lineup 3132, which led to rejection of the null hypothesis that $\\beta_3$ = 0.
 The network simulated from model M3 is found in panel $\\sqrt{16} - 1$, and the remaining panels show networks simulated from model M1.">>=
@@
 Another significance lineup for model M3, which resulted in failure to reject the null hypothesis, is shown in Figure~\ref{fig:lineup3133}.
-When viewing Figure~\ref{fig:lineup3132}, 26 of 31 viewers chose the plot from M3, while only 2 of 27 chose the plot from M3 when viewing Figure~\ref{fig:lineup3133}.
-The most common choice in the latter was panel two, which 16 of 27 viewers chose as the most different due to its large connected component, making it seem more complex than the others.
-In viewing these two lineups, it is evident that there is a large amount of variability.
-It is difficult to see that five of the six come from the same model when they can all look different in their own way.
-Thus, the variability in results is introduced through the null plots generated from M1, as not all simulations look alike.
-In addition, the necessarily small number of null plots do not give the viewer as complete of a view of the null model as the usual 19 null plots would.
+When viewing Figure~\ref{fig:lineup3132}, 26 of 31 viewers chose the alternative plot from M3, while only 2 of 27 chose the alternative plot from M3 when viewing Figure~\ref{fig:lineup3133}.
+The most common choice in the latter was panel two, which 16 of 27 viewers chose as the most different due to its large connected component that makes it seem more complex than the others.
+In viewing these two lineups, it is evident that there is a large amount of variability in networks simulated from CTMC models.
+It is difficult to see that five of the six networks come from the same model when they all look very different.
+The variability in significance test results is introduced through the null plots generated from M1.
+In addition, the small number of null plots do not give the viewer as complete of a view of the null model as the usual 19 null plots would.
+
 The results of the significance tests given in Table~\ref{tab:sigtesting} for $\beta_3$ and $\beta_4$ are not definitive.
@@
 For the test of $\beta_4$, one test is significant, one is decidedly not significant, and the third is significant at the level of 0.10.
-Thus, unlike the Wald-type tests described at the beginning of this section, there is no way to decisively reject or to fail to reject the null hypothesis that the parameter value is 0.
-We include all of the lineups shown to our participants in the appendix.
+Thus, unlike with the Wald-type tests described at the beginning of this section, we cannot decisively reject or to fail to reject the null hypothesis that the parameter value is 0.
+%We include all of the lineups shown to our participants in the appendix.
 
@@
 Most network models, other than the most simple, lack the necessary asymptotics for developing goodness-of-fit methods \citep{goldenberg09}.
-Some methods have been developed based on what \citeauthor{RSienamanual} call ``auxiliary statistics" such as the indegree or outdegree distribution on the nodes.
+Some simulation-based methods have been developed using what \citeauthor{RSienamanual} call \textit{auxiliary statistics} such as the indegree or outdegree distribution on the nodes.
 In \texttt{RSiena}, the \texttt{sienaGOF} function performs goodness-of-fit testing as follows:
 \begin{enumerate}
-\item Auxiliary statistics, such as the cumulative outdegree counts on the nodes, are computed on the observed data ($\mathbf{u}_d$) and on $N$ simulated observations from the model ($\mathbf{u}_1 \dots \mathbf{u}_N$).
-(Usually, $N=1000$)
-\item The mean vector, $\overline{\mathbf{u}}$ and covariance matrix, $\mathbf{S}$ of the statistics on the simulations from the model are computed, and the Mahalanobis distance, $d_M(\mathbf{u})$ from the observed statistics to the distribution of the simulated statistics is computed:
+\item Auxiliary statistics, such as the cumulative outdegree counts on the nodes, are computed on the observed data ($\mathbf{u}_d$) and on $N$ observations simulated from the model ($\mathbf{u}_1 \dots \mathbf{u}_N$).
+\item The mean $\overline{\mathbf{u}}$ and covariance matrix $\mathbf{S}$ are computed from the $N$ simulations, and the Mahalanobis distance, $d_M(\mathbf{u}_d)$ from the observed statistics to the distribution of the simulated statistics is computed:
 \begin{equation}\label{eq:mahal}
-d_M(\mathbf{u}) = \sqrt{(\mathbf{u} - \overline{\mathbf{u}})' S^{-1} (\mathbf{u} - \overline{\mathbf{u}})}
+d_M(\mathbf{u}_d) = \sqrt{(\mathbf{u}_d - \overline{\mathbf{u}})' \mathbf{S}^{-1} (\mathbf{u}_d - \overline{\mathbf{u}})}
 \end{equation}
 \item The Mahalanobis distance for each of the $N$ simulations is calculated and $d_M(\mathbf{u}_d)$ is compared to this distribution of distances.
-\item An empirical $p$-value is found by computing the proportion of simulated distances found in step 4 that are as large or larger than $d_M(\mathbf{u}_d)$.
-A model is thus considered a good fit to the data if $p$ is large.
-A plot comparing the data to the simulations is also considered, and a similar plot is shown in Figure~\ref{fig:gofsiena} for the outdegree distribution of small data set, shown in the points and connected lines, with the simulated values of $\mathbf{u}_d$ shown in boxplots and overlaid violin plots.
+\item An empirical $p$-value is found by computing the proportion of simulated distances found in step 4 that are as large or larger than $d_M(\mathbf{u}_d)$.% A model is thus considered a good fit to the data if $p$ is large.
 \end{enumerate}
-
-<<gofsiena, fig.height=3, fig.cap="An example of what a goodness-of-fit plot from \\texttt{RSiena} looks like.
+<<gofsiena, fig.height=2.5, fig.cap="An example of what a goodness-of-fit plot from \\texttt{RSiena} looks like.
 The overlaid boxplots and violin plots show the distribution of each of the outdegree count values on the simulated networks, and the red points and lines are the observed data values.", cache = TRUE>>=
@@
 @
-
-The \texttt{RSiena} software also provides a Rao score-type test for goodness-of-fit for assessing one or more parameters, the test statistic of which is compared to a Chi-square distribution with $P$ degrees of freedom, where $P$ has the same definition as in Section~\ref{sec:sigtest}.
+A plot comparing the data to the simulations is also considered, and a similar plot is shown in Figure~\ref{fig:gofsiena} for the outdegree distribution of small data set, shown in the points and connected lines, with the simulated values of $\mathbf{u}_d$ shown in boxplots and overlaid violin plots.
+The \texttt{RSiena} software also provides a Rao score-type test for goodness-of-fit for assessing one or more parameters, the test statistic of which is compared to a Chi-square distribution with $H$ degrees of freedom, where $H$ has the same definition as in Section~\ref{sec:sigtest}.
 For full detail on the score-type test, see \citet{scoretest}.
-
 These methods are both restriced: the \texttt{sienaGOF} method only considers \textit{one} measure on the data and simulations from the model, while the score-type tests only consider \textit{subsets} of parameters, ``nuisance parameters" in \citet{scoretest}, not the entire set of parameters.
@@
 Using the lineup protocol, we show each Amazon Mechanical Turk worker the data once, in a lineup with five other plots of simulated data from one of the models we chose.
-We examined four different models, M3, M4, M5, and M7, and examined three repetitions of each, for a total of 12 goodness-of-fit lineups.
-In each lineup, the ``null model" is one of the four models and the ``alternative" model is the true, unknown model that generated the senate network data.
+We examined four different models, M3, M4, M5, and M7, and examined three repetitions of each, for a total of 12 goodness-of-fit lineups shown to participants.
+In each lineup, the null model is one of the four models and the alternative model is the true, unknown model that generated the senate network data.
 The hypotheses for our goodness-of-fit tests are:
 \begin{description}
-\item $H_0$:  The senate network data come from (or could have come from) the null model, M$i$.
+\item $H_0$:  The senate network data come from the null model, M$i$.
 \item $H_A$: The senate network data do not come from the null model. 
@@
 On the contrary, if the lineup viewer picks one of the null plots, that is evidence in favor of the null hypothesis.
-Because the size of the lineups is small, the probability of picking the data by chance is high, $\frac{1}{6}$, but if \textit{many} independent viewers pick out the data from the nulls, the evidence in favor of the alternative hypothesis becomes stronger.
 Results from our MTurk goodness-of-fit plots are provided in Table~\ref{tab:gofstats}.
@@
 print_gof_stats$Model[c(2:3,5:6,8:9,11:12)] <- ""
-print_gof_stats$Model[2] <- "jtt party"
-print_gof_stats$Model[5] <- "jtt sex"
-print_gof_stats$Model[8] <- "stt party"
+#print_gof_stats$Model[2] <- "jtt party"
+#print_gof_stats$Model[5] <- "jtt sex"
+#print_gof_stats$Model[8] <- "stt party"
 print_gof_stats$`p-value` <- ifelse(print_gof_stats$`p-value` < .0001, "< 0.0001", sprintf("%.4f", print_gof_stats$`p-value`))
@@
 Only 7 of 20 viewers of this lineup selected the data plot as the most different from the others.
-The most commonly chosen panel was number four, which has a relatively simple structure compared to panels 2, 3, and 6 especially.">>=
+The most commonly chosen panel was number four, which has a relatively simple structure compared to the other panels.">>=
 load("data/se112adjmat.RDS")
@@
 For model M5, we can neither reject nor fail to reject the null hypothesis that the data come from model M5.
-This is evidence that the goodness-of-fit of network models cannot always be determined by one dimensional derived features, such as $p$-value shown on the $x$-axis in Figure~\ref{fig:gofsiena}.
+This is evidence that the goodness-of-fit of network models cannot always be determined by one dimensional derived features.%, such as $p$-value shown on the $x$-axis in Figure~\ref{fig:gofsiena}.
 %This leads us to believe that the high $p$-value observed in the second replicate is more of a fluke.
@@
 We could move the threshold from $\alpha = 0.05$ to $\alpha = 0.01$ or $\alpha = 0.001$, but more work needs to be done to determine which, if any of these, is appropriate.
+
+<<mostsiggof, fig.height = 4, fig.cap="The lineup resulting in the smallest $p$-value rejecting the null hypothesis. Surprisingly, this another repetition for M5 as the null model.">>=
+dat <- read_csv(paste0("data/jtts_gof_9_3.csv"))
+datplot <- unique(dat$ord[dat$sim == 1001])
+dat <- dat %>% filter(sim != 1001) %>% 
+      bind_rows(wave2) %>% 
+      mutate(ord = ifelse(is.na(ord), datplot, ord))
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
+        theme_net() + 
+        theme(panel.background = element_rect(color = 'black')) +
+        facet_wrap(~ord)
+@
+
+For the other models for which we tested goodness-of-fit, however, we \textit{do} have significant evidence from all three replicates to reject the null hypothesis that the null model generated the data.
+For models M3, M5, and M7, these goodness-of-fit tests have rejected the null hypotheses that the senate data come from these models.
+We hypothesized that the model with the most effects, M7, would be the best fit.
+However, as shown in Figure~\ref{fig:gofm71}, the model does not capture the overall structure very well at all.
+The rest of the goodness-of-fit lineups as shown to participants are provided in the appendix.
+
+%\hh{XXX we need to go throught the logic of the argument below.}\st{ XXX I reformulated the discussion per our convo today.  }
 
@@
 @
-For the other models for which we tested goodness-of-fit, however, we \textit{do} have significant evidence from all three replicates to reject the null hypothesis that the null model generated the data.
-For models M3, M5, and M7, these goodness-of-fit tests have rejected the null hypotheses that the senate data come from these models.
-We hypothesized that the model with the most effects, M7, would be the best fit.
-However, as shown in Figure~\ref{fig:gofm71}, the model does not capture the overall structure very well at all.
-The rest of the goodness-of-fit lineups as shown to participants are provided in the appendix.
-
-%\hh{XXX we need to go throught the logic of the argument below.}\st{ XXX I reformulated the discussion per our convo today.  }
-
-<<mostsiggof, fig.height = 4, fig.cap="The lineup resulting in the smallest $p$-value rejecting the null hypothesis. Surprisingly, this another repetition for M5 as the null model.">>=
-dat <- read_csv(paste0("data/jtts_gof_9_3.csv"))
-datplot <- unique(dat$ord[dat$sim == 1001])
-dat <- dat %>% filter(sim != 1001) %>% 
-      bind_rows(wave2) %>% 
-      mutate(ord = ifelse(is.na(ord), datplot, ord))
-ggplot(data = dat) + 
-        geom_net(aes(from_id = from, to_id = to), 
-                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
-                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
-                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
-        theme_net() + 
-        theme(panel.background = element_rect(color = 'black')) +
-        facet_wrap(~ord)
-@
 
@@
 
-A summary of the results from our experiment is shown as points in Figure~\ref{fig:predictglmm}.
+The results from the visual power piece of our experiment are shown in Figure~\ref{fig:predictglmm}.
 On the $x$ axis, we plot the value of the parameter of interest, and on the $y$ axis, the proportion of times the alternative data plot was picked out for each lineup.
@@
 
-We further explore this relationship between identification of the alternative data in the lineup and the parameter of interest, effect size, and lineup type with a generalized linear mixed model that provides us with an estimate of the power of the visual significance test.
-The response variable, $Y_{ijkm}$, is binary, indicating whether participant $m$ picked the alternative data plot in lineup type $j$, rep $k$, for effect $i$.
-There is one continuous covariate $x$, which is the centered and scaled size of the effect of interest from which the alternative data were simulated, the values of which are labeled ``easy", ``medium", and ``hard" in Table~\ref{tab:experdetail} according to how difficult we thought the Turk participants would find each lineup.
-In Equation~\ref{eq:glmm}, $i \in \{1, 2, 3, 4, 6\}$ corresponds to the effects $\beta_1,\dots, \beta_6$, respectively, $j \in \{-1,1\}$, and $k \in \{1,2,3\}$.
-We also include random effects in the model: one for each lineup, $\delta_{ijk}$, and one for each participant, $\epsilon_m$, and fit a hierarchical model given in Equation~\ref{eq:glmm}.
+We model the relationship between identification of the alternative data in the lineup and the parameter of interest, the effect size, and the lineup type with a generalized linear mixed model (GLMM) that provides us with an estimate of the power of the visual significance test.
+The response variable, $X_{klqr}$, is binary, indicating whether participant $r$ picked the alternative data plot in lineup type $q$, rep $\ell$, for effect $k$.
+There is one continuous covariate $v$, which is the centered and scaled size of the effect of interest from which the alternative data were simulated, the values of which are labeled ``easy", ``medium", and ``hard" in Table~\ref{tab:experdetail} according to how difficult we thought the Turk participants would find each lineup.
+In Equation~\ref{eq:glmm}, $k \in \{1, 2, 3, 4, 6\}$ corresponds to the effects $\beta_1,\dots, \beta_6$, respectively, $\ell \in \{-1,1\}$, and $q \in \{1,2,3\}$.
+We also include random effects in the model: one for each lineup, $\delta_{klq}$, and one for each participant, $\epsilon_r$, and fit a hierarchical model given in Equation~\ref{eq:glmm}.
 
@@
 \begin{split}
-Y_{ijkm} &\sim \text{Bernoulli}(\pi_{ijkm}) \\
-\text{logit}(\pi_{ijkm}) & = \alpha_{ij} + \gamma_{ij} x + \delta_{ijk} + \epsilon_{m} \label{eq:glmm} \\
-\delta_{ijk} & \stackrel{iid}{\sim} N(0, \sigma^2_{\delta}) \\
-\epsilon_{m} & \stackrel{iid}{\sim} N(0, \sigma^2_{\epsilon})
+X_{klqr} &\sim \text{Bernoulli}(\pi_{klqr}) \\
+\text{logit}(\pi_{klqr}) & = \eta_{kl} + \gamma_{kl} v + \delta_{klq} + \epsilon_{r} \label{eq:glmm} \\
+\delta_{klq} & \stackrel{iid}{\sim} N(0, \sigma^2_{\delta}) \\
+\epsilon_{r} & \stackrel{iid}{\sim} N(0, \sigma^2_{\epsilon})
 \end{split}
@@
 The results of fitting this model, including estimates of parameters, standard errors, $p$-values, and the odds ratio multipliers, using \texttt{glmer} from the \texttt{lme4} package are summarized in Table~\ref{tab:glmmests} \citep{lme4}.
-For each combination of parameter and lineup type, the expected value of the link function for a new lineup and a new observer with parameter value $x$ is
+For each combination of parameter and lineup type, the expected value of the link function for a new lineup and a new observer with parameter value $v$ is
 \begin{equation}
-E[\text{logit}(\pi_{ij})] = \alpha_{ij} + \gamma_{ij}x
+E[\text{logit}(\pi_{kl})] = \alpha_{kl} + \gamma_{kl}v
 \end{equation}
@@
 \begin{equation}
-\pi_{ij} = \frac{\exp\{\alpha_{ij} + \gamma_{ij}x\}}{1 + \exp\{\alpha_{ij} + \gamma_{ij}x\}}
+\pi_{kl} = \frac{\exp\{\alpha_{kl} + \gamma_{kl}v\}}{1 + \exp\{\alpha_{kl} + \gamma_{kl}x\}}
 \end{equation}
@@
 Parameter & \multicolumn{1}{r}{Estimate} & \multicolumn{1}{r}{Std Error} & $p$-value & Odds Multiplier \\ \hline\hline
-$\alpha_{1+}$ & \Sexpr{ests[7]} & \Sexpr{se[7]} & \Sexpr{pval[7]}$^{\ddagger}$ & \Sexpr{odds[7]} \\
-$\alpha_{1-}$ & \Sexpr{ests[1]} & \Sexpr{se[1]} & \Sexpr{pval[1]}$^{\dagger}$ & \Sexpr{odds[1]}\\
+$\eta_{1+}$ & \Sexpr{ests[7]} & \Sexpr{se[7]} & \Sexpr{pval[7]}$^{\ddagger}$ & \Sexpr{odds[7]} \\
+$\eta_{1-}$ & \Sexpr{ests[1]} & \Sexpr{se[1]} & \Sexpr{pval[1]}$^{\dagger}$ & \Sexpr{odds[1]}\\
 $\gamma_{1+}$ & \Sexpr{ests[19]} & \Sexpr{se[19]} & \Sexpr{pval[19]}$^{\ddagger}$ & \Sexpr{odds[19]} \\
 $\gamma_{1-}$  & \Sexpr{ests[13]} & \Sexpr{se[13]} & \Sexpr{pval[13]}$^{\dagger}$ & \Sexpr{odds[13]} \\ \hline
-$\alpha_{2+}$  & \Sexpr{ests[10]} & \Sexpr{se[10]} & \Sexpr{pval[10]} & \Sexpr{odds[10]} \\
-$\alpha_{2-}$  & \Sexpr{ests[4]} & \Sexpr{se[4]} & \Sexpr{pval[4]}$^{\ddagger}$ & \Sexpr{odds[4]} \\
+$\eta_{2+}$  & \Sexpr{ests[10]} & \Sexpr{se[10]} & \Sexpr{pval[10]} & \Sexpr{odds[10]} \\
+$\eta_{2-}$  & \Sexpr{ests[4]} & \Sexpr{se[4]} & \Sexpr{pval[4]}$^{\ddagger}$ & \Sexpr{odds[4]} \\
 $\gamma_{2+}$ & \Sexpr{ests[22]} & \Sexpr{se[22]} & \Sexpr{pval[22]}$^{*}$ & \Sexpr{odds[22]} \\
 $\gamma_{2-}$  & \Sexpr{ests[16]} & \Sexpr{se[16]} & \Sexpr{pval[16]}$^{\ddagger}$ & \Sexpr{odds[16]} \\ \hline
-$\alpha_{3+}$  & \Sexpr{ests[8]} & \Sexpr{se[8]} & \Sexpr{pval[8]}$^{\dagger}$ & \Sexpr{odds[8]} \\ 
-$\alpha_{3-}$ & \Sexpr{ests[2]} & \Sexpr{se[2]} & \Sexpr{pval[2]}$^{\dagger}$ & \Sexpr{odds[2]} \\
+$\eta_{3+}$  & \Sexpr{ests[8]} & \Sexpr{se[8]} & \Sexpr{pval[8]}$^{\dagger}$ & \Sexpr{odds[8]} \\ 
+$\eta_{3-}$ & \Sexpr{ests[2]} & \Sexpr{se[2]} & \Sexpr{pval[2]}$^{\dagger}$ & \Sexpr{odds[2]} \\
 $\gamma_{3+}$ & \Sexpr{ests[20]} & \Sexpr{se[20]} & \Sexpr{pval[20]}$^{\dagger}$ & \Sexpr{odds[20]}  \\
 $\gamma_{3-}$   & \Sexpr{ests[14]} & \Sexpr{se[14]} & \Sexpr{pval[14]}$^{*}$ & \Sexpr{odds[14]} \\ \hline
-$\alpha_{4+}$ & \Sexpr{ests[9]} & \Sexpr{se[9]} & \Sexpr{pval[9]}$^{**}$ & \Sexpr{odds[9]} \\
-$\alpha_{4-}$ & \Sexpr{ests[3]} & \Sexpr{se[3]} & \Sexpr{pval[3]}$^{**}$ & \Sexpr{odds[3]}  \\
+$\eta_{4+}$ & \Sexpr{ests[9]} & \Sexpr{se[9]} & \Sexpr{pval[9]}$^{**}$ & \Sexpr{odds[9]} \\
+$\eta_{4-}$ & \Sexpr{ests[3]} & \Sexpr{se[3]} & \Sexpr{pval[3]}$^{**}$ & \Sexpr{odds[3]}  \\
 $\gamma_{4+}$ & \Sexpr{ests[21]} & \Sexpr{se[21]} & \Sexpr{pval[21]}$^{**}$ & \Sexpr{odds[21]} \\ 
 $\gamma_{4-}$  & \Sexpr{ests[15]} & \Sexpr{se[15]} & \Sexpr{pval[15]} & \Sexpr{odds[15]} \\ \hline
-$\alpha_{5+}$  & \Sexpr{ests[12]} & \Sexpr{se[12]} & \Sexpr{pval[12]}$^{*}$ & \Sexpr{odds[12]} \\
-$\alpha_{5-}$  & \Sexpr{ests[6]} & \Sexpr{se[6]} & \Sexpr{pval[6]}$^{\ddagger}$ & \Sexpr{odds[6]} \\
+$\eta_{5+}$  & \Sexpr{ests[12]} & \Sexpr{se[12]} & \Sexpr{pval[12]}$^{*}$ & \Sexpr{odds[12]} \\
+$\eta_{5-}$  & \Sexpr{ests[6]} & \Sexpr{se[6]} & \Sexpr{pval[6]}$^{\ddagger}$ & \Sexpr{odds[6]} \\
 $\gamma_{5+}$ & \Sexpr{ests[24]} & \Sexpr{se[24]} & \Sexpr{pval[24]}$^{*}$ & \Sexpr{odds[24]} \\ 
 $\gamma_{5-}$ & \Sexpr{ests[18]} & \Sexpr{se[18]} & \Sexpr{pval[18]}$^{\ddagger}$ & \Sexpr{odds[18]} \\ \hline
-$\alpha_{6+}$  & \Sexpr{ests[11]} & \Sexpr{se[11]} & \Sexpr{pval[11]} & \Sexpr{odds[11]} \\
-$\alpha_{6-}$ & \Sexpr{ests[5]} & \Sexpr{se[5]} & \Sexpr{pval[5]}$^{\ddagger}$ & \Sexpr{odds[5]}  \\
+$\eta_{6+}$  & \Sexpr{ests[11]} & \Sexpr{se[11]} & \Sexpr{pval[11]} & \Sexpr{odds[11]} \\
+$\eta_{6-}$ & \Sexpr{ests[5]} & \Sexpr{se[5]} & \Sexpr{pval[5]}$^{\ddagger}$ & \Sexpr{odds[5]}  \\
 $\gamma_{6+}$ & \Sexpr{ests[23]} & \Sexpr{se[23]} & \Sexpr{pval[23]} & \Sexpr{odds[23]} \\
@@
 
-In Figure~\ref{fig:predictglmm}, we see a clear trend in all parameters except $\beta_1$ and $\beta_2$ that as the parameter value approaches zero from either side, the probability of picking the data plot in a lineup of size six descreases.
+In Figure~\ref{fig:predictglmm}, we see a clear trend in all parameters except $\beta_1$ and $\beta_2$ that as the parameter value approaches zero from either side, the probability of picking the data plot in a lineup of size six decreases.
+This supports our hypothesis shown in Figure~\ref{fig:hypothesis}.
 For $\beta_3$ and $\beta_5$, the slope of the fitted line is \textit{much} steeper for positive values of the parameter than for negative values, meaning that our participants perceived differences more often for postitive parameter values than for negative parameter values.
-This finding is similar to that of \citet{corrviz}, who found that people detect positive correlations sooner and better than negative correlations.
+This finding is similar to that of \citet{corrviz}, who found that people detect positive correlations better and at lower values than negative correlations.
 
@@
 We expand portions of Figure~\ref{fig:predictglmm} in Figures~\ref{fig:recipzoom}-\ref{fig:beta4zoom}.
-These figures show the same prediction regions as in Figure~\ref{fig:predictglmm}, plus some additional predictions outside of the data range shown in gray.
+These figures show the same prediction regions as in Figure~\ref{fig:predictglmm}, plus some additional predictions outside of the data range shown in light gray.
 Again, the points represent the results from the experiment.
@@
 This demonstrates that the visual signal of one plot from M4 among five plots from M1 is much stronger than that of one plot from M1 among five plots from M4.
-We posit that the latter is a more difficult task because it involves noticing a \textit{lack of structure} as opposed to the presence of more structure.
+We posit that the latter is a more difficult task because it involves noticing a \textit{lack of structure} as opposed to the presence of \textit{more} structure.
 We can see a similar effect in Figure~\ref{fig:beta5zoom}.
@@
 At a value of $\beta_5 = -20$, however, the model predicts this same probability to be about 0.40.
-This again demonstrates that the presence of structure is detected sooner and more frequently than the absence of strucutre.
+This again demonstrates that the \textit{presence} of structure is detected  more frequently and at smaller values than the absence of strucutre.
 
@@
 One of the lineups for the $\beta_4 = 6.681$, lineup type 1 scenario is given in Figure~\ref{fig:beta41}, and a corresponding lineup for the lineup type -1 scenario is given in Figure~\ref{fig:beta4neg1}.
-For identical values of the parameter, viewers had a harder time identifying the different plot when they were selected the most ``simple" structure, detecting M1 in five plots from the more complicated model, than they did identifying the most ``complex" structure, the plot from the more complicated model, from the five plots from M1.
-This result is also similar to that of \citet{corrviz} because it emphasizes the difficulty of picking out the absence of an effect relative to picking out the presence of an effect.
+For identical values of the parameter, viewers had a harder time identifying the different plot when they were selecting the most ``simple" structure, detecting M1 in five plots from the more complicated model, than they did identifying the most ``complex" structure, the plot from the more complicated model, from the five plots from M1.
+%This result is also similar to that of \citet{corrviz} because it emphasizes the difficulty of picking out the absence of an effect relative to picking out the presence of an effect.
 
@@
 % what did we do? what did we learn from it? 
-By using visual inference methods, we have developed new ways to perform significance and goodness-of-fit testing for a complicated and intractable set of statistical models for social network data.
+By using visual inference methods, we have developed new ways to perform significance and goodness-of-fit tests for a complicated, intractable set of statistical models for social network data.
 We have also developed a way to determine the power of these new visual tests.
@@
 The traditional methods only look at one piece or derived measure of a network model, whereas our methods look at the models holistically for a broader sense of what it means for a parameter to be significant or a model to be a good fit.
-By looking at an entire network simulated from a CTMC model side-by-side with other instances of networks simulated from another model, instead of singular %one-dimensional derived
-features, we develop an idea of the model in terms of the \textit{data} itself, instead of in terms of statistical summaries of the data.
-These methods place the model in the data space, instead of summarizing or compressing the data to place it in the model space.
+By looking at an entire network simulated from a network model side-by-side with other instances of networks simulated from a null model, instead of singular %one-dimensional derived
+features, we develop an idea of the model in terms of the \textit{data} itself, not in terms of statistical summaries of the data.
+These methods place the model in the data space: they don't summarize or compress the data to put it in the model space \citep{Wickham2015}.
 
@@
 These results do not come without limitations.
-In visual inference, the null plots are supposed to play the role of good representatives of the null model.
+In VI, the null plots are supposed to play the role of good representatives of the null model.
 Here, the number of null plots is reduced to five, which increases the variability seen in a single lineup dramatically, and can unfortunately lead to very different conclusions for the same lineup scenario, as we have seen.

### 2019-03-05 ASApaper/paper.Rnw
looks like i forgot to commit stuff
big push; almost done

@@
 All three tasks become more difficult as the model becomes more complex.
-We will explore the three assessments of one particularly complex set of models, continuous time Markov chain (CTMC) models, for dynamic social networks  \citep{saompaper}.
-In this paper, we propose new methods for significance and goodness-of-fit testing, as well as power calculation for CTMC models via the visual inference (VI) paradigm of \citet{Bujaetal}.
-With VI, we can look at entire datasets simulated from a model, instead of relying a single metric such as a $p$-value for a single parameter in the model or goodness-of-fit assessed by looking at outdegree.
+We will explore these three assessments of one particularly complex set of models, continuous time Markov chain (CTMC) models, for dynamic social networks  \citep{saompaper}.
+In this paper, we propose new methods for significance and goodness-of-fit testing, as well as power calculations for CTMC models via the visual inference (VI) paradigm of \citet{Bujaetal}.
+With VI, we can look at entire datasets simulated from a model, instead of relying a single metric such as a $p$-value.
 We conducted a VI experiment, with participants recruited via Amazon Mechanical Turk, to assess the significance of CTMC parameters, the fit of a CTMC model, and the visual power of CTMC parameters.
@@
 
-<<child-demo, child='paper-child-thesis-ch-2.Rnw'>>=
+<<child-demo, child='paper-child.Rnw'>>=
 @

### 2019-03-05 ASApaper/paper-child.Rnw
done with big revision

@@
 \end{equation}
-where $\boldsymbol{\beta} = (\beta_1, \dots, \beta_K)$ are additional model parameters, each associated with some statistics, $s_{i1}(x, \mathbf{Z})$, \dots, s_{iK}(x, \mathbf{Z})$, calculated for actor $i$ at the current network state $x$.
+where $\boldsymbol{\beta} = (\beta_1, \dots, \beta_K)$ are additional model parameters, each associated with some statistics, $s_{i1}(x, \mathbf{Z}), \dots, s_{iK}(x, \mathbf{Z})$, calculated for actor $i$ at the current network state $x$.
 The statistics can be very simple, like outdegree of the node, $s_i(x) = \sum_{i\neq j} x_{ij}$, or more complicated, like the {\it transitive triplets jumping to different covariate} statistic, $s_i(x, \mathbf{Z}) = \sum_{i \neq j \neq h} x_{ij}x_{ih}x_{hj} \cdot \mathbb{I}(z_i = z_h \neq z_j)$.
@@
 
-The objective function $f_{i}(x, \boldsymbol{\beta}, \mathbf{Z})$ and  random component $U$ combine to form the \textit{transition probability}, $p_{ij}$ of the network changing from its current state $x$ to the state $x(i\leadstoj)$, which is identical to $x$ except for $x_{ij}$: $x_{ij}(i \leadsto j) = 1 - x_{ij}$.
+The objective function $f_{i}(x, \boldsymbol{\beta}, \mathbf{Z})$ and  random component $U$ combine to form the \textit{transition probability}, $p_{ij}$ of the network changing from its current state $x$ to the state $x(i\leadsto j)$, which is identical to $x$ except for $x_{ij}$: $x_{ij}(i \leadsto j) = 1 - x_{ij}$.
 The transition probability is
@@
 Color represents party, shape represents gender, and size represents number of bills authored in a session.
-The Frucherman-Reingold layout is shown \citep{fr}.
+The Fruchterman-Reingold layout is shown \\citep{fr}.
 Drawn with the \\texttt{geomnet} R package.
@@
 In the picture which represents each effect, the dotted tie is encouraged to form if the estimate is positive, and discouraged to form if the estimate is negative.
-$^*$ - $sim^b_{ij} is defined in Equation~\ref{eq:similar} and $\overline{sim}^b = \frac{1}{n(n-1)}\sum_{i\neq j} sim^b_{ij}$ is the average bill similarity score between two senators.}
+$^*$ - $sim^b_{ij}$ is defined in Equation~\ref{eq:similar} and $\overline{sim}^b = \frac{1}{n(n-1)}\sum_{i\neq j} sim^b_{ij}$ is the average bill similarity score between two senators.}
 \end{table}
@@
 
-Because the size of the lineups is small, the probability of picking the data by chance is high, $\frac{1}{6}$, but if \textit{many} independent viewers pick out the data from the nulls, the evidence in favor of the alternative hypothesis becomes stronger.
+In this section, we present the results from our experiment and compare them to traditional statistical tests where applicable.
+Because each lineup shown to participants has only six plots, the probability of picking the data by chance is high, $\frac{1}{6}$, but if \textit{many} independent viewers pick out the data from the nulls, the evidence in favor of the alternative hypothesis becomes stronger.
+The $p$-values from the lineups were calculated using the \texttt{vinference} package by \citet{vinference}.
+This package contains methods to calculate \textit{Visual distributions} for lineup experiment data.
+The distribution depends on the number of evaluations of a plot, $L$, the size of the lineup, $P$, and the lineup scenario, which here is that each lineup containing the same data and the same set of null plots is shown to $L$ independent observers.
+The visual inference family of distributions is similar to the binomial distribution, but takes the dependency among the $P$ plots in a single lineup shown to multiple viewers into account.
 
@@
             pv = map2_dbl(npick, n, pV, m=6, scenario=3))
-names(summ_sig) <- c("Lineup ID", "parameter", "# Alt. Model Picks", "Total Views", "p-value")
+names(summ_sig) <- c("Lineup ID", "parameter", "\\# Alt. Model Picks", "Total Views", "p-value")
 summ_sig$parameter <- as.factor(summ_sig$parameter)
@@
 
-The $p$-values were calculated using the \texttt{vinference} package by \citet{vinference}.
-This package contains methods to calculate \textit{Visual distributions} for lineup experiment data.
-The distribution depends on the number of evaluations of a plot, $L$, the size of the lineup, $P$, and the lineup scenario, which here is that each lineup containing the same data and the same set of null plots is shown to $L$ independent observers.
-The visual inference family of distributions is similar to the binomial distribution, but takes the dependency among the $P$ plots in a single lineup shown to multiple viewers into account.
  %Using these $p$-values, all but one lineup results in a rejection of the null hypothesis at Type-I error rate of $\alpha = 0.05$. 
@@
 library(lattice)
-gof1 <- sienaGOF(ansnull, OutdegreeDistribution, varName ="friendship", join = FALSE)
+#gof1 <- sienaGOF(ansnull, OutdegreeDistribution, varName ="friendship", join = FALSE)
+gof1 <- readr::read_rds("data/gof-data.rds")
 sims1 <- data.frame(gof1$`Period 1`$Simulations)
@@
 
-<<predictglmm, fig.height=6, fig.cap='Predictions from our generalized linear mixed effects model given in Equation~\ref{eq:glmm}.
-The lines show the expected probability of detecting the alternative data in a lineup of size 6 for new observers of new lineups is plotted on the $y$-axis, and the size of the parameter of interest is on the $x$-axis.
+<<predictglmm, fig.height=6, fig.cap='Predictions from our GLMM given in Equation~\\ref{eq:glmm}.
+For new observers of new lineups, the lines show the expected probability of detecting the alternative data in a lineup of size 6 as a function of the parameter value.
 The proportions detected by our Turk participants for each lineup group are shown by the points, with the probability of picking out the data plot at random shown by a horizontal line at 1/6.
 The lineup marked as ``outlier" was removed from modeling.
-The panel for the reciprocity parameter, $\\beta_2$ is also presented in Figure~\\ref{fig:recipzoom} in more detail.'>>=
+Estimated parameter values shown by vertical dotted lines.'>>=
 newdata <- read_csv("data/newdata_pred_glmm.csv")
@@
 plotdata$param <- as.integer(as.character(plotdata$param))
-labdat <- data_frame(test_param = "jttp", param = 3, type2 = -1, x = -5, y = .9, label = "jttp_neg_hard_2")
+labdat <- data_frame(test_param = "jttp", param = 3, type2 = -1, x = -5, y = .9, label = "outlier")
+betahat <- tibble(param = 1:6, size = c(-4.903, 4.893, -3.45, 3.34, 10.091, 1.329))
 
 ggplot() + 
+  geom_vline(data = betahat, aes(xintercept = size), linetype = "dotted", color = "grey60") + 
+  geom_text(data = betahat, aes(x = size, y = .9, label = paste0("hat(beta)[", param, "]")), parse = T, hjust = 0, color = "grey60") + 
   geom_line(data = newdata, aes(x = size, y = predictfinal, color = as.factor(type2))) + 
@@
   geom_text(data = labdat, aes(x = x, y=y, label = label, color = as.factor(type2)),
-            show.legend = F) + 
+            show.legend = F, vjust = 2) + 
   geom_hline(yintercept = 1/6, linetype='dashed') +
@@
 Again, a horizontal line is drawn at 1/6, the chance of selecting the data plot at random.'>>=
+newdatabeta2 <- newdata %>% filter(test_param == "recip", category == "inside") %>% mutate(param2 = "beta[2]", size2 = sqrt(size)) %>% as.tibble()
+plotdatabeta2 <- plotdata %>% filter(test_param=="recip")%>% mutate(param2 = "beta[2]", size2 = sqrt(size)) %>% as.tibble()
+newdatabeta22 <-  newdata %>% filter(test_param == "recip", category == "outside")%>% mutate(param2 = "beta[2]", size2 = sqrt(size)) %>% as.tibble()
+newdatabeta23 <- newdata %>% filter(test_param == "recip", category == "outside2")%>% mutate(param2 = "beta[2]", size2 = sqrt(size)) %>% filter(!is.nan(size2)) %>% as.tibble()
+interceptlabel <- tibble(x = sqrt(4.893), y = .05, label = "hat(beta)[2]")
 ggplot() + 
-  geom_line(data = newdata %>% filter(test_param == "recip", category == "inside"), aes(x = sqrt(size), y = predictfinal, color = as.factor(type2)), size = 1.25) + 
-  geom_point(data = plotdata %>% filter(test_param=="recip"), aes(x = sqrt(size), y = datapick, color = as.factor(type2)), size = 2) + 
-  geom_line(data = newdata %>% filter(test_param == "recip", category == "outside"), aes(x = sqrt(size), y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
-  geom_line(data = newdata %>% filter(test_param == "recip", category == "outside2"), aes(x = sqrt(size), y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
+  geom_line(data = newdatabeta2, aes(x = size2, y = predictfinal, color = as.factor(type2)), size = 1.25) + 
+  geom_point(data = plotdatabeta2, aes(x = size2, y = datapick, color = as.factor(type2)), size = 2) + 
+  geom_line(data =newdatabeta22, aes(x = size2, y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
+  geom_line(data = newdatabeta23, aes(x = size2, y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
   #geom_text(data = plotdata, aes(x = size, y = datapick, color = as.factor(type2), label = pic_id)) +
@@
   #          show.legend = F) + 
-  geom_vline(xintercept = sqrt(4.893)) + 
+  geom_vline(xintercept = sqrt(4.893),  linetype = "dotted", color = "grey60") + 
+  geom_text(data = interceptlabel, aes(x =x, y = y, label = label), parse = T, inherit.aes = F, hjust = -.2) + 
   geom_hline(yintercept = 1/6, linetype='dotted') +
   scale_color_brewer(palette = "Dark2", name = "Lineup Type", labels = c("-", "+")) +  
-  facet_grid(~param) + 
-  labs(x = "Square root of parameter value", y = "Expected probability of detection\nin a lineup of size 6") + 
+  facet_wrap(~param2, labeller = label_parsed) + 
+  labs(x = "Square root of parameter value", y = "Expected probability of detection\nin a lineup of size 6", title = ) + 
   ThemeNoNet + 
-  theme(legend.position = 'bottom', strip.text = element_text(size = 10), text = element_text(size = rel(3)), legend.text = element_text(size = rel(3)), legend.key.size = unit(.25, "inches"))
+  theme(legend.position = 'bottom')
 @
@@
 According to the model, people will detect the effect at lower values and with greater frequency as the value increases when it is positive instead of negative.">>=
+newdatabeta5 <- newdata %>% filter(test_param == "simttb", category == "inside") %>% mutate(param = "beta[5]")
+plotdatabeta5 <- plotdata %>% filter(test_param=="simttb")%>% mutate(param = "beta[5]")
+newdatabeta52 <- newdata %>% filter(test_param == "simttb", category == "outside") %>% mutate(param = "beta[5]")
+newdatabeta53 <- newdata %>% filter(test_param == "simttb", category == "outside2") %>% mutate(param = "beta[5]")
+interceptdata <- tibble(x = 10.090829, y= .05, label = "hat(beta)[5]")
 ggplot() + 
-  geom_line(data = newdata %>% filter(test_param == "simttb", category == "inside"), aes(x = size, y = predictfinal, color = as.factor(type2)), size = 1.25) + 
-  geom_point(data = plotdata %>% filter(test_param=="simttb"), aes(x = size, y = datapick, color = as.factor(type2)), size = 2) + 
-  geom_line(data = newdata %>% filter(test_param == "simttb", category == "outside"), aes(x = size, y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
-  geom_line(data = newdata %>% filter(test_param == "simttb", category == "outside2"), aes(x = size, y = predictfinal, group = as.factor(type2)), color = "gray60",  alpha = .7) +
+  geom_line(data = newdatabeta5, aes(x = size, y = predictfinal, color = as.factor(type2)), size = 1.25) + 
+  geom_point(data = plotdatabeta5, aes(x = size, y = datapick, color = as.factor(type2)), size = 2) + 
+  geom_line(data = newdatabeta52, aes(x = size, y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
+  geom_line(data = newdatabeta53, aes(x = size, y = predictfinal, group = as.factor(type2)), color = "gray60",  alpha = .7) +
   #geom_text(data = plotdata, aes(x = size, y = datapick, color = as.factor(type2), label = pic_id)) +
@@
   #          show.legend = F) + 
-  geom_vline(xintercept = 10.090829) + 
+  geom_vline(xintercept = 10.090829, linetype = "dotted", color = "grey60") + 
+  geom_text(data = interceptdata, aes(x=x,y=y,label=label), parse = T, hjust = -.2) + 
   geom_hline(yintercept = 1/6, linetype='dotted') +
   scale_color_brewer(palette = "Dark2", name = "Lineup Type", labels = c("-", "+")) +
-  facet_grid(~param) + 
-  labs(x = "Parameter value", y = "Expected probability of detection in a lineup of size 6") + 
+  facet_grid(~param, labeller = label_parsed) + 
+  labs(x = "Parameter value", y = "Expected probability of detection\nin a lineup of size 6") + 
   ThemeNoNet + 
@@
 
-<<beta4zoom, fig.height=3, fig.cap='The bottom left panel of Figure~\\ref{fig:predictglmm} expanded to show greater detail.
+<<beta4zoom, fig.height=3, fig.cap='The bottom right panel of Figure~\\ref{fig:predictglmm} expanded to show greater detail.
 The parameter value is shown on the $x$-axis.
@@
 Reversing the lineup scenario was not symmetric as we hypothesized.'>>=
+newdatabeta6 <- newdata %>% filter(test_param == "jtts", category == "inside") %>% mutate(param = "beta[6]")
+plotdatabeta6 <- plotdata %>% filter(test_param=="jtts") %>% mutate(param = "beta[6]")
+newdatabeta62 <- newdata %>% filter(test_param == "jtts", category == "outside")%>% mutate(param = "beta[6]")
+newdatabeta63 <- newdata %>% filter(test_param == "jtts", category == "outside2") %>% mutate(param = "beta[6]")
+labellerdata <- tibble(x = 3.340302, y = .8, label = "hat(beta)[6]")
 ggplot() + 
-  geom_line(data = newdata %>% filter(test_param == "jtts", category == "inside"), aes(x = size, y = predictfinal, color = as.factor(type2)), size = 1.25) + 
-  geom_point(data = plotdata %>% filter(test_param=="jtts"), aes(x = size, y = datapick, color = as.factor(type2)), size = 2) + 
-  geom_line(data = newdata %>% filter(test_param == "jtts", category == "outside"), aes(x = size, y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
-  geom_line(data = newdata %>% filter(test_param == "jtts", category == "outside2"), aes(x = size, y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
+  geom_vline(xintercept = 3.340302, linetype = "dotted", color = "grey60") + 
+  geom_text(data = labellerdata, aes(x = x, y = y, label = label), parse =T, hjust = 1) + 
+  geom_line(data = newdatabeta6, aes(x = size, y = predictfinal, color = as.factor(type2)), size = 1.25) + 
+  geom_point(data = plotdatabeta6, aes(x = size, y = datapick, color = as.factor(type2)), size = 2) + 
+  geom_line(data = newdatabeta62, aes(x = size, y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
+  geom_line(data = newdatabeta63, aes(x = size, y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
   #geom_text(data = plotdata, aes(x = size, y = datapick, color = as.factor(type2), label = pic_id)) +
@@
   #          show.legend = F) + 
-  geom_vline(xintercept = 3.340302) + 
   geom_hline(yintercept = 1/6, linetype='dotted') +
   scale_color_brewer(palette = "Dark2", name = "Lineup Condition", labels = c("reverse", "regular")) +
-  facet_grid(~param) + 
-  labs(x = "Parameter value", y = "Expected probability of detection in a lineup of size 6") + 
+  facet_grid(~param, labeller = label_parsed) + 
+  labs(x = "Parameter value", y = "Expected probability of detection\nin a lineup of size 6") + 
   ThemeNoNet + 
@@
 In VI, the null plots are supposed to play the role of good representatives of the null model.
-Here, the number of null plots is reduced to five, which increases the variability seen in a single lineup dramatically, and can unfortunately lead to very different conclusions for the same lineup scenario, as we have seen.
+Here, the number of null plots is reduced to five, which has lead to very different conclusions for the same lineup scenario.
 Furthermore, these results do not generalize to all CTMC models or to one particular subset of CTMC models.
@@
 But given the limitations of the node-link visualization, the cognitive load of looking at a lineup is very high for the average observer.
-We would therefore like to explore larger datasets, different layout algorithms, and different ways of visualizing network data, such adjacency matrix visualizations, using visual inference to see if similar patterns emerge.
+More research is needed to apply these methods to larger datasets, different layout algorithms, and different ways of visualizing network data, such as adjacency matrix visualizations, using visual inference to see if similar findings emerge.
 

### 2019-07-11 ASApaper/paper-child.Rnw
making changes per hh comments from a while ago

@@
 % }
-When selecting and fitting statistical models, there are typically three key assessments of interest: significance tests of parameters, goodness-of-fit tests, and power calculations of the tests.
+When selecting and fitting statistical models, there are typically three key assessments of interest: significance tests of parameters, goodness-of-fit tests, and power calculations.
 With significance testing, the null hypothesis assumes that the data come from a simple model nested within the model of interest.
-We then conduct significance tests of one or more additional parameters to determine if they can explain some of the variability in the data.
-For goodness-of-fit tests, we examine one or more models of interest to assess how well they fit the data.
+We then conduct significance tests of one or more additional parameters to determine how much of the variability in the data they explain.
+For goodness-of-fit tests, we examine one or more models of interest to assess how well these models explain the data.
 Power then quantifies the ability of the hypothesis test to detect the difference between the null and alternative hypothesis.
-All three of these elements of statistical modeling are vital for inference.
+All three of these elements of statistical modeling are vital to ensure that we can draw valid conclusions from a model.
 %The more complicated the model, the harder it is to determine fit or to include or exclude a parameter.
 
-The more complex the model, however, the harder it is to assess significance, power, and goodness-of-fit.
+The more complex the model, however, the harder it typically is to assess significance, power, and goodness-of-fit.
 One particularly complicated family of models are those designed to model networks.
@@
 dependencies inherent to network data make them  difficult to model. %Even more challenging is the situation 
-This difficulty only increases when studying \textit{dynamic networks}, the same set of nodes and their changing relationships observed at many points in time, because of the added temporal dependence.
+This difficulty increases further when studying \textit{dynamic networks}, the same set of nodes and their changing relationships observed at many points in time, because of the added temporal dimension.
 Yet, researchers are often interested in modeling dynamic social networks, such as friendship networks among students or collaboration networks between legislators.
@@
 Even some of the simplest network models, however, lack the asymptotics required to perform traditional goodness-of-fit tests \citep{holland81}.
-In addition, direct maximum likelihood estimation of model parameters is frequently impossible due to the intractability of the models \citep{hummeletal}.
+In addition, a direct maximum likelihood estimation of model parameters is frequently impossible due to the intractability of the models \citep{hummeletal}.
 %\hh{is the previous sentence a quote? I'm not sure that I follow.} %\st{it's a paraphrase... shall I make a quote? Also I definitely botched a word. Hopefully makes more sense now? }
@@
 
-In order to circumvent some of these difficulties, we propose new methods for significance testing of parameters, goodness-of-fit testing, and power calculations %of these 
+In order to circumvent some of these difficulties, we propose a new approach for significance testing of parameters, goodness-of-fit testing, and power calculations %of these 
 for one family social network models: continuous time Markov-chain (CTMC) models for dynamic network data, as described in \citet{saompaper}.
 We use \textit{visual inference} in place of traditional statistical methods for social network models, such as $t$-tests for significance of parameters or outdegree distributions for goodness-of-fit tests.
-Visual inference (VI), introduced by \citet{Bujaetal}, will allow us to look at the \textit{entire} dataset simulated from a network model, whereas traditional methods use one-dimensional metrics derived from the network or a $p$-value for a parameter in the model.
+Visual inference (VI), introduced by \citet{Bujaetal}, allows us to look at the \textit{entire} dataset simulated from a network model, whereas traditional methods use one-dimensional metrics derived from the network or a $p$-value for a parameter in the model.
 By using VI to supplement traditional statistical tests, we gain insight into the role of the parameters in these CTMC models, and we gain the ability to assess the fit of the CTMCs to dynamic network data.%In addition, we use visual inference to better understand some parameters in a social network model, by looking at how different values for parameters affect the networks simulated from these models.
@@
 Full details on these models can be found in \citet{saompaper, snijders01, snijders:2010, snijders2010, snijdersetall:2007, snijders:2017}.
-These CTMCs are a family of models for dynamic network data that incorporate both network structure and node-level covariates to describe how a network changes over time \citep{saompaper}.
+CTMCs are a family of models for dynamic network data that incorporate both network structure and node-level covariates to describe how a network changes over time \citep{saompaper}.
 Traditional network models, such as exponential random graph models, usually only consider network structure.
-Social networks are ever-changing as relationships decay or grow in seemingly random ways, and the actors in them have characteristics that affect how they change their ties to other nodes in the network.
+Social networks are ever-changing as relationships decay or grow over time, and each actor in a network has characteristics that affect ties to other actors in the network.
 % Talk about SAOMs
@@
 CTMC models use network structure and node covariates to model the network change one tie at a time.
-There are two model pieces: a \textit{rate function} that dictates \textit{how often} changes in the network occur and an \textit{objective function} that determines \textit{what} those changes are.
-The variables of interest are the edges of the network, $x_{ij}$ between nodes $i$ and $j$, where $i \neq j \in \{1, 2, \dots, n\}$ and $n$ the number of nodes in the network.
+There are two model pieces: a \textit{rate function} that dictates \textit{how often} changes in the network occur and an \textit{objective function} that determines \textit{what} those changes are.% The variables of interest are the edges of the network, $x_{ij}$ between nodes $i$ and $j$, where $i \neq j \in \{1, 2, \dots, n\}$ and $n$ the number of nodes in the network.
 The $x_{ij}$ are directed and binary, i.e.
-  \begin{equation}\label{eq:edgevars}
-  x_{ij} =
-  \begin{cases}
-                                   1 & \text{if an edge from $i$ to $j$ exists} \\
-                                   0 & \text{otherwise}
-  \end{cases}
-  \end{equation}
-and self-referencing edges or loops are not allowed, i.e.  $x_{ii} = 0$ for all $i=1, 2, \dots, n$. 
-The network is observed $M \geq 2$ times at time points  $t_1 < ... < t_M$, and the entire network at time point $t_m$ is denoted as $x(t_m)$ for $m \in \{1, \dots, M\}$.
-
-\noindent \textbf{Rate Function:} All changes in the network variables $x_{ij}$ are treated as choices made by the actors in the network.
+%   \begin{equation}\label{eq:edgevars}
+%   x_{ij} =
+%   \begin{cases}
+%                                    1 & \text{if an edge from $i$ to $j$ exists} \\
+%                                    0 & \text{otherwise}
+%   \end{cases}
+%   \end{equation}
+% and self-referencing edges or loops are not allowed, i.e.  $x_{ii} = 0$ for all $i=1, 2, \dots, n$. 
+% The network is observed $M \geq 2$ times at time points  $t_1 < ... < t_M$, and the entire network at time point $t_m$ is denoted as $x(t_m)$ for $m \in \{1, \dots, M\}$.
+
+\noindent \textbf{Rate Function:} All changes in the network are treated as choices made by the actors in the network.
 The rate function dictates when changes are made and which actor can make them.
@@
 In general, the rate function can include structural and node covariate parameters into account so that each actor has a different rate of change.
-However, we choose a simple rate function that is constant over all nodes in a given time period, because we want to focus on interpreting the parameters of the objective function.
+However, we choose a simple rate function that is constant over all nodes in a given time period, because we focus on interpreting the parameters of the objective function.
 We denote the rate from $t_{m}$ to $t_{m+1}$ as $\alpha_m$ for $m = 1, \dots, M-1$.
 Using this notation, the \textit{waiting time} to the next chance for actor $i$ to make a change is exponentially distributed with expected value $\alpha_m^{-1}$.
-Since the rate is the same for all actors, the waiting time for \textit{any} actor to get the opportunity to change its set of ties is exponentially distributed with expected value $(n\alpha_m)^{-1}$.
+Since the rate is the same for all actors, the waiting time for \textit{any} actor to get the opportunity to change its set of ties is also exponentially distributed with expected value $(n\alpha_m)^{-1}$.
 
@@
 where $\boldsymbol{\beta} = (\beta_1, \dots, \beta_K)$ are additional model parameters, each associated with some statistics, $s_{i1}(x, \mathbf{Z}), \dots, s_{iK}(x, \mathbf{Z})$, calculated for actor $i$ at the current network state $x$.
-The statistics can be very simple, like outdegree of the node, $s_i(x) = \sum_{i\neq j} x_{ij}$, or more complicated, like the {\it transitive triplets jumping to different covariate} statistic, $s_i(x, \mathbf{Z}) = \sum_{i \neq j \neq h} x_{ij}x_{ih}x_{hj} \cdot \mathbb{I}(z_i = z_h \neq z_j)$.
-%, plus many more.
+T%he statistics can be very simple, like outdegree of the node, $s_{i1}(x) = \sum_{i\neq j} x_{ij}$, or more complicated, like the {\it transitive triplets jumping to different covariate} statistic, $s_i(x, \mathbf{Z}) = \sum_{i \neq j \neq h} x_{ij}x_{ih}x_{hj} \cdot \mathbb{I}(z_i = z_h \neq z_j)$.
+At least two parameters must be included in the objective function: density and reciprocity \citep{RSienamanual}.
+We denote the density, or out-degree, parameter by $\beta_1$ and the associated statistic as $s_{i1}(x)$ and we denote the reciprocity parameter by $\beta_2$ and the associated statistic as $s_{i2}(x) = \sum_{j} x_{ij}x_{ji}$.
+We will refer to the very simple model with only these two parameters in the objective function as model M1.
+We define additional parameters and models of interest in Section~\ref{sec:models}  %, plus many more.
 At last count, in the software we use to fit these models to network data,
-Version 1.2-3 of \texttt{RSiena} \citep{RSiena}, the software used to fit the models here, provides over 80 possible effects that can be included in the objective function.
-We discuss these and other statistics in more detail in Section~\ref{sec:models}.
+Version 1.2-3 of \texttt{RSiena} \citep{RSiena}, the software we use to fit CTMC models to data, provides over 80 possible effects that can be included in the objective function.
+%We discuss these and other statistics in more detail in Section~\ref{sec:models}.
 
@@
 \begin{equation}\label{eq:transprob}
-p_{ij} = \dfrac{\exp\{f_i(x(i \leadsto j), \boldsymbol{\beta}, \mathbf{Z})\}}{\sum_{h} \exp\{f_i(x(i \leadsto h), \boldsymbol{\beta}, \mathbf{Z})\}}
+p_{ij} = \dfrac{\exp\{f_i(x(i \leadsto j), \boldsymbol{\beta}, \mathbf{Z})\}}{\sum_{h} \exp\{f_i(x(i \leadsto h), \boldsymbol{\beta}, \mathbf{Z})\}},
 \end{equation}
-This probability dictates which edge change is made by the node $i$.
-Node $i$ can also choose to \textit{not} change at all, which is most likely when the value of Equation~\ref{eq:objective} is larger for $x$ than for any $x(i \leadsto j)$.
+dictating which edge node $i$ changes.
+%Node $i$ can also choose to \textit{not} change at all, which is most likely when the value of Equation~\ref{eq:objective} is larger for $x$ than for any $x(i \leadsto j)$.
 
 %\hh{XXX how is the probability computed in the case that no change is made? Why is the notation $j \equiv i$ used for no change?}\st{that's more for the computation. I'll adjust accordingly}
-
-According to \citet{RSienamanual}, at least two parameters must be included in the objective function: density and reciprocity.
-We denote the density, or out-degree, parameter by $\beta_1$ and the associated statistic as $s_{i1}(x) = \sum_{j} x_{ij}$.
-Similarly, we denote the reciprocity parameter by $\beta_2$ and the associated statistic as $s_{i2}(x) = \sum_{j} x_{ij}x_{ji}$.
-We will refer to the very simple model with only these two parameters in the objective function as model M1.
-We define additional models of interest in Section~\ref{sec:models}
 
@@
 Here, we use the lineup protocol to design significance, goodness-of-fit (GoF) and power tests.
-Under the protocol, we begin with a data set of interest to us, such as a network, a visualization of this data, such as a node-link diagram, and a model of interest that we think generated this data.
-There are two hypotheses considered: the null hypothes that the model of interest generated the data, and the alternative hypothesis that the data were not generated under this model.
-Then, $P-1$ sets of data are simulated from the null model, where $P$ is the lineup size.
+Under the protocol, we begin with a data set of interest to us, such as a network, a visualization of this data, such as a node-link diagram, and a model of interest.
+We consider two hypotheses: the null hypothesis that the model of interest generated the data, and the alternative hypothesis that the data were not generated under this model.
+To construct a lineup of size $P$, $P-1$ sets of data are simulated from the null model.
 Each of the $P-1$ simulated datasets are visualized in the same way as the data, and the plot of the data is placed randomly among the set of $P-1$ \textit{null plots}.
-Human observers then examine the lineup and identify the plot(s) that look most different from the others.
+Human observers then examine the lineup and identify the plot(s) that look(s) most different from the others.
 If an observer identifies the data plot, this is evidence against the null hypothesis.
-Since an observer has a chance of $\frac{1}{P}$ to pick the data plot from the lineup by simply guessing, the evidence grows in strength with the number of independent observers identifying the data plot.
+An observer has a chance of $P^{-1}$ to pick the data plot from the lineup by simply guessing.
+The evidence grows in strength with the number of independent observers identifying the data plot.
 
@@
 %In the experimental data that we analyzed the null generating methods used were permutation methods and direct simulation from a null model.
-The lineup protocol was formally tested in a head-to-head comparison with the equivalent conventional test in \citet{majumder:2011}.
+The lineup protocol was formally tested for linear models in a head-to-head comparison with the equivalent conventional test in \citet{majumder:2011}.
 The experiment utilized human subjects from Amazon's Mechanical Turk \citep{turk} and used simulation to control conditions.
-The results suggest that VI done in a controlled setting gives the same results as conventional tests.
+The results suggest that VI done in a controlled setting gives similar results to conventional tests.
 This is evidence that VI can be used when no conventional tests exist or when testing is difficult.
@@
 
-\noindent \textbf{Data:} The data we use are collaboration networks in the United States Senate during the $111^{th}$ through $114^{th}$ Congresses.
+The data we use are collaboration networks in the United States Senate during the $111^{th}$ through $114^{th}$ Congresses.
 These senates began on January 6, 2009, the start date of the $111^{th}$ Congress, and ended on January 3, 2017, the last date of the $114^{th}$ Congress Details of how this data can be downloaded are provided by François Briatte at \url{github.com/briatte/congress}.
-In the US Senate, senators often show support for a piece of legislation by cosponsoring a bill authored by one of their colleagues.
-In a cosponsorship network, ties are directed from senator $i$ to senator $j$ when senator $i$ signs on as a cosponsor to the bill that senator $j$ authored.
-There are many hundreds of ties between senators when they are connected in this way, so we simplify the network by computing a single value for each senator-senator collaboration called the \textit{weighted propensity to cosponsor} (WPC).
+In the US Senate, senators often show support for a piece of legislation by co-sponsoring a bill authored by one of their colleagues.
+In a co-sponsorship network, ties are directed from senator $i$ to senator $j$ when senator $i$ signs on as a co-sponsor to the bill that senator $j$ authored.
+There are many hundreds of ties between senators when they are connected in this way, so we simplify the network by computing a single value for each senator-senator collaboration called the \textit{weighted propensity to co-sponsor} (WPC).
 This value is defined in \citet{senate} as
@@
 \begin{equation}\label{eq:sen1}
-    WPC_{ij} = \dfrac{\sum\limits_{b=1}^{B_j} \frac{Y_{ij(b)}}{c_{j(b)}}}{\sum\limits_{b=1}^{B_j} \frac{1}{c_{j(b)}}}
+    WPC_{ij} = \sum\limits_{b=1}^{B_j} \frac{Y_{ij(b)}}{c_{j(b)}}\left(\sum\limits_{b=1}^{B_j} \frac{1}{c_{j(b)}}\right)^{-1}
 \end{equation}
-where $B_j$ is the number of bills in a congressional session authored by senator $j$, $c_{j(b)}$ is the number of cosponsors on senator $j$'s $b^{th}$ bill, where $b \in \{1,\dots, B_j\}$, and $Y_{ij(b)}$ is an indicator variable that senator $i$ cosponsored senator $j$'s $b^{th}$ bill.
-This measure ranges in value from 0 to 1, where $WPC_{ij} = 1$ if senator $i$ is a cosponsor on every one of senator $j$'s bills and $WPC_{ij} = 0$ if senator $i$ is never a cosponsor any of senator $j$'s bills.
+where $B_j$ is the number of bills in a congressional session authored by senator $j$, $c_{j(b)}$ is the number of co-sponsors on senator $j$'s $b^{th}$ bill, where $b \in \{1,\dots, B_j\}$, and $Y_{ij(b)}$ is an indicator variable that senator $i$ co-sponsored senator $j$'s $b^{th}$ bill.
+This measure ranges in value from 0 to 1, where $WPC_{ij} = 1$ if senator $i$ is a co-sponsor on every one of senator $j$'s bills and $WPC_{ij} = 0$ if senator $i$ is never a co-sponsor any of senator $j$'s bills.
 Because CTMC models require binary edges, we construct the edges as follows:
@@
 The node-link diagram for the senate network is shown in Figure~\ref{fig:senateAll}.
-We have labelled some of the nodes in the network whose names will be familiar to US readers, either because they are leaders in their party or they have run for president.
+We have labelled some of the nodes in the network whose names we think will be familiar the reader, either because they are leaders in their party or they have run for president.
 The size of the nodes represents how many bills the senator authored in a session, the color represents party affiliation, and the shape represents gender.
@@
 
-<<senateAll, fig.cap="The US senate collaboration network observed at four time points.
+<<senateAll, fig.height = 5, fig.cap="The US senate collaboration network observed at four time points.
 Color represents party, shape represents gender, and size represents number of bills authored in a session.
@@
 Drawn with the \\texttt{geomnet} R package.
-\\citep{geomnet, rstats}">>=
+\\citep{geomnet}">>=
 seobama <- read_csv("data/senateobamapres_gsw_25.csv")
@@
   "Lindsey O. Graham","Rand Paul", "Bernard Sanders", "Jim Webb", "Mitch McConnell",
-  "Harry M. Reid", "Hillary Rodham Clinton", "Amy Jean Klobuchar", "Elizabeth Warren")
+  "Harry M. Reid", "Amy Jean Klobuchar", "Elizabeth Warren")
 seobama$label <- as.factor(ifelse(seobama$source %in% to_label, seobama$source, ""))
 levels(seobama$label) <- c("", "Amy Klobuchar", "Bernie Sanders", "Elizabeth Warren",
-                           "Harry Reid", "Hillary Clinton", "Jim Webb", "John McCain", 
+                           "Harry Reid", "Jim Webb", "John McCain", 
                            "Joe Biden","Lindsey Graham", "Marco Rubio", "Mitch McConnell", 
@@
   scale_size_continuous(name = "Bills\nauthored", range = c(.75, 3)) + 
-  theme(legend.position = 'bottom', strip.text = element_text(size = 8)) + 
+  theme(strip.text = element_text(size = 8)) + 
   facet_wrap(~senate, nrow = 2, labeller = "label_both") + 
@@
 %For Senate 111, for instance, we see Hillary Clinton, serving out her second term in the senate until she became Secretary of State.
-She is isolated in Figure~\ref{fig:senateAll}, but in actuality, she had many cosponsors on only two pieces of legislation she authored in that short time.%, as is shown in Figure~\ref{fig:senateClinton}.
+She is isolated in Figure~\ref{fig:senateAll}, but in actuality, she had many co-sponsors on only two pieces of legislation she authored in that short time.%, as is shown in Figure~\ref{fig:senateClinton}.
 %We chose to remove Clinton and her edges from the network because they make the overall structure look vastly more connected than the other three senates.
-We suspect that because she had just recently run for President and had been selected by President-Elect Obama to be Secretary of State, the cosponsorships she garnered at the start of the new congress were largely symbolic, so the $111^{th}$ Senate without Hillary Clinton is more typical than the $111^{th}$ Senate with her.
+We suspect that because she had just recently run for President and had been selected by President-Elect Obama to be Secretary of State, the co-sponsorships she garnered at the start of the new congress were largely symbolic, so the $111^{th}$ Senate without Hillary Clinton is more typical than the $111^{th}$ Senate with her.
 In fact, she had several co-sponsors on both of her pieces of legislation, which meant she had the maximum WPC value of one with about a dozen other senators.
@@
 
-In legislative cosponsorship networks, it is known that party affiliation, reciprocity of relationships, and whether senators are ``workhorses" who author many bills or ``showhorses" who author few bills, are major influences on structure \citep{legnet}.
+In legislative co-sponsorship networks, it is known that party affiliation, reciprocity of relationships, and whether senators are ``workhorses" who author many bills or ``showhorses" who author few bills, are major influences on structure \citep{legnet}.
 We focus on these covariates, plus the additional gender covariate when choosing which CTMC models to fit to the data.
 
-\noindent \textbf{Models:} %In order to determine which models to examine, we first consider already well-known effects in legislative networks.
-We also selected some effects which were significant according to the built-in Wald-type tests in \texttt{RSiena} for application of our significance and goodness-of-fit methods.
-We examine a total of six models, each identified by its objective function below. The parameters in the objective function, $\beta_1, \dots, \beta_6$ are defined in Table~\ref{tab:effects}.
-\begin{enumerate}
-\item Model M1: $f_{i}(x, \boldsymbol{\beta}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x)$
-\item Model M3: $f_{i}(x, \boldsymbol{\beta}, \mathbf{p}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_3 s_{i3}(x, \mathbf{p})$
-\item Model M4: $f_{i}(x, \boldsymbol{\beta}, \mathbf{s}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_4 s_{i4}(x, \mathbf{s})$
-\item Model M5: $f_{i}(x, \boldsymbol{\beta}, \mathbf{b}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_5 s_{i5}(x, \mathbf{b})$
-\item Model M6: $f_{i}(x, \boldsymbol{\beta}, \mathbf{p}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_6 s_{i6}(x, \mathbf{p})$
-\item Model M7: $f_{i}(x, \boldsymbol{\beta}, \mathbf{p}, \mathbf{b}, \mathbf{s}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_4 s_{i4}(x, \mathbf{s}) + \beta_5 s_{i5}(x, \mathbf{b}) + \beta_6 s_{i6}(x, \mathbf{p})$
-\end{enumerate}
+\begin{table}
+\centering
+\scalebox{0.7}{
+\begin{tabular}{c|p{2cm}|p{2.1cm}|c|l|p{1.75cm}|p{2cm}}
+\toprule
+$\beta_k$ & Effect name & Interaction Variable & Formula & Picture &  Initial estimate &  Wald $p$-value \\
+\midrule
+$\beta_1$ & density & -- & $s_{i1}(x) = \sum_{j} x_{ij}$ & \includegraphics[width=.6in]{img/density.png} & 2.204 & NA \\ 
+$\beta_2$ & reciprocity & -- & $s_{i2}(x) = \sum_{j} x_{ij}x_{ji}$ & \includegraphics[width=.6in]{img/recip.png} & -4.903 & NA \\
+$\beta_3$ & jumping transitive triplet & party & $s_{i3}(x, \mathbf{p}) = \sum_{j\neq h} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(p_i = p_h \neq p_j)$ & \includegraphics[width=.6in]{img/jttp.png} & -5.884 & $<0.0001$\\
+$\beta_4$ & jumping transitive triplet & sex & $s_{i4}(x, \mathbf{s}) = \sum_{j\neq h} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(s_i = s_h \neq s_j)$ & \includegraphics[width=.6in]{img/jtts.png}  & 3.335 & 0.0002 \\
+$\beta_5$ & similarity transitive triplet & bills & $s_{i5}(x, \mathbf{b}) = \sum_{j} x_{ij}x_{ih}x_{hj}\cdot (sim^b_{ij} - \overline{sim}^b)^*$ & \includegraphics[width=1in]{img/simttb.png} & 9.821 & 0.0128 \\
+$\beta_6$ & same transtive triplet & party & $s_{i6}(x, \mathbf{p}) =\sum_{j} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(p_i = p_j)$ & \includegraphics[width=1in]{img/samettp.png}  & 1.306 & 0.0642 \\
+\bottomrule
+%\\
+%$\beta_7$ & same & party & $s_{i7}(x, \mathbf{p}) = \sum_j x_{ij}\mathbb{I}(p_i = p_j)$ & \includegraphics[width=.6in]{img/samep.png}  & 0.363 & 0.0074
+\end{tabular}
+}
+\caption{\label{tab:effects} The effects we used in the CTMC models fit to the senate data.
+In the picture which represents each effect, the dotted tie is encouraged to form if the estimate is positive, and discouraged to form if the estimate is negative.
+$^*$ - $sim^b_{ij}$ is defined in Equation~\ref{eq:similar} and $\overline{sim}^b = \frac{1}{n(n-1)}\sum_{i\neq j} sim^b_{ij}$ is the average bill similarity score between two senators.}
+\end{table}
+
+\begin{table}
+\centering
+\begin{tabular}{lcccccc}
+\toprule 
+Model & $s_{i1}$ & $s_{i2}$ & $s_{i3}$ & $s_{i4}$ & $s_{i5}$ & $s_{i6}$ \\
+\midrule
+M1 & 	\checkmark & 	\checkmark & -- & -- & -- & -- \\
+M3 & \checkmark & 	\checkmark & \checkmark & -- & -- & -- \\
+M4 & \checkmark & 	\checkmark & -- & \checkmark & -- & -- \\
+M5 & \checkmark & 	\checkmark & -- & -- & \checkmark & -- \\
+M6 & \checkmark & 	\checkmark & -- & -- & -- & \checkmark \\
+M7 & \checkmark & 	\checkmark & \checkmark & \checkmark & \checkmark & \checkmark \\
+\bottomrule
+\end{tabular}
+\caption{\label{tab:models}The models we use defined by the statistics in their objective functions. The corresponding parameter for $s_{ik}$ is $\beta_k$. Note there is no model M2.}
+\end{table}
+% 
+% \begin{enumerate}
+% \item Model M1: $f_{i}(x, \boldsymbol{\beta}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x)$
+% \item Model M3: $f_{i}(x, \boldsymbol{\beta}, \mathbf{p}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_3 s_{i3}(x, \mathbf{p})$
+% \item Model M4: $f_{i}(x, \boldsymbol{\beta}, \mathbf{s}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_4 s_{i4}(x, \mathbf{s})$
+% \item Model M5: $f_{i}(x, \boldsymbol{\beta}, \mathbf{b}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_5 s_{i5}(x, \mathbf{b})$
+% \item Model M6: $f_{i}(x, \boldsymbol{\beta}, \mathbf{p}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_6 s_{i6}(x, \mathbf{p})$
+% \item Model M7: $f_{i}(x, \boldsymbol{\beta}, \mathbf{p}, \mathbf{b}, \mathbf{s}) = \beta_1 s_{i1}(x) + \beta_2 s_{i2}(x) + \beta_4 s_{i4}(x, \mathbf{s}) + \beta_5 s_{i5}(x, \mathbf{b}) + \beta_6 s_{i6}(x, \mathbf{p})$
+% \end{enumerate}
 
@@
 %The parameters we use for the remainder of the paper are detailed in Table~\ref{tab:effects}.
-All models have the first two structural parameters, $\beta_1 and \beta_2$, included.
+
+\noindent \textbf{Models:} %In order to determine which models to examine, we first consider already well-known effects in legislative networks.
+We also selected some effects which were significant according to the built-in Wald-type tests in \texttt{RSiena} for application of our significance and goodness-of-fit methods.
+We examine a total of six models, each identified by its objective function.
+The parameters in the objective function, $\beta_1, \dots, \beta_6$ are defined in Table~\ref{tab:effects}.
+All models have the outdegree and reciprocity parameters, $\beta_1$ and $\beta_2$, included.
 Other effects were added one at a time so that the simplest model $M1$ is nested in each of the other models.
@@
 This parameter also considers transitive closures, but for senators of different genders.
-The positive value indicates that transitive ties between senators of different genders are more likley to form.
+The positive value indicates that transitive ties between senators of different genders are more likely to form.
 Next, we consider $\beta_5$, the covariate-related similarity score-weighted transitive triplets parameter for the number of bills authored by a senator.
@@
 The high positive estimate suggests senators are encouraged to collaborate with other senators who author about the same number of bills they do.
-This tendency of senators to cosponsor bills written by senators who are similarly ``prolific" corresponds to the tendency of senators to be either ``workhorses" or ``showhorses".
+This tendency of senators to co-sponsor bills written by senators who are similarly ``prolific" corresponds to the tendency of senators to be either ``workhorses" or ``showhorses".
 Senators known as workhorses author many pieces of legislation in a session, and largely stay out of the public arena.
@@
 
-\begin{table}
-\scalebox{0.8}{
-\begin{tabular}{c|p{2cm}|p{2.1cm}|c|l|p{1.75cm}|p{2cm}}
-$\beta_k$ & {\bf Effect name} & {\bf Interaction Variable} & {\bf Formula} & {\bf Picture} & {\bf Initial estimate} & {\bf Wald $p$-value} \\
-\hline 
-$\beta_1$ & density & -- & $s_{i1}(x) = \sum_{j} x_{ij}$ & \includegraphics[width=.6in]{img/density.png} & 2.204 & NA \\ 
-$\beta_2$ & reciprocity & -- & $s_{i2}(x) = \sum_{j} x_{ij}x_{ji}$ & \includegraphics[width=.6in]{img/recip.png} & -4.903 & NA \\
-$\beta_3$ & jumping transitive triplet & party & $s_{i3}(x, \mathbf{p}) = \sum_{j\neq h} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(p_i = p_h \neq p_j)$ & \includegraphics[width=.6in]{img/jttp.png} & -5.884 & $<0.0001$\\
-$\beta_4$ & jumping transitive triplet & sex & $s_{i4}(x, \mathbf{s}) = \sum_{j\neq h} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(s_i = s_h \neq s_j)$ & \includegraphics[width=.6in]{img/jtts.png}  & 3.335 & 0.0002 \\
-$\beta_5$ & similarity transitive triplet & bills & $s_{i5}(x, \mathbf{b}) = \sum_{j} x_{ij}x_{ih}x_{hj}\cdot (sim^b_{ij} - \overline{sim}^b)^*$ & \includegraphics[width=1in]{img/simttb.png} & 9.821 & 0.0128 \\
-$\beta_6$ & same transtive triplet & party & $s_{i6}(x, \mathbf{p}) =\sum_{j} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(p_i = p_j)$ & \includegraphics[width=1in]{img/samettp.png}  & 1.306 & 0.0642 %\\
-%$\beta_7$ & same & party & $s_{i7}(x, \mathbf{p}) = \sum_j x_{ij}\mathbb{I}(p_i = p_j)$ & \includegraphics[width=.6in]{img/samep.png}  & 0.363 & 0.0074
-\end{tabular}
-}
-\caption{\label{tab:effects} The effects we used in the CTMC models fit to the senate data.
-In the picture which represents each effect, the dotted tie is encouraged to form if the estimate is positive, and discouraged to form if the estimate is negative.
-$^*$ - $sim^b_{ij}$ is defined in Equation~\ref{eq:similar} and $\overline{sim}^b = \frac{1}{n(n-1)}\sum_{i\neq j} sim^b_{ij}$ is the average bill similarity score between two senators.}
-\end{table}
-
 We fit all six of our models in \texttt{RSiena} using Markov Chain Monte Carlo (MCMC) methods to approximate the method of moments estimates of the parameters.
 Because the estimation is done through MCMC simulation, we fit each model to the data 1,000 times.
-From the simulations in which the MC converged, which made up over 90\% of the fits for each model, we computed the mean of the the estimates of each parameter to get final estimates of $\hat{\boldsymbol{\beta}}$ for each model, which are given in Table~\ref{tab:fittedvalues}.
+From the simulations in which the MC converged, which made up over 90\% of the fits for each model, we computed the mean of the estimates of each parameter to get final estimates of $\hat{\boldsymbol{\beta}}$ for each model, which are shown in Table~\ref{tab:fittedvalues}.
 

### 2019-07-11 ASApaper/paper.Rnw
making changes per hh comments from a while ago

@@
 \usepackage{numprint}
+\usepackage{booktabs}
 
@@
 With VI, we can look at entire datasets simulated from a model, instead of relying a single metric such as a $p$-value.
-We conducted a VI experiment, with participants recruited via Amazon Mechanical Turk, to assess the significance of CTMC parameters, the fit of a CTMC model, and the visual power of CTMC parameters.
+We conducted a VI experiment, with participants recruited via Amazon Mechanical Turk, to assess the significance of CTMC parameters, the fit of a CTMC model, and the visual power of CTMC parameters, and found that ...
+TBD
 \end{abstract}

### 2019-07-11 ASApaper/paper-child.Rnw
more smol changes. struggling today.

@@
 \centering
-\begin{tabular}{|l|ccccccccc|}
-\hline
+\begin{tabular}{|lccccccccc|}
+\\toprule
 Model & $\hat{\alpha}_1$ & $\hat{\alpha}_2$ & $\hat{\alpha}_3$ & $\hat{\beta}_1$ & $\hat{\beta}_2$ & $\hat{\beta}_3$ & $\hat{\beta}_4$ & $\hat{\beta}_5$ & $\hat{\beta}_6$ \\
-\hline
-\hline
+\midrule
 M1 & \Sexpr{round(modelmeans$alpha1[1], 3)} & \Sexpr{round(modelmeans$alpha2[1], 3)} & \Sexpr{round(modelmeans$alpha3[1], 3)} & \Sexpr{round(modelmeans$beta1[1], 3)} & \Sexpr{round(modelmeans$beta2[1], 3)} & -- & -- & -- & --\\
@@
 M7 & \Sexpr{round(M6[1], 3)} & \Sexpr{round(M6[2], 3)} & \Sexpr{round(M6[3], 3)} & \Sexpr{round(M6[4], 3)} & \Sexpr{round(M6[5], 3)} & -- & \Sexpr{round(M6[6], 3)} & \Sexpr{round(M6[8], 3)} & \Sexpr{round(M6[7], 3)} \\
-\hline
+\bottomrule
 \end{tabular}
@@
 Typically, a lineup shows sets of 20 plots at a time, for example in \citet{loy:2015, vanderplas:2016}, but we determined that looking at 20 node-link diagrams at once is too difficult.
-We chose to present our participants with only six plots at a time in order to show the node-link diagrams in more detail and to reduce their cognitive load.
+We chose to present our participants with only six plots at a time in order to show the node-link diagrams in more detail and to reduce cognitive load.
 To construct a lineup, we simulate five networks from the null model and one network from the alternative model.
-An example of a lineup like those shown to our participants is givenå in Figure~\ref{fig:shiny}.
-In this lineup, model M4 is the alternative model (one plot), and model M1 is the null model (five plots).
+Several lineups shown to our participants are presented and discussed in Section~\ref{sec:res}.
+%In this lineup, model M4 is the alternative model (one plot), and model M1 is the null model (five plots).
 
 To simulate lineups from the models we used the \texttt{siena07} function in \texttt{RSiena} \citep{RSiena}.
-For the purposes of our experiment, we focus on simulating the second "wave" of data, the $112^{th}$ Senate network, and we condition on the first wave of data, $111^{th}$ Senate network.
+For the purposes of our experiment, we focus on simulating the second ``wave" of data, the $112^{th}$ Senate network, and we condition on the first wave of data, $111^{th}$ Senate network.
 Sections~\ref{sec:setupst} through~\ref{sec:setupvp} describe in detail how we constructed the lineups, which parameter values we simulate from, and why.
@@
 If an observer picks out the alternative data plot, that is evidence against the null hypothesis, while picking one of the null plots is evidence in favor of the null hypothesis.
-To avoid exhausting our participants, we chose only two parameters to test, $\beta_3$ and $\beta_4$.
-Thus, we compare the null model M1 to the alternative models M3 and M4, using three repetions of each hypothesis test.
+To avoid over-working our participants, we chose to test only two parameters, $\beta_3$ and $\beta_4$.
+These two had the smallest $p$-values from the significance tests, so we chose them for the visual significance test as well because we hypothesized they would be easier to pick out of the lineup.
+Thus, we compare the null model M1 to the alternative models M3 and M4, using three repetions of each hypothesis test in the experiment.
 
@@
 
-For the goodness-of-fit tests, we compare one model of interest, say M\textit{i} to the data. The hypothesis we use to generate lineups are 
+For the goodness-of-fit tests, we compare one model of interest to the data. The hypothesis we use to generate lineups are 
 \begin{description}
-\item $H_0$: The data come from model M\textit{i}
+\item $H_0$: The data come from the model of interest
 \item $H_A$: The data come from some other, unknown model
@@
 
-To generate the null plots, we simulate five networks from model M\textit{i} using the corresponding parameters in Table~\ref{tab:fittedvalues}.
-We pick a wave to focus on, wave two, which is the first simulated network, and among these five plots, we place a node-link diagram of the true second wave data.
+To generate the null plots, we simulate five networks the model of interest using the corresponding parameters in Table~\ref{tab:fittedvalues}.
+We pick a wave to focus on, wave two, which is the first simulated network, and among these five plots, we place a node-link diagram of the true second wave of data.
 We cannot show the data more than once to each participant.
@@
 <<hypothesis, fig.height=2, fig.cap="As the parameter increases in absolute value, more viewers of the lineup should pick the alternative data out of a lineup.
-Note that the significance test we construct in Section~\\ref{sec:setupst} is just one point on the line below, represented by the vertical dotted line labeled $\\hat{\\beta}$.
+Note that the significance test we construct in Section~\\ref{sec:setupst} is just one point on the curve, represented by the vertical dotted line labeled $\\hat{\\beta}$.
 The easy, medium, and hard lines represent how we determined which values of the parameters to show to our participants, and the horizontal dotted line shows the type-I error for one viewer of a lineup of size 6.">>=
@@
   geom_text(data = NULL, aes(x = -230, y = .5, label = "medium"), angle = 90, vjust = -.5, size = 3) +
-  geom_text(data = NULL, aes(x = -175, y = .05, label = "hard"), angle = 90, vjust = -.5, size = 3) + 
+  geom_text(data = NULL, aes(x = -175, y = .06, label = "hard"), angle = 90, vjust = -.5, size = 3) + 
   scale_y_continuous(name = "% detecting plot from\nthe alternative model", labels = paste0(c(0,25,50,75,100), "%")) + 
@@
 Figure~\ref{fig:hypothesis} shows a sketch of what the selection probability looks like hypothetically with varying effect size: the higher in absolute value the parameter is, the more likely participants are to select the alternative model out of the lineup.
-To determine the exact values of the six levels we want to test for each effect, we started with the estimates of the parameter at hand (see  Section~\ref{sec:models}), and used small negative and positive factors to determine at what point \textit{we} noticed the effect of the parameter in simulations from the changed models.
+To determine the exact values of the six levels we want to test for each effect, we started with the estimates of the parameter at hand (see  Table~\ref{tab:fittedvalues }), and used small negative and positive factors to determine at what point \textit{we} noticed the effect of the parameter in simulations from the changed models.
 In Figure~\ref{fig:hypothesis}, we demonstrate these values with the vertical lines labeled ``easy," ``medium," and ``hard." We expect most viewers to see the effects at the ``easy" values, and we expect very few, if any, viewers to see the effects at the ``hard" values.
@@
 We recruited 250 participants for our experiment through Amazon Mechanical Turk.
-Each participant was presented with some brief training material before beginning the experiment.
-Before we presented the lineups for the hypothesis tests, each person was shown two trial plots one where the alternative plot was the most different from the others due to its relatively \textit{complex} structure, while the other trial included an alternative plot that was most different from the others due to its comparatively \textit{simple} structure.
+Each participant was presented with some brief training material first, then before presenting the lineups for the hypothesis tests, each person was shown two trial plots one where the alternative plot was the most different from the others due to its relatively \textit{complex} structure, while the other trial included an alternative plot that was most different from the others due to its comparatively \textit{simple} structure.
 Only when participants were able to correctly identify the alternative plot from the trial lineups were they allowed to begin the experiment.
@@
 Each participant was randomly assigned 13 lineups to look at.
-They were asked to select one or more plots that they perceived as ``most different" from the others, and provide a reasoning for their choice: ``Most simple overall structure",``Most complex overall structure", or ``Other".
-If they selected other, they were required to describe their reasoning.
-The language in the reasoning is purposefully vague: we want participants to tell us what they see.
-We do not want to tell them what they ``should" be seeing.
-In this way, we can truly determine what  effects are \textit{noticeable}.
+They were asked to select one or more plots that they perceived as ``most different" from the others, and provide a reasoning for their choice: ``most simple overall structure",``most complex overall structure", or ``other".
+If they selected ``other", they were required to describe their reasoning.
+The language in the reasoning is purposefully vague to avoid contextual bias.
 
@@
 In this section, we present the results from our experiment and compare them to traditional statistical tests where applicable.
-Because each lineup shown to participants has only six plots, the probability of picking the data by chance is high, $\frac{1}{6}$, but if \textit{many} independent viewers pick out the data from the nulls, the evidence in favor of the alternative hypothesis becomes stronger.
+Because each lineup shown to participants has only six plots, the probability of picking the data by chance is high at 1 in 6, but if \textit{many} independent viewers pick out the data from the nulls, the evidence in favor of the alternative hypothesis becomes stronger.
 The $p$-values from the lineups were calculated using the \texttt{vinference} package by \citet{vinference}.
@@
             pv = map2_dbl(npick, n, pV, m=6, scenario=3))
-names(summ_sig) <- c("Lineup ID", "parameter", "\\# Alt. Model Picks", "Total Views", "p-value")
+names(summ_sig) <- c("Lineup ID", "parameter", "\\# Identified", "Total Views", "p-value")
 summ_sig$parameter <- as.factor(summ_sig$parameter)
@@
 xtable::print.xtable(xtable::xtable(summ_sig, caption = "Experiment results for the two parameters for which we performed significance tests.
+\\# Identified indicates the number of participants who selected the alternative model plot.
 There were three lineups for each parameter, so there are three results for each plot.", label = "tab:sigtesting", digits = c(0,0,0,0,0,5)), include.rownames = FALSE, sanitize.text.function=function(x){x})

### 2019-07-23 ASApaper/paper-child.Rnw
spell-checked, put pdf up for hh

@@
 
-The paper is outlined as follows: Section~\ref{subsec:saom} provides an introduction to the CTMC familty of models.
+The paper is outlined as follows: Section~\ref{subsec:saom} provides an introduction to the CTMC family of models.
 Section~\ref{subsec:vi} gives a basic overview of visual inference and the lineup protocol.
-In Section~\ref{sec:models}, we describe our example data and define our models of interst that we used to develop our VI methods.
+In Section~\ref{sec:models}, we describe our example data and define our models of interest that we used to develop our VI methods.
 Section~\ref{sec:expersetup} details how we designed a VI experiment for significance testing and goodness of fit procedures for CTMC models, and Section~\ref{sec:res} details the results of our experiment.
@@
 The node-link diagram for the senate network is shown in Figure~\ref{fig:senateAll}.
-We have labelled some of the nodes in the network whose names we think will be familiar the reader, either because they are leaders in their party or they have run for president.
+We labelled some of the nodes in the network whose names we think will be familiar the reader, either because they are leaders in their party or they have run for president.
 The size of the nodes represents how many bills the senator authored in a session, the color represents party affiliation, and the shape represents gender.
@@
 
-In legislative co-sponsorship networks, it is known that party affiliation, reciprocity of relationships, and whether senators are ``workhorses" who author many bills or ``showhorses" who author few bills, are major influences on structure \citep{legnet}.
+In legislative co-sponsorship networks, it is known that party affiliation, reciprocity of relationships, and whether senators are ``workhorses" who author many bills or ``show horses" who author few bills, are major influences on structure \citep{legnet}.
 We focus on these covariates, plus the additional gender covariate when choosing which CTMC models to fit to the data.
@@
 $\beta_5$ & similarity transitive triplet & bills & $s_{i5}(x, \mathbf{b}) = \sum_{j} x_{ij}x_{ih}x_{hj}\cdot (sim^b_{ij} - \overline{sim}^b)^*$ & \includegraphics[width=1in]{img/simttb.png} & 9.821 & 0.0128 \\
-$\beta_6$ & same transtive triplet & party & $s_{i6}(x, \mathbf{p}) =\sum_{j} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(p_i = p_j)$ & \includegraphics[width=1in]{img/samettp.png}  & 1.306 & 0.0642 \\
+$\beta_6$ & same transitive triplet & party & $s_{i6}(x, \mathbf{p}) =\sum_{j} x_{ij}x_{ih}x_{hj}\cdot \mathbb{I}(p_i = p_j)$ & \includegraphics[width=1in]{img/samettp.png}  & 1.306 & 0.0642 \\
 \bottomrule
@@
 \toprule 
-Model & $s_{i1}$ & $s_{i2}$ & $s_{i3}$ & $s_{i4}$ & $s_{i5}$ & $s_{i6}$ \\
+Model & $\beta_{1}$ & $\beta_{2}$ & $\beta_{3}$ & $\beta_{4}$ & $\beta_{5}$ & $\beta_{6}$ \\
 \midrule
@@
 \end{tabular}
-\caption{\label{tab:models}The models we use defined by the statistics in their objective functions. The corresponding parameter for $s_{ik}$ is $\beta_k$. Note there is no model M2.}
+\caption{\label{tab:models}The models we use defined by the parameters in their objective functions. The corresponding parameter for $s_{ik}$ is $\beta_k$. Note there is no model M2.}
 \end{table}
@@
 \end{equation}
-where $\max_{hk}|b_h - b_k|$ is the range of number of bills authored by senators, and $b_i$, $b_j$ are the number of bills authored by senators $i,j$ repectively in the senate period.
+where $\max_{hk}|b_h - b_k|$ is the range of number of bills authored by senators, and $b_i$, $b_j$ are the number of bills authored by senators $i,j$ respectively in the senate period.
 This effect was estimated at about 9.8 with standard error of 3.9.
 The high positive estimate suggests senators are encouraged to collaborate with other senators who author about the same number of bills they do.
-This tendency of senators to co-sponsor bills written by senators who are similarly ``prolific" corresponds to the tendency of senators to be either ``workhorses" or ``showhorses".
+This tendency of senators to co-sponsor bills written by senators who are similarly ``prolific" corresponds to the tendency of senators to be either ``workhorses" or ``show horses".
 Senators known as workhorses author many pieces of legislation in a session, and largely stay out of the public arena.
-The showhorse senators, on the other hand, author relatively few pieces of legislation, and tend to appear in the media very frequently.
+The show horse senators, on the other hand, author relatively few pieces of legislation, and tend to appear in the media very frequently.
 Finally, we found $\beta_6$, the same party transitive triplet effect to be significant, with a fitted value of 1.3 and standard error of 0.7, meaning that transitive relationships between senators tend to form when they are from the same party, exactly as we would expect in a legislative body in a country with deeply entrenched partisan divides.
@@
 \begin{tabular}{|lccccccccc|}
-\\toprule
+\toprule
 Model & $\hat{\alpha}_1$ & $\hat{\alpha}_2$ & $\hat{\alpha}_3$ & $\hat{\beta}_1$ & $\hat{\beta}_2$ & $\hat{\beta}_3$ & $\hat{\beta}_4$ & $\hat{\beta}_5$ & $\hat{\beta}_6$ \\
@@
 These two had the smallest $p$-values from the significance tests, so we chose them for the visual significance test as well because we hypothesized they would be easier to pick out of the lineup.
-Thus, we compare the null model M1 to the alternative models M3 and M4, using three repetions of each hypothesis test in the experiment.
+Thus, we compare the null model M1 to the alternative models M3 and M4, using three repetitions of each hypothesis test in the experiment.
 
@@
 For $\beta_4$ and $\beta_6$, null model M1 and the alternative model (M4 or M6) switch roles in the reversed lineups, i.e.
-five plots show data simluated from the laternative model and only one plot shows data from M1.}
+five plots show data simulated from the alternative model and only one plot shows data from M1.}
 \label{tab:experdetail}
@@
 \begin{equation}\label{eq:wald1}
-\frac{(\hat{\beta_k})^2}{var(\hat{\beta_k})},
+\frac{(\hat{\beta_k})^2}{var(\hat{\beta_k})} \sim \chi^2_1,
 \end{equation}
-which is compared to a Chi-square distribution with one degree of freedom \citep{RSienamanual}.
+\citep{RSienamanual}.
 Testing the significance of multiple parameters depends on the hypothesis we wish to test, and an $H \times K$ matrix, $\mathbf{A}$, must be appropriately designed to test the $H$ hypotheses of interest.
@@
 \begin{equation}\label{eq:wald2}
-(\mathbf{A}\boldsymbol{\hat{\beta}})' \hat{\Sigma}^{-1} \mathbf{A}\boldsymbol{\hat{\beta}},
+(\mathbf{A}\boldsymbol{\hat{\beta}})' \hat{\Sigma}^{-1} \mathbf{A}\boldsymbol{\hat{\beta}} \sim \chi^2_H,
 \end{equation}
-where $\hat{\Sigma}$ is the estimated covariance matrix of $\boldsymbol{\beta}$. This statistic is then compared to a Chi-square distribution with $H$ degrees of freedom.
+where $\hat{\Sigma}$ is the estimated covariance matrix of $\boldsymbol{\beta}$.
 
@@
 @
-A plot comparing the data to the simulations is also considered, and a similar plot is shown in Figure~\ref{fig:gofsiena} for the outdegree distribution of small data set, shown in the points and connected lines, with the simulated values of $\mathbf{u}_d$ shown in boxplots and overlaid violin plots.
+A plot comparing the data to the simulations is also considered, and a similar plot is shown in Figure~\ref{fig:gofsiena} for the outdegree distribution of small data set, shown in the points and connected lines, with the simulated values of $\mathbf{u}_d$ shown in box plots and overlaid violin plots.
 The \texttt{RSiena} software also provides a Rao score-type test for goodness-of-fit for assessing one or more parameters, the test statistic of which is compared to a Chi-square distribution with $H$ degrees of freedom, where $H$ has the same definition as in Section~\ref{sec:sigtest}.
 For full detail on the score-type test, see \citet{scoretest}.
-These methods are both restriced: the \texttt{sienaGOF} method only considers \textit{one} measure on the data and simulations from the model, while the score-type tests only consider \textit{subsets} of parameters, ``nuisance parameters" in \citet{scoretest}, not the entire set of parameters.
+These methods are both restricted: the \texttt{sienaGOF} method only considers \textit{one} measure on the data and simulations from the model, while the score-type tests only consider \textit{subsets} of parameters, ``nuisance parameters" in \citet{scoretest}, not the entire set of parameters.
 By using visual inference instead of more traditional statistical methods, we hope to perform a more \textit{holistic} goodness-of-fit test.
@@
 This supports our hypothesis shown in Figure~\ref{fig:hypothesis}.
-For $\beta_3$ and $\beta_5$, the slope of the fitted line is \textit{much} steeper for positive values of the parameter than for negative values, meaning that our participants perceived differences more often for postitive parameter values than for negative parameter values.
+For $\beta_3$ and $\beta_5$, the slope of the fitted line is \textit{much} steeper for positive values of the parameter than for negative values, meaning that our participants perceived differences more often for positive parameter values than for negative parameter values.
 This finding is similar to that of \citet{corrviz}, who found that people detect positive correlations better and at lower values than negative correlations.
@@
 At a value of $\beta_5 = -20$, however, the model predicts this same probability to be about 0.40.
-This again demonstrates that the \textit{presence} of structure is detected  more frequently and at smaller values than the absence of strucutre.
+This again demonstrates that the \textit{presence} of structure is detected  more frequently and at smaller values than the absence of structure.
 

### 2019-07-26 ASApaper/paper-child.Rnw
latest edits

@@
                   strip.background = element_rect(colour = "black", fill = "white"))
+turk22 <- read_csv("data/turk22-sig.csv")
 @
@@
 The more complex the model, however, the harder it typically is to assess significance, power, and goodness-of-fit.
-One particularly complicated family of models are those designed to model networks.
+\st{Citation?} One particularly complicated family of models are those designed to model networks.
 A \textit{network} is any set of things, such as people, or computers, that are connected in some way, through social relations or the Internet.
-In a social network, the people are the \textit{nodes} or \textit{actors}, and the relationships between the people are \textit{edges} or \textit{ties}.
-In this paper we use nodes and edges interchangeably with actors and ties, respectively.
+In a social network, people are \textit{nodes} or \textit{actors} connected by relationships, which are \textit{edges} or \textit{ties}.
+We use nodes and edges interchangeably with actors and ties, respectively.
 Models for networks are particularly complex, as
@@
 Even some of the simplest network models, however, lack the asymptotics required to perform traditional goodness-of-fit tests \citep{holland81}.
-In addition, a direct maximum likelihood estimation of model parameters is frequently impossible due to the intractability of the models \citep{hummeletal}.
+In addition, direct maximum likelihood estimation of model parameters is frequently impossible due to the intractability of the models \citep{hummeletal}.
 %\hh{is the previous sentence a quote? I'm not sure that I follow.} %\st{it's a paraphrase... shall I make a quote? Also I definitely botched a word. Hopefully makes more sense now? }
@@
 for one family social network models: continuous time Markov-chain (CTMC) models for dynamic network data, as described in \citet{saompaper}.
-We use \textit{visual inference} in place of traditional statistical methods for social network models, such as $t$-tests for significance of parameters or outdegree distributions for goodness-of-fit tests.
+We use \textit{visual inference} \st{to complement}%in place of
+traditional statistical methods for social network models, such as $t$-tests for significance of parameters or outdegree distributions for goodness-of-fit tests.
 Visual inference (VI), introduced by \citet{Bujaetal}, allows us to look at the \textit{entire} dataset simulated from a network model, whereas traditional methods use one-dimensional metrics derived from the network or a $p$-value for a parameter in the model.
@@
 % Talk about SAOMs
-CTMC models use the network structure and the node covariate information, and as such can be very complicated models.
-As network data are inherently complex, it can be difficult to interpret model parameters and their estimates.
-There are also many possible parameters to include in a CTMC, which makes parameter selection and goodness-of-fit testing challenging.
+CTMC models use the network structure and the node covariate information, which can lead to very complicated models.
+The model complexity and the inherent complexity of network data combine to make interpreting model parameters and their estimates veryh difficult.
+\st{citation?} There are also many possible parameters to include in a CTMC, which makes parameter selection and goodness-of-fit testing challenging.
 
 CTMC models use network structure and node covariates to model the network change one tie at a time.
-There are two model pieces: a \textit{rate function} that dictates \textit{how often} changes in the network occur and an \textit{objective function} that determines \textit{what} those changes are.% The variables of interest are the edges of the network, $x_{ij}$ between nodes $i$ and $j$, where $i \neq j \in \{1, 2, \dots, n\}$ and $n$ the number of nodes in the network.
+The model is hierarchical: first, the \textit{rate function} dictates \textit{how often} changes in the network occur, then the  \textit{objective function} determines \textit{what} those changes are.% The variables of interest are the edges of the network, $x_{ij}$ between nodes $i$ and $j$, where $i \neq j \in \{1, 2, \dots, n\}$ and $n$ the number of nodes in the network.
 The $x_{ij}$ are directed and binary, i.e.
@@
 
-\noindent \textbf{Rate Function:} All changes in the network are treated as choices made by the actors in the network.
-The rate function dictates when changes are made and which actor can make them.
-When chosen, the actor, $i$, gets a chance to make a change to one of the other nodes $j$.
+\noindent \textbf{Rate Function:}  The rate function dictates when network changes are made and which actor can make them.
+An actor, $i$, is chosen make a change in its ties one of the other nodes $j$.
 In general, the rate function can include structural and node covariate parameters into account so that each actor has a different rate of change.
-However, we choose a simple rate function that is constant over all nodes in a given time period, because we focus on interpreting the parameters of the objective function.
-We denote the rate from $t_{m}$ to $t_{m+1}$ as $\alpha_m$ for $m = 1, \dots, M-1$.
+However, we choose a simple rate function that is constant over all nodes in a given time period, because we focus on interpreting the parameters of the objective function which directly impact the overall network structure.
+We denote the rate from $t_{m}$ to $t_{m+1}$ as $\alpha_m$ for $m = 1, \dots, M-1$, where $M$ is the number of time points at which the network was observed.
 Using this notation, the \textit{waiting time} to the next chance for actor $i$ to make a change is exponentially distributed with expected value $\alpha_m^{-1}$.
@@
 
-\noindent \textbf{Objective Function:} After actor $i$ has been selected by the rate function to change, it randomly picks one of its current ties, $x_{ij}$, to change.
-The probability that actor $i$ changes its current tie to actor $j$ is determined by the \textit{objective function} of the model and a random error term $U$ with log-Weibull distribution \citep{modelsSnijders}.
+\noindent \textbf{Objective Function:} After actor $i$ is selected to make a change, it randomly picks one of its current ties, $x_{ij}$, to change.
+%The probability that actor $i$ changes its current tie to actor $j$ is determined by the \textit{objective function} of the model and a random error term $U$ with log-Weibull distribution \citep{modelsSnijders}.
 %\hh{XXX can you paraphrase the purpose of U in half a sentence?}
@@
 where $\boldsymbol{\beta} = (\beta_1, \dots, \beta_K)$ are additional model parameters, each associated with some statistics, $s_{i1}(x, \mathbf{Z}), \dots, s_{iK}(x, \mathbf{Z})$, calculated for actor $i$ at the current network state $x$.
-T%he statistics can be very simple, like outdegree of the node, $s_{i1}(x) = \sum_{i\neq j} x_{ij}$, or more complicated, like the {\it transitive triplets jumping to different covariate} statistic, $s_i(x, \mathbf{Z}) = \sum_{i \neq j \neq h} x_{ij}x_{ih}x_{hj} \cdot \mathbb{I}(z_i = z_h \neq z_j)$.
+%The statistics can be very simple, like outdegree of the node, $s_{i1}(x) = \sum_{i\neq j} x_{ij}$, or more complicated, like the {\it transitive triplets jumping to different covariate} statistic, $s_i(x, \mathbf{Z}) = \sum_{i \neq j \neq h} x_{ij}x_{ih}x_{hj} \cdot \mathbb{I}(z_i = z_h \neq z_j)$.
 At least two parameters must be included in the objective function: density and reciprocity \citep{RSienamanual}.
-We denote the density, or out-degree, parameter by $\beta_1$ and the associated statistic as $s_{i1}(x)$ and we denote the reciprocity parameter by $\beta_2$ and the associated statistic as $s_{i2}(x) = \sum_{j} x_{ij}x_{ji}$.
+We denote the density, or out-degree, parameter by $\beta_1$ and the associated statistic as $s_{i1}(x)=\sum_{j} x_{ij}$ and we denote the reciprocity parameter by $\beta_2$ and the associated statistic as $s_{i2}(x) = \sum_{j} x_{ij}x_{ji}$.
 We will refer to the very simple model with only these two parameters in the objective function as model M1.
@@
 
-The objective function $f_{i}(x, \boldsymbol{\beta}, \mathbf{Z})$ and  random component $U$ combine to form the \textit{transition probability}, $p_{ij}$ of the network changing from its current state $x$ to the state $x(i\leadsto j)$, which is identical to $x$ except for $x_{ij}$: $x_{ij}(i \leadsto j) = 1 - x_{ij}$.
+The objective function $f_{i}(x, \boldsymbol{\beta}, \mathbf{Z})$ dictates the \textit{transition probability}, $p_{ij}$ of the network changing from its current state $x$ to the state $x(i\leadsto j)$, which is identical to $x$ except for $x_{ij}$: $x_{ij}(i \leadsto j) = 1 - x_{ij}$.
 The transition probability is
@@
 dictating which edge node $i$ changes.
-%Node $i$ can also choose to \textit{not} change at all, which is most likely when the value of Equation~\ref{eq:objective} is larger for $x$ than for any $x(i \leadsto j)$.
+Thus, the actor is more likely to make changes that increase the value of their objective function, and no change is most likely when any change decreases the value of the objective function.%Node $i$ can also choose to \textit{not} change at all, which is most likely when the value of Equation~\ref{eq:objective} is larger for $x$ than for any $x(i \leadsto j)$.
 
@@
 
-Data visualizations are an important component of data analysis, providing a mechanism for discovering patterns in data.
+Data visualization is an important component of data analysis, providing a mechanism for discovering patterns in data.
 Pioneering research by \citet{gelman:2004}, \citet{Bujaetal} and \citet{majumder:2011} provide methods to quantify the significance of discoveries made from visualizations.
@@
 The Rorschach protocol helps to understand the extent of randomness in the null model.
-Here, we use the lineup protocol to design significance, goodness-of-fit (GoF) and power tests.
-Under the protocol, we begin with a data set of interest to us, such as a network, a visualization of this data, such as a node-link diagram, and a model of interest.
-We consider two hypotheses: the null hypothesis that the model of interest generated the data, and the alternative hypothesis that the data were not generated under this model.
+Here, we use the lineup protocol to design significance, goodness-of-fit (GoF) and power tests for the CTMC model.
+Under the lineup protocol, we begin with a data set of interest to us, such as a network, a visualization of this data, such as a node-link diagram, and a model of interest.
+We consider two hypotheses: the null hypothesis that the model of interest generated the data, and the alternative hypothesis that the data were not generated under the model of interest.
 To construct a lineup of size $P$, $P-1$ sets of data are simulated from the null model.
@@
 Human observers then examine the lineup and identify the plot(s) that look(s) most different from the others.
-If an observer identifies the data plot, this is evidence against the null hypothesis.
-An observer has a chance of $P^{-1}$ to pick the data plot from the lineup by simply guessing.
-The evidence grows in strength with the number of independent observers identifying the data plot.
+If an observer identifies the data plot, this is evidence against the null hypothesis, \st{though an observer has a chance of 1 in $P$ to pick the data plot at random.}  The evidence grows in strength with the number of independent observers identifying the data plot.
 
@@
 This measure ranges in value from 0 to 1, where $WPC_{ij} = 1$ if senator $i$ is a co-sponsor on every one of senator $j$'s bills and $WPC_{ij} = 0$ if senator $i$ is never a co-sponsor any of senator $j$'s bills.
-Because CTMC models require binary edges, we construct the edges as follows:
+To simplify the problem, we construct a network with binary edges as follows:
  \begin{equation}\label{eq:edgewpc}
@@
 \end{equation}
-so that only strong relationships between senators are examined.
+so that only strong ties between senators are in the network.
 For each of the four senate sessions, we have three node covariates: the party affiliation of each senator, the number of bills they authored in each session, and their gender.
@@
 
-In legislative co-sponsorship networks, it is known that party affiliation, reciprocity of relationships, and whether senators are ``workhorses" who author many bills or ``show horses" who author few bills, are major influences on structure \citep{legnet}.
-We focus on these covariates, plus the additional gender covariate when choosing which CTMC models to fit to the data.
-
 \begin{table}
@@
 }
-\caption{\label{tab:effects} The effects we used in the CTMC models fit to the senate data.
+\caption{\label{tab:effects} The effects we used in the CTMC models for the senate data.
 In the picture which represents each effect, the dotted tie is encouraged to form if the estimate is positive, and discouraged to form if the estimate is negative.
@@
 We also selected some effects which were significant according to the built-in Wald-type tests in \texttt{RSiena} for application of our significance and goodness-of-fit methods.
-We examine a total of six models, each identified by its objective function.
+We fit six models to the senate data, and each model is identified by the parameters its objective function as shown in Table~\ref{tab:models}.
 The parameters in the objective function, $\beta_1, \dots, \beta_6$ are defined in Table~\ref{tab:effects}.
@@
 The largest model is M7: each of the other five models is nested in it.
-We used Wald tests in \texttt{RSiena} as described in Section~\ref{sec:res} to perform significance tests on $\beta_3, \dots, \beta_6$ in models M3, ..., M6.
+We used Wald-type as described in Section~\ref{sec:res} to perform significance tests on $\beta_3, \dots, \beta_6$ in models M3,$\dots$, M6.
 The most significant effect is $\beta_3$, the jumping transitive triplet (JTT) parameter for the party covariate, which was estimated to be about -5.9 with a standard error of 0.11, resulting in a Wald $p$-value of less than 0.0001.
@@
 The high positive estimate suggests senators are encouraged to collaborate with other senators who author about the same number of bills they do.
-This tendency of senators to co-sponsor bills written by senators who are similarly ``prolific" corresponds to the tendency of senators to be either ``workhorses" or ``show horses".
+That senators tend co-sponsor bills written by senators who are similarly prolific corresponds to the tendency of senators to be either ``workhorses" or ``show horses" \citep{legnet}.
 Senators known as workhorses author many pieces of legislation in a session, and largely stay out of the public arena.
@@
 @
-
-We fit all six of our models in \texttt{RSiena} using Markov Chain Monte Carlo (MCMC) methods to approximate the method of moments estimates of the parameters.
-Because the estimation is done through MCMC simulation, we fit each model to the data 1,000 times.
-From the simulations in which the MC converged, which made up over 90\% of the fits for each model, we computed the mean of the estimates of each parameter to get final estimates of $\hat{\boldsymbol{\beta}}$ for each model, which are shown in Table~\ref{tab:fittedvalues}.
 
@@
 \end{tabular}
-\caption{\label{tab:fittedvalues} The final estimates from repeated estimation of our models of interest.
-When simulating from these models, these are the estimates that we will use unless otherwise stated.}
+\caption{\label{tab:fittedvalues} The mean estimates from repeated fits of our models of interest. When simulating from these models, these are the estimates that we will use unless otherwise stated.}
 \end{table}
 
+We fit all six of our models in \texttt{RSiena} using Markov Chain Monte Carlo (MCMC) methods to approximate the method of moments estimates of the parameters.
+Because the estimation is done through MCMC simulation, we fit each model to the data 1,000 times.
+From the simulations that converged, which made up over 90\% of the fits for each model, we computed the mean of the estimates of each parameter to get final estimates of $\hat{\boldsymbol{\beta}}$ for each model, which are shown in Table~\ref{tab:fittedvalues}.
+
 We want to determine the role that each of these parameters plays in the objective functions for the different models.
-So, we use the estimates given in Table~\ref{tab:fittedvalues} to simulate from models M1 through M6.
-We discuss the simulation procedure and how we use the simulations in Section~\ref{sec:expersetup}.
+We use the estimates given in Table~\ref{tab:fittedvalues} to simulate from models M1 through M6.
+Then, we examine the effects on the structure of the simulated networks with a large visual inference experiment to determine significant effects, goodness-of-fit, and visual power.
 
@@
 
-We want to assess CTMC models for social network data using the lineup protocol in three ways: (1) significance tests of parameters, (2) goodness-of-fit tests of a model, and (3) determining visual power of the effects.
-%visual detection of parameters.
+We want to assess CTMC models for the senate network data using the lineup protocol in three ways: with significance tests of parameters,  goodness-of-fit tests of a model, and determination of the visual power of the effects.
 Each one of these situations requires a different setup, which we describe in detail, making use of the lineup protocol defined in \citet{Bujaetal}.
@@
 We chose to present our participants with only six plots at a time in order to show the node-link diagrams in more detail and to reduce cognitive load.
-To construct a lineup, we simulate five networks from the null model and one network from the alternative model.
+%To construct a lineup, we simulate five networks from the null model and one network from the alternative model.  
 Several lineups shown to our participants are presented and discussed in Section~\ref{sec:res}.
@@
 For the purposes of our experiment, we focus on simulating the second ``wave" of data, the $112^{th}$ Senate network, and we condition on the first wave of data, $111^{th}$ Senate network.
-Sections~\ref{sec:setupst} through~\ref{sec:setupvp} describe in detail how we constructed the lineups, which parameter values we simulate from, and why.
+Sections~\ref{sec:setupst} through~\ref{sec:setupvp} describe in detail how we constructed the lineups, which parameter values used, and why.
 Lineups we created were shown to independent observers recruited through Amazon Mechanical Turk for feedback, and we provide more detail on the Turk setup in Section~\ref{sec:res}.
@@
 To get the simulations,
-
-\begin{figure}
-\includegraphics[width=.9\textwidth]{img/shinyappscreen.png}
-\caption{\label{fig:shiny} A screen shot of the web application we created to help design our lineup experiment.
-More details about this application are given in Section~\ref{sec:setupvp}.
-In the lineup, M5 is the alternative model with $\beta_5$ set to twice its estimated value given in Table~\ref{tab:fittedvalues}.
-One plot simulated from this model is placed at random among five observations simulated from the null model, M1.
-%Participants of the study are asked to identify the most different plot.
-%The other two tabs in the right side of the screen tell you which plot comes from the alternative model, and allow you to view and download the data that is plotted in the lineup.
-}
-\end{figure}
 
@@
 %\end{description}
-Under the null hypothesis, we assume that the model that generated the network data is M1, the simplest model presented in Section~\ref{sec:models}.
-In the lineup, there are five null plots, constructed from  five simulations from M1 of the second wave, with $\beta_1, \beta_2$ set to the corresponding estimates given in Table~\ref{tab:fittedvalues}.
-The alternative model is the model with $\beta_1, \beta_2$, and $\beta_k$ in the objective function.
-The alternative data plot is simulated from the appropriate model with values set to the estimates from Table~\ref{tab:fittedvalues}.
-If an observer picks out the alternative data plot, that is evidence against the null hypothesis, while picking one of the null plots is evidence in favor of the null hypothesis.
+Under the null hypothesis, we assume that the model that generated the network data is M1, the simplest model presented in Section~\ref{sec:models}, and the alternative model is the one with $\beta_k$ included.
+In the lineup, there are five null plots constructed from  five simulations from M1 of the second wave, with $\beta_1, \beta_2$ set to the corresponding estimates given in Table~\ref{tab:fittedvalues}.
+The alternative data plot is simulated from the appropriate model with $\beta_1, \beta_2$, and $\beta_k$ in the objective function at corresponding values given in Table~\ref{tab:fittedvalues}.
+We constructed the lineups in this way because we hypothesize that if an effect is significant, it will alter the appearance of the simulated networks to the extent that they can be visually distinguished from the networks simulated from the simple model M1.
+Thus, an observer picking the alternative data plot as most different is evidence against the null hypothesis, while picking one of the null plots is evidence in favor of the null hypothesis.
 To avoid over-working our participants, we chose to test only two parameters, $\beta_3$ and $\beta_4$.
 These two had the smallest $p$-values from the significance tests, so we chose them for the visual significance test as well because we hypothesized they would be easier to pick out of the lineup.
-Thus, we compare the null model M1 to the alternative models M3 and M4, using three repetitions of each hypothesis test in the experiment.
+We compare the null model M1 to the alternative models M3 and M4, using three repetitions of each hypothesis test in the experiment to determine if $\beta_3, \beta_4$ are significant.
 
@@
 
-For the goodness-of-fit tests, we compare one model of interest to the data. The hypothesis we use to generate lineups are 
+For the goodness-of-fit tests, we compare one model of interest to the second wave of sentate network data.
+This procedure is identical to that in \citet{Bujaetal} and \citet{majumder:2011}.
+The hypotheses for the goodness-of-fit tests are:
 \begin{description}
+\centering
 \item $H_0$: The data come from the model of interest
@@
 
-To generate the null plots, we simulate five networks the model of interest using the corresponding parameters in Table~\ref{tab:fittedvalues}.
-We pick a wave to focus on, wave two, which is the first simulated network, and among these five plots, we place a node-link diagram of the true second wave of data.
-We cannot show the data more than once to each participant.
-The models we chose for goodness-of-fit testing are M3, M4, M5, and M7, and each participant only saw one lineup for the goodness-of-fit tests.
+\noindent To generate the null plots, we simulate five second wave networks from the model of interest using the corresponding parameters in Table~\ref{tab:fittedvalues}.
+Among these five plots, we place the true second wave of data.
+We cannot show the data more than once to each participant because doing so would bias the results, so each participant only sees one goodness-of-fit lineup.
+The models we chose for goodness-of-fit testing are M3, M4, M5, and M7.
+If a participant selects the data as most different, that is evidence that the model of interest is not a good fit.
 
@@
 Using VI, we want to determine at what value an additional effect included in the model becomes noticeable.
-By \textit{noticeable}, we mean that the inclusion of the effect alters the appearance of networks simulated from the model so much that many independent viewers are able to pick out the one node-link diagram rendered from data simulated from the model \textit{with} the effect in a lineup among five plots simulated from the model \textit{without} the effect.
+By \textit{noticeable}, we mean that the inclusion of the effect alters the appearance of networks simulated from the model so much that many independent viewers are able to pick out plot containing the data simulated from the model \textit{with} the effect in a lineup among five plots simulated from the model \textit{without} the effect.
 In this way, we measure the visual power of a parameter.
@@
 
-In model M1, with only two parameters in the objective function, we varied both the density and reciprocity parameter values one at a time, keeping all other parameters at their fitted values given in Table~\ref{tab:fittedvalues}.
-In models M3 through M6, we vary the additional parameter, $\beta_3$ through $\beta_6$, respectively, while holding all other parameters at their value in Table~\ref{tab:fittedvalues}.
-We want to determine how the size of these parameters affects the overall structure of the network data simulated from the models M1 through M6, so we also vary the magnitude of the parameters in order to determine at what value the effects become noticeable.
+In model M1, with only two parameters in the objective function, we changed the density and reciprocity parameter values one at a time, keeping all other parameters constant.
+We look at some values greater and some values lower than the estimates of $\beta_1, \beta_2$ given in Table~\ref{tab:fittedvalues}.
+In models M3 through M6, we vary only the additional parameter, $\beta_3$ through $\beta_6$, respectively, while holding all other parameters constant.
+We want to determine how the size of these parameters affects the overall structure of the network data simulated from the models M1 through M6.
+%By varying  so we also vary the magnitude of the parameters in order to determine at what value the effects become noticeable.
 %in both negative and positive directions.
@@
 <<hypothesis, fig.height=2, fig.cap="As the parameter increases in absolute value, more viewers of the lineup should pick the alternative data out of a lineup.
-Note that the significance test we construct in Section~\\ref{sec:setupst} is just one point on the curve, represented by the vertical dotted line labeled $\\hat{\\beta}$.
-The easy, medium, and hard lines represent how we determined which values of the parameters to show to our participants, and the horizontal dotted line shows the type-I error for one viewer of a lineup of size 6.">>=
+Note that the significance test we construct in Section~\\ref{sec:setupst} is just one point on the curve.
+The easy, medium, and hard lines represent how we determined which values of the parameters to show to our participants, and the horizontal dotted line is the chance of picking the alternative plot at random.">>=
+# , represented by the vertical dotted line labeled $\\hat{\\beta}$
 x <- seq(-10, 10, .05)
@@
   geom_label(data = NULL, aes(x = 0, y = .9, label = "Null Model = Alt. Model"), size = 2) +
-  geom_vline(xintercept = 150, color = 'red', linetype = 'dotdash') + 
-  annotate('text', x = 160, y = .9, label = "hat(beta)",parse = TRUE, size = 2) + 
+ # geom_vline(xintercept = 150, color = 'red', linetype = 'dashed') + 
+#  annotate('text', x = 160, y = .9, label = "hat(beta)",parse = TRUE, size = 2) + 
   geom_vline(xintercept = c(-300,-230, -175), color = c('green', 'yellow', 'red')) + 
+  geom_vline(xintercept = c(300,230, 175), color = c('green', 'yellow', 'red')) + 
   geom_text(data = NULL, aes(x = -300, y = .75, label = "easy"), angle = 90, vjust = -.5, size = 3) + 
@@
   geom_text(data = NULL, aes(x = -175, y = .06, label = "hard"), angle = 90, vjust = -.5, size = 3) + 
+  geom_text(data = NULL, aes(x = 300, y = .75, label = "easy"), angle = 90, vjust = -.5, size = 3) + 
+  geom_text(data = NULL, aes(x = 230, y = .5, label = "medium"), angle = 90, vjust = -.5, size = 3) +
+  geom_text(data = NULL, aes(x = 175, y = .06, label = "hard"), angle = 90, vjust = -.5, size = 3) + 
   scale_y_continuous(name = "% detecting plot from\nthe alternative model", labels = paste0(c(0,25,50,75,100), "%")) + 
@@
 
-To determine when an effect becomes noticeable, we examine six different levels of the effect, three negative and three positive ones.
-Figure~\ref{fig:hypothesis} shows a sketch of what the selection probability looks like hypothetically with varying effect size: the higher in absolute value the parameter is, the more likely participants are to select the alternative model out of the lineup.
-To determine the exact values of the six levels we want to test for each effect, we started with the estimates of the parameter at hand (see  Table~\ref{tab:fittedvalues }), and used small negative and positive factors to determine at what point \textit{we} noticed the effect of the parameter in simulations from the changed models.
-In Figure~\ref{fig:hypothesis}, we demonstrate these values with the vertical lines labeled ``easy," ``medium," and ``hard." We expect most viewers to see the effects at the ``easy" values, and we expect very few, if any, viewers to see the effects at the ``hard" values.
-
-To decide on the values to use for each difficulty level, we constructed an online application that simulated the lineup protocol for us to be the guinea pigs in our own experiment \citep{theqs}.
+\begin{figure}
+\includegraphics[width=.9\textwidth]{img/shinyappscreen.png}
+\caption{\label{fig:shiny} A screen shot of the web application we created to help design our lineup experiment.% More details about this application are given in Section~\ref{sec:setupvp}.
+In the lineup, M5 is the alternative model with $\beta_5$ set to twice its estimated value given in Table~\ref{tab:fittedvalues}.
+One plot simulated from this model is placed at random among five observations simulated from the null model, M1.
+%Participants of the study are asked to identify the most different plot.
+%The other two tabs in the right side of the screen tell you which plot comes from the alternative model, and allow you to view and download the data that is plotted in the lineup.
+}
+\end{figure}
+
+To determine when an effect becomes noticeable, we examine six different levels of the effect: three negative and three positive.
+Figure~\ref{fig:hypothesis} shows a sketch of the hypothetical selection probability with varying effect size.
+We hypothesize that as the parameter increases in absolute value, an observer is more likely to select the alternative data plot out of the lineup.
+The six levels of the effect are vertical lines labeled ``easy," ``medium," and ``hard" in Figure~\ref{fig:hypothesis}.
+We expect most observers to pick out the alternative data plot at the ``easy" values, and we expect very few, if any, observers to pick out the alternative data plot at the ``hard" values.
+To help us decide on the parameter values to use for each effect at each difficulty level, we constructed an online application that simulated the lineup protocol for us to be the guinea pigs in our own experiment \citep{theqs}.
 A screen shot of the app we created with the \texttt{shiny} package by \citet{shiny} is shown in Figure~\ref{fig:shiny}.
-On the left side of the screen, the user\footnote{Please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/} to create lineups constructed from the models we present for this data for yourself.} can input the information necessary for creating a lineup from the models M1 through M7 for the data in Section~\ref{sec:models}:
-\begin{enumerate}
-\item Choose to simulate one plot to change the alternative model in the lineup, or simulate $P-1$ plots to change the null model in the lineup
-\item Choose the model from which to simulate and choose the wave of data to simulate. 
-\item If M1 is selected in (2), select whether to alter the density ($\beta_1$) or the reciprocity ($\beta_2$) parameter. 
-\item Choose the size of the lineup, $P$, the magnitude of the effect size, a random seed for reproducibility, and a layout algorithm to use for the node-link diagrams.
-There is also a checkbox if the user wishes the nodes to be colored by the size of the connected component to which they belong.
-\end{enumerate}
-All plots that are \textit{not} from the model specified by steps 1-4 above are simulated from model M1 with the estimates of the rate parameters, and $\beta_1$ and $\beta_2$ as given in Table~\ref{tab:fittedvalues}.
+On the left side of the screen, the user\footnote{Please visit \url{https://sctyner.shinyapps.io/saom_lineup_creation/} to create lineups constructed from the models we present for this data for yourself.} can input the information necessary for creating a lineup using network data simulated from models M1 through M7.
+%\begin{enumerate}
+%\item Choose to simulate one plot to change the alternative model in the lineup, or simulate $P-1$ plots to change %the null model in the lineup
+%\item Choose the model from which to simulate and choose the wave of data to simulate. 
+%\item If M1 is selected in (2), select whether to alter the density ($\beta_1$) or the reciprocity ($\beta_2$) %parameter. 
+%\item Choose the size of the lineup, $P$, the magnitude of the effect size, a random seed for reproducibility, and a %layout algorithm to use for the node-link diagrams.
+There is also a checkbox if the user wishes the nodes to be %colored by the size of the connected component to which they belong.
+%\end{enumerate}
+The plots in the lineup that are not specified by the user contain data simulated from model M1.
+%estimates of the rate parameters, and $\beta_1$ and $\beta_2$ as given in Table~\ref{tab:fittedvalues}.
 
@@
 \hline
-$\beta_4$ & pos & 8.351 & 6.6806 & 5.010 \\ % 5.316 & 3.9872 & 3.323 \\ 
-   & reverse & 6.681 & 5.0105 & 3.340 \\  %5.316 & 3.9872 & 3.323 \\ 
+$\beta_4$ & pos &  6.681 & 5.0105 & 3.340  \\ % 5.316 & 3.9872 & 3.323 \\ 
+   & reverse & 8.351 & 6.6806 & 5.010 \\  %5.316 & 3.9872 & 3.323 \\ 
   $\beta_6$ & pos & 5.316 & 3.9872 & 3.323 \\  %-30.272 & -20.1817 & -10.091 \\ 
@@
 
-Using our Shiny application, we settled on six parameter values to test for each of our six effects, $\beta_1, \dots \beta_6$.
+Using the Shiny application, we settled on six parameter values to test for each of the six effects, $\beta_1, \dots \beta_6$.
 All values of the parameters used in the experiment are given in Table~\ref{tab:experdetail}.
-In the case of both $\beta_4$ and $\beta_6$, we could not determine any negative parameter values that made the data simulated from M4 and M6 look different than null model simulations from model M1.
+In the case of both $\beta_4$ and $\beta_6$, we could not determine any negative parameter values that made the networks simulated from M4 and M6 look different than the networks simulated from model M1.
 We hypothesize that this is due to negative effects \textit{removing} visually interesting structural elements as opposed to \textit{adding} noticeable structural elements.
@@
 
-<<getexperdetail, results='asis'>>=
-turk22 <- read_csv("data/turk22-sig.csv")
+<<getexperdetail, results='asis', eval = FALSE>>=
+
 modelData_sig <- turk22 %>% 
@@
 We recruited 250 participants for our experiment through Amazon Mechanical Turk.
-Each participant was presented with some brief training material first, then before presenting the lineups for the hypothesis tests, each person was shown two trial plots one where the alternative plot was the most different from the others due to its relatively \textit{complex} structure, while the other trial included an alternative plot that was most different from the others due to its comparatively \textit{simple} structure.
+Each participant was first presented with some brief training material.
+Before presenting the lineups for the hypothesis tests, each participant was shown two trial plots: one where the alternative plot was the most different from the others due to its relatively \textit{complex} structure, while the other trial included an alternative plot that was most different from the others due to its comparatively \textit{simple} structure.
 Only when participants were able to correctly identify the alternative plot from the trial lineups were they allowed to begin the experiment.
@@
 Twelve of the 13 lineups that the participants saw were used for the significance testing and the visual power methods discussed in Sections~\ref{sec:setupst} and \ref{sec:setupvp}.
+Among the 12 lineups were four from each difficulty level (easy, medium, and hard), and two of the four were from each condition (negative-positive, or positive-reverse).
+Each participant saw each of the six models twice at some combination of difficulty and condition.
 %The six parameters, $\beta_1, \dots, \beta_6$ were set to three different values according to how difficult we thought picking the data plot from the lineup would be for our participants, and were also set to be less than or greater than the initial estimate, creating the lineup type variable.%\hh{what about the first 12 plots?
@@
 In this section, we present the results from our experiment and compare them to traditional statistical tests where applicable.
-Because each lineup shown to participants has only six plots, the probability of picking the data by chance is high at 1 in 6, but if \textit{many} independent viewers pick out the data from the nulls, the evidence in favor of the alternative hypothesis becomes stronger.
+Because each lineup shown to participants has only six plots, the probability of picking the data by chance is high at 1 in 6, but if many independent viewers pick out the data from the nulls, the evidence against the null hypothesis becomes stronger.
 The $p$-values from the lineups were calculated using the \texttt{vinference} package by \citet{vinference}.
-This package contains methods to calculate \textit{Visual distributions} for lineup experiment data.
+This package contains methods to calculate \textit{visual distributions} for lineup experiment data.
 The distribution depends on the number of evaluations of a plot, $L$, the size of the lineup, $P$, and the lineup scenario, which here is that each lineup containing the same data and the same set of null plots is shown to $L$ independent observers.
 The visual inference family of distributions is similar to the binomial distribution, but takes the dependency among the $P$ plots in a single lineup shown to multiple viewers into account.
+Suppose that out of $L$ observers of a lineup of size $P$, there are $\ell$ observers who pick out the alternative data plot.
+Then the corresponding $p$-value from the visual distribution gives the probability that $\ell$ or more out of $L$ independent observers would pick out the alternative data plot by chance.
 
@@
 \citep{RSienamanual}.
-Testing the significance of multiple parameters depends on the hypothesis we wish to test, and an $H \times K$ matrix, $\mathbf{A}$, must be appropriately designed to test the $H$ hypotheses of interest.
+%Testing the significance of multiple parameters depends on the hypothesis we wish to test, and an $H \times K$ matrix, $\mathbf{A}$, must be appropriately designed to test the $H$ hypotheses of interest.
 The null hypothesis is that $\mathbf{A}\boldsymbol{\beta} = \mathbf{0}$, and the test statistic is
-\begin{equation}\label{eq:wald2}
-(\mathbf{A}\boldsymbol{\hat{\beta}})' \hat{\Sigma}^{-1} \mathbf{A}\boldsymbol{\hat{\beta}} \sim \chi^2_H,
-\end{equation}
-where $\hat{\Sigma}$ is the estimated covariance matrix of $\boldsymbol{\beta}$.
-
-Both parameters we test for significance using the lineup protocol, $\beta_3$ and $\beta_4$, were determined to be statistically significant using Equation~\ref{eq:wald1}.
-The corresponding results from the significance tests we performed using the lineup protocol are given in Table~\ref{tab:sigtesting}.
-If enough participants pick out the alternative plot to result in a $p$-value less than 0.05, we reject the null hypothesis that the true value of the additional parameter, either $\beta_3$ or $\beta_4$, is equal to zero.
+% \begin{equation}\label{eq:wald2}
+% (\mathbf{A}\boldsymbol{\hat{\beta}})' \hat{\Sigma}^{-1} \mathbf{A}\boldsymbol{\hat{\beta}} \sim \chi^2_H,
+% \end{equation}
+% where $\hat{\Sigma}$ is the estimated covariance matrix of $\boldsymbol{\beta}$. 
+Both parameters we test for significance using the lineup protocol, $\beta_3$ and $\beta_4$, were determined to be statistically significant using Equation~\ref{eq:wald1} with $p$-values given in Table~\ref{tab:sigtesting}.
+For the visual significance test, if the number of participants who pick out the alternative plot results in a $p$-value less than 0.05, we reject the null hypothesis that the true value of the additional parameter, either $\beta_3$ or $\beta_4$, is equal to zero.
+If the null hypothesis is not rejected, then there is evidence that the additional parameter does not affect the overall structure of the network even though it was found to be sigificant with the Wald-type test.
+The $p$-values from the visual distribution for the significance tests of $\beta_3, \beta_4$ are given in Table~\ref{tab:sigtesting}.
+The results for each lineup, whether to reject or fail to reject the null hypothesis, vary, with one of three lineups from each parameter  group resulting in rejection of the null hypothesis.
 %\hh{XXX how do we interpret these  visual results?
@@
             pv = map2_dbl(npick, n, pV, m=6, scenario=3))
-names(summ_sig) <- c("Lineup ID", "parameter", "\\# Identified", "Total Views", "p-value")
+names(summ_sig) <- c("Lineup ID", "parameter", "\\# ID", "Total Views", "p-value")
 summ_sig$parameter <- as.factor(summ_sig$parameter)
@@
 xtable::print.xtable(xtable::xtable(summ_sig, caption = "Experiment results for the two parameters for which we performed significance tests.
-\\# Identified indicates the number of participants who selected the alternative model plot.
+\\# ID indicates the number of participants who identified the alternative data plot.
 There were three lineups for each parameter, so there are three results for each plot.", label = "tab:sigtesting", digits = c(0,0,0,0,0,5)), include.rownames = FALSE, sanitize.text.function=function(x){x})
@@
  %Using these $p$-values, all but one lineup results in a rejection of the null hypothesis at Type-I error rate of $\alpha = 0.05$.
-We see in Table~\ref{tab:sigtesting} that the $p$-values for visual inference here are highly variable.
 
@@
 
-<<lineup3133, fig.height=4, fig.cap="One of the lineups which failed to reject the null hypothesis that $\\beta_3$ = 0.
-The network simulated from model M3 is found in panel $\\sqrt{25} - 4$, and the remaining panels show networks simulated from model M1.">>=
+<<lineup3133, fig.height=4, fig.cap="Lineup 3133, which resulted in failure to reject the null hypothesis that $\\beta_3$ = 0.
+The network simulated from model M3 is found in panel $\\sqrt{25} - 4$, and the remaining panels show networks simulated from model M1.
+16 of 27 viewers picked plot two as most different.">>=
 dat <- read_csv("data/jttp_pos_hard_3.csv")
@@
 In viewing these two lineups, it is evident that there is a large amount of variability in networks simulated from CTMC models.
+The variability in significance test results is introduced through the null plots generated from M1.
 It is difficult to see that five of the six networks come from the same model when they all look very different.
-The variability in significance test results is introduced through the null plots generated from M1.
 In addition, the small number of null plots do not give the viewer as complete of a view of the null model as the usual 19 null plots would.
+In both lineups in Figures~\ref{fig:lineup3132} and ~\ref{fig:lineup3133}, most participants chose the plot with the largest connected component as the most different.
 
@@
 Most network models, other than the most simple, lack the necessary asymptotics for developing goodness-of-fit methods \citep{goldenberg09}.
-Some simulation-based methods have been developed using what \citeauthor{RSienamanual} call \textit{auxiliary statistics} such as the indegree or outdegree distribution on the nodes.
+Some simulation-based methods have been developed using what auxiliary statistics such as the indegree or outdegree distribution on the nodes \citep{RSienamanual}.
 In \texttt{RSiena}, the \texttt{sienaGOF} function performs goodness-of-fit testing as follows:
 \begin{enumerate}
-\item Auxiliary statistics, such as the cumulative outdegree counts on the nodes, are computed on the observed data ($\mathbf{u}_d$) and on $N$ observations simulated from the model ($\mathbf{u}_1 \dots \mathbf{u}_N$).
+\item Auxiliary statistics, such as the cumulative outdegree distribution on the nodes, are computed on the observed data ($\mathbf{u}_d$) and on $N$ observations simulated from the model ($\mathbf{u}_1 \dots \mathbf{u}_N$).
 \item The mean $\overline{\mathbf{u}}$ and covariance matrix $\mathbf{S}$ are computed from the $N$ simulations, and the Mahalanobis distance, $d_M(\mathbf{u}_d)$ from the observed statistics to the distribution of the simulated statistics is computed:
@@
 <<gofsiena, fig.height=2.5, fig.cap="An example of what a goodness-of-fit plot from \\texttt{RSiena} looks like.
-The overlaid boxplots and violin plots show the distribution of each of the outdegree count values on the simulated networks, and the red points and lines are the observed data values.", cache = TRUE>>=
+The overlaid boxplots and violin plots show the distribution of each of the outdegree count values on the simulated networks, and the red points and lines are the observed data values.", cache = TRUE, eval=FALSE>>=
 load("data/ansnullpaper.rda")
@@
 @
-A plot comparing the data to the simulations is also considered, and a similar plot is shown in Figure~\ref{fig:gofsiena} for the outdegree distribution of small data set, shown in the points and connected lines, with the simulated values of $\mathbf{u}_d$ shown in box plots and overlaid violin plots.
-The \texttt{RSiena} software also provides a Rao score-type test for goodness-of-fit for assessing one or more parameters, the test statistic of which is compared to a Chi-square distribution with $H$ degrees of freedom, where $H$ has the same definition as in Section~\ref{sec:sigtest}.
+%A plot comparing the data to the simulations is also considered, and a similar plot is shown in Figure~\ref{fig:gofsiena} for the outdegree distribution of small data set, shown in the points and connected lines, with the simulated values of $\mathbf{u}_d$ shown in box plots and overlaid violin plots.
+%The \texttt{RSiena} software also provides a Rao score-type test for goodness-of-fit for assessing one or more parameters, the test statistic of which is compared to a Chi-square distribution with $H$ degrees of freedom, where $H$ has the same definition as in Section~\ref{sec:sigtest}.
 For full detail on the score-type test, see \citet{scoretest}.
-These methods are both restricted: the \texttt{sienaGOF} method only considers \textit{one} measure on the data and simulations from the model, while the score-type tests only consider \textit{subsets} of parameters, ``nuisance parameters" in \citet{scoretest}, not the entire set of parameters.
-By using visual inference instead of more traditional statistical methods, we hope to perform a more \textit{holistic} goodness-of-fit test.
+\noindent This goodness-of-fit test is very limited because it only considers one measure on the data and simulations at a time.
+To assess goodness-of-fit on the whole network, many auxiliary statistics need to be considered.
+Thus, by using visual inference instead of more traditional statistical methods, we construct a more holistic goodness-of-fit test.
 
@@
 In each lineup, the null model is one of the four models and the alternative model is the true, unknown model that generated the senate network data.
-The hypotheses for our goodness-of-fit tests are:
-\begin{description}
-\item $H_0$:  The senate network data come from the null model, M$i$.
-\item $H_A$: The senate network data do not come from the null model. 
-\end{description}
+%The hypotheses for our goodness-of-fit tests are:
+% \begin{description}
+% \item $H_0$:  The senate network data come from the null model, M$i$.
+% \item $H_A$: The senate network data do not come from the null model. 
+% \end{description}
 If a lineup viewer picks out the data among the five simulations from the null model, it is evidence against the null hypothesis.
@@
 The $p$-values were again computed using the \texttt{vinference} package by \citet{vinference}.
+In all lineups except replicate 2 of M4, the null hypothesis that the data were generated by the model is rejected.
+Thus, none of the models we have chosen are a good fit to the data.
 The lineup that resulted in a failure to reject the null hypothesis is shown in Figure~\ref{fig:failgof}.
-The null model in this lineup is M5, and the senate data is shown in panel number $3^2 - 7$.
+The null model in this lineup is M4, and the senate data is shown in panel number $3^2 - 7$.
 However, the panel most participants chose was number four, and the most common reasoning for that choice was that it had the most simple structure.
 Some of the other panels, such as three and six, in Figure~\ref{fig:failgof} have large connected components that are similar in size to the connected component of the data plot shown in panel two.
-Thus, model M5 is sometimes capable of capturing the network structure of the senate collaboration data.
-
-<<failgof, fig.height=4, fig.cap="The goodness-of-fit lineup that failed to reject the null hypothesis.
-The null model for this lineup is M5.
+Thus, model M4 is sometimes capable of capturing the network structure of the senate collaboration data.
+
+<<failgof, fig.height=4, fig.cap="The goodness-of-fit lineup which resulted in failure to reject the null hypothesis.
+The null model for this lineup is M4.
 Only 7 of 20 viewers of this lineup selected the data plot as the most different from the others.
@@
 
-The smallest $p$-value for one of the goodness-of-fit lineups was for the third replicate of the null model M5.
-This result contrasts with our previous finding that the only lineup to fail to reject the null was also when the null model was M5.
+The smallest $p$-value for one of the goodness-of-fit lineups was for the third replicate of the null model M4.
+This result contrasts with our previous finding that the only lineup to fail to reject the null was also when the null model was M4.
 This lineup is shown in Figure~\ref{fig:mostsiggof}.
-In the remaining replicate of M5 as the null model, 13 of 16 viewers identified the data plot, corresponding to a $p$-values of less than $0.0001$, just like the third replicate.
-This variability in results is similar to the variability we found in Section~\ref{sec:sigtest}.
-This variability is again introduced through the plots simulated from null model, and does not provide us with a clear cut decision resulting the hypothesis test.
-For model M5, we can neither reject nor fail to reject the null hypothesis that the data come from model M5.
+In the remaining replicate of M4 as the null model, 13 of 16 viewers identified the data plot, corresponding to a $p$-values of less than $0.0001$, just like the third replicate.
+This variability in results is similar to the variability we found in Section~\ref{sec:sigtest}, and does not provide us with a clear cut result from the goodness-of-fit tests.
+For model M4, we can neither reject nor fail to reject the null hypothesis that the data come from model M4.
 This is evidence that the goodness-of-fit of network models cannot always be determined by one dimensional derived features.%, such as $p$-value shown on the $x$-axis in Figure~\ref{fig:gofsiena}.
@@
 
-<<mostsiggof, fig.height = 4, fig.cap="The lineup resulting in the smallest $p$-value rejecting the null hypothesis. Surprisingly, this another repetition for M5 as the null model.">>=
+<<mostsiggof, fig.height = 4, fig.cap="The lineup resulting in the smallest $p$-value rejecting the null hypothesis.
+Surprisingly, this another repetition for M5 as the null model.
+The data are in panel 6.">>=
 dat <- read_csv(paste0("data/jtts_gof_9_3.csv"))
@@
 
-<<gofm71, fig.height=4, fig.cap="One repitition of a goodness-of-fit lineup testing modle M7.
+<<gofm71, fig.height=4, fig.cap="One repitition of a goodness-of-fit lineup testing model M7.
 The senate data are shown in panel two, and it is evident that none of the other five panels, which show data simulated from model M7, come close to creating the large connected component that is central to the structure of the senate data.">>=
@@
 
-We believe this goodness-of-fit testing method holds promise for the future of social network analysis.
+We believe this goodness-of-fit testing method holds promise for the future of network analysis.
 The participants in our experiments are very good overall at picking out the data when it is noticeably different from the null plots in the lineups.
 In addition, as in replicate three for null model M4, when the null plots contain similarly sized structures as the data plot, our participants have a hard time distinguishing the data.
-We believe that running these tests multiple times using several different sets of null models to adequately explore the possible structures generated by the models is a step in the right direction for a more comprehensive goodness-of-fit test for network models.
+We believe that running these tests multiple times using several different sets of null models to adequately explore the possible structures generated by the models \st{results in} a more comprehensive goodness-of-fit test for network models.
 
@@
 
-The results from the visual power piece of our experiment are shown in Figure~\ref{fig:predictglmm}.
-On the $x$ axis, we plot the value of the parameter of interest, and on the $y$ axis, the proportion of times the alternative data plot was picked out for each lineup.
+The results from the visual power part of our experiment are shown in Figure~\ref{fig:predictglmm}.
+On the $x$ axis, we plot the value of the parameter of interest, and on the $y$ axis, the proportion of times the alternative data plot was picked out from each lineup.
 The results are split into groups based on the value of the parameter and the lineup type.
@@
 
-We model the relationship between identification of the alternative data in the lineup and the parameter of interest, the effect size, and the lineup type with a generalized linear mixed model (GLMM) that provides us with an estimate of the power of the visual significance test.
-The response variable, $X_{klqr}$, is binary, indicating whether participant $r$ picked the alternative data plot in lineup type $q$, rep $\ell$, for effect $k$.
+We model identification of the alternative data in the lineup by the parameter of interest, the effect size, and the lineup type with a generalized linear mixed model (GLMM) that provides us with an estimate of the power of the visual significance test.
+The visual power is the probability of detecting the parameter in a lineup of size six.
+The response variable in our model, $X_{k\ell qr}$, is binary, indicating whether participant $r$ picked the alternative data plot in lineup type $\ell$, rep $q$, for effect $k$, and it follows a Bernoulli distribution with probability $\pi_{klqr}$.
+We model $\text{logit}(\pi_{klqr})$ as a sum of the effects for parameter and lineup type, plus random effects for each lineup, $\delta_{klq}$, and for each participant, $\epsilon_r$.
 There is one continuous covariate $v$, which is the centered and scaled size of the effect of interest from which the alternative data were simulated, the values of which are labeled ``easy", ``medium", and ``hard" in Table~\ref{tab:experdetail} according to how difficult we thought the Turk participants would find each lineup.
-In Equation~\ref{eq:glmm}, $k \in \{1, 2, 3, 4, 6\}$ corresponds to the effects $\beta_1,\dots, \beta_6$, respectively, $\ell \in \{-1,1\}$, and $q \in \{1,2,3\}$.
-We also include random effects in the model: one for each lineup, $\delta_{klq}$, and one for each participant, $\epsilon_r$, and fit a hierarchical model given in Equation~\ref{eq:glmm}.
+The complete hierarchical model is given in Equation~\ref{eq:glmm}, where $k \in \{1, 2, 3, 4, 5, 6\}$ corresponds to the effects $\beta_1,\dots, \beta_6$, respectively, $\ell \in \{-1,1\}$, and $q \in \{1,2,3\}$.
 
@@
 
-The results of fitting this model, including estimates of parameters, standard errors, $p$-values, and the odds ratio multipliers, using \texttt{glmer} from the \texttt{lme4} package are summarized in Table~\ref{tab:glmmests} \citep{lme4}.
+We fit the hierarchical model with \texttt{glmer} from the \texttt{lme4} package \citep{lme4}.
+The parameter estimates, standard errors, $p$-values, and the odds ratio multipliers from the model  are summarized in Table~\ref{tab:glmmests}.
 For each combination of parameter and lineup type, the expected value of the link function for a new lineup and a new observer with parameter value $v$ is
 \begin{equation}
-E[\text{logit}(\pi_{kl})] = \alpha_{kl} + \gamma_{kl}v
+E[\text{logit}(\pi_{kl})] = \eta_{kl} + \gamma_{kl}v
 \end{equation}
@@
   geom_vline(data = betahat, aes(xintercept = size), linetype = "dotted", color = "grey60") + 
-  geom_text(data = betahat, aes(x = size, y = .9, label = paste0("hat(beta)[", param, "]")), parse = T, hjust = 0, color = "grey60") + 
+  geom_text(data = betahat, aes(x = size, y = Inf, label = paste0("hat(beta)[", param, "]")), parse = T, color = "grey60", hjust = -.5, vjust = 1.1) + 
   geom_line(data = newdata, aes(x = size, y = predictfinal, color = as.factor(type2))) + 
@@
 
-In Figure~\ref{fig:predictglmm}, we see a clear trend in all parameters except $\beta_1$ and $\beta_2$ that as the parameter value approaches zero from either side, the probability of picking the data plot in a lineup of size six decreases.
+In Figure~\ref{fig:predictglmm}, we see a clear trend in $\beta_2$ through $\beta_6$ that as the parameter value approaches zero from either side, the probability of picking the data plot in a lineup of size six decreases.
 This supports our hypothesis shown in Figure~\ref{fig:hypothesis}.
@@
 At a value of $\beta_5 = -20$, however, the model predicts this same probability to be about 0.40.
-This again demonstrates that the \textit{presence} of structure is detected  more frequently and at smaller values than the absence of structure.
+This again demonstrates that the presence of structure is detected  more frequently and at smaller values than the absence of structure.
 
@@
 
-<<beta4neg1,fig.height=4, fig.cap='In our experiment, 41.4\\% of viewers of this plot selected the plot from the alternative model, M1.
-The ``reverse" of this lineup is given in Figure~\\ref{fig:beta41}, where 52.8\\% of viewers selected the plot from the alternative model, M1.
-Here, the alternative plot is $\\sqrt{25} - 1$.' >>=
+<<beta4neg1,fig.height=4, fig.cap='In our experiment, 41.4\\% people viewing this lineup selected the plot from the alternative model, M1.
+Here, the alternative plot is $\\sqrt{25} - 1$.
+The other five plots were simulated from model M4.' >>=
 dat <- read_csv("data/jtts_neg_med_2.csv")
@@
 For $\beta_4$ and $\beta_6$, where one plot simulated from M1 was placed among five plots from the corresponding model, we see that the predictions for the reverse lineup type (-1), are less than the standard lineup type (1) for all values of the parameter that we have.
-This contradicts our hypothesis for this scenario, which was that these two scenarios would perform similarly.
-One of the lineups for the $\beta_4 = 6.681$, lineup type 1 scenario is given in Figure~\ref{fig:beta41}, and a corresponding lineup for the lineup type -1 scenario is given in Figure~\ref{fig:beta4neg1}.
+This contradicts our hypothesis for this scenario, which was that these two lineup types would perform similarly.
+A lineup type 1 scenario is given in Figure~\ref{fig:beta41}, and a corresponding lineup for the lineup type -1 scenario is given in Figure~\ref{fig:beta4neg1}, where the parameter of interest is $\beta_4 = 6.681$.
 For identical values of the parameter, viewers had a harder time identifying the different plot when they were selecting the most ``simple" structure, detecting M1 in five plots from the more complicated model, than they did identifying the most ``complex" structure, the plot from the more complicated model, from the five plots from M1.
-%This result is also similar to that of \citet{corrviz} because it emphasizes the difficulty of picking out the absence of an effect relative to picking out the presence of an effect.
+This result is also similar to that of \citet{corrviz} because it emphasizes the difficulty of picking out the absence of an effect relative to picking out the presence of an effect.
+Overall, the visual power of an effect increases as the effect size increases in absolute value, but the power is not symmetric around zero because stong positive effects are noticeable sooner than strong negative effects.
 
@@
 % what did we do? what did we learn from it? 
-By using visual inference methods, we have developed new ways to perform significance and goodness-of-fit tests for a complicated, intractable set of statistical models for social network data.
+By using visual inference methods, we have developed new ways to perform significance and goodness-of-fit tests for a complicated family of statistical models for social network data.
 We have also developed a way to determine the power of these new visual tests.
@@
 features, we develop an idea of the model in terms of the \textit{data} itself, not in terms of statistical summaries of the data.
-These methods place the model in the data space: they don't summarize or compress the data to put it in the model space \citep{Wickham2015}.
-
-Furthermore, we have found the visual power of some effects in the object function of a CTMC model for this particular senate data example,  and we have shown that, for the same effects, there is a lot of variability in results from significance and goodness of fit tests.
+%These methods place the model in the data space: they don't summarize or compress the data to put it in the model space \citep{Wickham2015}.
+
+Furthermore, we have found the visual power of some effects in the objective function of a CTMC model for this particular senate data example,  and we have shown that, for the same effects, there is a lot of variability in results from significance and goodness of fit tests.
 Because the visual tests we performed show a great deal of variability, we can see that the decisions with respect to the significance of a parameter or the goodness of fit of a model to data are not as cut-and-dried as the more traditional methods would have us believe.
@@
 
-We these methods will be applied to different types of network data and different types of network models.
+We hope these methods will be applied to different types of network data and different types of network models.
 But given the limitations of the node-link visualization, the cognitive load of looking at a lineup is very high for the average observer.

### 2019-07-26 ASApaper/paper.Rnw
latest edits

--- 2019-03-05 ASApaper/paper.Rnw
@@
 With VI, we can look at entire datasets simulated from a model, instead of relying a single metric such as a $p$-value.
-We conducted a VI experiment, with participants recruited via Amazon Mechanical Turk, to assess the significance of CTMC parameters, the fit of a CTMC model, and the visual power of CTMC parameters.
+We conducted a VI experiment, with participants recruited via Amazon Mechanical Turk, to assess the significance of CTMC parameters, the fit of a CTMC model, and the visual power of CTMC parameters, and found visual inference can be used to supplement traditional statistical tests for network models.
 \end{abstract}

### 2019-08-06 ASApaper/paper-child.Rnw
moving stuff to supp mat

@@
 The more complex the model, however, the harder it typically is to assess significance, power, and goodness-of-fit.
-\st{Citation?} One particularly complicated family of models are those designed to model networks.
+One particularly complicated family of models are those designed to model networks.
 A \textit{network} is any set of things, such as people, or computers, that are connected in some way, through social relations or the Internet.
@@
 for one family social network models: continuous time Markov-chain (CTMC) models for dynamic network data, as described in \citet{saompaper}.
-We use \textit{visual inference} \st{to complement}%in place of
+We use \textit{visual inference} \st{to complement} %in place of
 traditional statistical methods for social network models, such as $t$-tests for significance of parameters or outdegree distributions for goodness-of-fit tests.
@@
 The data we use are collaboration networks in the United States Senate during the $111^{th}$ through $114^{th}$ Congresses.
-These senates began on January 6, 2009, the start date of the $111^{th}$ Congress, and ended on January 3, 2017, the last date of the $114^{th}$ Congress Details of how this data can be downloaded are provided by François Briatte at \url{github.com/briatte/congress}.
-In the US Senate, senators often show support for a piece of legislation by co-sponsoring a bill authored by one of their colleagues.
-In a co-sponsorship network, ties are directed from senator $i$ to senator $j$ when senator $i$ signs on as a co-sponsor to the bill that senator $j$ authored.
-There are many hundreds of ties between senators when they are connected in this way, so we simplify the network by computing a single value for each senator-senator collaboration called the \textit{weighted propensity to co-sponsor} (WPC).
-This value is defined in \citet{senate} as
-
-\begin{equation}\label{eq:sen1}
-    WPC_{ij} = \sum\limits_{b=1}^{B_j} \frac{Y_{ij(b)}}{c_{j(b)}}\left(\sum\limits_{b=1}^{B_j} \frac{1}{c_{j(b)}}\right)^{-1}
-\end{equation}
-where $B_j$ is the number of bills in a congressional session authored by senator $j$, $c_{j(b)}$ is the number of co-sponsors on senator $j$'s $b^{th}$ bill, where $b \in \{1,\dots, B_j\}$, and $Y_{ij(b)}$ is an indicator variable that senator $i$ co-sponsored senator $j$'s $b^{th}$ bill.
-This measure ranges in value from 0 to 1, where $WPC_{ij} = 1$ if senator $i$ is a co-sponsor on every one of senator $j$'s bills and $WPC_{ij} = 0$ if senator $i$ is never a co-sponsor any of senator $j$'s bills.
-To simplify the problem, we construct a network with binary edges as follows:
- \begin{equation}\label{eq:edgewpc}
-  x_{ij} =
-\begin{cases}
-                                   1 & WPC_{ij} > 0.25 \\
-                                   0 & WPC_{ij} \leq 0.25
-\end{cases}
-\end{equation}
-so that only strong ties between senators are in the network.
+These senates began on January 6, 2009, the start date of the $111^{th}$ Congress, and ended on January 3, 2017, the last date of the $114^{th}$ Congress.
 For each of the four senate sessions, we have three node covariates: the party affiliation of each senator, the number of bills they authored in each session, and their gender.
@@
 ggplot(data = seobama2) + 
-  geom_net(directed = T, labelon=T, arrowsize = .25, singletons= T, fiteach = T, linewidth = .25, layout.alg = 'fruchtermanreingold', fontsize = 2, repel=T, labelcolour = "grey40", 
+  geom_net(directed = T, labelon=T, arrowsize = .25, singletons= T, fiteach = T, linewidth = .25, layout.alg = 'fruchtermanreingold', fontsize = 2.5, repel=F, labelcolour = "black", vjust = -.8,
            aes(from_id = source, to_id = target, color = party, 
@@
 
+%\begin{wraptable}{r}{6.5cm}
 \begin{table}
-\centering
 \begin{tabular}{lcccccc}
@@
 M6 & \checkmark & 	\checkmark & -- & -- & -- & \checkmark \\
-M7 & \checkmark & 	\checkmark & \checkmark & \checkmark & \checkmark & \checkmark \\
+M7 & \checkmark & 	\checkmark & -- & \checkmark & \checkmark & \checkmark \\
 \bottomrule
@@
 \end{table}
+%\end{wraptable}
 % 
@@
 
-\section{Experiment Design}\label{sec:expersetup}
+\section{Methodology}\label{sec:expersetup}
 
@@
 
+We recruited 250 participants for our experiment through Amazon Mechanical Turk.
+Each participant was first presented with some brief training material.
+Before presenting the lineups for the hypothesis tests, each participant was shown two trial plots: one where the alternative plot was the most different from the others due to its relatively \textit{complex} structure, while the other trial included an alternative plot that was most different from the others due to its comparatively \textit{simple} structure.
+Only when participants were able to correctly identify the alternative plot from the trial lineups were they allowed to begin the experiment.
+
+Each participant was randomly assigned 13 lineups to look at.
+They were asked to select one or more plots that they perceived as ``most different" from the others, and provide a reasoning for their choice: ``most simple overall structure",``most complex overall structure", or ``other".
+If they selected ``other", they were required to describe their reasoning.
+The language in the reasoning is purposefully vague to avoid contextual bias.
+
+Twelve of the 13 lineups that the participants saw were used for the significance testing and the visual power methods discussed in Sections~\ref{sec:setupst} and \ref{sec:setupvp}.
+Among the 12 lineups were four from each difficulty level (easy, medium, and hard), and two of the four were from each condition (negative-positive, or positive-reverse).
+Each participant saw each of the six models twice at some combination of difficulty and condition.
+%The six parameters, $\beta_1, \dots, \beta_6$ were set to three different values according to how difficult we thought picking the data plot from the lineup would be for our participants, and were also set to be less than or greater than the initial estimate, creating the lineup type variable.%\hh{what about the first 12 plots?
+6 parameters, 3 levels, 2 directions}
+The last lineups shown to participants contained the true data from the 112th senate shown in Section~\ref{sec:models} placed among five other plots from  one of the models M3, M4, M5, and M7 as discussed in Section~\ref{sec:gofsetup}.
+Upon completion of the 13 lineups, each participant was paid \$1.75
+
+% future:  which took an average of 308 seconds per person, 
 \subsection{Significance Testing}\label{sec:setupst}
@@
 
-\subsection{Execution}
-
-We recruited 250 participants for our experiment through Amazon Mechanical Turk.
-Each participant was first presented with some brief training material.
-Before presenting the lineups for the hypothesis tests, each participant was shown two trial plots: one where the alternative plot was the most different from the others due to its relatively \textit{complex} structure, while the other trial included an alternative plot that was most different from the others due to its comparatively \textit{simple} structure.
-Only when participants were able to correctly identify the alternative plot from the trial lineups were they allowed to begin the experiment.
-
-Each participant was randomly assigned 13 lineups to look at.
-They were asked to select one or more plots that they perceived as ``most different" from the others, and provide a reasoning for their choice: ``most simple overall structure",``most complex overall structure", or ``other".
-If they selected ``other", they were required to describe their reasoning.
-The language in the reasoning is purposefully vague to avoid contextual bias.
-
-Twelve of the 13 lineups that the participants saw were used for the significance testing and the visual power methods discussed in Sections~\ref{sec:setupst} and \ref{sec:setupvp}.
-Among the 12 lineups were four from each difficulty level (easy, medium, and hard), and two of the four were from each condition (negative-positive, or positive-reverse).
-Each participant saw each of the six models twice at some combination of difficulty and condition.
-%The six parameters, $\beta_1, \dots, \beta_6$ were set to three different values according to how difficult we thought picking the data plot from the lineup would be for our participants, and were also set to be less than or greater than the initial estimate, creating the lineup type variable.%\hh{what about the first 12 plots?
-6 parameters, 3 levels, 2 directions}
-The last lineups shown to participants contained the true data from the 112th senate shown in Section~\ref{sec:models} placed among five other plots from  one of the models M3, M4, M5, and M7 as discussed in Section~\ref{sec:gofsetup}.
-Upon completion of the 13 lineups, which took an average of 308 seconds per person, each participant was paid \$1.75
+%\subsection{Execution}
 
@@
 Some simulation-based methods have been developed using what auxiliary statistics such as the indegree or outdegree distribution on the nodes \citep{RSienamanual}.
-In \texttt{RSiena}, the \texttt{sienaGOF} function performs goodness-of-fit testing as follows:
-\begin{enumerate}
-\item Auxiliary statistics, such as the cumulative outdegree distribution on the nodes, are computed on the observed data ($\mathbf{u}_d$) and on $N$ observations simulated from the model ($\mathbf{u}_1 \dots \mathbf{u}_N$).
-\item The mean $\overline{\mathbf{u}}$ and covariance matrix $\mathbf{S}$ are computed from the $N$ simulations, and the Mahalanobis distance, $d_M(\mathbf{u}_d)$ from the observed statistics to the distribution of the simulated statistics is computed:
-\begin{equation}\label{eq:mahal}
-d_M(\mathbf{u}_d) = \sqrt{(\mathbf{u}_d - \overline{\mathbf{u}})' \mathbf{S}^{-1} (\mathbf{u}_d - \overline{\mathbf{u}})}
-\end{equation}
-\item The Mahalanobis distance for each of the $N$ simulations is calculated and $d_M(\mathbf{u}_d)$ is compared to this distribution of distances.
-\item An empirical $p$-value is found by computing the proportion of simulated distances found in step 4 that are as large or larger than $d_M(\mathbf{u}_d)$.% A model is thus considered a good fit to the data if $p$ is large.
-\end{enumerate}
-<<gofsiena, fig.height=2.5, fig.cap="An example of what a goodness-of-fit plot from \\texttt{RSiena} looks like.
-The overlaid boxplots and violin plots show the distribution of each of the outdegree count values on the simulated networks, and the red points and lines are the observed data values.", cache = TRUE, eval=FALSE>>=
-load("data/ansnullpaper.rda")
-library(lattice)
-#gof1 <- sienaGOF(ansnull, OutdegreeDistribution, varName ="friendship", join = FALSE)
-gof1 <- readr::read_rds("data/gof-data.rds")
-sims1 <- data.frame(gof1$`Period 1`$Simulations)
-dat1 <- gof1$`Period 1`$Observations
-dat1 %>% data.frame() %>% gather(outdegree, val) %>%
-  mutate(outdegree = parse_number(outdegree)-1) -> dat1
-p1 <- sims1 %>%
-  mutate(sim = row_number()) %>%
-  gather(outdegree, val, X1:X9) %>%
-  mutate(outdegree = parse_number(outdegree) - 1) %>%
-  filter(outdegree <= 6) %>%
-  ggplot(aes(x = outdegree, y = val)) + 
-  geom_boxplot(aes(group = outdegree), size = .5, outlier.shape = "x", outlier.size = 5) + 
-  geom_violin(aes(group = outdegree), bw = 2, fill = NA) + 
-  geom_point(data = dat1, color = 'red') + 
-  geom_line(data= dat1, color = 'red') + 
-  geom_text(data = dat1, aes(label = val), hjust = -.5) + 
-  labs(x = "Outdegree (p = 0.154)", y = "Statistic", 
-       title = "Goodness-of-Fit: Outdegree distribution period 1") + 
-  ThemeNoNet
-p1
+In \texttt{RSiena},
 @
@@
 
-\npdecimalsign{.}
-%\nprounddigits{3}
-
-\begin{table}
-\centering
-\begin{tabular}{cn{5}{2}n{5}{3}rr}\hline
-Parameter & \multicolumn{1}{r}{Estimate} & \multicolumn{1}{r}{Std Error} & $p$-value & Odds Multiplier \\ \hline\hline
-$\eta_{1+}$ & \Sexpr{ests[7]} & \Sexpr{se[7]} & \Sexpr{pval[7]}$^{\ddagger}$ & \Sexpr{odds[7]} \\
-$\eta_{1-}$ & \Sexpr{ests[1]} & \Sexpr{se[1]} & \Sexpr{pval[1]}$^{\dagger}$ & \Sexpr{odds[1]}\\
-$\gamma_{1+}$ & \Sexpr{ests[19]} & \Sexpr{se[19]} & \Sexpr{pval[19]}$^{\ddagger}$ & \Sexpr{odds[19]} \\
-$\gamma_{1-}$  & \Sexpr{ests[13]} & \Sexpr{se[13]} & \Sexpr{pval[13]}$^{\dagger}$ & \Sexpr{odds[13]} \\ \hline
-$\eta_{2+}$  & \Sexpr{ests[10]} & \Sexpr{se[10]} & \Sexpr{pval[10]} & \Sexpr{odds[10]} \\
-$\eta_{2-}$  & \Sexpr{ests[4]} & \Sexpr{se[4]} & \Sexpr{pval[4]}$^{\ddagger}$ & \Sexpr{odds[4]} \\
-$\gamma_{2+}$ & \Sexpr{ests[22]} & \Sexpr{se[22]} & \Sexpr{pval[22]}$^{*}$ & \Sexpr{odds[22]} \\
-$\gamma_{2-}$  & \Sexpr{ests[16]} & \Sexpr{se[16]} & \Sexpr{pval[16]}$^{\ddagger}$ & \Sexpr{odds[16]} \\ \hline
-$\eta_{3+}$  & \Sexpr{ests[8]} & \Sexpr{se[8]} & \Sexpr{pval[8]}$^{\dagger}$ & \Sexpr{odds[8]} \\ 
-$\eta_{3-}$ & \Sexpr{ests[2]} & \Sexpr{se[2]} & \Sexpr{pval[2]}$^{\dagger}$ & \Sexpr{odds[2]} \\
-$\gamma_{3+}$ & \Sexpr{ests[20]} & \Sexpr{se[20]} & \Sexpr{pval[20]}$^{\dagger}$ & \Sexpr{odds[20]}  \\
-$\gamma_{3-}$   & \Sexpr{ests[14]} & \Sexpr{se[14]} & \Sexpr{pval[14]}$^{*}$ & \Sexpr{odds[14]} \\ \hline
-$\eta_{4+}$ & \Sexpr{ests[9]} & \Sexpr{se[9]} & \Sexpr{pval[9]}$^{**}$ & \Sexpr{odds[9]} \\
-$\eta_{4-}$ & \Sexpr{ests[3]} & \Sexpr{se[3]} & \Sexpr{pval[3]}$^{**}$ & \Sexpr{odds[3]}  \\
-$\gamma_{4+}$ & \Sexpr{ests[21]} & \Sexpr{se[21]} & \Sexpr{pval[21]}$^{**}$ & \Sexpr{odds[21]} \\ 
-$\gamma_{4-}$  & \Sexpr{ests[15]} & \Sexpr{se[15]} & \Sexpr{pval[15]} & \Sexpr{odds[15]} \\ \hline
-$\eta_{5+}$  & \Sexpr{ests[12]} & \Sexpr{se[12]} & \Sexpr{pval[12]}$^{*}$ & \Sexpr{odds[12]} \\
-$\eta_{5-}$  & \Sexpr{ests[6]} & \Sexpr{se[6]} & \Sexpr{pval[6]}$^{\ddagger}$ & \Sexpr{odds[6]} \\
-$\gamma_{5+}$ & \Sexpr{ests[24]} & \Sexpr{se[24]} & \Sexpr{pval[24]}$^{*}$ & \Sexpr{odds[24]} \\ 
-$\gamma_{5-}$ & \Sexpr{ests[18]} & \Sexpr{se[18]} & \Sexpr{pval[18]}$^{\ddagger}$ & \Sexpr{odds[18]} \\ \hline
-$\eta_{6+}$  & \Sexpr{ests[11]} & \Sexpr{se[11]} & \Sexpr{pval[11]} & \Sexpr{odds[11]} \\
-$\eta_{6-}$ & \Sexpr{ests[5]} & \Sexpr{se[5]} & \Sexpr{pval[5]}$^{\ddagger}$ & \Sexpr{odds[5]}  \\
-$\gamma_{6+}$ & \Sexpr{ests[23]} & \Sexpr{se[23]} & \Sexpr{pval[23]} & \Sexpr{odds[23]} \\
-$\gamma_{6-}$ & \Sexpr{ests[17]} & \Sexpr{se[17]} & \Sexpr{pval[17]}$^{\ddagger}$ & \Sexpr{odds[17]} \\ \hline
-$\sigma^2_{\delta}$ & \Sexpr{round(sigma_delta, 4)} & \multicolumn{1}{r}{--} & -- & --\\
-$\sigma^2_{\epsilon}$ &  \Sexpr{round(sigma_epsilon, 4)}  & \multicolumn{1}{r}{--} & -- & --\\
-\hline 
-\end{tabular}
-\caption{\label{tab:glmmests}Summary of the results from fitting the model given in Equation~\ref{eq:glmm}. Significance levels: * - $< 0.10$; ** - $<0.05$; $\dagger$ - $<0.01$; $\ddagger$ - $<0.001$}
-\end{table}
-
-% \begin{table}
-% \centering
-% \begin{tabular}{lcccc}
-% Parameter & Estimate & Std Error & $p$-value & Odds Multiplier \\
-% $\mu$ & \Sexpr{ests[1]} & \Sexpr{se[1]} & \Sexpr{pval[1]} & \Sexpr{odds[1]}\\
-% $\alpha_1$ & \Sexpr{ests[2]} & \Sexpr{se[2]} & \Sexpr{pval[2]} & \Sexpr{odds[2]} \\
-% $\alpha_2$ & \Sexpr{ests[3]} & \Sexpr{se[3]} & \Sexpr{pval[3]} & \Sexpr{odds[3]}  \\
-% $\alpha_3$  & \Sexpr{ests[4]} & \Sexpr{se[4]} & \Sexpr{pval[4]} & \Sexpr{odds[4]} \\
-% $\alpha_4$ & \Sexpr{ests[5]} & \Sexpr{se[5]} & \Sexpr{pval[5]} & \Sexpr{odds[5]}  \\
-% $\alpha_5$  & \Sexpr{ests[6]} & \Sexpr{se[6]} & \Sexpr{pval[6]} & \Sexpr{odds[6]} \\
-% $\theta$  & \Sexpr{ests[7]} & \Sexpr{se[7]} & \Sexpr{pval[7]} & \Sexpr{odds[7]} \\
-% $\gamma$  & \Sexpr{ests[8]} & \Sexpr{se[8]} & \Sexpr{pval[8]} & \Sexpr{odds[8]} \\ 
-% $(\alpha\theta)_1$  & \Sexpr{ests[9]} & \Sexpr{se[9]} & \Sexpr{pval[9]} & \Sexpr{odds[9]} \\
-% $(\alpha\theta)_2$  & \Sexpr{ests[10]} & \Sexpr{se[10]} & \Sexpr{pval[10]} & \Sexpr{odds[10]} \\
-% $(\alpha\theta)_3$  & \Sexpr{ests[11]} & \Sexpr{se[11]} & \Sexpr{pval[11]} & \Sexpr{odds[11]} \\
-% $(\alpha\theta)_4$  & \Sexpr{ests[12]} & \Sexpr{se[12]} & \Sexpr{pval[12]} & \Sexpr{odds[12]} \\
-% $(\alpha\theta)_5$  & \Sexpr{ests[13]} & \Sexpr{se[13]} & \Sexpr{pval[13]} & \Sexpr{odds[13]} \\
-% $(\alpha\gamma)_1$  & \Sexpr{ests[14]} & \Sexpr{se[14]} & \Sexpr{pval[14]} & \Sexpr{odds[14]} \\
-% $(\alpha\gamma)_2$  & \Sexpr{ests[15]} & \Sexpr{se[15]} & \Sexpr{pval[15]} & \Sexpr{odds[15]} \\
-% $(\alpha\gamma)_3$  & \Sexpr{ests[16]} & \Sexpr{se[16]} & \Sexpr{pval[16]} & \Sexpr{odds[16]} \\
-% $(\alpha\gamma)_4$  & \Sexpr{ests[17]} & \Sexpr{se[17]} & \Sexpr{pval[17]} & \Sexpr{odds[17]} \\
-% $(\alpha\gamma)_5$  & \Sexpr{ests[18]} & \Sexpr{se[18]} & \Sexpr{pval[18]} & \Sexpr{odds[18]} \\
-% $\theta\gamma$  & \Sexpr{ests[19]} & \Sexpr{se[19]} & \Sexpr{pval[19]} & \Sexpr{odds[19]} \\
-% $(\alpha\theta\gamma)_1$ & \Sexpr{ests[20]} & \Sexpr{se[20]} & \Sexpr{pval[20]} & \Sexpr{odds[20]}  \\
-% $(\alpha\theta\gamma)_2$  & \Sexpr{ests[21]} & \Sexpr{se[21]} & \Sexpr{pval[21]} & \Sexpr{odds[21]} \\
-% $(\alpha\theta\gamma)_3$  & \Sexpr{ests[22]} & \Sexpr{se[22]} & \Sexpr{pval[22]} & \Sexpr{odds[22]} \\
-% $(\alpha\theta\gamma)_4$  & \Sexpr{ests[23]} & \Sexpr{se[23]} & \Sexpr{pval[23]} & \Sexpr{odds[23]} \\
-% $(\alpha\theta\gamma)_5$  & \Sexpr{ests[24]} & \Sexpr{se[24]} & \Sexpr{pval[24]} & \Sexpr{odds[24]} \\ 
-% $\sigma^2_{\delta}$ & 0.5638 \\
-% $\sigma^2_{\epsilon}$ & 0.3416
-% \end{tabular}
-% \caption{\label{tab:glmmests}Summary of the results from fitting the model given in Equation~\ref{eq:glmm}}
-% \end{table}
-
 <<predictglmm, fig.height=6, fig.cap='Predictions from our GLMM given in Equation~\\ref{eq:glmm}.

### 2019-08-06 ASApaper/supplement.Rnw
moving stuff to supp mat

\documentclass[12pt]{article}
% \usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
% \geometry{letterpaper}                   % ... or a4paper or a5paper or ...
%\usepackage{graphicx}
\usepackage[font=small,skip=5pt]{caption}
\usepackage{subcaption}
%\usepackage{animate}
\usepackage{afterpage}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{amsfonts}
%\usepackage{color}
\usepackage{multirow}
\usepackage{rotating}
\usepackage[dvipsnames,svgnames,table]{xcolor}
\usepackage{xr}
\usepackage{xr-hyper}
\usepackage{hyperref}
\usepackage{numprint}
\usepackage{booktabs}

%\usepackage{appendix}

\graphicspath{{sm-figure/}}
% \usepackage{endfloat} % Figures to the end of the document
\newcommand{\st}[1]{{\color{orange} #1}}
\newcommand{\hh}[1]{{\color{magenta} #1}}

\DeclareGraphicsRule{.tif}{png}{.png}{`convert #1 `dirname #1`/`basename #1 .tif`.png}
%---------------------------------------------------
%                 Editing Commands
%\newcommand{\hh}[1]{{\color{magenta} #1}}
%\newcommand{\st}[1]{{\color{orange} #1}}

%---------------------------------------------------
%                 Placing Figures
\renewcommand{\topfraction}{0.99}	% max fraction of floats at top
\renewcommand{\bottomfraction}{0.8}	% max fraction of floats at bottom
\renewcommand{\textfraction}{0.05}	% allow minimal text w. figs

%---------------------------------------------------
% Define new environment
\newtheorem{theorem}{Theorem}[section]
\newtheorem{algorithm}[theorem]{Algorithm}
%---------------------------------------------------

%\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

\externaldocument{paper}

\begin{document}

\title{\bf Supplementary Materials: Visual Inference for a Social Network Model}
  \author{Samantha Tyner\thanks{
    Department of Statistics and Statistical Laboratory, Iowa State University\\
    and \\
    Heike Hofmann\\
    Department of Statistics and Statistical Laboratory, Iowa State University}
  \maketitle
  }

   \bigskip
  \bigskip
  \bigskip
  \begin{center}
    {\LARGE\bf Supplementary Materials: Visual Inference for a Social Network Model}
\end{center}

\spacingset{1.45} % DON'T change the spacing!
<<setup-ch-2, echo = FALSE, message = FALSE, warning = FALSE>>=
options(replace.assign=TRUE,width=70, digits=3)
require(knitr)
opts_chunk$set(fig.path='sm-figure/', cache.path='sm-cache/', fig.align='center', fig.pos='h', out.width='.99\\textwidth', par=TRUE, cache=FALSE, concordance=TRUE, autodep=TRUE, message=F, warning=F, echo = FALSE, dev="cairo_pdf", fig.width = 6, fig.height = 6)#, root.dir = "~/Desktop/Dissertation/SAOM-removing-blindfold/")
@
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% intro: talk about the sections of the appendix
Blah

\appendix % to get sections with letters instead of numbers
\section{Model Details}\label{sec:sm-model}

The continuous-time Markov chain (CTMC) family of models we use were first introduced in \citet{saompaper}. We describe their basic structure here, and full detail can be found in \citet{saompaper, snijders:2010}.

\noindent \textbf{Rate Function:}  The rate function dictates when network changes are made and which actor can make them. An actor, $i$, is chosen make a change in its ties one of the other nodes $j$. In general, the rate function can include structural and node covariate parameters into account so that each actor has a different rate of change. However, we choose a simple rate function that is constant over all nodes in a given time period, because we focus on interpreting the parameters of the objective function which directly impact the overall network structure. We denote the rate from $t_{m}$ to $t_{m+1}$ as $\alpha_m$ for $m = 1, \dots, M-1$, where $M$ is the number of time points at which the network was observed. Using this notation, the \textit{waiting time} to the next chance for actor $i$ to make a change is exponentially distributed with expected value $\alpha_m^{-1}$. Since the rate is the same for all actors, the waiting time for \textit{any} actor to get the opportunity to change its set of ties is also exponentially distributed with expected value $(n\alpha_m)^{-1}$.

\noindent \textbf{Objective Function:} After actor $i$ is selected to make a change, it randomly picks one of its current ties, $x_{ij}$, to change. %The probability that actor $i$ changes its current tie to actor $j$ is determined by the \textit{objective function} of the model and a random error term $U$ with log-Weibull distribution \citep{modelsSnijders}.  %\hh{XXX can you paraphrase the purpose of U in half a sentence?}
%The random component is assumed to have log-Weibull distribution with location parameter $\mu = 0$ and scale parameter $\sigma = 0$ (see \citet{modelsSnijders}), so the probability density function of $U$ is
%\begin{equation}\label{eq:logweibull}
%g(u) = \exp\{-(u + \exp\{-u\})\}.
%\end{equation}
%The objective function, $f_i$, is the driving force of a SAOM. 
Actor $i$ aims to maximize the objective function $f_i$ given the current state of the network, $x$ and the node-level covariates, $\mathbf{Z}$. This function is defined as:% which has the form
\begin{equation}\label{eq:objective}
f_i(x, \boldsymbol{\beta}, \mathbf{Z}) = \sum_{k = 1}^K \beta_k s_{ik}(x, \mathbf{Z}),
\end{equation}
where $\boldsymbol{\beta} = (\beta_1, \dots, \beta_K)$ are additional model parameters, each associated with some statistics, $s_{i1}(x, \mathbf{Z}), \dots, s_{iK}(x, \mathbf{Z})$, calculated for actor $i$ at the current network state $x$. %The statistics can be very simple, like outdegree of the node, $s_{i1}(x) = \sum_{i\neq j} x_{ij}$, or more complicated, like the {\it transitive triplets jumping to different covariate} statistic, $s_i(x, \mathbf{Z}) = \sum_{i \neq j \neq h} x_{ij}x_{ih}x_{hj} \cdot \mathbb{I}(z_i = z_h \neq z_j)$. 
At least two parameters must be included in the objective function: density and reciprocity \citep{RSienamanual}. We denote the density, or out-degree, parameter by $\beta_1$ and the associated statistic as $s_{i1}(x)=\sum_{j} x_{ij}$ and we denote the reciprocity parameter by $\beta_2$ and the associated statistic as $s_{i2}(x) = \sum_{j} x_{ij}x_{ji}$. We will refer to the very simple model with only these two parameters in the objective function as model M1. We define additional parameters and models of interest in Section~\ref{sec:models}  %, plus many more. At last count, in the software we use to fit these models to network data, 
Version 1.2-3 of \texttt{RSiena} \citep{RSiena}, the software we use to fit CTMC models to data, provides over 80 possible effects that can be included in the objective function. %We discuss these and other statistics in more detail in Section~\ref{sec:models}.

The objective function $f_{i}(x, \boldsymbol{\beta}, \mathbf{Z})$ dictates the \textit{transition probability}, $p_{ij}$ of the network changing from its current state $x$ to the state $x(i\leadsto j)$, which is identical to $x$ except for $x_{ij}$: $x_{ij}(i \leadsto j) = 1 - x_{ij}$. The transition probability is
\begin{equation}\label{eq:transprob}
p_{ij} = \dfrac{\exp\{f_i(x(i \leadsto j), \boldsymbol{\beta}, \mathbf{Z})\}}{\sum_{h} \exp\{f_i(x(i \leadsto h), \boldsymbol{\beta}, \mathbf{Z})\}},
\end{equation}
dictating which edge node $i$ changes. Thus, the actor is more likely to make changes that increase the value of their objective function, and no change is most likely when any change decreases the value of the objective function.

\subsection{Goodness-of-Fit Testing}

The software \texttt{RSiena} contains methods for performing goodness-of-fit tests for the CTMC models. The \texttt{sienaGOF()} function performs goodness-of-fit testing as follows:
\begin{enumerate}
\item Auxiliary statistics, such as the cumulative outdegree distribution on the nodes, are computed on the observed data ($\mathbf{u}_d$) and on $N$ observations simulated from the model ($\mathbf{u}_1 \dots \mathbf{u}_N$). 
\item The mean $\overline{\mathbf{u}}$ and covariance matrix $\mathbf{S}$ are computed from the $N$ simulations, and the Mahalanobis distance, $d_M(\mathbf{u}_d)$ from the observed statistics to the distribution of the simulated statistics is computed:
\begin{equation}\label{eq:mahal}
d_M(\mathbf{u}_d) = \sqrt{(\mathbf{u}_d - \overline{\mathbf{u}})' \mathbf{S}^{-1} (\mathbf{u}_d - \overline{\mathbf{u}})}
\end{equation}
\item The Mahalanobis distance for each of the $N$ simulations is calculated and $d_M(\mathbf{u}_d)$ is compared to this distribution of distances.
\item An empirical $p$-value is found by computing the proportion of simulated distances found in step 4 that are as large or larger than $d_M(\mathbf{u}_d)$.% A model is thus considered a good fit to the data if $p$ is large.
\end{enumerate}
<<gofsiena, fig.height=2.5, fig.cap="An example of what a goodness-of-fit plot from \\texttt{RSiena} looks like. The overlaid boxplots and violin plots show the distribution of each of the outdegree count values on the simulated networks, and the red points and lines are the observed data values.", echo=FALSE>>=
library(RSiena)
library(tidyverse)
load("data/ansnullpaper.rda")
library(lattice)
#gof1 <- sienaGOF(ansnull, OutdegreeDistribution, varName ="friendship", join = FALSE)
gof1 <- readr::read_rds("data/gof-data.rds")
sims1 <- data.frame(gof1$`Period 1`$Simulations)
dat1 <- gof1$`Period 1`$Observations
dat1 %>% data.frame() %>% gather(outdegree, val) %>%
  mutate(outdegree = parse_number(outdegree)-1) -> dat1
p1 <- sims1 %>%
  mutate(sim = row_number()) %>%
  gather(outdegree, val, X1:X9) %>%
  mutate(outdegree = parse_number(outdegree) - 1) %>%
  filter(outdegree <= 6) %>%
  ggplot(aes(x = outdegree, y = val)) + 
  geom_boxplot(aes(group = outdegree), size = .5) + 
  geom_violin(aes(group = outdegree), bw = 2, fill = NA) + 
  geom_point(data = dat1, color = 'red', alpha = .5) + 
  geom_line(data= dat1, color = 'red', alpha=.5) + 
  geom_text(data = dat1, aes(label = val), hjust = -.5) + 
  labs(x = "Outdegree (p = 0.154)", y = "Statistic", 
       title = "Goodness-of-Fit: Outdegree distribution period 1") 
p1
@

\section{Data}\label{sec:sm-data}

 Details of how this data can be downloaded are provided by François Briatte at \url{github.com/briatte/congress}. In the US Senate, senators often show support for a piece of legislation by co-sponsoring a bill authored by one of their colleagues. In a co-sponsorship network, ties are directed from senator $i$ to senator $j$ when senator $i$ signs on as a co-sponsor to the bill that senator $j$ authored. There are many hundreds of ties between senators when they are connected in this way, so we simplify the network by computing a single value for each senator-senator collaboration called the \textit{weighted propensity to co-sponsor} (WPC). This value is defined in \citet{senate} as

\begin{equation}\label{eq:sen1}
    WPC_{ij} = \sum\limits_{b=1}^{B_j} \frac{Y_{ij(b)}}{c_{j(b)}}\left(\sum\limits_{b=1}^{B_j} \frac{1}{c_{j(b)}}\right)^{-1}
\end{equation}
where $B_j$ is the number of bills in a congressional session authored by senator $j$, $c_{j(b)}$ is the number of co-sponsors on senator $j$'s $b^{th}$ bill, where $b \in \{1,\dots, B_j\}$, and $Y_{ij(b)}$ is an indicator variable that senator $i$ co-sponsored senator $j$'s $b^{th}$ bill. This measure ranges in value from 0 to 1, where $WPC_{ij} = 1$ if senator $i$ is a co-sponsor on every one of senator $j$'s bills and $WPC_{ij} = 0$ if senator $i$ is never a co-sponsor any of senator $j$'s bills. To simplify the problem, we construct a network with binary edges as follows: 
 \begin{equation}\label{eq:edgewpc}
  x_{ij} =
\begin{cases}
                                   1 & WPC_{ij} > 0.25 \\
                                   0 & WPC_{ij} \leq 0.25
\end{cases}
\end{equation}
so that only strong ties between senators are in the network.

\section{Results}

\subsection{Visual Power}\label{sec:sm-res-vp}

<<glmmres>>=
load("data/finalglmm31-2.RDA")
mod <- model2randomscalesize
res <- summary(mod)[[10]]
ests <- as.numeric(round(res[,1], 4))
oddsidx <- which(exp(ests) < .0001)
oddsidx2 <- which(exp(ests) > 10000)
odds <- round(exp(ests), 4)
odds[oddsidx] <- "$<10^{-4}$"
odds[oddsidx2] <- "$>10^4$"
se <- as.numeric(round(res[,2], 4))
pvalidx <- which(res[,4] < .0001)
pval <- sprintf("%.4f",round(res[,4],4))
pval[pvalidx] <- "$<$0.0001"
sigma_epsilon <- summary(model2randomscalesize)[[13]]$nick_name[[1]]
sigma_delta <- summary(model2randomscalesize)[[13]]$pic_id[[1]]
@

\npdecimalsign{.}
%\nprounddigits{3}

\begin{table}
\centering
\begin{tabular}{cn{5}{2}n{5}{3}rr}\hline
Parameter & \multicolumn{1}{r}{Estimate} & \multicolumn{1}{r}{Std Error} & $p$-value & Odds Multiplier \\ \hline\hline
$\eta_{1+}$ & \Sexpr{ests[7]} & \Sexpr{se[7]} & \Sexpr{pval[7]}$^{\ddagger}$ & \Sexpr{odds[7]} \\
$\eta_{1-}$ & \Sexpr{ests[1]} & \Sexpr{se[1]} & \Sexpr{pval[1]}$^{\dagger}$ & \Sexpr{odds[1]}\\
$\gamma_{1+}$ & \Sexpr{ests[19]} & \Sexpr{se[19]} & \Sexpr{pval[19]}$^{\ddagger}$ & \Sexpr{odds[19]} \\
$\gamma_{1-}$  & \Sexpr{ests[13]} & \Sexpr{se[13]} & \Sexpr{pval[13]}$^{\dagger}$ & \Sexpr{odds[13]} \\ \hline
$\eta_{2+}$  & \Sexpr{ests[10]} & \Sexpr{se[10]} & \Sexpr{pval[10]} & \Sexpr{odds[10]} \\
$\eta_{2-}$  & \Sexpr{ests[4]} & \Sexpr{se[4]} & \Sexpr{pval[4]}$^{\ddagger}$ & \Sexpr{odds[4]} \\
$\gamma_{2+}$ & \Sexpr{ests[22]} & \Sexpr{se[22]} & \Sexpr{pval[22]}$^{*}$ & \Sexpr{odds[22]} \\
$\gamma_{2-}$  & \Sexpr{ests[16]} & \Sexpr{se[16]} & \Sexpr{pval[16]}$^{\ddagger}$ & \Sexpr{odds[16]} \\ \hline
$\eta_{3+}$  & \Sexpr{ests[8]} & \Sexpr{se[8]} & \Sexpr{pval[8]}$^{\dagger}$ & \Sexpr{odds[8]} \\ 
$\eta_{3-}$ & \Sexpr{ests[2]} & \Sexpr{se[2]} & \Sexpr{pval[2]}$^{\dagger}$ & \Sexpr{odds[2]} \\
$\gamma_{3+}$ & \Sexpr{ests[20]} & \Sexpr{se[20]} & \Sexpr{pval[20]}$^{\dagger}$ & \Sexpr{odds[20]}  \\
$\gamma_{3-}$   & \Sexpr{ests[14]} & \Sexpr{se[14]} & \Sexpr{pval[14]}$^{*}$ & \Sexpr{odds[14]} \\ \hline
$\eta_{4+}$ & \Sexpr{ests[9]} & \Sexpr{se[9]} & \Sexpr{pval[9]}$^{**}$ & \Sexpr{odds[9]} \\
$\eta_{4-}$ & \Sexpr{ests[3]} & \Sexpr{se[3]} & \Sexpr{pval[3]}$^{**}$ & \Sexpr{odds[3]}  \\
$\gamma_{4+}$ & \Sexpr{ests[21]} & \Sexpr{se[21]} & \Sexpr{pval[21]}$^{**}$ & \Sexpr{odds[21]} \\ 
$\gamma_{4-}$  & \Sexpr{ests[15]} & \Sexpr{se[15]} & \Sexpr{pval[15]} & \Sexpr{odds[15]} \\ \hline
$\eta_{5+}$  & \Sexpr{ests[12]} & \Sexpr{se[12]} & \Sexpr{pval[12]}$^{*}$ & \Sexpr{odds[12]} \\
$\eta_{5-}$  & \Sexpr{ests[6]} & \Sexpr{se[6]} & \Sexpr{pval[6]}$^{\ddagger}$ & \Sexpr{odds[6]} \\
$\gamma_{5+}$ & \Sexpr{ests[24]} & \Sexpr{se[24]} & \Sexpr{pval[24]}$^{*}$ & \Sexpr{odds[24]} \\ 
$\gamma_{5-}$ & \Sexpr{ests[18]} & \Sexpr{se[18]} & \Sexpr{pval[18]}$^{\ddagger}$ & \Sexpr{odds[18]} \\ \hline
$\eta_{6+}$  & \Sexpr{ests[11]} & \Sexpr{se[11]} & \Sexpr{pval[11]} & \Sexpr{odds[11]} \\
$\eta_{6-}$ & \Sexpr{ests[5]} & \Sexpr{se[5]} & \Sexpr{pval[5]}$^{\ddagger}$ & \Sexpr{odds[5]}  \\
$\gamma_{6+}$ & \Sexpr{ests[23]} & \Sexpr{se[23]} & \Sexpr{pval[23]} & \Sexpr{odds[23]} \\
$\gamma_{6-}$ & \Sexpr{ests[17]} & \Sexpr{se[17]} & \Sexpr{pval[17]}$^{\ddagger}$ & \Sexpr{odds[17]} \\ \hline
$\sigma^2_{\delta}$ & \Sexpr{round(sigma_delta, 4)} & \multicolumn{1}{r}{--} & -- & --\\
$\sigma^2_{\epsilon}$ &  \Sexpr{round(sigma_epsilon, 4)}  & \multicolumn{1}{r}{--} & -- & --\\
\hline 
\end{tabular}
\caption{\label{tab:glmmests}Summary of the results from fitting the model given in Equation~\ref{eq:glmm}. Significance levels: * - $< 0.10$; ** - $<0.05$; $\dagger$ - $<0.01$; $\ddagger$ - $<0.001$}
\end{table}

%%%%%%%%%%%%%%%%%%%%%% References %%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{asa}
\bibliography{bibliography}

\end{document}

### 2019-08-06 ASApaper/paper-child.Rnw
fix stuff with aux file & cross refs

@@
 
-\noindent \textbf{Rate Function:}  The rate function dictates when network changes are made and which actor can make them.
-An actor, $i$, is chosen make a change in its ties one of the other nodes $j$.
-In general, the rate function can include structural and node covariate parameters into account so that each actor has a different rate of change.
-However, we choose a simple rate function that is constant over all nodes in a given time period, because we focus on interpreting the parameters of the objective function which directly impact the overall network structure.
-We denote the rate from $t_{m}$ to $t_{m+1}$ as $\alpha_m$ for $m = 1, \dots, M-1$, where $M$ is the number of time points at which the network was observed.
-Using this notation, the \textit{waiting time} to the next chance for actor $i$ to make a change is exponentially distributed with expected value $\alpha_m^{-1}$.
-Since the rate is the same for all actors, the waiting time for \textit{any} actor to get the opportunity to change its set of ties is also exponentially distributed with expected value $(n\alpha_m)^{-1}$.
-
-\noindent \textbf{Objective Function:} After actor $i$ is selected to make a change, it randomly picks one of its current ties, $x_{ij}$, to change.
-%The probability that actor $i$ changes its current tie to actor $j$ is determined by the \textit{objective function} of the model and a random error term $U$ with log-Weibull distribution \citep{modelsSnijders}.
-%\hh{XXX can you paraphrase the purpose of U in half a sentence?}
-%The random component is assumed to have log-Weibull distribution with location parameter $\mu = 0$ and scale parameter $\sigma = 0$ (see \citet{modelsSnijders}), so the probability density function of $U$ is
-%\begin{equation}\label{eq:logweibull}
-%g(u) = \exp\{-(u + \exp\{-u\})\}.
-%\end{equation}
-%The objective function, $f_i$, is the driving force of a SAOM. 
-Actor $i$ aims to maximize the objective function $f_i$ given the current state of the network, $x$ and the node-level covariates, $\mathbf{Z}$. This function is defined as:% which has the form
-\begin{equation}\label{eq:objective}
-f_i(x, \boldsymbol{\beta}, \mathbf{Z}) = \sum_{k = 1}^K \beta_k s_{ik}(x, \mathbf{Z}),
-\end{equation}
-where $\boldsymbol{\beta} = (\beta_1, \dots, \beta_K)$ are additional model parameters, each associated with some statistics, $s_{i1}(x, \mathbf{Z}), \dots, s_{iK}(x, \mathbf{Z})$, calculated for actor $i$ at the current network state $x$.
-%The statistics can be very simple, like outdegree of the node, $s_{i1}(x) = \sum_{i\neq j} x_{ij}$, or more complicated, like the {\it transitive triplets jumping to different covariate} statistic, $s_i(x, \mathbf{Z}) = \sum_{i \neq j \neq h} x_{ij}x_{ih}x_{hj} \cdot \mathbb{I}(z_i = z_h \neq z_j)$.
-At least two parameters must be included in the objective function: density and reciprocity \citep{RSienamanual}.
-We denote the density, or out-degree, parameter by $\beta_1$ and the associated statistic as $s_{i1}(x)=\sum_{j} x_{ij}$ and we denote the reciprocity parameter by $\beta_2$ and the associated statistic as $s_{i2}(x) = \sum_{j} x_{ij}x_{ji}$.
-We will refer to the very simple model with only these two parameters in the objective function as model M1.
-We define additional parameters and models of interest in Section~\ref{sec:models}  %, plus many more.
-At last count, in the software we use to fit these models to network data,
-Version 1.2-3 of \texttt{RSiena} \citep{RSiena}, the software we use to fit CTMC models to data, provides over 80 possible effects that can be included in the objective function.
-%We discuss these and other statistics in more detail in Section~\ref{sec:models}.
-
-The objective function $f_{i}(x, \boldsymbol{\beta}, \mathbf{Z})$ dictates the \textit{transition probability}, $p_{ij}$ of the network changing from its current state $x$ to the state $x(i\leadsto j)$, which is identical to $x$ except for $x_{ij}$: $x_{ij}(i \leadsto j) = 1 - x_{ij}$.
-The transition probability is
-\begin{equation}\label{eq:transprob}
-p_{ij} = \dfrac{\exp\{f_i(x(i \leadsto j), \boldsymbol{\beta}, \mathbf{Z})\}}{\sum_{h} \exp\{f_i(x(i \leadsto h), \boldsymbol{\beta}, \mathbf{Z})\}},
-\end{equation}
-dictating which edge node $i$ changes.
-Thus, the actor is more likely to make changes that increase the value of their objective function, and no change is most likely when any change decreases the value of the objective function.%Node $i$ can also choose to \textit{not} change at all, which is most likely when the value of Equation~\ref{eq:objective} is larger for $x$ than for any $x(i \leadsto j)$.
-
-%\hh{XXX how is the probability computed in the case that no change is made? Why is the notation $j \equiv i$ used for no change?}\st{that's more for the computation. I'll adjust accordingly}
-
 \subsection{Visual Inference}\label{subsec:vi}

### 2019-08-08 ASApaper/supplement.Rnw
add all lineup results in appendix + images

@@
 \usepackage{booktabs}
+\usepackage{longtable}
 
@@
 
+\subsection{Lineup Summaries}
+
+<<res-table-lus, results = "asis">>=
+# wrap_pvsim <- function(x, k, m = 6){
+#   res <- vinference::pVsim(x = x, K = k, m = m, scenario = 3, target = 1)
+#   res[2]
+# }
+
+# turk22 <- read_csv("data/turk22.csv")
+# res_sim_pvals <- turk22 %>% group_by(pic_id) %>% summarise(n_datapick = sum(datapick), n_views = n()) %>% 
+#   separate(pic_id, into = c("difficulty", "lu_type", "param", "rep"), sep = 1:3) %>% 
+#   group_by(difficulty, lu_type, param) %>% 
+#   mutate(n_datapick_c = sum(n_datapick), n_views_c = sum(n_views)) %>% 
+#   mutate(single_pv =  map2_dbl(n_datapick, n_views, wrap_pvsim))
+# res_sim_pvals2 <- res_sim_pvals %>% 
+#   mutate(combined_pv =  map2_dbl(n_datapick_c, n_views_c, wrap_pvsim))
+
+sig_marks <- function(p){
+  if(p < .1 & p >= .05){
+    return("^{*}$")
+  }
+   if(p < .05 & p >= .01){
+    return("^{**}$")
+   }
+   if(p < .01 & p >= .001){
+    return("^{\\dagger}$")
+   }
+   if(p < .001){
+    return("^{\\ddagger}$")
+   } else{
+    return("$")
+  }
+}
+
+lineup_pvals <- read_csv("data/vinf_pvals_res.csv")
+pvals_table <- lineup_pvals %>% 
+  mutate(sig_ind = map_chr(single_pv, sig_marks)) %>%
+  mutate(res = str_c("$",n_datapick,"/", n_views, sig_ind, collapse = NULL)) %>% 
+  group_by(difficulty, lu_type, param) %>% 
+  mutate(combined_pv = round(mean(combined_pv), 3)) %>%
+  select(difficulty:rep, combined_pv, res) %>% 
+  spread(rep, res) %>% select(1:3, 5:7, 4) %>% 
+  mutate(sig_ind_c = map_chr(combined_pv, sig_marks))
+pvals_table$combined_pv[which(pvals_table$combined_pv < 0.0001)] <- "$< 10^{-4}$"
+pvals_table$combined_pv <- paste0(pvals_table$combined_pv, "$", pvals_table$sig_ind_c)
+pvals_table$combined_pv[which(pvals_table$sig_ind_c == "$")] <- str_remove_all(pvals_table$combined_pv[which(pvals_table$sig_ind_c == "$")], "\\$")
+
+pvals_table %>% select(1:7) %>% 
+  arrange(param, lu_type, difficulty) %>% ungroup %>%  
+  mutate(lu_type = recode(lu_type, `1` = "1", `2` = "-1", `9` = "GoF"), 
+         difficulty = recode(difficulty, `1` = "Easy", `2` = "Med.", `3` = "Hard", `9` = "GoF"),
+         param = recode(param, `1`= "$\\beta_1$", `2`="$\\beta_2$", `3`="$\\beta_3$", `4`="$\\beta_4$", `5`="$\\beta_5$", `6`="$\\beta_6$", `7` = "M7")) -> pretty_table
+
+lus <- tibble(filename = list.files("img/lineup-pngs/"))
+
+lus %>% mutate(filename2 = filename) %>% separate(filename2, into = c("pic", "id", "altplot", "png")) %>% 
+  select(filename, id, altplot) %>% separate(altplot, into = c("plot", "alt_id"), sep = 4) %>% 
+  select(filename, id, alt_id) %>% 
+  separate(id, into = c("difficulty", "lu_type", "param", "rep"), sep = 1:3) %>% 
+  arrange(param, lu_type, difficulty) -> lus
+
+pretty_table <- mutate(pretty_table, pic_file = str_remove(lus$filename, ".png"))
+
+insert_image <- pretty_table %>% glue::glue_data("\\includegraphics[width=3cm]{{img/lineup-pngs/{pic_file}}}")
+
+pretty_table %>% mutate(pic_insert = insert_image) %>% select(9, 1:7) %>%
+  knitr::kable(format = "latex", escape = F, col.names = c("Lineup", "Difficulty", "Type", "Param.", "Rep 1", "Rep 2", "Rep 3", "Global p-val."), longtable=T, caption = "Here is a caption.
+Put stuff here later", booktabs=T)
+
+#pic_files <- turk22 %>% select(pic_id, pic_name) %>% distinct()
+
+#magick::image_read_svg(path = paste0("/Users/sctyner/Desktop/Dissertation/lineups/experiments/turk22/plots/", pic_files$pic_name[1]))
+
+@
+
+<<res-table-lus-reas, results = "asis", eval = FALSE>>=
+# ok, Heike was right. delving into the reasons is NOT worth the effort. 
+turk22 <- read_csv("data/turk22.csv")
+turk22 %>% group_by(pic_id, choice_reason) %>% 
+  mutate(choice_reason2 = str_detect(choice_reason, "Other")) %>% 
+  mutate(choice_reason3 = ifelse(choice_reason2, "Other", choice_reason)) %>% 
+  group_by(pic_id, choice_reason3) %>% count()
+mydates <- turk22$start_time
+class(mydates) <- c('POSIXt','POSIXct')
+turk22 %>%  mutate(time2 = mydates) %>% #why are there two???????
+  filter(choice_reason == "Most complex overall structure" | choice_reason == "Most complicated overall structure" ) %>%
+  ggplot() + 
+  geom_histogram(aes(x = as.Date(time2), fill = choice_reason))
+
+turk22 %>% group_by(pic_id, choice_reason) %>% 
+  mutate(choice_reason2 = str_detect(choice_reason, "Other")) %>% 
+  mutate(choice_reason3 = ifelse(choice_reason2, "Other", choice_reason)) %>% 
+  ggplot(aes(x = start_time, y = end_time, color = choice_reason3)) + 
+  geom_point() + 
+  geom_abline() + 
+  facet_wrap(~choice_reason3)
+
+@
+
 \subsection{Visual Power}\label{sec:sm-res-vp}
@@
 res <- summary(mod)[[10]]
-ests <- as.numeric(round(res[,1], 4))
+ests <- as.numeric(round(res[,1], 2))
 oddsidx <- which(exp(ests) < .0001)
 oddsidx2 <- which(exp(ests) > 10000)
-odds <- round(exp(ests), 4)
+odds <- ifelse(exp(ests) > 1, round(exp(ests), 2), round(exp(ests), 3))
 odds[oddsidx] <- "$<10^{-4}$"
 odds[oddsidx2] <- "$>10^4$"
-se <- as.numeric(round(res[,2], 4))
+se <- as.numeric(round(res[,2],2))
 pvalidx <- which(res[,4] < .0001)
-pval <- sprintf("%.4f",round(res[,4],4))
-pval[pvalidx] <- "$<$0.0001"
+pval <- sprintf("%.4f",round(res[,4],3))
+pval[pvalidx] <- "$<10^{-4}$"
 sigma_epsilon <- summary(model2randomscalesize)[[13]]$nick_name[[1]]
 sigma_delta <- summary(model2randomscalesize)[[13]]$pic_id[[1]]
+@
+
+<<glmres_predict, eval = FALSE>>=
+predict(model2randomscalesize, type = "link")
+predict(model2randomscalesize, type = "response")
+
+str(model2randomscalesize)
+
  @

### 2019-08-08 ASApaper/paper-child.Rnw
move effects table from paper to supplement. something is funky with fonts in pdfs on plots

@@
 
-\begin{table}
-\centering
-\begin{tabular}{|lccccccccc|}
-\toprule
-Model & $\hat{\alpha}_1$ & $\hat{\alpha}_2$ & $\hat{\alpha}_3$ & $\hat{\beta}_1$ & $\hat{\beta}_2$ & $\hat{\beta}_3$ & $\hat{\beta}_4$ & $\hat{\beta}_5$ & $\hat{\beta}_6$ \\
-\midrule
-M1 & \Sexpr{round(modelmeans$alpha1[1], 3)} & \Sexpr{round(modelmeans$alpha2[1], 3)} & \Sexpr{round(modelmeans$alpha3[1], 3)} & \Sexpr{round(modelmeans$beta1[1], 3)} & \Sexpr{round(modelmeans$beta2[1], 3)} & -- & -- & -- & --\\
-M3 & \Sexpr{round(modelmeans$alpha1[2], 3)} & \Sexpr{round(modelmeans$alpha2[2], 3)} & \Sexpr{round(modelmeans$alpha3[2], 3)} & \Sexpr{round(modelmeans$beta1[2], 3)} & \Sexpr{round(modelmeans$beta2[2], 3)} & \Sexpr{round(modelmeans$beta3[2], 3)} & -- & -- & --\\
-M4 & \Sexpr{round(modelmeans$alpha1[3], 3)} & \Sexpr{round(modelmeans$alpha2[3], 3)} & \Sexpr{round(modelmeans$alpha3[3], 3)} & \Sexpr{round(modelmeans$beta1[3], 3)} & \Sexpr{round(modelmeans$beta2[3], 3)} & -- & \Sexpr{round(modelmeans$beta4[3], 3)} & -- & --\\
-M5 & \Sexpr{round(modelmeans$alpha1[5], 3)} & \Sexpr{round(modelmeans$alpha2[5], 3)} & \Sexpr{round(modelmeans$alpha3[5], 3)} & \Sexpr{round(modelmeans$beta1[5], 3)}& \Sexpr{round(modelmeans$beta2[5], 3)} & -- & -- & \Sexpr{round(modelmeans$beta5[5], 3)} & -- \\
-M6 & \Sexpr{round(modelmeans$alpha1[4], 3)} & \Sexpr{round(modelmeans$alpha2[4], 3)} & \Sexpr{round(modelmeans$alpha3[4], 3)} & \Sexpr{round(modelmeans$beta1[4], 3)} & \Sexpr{round(modelmeans$beta2[4], 3)} & -- & -- & -- & \Sexpr{round(modelmeans$beta6[4], 3)}\\
-M7 & \Sexpr{round(M6[1], 3)} & \Sexpr{round(M6[2], 3)} & \Sexpr{round(M6[3], 3)} & \Sexpr{round(M6[4], 3)} & \Sexpr{round(M6[5], 3)} & -- & \Sexpr{round(M6[6], 3)} & \Sexpr{round(M6[8], 3)} & \Sexpr{round(M6[7], 3)} \\
-\bottomrule
-\end{tabular}
-\caption{\label{tab:fittedvalues} The mean estimates from repeated fits of our models of interest. When simulating from these models, these are the estimates that we will use unless otherwise stated.}
-\end{table}
-
 We fit all six of our models in \texttt{RSiena} using Markov Chain Monte Carlo (MCMC) methods to approximate the method of moments estimates of the parameters.

### 2019-08-08 ASApaper/supplement.Rnw
move effects table from paper to supplement. something is funky with fonts in pdfs on plots

@@
 Thus, the actor is more likely to make changes that increase the value of their objective function, and no change is most likely when any change decreases the value of the objective function.
+
+\subsection{Model Effects}
+
+The similarity measure is computed as: 
+\begin{equation}\label{eq:similar}
+sim^b_{ij} = \frac{\max_{hk}|b_h - b_k| - |b_i - b_j|}{\max_{hk}|b_h - b_k|} 
+\end{equation}
+where $\max_{hk}|b_h - b_k|$ is the range of number of bills authored by senators, and $b_i$, $b_j$ are the number of bills authored by senators $i,j$ repectively in the senate period.
+
+The fitted values from repeated converged simulations are
+
+<<modelestimates>>=
+load("data/allModelMeans.RDS")
+modelMeanEsts %>% unnest(ests) %>% mutate(param = "rate") -> modelmeans
+modelmeans$param[c(1,6,12, 18,24, 30)] <- "alpha1"
+modelmeans$param[c(1,6,12, 18,24, 30)+1] <- "alpha2"
+modelmeans$param[c(1,6,12, 18,24, 30)+2] <- "alpha3"
+modelmeans$param[c(1,6,12, 18,24, 30)+3] <- "beta1"
+modelmeans$param[c(1,6,12, 18,24, 30)+4] <- "beta2"
+modelmeans$param[modelmeans$param == "rate"] <- c("beta3", 'beta4', 'samep', 'beta6', 'beta5')
+modelmeans %>% filter(model != "samep") %>% spread(param, ests) -> modelmeans
+M6 <- c(2.4405048,2.4594403,2.2098176,-4.9232775,4.8916183,2.3743720,0.2047038,6.9661589)
+@
+
+\begin{table}
+\centering
+\begin{tabular}{|lccccccccc|}
+\toprule
+Model & $\hat{\alpha}_1$ & $\hat{\alpha}_2$ & $\hat{\alpha}_3$ & $\hat{\beta}_1$ & $\hat{\beta}_2$ & $\hat{\beta}_3$ & $\hat{\beta}_4$ & $\hat{\beta}_5$ & $\hat{\beta}_6$ \\
+\midrule
+M1 & \Sexpr{round(modelmeans$alpha1[1], 3)} & \Sexpr{round(modelmeans$alpha2[1], 3)} & \Sexpr{round(modelmeans$alpha3[1], 3)} & \Sexpr{round(modelmeans$beta1[1], 3)} & \Sexpr{round(modelmeans$beta2[1], 3)} & -- & -- & -- & --\\
+M3 & \Sexpr{round(modelmeans$alpha1[2], 3)} & \Sexpr{round(modelmeans$alpha2[2], 3)} & \Sexpr{round(modelmeans$alpha3[2], 3)} & \Sexpr{round(modelmeans$beta1[2], 3)} & \Sexpr{round(modelmeans$beta2[2], 3)} & \Sexpr{round(modelmeans$beta3[2], 3)} & -- & -- & --\\
+M4 & \Sexpr{round(modelmeans$alpha1[3], 3)} & \Sexpr{round(modelmeans$alpha2[3], 3)} & \Sexpr{round(modelmeans$alpha3[3], 3)} & \Sexpr{round(modelmeans$beta1[3], 3)} & \Sexpr{round(modelmeans$beta2[3], 3)} & -- & \Sexpr{round(modelmeans$beta4[3], 3)} & -- & --\\
+M5 & \Sexpr{round(modelmeans$alpha1[5], 3)} & \Sexpr{round(modelmeans$alpha2[5], 3)} & \Sexpr{round(modelmeans$alpha3[5], 3)} & \Sexpr{round(modelmeans$beta1[5], 3)}& \Sexpr{round(modelmeans$beta2[5], 3)} & -- & -- & \Sexpr{round(modelmeans$beta5[5], 3)} & -- \\
+M6 & \Sexpr{round(modelmeans$alpha1[4], 3)} & \Sexpr{round(modelmeans$alpha2[4], 3)} & \Sexpr{round(modelmeans$alpha3[4], 3)} & \Sexpr{round(modelmeans$beta1[4], 3)} & \Sexpr{round(modelmeans$beta2[4], 3)} & -- & -- & -- & \Sexpr{round(modelmeans$beta6[4], 3)}\\
+M7 & \Sexpr{round(M6[1], 3)} & \Sexpr{round(M6[2], 3)} & \Sexpr{round(M6[3], 3)} & \Sexpr{round(M6[4], 3)} & \Sexpr{round(M6[5], 3)} & -- & \Sexpr{round(M6[6], 3)} & \Sexpr{round(M6[8], 3)} & \Sexpr{round(M6[7], 3)} \\
+\bottomrule
+\end{tabular}
+\caption{\label{tab:fittedvalues} The mean estimates from repeated fits of our models of interest. When simulating from these models, these are the estimates that we will use unless otherwise stated.}
+\end{table}
 

### 2019-08-08 ASApaper/paper-child.Rnw
paper down to 24 pages. now do the hard part of making the paper and the supplement coherent.

@@
 
-<<sigtesting, results='asis'>>=
+<<sigtesting, results='asis', eval = FALSE>>=
 turk22_sig <- turk22 %>% 
@@
 
-<<lineup3132, fig.height=4, fig.cap="Lineup 3132, which led to rejection of the null hypothesis that $\\beta_3$ = 0.
-The network simulated from model M3 is found in panel $\\sqrt{16} - 1$, and the remaining panels show networks simulated from model M1.">>=
-dat <- read_csv("data/jttp_pos_hard_2.csv")
-ggplot(data = dat) + 
-        geom_net(aes(from_id = from, to_id = to), 
-                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
-                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
-                 color = 'black', arrowgap = .015, arrowsize = .3, size =1) + 
-        ThemeNet + 
-        theme(panel.background = element_rect(color = 'black')) +
-        facet_wrap(~ord)
-@
-
-<<lineup3133, fig.height=4, fig.cap="Lineup 3133, which resulted in failure to reject the null hypothesis that $\\beta_3$ = 0.
-The network simulated from model M3 is found in panel $\\sqrt{25} - 4$, and the remaining panels show networks simulated from model M1.
-16 of 27 viewers picked plot two as most different.">>=
-dat <- read_csv("data/jttp_pos_hard_3.csv")
-ggplot(data = dat) + 
-        geom_net(aes(from_id = from, to_id = to), 
-                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
-                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
-                 color = 'black', arrowgap = .015, arrowsize = .3, size =1) + 
-        ThemeNet + 
-        theme(panel.background = element_rect(color = 'black')) +
-        facet_wrap(~ord)
-@
-
 The lineup for significance testing of $\beta_3$ which resulted in a very small $p$-value and rejection of the null hypothesis is shown in Figure~\ref{fig:lineup3132}.
@@
 
-<<goftab, cache = TRUE, results='asis'>>=
+<<goftab, cache = TRUE, results='asis', eval = FALSE>>=
 turk22_gof <- read_csv("data/turk22-gof.csv")
@@
 
-<<failgof, fig.height=4, fig.cap="The goodness-of-fit lineup which resulted in failure to reject the null hypothesis.
-The null model for this lineup is M4.
-Only 7 of 20 viewers of this lineup selected the data plot as the most different from the others.
-The most commonly chosen panel was number four, which has a relatively simple structure compared to the other panels.">>=
-load("data/se112adjmat.RDS")
-wave2 <- data.frame(se112adj)
-senators <- data.frame(id = colnames(wave2), number = 1:155)
-wave2$from <- colnames(wave2)
-wave2 <- wave2 %>% gather(to, val, Alan.Stuart.Franken:William.Cowan) %>% filter(val > 0)
-wave2 <- merge(wave2, senators, by.x = "from", by.y = "id", all = T)
-wave2$from <- senators$number[match(wave2$from, senators$id)]
-wave2$to <- senators$number[match(wave2$to, senators$id)]
-wave2$from <- paste0("V", wave2$from)
-wave2$to <- ifelse(is.na(wave2$to), NA, paste0("V", wave2$to))
-wave2 <- wave2 %>% mutate(sim = 1001, model = "data", wave = 1) %>% select(-c(val, number))
-dat <- read_csv(paste0("data/jtts_gof_9_2.csv"))
-datplot <- unique(dat$ord[dat$sim == 1001])
-dat <- dat %>% filter(sim != 1001) %>% 
-      bind_rows(wave2) %>% 
-      mutate(ord = ifelse(is.na(ord), datplot, ord))
-ggplot(data = dat) + 
-        geom_net(aes(from_id = from, to_id = to), 
-                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
-                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
-                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
-        ThemeNet + 
-        theme(panel.background = element_rect(color = 'black')) +
-        facet_wrap(~ord)
-@
-
 The smallest $p$-value for one of the goodness-of-fit lineups was for the third replicate of the null model M4.
@@
 
-<<mostsiggof, fig.height = 4, fig.cap="The lineup resulting in the smallest $p$-value rejecting the null hypothesis.
-Surprisingly, this another repetition for M5 as the null model.
-The data are in panel 6.">>=
-dat <- read_csv(paste0("data/jtts_gof_9_3.csv"))
-datplot <- unique(dat$ord[dat$sim == 1001])
-dat <- dat %>% filter(sim != 1001) %>% 
-      bind_rows(wave2) %>% 
-      mutate(ord = ifelse(is.na(ord), datplot, ord))
-ggplot(data = dat) + 
-        geom_net(aes(from_id = from, to_id = to), 
-                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
-                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
-                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
-        ThemeNet + 
-        theme(panel.background = element_rect(color = 'black')) +
-        facet_wrap(~ord)
-@
-
 For the other models for which we tested goodness-of-fit, however, we \textit{do} have significant evidence from all three replicates to reject the null hypothesis that the null model generated the data.
@@
 %\hh{XXX we need to go throught the logic of the argument below.}\st{ XXX I reformulated the discussion per our convo today.  }
-
-<<gofm71, fig.height=4, fig.cap="One repitition of a goodness-of-fit lineup testing model M7.
-The senate data are shown in panel two, and it is evident that none of the other five panels, which show data simulated from model M7, come close to creating the large connected component that is central to the structure of the senate data.">>=
-dat <- read_csv(paste0("data/bigmod_gof_9_1.csv"))
-datplot <- unique(dat$ord[dat$sim == 1001])
-dat <- dat %>% filter(sim != 1001) %>% 
-      bind_rows(wave2) %>% 
-      mutate(ord = ifelse(is.na(ord), datplot, ord))
-ggplot(data = dat) + 
-        geom_net(aes(from_id = from, to_id = to), 
-                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
-                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
-                 color = 'black', arrowgap = .015, arrowsize = .3, size =1) + 
-        ThemeNet + 
-        theme(panel.background = element_rect(color = 'black')) +
-        facet_wrap(~ord)
-@
 
@@
 
-<<recipzoom, fig.height=3, fig.cap='The top middle panel of Figure~\\ref{fig:predictglmm} expanded to show greater detail.
-The square root of the parameter value is shown on the $x$-axis.
-For this parameter, as its value approaches zero, the probability of identifying the alternate data model decreases, then increases, which is noticeably different from the pattern exhibited by the others.
-Again, a horizontal line is drawn at 1/6, the chance of selecting the data plot at random.'>>=
-newdatabeta2 <- newdata %>% filter(test_param == "recip", category == "inside") %>% mutate(param2 = "beta[2]", size2 = sqrt(size)) %>% as.tibble()
-plotdatabeta2 <- plotdata %>% filter(test_param=="recip")%>% mutate(param2 = "beta[2]", size2 = sqrt(size)) %>% as.tibble()
-newdatabeta22 <-  newdata %>% filter(test_param == "recip", category == "outside")%>% mutate(param2 = "beta[2]", size2 = sqrt(size)) %>% as.tibble()
-newdatabeta23 <- newdata %>% filter(test_param == "recip", category == "outside2")%>% mutate(param2 = "beta[2]", size2 = sqrt(size)) %>% filter(!is.nan(size2)) %>% as.tibble()
-interceptlabel <- tibble(x = sqrt(4.893), y = .05, label = "hat(beta)[2]")
-ggplot() + 
-  geom_line(data = newdatabeta2, aes(x = size2, y = predictfinal, color = as.factor(type2)), size = 1.25) + 
-  geom_point(data = plotdatabeta2, aes(x = size2, y = datapick, color = as.factor(type2)), size = 2) + 
-  geom_line(data =newdatabeta22, aes(x = size2, y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
-  geom_line(data = newdatabeta23, aes(x = size2, y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
-  #geom_text(data = plotdata, aes(x = size, y = datapick, color = as.factor(type2), label = pic_id)) +
-  #geom_text(data = labdat, aes(x = x, y=y, label = label, color = as.factor(type2)),
-  #          show.legend = F) + 
-  geom_vline(xintercept = sqrt(4.893),  linetype = "dotted", color = "grey60") + 
-  geom_text(data = interceptlabel, aes(x =x, y = y, label = label), parse = T, inherit.aes = F, hjust = -.2) + 
-  geom_hline(yintercept = 1/6, linetype='dotted') +
-  scale_color_brewer(palette = "Dark2", name = "Lineup Type", labels = c("-", "+")) +  
-  facet_wrap(~param2, labeller = label_parsed) + 
-  labs(x = "Square root of parameter value", y = "Expected probability of detection\nin a lineup of size 6", title = ) + 
-  ThemeNoNet + 
-  theme(legend.position = 'bottom')
-@
-
-<<beta5zoom, fig.height=3, fig.cap="The bottom middle panel of Figure~\\ref{fig:predictglmm} expanded to show greater detail.
-The parameter value is shown on the $x$-axis.
-This parameter most closely follows our hypothesis shown in Figure~\\ref{fig:hypothesis}.
-However, the result is not symmetric.
-According to the model, people will detect the effect at lower values and with greater frequency as the value increases when it is positive instead of negative.">>=
-newdatabeta5 <- newdata %>% filter(test_param == "simttb", category == "inside") %>% mutate(param = "beta[5]")
-plotdatabeta5 <- plotdata %>% filter(test_param=="simttb")%>% mutate(param = "beta[5]")
-newdatabeta52 <- newdata %>% filter(test_param == "simttb", category == "outside") %>% mutate(param = "beta[5]")
-newdatabeta53 <- newdata %>% filter(test_param == "simttb", category == "outside2") %>% mutate(param = "beta[5]")
-interceptdata <- tibble(x = 10.090829, y= .05, label = "hat(beta)[5]")
-ggplot() + 
-  geom_line(data = newdatabeta5, aes(x = size, y = predictfinal, color = as.factor(type2)), size = 1.25) + 
-  geom_point(data = plotdatabeta5, aes(x = size, y = datapick, color = as.factor(type2)), size = 2) + 
-  geom_line(data = newdatabeta52, aes(x = size, y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
-  geom_line(data = newdatabeta53, aes(x = size, y = predictfinal, group = as.factor(type2)), color = "gray60",  alpha = .7) +
-  #geom_text(data = plotdata, aes(x = size, y = datapick, color = as.factor(type2), label = pic_id)) +
-  #geom_text(data = labdat, aes(x = x, y=y, label = label, color = as.factor(type2)),
-  #          show.legend = F) + 
-  geom_vline(xintercept = 10.090829, linetype = "dotted", color = "grey60") + 
-  geom_text(data = interceptdata, aes(x=x,y=y,label=label), parse = T, hjust = -.2) + 
-  geom_hline(yintercept = 1/6, linetype='dotted') +
-  scale_color_brewer(palette = "Dark2", name = "Lineup Type", labels = c("-", "+")) +
-  facet_grid(~param, labeller = label_parsed) + 
-  labs(x = "Parameter value", y = "Expected probability of detection\nin a lineup of size 6") + 
-  ThemeNoNet + 
-  theme(legend.position = 'bottom', strip.text = element_text(size = 10), text = element_text(size = rel(3)), legend.text = element_text(size = rel(3)), legend.key.size = unit(.25, "inches"))
-@
-
-<<beta4zoom, fig.height=3, fig.cap='The bottom right panel of Figure~\\ref{fig:predictglmm} expanded to show greater detail.
-The parameter value is shown on the $x$-axis.
-The ``reverse" lineup has a much flatter slope than the ``regular" lineup, which means the participants had a harder time detecting a more simple M1 structure among many more complex M4 structures.
-Reversing the lineup scenario was not symmetric as we hypothesized.'>>=
-newdatabeta6 <- newdata %>% filter(test_param == "jtts", category == "inside") %>% mutate(param = "beta[6]")
-plotdatabeta6 <- plotdata %>% filter(test_param=="jtts") %>% mutate(param = "beta[6]")
-newdatabeta62 <- newdata %>% filter(test_param == "jtts", category == "outside")%>% mutate(param = "beta[6]")
-newdatabeta63 <- newdata %>% filter(test_param == "jtts", category == "outside2") %>% mutate(param = "beta[6]")
-labellerdata <- tibble(x = 3.340302, y = .8, label = "hat(beta)[6]")
-ggplot() + 
-  geom_vline(xintercept = 3.340302, linetype = "dotted", color = "grey60") + 
-  geom_text(data = labellerdata, aes(x = x, y = y, label = label), parse =T, hjust = 1) + 
-  geom_line(data = newdatabeta6, aes(x = size, y = predictfinal, color = as.factor(type2)), size = 1.25) + 
-  geom_point(data = plotdatabeta6, aes(x = size, y = datapick, color = as.factor(type2)), size = 2) + 
-  geom_line(data = newdatabeta62, aes(x = size, y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
-  geom_line(data = newdatabeta63, aes(x = size, y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
-  #geom_text(data = plotdata, aes(x = size, y = datapick, color = as.factor(type2), label = pic_id)) +
-  #geom_text(data = labdat, aes(x = x, y=y, label = label, color = as.factor(type2)),
-  #          show.legend = F) + 
-  geom_hline(yintercept = 1/6, linetype='dotted') +
-  scale_color_brewer(palette = "Dark2", name = "Lineup Condition", labels = c("reverse", "regular")) +
-  facet_grid(~param, labeller = label_parsed) + 
-  labs(x = "Parameter value", y = "Expected probability of detection\nin a lineup of size 6") + 
-  ThemeNoNet + 
-  theme(legend.position = 'bottom', strip.text = element_text(size = 10), text = element_text(size = rel(3)), legend.text = element_text(size = rel(3)), legend.key.size = unit(.25, "inches"))
-@
-
-We expand portions of Figure~\ref{fig:predictglmm} in Figures~\ref{fig:recipzoom}-\ref{fig:beta4zoom}.
-These figures show the same prediction regions as in Figure~\ref{fig:predictglmm}, plus some additional predictions outside of the data range shown in light gray.
-Again, the points represent the results from the experiment.
-In all three of these figures, the lack of symmetry is apparent.
-In the reverse lineup scenario shown in Figure~\ref{fig:beta4zoom}, the probability of prediction is consistently far less than the probability of prediction in the regular lineup scenario.
-This demonstrates that the visual signal of one plot from M4 among five plots from M1 is much stronger than that of one plot from M1 among five plots from M4.
-We posit that the latter is a more difficult task because it involves noticing a \textit{lack of structure} as opposed to the presence of \textit{more} structure.
-We can see a similar effect in Figure~\ref{fig:beta5zoom}.
-At a value of $\beta_5 = 20$, the model predicts a probability of about 0.60 that a new viewer of a new lineup will identify the alternative data plot.
-At a value of $\beta_5 = -20$, however, the model predicts this same probability to be about 0.40.
-This again demonstrates that the presence of structure is detected  more frequently and at smaller values than the absence of structure.
-
-<<beta41, fig.height = 4, fig.cap='In our experiment, 52.8\\% of viewers of this plot selected the plot from the alternative model, M4.
-The ``reverse" of this lineup is given in Figure~\\ref{fig:beta4neg1}, where 41.4\\% of viewers selected the plot from the alternative model, M1.
-Here, the alternative plot is $\\sqrt{25} - 3$.'>>=
-dat <- read_csv("data/jtts_pos_easy_3.csv")
-# "answer" is 2
-ggplot(data = dat) + 
-        geom_net(aes(from_id = from, to_id = to), 
-                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
-                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
-                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
-        ThemeNet + 
-        theme(panel.background = element_rect(color = 'black')) +
-        facet_wrap(~ord)
-@
-
-<<beta4neg1,fig.height=4, fig.cap='In our experiment, 41.4\\% people viewing this lineup selected the plot from the alternative model, M1.
-Here, the alternative plot is $\\sqrt{25} - 1$.
-The other five plots were simulated from model M4.' >>=
-dat <- read_csv("data/jtts_neg_med_2.csv")
-# 'answer' is 4
-ggplot(data = dat) + 
-        geom_net(aes(from_id = from, to_id = to), 
-                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
-                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
-                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
-        ThemeNet + 
-        theme(panel.background = element_rect(color = 'black')) +
-        facet_wrap(~ord)
-@
-
 For $\beta_4$ and $\beta_6$, where one plot simulated from M1 was placed among five plots from the corresponding model, we see that the predictions for the reverse lineup type (-1), are less than the standard lineup type (1) for all values of the parameter that we have.

### 2019-08-08 ASApaper/supplement.Rnw
paper down to 24 pages. now do the hard part of making the paper and the supplement coherent.

@@
 @
+
+<<pkgs-sm>>=
+library(tidyverse)
+library(RSiena)
+library(geomnet)
+library(vinference)
+turk22 <- read_csv("data/turk22-sig.csv")
+newdata <- read_csv("data/newdata_pred_glmm.csv")
+newdata$param <- newdata$test_param
+newdata$param <- as.factor(newdata$test_param)
+levels(newdata$param) <-  c(1,3,4,2,6,5) 
+newdata$param <- as.integer(as.character(newdata$param))
+plotdata <- turk22 %>% mutate(type2 = ifelse(sign == 0, sign(initialEst), sign),
+                       type2 = ifelse(type == "one", type2, -1)) %>% 
+  group_by(pic_id, size, test_param, sign, type2,alt_model) %>% summarize(
+    datapick = mean(datapick)
+)
+plotdata$param <- plotdata$test_param
+plotdata$param <- as.factor(plotdata$param)
+levels(plotdata$param) <- c(1,3,4,2,6,5) 
+plotdata$param <- as.integer(as.character(plotdata$param))
+labdat <- data_frame(test_param = "jttp", param = 3, type2 = -1, x = -5, y = .9, label = "outlier")
+betahat <- tibble(param = 1:6, size = c(-4.903, 4.893, -3.45, 3.34, 10.091, 1.329))
+ThemeNoNet <- theme_bw() %+replace% 
+            theme(plot.title = element_text(size = 12,
+                                            face = 'plain', 
+                                            angle = 0, 
+                                            family="Times New Roman"), 
+                  axis.title.x = element_text(size = 12,
+                                            face = 'plain', 
+                                            angle = 0,
+                                            family="Times New Roman"),
+                  axis.title.y = element_text(size = 12,
+                                            face = 'plain', 
+                                            angle = 90,
+                                            family="Times New Roman"),
+                  axis.text.x = element_text(size = 12,
+                                            face = 'plain', 
+                                            angle = 0,
+                                            family="Times New Roman"), 
+                  # axis.text.x.top = element_text(size = 12,
+                  #                           face = 'plain', 
+                  #                           angle = 0,
+                  #                           family="Times New Roman"), 
+                  # axis.text.x.bottom = element_text(size = 12,
+                  #                           face = 'plain', 
+                  #                           angle = 0,
+                  #                           family="Times New Roman"), 
+                  axis.text.y = element_text(size = 12,
+                                            face = 'plain', 
+                                            angle = 0,
+                                            family="Times New Roman"),
+                  # axis.text.y.left = element_text(size = 12,
+                  #                           face = 'plain', 
+                  #                           angle = 0,
+                  #                           family="Times New Roman"),
+                  # axis.text.y.right = element_text(size = 12,
+                  #                           face = 'plain', 
+                  #                           angle = 0,
+                  #                           family="Times New Roman"),
+                  strip.text.x = element_text(size = 12,
+                                            face = 'plain', 
+                                            angle = 0, 
+                                            family="Times New Roman",
+                                            margin = margin(t = 3, r = 0, b = 3, l = 0, unit = "pt")),
+                  strip.text.y =element_text(size = 12,
+                                            face = 'plain', 
+                                            angle = 90,
+                                            family="Times New Roman",
+                                            margin = margin(t = 3, r = 3, b = 3, l = 3, unit = "pt")),
+                  strip.background = element_rect(colour = "black", fill = "white")
+                    )
+ThemeNet <- theme_net() %+replace% 
+            theme(plot.title = element_text(size = 12,
+                                            face = 'plain', angle = 0, family = "Times New Roman"), 
+                  strip.text.x = element_text(size = 12,
+                                            face = 'plain', angle = 0, family = "Times New Roman", 
+                                            margin = margin(t = 3, r = 0, b = 3, l = 0, unit = "pt")),
+                  strip.text.y =element_text(size = 12,
+                                            face = 'plain', angle = 0, family = "Times New Roman",margin = margin(t = 3, r = 0, b = 3, l = 0, unit = "pt")),
+                  panel.border = element_rect(fill = NA, color = 'black'),
+                  strip.background = element_rect(colour = "black", fill = "white"))
+@
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\section{Lineups}
+
+<<lineup3132, fig.height=4, fig.cap="Lineup 3132, which led to rejection of the null hypothesis that $\\beta_3$ = 0.
+The network simulated from model M3 is found in panel $\\sqrt{16} - 1$, and the remaining panels show networks simulated from model M1.">>=
+dat <- read_csv("data/jttp_pos_hard_2.csv")
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size =1) + 
+        ThemeNet + 
+        theme(panel.background = element_rect(color = 'black')) +
+        facet_wrap(~ord)
+@
+
+<<lineup3133, fig.height=4, fig.cap="Lineup 3133, which resulted in failure to reject the null hypothesis that $\\beta_3$ = 0.
+The network simulated from model M3 is found in panel $\\sqrt{25} - 4$, and the remaining panels show networks simulated from model M1.
+16 of 27 viewers picked plot two as most different.">>=
+dat <- read_csv("data/jttp_pos_hard_3.csv")
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size =1) + 
+        ThemeNet + 
+        theme(panel.background = element_rect(color = 'black')) +
+        facet_wrap(~ord)
+@
+
+<<failgof, fig.height=4, fig.cap="The goodness-of-fit lineup which resulted in failure to reject the null hypothesis.
+The null model for this lineup is M4.
+Only 7 of 20 viewers of this lineup selected the data plot as the most different from the others.
+The most commonly chosen panel was number four, which has a relatively simple structure compared to the other panels.">>=
+load("data/se112adjmat.RDS")
+wave2 <- data.frame(se112adj)
+senators <- data.frame(id = colnames(wave2), number = 1:155)
+wave2$from <- colnames(wave2)
+wave2 <- wave2 %>% gather(to, val, Alan.Stuart.Franken:William.Cowan) %>% filter(val > 0)
+wave2 <- merge(wave2, senators, by.x = "from", by.y = "id", all = T)
+wave2$from <- senators$number[match(wave2$from, senators$id)]
+wave2$to <- senators$number[match(wave2$to, senators$id)]
+wave2$from <- paste0("V", wave2$from)
+wave2$to <- ifelse(is.na(wave2$to), NA, paste0("V", wave2$to))
+wave2 <- wave2 %>% mutate(sim = 1001, model = "data", wave = 1) %>% select(-c(val, number))
+dat <- read_csv(paste0("data/jtts_gof_9_2.csv"))
+datplot <- unique(dat$ord[dat$sim == 1001])
+dat <- dat %>% filter(sim != 1001) %>% 
+      bind_rows(wave2) %>% 
+      mutate(ord = ifelse(is.na(ord), datplot, ord))
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
+        ThemeNet + 
+        theme(panel.background = element_rect(color = 'black')) +
+        facet_wrap(~ord)
+@
+
+<<mostsiggof, fig.height = 4, fig.cap="The lineup resulting in the smallest $p$-value rejecting the null hypothesis.
+Surprisingly, this another repetition for M5 as the null model.
+The data are in panel 6.">>=
+dat <- read_csv(paste0("data/jtts_gof_9_3.csv"))
+datplot <- unique(dat$ord[dat$sim == 1001])
+dat <- dat %>% filter(sim != 1001) %>% 
+      bind_rows(wave2) %>% 
+      mutate(ord = ifelse(is.na(ord), datplot, ord))
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
+        ThemeNet + 
+        theme(panel.background = element_rect(color = 'black')) +
+        facet_wrap(~ord)
+@
+
+<<gofm71, fig.height=4, fig.cap="One repitition of a goodness-of-fit lineup testing model M7.
+The senate data are shown in panel two, and it is evident that none of the other five panels, which show data simulated from model M7, come close to creating the large connected component that is central to the structure of the senate data.">>=
+dat <- read_csv(paste0("data/bigmod_gof_9_1.csv"))
+datplot <- unique(dat$ord[dat$sim == 1001])
+dat <- dat %>% filter(sim != 1001) %>% 
+      bind_rows(wave2) %>% 
+      mutate(ord = ifelse(is.na(ord), datplot, ord))
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size =1) + 
+        ThemeNet + 
+        theme(panel.background = element_rect(color = 'black')) +
+        facet_wrap(~ord)
+@
+
+<<beta41, fig.height = 4, fig.cap='In our experiment, 52.8\\% of viewers of this plot selected the plot from the alternative model, M4.
+The ``reverse" of this lineup is given in Figure~\\ref{fig:beta4neg1}, where 41.4\\% of viewers selected the plot from the alternative model, M1.
+Here, the alternative plot is $\\sqrt{25} - 3$.'>>=
+dat <- read_csv("data/jtts_pos_easy_3.csv")
+# "answer" is 2
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
+        ThemeNet + 
+        theme(panel.background = element_rect(color = 'black')) +
+        facet_wrap(~ord)
+@
+
+<<beta4neg1,fig.height=4, fig.cap='In our experiment, 41.4\\% people viewing this lineup selected the plot from the alternative model, M1.
+Here, the alternative plot is $\\sqrt{25} - 1$.
+The other five plots were simulated from model M4.' >>=
+dat <- read_csv("data/jtts_neg_med_2.csv")
+# 'answer' is 4
+ggplot(data = dat) + 
+        geom_net(aes(from_id = from, to_id = to), 
+                 arrow = arrow(type = 'open', length = unit(2, "points") ), 
+                 linewidth = .25, singletons = T, fiteach = T, directed = T, 
+                 color = 'black', arrowgap = .015, arrowsize = .3, size = .5) + 
+        ThemeNet + 
+        theme(panel.background = element_rect(color = 'black')) +
+        facet_wrap(~ord)
+@
+
 \section{Results}
@@
 
+\subsection{Significance Testing}
+
+<<sigtesting, results='asis'>>=
+turk22_sig <- turk22 %>% 
+  mutate(type2 = ifelse(sign == 0, sign(initialEst), sign),
+         type2 = ifelse(type == "one", type2, -1)) %>% 
+  filter(param_value == 1)
+summ_sig <- turk22_sig %>% group_by(pic_id, test_param) %>% 
+  summarize(npick = sum(datapick), n = n(), 
+            pv = map2_dbl(npick, n, pV, m=6, scenario=3))
+names(summ_sig) <- c("Lineup ID", "parameter", "\\# ID", "Total Views", "p-value")
+summ_sig$parameter <- as.factor(summ_sig$parameter)
+levels(summ_sig$parameter) <- c("$ \\beta_3 $", "$ \\beta_4$ ")
+xtable::print.xtable(xtable::xtable(summ_sig, caption = "Experiment results for the two parameters for which we performed significance tests.
+\\# ID indicates the number of participants who identified the alternative data plot.
+There were three lineups for each parameter, so there are three results for each plot.", label = "tab:sigtesting", digits = c(0,0,0,0,0,5)), include.rownames = FALSE, sanitize.text.function=function(x){x})
+@
+
+\subsection{GoF}
+
+<<goftab, cache = FALSE, results='asis'>>=
+turk22_gof <- read_csv("data/turk22-gof.csv")
+turk22_gof_stats <- turk22_gof %>% group_by(pic_id) %>% 
+  summarize(npickdata = sum(datapick), total = n(), 
+          pvinf = map2_dbl(npickdata, total, pV, m = 6, scenario = 3))
+turk22_gof_stats %>% separate(pic_id, into = c("discard", "model", "rep"), sep = c(2,3)) -> print_gof_stats
+print_gof_stats %>% select(-discard) %>% mutate(model = paste0("M", (as.integer(model))))-> print_gof_stats
+names(print_gof_stats) <- c("Model", "Replicate", "Data Picks", "Total Viewers", "p-value")
+print_gof_stats$Model[c(2:3,5:6,8:9,11:12)] <- ""
+#print_gof_stats$Model[2] <- "jtt party"
+#print_gof_stats$Model[5] <- "jtt sex"
+#print_gof_stats$Model[8] <- "stt party"
+print_gof_stats$`p-value` <- ifelse(print_gof_stats$`p-value` < .0001, "< 0.0001", sprintf("%.4f", print_gof_stats$`p-value`))
+print(xtable::xtable(print_gof_stats, label = "tab:gofstats", caption = "An overview of the results from the 12 goodness-of-fit lineup tests.", digits = 4,
+                      align="rlrrrr"), caption.placement = "bottom", include.rownames = FALSE, hline.after = c(-1,0,0,3,6,9,nrow(print_gof_stats)))
+@
+
 \subsection{Visual Power}\label{sec:sm-res-vp}
@@
 <<glmmres>>=
+newdata <- read_csv("data/newdata_pred_glmm.csv")
+newdata$param <- newdata$test_param
+newdata$param <- as.factor(newdata$test_param)
+levels(newdata$param) <-  c(1,3,4,2,6,5) 
+newdata$param <- as.integer(as.character(newdata$param))
+plotdata <- turk22 %>% mutate(type2 = ifelse(sign == 0, sign(initialEst), sign),
+                       type2 = ifelse(type == "one", type2, -1)) %>% 
+  group_by(pic_id, size, test_param, sign, type2,alt_model) %>% summarize(
+    datapick = mean(datapick)
+)
+plotdata$param <- plotdata$test_param
+plotdata$param <- as.factor(plotdata$param)
+levels(plotdata$param) <- c(1,3,4,2,6,5) 
+plotdata$param <- as.integer(as.character(plotdata$param))
+labdat <- data_frame(test_param = "jttp", param = 3, type2 = -1, x = -5, y = .9, label = "outlier")
+betahat <- tibble(param = 1:6, size = c(-4.903, 4.893, -3.45, 3.34, 10.091, 1.329))
 load("data/finalglmm31-2.RDA")
@@
 
+<<morepredicts>>=
+alphas <- (coef(mod)[[1]][,-1] %>% unique())[1:12] 
+betas <- (coef(mod)[[1]][,-1] %>% unique())[13:24] 
+names(betas) <- str_replace(str_replace(names(betas), "lps_param", ""), ":centersize", "")
+names(betas) <- str_replace(names(betas), ".-1", "neg")
+names(betas) <- str_replace(names(betas), ".1", "pos")
+names(alphas) <- str_replace(names(alphas), "lps_param", "")
+names(alphas) <- str_replace(names(alphas), ".-1", "neg")
+names(alphas) <- str_replace(names(alphas), ".1", "pos")
+source("code/newpredictions.R")
+newdata <- newdata3
+newdata$param <- newdata$test_param
+newdata$param <- as.factor(newdata$test_param)
+#levels(newdata$param) <-  c(1,3,4,2,6,5) 
+#newdata$param <- as.integer(as.character(newdata$param))
+plotdata <- turk22 %>% mutate(type2 = ifelse(sign == 0, sign(initialEst), sign),
+                       type2 = ifelse(type == "one", type2, -1)) %>% 
+  group_by(pic_id, size, test_param, sign, type2,alt_model) %>% summarize(
+    datapick = mean(datapick)
+)
+plotdata$param <- plotdata$test_param
+plotdata$param <- as.factor(plotdata$param)
+#levels(plotdata$param) <- c(1,3,4,2,6,5) 
+#plotdata$param <- as.integer(as.character(plotdata$param))
+labdat <- data_frame(test_param = "jttp", param = 3, type2 = -1, x = -5, y = .9, label = "jttp_neg_hard_2")
+@
+
+<<recipzoom, fig.height=3, fig.cap='The top middle panel of Figure~\\ref{fig:predictglmm} expanded to show greater detail.
+The square root of the parameter value is shown on the $x$-axis.
+For this parameter, as its value approaches zero, the probability of identifying the alternate data model decreases, then increases, which is noticeably different from the pattern exhibited by the others.
+Again, a horizontal line is drawn at 1/6, the chance of selecting the data plot at random.'>>=
+library(dplyr)
+newdatabeta2 <- newdata %>% filter(test_param == "recip", category == "inside") %>% mutate(param2 = "beta[2]", size2 = sqrt(size)) %>% as.tibble()
+plotdatabeta2 <- plotdata %>% filter(test_param=="recip")%>% mutate(param2 = "beta[2]", size2 = sqrt(size)) %>% as.tibble()
+newdatabeta22 <-  newdata %>% filter(test_param == "recip", category == "outside")%>% mutate(param2 = "beta[2]", size2 = sqrt(size)) %>% as.tibble()
+newdatabeta23 <- newdata %>% filter(test_param == "recip", category == "outside2")%>% mutate(param2 = "beta[2]", size2 = sqrt(size)) %>% filter(!is.nan(size2)) %>% as.tibble()
+interceptlabel <- tibble(x = sqrt(4.893), y = .05, label = "hat(beta)[2]")
+ggplot() + 
+  geom_line(data = newdatabeta2, aes(x = size2, y = predictfinal, color = as.factor(type2)), size = 1.25) + 
+  geom_point(data = plotdatabeta2, aes(x = size2, y = datapick, color = as.factor(type2)), size = 2) + 
+  geom_line(data =newdatabeta22, aes(x = size2, y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
+  geom_line(data = newdatabeta23, aes(x = size2, y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
+  #geom_text(data = plotdata, aes(x = size, y = datapick, color = as.factor(type2), label = pic_id)) +
+  #geom_text(data = labdat, aes(x = x, y=y, label = label, color = as.factor(type2)),
+  #          show.legend = F) + 
+  geom_vline(xintercept = sqrt(4.893),  linetype = "dotted", color = "grey60") + 
+  geom_text(data = interceptlabel, aes(x =x, y = y, label = label), parse = T, inherit.aes = F, hjust = -.2) + 
+  geom_hline(yintercept = 1/6, linetype='dotted') +
+  scale_color_brewer(palette = "Dark2", name = "Lineup Type", labels = c("-", "+")) +  
+  facet_wrap(~param2, labeller = label_parsed) + 
+  labs(x = "Square root of parameter value", y = "Expected probability of detection\nin a lineup of size 6", title = ) + 
+  ThemeNoNet + 
+  theme(legend.position = 'bottom')
+@
+
+<<beta5zoom, fig.height=3, fig.cap="The bottom middle panel of Figure~\\ref{fig:predictglmm} expanded to show greater detail.
+The parameter value is shown on the $x$-axis.
+This parameter most closely follows our hypothesis shown in Figure~\\ref{fig:hypothesis}.
+However, the result is not symmetric.
+According to the model, people will detect the effect at lower values and with greater frequency as the value increases when it is positive instead of negative.">>=
+newdatabeta5 <- newdata %>% filter(test_param == "simttb", category == "inside") %>% mutate(param = "beta[5]")
+plotdatabeta5 <- plotdata %>% filter(test_param=="simttb")%>% mutate(param = "beta[5]")
+newdatabeta52 <- newdata %>% filter(test_param == "simttb", category == "outside") %>% mutate(param = "beta[5]")
+newdatabeta53 <- newdata %>% filter(test_param == "simttb", category == "outside2") %>% mutate(param = "beta[5]")
+interceptdata <- tibble(x = 10.090829, y= .05, label = "hat(beta)[5]")
+ggplot() + 
+  geom_line(data = newdatabeta5, aes(x = size, y = predictfinal, color = as.factor(type2)), size = 1.25) + 
+  geom_point(data = plotdatabeta5, aes(x = size, y = datapick, color = as.factor(type2)), size = 2) + 
+  geom_line(data = newdatabeta52, aes(x = size, y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
+  geom_line(data = newdatabeta53, aes(x = size, y = predictfinal, group = as.factor(type2)), color = "gray60",  alpha = .7) +
+  #geom_text(data = plotdata, aes(x = size, y = datapick, color = as.factor(type2), label = pic_id)) +
+  #geom_text(data = labdat, aes(x = x, y=y, label = label, color = as.factor(type2)),
+  #          show.legend = F) + 
+  geom_vline(xintercept = 10.090829, linetype = "dotted", color = "grey60") + 
+  geom_text(data = interceptdata, aes(x=x,y=y,label=label), parse = T, hjust = -.2) + 
+  geom_hline(yintercept = 1/6, linetype='dotted') +
+  scale_color_brewer(palette = "Dark2", name = "Lineup Type", labels = c("-", "+")) +
+  facet_grid(~param, labeller = label_parsed) + 
+  labs(x = "Parameter value", y = "Expected probability of detection\nin a lineup of size 6") + 
+  ThemeNoNet + 
+  theme(legend.position = 'bottom', strip.text = element_text(size = 10), text = element_text(size = rel(3)), legend.text = element_text(size = rel(3)), legend.key.size = unit(.25, "inches"))
+@
+
+<<beta4zoom, fig.height=3, fig.cap='The bottom right panel of Figure~\\ref{fig:predictglmm} expanded to show greater detail.
+The parameter value is shown on the $x$-axis.
+The ``reverse" lineup has a much flatter slope than the ``regular" lineup, which means the participants had a harder time detecting a more simple M1 structure among many more complex M4 structures.
+Reversing the lineup scenario was not symmetric as we hypothesized.'>>=
+newdatabeta6 <- newdata %>% filter(test_param == "jtts", category == "inside") %>% mutate(param = "beta[6]")
+plotdatabeta6 <- plotdata %>% filter(test_param=="jtts") %>% mutate(param = "beta[6]")
+newdatabeta62 <- newdata %>% filter(test_param == "jtts", category == "outside")%>% mutate(param = "beta[6]")
+newdatabeta63 <- newdata %>% filter(test_param == "jtts", category == "outside2") %>% mutate(param = "beta[6]")
+labellerdata <- tibble(x = 3.340302, y = .8, label = "hat(beta)[6]")
+ggplot() + 
+  geom_vline(xintercept = 3.340302, linetype = "dotted", color = "grey60") + 
+  geom_text(data = labellerdata, aes(x = x, y = y, label = label), parse =T, hjust = 1) + 
+  geom_line(data = newdatabeta6, aes(x = size, y = predictfinal, color = as.factor(type2)), size = 1.25) + 
+  geom_point(data = plotdatabeta6, aes(x = size, y = datapick, color = as.factor(type2)), size = 2) + 
+  geom_line(data = newdatabeta62, aes(x = size, y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
+  geom_line(data = newdatabeta63, aes(x = size, y = predictfinal, group = as.factor(type2)), color = "gray60", alpha = .7) +
+  #geom_text(data = plotdata, aes(x = size, y = datapick, color = as.factor(type2), label = pic_id)) +
+  #geom_text(data = labdat, aes(x = x, y=y, label = label, color = as.factor(type2)),
+  #          show.legend = F) + 
+  geom_hline(yintercept = 1/6, linetype='dotted') +
+  scale_color_brewer(palette = "Dark2", name = "Lineup Condition", labels = c("reverse", "regular")) +
+  facet_grid(~param, labeller = label_parsed) + 
+  labs(x = "Parameter value", y = "Expected probability of detection\nin a lineup of size 6") + 
+  ThemeNoNet + 
+  theme(legend.position = 'bottom', strip.text = element_text(size = 10), text = element_text(size = rel(3)), legend.text = element_text(size = rel(3)), legend.key.size = unit(.25, "inches"))
+@
+
+We expand portions of Figure~\ref{fig:predictglmm} in Figures~\ref{fig:recipzoom}-\ref{fig:beta4zoom}.
+These figures show the same prediction regions as in Figure~\ref{fig:predictglmm}, plus some additional predictions outside of the data range shown in light gray.
+Again, the points represent the results from the experiment.
+In all three of these figures, the lack of symmetry is apparent.
+In the reverse lineup scenario shown in Figure~\ref{fig:beta4zoom}, the probability of prediction is consistently far less than the probability of prediction in the regular lineup scenario.
+This demonstrates that the visual signal of one plot from M4 among five plots from M1 is much stronger than that of one plot from M1 among five plots from M4.
+We posit that the latter is a more difficult task because it involves noticing a \textit{lack of structure} as opposed to the presence of \textit{more} structure.
+We can see a similar effect in Figure~\ref{fig:beta5zoom}.
+At a value of $\beta_5 = 20$, the model predicts a probability of about 0.60 that a new viewer of a new lineup will identify the alternative data plot.
+At a value of $\beta_5 = -20$, however, the model predicts this same probability to be about 0.40.
+This again demonstrates that the presence of structure is detected  more frequently and at smaller values than the absence of structure.
+
 %%%%%%%%%%%%%%%%%%%%%% References %%%%%%%%%%%%%%%%%%%%%%%%
